### 2021-02-23 sections/1-introduction.tex
setup

\section{Introduction}
\label{sec:introduction}

Write intro \dots

Cite CHESS~\cite{ishaq2019clustered}

### 2021-02-23 sections/2-methods.tex
setup

\section{Methods}
\label{sec:methods}

Write about methods \dots

### 2021-02-23 sections/3-results.tex
setup

\section{Results}
\label{sec:results}

Write about results \dots

### 2021-02-23 sections/4-discussion.tex
setup

\section{Discussion}
\label{sec:discussion}

Discuss results and future work \dots

Include figures~\ref{fig:discussion:umap-annthyroid-euclidean}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
    \label{fig:discussion:umap-annthyroid-euclidean}
\end{figure}

### 2021-02-23 sections/2-methods.tex
wrote up proof for how cluster radii decrease

@@
 Write about methods \dots
+
+\subsection{Proofs}
+\label{subserc:methods:proofs}
+
+\subsubsection{Guaranteed decreases in Radii}
+\label{subsubsec:methods:proofs:radii-decrease}
+
+We take a dataset to be any set $X$ with $n$ instances in $\mathcal{D}$ embedding dimensions where the data follow a low-dimensional manifold with fractal dimension $d$ where $d \ll \mathcal{D}$.
+
+We take a distance function to be $f: X \times X \mapsto \mathbb{R}$ with the following properties:
+
+\begin{itemize}
+    \item $f(x, y) = 0 \ iff \ x = y$, identity
+    \item $f(x, y) = f(y, x) \ \forall x, y \in X$, symmetry
+    \item $f(x - z, y - z) = f(x, y) \ \forall x, y, z \in X$, invariance under translation 
+\end{itemize}
+
+We show that the radii of clusters are guaranteed to decrease after every $d$ partitions (at most).
+
+Assume that the data follow a $d$-dimensional distribution embedded in $\mathcal{D}$-dimensional space.
+We can describe this distribution choosing some set of $d$ mutually orthagonal axes.
+Let $2r$ be the maximum distance between any pair of points along any axis from any choice of mutually orthagonal axes for our $d$-dimensional distribution.
+The cluster containing all these points will have a radius of $r$.
+In the worst case, the distribution is a $d$-sphere and $2r$ is the maximum distance along every axis from every choice of axes.
+
+The partition algorithm will choose the maximally distant pair of points and, thus, the axis defined by those points as the axis for partitioning the points into two child clusters.
+For points in either child cluster, the maximum distance between any pair of points along that axis will have been reduced to $r$ or less.
+If $d > 1$ then the child clusters will have a radius $r_{child}$ such that $r_{child} \leq r \sqrt{2}$.
+By partitioning the cluster in this way, we have taken one axis and reduced the maximum distance along that axis from $2r$ to $r$.
+However, as per the distribution of data as discussed earlier, there may be other axes along which the maximum pairwise distance is still $2r$.
+We, thus, recursively partition each child cluster.
+
+With each recursive application of Partition, we consume one axis along which the maximum distance was $2r$ and reduce the maximum distance along that axis to $r$ in the child clusters.
+After $d$ recursive applications of partition, we will have exhausted the $d$ axes with the maximum distance of $2r$.
+After these $d$ partitions, the points in each child cluster will have a maximum pairwise distance of $r$ along any axis.
+Thus the radii of those child clusters will be at most $\frac{r}{2}$.

### 2021-03-09 sections/1-introduction.tex
wrote outline

@@
 
-Cite CHESS~\cite{ishaq2019clustered}
+Growth in size and types of datasets:
+\begin{itemize}
+    \item SDSS APOGEE2 data. Near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions. $R=22,500 = \frac{\lambda}{\Delta \lambda}$ for $\lambda \in [1.51, 1.70) \mu m$
+    \item SDSS MaNGA hexagon shaped images. 10,000 galaxies with 2,700 deg$^2$, resolution $R=2,000$ for $\lambda \in [360, 1000) nm$
+    \item Silva 18S ribosomal DNA sequences of approximately 2.25 million genomes with aligned length of 50,000 letters. TODO: Look into other variants in the dataset particularly NR99.
+    \item Datasets from ANN-Benchmarks suite
+\end{itemize}
+
+Recent approaches in dealing with exponential growth of data:
+\begin{itemize}
+    \item LSH
+    \item clever indexing e.g. FM Index.
+    \item Entropy-scaling search~\cite{yu2015entropy} and CHESS~\cite{ishaq2019clustered}.
+\end{itemize}
+
+This paper focuses on:
+\begin{itemize}
+    \item Improvements and extensions of CHESS.
+    \item Accelerated $\rho$-nearest and k-nearest neighbors search.
+    \item Entropy-scaling (?? is it though?) compression using the hierarchical clustering.
+    \item Proof for how cluster-radii, and thus compression deltas, fall as we descend along the tree.
+    \item Information theoretic bounds on compression performance. We compress along the dimensionality of data rather than the cardinality of data.
+    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, wasserstein, hamming, levenshtein, needleman-wunch, and any user-defined function.
+\end{itemize}
+
+Comparisons for:
+\begin{itemize}
+    \item Search time vs FALCONN and other in ann-benchmarks suite.
+    \item Compression performance of CAKES+gzip vs just gzip.
+\end{itemize}

### 2021-03-09 sections/2-methods.tex
wrote outline

@@
 
+Talk about manifold hypothesis.
+
+\subsection{Definitions}
+\label{subsec:methods:definitions}
+
+\subsubsection{Dataset}
+\label{subsubsec:methods:definitions:dataset}
+
+General datasets with high dimensionality and high cardinality.
+We take a dataset to be any set $X$ with $n$ instances in $\mathcal{D}$ embedding dimensions where the data follow a low-dimensional manifold with fractal dimension $d$ where $d \ll \mathcal{D}$.
+
+\subsubsection{Distance Functions}
+\label{subsubsec:methods:definitions:distance-functions}
+
+General distance function which deterministically maps a pair of instances to a non-negative real number:
+\begin{itemize}
+    \item $dist: \mathbb{X} \times \mathbb{X} \mapsto \mathbb{R}^+$
+    \item $dist(x, y) = 0 \Leftrightarrow x = y$, identity
+    \item $dist(x, y) = dist(y, x) \ \forall x, y \in X$, symmetry
+    \item $dist(x - z, y - z) = dist(x, y) \ \forall x, y, z \in X$, invariance under translation, not valid for cosine distance or when subtraction is not defined.
+\end{itemize}
+
+\subsubsection{Local Fractal Dimension}
+\label{subsubsec:methods:definitions:local-fractal-dimension}
+
+LFD is a discrete approximation to Hausdorff Dimension:
+
+\begin{gather}
+    \log_2\bigg(\frac{|B_D(q, r_1)|}{|B_D(q, r_2)|}\bigg)
+    \label{fractal-dimension}
+\end{gather}
+
+where $B_D(q,r)$ is the set of instances contained in a ball on the dataset $D$ of radius $r$ centered on an instance $q$.
+We approximate the local fractal dimension by computing it for a radius $r_1$ and a smaller radius $r_2=\frac{r_1}{2}$.
+
+\subsubsection{Cluster}
+\label{subsubsec:methods:definitions:cluster}
+
+A \textit{cluster} is a set of instances with a \textit{center} (the geometric median) and a \textit{radius} (the maximum distance of any instance in the cluster from the cluster-center).
+Each cluster is either a leaf-cluster or has two child-clusters in the same way as nodes do in a binary tree.
+
+\subsubsection{Metric Entropy}
+\label{subsubsec:methods:definitions:metric-entropy}
+
+Same as in CHESS.
+
 \subsection{Proofs}
-\label{subserc:methods:proofs}
+\label{subsec:methods:proofs}
 
@@
 
-We take a dataset to be any set $X$ with $n$ instances in $\mathcal{D}$ embedding dimensions where the data follow a low-dimensional manifold with fractal dimension $d$ where $d \ll \mathcal{D}$.
-
-We take a distance function to be $f: X \times X \mapsto \mathbb{R}$ with the following properties:
-
-\begin{itemize}
-    \item $f(x, y) = 0 \ iff \ x = y$, identity
-    \item $f(x, y) = f(y, x) \ \forall x, y \in X$, symmetry
-    \item $f(x - z, y - z) = f(x, y) \ \forall x, y, z \in X$, invariance under translation 
-\end{itemize}
-
-We show that the radii of clusters are guaranteed to decrease after every $d$ partitions (at most).
+We show that the radii of clusters are guaranteed to decrease after, at most, every $d$ recursive partitions.
 
 Assume that the data follow a $d$-dimensional distribution embedded in $\mathcal{D}$-dimensional space.
-We can describe this distribution choosing some set of $d$ mutually orthagonal axes.
+We can describe this distribution by choosing some set of $d$ mutually orthagonal axes.
 Let $2r$ be the maximum distance between any pair of points along any axis from any choice of mutually orthagonal axes for our $d$-dimensional distribution.
-The cluster containing all these points will have a radius of $r$.
+The cluster containing these points will have a radius of $r$.
 In the worst case, the distribution is a $d$-sphere and $2r$ is the maximum distance along every axis from every choice of axes.
 
-The partition algorithm will choose the maximally distant pair of points and, thus, the axis defined by those points as the axis for partitioning the points into two child clusters.
+The partition algorithm will choose the maximally distant pair of points and, thus, the axis defined by those points for partitioning the cluster into two child clusters.
 For points in either child cluster, the maximum distance between any pair of points along that axis will have been reduced to $r$ or less.
@@
 After these $d$ partitions, the points in each child cluster will have a maximum pairwise distance of $r$ along any axis.
-Thus the radii of those child clusters will be at most $\frac{r}{2}$.
+Thus the radii of those child clusters will be at most $\frac{r}{\sqrt{2}}$.
+
+\subsubsection{Information-Theoretic properties of Compression}
+\label{subsubsec:methods:proofs:information-theoretic-properties-of-compression}
+
+TODO, TODO, TODO!
+
+\subsection{Algorithms}
+\label{subsec:methods:algorithms}
+
+\subsubsection{Partition}
+\label{subsubsec:methods:algorithms:partition}
+
+Partition algorithm for building the tree.
+Lift this from the CHAODA paper.
+
+% latex does not seem to like math symbols in this heading
+\subsubsection{$\rho$-Nearest Neighbors Search}
+\label{subsubsec:methods:algorithms:rnn-search}
+
+Modified binary search.
+Lift this from CHESS paper.
+
+\subsection{k-Nearest Neighbors Search}
+\label{subsubsec:methods:algorithms:knn-search}
+
+Repeated $\rho$-nearest neighbors search with modified radius with each iteration guided by lfd of query ball.
+Use from CHESS paper as a starting point.
+
+\subsubsection{Compression}
+\label{subsubsec:methods:algorithms:compression}
+
+Encode and Decode algorithms.
+These will be specific for each dataset and/or distance function.
+
+\subsection{Datasets and Distance Functions}
+\label{subsec:methods:datasets-and-distance-functions}
+
+Provide details on the datasets we use for benchmarks.
+
+Add t-SNE or UMAP embeddings of datasets. For example~\ref{fig:discussion:umap-annthyroid-euclidean}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
+    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
+    \label{fig:discussion:umap-annthyroid-euclidean}
+\end{figure}
+
+\subsubsection{APOGEE-2}
+\label{subsec:methods:datasets:apogee-2}
+
+euclidean, cosine, wasserstein-1d
+
+\dots
+
+\subsubsection{MaNGA}
+\label{subsec:methods:datasets:manga}
+
+euclidean, wasserstein-2d
+
+\dots
+
+\subsubsection{Silva 18S and variants}
+\label{subsec:methods:datasets:silva-18s}
+
+hamming, levenshtein
+
+\dots
+
+\subsubsection{ANN-Benchmarks suite}
+\label{subsec:methods:datasets:ann-benchmarks-suite}
+
+deep1b: cosine
+
+fashion-mnist: euclidean
+
+gist: euclidean
+
+glove: cosine
+
+kosarak: jaccard
+
+mnist: euclidean
+
+nytimes: cosine
+
+sift: euclidean
+
+last.fm: cosine
+
+\dots
+
+\subsection{Hardware}
+\label{subsec:methods:hardware}
+
+Specs of Ark, and maybe M1 MacBook Air and Dell XPS.

### 2021-03-09 sections/4-discussion.tex
wrote outline

@@
 
-Include figures~\ref{fig:discussion:umap-annthyroid-euclidean}
+CLAM-CAKES.
+This is most effective when datasets exhibit low metric entropy and low local fractal dimension.
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
-    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
-    \label{fig:discussion:umap-annthyroid-euclidean}
-\end{figure}
+Unlike LSH, this is extensible to any user provided distance function.
+
+Exactness of search and metric vs non-metric.
+
+\subsection{Future Works}
+\label{subsec:results:future-works}
+
+GPU-acceleration for the more computationally intensive distance functions like wasserstein.
+
+Live-updates as new points are added to the dataset.
+For example, build tree with a subsample of the full data and then add the rest of the instances to simulate live-updates.
+
+More distance functions. For example, maximal-common-subgraph for molecular structures.
+
+TODO: Provide link to source code on github.

### 2021-04-07 sections/2-methods.tex
notes
hierarchical extension of 2020 Berger et. al. paper

--- 2021-02-23 sections/2-methods.tex
@@
 
-Write about methods \dots
+\subsection{Partition}
+\label{subsec:methods:partition}
+
+Partition algorithm for building the tree.
+Lift this from the CHAODA paper.
+
+Exact partition costing $\mathcal{O}(n^2)$ vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+
+% latex does not seem to like math symbols in this heading
+\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
+\label{subsec:methods:rnn-search}
+
+Modified binary search.
+Lift this from CHESS paper.
+
+\subsection{k-Nearest Neighbors Search}
+\label{subsec:methods:knn-search}
+
+Repeated $\rho$-NN search with where radius is modified with each iteration as guided by lfd of query ball.
+
+Lift from CHESS paper as a starting point.
+
+\subsection{Compression}
+\label{subsec:methods:compression}
+
+Encode and Decode algorithms.
+These will be specific for each dataset and/or distance function.
+
+% discrete vs continuous, and small alphabet vs large alphabet in discrete

### 2021-04-07 sections/3-info-theory.tex
hierarchical extension of 2020 Berger et. al. paper

\section{Information Theoretic Properties}
\label{sec:information-theoretic-properties}

We extend the work from~\cite{berger2020levenshtein} for search and compression using a flat clustering to our case of a hierarchical clustering.
We further generalize this for any dataset and distance function described below.

We take a \textbf{dataset} $X$ with $n$ instances embedded in a $\mathcal{D}$-dimensional Banach Space defined by a distance function $f: X \times X \mapsto \mathbb{R}$.
We assume that the dataset follows the Manifold Hypothesis~\cite{fefferman2016testing}, i.e. instances in the dataset lie on a $d$-dimensional manifold where $d \ll \mathcal{D}$.

The \textbf{distance function} must have certain properties.
Specifically, $\forall x, y, z \in X$:

\begin{enumerate}[i.]
    \item $f(x, y) \mapsto \mathbb{R}$ is deterministic.
    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$, i.e. distance values are non-negative.
    \item $f(x, y) = f(y, x)$, i.e. symmetry.
    \item $f(x - z, y - z) = f(x, y)$, i.e. invariance under translation or encoding.
    \item $f(x, z) \leq f(x, y) + f(y, z)$, i.e. the triangle inequality holds.
\end{enumerate}

% TODO: Add note about the triangle inequality.
%  Not always needed, without it, search might miss some hits.

We define a \textbf{Metric Ball} as $B(q, r) = \{ x \in X \ | \ f(q, x) \leq r \}$, i.e. a ball containing all instances in the dataset within distance $r$ from an instance $q$.
We also define a \textbf{Cluster} as $C(c, r)$ analogous to a Metric Ball with the caveat that if some clusters have overlapping volumes then any instances in that overlapping volume are assigned to only one of the clusters.
Thus $C_i \cap C_j = \phi \Leftrightarrow i \neq j$ and $\bigcup C = X$.

The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e. the maximum number of $\epsilon$-separated instances in the dataset.

The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $\bigcup_{i = 1}^{n} C(c_i, \epsilon) = X$, i.e. the clusters $C_i$ collectively cover $X$.

Clearly, $N_{2\epsilon}^{ent}(X) \leq N_{\epsilon}^{int}(X) \leq N_{\epsilon}^{ent}(X)$.
Therefore, the Metric Entropy and the Internal Covering Number of a dataset are equivalent measures.

The \textbf{Fractal Dimension} (Minkowski Dimension) of a dataset is
\begin{align*}
    d(X) &= \lim_{\epsilon \rightarrow 0} \frac{\log(N_{\epsilon}^{int}(X))}{\log(\frac{1}{\epsilon})}
\end{align*}

However, if $|X|$ is finite then $d(X) = 0$.
To remedy this, we define the \textbf{Local Fractal Dimension}, as in~\cite{berger2020levenshtein}, for range $r$ and scale $s$ as
\begin{align*}
    d(X, r, s) &= \max_{x \in X} \ \log \frac{|B(x, r + s)| \ / \ |B(x, s)|}{(r + s) \ / \ s}
\end{align*}

For a cluster, $C_i(c_i, r_i)$, its local fractal dimension is equivalently defined as
\begin{align}
    \label{eq:local-fractal-dimension}
    d(C_i) &= \log_2 \ \frac{|B(c_i, r_i)|}{|B(c_i, \frac{r_i}{2})|}
\end{align}

Given an \textbf{Internal Covering} of $X$, we can specify any $\bar{x} \in X$to within error $\epsilon$ by specifying $x_i$ such that $\bar{x} \in B(x_i, \epsilon)$.
This it takes $\mathcal{O} \big( \log N_{\epsilon}^{ent}(X) \big)$ bits of information to specify $\bar{x}$ to within error $\epsilon$.

Given two instances $x_1, x_2 \in X$, we can construct $x_2 - x_1$, a \textbf{minimal encoding} of $x_2$ in terms of $x_1$, such that the number of bits needed to specify the encoding scales linearly with $f(x_1, x_2)$.
We would need to construct a method for construction such \textit{minimal encodings} for each pairing of $(X, f)$.
See Sections~\ref{subsec:methods:compression} and~\ref{sec:datasets-and-distance-functions}.

The \textbf{Shannon Entropy} of an instance $x$, denoted by $H(x)$, is the number of bits needed to exactly specify $x$.
Given an encoding, $x_2 - x_1$, of an instance $x_2$ in terms of another instance $x_1$, let $H(x_2 - x_1)$ be the Shannon Entropy of the encoding, i.e. the number of bits needed to store the encoding.
Therefore, \[ H \big( \{ x_1, x_2 \} \big) \leq H(x_1) + H(x_2 - x_1) \] where $\{ x_1, x_2 \}$ is the concatenation of $x_1$ and $x_2$.

For example, given a dataset of genomic strings with the Levenshtein edit distance as the metric, let $x_1$ and $x_2$ be two strings in the dataset.
Let $x_2 - x_1$ be the list of edits to convert $x_1$ into $x_2$ as used for the Levenshtein distance between $x_1$ and $x_2$.
The length of this edit list is equal to the Levenshtein edit distance i.e. $|x_2 - x_1| = f(x_1, x_2)$.
Each such edit can be stored using $\mathcal{O}(\log m)$ bits where $m$ is the maximum length of any string in the dataset.
Thus, the total number of bits needed for the encoding is $\mathcal{O} \big( f(x_1, x_2) \big)$.

Hence, for $|X| < \inf$,
\begin{align*}
    H \big( C(c, r) \big) &\leq H(c) + \sum_{x \in C} H(x - c) \\
    &\leq H(c) + |C(c, r)| \cdot H(y - c) \\ \\
    \text{where} \ y = \argmax_{x \in C}& \ f(x, c) \text{, or equivalently} \ f(y, c) = r.
\end{align*}

For a \textbf{Hierarchical Clustering}, let the cluster $C$ have parent $C_{parent}$, children $C_{left}$ and $C_{right}$, center $c \in C$, radius $r$, and an instance $y \in C$ such that $f(y, c) = r$.

We have two choices for encoding the instances in a cluster:
either each instance is directly encoded in terms of the cluster center, 
or each child-cluster is encoded in terms of the child-center and the child-centers are encoded in terms of the cluster-center.

Thus the Shannon Entropy of a cluster encoding is given by
\begin{align}
    \label{eq:hierarchical-shannon-entropy}
    H(C) &\leq H(c - c_{parent}) + \min \begin{cases}
        |C| \cdot H(y - c) \\
        H(C_{left}) + H(C_{right})
    \end{cases}
\end{align}

where the center of the root cluster is encoded directly rather than in terms of another instance.

\subsection{Guaranteed Decreases in Radii}
\label{subsec:methods:radii-decrease}

We show that the radii of clusters are guaranteed to decrease after, at most, every $d$ recursive partitions.

The Manifold Hypothesis assumption indicates that the instances in the dataset follow a $d$-dimensional distribution where $d \ll D$, the dimensionality of the embedding space.
We measure the dimensionality of this low-dimensional distribution using the local fractal dimension of the dataset.
Thus, we can describe this distribution by choosing some set of $d$ mutually orthogonal axes.

Let $2r$ be the maximum pairwise distance among the instances in the dataset.
In the worst case, e.g. with the distribution filling a $d$-sphere, our chosen $d$ axes will be such that $2r$ is the maximum pairwise distance \textit{along} each axis.

The Partition algorithm will select a maximally distant pair of points to use as poles, i.e. it will choose one of the $d$ axes along which to partition the cluster into two children.
After one application of this Partition, the maximum pairwise distance along that axis will be bounded above by $r$.
The next recursive application of Partition will select another of the $d$ axes.
Thus, after at most $d$ Partitions, the maximum pairwise distance along each of the $d$ axes will be bounded above by $r$.
The overall maximum pairwise distance, \textit{i.e. not restricted to be along any one axis}, will be bounded above by $r \sqrt{2}$, for example, between two instances that are at the extrema of two different axes.

Thus, starting with a Cluster $C$ of radius $r$, after at most $d$ Partitions, the descendants of $C$ will each have radius bounded above by $\frac{r}{\sqrt{2}}$.
In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{1}{\sqrt{2}}$ after at most $d$ Partitions where $d$ is the local fractal dimension of the manifold occupied by the data.

\subsection{Back to Shannon Entropy}
\label{subsec:methods:back-to-shannon-entropy}

Starting with the Hierarchical Shannon Entropy defined in Equation~\ref{eq:hierarchical-shannon-entropy}, we can now further analyze the case of encoding a cluster using its children.

In the worst case, we would have a perfectly balanced clustering.
Starting with cluster $C$ and applying $d$ recursive partitions would give us $2^d$ descendants each with $|C| \cdot 2^{-d}$ instances.
If cluster $C$ has radius $r$ then each of the $2^d$ descendants will have radius $\frac{r}{\sqrt{2}}$.
In the general case, we can assume that the descendants have $\mathcal{O}(|C| \cdot 2^{-d})$ instances each.

Note that, in practice, we never see a perfectly balanced clustering.
Instead, Partition produces unbalanced trees due the varying density of sampling of the instances in different regions of the manifold.
Further, the cluster radii decrease by a factor much larger than $\frac{1}{\sqrt{2}}$ and they do so every one or two partitions rather than after $d$ partitions.

Let $T(C, d)$ be the subtree rooted at $C$ with depth $d$ and $J$ leaves.
Since we cannot partition a cluster containing only one instance, we know that $J \leq 2^d \leq |C|$.
Let the leaves be indexed with integers $j \in [1, J]$.
Let the center of each child cluster be encoded in terms of the center of its parent cluster.
Therefore, $H \big( T(C, d) \big)$, the Shannon Entropy of $T(C, d)$, is the sum of all encodings of child-centers in terms of their parent-centers..

Thus, the second case in Equation~\ref{eq:hierarchical-shannon-entropy} becomes:
\begin{align*}
    & H \big( T(C, d) \big) + \sum H(C_j) \\
    = \ & H \big( T(C, d) \big) + \sum |C_j| \cdot H(y_j - c_j)
\end{align*}

By the scaling behavior of cluster radii, as discussed in Section~\ref{subsec:methods:radii-decrease}, we know $H(y_j - c_j) \leq \frac{1}{\sqrt{2}} H(y - c)$.
Thus, we continue:
\begin{align*}
    \leq \ & H \big( T(C, d) \big) + \frac{1}{\sqrt{2}} \cdot \sum |C_j| \cdot H(y - c) \\
    = \ & H \big( T(C, d) \big) + \frac{1}{\sqrt{2}} \cdot |C| \cdot H(y - c)
\end{align*}

So we rewrite Equation~\ref{eq:hierarchical-shannon-entropy} as
\begin{align}
    \label{eq:hierarchical-shannon-entropy-2}
    H(C) &\leq H(c - c_{parent}) + \min \begin{cases}
        |C| \cdot H(y - c) \\
        |C| \cdot H(y - c) \cdot \frac{1}{\sqrt{2}} + H \big( T(C, d) \big) 
    \end{cases}
\end{align}

Thus, we have a measurable trade-off when encoding the instances in a cluster as to whether to encode in terms of the cluster-center or to encode using the subtree rooted at that cluster up to a low depth guided by the local fractal fractal dimension of that cluster.

In practice, for low depths in the tree, using subtrees to encode the instances produces great gains in the number of bits used.
As we continue deeper into the full tree, we start seeing diminishing returns for using subtrees.
Eventually, either the two cases equalize or the first cases becomes the better choice.
At such a depth in the tree, we encode the instances in a cluster using the cluster center rather than the subtree.
Notably, such a depth need not be uniform across the tree and can adapt to the sampling density in different regions of the manifold.

### 2021-04-07 sections/4-datasets.tex
hierarchical extension of 2020 Berger et. al. paper

\section{Datasets and Distance Functions}
\label{sec:datasets-and-distance-functions}

Provide details on the datasets we use for benchmarks.

Add t-SNE or UMAP embeddings of datasets. For example~\ref{fig:discussion:umap-annthyroid-euclidean}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
    \label{fig:discussion:umap-annthyroid-euclidean}
\end{figure}

\subsection{APOGEE-2}
\label{subsec:datasets:apogee-2}

euclidean, cosine, wasserstein-1d

\dots

\subsection{MaNGA}
\label{subsec:datasets:manga}

euclidean, wasserstein-2d

\dots

\subsection{Silva 18S and variants}
\label{subsec:datasets:silva-18s}

hamming, levenshtein

\dots

\subsection{ANN-Benchmarks suite}
\label{subsec:datasets:ann-benchmarks-suite}

deep1b: cosine

fashion-mnist: euclidean

gist: euclidean

glove: cosine

kosarak: jaccard

mnist: euclidean

nytimes: cosine

sift: euclidean

last.fm: cosine

\dots

### 2021-04-07 sections/5-results.tex
hierarchical extension of 2020 Berger et. al. paper

\section{Results}
\label{sec:results}

Write about results \dots

\subsection{Asymptotic Complexity}
\label{subsec:results:asymptotic-complexity}

Analyze:
\begin{itemize}
    \item Building the tree.
    \item $\rho$-nearest neighbors search.
    \item k-nearest neighbors search.
    \item compression:
    \begin{itemize}
        \item encode (novel contribution)
        \item serialize
        \item compress
        \item decompress
        \item deserialize
        \item decode (novel contribution)
    \end{itemize}
\end{itemize}

\subsection{Benchmarks}
\label{subsec:results:benchmarks}

Specs of Ark, and maybe M1 MacBook Air and Dell XPS.

Provide link to source code on GitHub.

For each pair $(X, f)$ and a small selection of depths, measure:
\begin{enumerate}[1.]
    \item linear search vs accelerated search:
    \begin{enumerate}[i.]
        \item runtime performance,
        \item speedup factor, and
        \item number of distance comparisons.
    \end{enumerate}
    \item Compression CAKES + zlib vs just zlib, also substitute other libraries for zlib:
    \begin{enumerate}[i.]
        \item just the encoding vs the full dataset,
        \item encoding + serialization, and
        \item encoding + serialization + zlib.
    \end{enumerate}
\end{enumerate}

Find the optimal clusters for compression.
Compare linear search vs accelerated search terminated at these optimal clusters.

### 2021-04-07 sections/2-info-theory.tex
typos, reorganization, and more

--- 2021-04-07 sections/3-info-theory.tex
@@
-\section{Information Theoretic Properties}
-\label{sec:information-theoretic-properties}
+\section{Information Theoretic Analysis}
+\label{sec:information-theoretic-analysis}
 
@@
 We further generalize this for any dataset and distance function described below.
+We also provide relevant algorithms for some parings of datasets and distance functions in Section~\ref{subsec:methods:compression}.
 
@@
 \begin{enumerate}[i.]
-    \item $f(x, y) \mapsto \mathbb{R}$ is deterministic.
-    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$, i.e. distance values are non-negative.
+    \item $f: X \times X \mapsto \mathbb{R}$ is deterministic.
+    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$.
     \item $f(x, y) = f(y, x)$, i.e. symmetry.
-    \item $f(x - z, y - z) = f(x, y)$, i.e. invariance under translation or encoding.
-    \item $f(x, z) \leq f(x, y) + f(y, z)$, i.e. the triangle inequality holds.
+    \item $f(x, z) \leq f(x, y) + f(y, z)$, i.e. the triangle inequality.
 \end{enumerate}
@@
 We also define a \textbf{Cluster} as $C(c, r)$ analogous to a Metric Ball with the caveat that if some clusters have overlapping volumes then any instances in that overlapping volume are assigned to only one of the clusters.
-Thus $C_i \cap C_j = \phi \Leftrightarrow i \neq j$ and $\bigcup C = X$.
-
-The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e.
-the maximum number of $\epsilon$-separated instances in the dataset.
-
-The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $\bigcup_{i = 1}^{n} C(c_i, \epsilon) = X$, i.e.
-the clusters $C_i$ collectively cover $X$.
-
-Clearly, $N_{2\epsilon}^{ent}(X) \leq N_{\epsilon}^{int}(X) \leq N_{\epsilon}^{ent}(X)$.
-Therefore, the Metric Entropy and the Internal Covering Number of a dataset are equivalent measures.
+Thus $C_i \cap C_j = \phi \Leftrightarrow i \neq j$ and $\bigcup C_i = X$.
 
@@
 \begin{align*}
-    d(X) &= \lim_{\epsilon \rightarrow 0} \frac{\log(N_{\epsilon}^{int}(X))}{\log(\frac{1}{\epsilon})}
+    d(X) &= \lim_{\epsilon \rightarrow 0} \frac{\log \big( N_{\epsilon}^{int}(X) \big) }{\log \big( \frac{1}{\epsilon} \big)}
 \end{align*}
@@
 
-For a cluster, $C_i(c_i, r_i)$, its local fractal dimension is equivalently defined as
+For a cluster, $C(c, r)$, its local fractal dimension is equivalently defined as
 \begin{align}
     \label{eq:local-fractal-dimension}
-    d(C_i) &= \log_2 \ \frac{|B(c_i, r_i)|}{|B(c_i, \frac{r_i}{2})|}
+    d(C) &= \log_2 \ \frac{|C(c, r)|}{|C(c, \frac{r}{2})|}
 \end{align}
 
-Given an \textbf{Internal Covering} of $X$, we can specify any $\bar{x} \in X$to within error $\epsilon$ by specifying $x_i$ such that $\bar{x} \in B(x_i, \epsilon)$.
+The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $n$ such that $\{x_1, \dots, x_n\} \subseteq X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e.
+the maximum number of $\epsilon$-separated instances in the dataset.
+
+The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $\bigcup_{i = 1}^{n} C(c_i, \epsilon) = X$, i.e.
+the clusters $C_i$ collectively cover $X$.
+
+As in~\cite{berger2020levenshtein}, $N_{2\epsilon}^{ent}(X) \leq N_{\epsilon}^{int}(X) \leq N_{\epsilon}^{ent}(X)$, i.e. the Metric Entropy bounds the Internal Covering Number.
+Thus, they are equivalent measures.
+
+Given an \textbf{Internal Covering} of $X$, we can specify any $\bar{x} \in X$ to within error $\epsilon$ by specifying $x_i$ such that $\bar{x} \in B(x_i, \epsilon)$.
 This it takes $\mathcal{O} \big( \log N_{\epsilon}^{ent}(X) \big)$ bits of information to specify $\bar{x}$ to within error $\epsilon$.
@@
 We would need to construct a method for construction such \textit{minimal encodings} for each pairing of $(X, f)$.
-See Sections~\ref{subsec:methods:compression} and~\ref{sec:datasets-and-distance-functions}.
+See Sections~\ref{subsec:methods:compression} and~\ref{sec:datasets-and-distance-functions} for some examples.
 
@@
 For example, given a dataset of genomic strings with the Levenshtein edit distance as the metric, let $x_1$ and $x_2$ be two strings in the dataset.
-Let $x_2 - x_1$ be the list of edits to convert $x_1$ into $x_2$ as used for the Levenshtein distance between $x_1$ and $x_2$.
+Let $x_2 - x_1$ be the shortest list of edits to convert $x_1$ into $x_2$.
 The length of this edit list is equal to the Levenshtein edit distance i.e. $|x_2 - x_1| = f(x_1, x_2)$.
-Each such edit can be stored using $\mathcal{O}(\log m)$ bits where $m$ is the maximum length of any string in the dataset.
-Thus, the total number of bits needed for the encoding is $\mathcal{O} \big( f(x_1, x_2) \big)$.
+As in~\cite{berger2020levenshtein}, each such edit can be stored using $\mathcal{O}(\log m)$ bits where $m$ is the maximum length of any string in the dataset.
+Thus, the total number of bits needed for each such encoding is $\mathcal{O} \big( f(x_1, x_2) \big)$.
 
-Hence, for $|X| < \inf$,
+Hence, for $|X| < \infty$,
 \begin{align*}
-    H \big( C(c, r) \big) &\leq H(c) + \sum_{x \in C} H(x - c) \\
-    &\leq H(c) + |C(c, r)| \cdot H(y - c) \\ \\
-    \text{where} \ y = \argmax_{x \in C}& \ f(x, c) \text{, or equivalently} \ f(y, c) = r.
+    H (C) &\leq H(c) + \sum_{x \in C} H(x - c) \\
+    &\leq H(c) + |C| \cdot H(y - c) \\ \\
+    \text{where} \ y &= \argmax_{x \in C} \ f(x, c) \text{, or equivalently} \ f(y, c) = r.
 \end{align*}
@@
 We have two choices for encoding the instances in a cluster:
-either each instance is directly encoded in terms of the cluster center, 
-or each child-cluster is encoded in terms of the child-center and the child-centers are encoded in terms of the cluster-center.
+either each instance can be encoded in terms of the cluster center, 
+or each child-cluster can be encoded in terms of the child-center and the child-centers encoded in terms of the cluster-center.
 
@@
     \label{eq:hierarchical-shannon-entropy}
-    H(C) &\leq H(c - c_{parent}) + \min \begin{cases}
+    H(C) &\leq H(c - c_{parent}) + min \begin{cases}
         |C| \cdot H(y - c) \\
@@
 
-\subsection{Guaranteed Decreases in Radii}
-\label{subsec:methods:radii-decrease}
+\subsection{Scaling Behavior of Cluster Radii}
+\label{subsec:methods:radii-scaling-behavior}
 
-We show that the radii of clusters are guaranteed to decrease after, at most, every $d$ recursive partitions.
+We show that the radii of clusters are guaranteed to decrease after, at most, every $d$ recursive applications of Partition.
+The Partition algorithm is described in Section~\ref{subsec:methods:partition}.
 
-The Manifold Hypothesis assumption indicates that the instances in the dataset follow a $d$-dimensional distribution where $d \ll D$, the dimensionality of the embedding space.
-We measure the dimensionality of this low-dimensional distribution using the local fractal dimension of the dataset.
+The Manifold Hypothesis indicates that the instances in the dataset follow a $d$-dimensional distribution with $d \ll D$ where $D$ is the dimensionality of the embedding space.
+We can measure the dimensionality of this low-dimensional distribution using the local fractal dimension of the dataset.
 Thus, we can describe this distribution by choosing some set of $d$ mutually orthogonal axes.
@@
 Let $2r$ be the maximum pairwise distance among the instances in the dataset.
-In the worst case, e.g. with the distribution filling a $d$-sphere, our chosen $d$ axes will be such that $2r$ is the maximum pairwise distance \textit{along} each axis.
+In the worst case, e.g. with the distribution shaped like a $d$-sphere, our chosen axes will be such that $2r$ is the maximum pairwise distance \textit{along every axis}.
+Such a distribution would produce a balanced clustering.
 
-The Partition algorithm will select a maximally distant pair of points to use as poles, i.e. it will choose one of the $d$ axes along which to partition the cluster into two children.
-After one application of this Partition, the maximum pairwise distance along that axis will be bounded above by $r$.
-The next recursive application of Partition will select another of the $d$ axes.
-Thus, after at most $d$ Partitions, the maximum pairwise distance along each of the $d$ axes will be bounded above by $r$.
+Partition will select a maximally distant pair of points to use as poles, i.e. it will choose one of the $d$ axes along which to split the cluster into two children.
+After one Partition, the maximum pairwise distance along that axis will be bounded above by $r$.
+The next recursive Partition will select another of the $d$ axes.
+Thus, after at most $d$ Partitions, the maximum pairwise distance along each axis will be bounded above by $r$.
 The overall maximum pairwise distance, \textit{i.e.
-not restricted to be along any one axis}, will be bounded above by $r \sqrt{2}$, for example, between two instances that are at the extrema of two different axes.
+not restricted to be along one axis}, will be bounded above by $r \sqrt{2}$, for example, between two instances that lie at the extrema of different axes.
 
@@
 In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{1}{\sqrt{2}}$ after at most $d$ Partitions where $d$ is the local fractal dimension of the manifold occupied by the data.
+
+Note that, in practice, we never see a balanced clustering.
+Instead, Partition produces unbalanced trees due the varying density of sampling in different regions of the manifold.
+Further, the cluster radii decrease by a factor much larger than $\frac{1}{\sqrt{2}}$ and they do so every one or two partitions rather than after $d$ partitions.
+See Section~\ref{sec:results} for results on several datasets.
 
@@
 
-In the worst case, we would have a perfectly balanced clustering.
-Starting with cluster $C$ and applying $d$ recursive partitions would give us $2^d$ descendants each with $|C| \cdot 2^{-d}$ instances.
+In the worst case, we would have a balanced clustering.
+Starting with cluster $C$ and applying $d$ recursive partitions would give us $2^d$ descendants with $|C| \cdot 2^{-d}$ instances each.
 If cluster $C$ has radius $r$ then each of the $2^d$ descendants will have radius $\frac{r}{\sqrt{2}}$.
-In the general case, we can assume that the descendants have $\mathcal{O}(|C| \cdot 2^{-d})$ instances each.
 
-Note that, in practice, we never see a perfectly balanced clustering.
-Instead, Partition produces unbalanced trees due the varying density of sampling of the instances in different regions of the manifold.
-Further, the cluster radii decrease by a factor much larger than $\frac{1}{\sqrt{2}}$ and they do so every one or two partitions rather than after $d$ partitions.
-
-Let $T(C, d)$ be the subtree rooted at $C$ with depth $d$ and $J$ leaves.
-Since we cannot partition a cluster containing only one instance, we know that $J \leq 2^d \leq |C|$.
+In the general case, let $T(C, d)$ be the subtree rooted at $C$ with depth $d$ and $J$ leaves.
+Since we cannot partition a cluster containing only one instance, we can bound $J$ and $d$, i.e. $J \leq 2^d \leq |C|$.
 Let the leaves be indexed with integers $j \in [1, J]$.
-Let the center of each child cluster be encoded in terms of the center of its parent cluster.
-Therefore, $H \big( T(C, d) \big)$, the Shannon Entropy of $T(C, d)$, is the sum of all encodings of child-centers in terms of their parent-centers..
+Let the center of each child-cluster be encoded in terms of the center of its parent-cluster.
+Therefore, $H(T)$, the Shannon Entropy of $T(C, d)$, is the sum of all encodings of child-centers in terms of their parent-centers.
 
@@
 \begin{align*}
-    & H \big( T(C, d) \big) + \sum H(C_j) \\
-    = \ & H \big( T(C, d) \big) + \sum |C_j| \cdot H(y_j - c_j)
+    & H(T) + \sum H(C_j) \\
+    = \ & H(T) + \sum |C_j| \cdot H(y_j - c_j)
 \end{align*}
 
-By the scaling behavior of cluster radii, as discussed in Section~\ref{subsec:methods:radii-decrease}, we know $H(y_j - c_j) \leq \frac{1}{\sqrt{2}} H(y - c)$.
+From the analysis of the scaling behavior of cluster-radii in Section~\ref{subsec:methods:radii-scaling-behavior}, we have $H(y_j - c_j) \leq \frac{1}{\sqrt{2}} H(y - c)$.
 Thus, we continue:
 \begin{align*}
-    \leq \ & H \big( T(C, d) \big) + \frac{1}{\sqrt{2}} \cdot \sum |C_j| \cdot H(y - c) \\
-    = \ & H \big( T(C, d) \big) + \frac{1}{\sqrt{2}} \cdot |C| \cdot H(y - c)
+    \leq \ & H(T) + \frac{1}{\sqrt{2}} \cdot \sum |C_j| \cdot H(y - c) \\
+    = \ & H(T) + \frac{1}{\sqrt{2}} \cdot |C| \cdot H(y - c)
 \end{align*}
@@
         |C| \cdot H(y - c) \\
-        |C| \cdot H(y - c) \cdot \frac{1}{\sqrt{2}} + H \big( T(C, d) \big) 
+        |C| \cdot H(y - c) \cdot \frac{1}{\sqrt{2}} + H(T)
     \end{cases}
@@
 
-In practice, for low depths in the tree, using subtrees to encode the instances produces great gains in the number of bits used.
-As we continue deeper into the full tree, we start seeing diminishing returns for using subtrees.
-Eventually, either the two cases equalize or the first cases becomes the better choice.
-At such a depth in the tree, we encode the instances in a cluster using the cluster center rather than the subtree.
-Notably, such a depth need not be uniform across the tree and can adapt to the sampling density in different regions of the manifold.
+For low depths in the tree, we expect that using subtrees to encode the instances will lead to great savings in the number of bits used.
+As we continue deeper into the tree, we would start seeing diminishing returns for using subtrees.
+Eventually, either the two cases will equalize or the first case will become the better choice.
+At such a depth in the tree, we can encode the instances in a cluster using the cluster center rather than the subtree.
+The center of that cluster can be encoded in terms of the center of its parent, and so on back up the tree to the root.
+
+Notably, such a depth need not be uniform across the tree.
+Indeed, the depth can adapt to the variations in sampling density and local fractal dimension in different regions of the manifold.

### 2021-04-07 sections/3-methods.tex
typos, reorganization, and more

--- 2021-04-07 sections/2-methods.tex
@@
 \label{sec:methods}
+
+Write about methods and provide asymptotic complexity analysis of each method.
 
@@
 
-Exact partition costing $\mathcal{O}(n^2)$ vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
 
-% latex does not seem to like math symbols in this heading
 \subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
@@
 
-\subsection{k-Nearest Neighbors Search}
+Complexity is the same as in CHESS.
+
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
 \label{subsec:methods:knn-search}
@@
 
+Complexity is some factor times CHESS-search.
+
 \subsection{Compression}
@@
 Encode and Decode algorithms.
-These will be specific for each dataset and/or distance function.
+These could be, and probably will be, different for each $(X, f)$ pairing.
+
+Asymptotic complexity analysis for:
+\begin{enumerate}[i.]
+    \item encode (novel contribution)
+    \item serialize
+    \item compress
+    \item decompress
+    \item deserialize
+    \item decode (novel contribution)
+\end{enumerate}
 

### 2021-04-07 sections/6-discussion.tex
hierarchical extension of 2020 Berger et. al. paper
typos, reorganization, and more

--- 2021-03-09 sections/4-discussion.tex
@@
 
+Is it the case that the optimal clusters for compression are also the optimal clusters for accelerated search?
+
+What kind of trade-offs can we offer for space-savings of compression vs time-savings of accelerated-search in the compressed space?
+
 \subsection{Future Works}
@@
 
-More distance functions. For example, maximal-common-subgraph for molecular structures.
+More distance functions. For example, Tanimoto Distance using maximal-common-subgraph for molecular structures.
 
-TODO: Provide link to source code on github.
+Better methods for constructing minimal encodings, perhaps using domain expertise.

### 2021-04-07 sections/2-info-theory.tex
uniformity assumption, tree equation, and minor cleanup

@@
 We assume that the dataset follows the Manifold Hypothesis~\cite{fefferman2016testing}, i.e. instances in the dataset lie on a $d$-dimensional manifold where $d \ll \mathcal{D}$.
+We also make some uniformity assumptions for the dataset;
+specifically that $|X|$ is finite and the minimum and maximum sampling densities in the dataset differ by some constant multiplicative factor.
 
-The \textbf{distance function} must have certain properties.
-Specifically, $\forall x, y, z \in X$:
-
+The \textbf{distance function} must have certain properties;
+specifically, $\forall x, y, z \in X$:
 \begin{enumerate}[i.]
-    \item $f: X \times X \mapsto \mathbb{R}$ is deterministic.
+    \item $f$ is deterministic.
     \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$.
@@
 \end{align*}
+However, if $|X|$ is finite then $d(X) = 0$.
 
-However, if $|X|$ is finite then $d(X) = 0$.
 To remedy this, we define the \textbf{Local Fractal Dimension}, as in~\cite{berger2020levenshtein}, for range $r$ and scale $s$ as
@@
 
-In the general case, let $T(C, d)$ be the subtree rooted at $C$ with depth $d$ and $J$ leaves.
-Since we cannot partition a cluster containing only one instance, we can bound $J$ and $d$, i.e. $J \leq 2^d \leq |C|$.
-Let the leaves be indexed with integers $j \in [1, J]$.
+In the general case, let $T(C, d)$ be the subtree rooted at $C$ with depth $d$ and up-to $2^d$ leaves.
+Since we cannot partition a cluster containing only one instance, we can bound the number of leaves, i.e. $2^d \leq |C|$.
+Let the leaves be indexed with integers $j \in [1, 2^{d + 1} - 1]$ as obtained by a breadth-first traversal of the tree, i.e.
+the root has index $1$ and a cluster with index $i$ has children with indices $2i$ and $2i + 1$.
 Let the center of each child-cluster be encoded in terms of the center of its parent-cluster.
-Therefore, $H(T)$, the Shannon Entropy of $T(C, d)$, is the sum of all encodings of child-centers in terms of their parent-centers.
+Therefore, $H(T)$, the Shannon Entropy of $T(C, d)$, is given by
+\begin{align*}
+    H(T) &= H(c_2 - c_1) + H(c_3 - c_1) + H(c_4 - c_2) + H(c_5 - c_2) \\
+    & \ \ \ \ + H(c_6 - c_3) + H(c_7 - c_3) + \dots + H(c_{2^{d + 1} - 1} - c_{2^d})
+\end{align*}
+i.e. the sum of all encodings of child-centers in terms of their parent-centers, where $c_1$ is the center of the root cluster.
 
-Thus, the second case in Equation~\ref{eq:hierarchical-shannon-entropy} becomes:
+Let us re-index the \textit{leaves} in $T(C, d)$ with integers $j \in [1, 2^d]$.
+The second case in Equation~\ref{eq:hierarchical-shannon-entropy} becomes:
 \begin{align*}
@@
 
-From the analysis of the scaling behavior of cluster-radii in Section~\ref{subsec:methods:radii-scaling-behavior}, we have $H(y_j - c_j) \leq \frac{1}{\sqrt{2}} H(y - c)$.
+From the analysis in Section~\ref{subsec:methods:radii-scaling-behavior}, we have that $H(y_j - c_j) \leq \frac{1}{\sqrt{2}} H(y - c) \ \forall \ j \in [1, 2^d]$.
 Thus, we continue:
 \begin{align*}
-    \leq \ & H(T) + \frac{1}{\sqrt{2}} \cdot \sum |C_j| \cdot H(y - c) \\
-    = \ & H(T) + \frac{1}{\sqrt{2}} \cdot |C| \cdot H(y - c)
+    \leq \ & H(T) + \frac{1}{\sqrt{2}} \cdot H(y - c) \cdot \sum |C_j| \\
+    = \ & H(T) + \frac{1}{\sqrt{2}} \cdot H(y - c) \cdot |C|
 \end{align*}
 
-So we rewrite Equation~\ref{eq:hierarchical-shannon-entropy} as
+We can finally rewrite Equation~\ref{eq:hierarchical-shannon-entropy} as
 \begin{align}
@@
 
-Thus, we have a measurable trade-off when encoding the instances in a cluster as to whether to encode in terms of the cluster-center or to encode using the subtree rooted at that cluster up to a low depth guided by the local fractal fractal dimension of that cluster.
+To encode an entire dataset, we start with the dataset contained in a single root cluster.
+We then check between the two cases for Equation~\ref{eq:hierarchical-shannon-entropy-2}.
+If the second case provides the smaller Shannon Entropy, then we recursively apply Equation~\ref{eq:hierarchical-shannon-entropy-2} to the leaves of $T(C,d)$.
 
-For low depths in the tree, we expect that using subtrees to encode the instances will lead to great savings in the number of bits used.
+Thus, we have a measurable trade-off when encoding the instances in a cluster.
+This trade-off is calculated using the subtree rooted at that cluster up to a low depth equal to the local fractal fractal dimension of that cluster.
+
+When starting out encoding the dataset in this fashion, for low depths in the tree, we expect that using subtrees to encode the instances will lead to great savings in the number of bits used.
 As we continue deeper into the tree, we would start seeing diminishing returns for using subtrees.
-Eventually, either the two cases will equalize or the first case will become the better choice.
-At such a depth in the tree, we can encode the instances in a cluster using the cluster center rather than the subtree.
-The center of that cluster can be encoded in terms of the center of its parent, and so on back up the tree to the root.
+Eventually, either the difference will be so small that it is not worth the computational cost of Partition, or the two cases will equalize, or the first case will become the better choice.
+At such a depth in the tree, we can encode the instances in a cluster using the cluster-center rather than the subtree.
+That cluster-center can be encoded in terms of the center of its parent and so on back up the tree to the root.
 
 Notably, such a depth need not be uniform across the tree.
-Indeed, the depth can adapt to the variations in sampling density and local fractal dimension in different regions of the manifold.
+Indeed, the depth can adapt to any variations in sampling density and local fractal dimension in different regions of the manifold.

### 2021-04-07 sections/2-info-theory.tex
fixes and typos

@@
 
-The \textbf{distance function} must have certain properties;
-specifically, $\forall x, y, z \in X$:
+The \textbf{distance function} must have must be such that $\forall x, y, z \in X$:
 \begin{enumerate}[i.]
@@
 We also define a \textbf{Cluster} as $C(c, r)$ analogous to a Metric Ball with the caveat that if some clusters have overlapping volumes then any instances in that overlapping volume are assigned to only one of the clusters.
-Thus $C_i \cap C_j = \phi \Leftrightarrow i \neq j$ and $\bigcup C_i = X$.
+Thus $\bigcup C_i = X$ and $i \neq j \Rightarrow C_i \cap C_j = \phi$.
+
+The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $n$ such that $\{x_1, \dots, x_n\} \subseteq X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e.
+the maximum number of $\epsilon$-separated instances in the dataset.
+
+The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $\bigcup_{i = 1}^{n} C(c_i, \epsilon) = X$, i.e.
+the minimum number of $\epsilon$-radius clusters that collectively cover $X$.
+
+As in~\cite{berger2020levenshtein}, $N_{2\epsilon}^{ent}(X) \leq N_{\epsilon}^{int}(X) \leq N_{\epsilon}^{ent}(X)$, i.e. the Metric Entropy bounds the Internal Covering Number.
+Thus, they are equivalent measures.
+
+Given an \textbf{Internal Covering} of $X$, we can specify any $\bar{x} \in X$ to within error $\epsilon$ by specifying $x_i$ such that $\bar{x} \in B(x_i, \epsilon)$.
+Thus it takes $\mathcal{O} \big( \log N_{\epsilon}^{ent}(X) \big)$ bits of information to specify $\bar{x}$ to within error $\epsilon$.
+
+Given two instances $x_1, x_2 \in X$, we can construct $x_2 - x_1$, a \textbf{minimal encoding} of $x_2$ in terms of $x_1$, such that the number of bits needed to specify the encoding scales linearly with $f(x_1, x_2)$.
+We would need a potentially different method for constructing such \textit{minimal encodings} for each pairing of $(X, f)$.
+See Sections~\ref{subsec:methods:compression} and~\ref{sec:datasets-and-distance-functions} for some examples.
 
@@
 
-The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $n$ such that $\{x_1, \dots, x_n\} \subseteq X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e.
-the maximum number of $\epsilon$-separated instances in the dataset.
+The \textbf{Shannon Entropy} of an instance $x$, denoted by $H(x)$, is the number of bits needed to exactly specify $x$.
 
-The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $\bigcup_{i = 1}^{n} C(c_i, \epsilon) = X$, i.e.
-the clusters $C_i$ collectively cover $X$.
-
-As in~\cite{berger2020levenshtein}, $N_{2\epsilon}^{ent}(X) \leq N_{\epsilon}^{int}(X) \leq N_{\epsilon}^{ent}(X)$, i.e. the Metric Entropy bounds the Internal Covering Number.
-Thus, they are equivalent measures.
-
-Given an \textbf{Internal Covering} of $X$, we can specify any $\bar{x} \in X$ to within error $\epsilon$ by specifying $x_i$ such that $\bar{x} \in B(x_i, \epsilon)$.
-This it takes $\mathcal{O} \big( \log N_{\epsilon}^{ent}(X) \big)$ bits of information to specify $\bar{x}$ to within error $\epsilon$.
-
-Given two instances $x_1, x_2 \in X$, we can construct $x_2 - x_1$, a \textbf{minimal encoding} of $x_2$ in terms of $x_1$, such that the number of bits needed to specify the encoding scales linearly with $f(x_1, x_2)$.
-We would need to construct a method for construction such \textit{minimal encodings} for each pairing of $(X, f)$.
-See Sections~\ref{subsec:methods:compression} and~\ref{sec:datasets-and-distance-functions} for some examples.
-
-The \textbf{Shannon Entropy} of an instance $x$, denoted by $H(x)$, is the number of bits needed to exactly specify $x$.
 Given an encoding, $x_2 - x_1$, of an instance $x_2$ in terms of another instance $x_1$, let $H(x_2 - x_1)$ be the Shannon Entropy of the encoding, i.e.
 the number of bits needed to store the encoding.
-Therefore, \[ H \big( \{ x_1, x_2 \} \big) \leq H(x_1) + H(x_2 - x_1) \] where $\{ x_1, x_2 \}$ is the concatenation of $x_1$ and $x_2$.
+Therefore,
+\begin{align*}
+    H \big( \{ x_1, x_2 \} \big) \leq H(x_1) + H(x_2 - x_1)
+\end{align*}
+where $\{ x_1, x_2 \}$ is the concatenation of $x_1$ and $x_2$.
 
 For example, given a dataset of genomic strings with the Levenshtein edit distance as the metric, let $x_1$ and $x_2$ be two strings in the dataset.
-Let $x_2 - x_1$ be the shortest list of edits to convert $x_1$ into $x_2$.
+Let $x_2 - x_1$ be a (possibly non-unique) shortest list of edits to convert $x_1$ into $x_2$.
 The length of this edit list is equal to the Levenshtein edit distance i.e. $|x_2 - x_1| = f(x_1, x_2)$.
@@
 
-Hence, for $|X| < \infty$,
+Let $H(C)$ be the Shannon Entropy of the cluster $C(c, r)$.
+Hence,
 \begin{align*}
@@
 
-For a \textbf{Hierarchical Clustering}, let the cluster $C$ have parent $C_{parent}$, children $C_{left}$ and $C_{right}$, center $c \in C$, radius $r$, and an instance $y \in C$ such that $f(y, c) = r$.
+For a \textbf{Hierarchical Clustering}, let the cluster $C$ have 
+parent $C_{parent}$, 
+children $C_{left}$ and $C_{right}$, 
+center $c \in C$, 
+radius $r \in \mathbb{R}^+$, 
+and an instance $y \in C$ such that $f(y, c) = r$.
 
 We have two choices for encoding the instances in a cluster:
-either each instance can be encoded in terms of the cluster center, 
+either each instance can be encoded in terms of the cluster-center, 
 or each child-cluster can be encoded in terms of the child-center and the child-centers encoded in terms of the cluster-center.
@@
 Let $2r$ be the maximum pairwise distance among the instances in the dataset.
-In the worst case, e.g. with the distribution shaped like a $d$-sphere, our chosen axes will be such that $2r$ is the maximum pairwise distance \textit{along every axis}.
-Such a distribution would produce a balanced clustering.
+We choose the axes such that the two points that are $2r$ apart lie along one of the axes.
+Thus, a $d$-dimensional hyper-sphere of radius $r$ would bound the dataset.
+In the worst case, e.g. with a distribution that fills the $d$-sphere, our axes will be such that $2r$ is the maximum pairwise distance \textit{along every axis}.
+Such a distribution would also produce a balanced clustering.
 
@@
 Thus, after at most $d$ Partitions, the maximum pairwise distance along each axis will be bounded above by $r$.
-The overall maximum pairwise distance, \textit{i.e.
-not restricted to be along one axis}, will be bounded above by $r \sqrt{2}$, for example, between two instances that lie at the extrema of different axes.
+The overall maximum pairwise distance, i.e.
+\textit{not restricted} to be along one axis, will be bounded above by $r \sqrt{2}$ by, for example, two instances that lie at the extrema of different axes.
 
@@
 Note that, in practice, we never see a balanced clustering.
-Instead, Partition produces unbalanced trees due the varying density of sampling in different regions of the manifold.
+Instead, Partition produces unbalanced trees due to the varying density of sampling in different regions of the manifold and the low-dimensional ``shape'' of the manifold.
 Further, the cluster radii decrease by a factor much larger than $\frac{1}{\sqrt{2}}$ and they do so every one or two partitions rather than after $d$ partitions.
-See Section~\ref{sec:results} for results on several datasets.
+See Section~\ref{sec:results} for observations of radii-scaling on several datasets.
 
@@
 Let the leaves be indexed with integers $j \in [1, 2^{d + 1} - 1]$ as obtained by a breadth-first traversal of the tree, i.e.
-the root has index $1$ and a cluster with index $i$ has children with indices $2i$ and $2i + 1$.
+the root has index $1$, a cluster with index $i < 2^d$ has children with indices $2i$ and $2i + 1$, and a cluster with index $i \geq 2^d$ is a leaf-cluster with no children.
 Let the center of each child-cluster be encoded in terms of the center of its parent-cluster.
@@
 
-We can finally rewrite Equation~\ref{eq:hierarchical-shannon-entropy} as
+Finally, we rewrite Equation~\ref{eq:hierarchical-shannon-entropy} as
 \begin{align}
@@
 As we continue deeper into the tree, we would start seeing diminishing returns for using subtrees.
-Eventually, either the difference will be so small that it is not worth the computational cost of Partition, or the two cases will equalize, or the first case will become the better choice.
+Eventually, either 
+the difference will be so small so as not to be worth the computational cost of Partition, 
+or the two cases will equalize, 
+or the first case will become the better choice.
 At such a depth in the tree, we can encode the instances in a cluster using the cluster-center rather than the subtree.
-That cluster-center can be encoded in terms of the center of its parent and so on back up the tree to the root.
+The center of that cluster can be encoded in terms of the center of its parent and so on back up the tree to the root.
 

### 2021-05-21 sections/2-info-theory.tex
A comment and a few additions

@@
 We further generalize this for any dataset and distance function described below.
-We also provide relevant algorithms for some parings of datasets and distance functions in Section~\ref{subsec:methods:compression}.
+We also provide relevant algorithms for some pairings of datasets and distance functions in Section~\ref{subsec:methods:compression}.
 
-We take a \textbf{dataset} $X$ with $n$ instances embedded in a $\mathcal{D}$-dimensional Banach Space defined by a distance function $f: X \times X \mapsto \mathbb{R}$.
-We assume that the dataset follows the Manifold Hypothesis~\cite{fefferman2016testing}, i.e. instances in the dataset lie on a $d$-dimensional manifold where $d \ll \mathcal{D}$.
+We take a \textbf{dataset} $X$ with $n$ instances embedded in a $\mathcal{D}$-dimensional Banach space defined by a distance function $f: X \times X \mapsto \mathbb{R}$.
+We assume that the dataset follows the Manifold Hypothesis~\cite{fefferman2016testing}, i.e. instances in the dataset mostly lie on a $d$-dimensional manifold where $d \ll \mathcal{D}$.
 We also make some uniformity assumptions for the dataset;
@@
 
-The \textbf{distance function} must have must be such that $\forall x, y, z \in X$:
+The \textbf{distance function} must be a metric; that is it must have the properties that $\forall x, y, z \in X$:
 \begin{enumerate}[i.]
@@
 where the center of the root cluster is encoded directly rather than in terms of another instance.
+
+% Wait, can't we encode any cluster at depth d with at most d bits?
+For a binary tree, arbitrarily choose 0 for left and 1 for right, and a length-d traversal uniquely identifies a cluster at depth d.
+Not sure if this is a weaker upper bound, but we should mention it.
+For n flat clusters, lg n bits identifies a unique one; for n leaves, lg n bits identifies a leaf, but an internal cluster at depth d only needs lg d bits.
+% Actually I see that you sort of get at this in the "back to shannon entropy" subsection, but maybe it could be explained intuitively as I did above?
 

### 2021-05-21 sections/5-results.tex
typos, reorganization, and more
A comment and a few additions

@@
 
-\subsection{Asymptotic Complexity}
-\label{subsec:results:asymptotic-complexity}
+Specs of Ark, and maybe M1 MacBook Air and Dell XPS.
 
-Analyze:
-\begin{itemize}
-    \item Building the tree.
-    \item $\rho$-nearest neighbors search.
-    \item k-nearest neighbors search.
-    \item compression:
-    \begin{itemize}
-        \item encode (novel contribution)
-        \item serialize
-        \item compress
-        \item decompress
-        \item deserialize
-        \item decode (novel contribution)
-    \end{itemize}
-\end{itemize}
+Ark: a 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM and CentOS 7 Linux with kernel 3.10.0-1127.13.1.el7.x86\_64 \#1 SMP. Rust 1.52, Python 3.6.8
 
-\subsection{Benchmarks}
-\label{subsec:results:benchmarks}
-
-Specs of Ark, and maybe M1 MacBook Air and Dell XPS.
+Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 11.3. Rust 1.52, Python 3.9.4
 
@@
     \end{enumerate}
+    \item Measure time taken for accelerated-search in the compressed space and decompression of the hits with different combinations of serde and codec libraries.
 \end{enumerate}

### 2021-10-27 sections/2-info-theory.tex
comment on hierarchical encodings
fix bug in equation
names

@@
 
-The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $n$ such that $\{x_1, \dots, x_n\} \subseteq X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e.
+The \textbf{Metric Entropy} (Kolmogorov Entropy) of the dataset, denoted by $N_{\epsilon}^{ent}(X)$, is the maximum $k$ such that $\{x_1, \dots, x_k\} \subseteq X$ and $f(x_i, x_j) \geq \epsilon \ \forall \ i \neq j$, i.e.
 the maximum number of $\epsilon$-separated instances in the dataset.
 
-The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $n$ such that $\{x_1, \dots, x_n\} \subset X$ and $\bigcup_{i = 1}^{n} C(c_i, \epsilon) = X$, i.e.
+The \textbf{Internal Covering Number} of the dataset, denoted by $N_{\epsilon}^{int}(X)$, is the minimum $k$ such that $\{x_1, \dots, x_k\} \subset X$ and $\bigcup_{i = 1}^{k} B(c_i, \epsilon) = X$, i.e.
 the minimum number of $\epsilon$-radius clusters that collectively cover $X$.
@@
 \begin{align*}
-    d(X) &= \lim_{\epsilon \rightarrow 0} \frac{\log \big( N_{\epsilon}^{int}(X) \big) }{\log \big( \frac{1}{\epsilon} \big)}
+    d(X) &= \lim_{\epsilon \rightarrow 0} \frac{\log \big( N_{\epsilon}^{ent}(X) \big) }{\log \big( \frac{1}{\epsilon} \big)}
 \end{align*}
@@
 
+% Noah
 % Wait, can't we encode any cluster at depth d with at most d bits?
@@
 % Actually I see that you sort of get at this in the "back to shannon entropy" subsection, but maybe it could be explained intuitively as I did above?
+
+% Najib
+% Such a traversal identifies only which cluster it would be.
+That traversal does not encode (or compress) the cluster-center.
+For discussing compression and Shannon entropy, we want to encode/compress cluster-centers using the hierarchical relations from the tree in such a way that the child-center is recoverable from a parent-center and the encoding.
+% I see that the term `encode' is overloaded here.
 

### 2022-06-13 sections/2-info-theory.tex
wip: new knn algorithm outline

@@
 
-The \textbf{distance function} must be a metric; that is it must have the properties that $\forall x, y, z \in X$:
+The following analysis assumes that the \textbf{distance function} is a metric;
+i.e. $\forall x, y, z \in X$:
 \begin{enumerate}[i.]
     \item $f$ is deterministic.
-    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$.
+    \item $f(x, y) = 0 \Leftrightarrow x = y$ and $f(x, y) > 0 \Leftrightarrow x \neq y$, i.e. identity.
     \item $f(x, y) = f(y, x)$, i.e. symmetry.
@@
 \end{enumerate}
+
+See Section~\ref{sec:discussion} for a discussion on what happens with distance functions for which these properties do not hold.
 

### 2022-06-13 sections/3-methods.tex
A comment
wip: new knn algorithm outline

@@
 
-Modified binary search.
-Lift this from CHESS paper.
+Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.
+
+Modified binary search to identify candidate leaf clusters.
+Exhaustive leaf search over those leaves.
+
+Lift this from CHESS paper. Be sure to cite CHESS here.
 
@@
 
-Repeated $\rho$-NN search with where radius is modified with each iteration as guided by lfd of query ball.
+Given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
 
-Lift from CHESS paper as a starting point.
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
+Let $\delta^0 = f(q, c)$ be the distance from the query to the cluster center.
+Let $\delta^1 = \delta^0 + r$ be the distance from the query to the potentially farthest instance in the Cluster.
+Let $\delta^2 = max(0, \delta^0 - r)$ be the distance from the query to the potentially closest instance in the Cluster.
 
-Complexity is some factor times CHESS-search.
+\subsubsection{Tree Search}
+\label{subsubsec:methods:knn-search:tree-search}
+
+Let $candidates$ be a priority queue sorted by non-decreasing $\delta^1$.
+Initialize $candidates$ with the root cluster.
+Replace clusters in $candidates$ by their children until $candidates$ has at least $k$ clusters.
+
+Take the $k^{th}$ candidate.
+Any cluster with $\delta^2 \leq \delta^2_k$ could contain one of the $k$-nearest neighbors of $q$.
+Keep all such clusters in $candidates$ and exclude all other clusters.
+Expand search to the children of all $candidates$ so far and repeat until all $candidates$ are leaf clusters.
+
+Sort, as a priority queue, $candidates$ by $\delta^2$ and break ties by $\delta^0$.
+Proceed to Leaf Search with these candidates.
+
+\subsubsection{Leaf Search}
+\label{subsubsec:methods:knn-search:leaf-search}
+
+Let $hits$ be a a fixed size priority queue with $k$ isntances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
+Initialize $hits$ with the centers of $candidates$.
+
+Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
+Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
+Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
+Repeat until $candidates$ is empty.
+
+\subsubsection{Complexity}
+\label{subsubsec:methods:knn-search:complexity}
+
+This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
+It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.
 

### 2022-06-16 sections/3-methods.tex
PQ discussion

@@
 
+Add each cluster to a priority queue $m$ times, where $m$ is the cardinality of the cluster, and the priority is the $\delta^1$ distance of that cluster. 
+Then, any cluster whose $\delta^2$ is greater than the $k^{th}$ highest $\delta^1$.
+
 Let $candidates$ be a priority queue sorted by non-decreasing $\delta^1$.
@@
 Take the $k^{th}$ candidate.
-Any cluster with $\delta^2 \leq \delta^2_k$ could contain one of the $k$-nearest neighbors of $q$.
+Any cluster with $\delta^2 \leq \delta^1_k$ could contain one of the $k$-nearest neighbors of $q$.
 Keep all such clusters in $candidates$ and exclude all other clusters.

### 2022-07-07 sections/1-introduction.tex
Some notes, new backronym, some citations

@@
 Write intro \dots
+
+- introduce kNN~\cite{fix1952discriminatory, cover1967nearest} as one of the most pervasive supervised classification methods in use
+- performance difficulties of naive kNN, and introduce existing improvements such as locality-sensitive hashing~\cite{bingham2001random}
+- Introduce Compressed Accelerated K-nearest-neighbor Entropy-scaling Search (CAKES)
+% note new backronym (NMD)
 

### 2022-07-07 sections/5-results.tex
Some notes, new backronym, some citations

@@
 
-Ark: a 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM and CentOS 7 Linux with kernel 3.10.0-1127.13.1.el7.x86\_64 \#1 SMP. Rust 1.52, Python 3.6.8
+Ark: a 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM and CentOS 7 Linux with kernel 5.15.50-1-MANJARO \#1 SMP PREEMPT. Rust 1.62, Python 3.10.5
 
-Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 11.3. Rust 1.52, Python 3.9.4
+Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 12.1. Rust 1.62, Python 3.9.4
 

### 2023-08-08 sections/3-methods.tex
added writeup for expanding threshold variant of knn

--- 2022-06-13 sections/3-methods.tex
@@
 Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
-Let $\delta^0 = f(q, c)$ be the distance from the query to the cluster center.
-Let $\delta^1 = \delta^0 + r$ be the distance from the query to the potentially farthest instance in the Cluster.
-Let $\delta^2 = max(0, \delta^0 - r)$ be the distance from the query to the potentially closest instance in the Cluster.
+Let $\delta = f(q, c)$ be the distance from the query to the cluster center.
+Let $\delta_{max} = \delta + r$ be the distance from the query to the potentially farthest instance in the Cluster.
+Let $\delta_{min} = max(0, \delta - r)$ be the distance from the query to the potentially closest instance in the Cluster.
 
-\subsubsection{Tree Search}
-\label{subsubsec:methods:knn-search:tree-search}
+\subsubsection{Expanding Threshold}
+\label{subsubsec:methods:knn-search:expanding-threshold}
 
-Let $candidates$ be a priority queue sorted by non-decreasing $\delta^1$.
+Create a priority queue, $candidates$, wherein the top priority candidate is the one with the smallest $\delta_{min}$.
 Initialize $candidates$ with the root cluster.
-Replace clusters in $candidates$ by their children until $candidates$ has at least $k$ clusters.
+Create a second priority queue, $hits$, wherein the top priority hit is the 
+one with the largest $\delta$. This queue starts empty.
 
-Take the $k^{th}$ candidate.
-Any cluster with $\delta^2 \leq \delta^2_k$ could contain one of the $k$-nearest neighbors of $q$.
-Keep all such clusters in $candidates$ and exclude all other clusters.
-Expand search to the children of all $candidates$ so far and repeat until all $candidates$ are leaf clusters.
+While $hits$ has fewer than $k$ clusters or the $\delta$ of the top priority hit is greater 
+than the $\delta_{min}$ of the top priority candidate, do the following:
+\begin{enumerate}
+\item While the top priority candidate isn't a leaf, replace the top priority 
+candidate with its children.
+\item Remove the top priority candidate from $candidates$ and add all of its 
+points to $hits$. 
+\item Remove points from $hits$ until $hits$ has $k$ points. 
+\end{enumerate}
 
-Sort, as a priority queue, $candidates$ by $\delta^2$ and break ties by $\delta^0$.
-Proceed to Leaf Search with these candidates.
+Extract hits.
+
+\begin{algorithm} % enter the algorithm environment
+\caption{Expanding Threshold} % give the algorithm a caption
+\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
+\begin{algorithmic}[1] % enter the algorithmic environment
+    \REQUIRE $tree$, $query$, $k$
+    \STATE $candidates \leftarrow$ priority queue
+    \STATE $hits \leftarrow$ priority queue
+    \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
+        \WHILE{$!candidates.peek().isLeaf$}
+            \STATE $[l, r] \leftarrow candidates.pop().children()$
+            \STATE $candidates.push([l, r])$
+        \ENDWHILE
+        \STATE $leaf \leftarrow candidates.pop()$
+        \STATE $hits.push(leaf)$
+        \WHILE{$|hits| > k$}
+            \STATE $hits.pop()$
+        \ENDWHILE
+    \ENDWHILE
+    \STATE Return $hits$
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Sieve V1}
+\label{subsubsec:methods:knn-search:sieve-v1}
+Add explanation of the sieve v1 algorithm.
+
+\subsubsection{Sieve V2}
+\label{subsubsec:methods:knn-search:sieve-v2}
+Add explanation of the sieve v2 algorithm.
+
+\subsubsection{Repeated Rnn}
+\label{subsubsec:methods:knn-search:repeated-rnn}
+Add explanation of the repeated rnn algorithm.
 

### 2023-08-12 sections/3-methods.tex
updates to sieve v1 alg

@@
 
-Given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
+The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
 
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
-Let $\delta = f(q, c)$ be the distance from the query to the cluster center.
-Let $\delta_{max} = \delta + r$ be the distance from the query to the potentially farthest instance in the Cluster.
-Let $\delta_{min} = max(0, \delta - r)$ be the distance from the query to the potentially closest instance in the Cluster.
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
+\begin{itemize}
+    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
+    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
+    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
+\end{itemize}
 
@@
 
-Create a priority queue, $candidates$, wherein the top priority candidate is the one with the smallest $\delta_{min}$.
-Initialize $candidates$ with the root cluster.
-Create a second priority queue, $hits$, wherein the top priority hit is the 
-one with the largest $\delta$. This queue starts empty.
+Let $candidates$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $candidates$ with the root cluster.
+Let $hits$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
 
@@
 
-Extract hits.
+Extract $hits$.
 
@@
     \STATE $candidates \leftarrow$ priority queue
+    \STATE $candidates.push(tree.root)$
     \STATE $hits \leftarrow$ priority queue
@@
 \label{subsubsec:methods:knn-search:sieve-v1}
-Add explanation of the sieve v1 algorithm.
+Let $candidates$ be a vector of $clusters$s sorted by non-decreasing $\delta_{max}$. Initialize $candidates$ with  
+the root cluster. 
+Let $hits$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
+
+In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{minmax}$ less than $\tau$ 
+can contain one of the $k$ nearest neighbors.
+
+Repeat the following: 
+\begin{enumerate}
+\item Find the cluster $C_{\tau}$ in $candidates$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
+\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
+\item Remove from $hits$ any points which are a distance greater than $\tau$ from the query. 
+\item If no cluster in $candidates$ has a $\delta_{max}$ less than or equal to $\tau$, break.
+\item Remove from $candidates$ any cluster whose $\delta_{min}$ is greater than $\tau$.
+\item Transfer from $candidates$ to $hits$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
+\item Replace all non-leaf clusters in $candidates$ with their children. 
+\end{enumerate}
+
+At this stage, if any clusters remain in $candidates$, we add all their points to $hits$. 
+Extract $hits$.
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{Sieve V1} % give the algorithm a caption
+    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[2] % enter the algorithmic environment
+        \REQUIRE $tree$, $query$, $k$
+        \STATE $candidates \leftarrow$ priority queue
+        \STATE $candidates.push(tree.root)$
+        \STATE $hits \leftarrow$ priority queue
+        \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
+            \WHILE{$!candidates.peek().isLeaf$}
+                \STATE $[l, r] \leftarrow candidates.pop().children()$
+                \STATE $candidates.push([l, r])$
+            \ENDWHILE
+            \STATE $leaf \leftarrow candidates.pop()$
+            \STATE $hits.push(leaf)$
+            \WHILE{$|hits| > k$}
+                \STATE $hits.pop()$
+            \ENDWHILE
+        \ENDWHILE
+        \STATE Return $hits$
+    \end{algorithmic}
+    \end{algorithm}
 
@@
 
-Let $hits$ be a a fixed size priority queue with $k$ isntances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
+Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
 Initialize $hits$ with the centers of $candidates$.

### 2023-08-18 sections/1-introduction.tex
struggles with intro and methods

--- 2021-03-09 sections/1-introduction.tex
@@
 
-Write intro \dots
+Often dubbed "the Big Data explosion", rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
+For example, the Sloan Digital Sky Survey (SDSS) has collected data on over 500 million celestial objects~\cite{alam2015eleventh}. Or, as another 
+example, the National Center for Biotechnology Information (NCBI) has collected over 2.25 million ribosomal DNA sequences~\cite{silva2014silva}. Finally, 
+as a third example, the Large Hadron Collider (LHC) has collected over 100 petabytes of data~\cite{evans2008lhc}. Or, as a fourth example, the 
+National Oceanic and Atmospheric Administration (NOAA) has collected over 20 petabytes of data~\cite{noaa2019noaa}. 
+In many cases, the influx of data outpaces the rate of improvement in computing capacity predicted by Moore's Law~\cite{moore1965cramming}, indicating 
+that computer hardware will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
+requires better algorithms.
+
+The ability to perform similarity search on these datasets enables a variety of applications, including recommendation and classification systems. 
+Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
+understanding and verifying the validity of model outputs. 
+As the sizes and dimensionalities of datasets have grown, efficient and accurate similarity search has become excessively challenging; 
+state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.
+In particular, k-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.
+
+Naive implementations of kNN, whose asymptotic complexity is a function of the dataset's cardinality, 
+prove prohibitively slow for large datasets. While algorithms for kNN on large datasets exist, most of them are approximate and fail to 
+exploit the inherent structure and redundancy often present in big datasets. While approximate search is sufficient for many applications, the 
+need for efficient, exact search remains.
+
+Recent approaches to tackle the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
+clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy}. 
+Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties of the dataset,
+rather than its cardinality. In 2019, we introduced CHESS (CLustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
+neighbors search to a hierarchical clustering approach. In this paper, we introduce CLAM (Clustered Learning of Approximate Manifolds), and CAKES (CLAM-Accelerated K-nearest-neighbor 
+Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm (figure out and explain improvements upon CHESS present in CLAM if there are any algorithmic ones)
+that can facilitate a variety of algorithms for big data. Using the tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
+k-nearest neighbors search in addition to $\rho$-nearest neighbors search.
+
+At some point, need to address the fact that exact search is important for small k and approxiate 
+is maybe more ideal for large k. Try to find citation for this.
 
@@
     \item Accelerated $\rho$-nearest and k-nearest neighbors search.
-    \item Entropy-scaling (?? is it though?) compression using the hierarchical clustering.
-    \item Proof for how cluster-radii, and thus compression deltas, fall as we descend along the tree.
-    \item Information theoretic bounds on compression performance. We compress along the dimensionality of data rather than the cardinality of data.
-    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, wasserstein, hamming, levenshtein, needleman-wunch, and any user-defined function.
+    \item Proof for how cluster-radii fall as we descend along the tree.
+    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, hamming, levenshtein, needleman-wunch, and any user-defined function.
 \end{itemize}
@@
 \begin{itemize}
-    \item Search time vs FALCONN and other in ann-benchmarks suite.
-    \item Compression performance of CAKES+gzip vs just gzip.
+    \item Search time vs FAISS and FALCONN and other in ann-benchmarks suite.
 \end{itemize}

### 2023-08-18 sections/3-methods.tex
struggles with intro and methods

@@
 
-\subsection{Partition}
-\label{subsec:methods:partition}
+\subsection{Dataset and Distance Function}
+\label{subsec:methods:dataset-and-distance-function}
 
-Partition algorithm for building the tree.
-Lift this from the CHAODA paper.
+We start with a dataset $\textbf{X} = \{x_1 \dots x_n\}$ with $n$ points and a distance function $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+$.
+Given two points from the dataset, the distance function deterministically returns a non-negative real number.
+We require that the distance function be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and
+that the distance between two points $x$ and $y$ be zero if and only if $x = y$. When, in addition to these constraints, we also require that the distance function obey 
+the triangle inequality (i.e., when the distance function is a metric), we can guarantee that our search algorithms have perfect recall.
 
-Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+CLAM assumes the ``manifold hypothesis''~\cite{fefferman2016testing}, which states that datasets collected from constrained generating phenomena that are embedded in a high-dimensional space typically only occupy a low-dimensional manifold in that space.
+In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
+While we often Euclidean notions, such as voids and volumes, to talk about geometric and topological
+properties of clusters and of the manifold, CLAM does not rely on such notions; they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.
+
+\subsection{Clustering}
+\label{subsec:methods:clustering}
+
+We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
+similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of 
+poles. For a cluster with $m$ points, we begin by computing a 
+random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
+cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
+The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
+from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
+left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
+points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
+those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
+repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
+is met.
+
+\begin{algorithm} % enter the algorithm environment
+\caption{Partition} % give the algorithm a caption
+\label{alg:partition} % and a label for \ref{} commands later in the document
+\begin{algorithmic}[1] % enter the algorithmic environment
+    \REQUIRE $cluster$
+    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
+    \STATE $seeds \leftarrow m$ random points from $cluster.points$
+    \STATE $c \leftarrow$ geometric median of $seeds$
+    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
+    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
+    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
+    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
+    \IF{$|left| > 1$}
+        \STATE Partition($left$)
+    \ENDIF
+    \IF{$|right| > 1$}
+        \STATE Partition($right$)
+    \ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection {Proof of Eventually Decreasing Cluster Radii}
+\subsection{Dataset and Distance Function}
+\label{subsec:methods:dataset-and-distance-function}
+
+We start with a dataset $\textbf{X} = \{x_1 \dots x_n\}$ with $n$ points and a distance function $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+$.
+Given two points from the dataset, the distance function deterministically returns a non-negative real number.
+We require that the distance function be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and
+that the distance between two points $x$ and $y$ be zero if and only if $x = y$. When, in addition to these constraints, we also require that the distance function obey 
+the triangle inequality (i.e., when the distance function is a metric), we can guarantee that our search algorithms have perfect recall.
+
+CLAM assumes the ``manifold hypothesis''~\cite{fefferman2016testing}, which states that datasets collected from constrained generating phenomena that are embedded in a high-dimensional space typically only occupy a low-dimensional manifold in that space.
+In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
+While we often Euclidean notions, such as voids and volumes, to talk about geometric and topological
+properties of clusters and of the manifold, CLAM does not rely on such notions; they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.
+
+\subsection{Clustering}
+\label{subsec:methods:clustering}
+
+We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
+similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of 
+poles. For a cluster with $m$ points, we begin by computing a 
+random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
+cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
+The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
+from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
+left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
+points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
+those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
+repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
+is met.
+
+\begin{algorithm} % enter the algorithm environment
+\caption{Partition} % give the algorithm a caption
+\label{alg:partition} % and a label for \ref{} commands later in the document
+\begin{algorithmic}[1] % enter the algorithmic environment
+    \REQUIRE $cluster$
+    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
+    \STATE $seeds \leftarrow m$ random points from $cluster.points$
+    \STATE $c \leftarrow$ geometric median of $seeds$
+    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
+    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
+    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
+    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
+    \IF{$|left| > 1$}
+        \STATE Partition($left$)
+    \ENDIF
+    \IF{$|right| > 1$}
+        \STATE Partition($right$)
+    \ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection {Proof of Eventually Decreasing Cluster Radii}
+\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}
+
+\subsubsection {Complexity}
+\label{subsubsec:methods:clustering:complexity}
+
+Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
 Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
@@
 \label{subsubsec:methods:knn-search:expanding-threshold}
+In this variant of KNN, we repeatedly
 
-Let $candidates$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $candidates$ with the root cluster.
-Let $hits$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
+Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
+Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
 
-While $hits$ has fewer than $k$ clusters or the $\delta$ of the top priority hit is greater 
-than the $\delta_{min}$ of the top priority candidate, do the following:
+While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
+than $\delta_{min}$ of the top priority element in $Q$, do the following:
 \begin{enumerate}
-\item While the top priority candidate isn't a leaf, replace the top priority 
-candidate with its children.
-\item Remove the top priority candidate from $candidates$ and add all of its 
-points to $hits$. 
-\item Remove points from $hits$ until $hits$ has $k$ points. 
+\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
+\item Remove $q$ from $Q$ and add all of its points to $P$. 
+\item Remove points from $P$ until $P$ has $k$ points. 
 \end{enumerate}
 
-Extract $hits$.
+Return $P$.
 
@@
     \REQUIRE $tree$, $query$, $k$
-    \STATE $candidates \leftarrow$ priority queue
-    \STATE $candidates.push(tree.root)$
-    \STATE $hits \leftarrow$ priority queue
-    \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
-        \WHILE{$!candidates.peek().isLeaf$}
-            \STATE $[l, r] \leftarrow candidates.pop().children()$
-            \STATE $candidates.push([l, r])$
+    \STATE $Q \leftarrow$ priority queue
+    \STATE $Q.push(tree.root)$
+    \STATE $P \leftarrow$ priority queue
+    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
+        \WHILE{$!Q.min.isLeaf$}
+            \STATE $[l, r] \leftarrow Q.extractMin().children()$
+            \STATE $Q.push([l, r])$
         \ENDWHILE
-        \STATE $leaf \leftarrow candidates.pop()$
-        \STATE $hits.push(leaf)$
-        \WHILE{$|hits| > k$}
-            \STATE $hits.pop()$
+        \STATE $leaf \leftarrow Q.extractMin()$
+        \STATE $P.push(leaf)$
+        \WHILE{$|P| > k$}
+            \STATE $P.extractMax()$
         \ENDWHILE
     \ENDWHILE
-    \STATE Return $hits$
+    \STATE Return $P$
 \end{algorithmic}
@@
 \label{subsubsec:methods:knn-search:sieve-v1}
-Let $candidates$ be a vector of $clusters$s sorted by non-decreasing $\delta_{max}$. Initialize $candidates$ with  
+
+In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
+can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
+
+Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
 the root cluster. 
-Let $hits$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
-
-In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{minmax}$ less than $\tau$ 
-can contain one of the $k$ nearest neighbors.
+Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
 
@@
 \begin{enumerate}
-\item Find the cluster $C_{\tau}$ in $candidates$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
+\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
 \item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
-\item Remove from $hits$ any points which are a distance greater than $\tau$ from the query. 
-\item If no cluster in $candidates$ has a $\delta_{max}$ less than or equal to $\tau$, break.
-\item Remove from $candidates$ any cluster whose $\delta_{min}$ is greater than $\tau$.
-\item Transfer from $candidates$ to $hits$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
-\item Replace all non-leaf clusters in $candidates$ with their children. 
+\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
+\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
+\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
+\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
+\item Replace all non-leaf clusters in $Q$ with their children. 
 \end{enumerate}
 
-At this stage, if any clusters remain in $candidates$, we add all their points to $hits$. 
-Extract $hits$.
+At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
+Extract $P$.
 
@@
         \REQUIRE $tree$, $query$, $k$
-        \STATE $candidates \leftarrow$ priority queue
-        \STATE $candidates.push(tree.root)$
-        \STATE $hits \leftarrow$ priority queue
-        \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
-            \WHILE{$!candidates.peek().isLeaf$}
-                \STATE $[l, r] \leftarrow candidates.pop().children()$
-                \STATE $candidates.push([l, r])$
+        \STATE $Q \leftarrow$ priority queue
+        \STATE $Q.push(tree.root)$
+        \STATE $P \leftarrow$ priority queue
+        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
+            \WHILE{$!Q.extractMax().isLeaf$}
+                \STATE $[l, r] \leftarrow Q.extractMax().children()$
+                \STATE $Q.push([l, r])$
             \ENDWHILE
-            \STATE $leaf \leftarrow candidates.pop()$
-            \STATE $hits.push(leaf)$
-            \WHILE{$|hits| > k$}
-                \STATE $hits.pop()$
+            \STATE $leaf \leftarrow Q.extractMax()$
+            \STATE $P.push(leaf)$
+            \WHILE{$|P| > k$}
+                \STATE $P.extractMax()$
             \ENDWHILE
         \ENDWHILE
-        \STATE Return $hits$
+        \STATE Return $P$
     \end{algorithmic}
@@
 
-\subsection{Compression}
-\label{subsec:methods:compression}
+\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}
 
-Encode and Decode algorithms.
-These could be, and probably will be, different for each $(X, f)$ pairing.
+\subsubsection {Complexity}
 
-Asymptotic complexity analysis for:
-\begin{enumerate}[i.]
-    \item encode (novel contribution)
-    \item serialize
-    \item compress
-    \item decompress
-    \item deserialize
-    \item decode (novel contribution)
+\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
+\label{subsec:methods:rnn-search}
+
+Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.
+
+Modified binary search to identify candidate leaf clusters.
+Exhaustive leaf search over those leaves.
+
+Lift this from CHESS paper. Be sure to cite CHESS here.
+
+Complexity is the same as in CHESS.
+
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
+\label{subsec:methods:knn-search}
+
+The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
+
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
+\begin{itemize}
+    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
+    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
+    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
+\end{itemize}
+
+\subsubsection{Expanding Threshold}
+\label{subsubsec:methods:knn-search:expanding-threshold}
+In this variant of KNN, we repeatedly
+
+Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
+Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
+
+While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
+than $\delta_{min}$ of the top priority element in $Q$, do the following:
+\begin{enumerate}
+\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
+\item Remove $q$ from $Q$ and add all of its points to $P$. 
+\item Remove points from $P$ until $P$ has $k$ points. 
 \end{enumerate}
 
-% discrete vs continuous, and small alphabet vs large alphabet in discrete
+Return $P$.
+
+\begin{algorithm} % enter the algorithm environment
+\caption{Expanding Threshold} % give the algorithm a caption
+\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
+\begin{algorithmic}[1] % enter the algorithmic environment
+    \REQUIRE $tree$, $query$, $k$
+    \STATE $Q \leftarrow$ priority queue
+    \STATE $Q.push(tree.root)$
+    \STATE $P \leftarrow$ priority queue
+    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
+        \WHILE{$!Q.min.isLeaf$}
+            \STATE $[l, r] \leftarrow Q.extractMin().children()$
+            \STATE $Q.push([l, r])$
+        \ENDWHILE
+        \STATE $leaf \leftarrow Q.extractMin()$
+        \STATE $P.push(leaf)$
+        \WHILE{$|P| > k$}
+            \STATE $P.extractMax()$
+        \ENDWHILE
+    \ENDWHILE
+    \STATE Return $P$
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Sieve V1}
+\label{subsubsec:methods:knn-search:sieve-v1}
+
+In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
+can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
+
+Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
+the root cluster. 
+Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
+
+Repeat the following: 
+\begin{enumerate}
+\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
+\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
+\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
+\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
+\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
+\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
+\item Replace all non-leaf clusters in $Q$ with their children. 
+\end{enumerate}
+
+At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
+Extract $P$.
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{Sieve V1} % give the algorithm a caption
+    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[2] % enter the algorithmic environment
+        \REQUIRE $tree$, $query$, $k$
+        \STATE $Q \leftarrow$ priority queue
+        \STATE $Q.push(tree.root)$
+        \STATE $P \leftarrow$ priority queue
+        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
+            \WHILE{$!Q.extractMax().isLeaf$}
+                \STATE $[l, r] \leftarrow Q.extractMax().children()$
+                \STATE $Q.push([l, r])$
+            \ENDWHILE
+            \STATE $leaf \leftarrow Q.extractMax()$
+            \STATE $P.push(leaf)$
+            \WHILE{$|P| > k$}
+                \STATE $P.extractMax()$
+            \ENDWHILE
+        \ENDWHILE
+        \STATE Return $P$
+    \end{algorithmic}
+    \end{algorithm}
+
+\subsubsection{Sieve V2}
+\label{subsubsec:methods:knn-search:sieve-v2}
+Add explanation of the sieve v2 algorithm.
+
+\subsubsection{Repeated Rnn}
+\label{subsubsec:methods:knn-search:repeated-rnn}
+Add explanation of the repeated rnn algorithm.
+
+\subsubsection{Leaf Search}
+\label{subsubsec:methods:knn-search:leaf-search}
+
+Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
+Initialize $hits$ with the centers of $candidates$.
+
+Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
+Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
+Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
+Repeat until $candidates$ is empty.
+
+\subsubsection{Complexity}
+\label{subsubsec:methods:knn-search:complexity}
+
+Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
+
+This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
+It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.

### 2023-08-21 sections/1-introduction.tex
lots of changes

@@
 \label{sec:introduction}
+Often dubbed "the Big Data explosion", rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
+The Sloan Digital Sky Survey (SDSS), for example, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
+Add sentence about MaNGA here: 10,000 galaxies with 2,700 deg$^2$, resolution $R=2,000$ for $\lambda \in [360, 1000) nm$.
 
-Often dubbed "the Big Data explosion", rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
-For example, the Sloan Digital Sky Survey (SDSS) has collected data on over 500 million celestial objects~\cite{alam2015eleventh}. Or, as another 
-example, the National Center for Biotechnology Information (NCBI) has collected over 2.25 million ribosomal DNA sequences~\cite{silva2014silva}. Finally, 
-as a third example, the Large Hadron Collider (LHC) has collected over 100 petabytes of data~\cite{evans2008lhc}. Or, as a fourth example, the 
-National Oceanic and Atmospheric Administration (NOAA) has collected over 20 petabytes of data~\cite{noaa2019noaa}. 
-In many cases, the influx of data outpaces the rate of improvement in computing capacity predicted by Moore's Law~\cite{moore1965cramming}, indicating 
+Biology has also seen an influx of large, high-dimensional datasets. Silva 18S contains ribosomal DNA sequences of approximately 2.25 
+million genomes with aligned length of 50,000 letters. (Do we need to look at NR99??)
+
+In many cases, the increase in data outpaces the rate of improvement in computing capacity predicted by Moore's Law~\cite{moore1965cramming}, indicating 
 that computer hardware will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
@@
 prove prohibitively slow for large datasets. While algorithms for kNN on large datasets exist, most of them are approximate and fail to 
-exploit the inherent structure and redundancy often present in big datasets. While approximate search is sufficient for many applications, the 
+exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
 need for efficient, exact search remains.
 
-Recent approaches to tackle the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
-clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy}. 
+Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
+clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
 Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
@@
 
-At some point, need to address the fact that exact search is important for small k and approxiate 
+At some point, need to address the fact that exact search is important for small k and approximate 
 is maybe more ideal for large k. Try to find citation for this.
-
-Growth in size and types of datasets:
-\begin{itemize}
-    \item SDSS APOGEE2 data. Near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions. $R=22,500 = \frac{\lambda}{\Delta \lambda}$ for $\lambda \in [1.51, 1.70) \mu m$
-    \item SDSS MaNGA hexagon shaped images. 10,000 galaxies with 2,700 deg$^2$, resolution $R=2,000$ for $\lambda \in [360, 1000) nm$
-    \item Silva 18S ribosomal DNA sequences of approximately 2.25 million genomes with aligned length of 50,000 letters. TODO: Look into other variants in the dataset particularly NR99.
-    \item Datasets from ANN-Benchmarks suite
-\end{itemize}
-
-Recent approaches in dealing with exponential growth of data:
-\begin{itemize}
-    \item LSH
-    \item clever indexing e.g. FM Index.
-    \item Entropy-scaling search~\cite{yu2015entropy} and CHESS~\cite{ishaq2019clustered}.
-\end{itemize}
 

### 2023-08-21 sections/2-methods.tex
lots of changes

@@
 
-\subsection{Partition}
-\label{subsec:methods:partition}
+In this manuscript, we are primarily concerned with $k$-nearest neighbors search in a finite-dimensional 
+vector space. Some of the algorithms we present use a modified version of the $\rho$-nearest neighbors search 
+algorithm described in CHESS~\cite{ishaq2019clustered} to conduct $k$-NN search. (TODO: make second section less 
+vague once we determine exactly what we're doing with sharding and indexing.)
 
-Partition algorithm for building the tree.
-Lift this from the CHAODA paper.
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of 
+an organism, the vector representation of a single image, or any entity on which we can define a \emph{distance function}).
 
-Exact partition costing $\mathcal{O}(n^2)$ vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+We define a \emph{distance function} $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
+given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
+be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
+When, in addition to these constraints, that the distance function obeys
+the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
+Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
+Euclidean (L2) and Cosine distances. With biological or string data on the other hand, Hamming and Levenshtein distances are more appropriate.
 
-% latex does not seem to like math symbols in this heading
+With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
+Given a query $q$, along with a distance function $f$ defined on a dataset $ \textbf{X}$, $k$-nearest neighbors search aims to: 
+for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$; that is, find the $k$ points $x \in \textbf{X}$ with the smallest $f(q, x)$.
+We also have the $\rho$-nearest neighbors search problem: for a given radius $\rho$, find all points in $\textbf{X}$ that are a distance of $\rho$ 
+or less from $q$; that is, all points $x \in \textbf{X}$ such that $f(q, x) \leq \rho$.
+
+Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some region of the dataset, 
+which we define as: 
+$$\text{log}_2 (\frac{|B_D(q, r_1)|}{|B_D(q, r_2)|})$$
+where $B_D(q, r)$ is the set of points contained in a ball of radius $r$ 
+centered on a point $q$ in the dataset $D$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
+We stress that this concept of local fractal dimension differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
+consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
+However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
+manifold. The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
+The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.
+
+We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
+The \emph{center} is the geometric median of the set of points in the \emph{cluster}, and so it is a real data point. The \emph{radius} is the
+maximum distance from a point in the cluster to its \emph{center}. We estimate \emph{local fractal dimension} at the cluster radius and half
+the cluster radius. Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in
+a binary tree has two child nodes. We define the \emph{metric entropy} of a data set under a hierarchical clustering scheme as a refinement of [7], where
+metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a binary, divisive clustering 
+approach, but with early stopping criteria; clusters with fewer than a threshold number of points are not further split. 
+Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of
+leaf clusters.
+
+\subsection{Clustering}
+\label{subsec:methods:clustering}
+
+We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
+similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
+improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
+such as local fractal dimension (see Section ADD SECTION: $k$-NN by Repeated $\rho$-NN).
+
+CLAM assumes the manifold hypothesis. 
+In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy 
+a $d$-dimensional manifold, where $d \ll D$. 
+While we sometimes use Euclidean notions, such as voids and volumes, to talk about geometric and topological 
+properties of clusters and of the manifold, CLAM does not rely on such notions; 
+they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.
+
+For a cluster with $m$ points, we begin by computing a 
+random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
+cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
+The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
+from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
+left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
+points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
+those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
+repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
+is met.
+
+\begin{algorithm} % enter the algorithm environment
+\caption{Partition} % give the algorithm a caption
+\label{alg:partition} % and a label for \ref{} commands later in the document
+\begin{algorithmic}[1] % enter the algorithmic environment
+    \REQUIRE $cluster$
+    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
+    \STATE $seeds \leftarrow m$ random points from $cluster.points$
+    \STATE $c \leftarrow$ geometric median of $seeds$
+    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
+    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
+    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
+    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
+    \IF{$|left| > 1$}
+        \STATE Partition($left$)
+    \ENDIF
+    \IF{$|right| > 1$}
+        \STATE Partition($right$)
+    \ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection {Guaranteed Decrease in Cluster Radii}
+\label{subsubsec:methods:guaranteed-decrease-in-cluster-radii}
+
+We show that cluster radius is guaranteed to decrease after at most $d$ partitions.
+
+Assume that we have a dataset which follows a $d$-dimensional distribution embedded in $\mathcal{D}$-dimensional space.
+We can describe this distribution choosing some set of $d$ mutually orthogonal axes.
+Let $m$ denote the maximum distance between any pair of points along any axis from any choice of mutually orthogonal axes for our $d$-dimensional distribution.
+In the worst case scenario, the distribution of our data is a $d$-sphere and $m$ is the maximum distance between two points along every axis for every choice of axes.
+In this scenario, the cluster radius is $r = \frac{m}{2}$.
+
+Given a cluster $\mathcal{C}$, our partition algorithm will choose the left and right poles, 
+and the axis defined by those poles will be the axis for partitioning $\mathcal{C}$'s points into two child clusters.
+For points in each child cluster, the maximum distance between any pair of points along the \emph{axis of partitioning} will have been reduced to at most $\frac{m}{2}$.
+Though this bound holds for points on the axis of partitioning, it does not serve as a bound for the maximum distance between \emph{any} pair of points in $\mathcal{C}$.
+To see this, consider the two dimensional scenario where the points in $\mathcal{C}$ are distributed in a circle of radius $r$. Clearly, the maximum possible distance between 
+any two points in $\mathcal{C}$ is $2r$, and occurs when the points fall on opposite ends of one of its diameters. Now consider a horizontal diameter of the circle, 
+and let $l$ denote the leftmost point on this diameter. Next, consider the orthogonal vertical diameter, and let $x$ denote the uppermost point on this diameter. When we 
+partition along the horizontal diameter, $l$ and $x$ will be assigned to the left child (as mentioned in ADD SECTION, we assign equidistant points to the left child). 
+This reduces the maximum possible distance between any two points on the axis of partition -- the horizontal diameter -- to $r$. However, by considering the 
+right triangle formed by $l$, $x$, and the center of the circle, we can see that the distance between 
+$l$ and $x$ is $\sqrt{2}{r}$.
+
+If $d > 1$ then the child clusters will have a maximal chord $m_{child}$ such that $m_{child} \leq m$.
+By partitioning the cluster in this way, we have taken one axis and reduced the maximum distance along that axis from $m$ to $\frac{m}{2}$.
+However, as per the distribution of data as discussed earlier, there may be other axes along which the maximum pairwise distance is still $m$.
+We, thus, recursively partition each child cluster.
+
+With each recursive application of Partition, we consume one axis along which the maximum distance was $m$ and reduce the maximum distance along that axis to $\frac{m}{2}$ in the child clusters.
+After $d$ recursive applications of partition, we will have exhausted the $d$ axes with the maximum distance of $m$.
+After these $d$ partitions, the points in each child cluster will have a maximum pairwise distance of $\frac{m}{2}$ along any axis.
+Thus the radii of those child clusters will be at most $\frac{m}{4}$.
+
+\subsubsection {Complexity}
+\label{subsubsec:methods:clustering:complexity}
+
+Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
+
+\subsection {Index-Building}
+\begin{itemize}
+    \item sharded search 
+    \item index building 
+    \item optimization of number of shards 
+\end{itemize}
+
 \subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
@@
 
-Modified binary search.
-Lift this from CHESS paper.
+Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.
 
-\subsection{k-Nearest Neighbors Search}
+We conduct ranged-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
+with some implementation improvements including identification of candidate clusters which are leaves and 
+conducting exhaustive search over those leaves.
+
+The asymptotic complexity of $\rho$-nearest neighbors is the same as in CHESS~\cite{ishaq2019clustered}.
+
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
 \label{subsec:methods:knn-search}
 
-Repeated $\rho$-NN search with where radius is modified with each iteration as guided by lfd of query ball.
+The $k$-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
+We present five variants of $k$-nearest neighbors search: Greedy Search, K-NN by Repeated $\rho$-NN, Sieve Search I, Sieve Search II, and 
+Linear Search. We use a preliminary auto-tuning function to predict the variant which will perform 
+best for a given query, dataset, and value of $k$. We then proceed with search using that variant.
 
-Lift from CHESS paper as a starting point.
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-NN variants make use of the following cluster 
+properties:
+\begin{itemize}
+    \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
+    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in $C$.
+    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in $C$.
+\end{itemize}
 
-\subsection{Compression}
-\label{subsec:methods:compression}
+We sometimes also use the notation $\delta = f(x, q)$ with points instead of clusters to refer to the distance from a point $x$ to the query $q$.
 
-Encode and Decode algorithms.
-These will be specific for each dataset and/or distance function.
+\subsubsection{Greedy Search}
+\label{subsubsec:methods:knn-search:greedy-search}
+Greedy Search repeatedly partitions the cluster with the lowest $\delta_{min}$ into its children 
+until it is a leaf, removing it and adding its points to a fixed size priority queue of points. Search terminates 
+when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining
+cluster.
 
-% discrete vs continuous, and small alphabet vs large alphabet in discrete
+Formally, let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
+Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
+
+While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
+than $\delta_{min}$ of the top priority element in $Q$, do the following:
+\begin{enumerate}
+\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
+\item Remove $q$ from $Q$ and add all of its points to $P$. 
+\item Remove points from $P$ until $P$ has $k$ points. 
+\end{enumerate}
+
+Return $P$.
+
+\begin{algorithm} 
+\caption{Greedy Search} 
+\label{alg:greedy_search} 
+\begin{algorithmic}[1]
+    \REQUIRE $tree$, $query$, $k$
+    \STATE $Q \leftarrow$ priority queue
+    \STATE $Q.push(tree.root)$
+    \STATE $P \leftarrow$ priority queue
+    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
+        \WHILE{$!Q.min.isLeaf$}
+            \STATE $[l, r] \leftarrow Q.extractMin().children()$
+            \STATE $Q.push([l, r])$
+        \ENDWHILE
+        \STATE $leaf \leftarrow Q.extractMin()$
+        \STATE $P.push(leaf)$
+        \WHILE{$|P| > k$}
+            \STATE $P.extractMax()$
+        \ENDWHILE
+    \ENDWHILE
+    \STATE Return $P$
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{$k$-NN by Repeated $\rho$-NN}
+\label{subsubsec:methods:knn-search:repeated-rnn}
+
+Hence its name, $k$-NN by Repeated $\rho$-NN repeatedly performs 
+$\rho$-nearest neighbors search, increasing the radius until $k$ neighbors
+have been found.
+
+Search starts with a radius $m$ equal to the radius of the cluster tree divided by
+the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
+If no points are within a distance $m$ of the query, we increase the radius by a factor of 
+2 and perform $rho$-NN search again, repeating until at least one point is found.
+
+Once we have found at least one point, we continue to perform $\rho$-NN search, but instead of 
+increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
+by the local fractal dimension of the clusters containing the points found so far. In particular, 
+we increase the radius by a factor of $\text{max}(2, \frac{k}{|P|^{\frac{1}{\mu}}}$), where $|P|$ is the number of 
+points found so far and $\mu$ is the mean local fractal dimension of the clusters containing the points found so far.
+Since the local fractal dimension is memoized at clustering time, this poses no additional cost at search time.
+Once $|P| >= k$, we sort the points in $P$ by $\delta$ and return the $k$ closest points.
+
+TODO: 
+- add intutiion for factor 
+- introduce P variable sooner
+
+\subsubsection{Sieve Search I}
+\label{subsubsec:methods:knn-search:sieve}
+
+\subsubsection{Sieve Search II}
+
+In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
+can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
+
+Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
+the root cluster. 
+Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
+
+Repeat the following: 
+\begin{enumerate}
+\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
+\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
+\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
+\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
+\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
+\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
+\item Replace all non-leaf clusters in $Q$ with their children. 
+\end{enumerate}
+
+At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
+Extract $P$.
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{Sieve V1} % give the algorithm a caption
+    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[2] % enter the algorithmic environment
+        \REQUIRE $tree$, $query$, $k$
+        \STATE $Q \leftarrow$ priority queue
+        \STATE $Q.push(tree.root)$
+        \STATE $P \leftarrow$ priority queue
+        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
+            \WHILE{$!Q.extractMax().isLeaf$}
+                \STATE $[l, r] \leftarrow Q.extractMax().children()$
+                \STATE $Q.push([l, r])$
+            \ENDWHILE
+            \STATE $leaf \leftarrow Q.extractMax()$
+            \STATE $P.push(leaf)$
+            \WHILE{$|P| > k$}
+                \STATE $P.extractMax()$
+            \ENDWHILE
+        \ENDWHILE
+        \STATE Return $P$
+    \end{algorithmic}
+    \end{algorithm}
+
+\subsubsection{Linear Search}
+\label{subsubsec:methods:knn-search:leaf-search}
+
+Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
+Initialize $hits$ with the centers of $candidates$.
+
+Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
+Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
+Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
+Repeat until $candidates$ is empty.
+
+\subsubsection{Complexity}
+\label{subsubsec:methods:knn-search:complexity}
+
+This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
+It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.
+
+\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}
+
+\subsubsection {Complexity}
+
+\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
+\label{subsec:methods:rnn-search}
+
+Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.
+
+Modified binary search to identify candidate leaf clusters.
+Exhaustive leaf search over those leaves.
+
+Lift this from CHESS paper. Be sure to cite CHESS here.
+
+Complexity is the same as in CHESS.
+
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
+\label{subsec:methods:knn-search}
+
+The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
+
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
+\begin{itemize}
+    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
+    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
+    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
+\end{itemize}
+
+\subsubsection{Expanding Threshold}
+\label{subsubsec:methods:knn-search:expanding-threshold}
+In this variant of KNN, we repeatedly
+
+Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
+Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
+
+While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
+than $\delta_{min}$ of the top priority element in $Q$, do the following:
+\begin{enumerate}
+\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
+\item Remove $q$ from $Q$ and add all of its points to $P$. 
+\item Remove points from $P$ until $P$ has $k$ points. 
+\end{enumerate}
+
+Return $P$.
+
+\begin{algorithm} % enter the algorithm environment
+\caption{Expanding Threshold} % give the algorithm a caption
+\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
+\begin{algorithmic}[1] % enter the algorithmic environment
+    \REQUIRE $tree$, $query$, $k$
+    \STATE $Q \leftarrow$ priority queue
+    \STATE $Q.push(tree.root)$
+    \STATE $P \leftarrow$ priority queue
+    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
+        \WHILE{$!Q.min.isLeaf$}
+            \STATE $[l, r] \leftarrow Q.extractMin().children()$
+            \STATE $Q.push([l, r])$
+        \ENDWHILE
+        \STATE $leaf \leftarrow Q.extractMin()$
+        \STATE $P.push(leaf)$
+        \WHILE{$|P| > k$}
+            \STATE $P.extractMax()$
+        \ENDWHILE
+    \ENDWHILE
+    \STATE Return $P$
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Sieve V1}
+\label{subsubsec:methods:knn-search:sieve-v1}
+
+In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
+can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
+
+Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
+the root cluster. 
+Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
+
+Repeat the following: 
+\begin{enumerate}
+\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
+\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
+\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
+\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
+\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
+\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
+\item Replace all non-leaf clusters in $Q$ with their children. 
+\end{enumerate}
+
+At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
+Extract $P$.
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{Sieve V1} % give the algorithm a caption
+    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[2] % enter the algorithmic environment
+        \REQUIRE $tree$, $query$, $k$
+        \STATE $Q \leftarrow$ priority queue
+        \STATE $Q.push(tree.root)$
+        \STATE $P \leftarrow$ priority queue
+        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
+            \WHILE{$!Q.extractMax().isLeaf$}
+                \STATE $[l, r] \leftarrow Q.extractMax().children()$
+                \STATE $Q.push([l, r])$
+            \ENDWHILE
+            \STATE $leaf \leftarrow Q.extractMax()$
+            \STATE $P.push(leaf)$
+            \WHILE{$|P| > k$}
+                \STATE $P.extractMax()$
+            \ENDWHILE
+        \ENDWHILE
+        \STATE Return $P$
+    \end{algorithmic}
+    \end{algorithm}
+
+\subsubsection{Sieve V2}
+\label{subsubsec:methods:knn-search:sieve-v2}
+Add explanation of the sieve v2 algorithm.
+
+\subsubsection{Repeated Rnn}
+\label{subsubsec:methods:knn-search:repeated-rnn}
+Add explanation of the repeated rnn algorithm.
+
+\subsubsection{Leaf Search}
+\label{subsubsec:methods:knn-search:leaf-search}
+
+Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
+Initialize $hits$ with the centers of $candidates$.
+
+Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
+Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
+Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
+Repeat until $candidates$ is empty.
+
+\subsubsection{Complexity}
+\label{subsubsec:methods:knn-search:complexity}
+
+Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
+Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
+
+This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
+It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.

### 2023-08-21 sections/4-results.tex
lots of changes

--- 2022-07-07 sections/5-results.tex
@@
     \end{enumerate}
-    \item Compression CAKES + zlib vs just zlib, also substitute other libraries for zlib:
-    \begin{enumerate}[i.]
-        \item just the encoding vs the full dataset,
-        \item encoding + serialization, and
-        \item encoding + serialization + zlib.
 \end{enumerate}
-    \item Measure time taken for accelerated-search in the compressed space and decompression of the hits with different combinations of serde and codec libraries.
-\end{enumerate}
-
-Find the optimal clusters for compression.
-Compare linear search vs accelerated search terminated at these optimal clusters.

### 2023-08-21 sections/2-methods.tex
repeated rnn mostly

@@
 vector space. Some of the algorithms we present use a modified version of the $\rho$-nearest neighbors search 
-algorithm described in CHESS~\cite{ishaq2019clustered} to conduct $k$-NN search. (TODO: make second section less 
+algorithm described in CHESS~\cite{ishaq2019clustered} to conduct $k$-NN search. (TODO: make second sentence less 
 vague once we determine exactly what we're doing with sharding and indexing.)
@@
 
-Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some region of the dataset, 
+Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} of some region of the dataset, 
 which we define as: 
-$$\text{log}_2 (\frac{|B_D(q, r_1)|}{|B_D(q, r_2)|})$$
+$$\text{log}_2 \left( \frac{|B_D(q, r_1)|}{|B_D(q, r_2)|} \right)$$
 where $B_D(q, r)$ is the set of points contained in a ball of radius $r$ 
@@
 leaf clusters.
+
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following cluster 
+properties:
+\begin{itemize}
+    \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
+    \item $\delta_{max} = \delta + r$ is the distance from the query to the theoretically farthest possible instance in $C$.
+    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest possible instance in $C$.
+\end{itemize}
+
+We sometimes also use the above notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
+from a point $x$ to the query $q$. Since we can think of a point as having radius 0, for points we have that $\delta_{max} = \delta = \delta_{min}$.
 
@@
 \label{subsubsec:methods:guaranteed-decrease-in-cluster-radii}
+
+TODO: This needs some doctoring.
 
@@
 
+\begin{algorithm} 
+    \caption{$\rho$-NN(\emph{clusters, query, r})} 
+    \label{alg:rnn} 
+    \begin{algorithmic}[2]
+        \REQUIRE $r \geq 0$
+        \REQUIRE $clusters \neq \emptyset$
+        \STATE $results \leftarrow \emptyset$
+        \IF{$clusters.left$}
+            \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
+                \STATE $\rho$-NN($clusters.left, query, r$)
+            \ENDIF
+        \ENDIF
+        \IF{$clusters.right$}
+            \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
+                \STATE $\rho$-NN($clusters.right, query, r$)
+            \ENDIF
+        \ENDIF
+        \IF{$\neg$$clusters.left$ $\land \neg$$clusters.right$}
+            \FOR{$p \in clusters.points$}
+                \IF{$p.\delta \leq r$}
+                    \STATE $results \leftarrow r$
+                \ENDIF
+            \ENDFOR
+        \ENDIF
+        \STATE Return $results$
+    \end{algorithmic}
+    \end{algorithm}
+
 The asymptotic complexity of $\rho$-nearest neighbors is the same as in CHESS~\cite{ishaq2019clustered}.
@@
 
-The $k$-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
-We present five variants of $k$-nearest neighbors search: Greedy Search, K-NN by Repeated $\rho$-NN, Sieve Search I, Sieve Search II, and 
-Linear Search. We use a preliminary auto-tuning function to predict the variant which will perform 
+In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, K-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
+We use a preliminary auto-tuning function (TODO: do we though?) to predict the variant which will perform 
 best for a given query, dataset, and value of $k$. We then proceed with search using that variant.
-
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-NN variants make use of the following cluster 
-properties:
-\begin{itemize}
-    \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
-    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in $C$.
-    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in $C$.
-\end{itemize}
-
-We sometimes also use the notation $\delta = f(x, q)$ with points instead of clusters to refer to the distance from a point $x$ to the query $q$.
 
@@
 \begin{algorithm} 
-\caption{Greedy Search} 
+\caption{GreedySearch(\emph{tree, query, k})} 
 \label{alg:greedy_search} 
-\begin{algorithmic}[1]
-    \REQUIRE $tree$, $query$, $k$
+\begin{algorithmic}[3]
     \STATE $Q \leftarrow$ priority queue
@@
 \label{subsubsec:methods:knn-search:repeated-rnn}
-
-Hence its name, $k$-NN by Repeated $\rho$-NN repeatedly performs 
-$\rho$-nearest neighbors search, increasing the radius until $k$ neighbors
+Hence the name, with $k$-NN by Repeated $\rho$-NN we repeatedly perform
+$\rho$-nearest neighbors search, increasing the search radius until $k$ neighbors
 have been found.
 
+Let $P$ be the set of nearest neighbors found thus far.
 Search starts with a radius $m$ equal to the radius of the cluster tree divided by
@@
 If no points are within a distance $m$ of the query, we increase the radius by a factor of 
-2 and perform $rho$-NN search again, repeating until at least one point is found.
+2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
+until $|P| > 0$.
 
-Once we have found at least one point, we continue to perform $\rho$-NN search, but instead of 
+Once $|P| > 0$, we continue to perform $\rho$-NN search, but instead of 
 increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
 by the local fractal dimension of the clusters containing the points found so far. In particular, 
-we increase the radius by a factor of $\text{max}(2, \frac{k}{|P|^{\frac{1}{\mu}}}$), where $|P|$ is the number of 
-points found so far and $\mu$ is the mean local fractal dimension of the clusters containing the points found so far.
-Since the local fractal dimension is memoized at clustering time, this poses no additional cost at search time.
+we increase the radius by a factor of 
+$$\text{min}\left(2, \frac{k}{|P|^{\frac{1}{\mu}}}\right)$$
+where $\mu$ is the mean local fractal dimension (abbreviated lfd in algorithm below) of the clusters containing the points found so far.
+Intuitively, the factor by which we increase the radius should be \emph{inversely} proportional to the number of points found so far. 
+Additionally, when the local fractal dimension near the query and found neighbors is low, this suggests that the data 
+are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} proportional to the 
+local fractal dimension. Since the local fractal dimension is memoized at clustering time, this poses no additional cost at search time.
 Once $|P| >= k$, we sort the points in $P$ by $\delta$ and return the $k$ closest points.
 
-TODO: 
-- add intutiion for factor 
-- introduce P variable sooner
+\begin{algorithm} % enter the algorithm environment
+    \caption{Repeated$\rho$-NN(\emph{tree, query, k})} % give the algorithm a caption
+    \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[4] % enter the algorithmic environment
+        \STATE $P \leftarrow$ $\emptyset$
+        \STATE $m \leftarrow$ $\frac{tree.radius}{tree.cardinality}$
+        \WHILE {$|P| = 0$}
+            \STATE $P.push(\rho$-NN$(tree, query, m)$)
+            \STATE $m \leftarrow 2m$
+        \ENDWHILE
+        \WHILE {$|P| < k$}
+            \STATE $clusters \leftarrow \{ C: \exists p \in P \land p \in C \}$
+            \STATE $\mu \leftarrow \frac{1}{|clusters|} \sum_{C \in clusters} C.lfd$
+            \STATE $m \leftarrow \text{min}\left(2, \frac{k}{|P|^{\frac{1}{\mu}}}\right)$
+            \STATE $P.push(\rho$-NN$(tree, query, m)$)
+        \ENDWHILE
+        \STATE $P.sort()$
+        \STATE Return $P[0..k]$
+    \end{algorithmic}
+    \end{algorithm}
 
@@
 \label{subsubsec:methods:knn-search:sieve}
-
-\subsubsection{Sieve Search II}
-
-In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
+With Sieve Search I, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
 can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
 
-Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
-the root cluster. 
-Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
+We start by letting $Q$ be a vector of clusters and points sorted by non-decreasing $\delta_{max}$ and initializing $Q$ with the root cluster. 
+While we haven't found a threshold which guarantees exactly $k$ points, we do the following:
 
-Repeat the following: 
+TODO: Fix this to reflect new algorithm.
+
 \begin{enumerate}
@@
 \item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
-\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
-\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
 \item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{Sieve V1} % give the algorithm a caption
+    \caption{Sieve } % give the algorithm a caption
     \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
@@
 
+\subsubsection{Sieve Search II}
+\label{subsubsec:methods:knn-search:sieve2}
+
 \subsubsection{Linear Search}
@@
 It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.
-
-\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}
-
-\subsubsection {Complexity}
-
-\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
-\label{subsec:methods:rnn-search}
-
-Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.
-
-Modified binary search to identify candidate leaf clusters.
-Exhaustive leaf search over those leaves.
-
-Lift this from CHESS paper. Be sure to cite CHESS here.
-
-Complexity is the same as in CHESS.
-
-\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
-\label{subsec:methods:knn-search}
-
-The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.
-
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
-\begin{itemize}
-    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
-    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
-    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
-\end{itemize}
-
-\subsubsection{Expanding Threshold}
-\label{subsubsec:methods:knn-search:expanding-threshold}
-In this variant of KNN, we repeatedly
-
-Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
-Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
-
-While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
-than $\delta_{min}$ of the top priority element in $Q$, do the following:
-\begin{enumerate}
-\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
-\item Remove $q$ from $Q$ and add all of its points to $P$. 
-\item Remove points from $P$ until $P$ has $k$ points. 
-\end{enumerate}
-
-Return $P$.
-
-\begin{algorithm} % enter the algorithm environment
-\caption{Expanding Threshold} % give the algorithm a caption
-\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
-\begin{algorithmic}[1] % enter the algorithmic environment
-    \REQUIRE $tree$, $query$, $k$
-    \STATE $Q \leftarrow$ priority queue
-    \STATE $Q.push(tree.root)$
-    \STATE $P \leftarrow$ priority queue
-    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
-        \WHILE{$!Q.min.isLeaf$}
-            \STATE $[l, r] \leftarrow Q.extractMin().children()$
-            \STATE $Q.push([l, r])$
-        \ENDWHILE
-        \STATE $leaf \leftarrow Q.extractMin()$
-        \STATE $P.push(leaf)$
-        \WHILE{$|P| > k$}
-            \STATE $P.extractMax()$
-        \ENDWHILE
-    \ENDWHILE
-    \STATE Return $P$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Sieve V1}
-\label{subsubsec:methods:knn-search:sieve-v1}
-
-In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
-can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
-
-Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
-the root cluster. 
-Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.
-
-Repeat the following: 
-\begin{enumerate}
-\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
-\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
-\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
-\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
-\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
-\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
-\item Replace all non-leaf clusters in $Q$ with their children. 
-\end{enumerate}
-
-At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
-Extract $P$.
-
-\begin{algorithm} % enter the algorithm environment
-    \caption{Sieve V1} % give the algorithm a caption
-    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[2] % enter the algorithmic environment
-        \REQUIRE $tree$, $query$, $k$
-        \STATE $Q \leftarrow$ priority queue
-        \STATE $Q.push(tree.root)$
-        \STATE $P \leftarrow$ priority queue
-        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
-            \WHILE{$!Q.extractMax().isLeaf$}
-                \STATE $[l, r] \leftarrow Q.extractMax().children()$
-                \STATE $Q.push([l, r])$
-            \ENDWHILE
-            \STATE $leaf \leftarrow Q.extractMax()$
-            \STATE $P.push(leaf)$
-            \WHILE{$|P| > k$}
-                \STATE $P.extractMax()$
-            \ENDWHILE
-        \ENDWHILE
-        \STATE Return $P$
-    \end{algorithmic}
-    \end{algorithm}
-
-\subsubsection{Sieve V2}
-\label{subsubsec:methods:knn-search:sieve-v2}
-Add explanation of the sieve v2 algorithm.
-
-\subsubsection{Repeated Rnn}
-\label{subsubsec:methods:knn-search:repeated-rnn}
-Add explanation of the repeated rnn algorithm.
 

### 2023-08-22 sections/1-introduction.tex
edits to intro and sieve algorithm

@@
 \label{sec:introduction}
-Often dubbed "the Big Data explosion", rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
-The Sloan Digital Sky Survey (SDSS), for example, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
-Add sentence about MaNGA here: 10,000 galaxies with 2,700 deg$^2$, resolution $R=2,000$ for $\lambda \in [360, 1000) nm$.
+Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
+This is especially true in the field of astronomy. 
+The Sloan Digital Sky Survey (SDSS), for example, in its APOGEE dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
+Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.
 
 Biology has also seen an influx of large, high-dimensional datasets. Silva 18S contains ribosomal DNA sequences of approximately 2.25 
-million genomes with aligned length of 50,000 letters. (Do we need to look at NR99??)
+million genomes with aligned length of 50,000 letters. (TODO: Do we need to talk about NR99; it seems like more of the same but 
+with removed redundancy; not sure if that is helpful to us...)
 
-In many cases, the increase in data outpaces the rate of improvement in computing capacity predicted by Moore's Law~\cite{moore1965cramming}, indicating 
-that computer hardware will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
+In many fields, data collection rates outpace the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
+that computer architecture will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
 requires better algorithms.
 
-The ability to perform similarity search on these datasets enables a variety of applications, including recommendation and classification systems. 
+Many researchers would like the ability to perform similarity search on these datasets. 
+Similatity search enables a variety of applications, including recommendation and classification systems. 
 Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
 understanding and verifying the validity of model outputs. 
-As the sizes and dimensionalities of datasets have grown, efficient and accurate similarity search has become excessively challenging; 
-state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.
-In particular, k-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.
+As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
+even state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.
 
+In particular, $k$-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.
 Naive implementations of kNN, whose asymptotic complexity is a function of the dataset's cardinality, 
-prove prohibitively slow for large datasets. While algorithms for kNN on large datasets exist, most of them are approximate and fail to 
+prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
 exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
-need for efficient, exact search remains.
+need for efficient, \emph{exact} search remains.
 
@@
 Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties of the dataset,
-rather than its cardinality. In 2019, we introduced CHESS (CLustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
+\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
+rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
 neighbors search to a hierarchical clustering approach. In this paper, we introduce CLAM (Clustered Learning of Approximate Manifolds), and CAKES (CLAM-Accelerated K-nearest-neighbor 
-Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm (figure out and explain improvements upon CHESS present in CLAM if there are any algorithmic ones)
+Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm 
 that can facilitate a variety of algorithms for big data. Using the tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
-k-nearest neighbors search in addition to $\rho$-nearest neighbors search.
+$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search.
 
-At some point, need to address the fact that exact search is important for small k and approximate 
-is maybe more ideal for large k. Try to find citation for this.
-
-This paper focuses on:
-\begin{itemize}
-    \item Improvements and extensions of CHESS.
-    \item Accelerated $\rho$-nearest and k-nearest neighbors search.
-    \item Proof for how cluster-radii fall as we descend along the tree.
-    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, hamming, levenshtein, needleman-wunch, and any user-defined function.
-\end{itemize}
-
-Comparisons for:
-\begin{itemize}
-    \item Search time vs FAISS and FALCONN and other in ann-benchmarks suite.
-\end{itemize}
+This paper focuses on improvements to CHESS, accelerated $\rho$- and $k$-nearest neighbors search, proof of a geometric guarantee for clusters, 
+and the implementation of a variety of distance functions, including Euclidean, Manhattan, Cosine, Jaccard, Hamming, Levenshtein, Needleman-Wunch, 
+and any user-defined function. We also provide a comparison of CAKES to FAISS~\cite{johnson2019billion}, FALCONN~\cite{andoni2015practical}, and
+other algorithms in the ann-benchmarks suite~\cite{wang2017ann}.

### 2023-08-22 sections/2-methods.tex
edits to intro and sieve algorithm

@@
 given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
-be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
-When, in addition to these constraints, that the distance function obeys
+be symmetric ($f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
+When, in addition to these constraints, the distance function obeys
 the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
@@
 
-With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
-Given a query $q$, along with a distance function $f$ defined on a dataset $ \textbf{X}$, $k$-nearest neighbors search aims to: 
-for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$; that is, find the $k$ points $x \in \textbf{X}$ with the smallest $f(q, x)$.
-We also have the $\rho$-nearest neighbors search problem: for a given radius $\rho$, find all points in $\textbf{X}$ that are a distance of $\rho$ 
-or less from $q$; that is, all points $x \in \textbf{X}$ such that $f(q, x) \leq \rho$.
-
-Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} of some region of the dataset, 
+Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point in the dataset, 
 which we define as: 
-$$\text{log}_2 \left( \frac{|B_D(q, r_1)|}{|B_D(q, r_2)|} \right)$$
-where $B_D(q, r)$ is the set of points contained in a ball of radius $r$ 
-centered on a point $q$ in the dataset $D$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
-We stress that this concept of local fractal dimension differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
+$$\text{log}_2 \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right)$$
+where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ 
+centered at a point $q$ in the dataset $\textbf{X}$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
+We stress that this concept differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
 consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
 However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
-manifold. The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
+manifold within the 8,575-dimensional embedding space.
+The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
 The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.
@@
 
+With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
+Given a query $q$, along with a distance function $f$ defined on a dataset $ \textbf{X}$, $k$-nearest neighbors search aims to: 
+for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$; that is, find the $k$ points $x \in \textbf{X}$ with the smallest $f(q, x)$.
+We also have the $\rho$-nearest neighbors search problem: for a given radius $\rho$, find all points in $\textbf{X}$ that are a distance of $\rho$ 
+or less from $q$; that is, all points $x \in \textbf{X}$ such that $f(q, x) \leq \rho$.
+
 Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following cluster 
@@
 
-We sometimes also use the above notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
-from a point $x$ to the query $q$. Since we can think of a point as having radius 0, for points we have that $\delta_{max} = \delta = \delta_{min}$.
+We sometimes also use this notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
+from a point $x$ to the query $q$. Since we can think of a point as having radius $r=0$, for points we have that $\delta_{max} = \delta = \delta_{min}$.
 
@@
 improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
-such as local fractal dimension (see Section ADD SECTION: $k$-NN by Repeated $\rho$-NN).
+such as local fractal dimension (see Section 2.4.2: $k$-NN by Repeated $\rho$-NN).
 
@@
 The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
-from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
+from $l$ is designated the right pole. We then partition the cluster into a left child and right child, where the 
 left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
@@
 
-Assume that we have a dataset which follows a $d$-dimensional distribution embedded in $\mathcal{D}$-dimensional space.
+Assume that we have a dataset which follows a $d$-dimensional distribution embedded in $D$-dimensional space.
 We can describe this distribution choosing some set of $d$ mutually orthogonal axes.
@@
 
-Given a cluster $\mathcal{C}$, our partition algorithm will choose the left and right poles, 
-and the axis defined by those poles will be the axis for partitioning $\mathcal{C}$'s points into two child clusters.
+Given a cluster $C$, we choose the left and right poles, 
+and the axis defined by those poles will be the axis for partitioning  $C$'s points into two child clusters.
 For points in each child cluster, the maximum distance between any pair of points along the \emph{axis of partitioning} will have been reduced to at most $\frac{m}{2}$.
-Though this bound holds for points on the axis of partitioning, it does not serve as a bound for the maximum distance between \emph{any} pair of points in $\mathcal{C}$.
-To see this, consider the two dimensional scenario where the points in $\mathcal{C}$ are distributed in a circle of radius $r$. Clearly, the maximum possible distance between 
-any two points in $\mathcal{C}$ is $2r$, and occurs when the points fall on opposite ends of one of its diameters. Now consider a horizontal diameter of the circle, 
+Though this bound holds for points on the axis of partitioning, it does not serve as a bound for the maximum distance between \emph{any} pair of points in  $C$.
+To see this, consider the two dimensional scenario where the points in  $C$ are distributed in a circle of radius $r$. Clearly, the maximum possible distance between 
+any two points in  $C$ is $2r$, and occurs when the points fall on opposite ends of one of its diameters. Now consider a horizontal diameter of the circle, 
 and let $l$ denote the leftmost point on this diameter. Next, consider the orthogonal vertical diameter, and let $x$ denote the uppermost point on this diameter. When we 
-partition along the horizontal diameter, $l$ and $x$ will be assigned to the left child (as mentioned in ADD SECTION, we assign equidistant points to the left child). 
-This reduces the maximum possible distance between any two points on the axis of partition -- the horizontal diameter -- to $r$. However, by considering the 
+partition along the horizontal diameter, $l$ and $x$ will be assigned to the left child (as mentioned in Section 2, we assign equidistant points to the left child). 
+This reduces the maximum possible distance between any two points on the axis of partition --the horizontal diameter-- to $r$. However, by considering the 
 right triangle formed by $l$, $x$, and the center of the circle, we can see that the distance between 
@@
 
-Given a query $q$ and a search radius $\rho$, find all $x \in X$ s.t. $f(q, x) \leq \rho$.
-
-We conduct ranged-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
+We conduct $\rho$-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
 with some implementation improvements including identification of candidate clusters which are leaves and 
@@
 
-In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, K-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
+In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
 We use a preliminary auto-tuning function (TODO: do we though?) to predict the variant which will perform 
@@
 \label{subsubsec:methods:knn-search:greedy-search}
-Greedy Search repeatedly partitions the cluster with the lowest $\delta_{min}$ into its children 
-until it is a leaf, removing it and adding its points to a fixed size priority queue of points. Search terminates 
-when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining
-cluster.
+Greedy Search repeatedly partitions the cluster with the smallest $\delta_{min}$ into its children 
+until the cluster with the smallest $\delta_{min}$ is a leaf. We then remove it and add its points to a fixed size priority queue of points. 
+Search terminates when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining cluster.
 
@@
 \begin{enumerate}
-\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
+\item While the top priority element $q$ in $Q$ is not a leaf, replace $q$ with its children.
 \item Remove $q$ from $Q$ and add all of its points to $P$. 
@@
 \end{enumerate}
-
-Return $P$.
+After this process, the points left in $P$ are the $k$ nearest neighbors of $q$.
 
@@
     \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
-        \WHILE{$!Q.min.isLeaf$}
-            \STATE $[l, r] \leftarrow Q.extractMin().children()$
+        \WHILE{$\neg Q.min.isLeaf$}
+            \STATE $[l, r] \leftarrow Q.extractMin().children$
             \STATE $Q.push([l, r])$
@@
 \label{subsubsec:methods:knn-search:repeated-rnn}
-Hence the name, with $k$-NN by Repeated $\rho$-NN we repeatedly perform
+In this $k$-NN variant, we repeatedly perform
 $\rho$-nearest neighbors search, increasing the search radius until $k$ neighbors
@@
 We start by letting $Q$ be a vector of clusters and points sorted by non-decreasing $\delta_{max}$ and initializing $Q$ with the root cluster. 
-While we haven't found a threshold which guarantees exactly $k$ points, we do the following:
-
-TODO: Fix this to reflect new algorithm.
+While we haven't found a threshold which guarantees exactly $k$ points, we repeat the following:
 
@@
 \item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
-\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
-\item Replace all non-leaf clusters in $Q$ with their children. 
+\item For each cluster left in $Q$, if it is a leaf or has $k$ or fewer points, remove it and all of its points to $Q$.
+\item Replace all remaining non-leaf clusters in $Q$ with their children. 
 \end{enumerate}
 
-At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
-Extract $P$.
+The points remaining in $Q$ are the $k$ nearest neighbors.
+
+TODO: Add find kth algorithm.
 
 \begin{algorithm} % enter the algorithm environment
-    \caption{Sieve } % give the algorithm a caption
-    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[2] % enter the algorithmic environment
+    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
+    \label{alg:sieve1} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[5] % enter the algorithmic environment
         \REQUIRE $tree$, $query$, $k$
@@
         \STATE $Q.push(tree.root)$
-        \STATE $P \leftarrow$ priority queue
-        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
-            \WHILE{$!Q.extractMax().isLeaf$}
-                \STATE $[l, r] \leftarrow Q.extractMax().children()$
+        \WHILE{$|Q| > k$}
+            \STATE $C_{\tau} \leftarrow Q.findKth.\delta_{max}$
+            \FOR {$q \in Q$}
+                \IF {$q.\delta_{min} > \tau$}
+                    \STATE $Q.remove(q)$
+                \ENDIF
+            \ENDFOR
+            \FOR {$q \in Q$}
+                \IF {$q.isLeaf \lor |q.cardinality| \leq k$}
+                    \STATE $Q.push(q.points)$
+                \ELSE
+                    \STATE $[l, r] \leftarrow q.children$
                     \STATE $Q.push([l, r])$   
+                \ENDIF
+                \STATE $Q.remove(q)$
+            \ENDFOR 
         \ENDWHILE
-            \STATE $leaf \leftarrow Q.extractMax()$
-            \STATE $P.push(leaf)$
-            \WHILE{$|P| > k$}
-                \STATE $P.extractMax()$
-            \ENDWHILE
-        \ENDWHILE
-        \STATE Return $P$
+        \STATE Return $Q$
     \end{algorithmic}
@@
 It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.
-
-\subsubsection{Leaf Search}
-\label{subsubsec:methods:knn-search:leaf-search}
-
-Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
-Initialize $hits$ with the centers of $candidates$.
-
-Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
-Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
-Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
-Repeat until $candidates$ is empty.
-
-\subsubsection{Complexity}
-\label{subsubsec:methods:knn-search:complexity}
-
-Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
-Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
-
-This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
-It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.

### 2023-08-23 sections/2-methods.tex
added quicksort partition and better detail in sieve algorithm

@@
     \item sharded search
+    \item dataset depth-first reordering 
     \item index building 
@@
 
-We start by letting $Q$ be a vector of clusters and points sorted by non-decreasing $\delta_{max}$ and initializing $Q$ with the root cluster. 
-While we haven't found a threshold which guarantees exactly $k$ points, we repeat the following:
+We start by letting $Q$ be a vector initialized with the root cluster. This vector will maintain the 
+clusters and points which are still in contention for being one of the nearest neighbors.
+While $Q$ contains at least one cluster (i.e., $Q$ is not a vector of only points), we repeat the process described in 
+the following paragraphs.
 
-\begin{enumerate}
-\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
-\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
-\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
-\item For each cluster left in $Q$, if it is a leaf or has $k$ or fewer points, remove it and all of its points to $Q$.
-\item Replace all remaining non-leaf clusters in $Q$ with their children. 
-\end{enumerate}
+First, 
+we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{max}$ such that 
+$q_{\tau}$ and all elements in $Q$ with smaller $\delta_{max}$ collectively contain at least $k$ points.
 
-The points remaining in $Q$ are the $k$ nearest neighbors.
+To find this element, 
+we use a process similar to the Partition algorithm used in Quicksort (CITE: original Quicksort paper and reference wherever it is 
+in our paper), adjusted to account for the varying cardinalities of elements in $Q$. Our modified algorithm finds the smallest index $i$ such 
+that all elements in $Q$ with $\delta_{max}$ closer to or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater
+than or equal to $k$. In the process of finding this index, we alter the order of $Q$ so that all elements to left
+of index $i$ have $\delta_{max}$ less than or equal to $Q[i]$'s $\delta_{max}$ and all elements to right of index $i$
+have $\delta_{max}$ greater than $Q[i]$'s $\delta_{max}$. We consider $Q[i]$ to be our $q_{\tau}$, and our threshold 
+$\tau$ to be $Q[i]$'s $\delta_{max}$.
 
-TODO: Add find kth algorithm.
+We then remove from $Q$ any element whose $\delta_{min}$ is greater than $\tau$. Then, for each element left 
+in $Q$, if it is a leaf or contains $k$ or fewer points, we remove it and all of its points to $Q$. Finally, we
+replace all remaining non-leaf clusters in $Q$ with their children.
+
+We complete this process until $Q$ contains only points. At this point, we 
+just use [INSERT ref to quicksort partition in paper] to find the $k$th 
+nearest neighbor. Since this algorithm will also elements on the correct side of 
+the index of the $k$th nearest neighbor, the first $k$ elements in $Q$ a are the $k$-nearest neighbors.
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{QuicksortPartition(\emph{Q, k, l, r})} % give the algorithm a caption
+    \label{alg:quicksort-partition} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[5] % enter the algorithmic environment
+        \IF{$l \geq r$}
+            \STATE Return $min(l, r)$
+        \ELSE 
+            \STATE $pivot = l + (r - l) / 2$
+            \STATE $Q.swap(pivot, r)$
+            \STATE $a \leftarrow l$
+            \STATE $b \leftarrow l$
+            \WHILE {$b < r$}
+                \IF {$Q[b].\delta_{max} < Q[r].\delta_{max}$}
+                    \STATE $Q.swap(a, b)$
+                    \STATE $a \leftarrow a + 1$
+                \ENDIF
+                \STATE $b \leftarrow b + 1$
+            \ENDWHILE
+            \STATE $Q.swap(a, r)$
+            \STATE $p \leftarrow r$
+            \STATE $g \leftarrow \sum_{i=0}^{p} Q[i].cardinality$
+            \IF {$g = k$}
+                \STATE Return $p$
+            \ELSIF {$g < k$}
+                \STATE $QuicksortPartition(Q, k, p + 1, r)$
+            \ELSE 
+                \IF {$p > 0 \land g > k + Q[p - 1].cardinality$}
+                    \STATE $QuicksortPartition(Q, k, l, p - 1)$
+                \ELSIF {$p > 0 \land g = k + Q[p - 1].cardinality$}
+                    \STATE Return $p - 1$
+                \ELSE
+                    \STATE Return $p$
+                \ENDIF
+            \ENDIF
+        \ENDIF
+    \end{algorithmic}
+    \end{algorithm}
 
@@
     \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
-    \label{alg:sieve1} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[5] % enter the algorithmic environment
+    \label{alg:sieve} % and a label for \ref{} commands later in the document
+    \begin{algorithmic}[6] % enter the algorithmic environment
         \REQUIRE $tree$, $query$, $k$
-        \STATE $Q \leftarrow$ priority queue
-        \STATE $Q.push(tree.root)$
+        \STATE $Q \leftarrow$ [$tree.root$]
         \WHILE{$|Q| > k$}
-            \STATE $C_{\tau} \leftarrow Q.findKth.\delta_{max}$
+            \STATE $i \leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
+            \STATE $\tau \leftarrow Q[i].\delta_{max}$
             \FOR {$q \in Q$}

### 2023-08-24 sections/1-introduction.tex
some edits to intro

@@
 Biology has also seen an influx of large, high-dimensional datasets. Silva 18S contains ribosomal DNA sequences of approximately 2.25 
-million genomes with aligned length of 50,000 letters. (TODO: Do we need to talk about NR99; it seems like more of the same but 
-with removed redundancy; not sure if that is helpful to us...)
+million genomes with aligned length of 50,000 letters.
+
+% (TODO: Do we need to talk about NR99; it seems like more of the same but 
+% with removed redundancy; not sure if that is helpful to us...) 
+% No we don't... I'll explain why if you want.
 
@@
 
-In particular, $k$-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.
-Naive implementations of kNN, whose asymptotic complexity is a function of the dataset's cardinality, 
-prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
+In particular, $k$-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}; most typically classification relies upon a majority vote among the $k$ nearest neighbors to a query.
+Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
+prove prohibitively slow for large datasets. While algorithms for fast kNN on large datasets exist, most of them are approximate and do not 
 exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
 need for efficient, \emph{exact} search remains.
+In particular, while approximate kNN may agree with exact kNN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.
 
@@
 rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
-neighbors search to a hierarchical clustering approach. In this paper, we introduce CLAM (Clustered Learning of Approximate Manifolds), and CAKES (CLAM-Accelerated K-nearest-neighbor 
-Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm 
-that can facilitate a variety of algorithms for big data. Using the tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
+neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated K-nearest-neighbor 
+Entropy-scaling Search).
+CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered} is a refinement of the CHESS clustering algorithm.
+Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform
 $k$-nearest neighbors search in addition to $\rho$-nearest neighbors search.

### 2023-08-28 sections/1-introduction.tex
forgot to push a bunch of stuff so here we go

@@
 \label{sec:introduction}
-Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
+Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
 This is especially true in the field of astronomy. 
 The Sloan Digital Sky Survey (SDSS), for example, in its APOGEE dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
-Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.
+Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$. [add sentence about big ANN competition]
 
@@
 
-In many fields, data collection rates outpace the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
-that computer architecture will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
+In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
+that computer architecture improvements will not "catch up" to our computational needs in the near future. As a result, analysis of these large datasets 
 requires better algorithms.
 
-Many researchers would like the ability to perform similarity search on these datasets. 
-Similatity search enables a variety of applications, including recommendation and classification systems. 
+Many researchers are especially interested in performing similarity search on these datasets. 
+Similarity search enables a variety of applications, including recommendation and classification systems. 
 Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
@@
 
-In particular, $k$-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}; most typically classification relies upon a majority vote among the $k$ nearest neighbors to a query.
+In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}; 
 Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
-prove prohibitively slow for large datasets. While algorithms for fast kNN on large datasets exist, most of them are approximate and do not 
-exploit the inherent structure and redundancy often present in big datasets. Approximate search is sufficient for many applications, but the 
+prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
+exploit the geometric and topological structure inherent in big datasets. Approximate search is sufficient for some applications, but the 
 need for efficient, \emph{exact} search remains.
-In particular, while approximate kNN may agree with exact kNN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.
+In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.
 
@@
 clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
-Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
+Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in big datasets.
 Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
@@
 rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
-neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated K-nearest-neighbor 
+neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor 
 Entropy-scaling Search).
-CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered} is a refinement of the CHESS clustering algorithm.
+CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm.
 Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
-$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search.
+$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search (also known as ranged nearest neighbors search).
 
-This paper focuses on improvements to CHESS, accelerated $\rho$- and $k$-nearest neighbors search, proof of a geometric guarantee for clusters, 
-and the implementation of a variety of distance functions, including Euclidean, Manhattan, Cosine, Jaccard, Hamming, Levenshtein, Needleman-Wunch, 
-and any user-defined function. We also provide a comparison of CAKES to FAISS~\cite{johnson2019billion}, FALCONN~\cite{andoni2015practical}, and
-other algorithms in the ann-benchmarks suite~\cite{wang2017ann}.
+This paper focuses on accelerated $\rho$- and $k$-nearest neighbors search. We explain some 
+implementation improvements to CHESS, as well as improved genericity across distance functions. 
+We also include proof of some geometric guarantees for clusters. Finally, we provide a comparison of 
+CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} on several datasets 
+from the ann and big ann benchmarks suites.

### 2023-08-28 sections/2-methods.tex
forgot to push a bunch of stuff so here we go

@@
 
-In this manuscript, we are primarily concerned with $k$-nearest neighbors search in a finite-dimensional 
-vector space. Some of the algorithms we present use a modified version of the $\rho$-nearest neighbors search 
-algorithm described in CHESS~\cite{ishaq2019clustered} to conduct $k$-NN search. (TODO: make second sentence less 
-vague once we determine exactly what we're doing with sharding and indexing.)
+In this manuscript, we are primarily concerned with $k$-nearest neighbors in a finite-dimensional 
+vector space.
 
@@
 Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
-Euclidean (L2) and Cosine distances. With biological or string data on the other hand, Hamming and Levenshtein distances are more appropriate.
+Euclidean (L2) and Cosine distances. With biological or string data, Hamming and Levenshtein distances are more appropriate.
 
@@
 The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.
+[TODO: add example for GreenGenes and add lfd plots from CHESS]
 
 We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
-The \emph{center} is the geometric median of the set of points in the \emph{cluster}, and so it is a real data point. The \emph{radius} is the
+The \emph{center} is the geometric median of a sample of points in the \emph{cluster}, and so it is a real data point. The \emph{radius} is the
 maximum distance from a point in the cluster to its \emph{center}. We estimate \emph{local fractal dimension} at the cluster radius and half
@@
 a binary tree has two child nodes. We define the \emph{metric entropy} of a data set under a hierarchical clustering scheme as a refinement of [7], where
-metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a binary, divisive clustering 
-approach, but with early stopping criteria; clusters with fewer than a threshold number of points are not further split. 
+metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a hierarchical, divisive clustering 
+approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g. a specified cardinality or radius) are not further split. 
 Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of
@@
 
+[TODO: Do set comprehension definitions for these and then prose refinement]
 With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
@@
 
+[TODO: talk about singleltons or whatever and overloading of notation]
 We sometimes also use this notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
@@
 
+[TODO: take out memoization and talk about reordering instead]
 We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
@@
 
+[TODO: use cluster C and |C| instead here; make sure you mention how radius is calculated; sampling, center, radius, poles; 
+define geometric median]
 For a cluster with $m$ points, we begin by computing a 
@@
 \end{algorithm}
+
+\subsubsection {Dataset Reordering}
+
+[2 tiers of improvement]
+
+CLAM-CAKES also reorders the dataset to improve the tree-building process 
+used in CHESS. In CHESS, each cluster stored a vector of indices into the dataset, which 
+were used to retrieve its points during search. Thus, for a dataset with cardinality $n$, 
+at each depth in the tree, we would store a total of $n$ indices. Since we have approximately
 
@@
     \item sharded search
-    \item dataset depth-first reordering 
     \item index building 
@@
 
+[TODO: delete this and talk about overlapping children improvement instead]
 We conduct $\rho$-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
@@
 
-In this section, we present four novel variants of $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
-We use a preliminary auto-tuning function (TODO: do we though?) to predict the variant which will perform 
+[TODO: reorder so rho nn is first, then sieves, then Greedy Sieve]
+In this section, we present four novel algorithms for $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
+We also have an implementation of linear search. We use a preliminary auto-tuning function to predict the variant which will perform 
 best for a given query, dataset, and value of $k$. We then proceed with search using that variant.
 
+[TODO: tighten up or omit those opening paragraphs which roughly explain the algorithm]
+
+[TODO: call queues of points H]
 \subsubsection{Greedy Search}
@@
 
+[TODO: replace with pop and push language and start with initial pop]
 While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
@@
 After this process, the points left in $P$ are the $k$ nearest neighbors of $q$.
+
+[TODO: fix to use priority queue language]
 
@@
 \label{subsubsec:methods:knn-search:repeated-rnn}
-In this $k$-NN variant, we repeatedly perform
-$\rho$-nearest neighbors search, increasing the search radius until $k$ neighbors
+In this algorithm, we perform
+$\rho$-nearest neighbors search at [TODO: add radius], repeatedly increasing the search radius until $k$ neighbors
 have been found.
 
+[TODO: change P to H]
 Let $P$ be the set of nearest neighbors found thus far.
@@
 
+[TODO: Add number next to all equations and fix factor so that it is correct (k should be raised too)]
 Once $|P| > 0$, we continue to perform $\rho$-NN search, but instead of 
@@
 
+[TODO: change sieve II to sieev with separate centers]
 \subsubsection{Sieve Search I}
 \label{subsubsec:methods:knn-search:sieve}
+[TODO: make writeup here tighter and more consistent with what's in the algorithm; talk 
+about enum variants here]
 With Sieve Search I, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
@@
 \label{subsubsec:methods:knn-search:sieve2}
+Sieve Search II is exactly the same as Sieve Search I, but with the following modification: clusters 
+in $Q$ are represented twice-- once as their center and once as the rest of their points. 
+Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster 
+center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, 
+thus allowing us to eliminate some clusters from $Q$ earlier in the process.
 
-\subsubsection{Linear Search}
-\label{subsubsec:methods:knn-search:leaf-search}
-
-Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
-Initialize $hits$ with the centers of $candidates$.
-
-Let $f_k$ be the distance to the $k^{th}$ farthest instance so far.
-Exclude all $candidates$ for which $\delta^2 > f_k$, i.e. their closest possible instance is farthest that the $k^{th}$ farthest instance found so far.
-Remove the closest cluster from $candidates$ and add all its contained instances to $hits$.
-Repeat until $candidates$ is empty.
-
-\subsubsection{Complexity}
-\label{subsubsec:methods:knn-search:complexity}
-
-This is at-least on-par with the complexity for $\rho$-nearest neighbors search from CHESS.
-It is potentially better because we effectively shrink the search radius to adapt to the $k$ closest instances found during the algorithm.
+\subsubsection{Complexity of $k$-NN Search}
+\label{paragraph:methods:knn-complexity}
+Add asymptotic complexity of various $k$-NN algorithms here.

### 2023-08-28 sections/3-datasets.tex
forgot to push a bunch of stuff so here we go

\section{Datasets and Distance Functions}
\label{sec:datasets-and-distance-functions}

Provide details on the datasets we use for benchmarks.

Add interesting t-SNE or UMAP embeddings of datasets.
For example~\ref{fig:discussion:umap-annthyroid-euclidean}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
    \label{fig:discussion:umap-annthyroid-euclidean}
\end{figure}

\subsection{APOGEE-2}
\label{subsec:datasets:apogee-2}
The Sloan Digital Sky Survey (SDSS) [3] Apache Point Observatory Galaxy Evolution
Experiment (APOGEE) [4] contains near-infrared
spectra of approximately 130,000 stellar objects in the Milky
Way galaxy. Each spectral datum describes a star's flux in
Janskys ($1 Jy = 10^{-26} W \dot m^{-2} \dot Hz^{-1}$) at each of 8,575 wavelengths ranging from $\lambda$ = 1.51 µm to 1.70 µm. Thus,
Thus, each datum is a real-valued vector in $\mathbb{R}_{+}^{8575}$.

Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history.

TODO: Add details about why this dataset is a good candidate for CAKES.
TODO: Talk about distance functions used with this dataset: euclidean, cosine, Wasserstein-1d

\subsection{MaNGA}
\label{subsec:datasets:manga}
SDSS's Mapping Nearby Galaxies at APO (MaNGA) [CITE] contains spectral measurements across the face of each of 
about 10,000 galaxies across 2,700 deg$^2$.

Like APOGEE, MaNGA helps astronomers understand the history of galaxies throughout their entire lifespan.

TODO: Add details about why this dataset is a good candidate for CAKES. 
TODO: Talk about distance functions used with this dataset: euclidean, cosine, Wasserstein-2d

\subsection{Silva 18S}
\label{subsec:datasets:silva-18s}

TODO: Which Silva dataset exactly are we using and what is its cardinality and dimensionality?

TODO: Add details about why this dataset is a good candidate for CAKES.
TODO: Talk about distance functions used with this dataset: hamming, levenshtein

\dots

\subsection{ANN-Benchmarks suite}
\label{subsec:datasets:ann-benchmarks-suite}

deep1b: cosine

fashion-mnist: euclidean

gist: euclidean

glove: cosine

kosarak: jaccard

mnist: euclidean

nytimes: cosine

sift: euclidean

last.fm: cosine

\dots

### 2023-08-29 sections/2-methods.tex
dataset reordering

@@
 which we define as: 
-$$\text{log}_2 \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right)$$
+\begin{equation} \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) } \label{1} \end{equation}
 where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ 
@@
 
-[TODO: Do set comprehension definitions for these and then prose refinement]
 With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
-Given a query $q$, along with a distance function $f$ defined on a dataset $ \textbf{X}$, $k$-nearest neighbors search aims to: 
-for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$; that is, find the $k$ points $x \in \textbf{X}$ with the smallest $f(q, x)$.
-We also have the $\rho$-nearest neighbors search problem: for a given radius $\rho$, find all points in $\textbf{X}$ that are a distance of $\rho$ 
-or less from $q$; that is, all points $x \in \textbf{X}$ such that $f(q, x) \leq \rho$.
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
+the set $\{x \in \textbf{X}: f(q, x) \leq f(q, x_k) \land x_k \text{ is the $k$th nearest neighbor of $q$ in $\textbf{X}$}\}$; that is,
+for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
+We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
+find all points in $\textbf{X}$ that at most a distance $\rho$ from $q$.
 
@@
 
-[TODO: talk about singleltons or whatever and overloading of notation]
-We sometimes also use this notation with points instead of clusters. For example, we sometimes use $\delta = f(x, q)$ to refer to the distance 
-from a point $x$ to the query $q$. Since we can think of a point as having radius $r=0$, for points we have that $\delta_{max} = \delta = \delta_{min}$.
+We define \emph{singletons} as clusters which either contain a single point (i.e., have cardinality 1) or which contain 
+many instances of the same point (i.e., have cardinality greater than 1 but contain only one unique point). Singletons clearly 
+have zero radius, and so $\delta_{max} = \delta_{min} = \delta$ for these clusters. Hence, we sometimes overload the above 
+notation to refer to distance from a query to a point. Similarly, we have that $\delta_{max} = \delta = \delta_{min}$ for 
+individual points.
 
@@
 
-[TODO: take out memoization and talk about reordering instead]
 We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
 similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
-improvements: better selection of poles for partitioning and memoization of useful cluster properties, 
-such as local fractal dimension (see Section 2.4.2: $k$-NN by Repeated $\rho$-NN).
+improvements: better selection of poles for partitioning and depth-first dataset reordering 
+(see Section ADD SECTION: Dataset Reordering).
 
@@
 
-[TODO: use cluster C and |C| instead here; make sure you mention how radius is calculated; sampling, center, radius, poles; 
-define geometric median]
-For a cluster with $m$ points, we begin by computing a 
-random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
-cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
-The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
-from $l$ is designated the right pole. We then partition the cluster into a left child and right child, where the 
+For a cluster $C$ with $|C|$ points, we begin by computing a 
+random sample of $\sqrt{|C|}$ points, and computing pairwise distances 
+between all points in the sample. Based on those pairwise distances, we compute the \emph{geometric median} of this sample; 
+that is, the point which minimizes the sum of distances to all other points in the sample. This geometric median 
+is $C$'s \emph{center}. The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
+The point $l$ which is furthest from the center is designated the \emph{left pole}, and the point $r$ which is furthest
+from $l$ is designated the \emph{right pole}. We then partition the cluster into a \emph{left child} and \emph{right child}, where the 
 left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
@@
 
-\subsubsection {Dataset Reordering}
-
-[2 tiers of improvement]
+\subsubsection {Dataset Depth-First Reordering}
 
 CLAM-CAKES also reorders the dataset to improve the tree-building process 
-used in CHESS. In CHESS, each cluster stored a vector of indices into the dataset, which 
-were used to retrieve its points during search. Thus, for a dataset with cardinality $n$, 
-at each depth in the tree, we would store a total of $n$ indices. Since we have approximately
+used in CHESS. In CHESS, each cluster stored a vector of indices of its points. 
+We use these indices during search to retrieve a cluster's points from the dataset. 
+Though this approach allowed us to retrieve the points in constant 
+time, its memory cost was prohibitively high; since each cluster stores indices of all of its
+points, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
+With $\log{}n$ levels in the tree, this approach had $\mathcal{O}(n\log{}n)$ memory cost.
+
+A later improvement reduced the memory cost to $\mathcal{O}(n)$ by storing only the indices of the points in the leaves.
+This approach, however, introduced an $\mathcal{O}(n)$ time cost every time we needed to find the indices for 
+a non-leaf cluster, since this required a traversal of the subtree rooted at that cluster.
+
+In this work, we introduce a new approach, wherein after building the cluster tree, we reorder the dataset 
+so that points are stored in a depth-first order based on the tree. Then, within each cluster, we need only 
+store the cluster's cardinality, and an \emph{offset} to access the cluster's points from the dataset.
+For left leaf clusters, the offset is the index of the first point in the cluster. Right leaves have an offset 
+equal to their sibling's offset plus the their sibling's cardinality. Parents have the same offset as their left child.
+With $\mathcal{O}(n)$ memory cost and $\mathcal{O}(1)$ time cost for retrieving a points during search, 
+dataset depth-first reordering offers significant improvements over the previous approaches.
 
@@
 [TODO: delete this and talk about overlapping children improvement instead]
-We conduct $\rho$-nearest neighbors search similarly as described in CHESS~\cite{ishaq2019clustered}, but 
+We conduct $\rho$-nearest neighbors search as described in CHESS~\cite{ishaq2019clustered}, but 
 with some implementation improvements including identification of candidate clusters which are leaves and 

### 2023-08-31 sections/1-introduction.tex
getting there

@@
 In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}; 
-Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
+naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
 prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
@@
 
-This paper focuses on accelerated $\rho$- and $k$-nearest neighbors search. We explain some 
+This paper focuses on four novel accelerated $k$-nearest neighbors search algorithms. We explain some 
 implementation improvements to CHESS, as well as improved genericity across distance functions. 

### 2023-08-31 sections/2-methods.tex
getting there

@@
 In this manuscript, we are primarily concerned with $k$-nearest neighbors in a finite-dimensional 
-vector space.
-
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of 
+vector space. Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of 
 an organism, the vector representation of a single image, or any entity on which we can define a \emph{distance function}).
@@
 given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
-be symmetric ($f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
+be symmetric (i.e., $f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
 When, in addition to these constraints, the distance function obeys
@@
 The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.
-[TODO: add example for GreenGenes and add lfd plots from CHESS]
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=2.5in]{images/lfd_plots/lfd_plot_apogee.png}
+    \caption{Mean fractal dimension of APOGEE clusters as a function of
+    depth when clustered based on L2 norm. Each plot line represents a distinct
+    decile of fractal dimension. Beyond a depth of 56, no clusters are further
+    divided due to the minimum-cluster-cardinality stopping criteria.}
+    \label{fig:methods:lfd_apogee}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=2.5in]{images/lfd_plots/lfd_plot_greengenes.png}
+    \caption{Mean fractal dimension of GreenGenes clusters as a function of
+    depth when clustered based on Hamming distance. Each plot line represents
+    a distinct decile of fractal dimension.}
+    \label{fig:methods:lfd_greengenes}
+\end{figure}
+
+Figure~\ref{fig:methods:lfd_apogee} shows the mean local fractal dimension of the APOGEE dataset at each level of hierarchical clustering, 
+under the L2 (Euclidean) distance metric. Each plot line represents a decile of fractal dimension; note that, other than the most extreme 10\% of clusters,
+virtually all clusters have a local fractal dimension of less than $2 \ll 8,575$. This suggests that APOGEE is a good candidate for use with CAKES.
+
+Figure~\ref{fig:methods:lfd_greengenes} shows the mean local fractal dimension of the GreenGenes dataset at each level of hierarchical clustering,
+under the Hamming distance metric. Each plot line represents a decile of fractal dimension; once again, other than the most extreme 10\% of clusters,
+virtually all clusters have a local fractal dimension of less than $2 \ll 2,250$. This suggests that GreenGenes is a good candidate for use with CAKES.
 
@@
 
-With these concepts defined, we can now pose the $k$- and $\rho$ nearest neighbors search problems.
+With these concepts defined, we can now pose the $k$- and $\rho$- nearest neighbors search problems.
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $\{x \in \textbf{X}: f(q, x) \leq f(q, x_k) \land x_k \text{ is the $k$th nearest neighbor of $q$ in $\textbf{X}$}\}$; that is,
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in X \setminus S_q$, $f(q, x) < \max\{f(y, q): y \in S_q \}$; that is,
 for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
 We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
-find all points in $\textbf{X}$ that at most a distance $\rho$ from $q$.
+find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
 
@@
 have zero radius, and so $\delta_{max} = \delta_{min} = \delta$ for these clusters. Hence, we sometimes overload the above 
-notation to refer to distance from a query to a point. Similarly, we have that $\delta_{max} = \delta = \delta_{min}$ for 
-individual points.
+notation to refer to distance from a query to a point; in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
 
@@
 
+\subsubsection {Cluster Partitioning}
+
 For a cluster $C$ with $|C|$ points, we begin by computing a 
-random sample of $\sqrt{|C|}$ points, and computing pairwise distances 
+random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances 
 between all points in the sample. Based on those pairwise distances, we compute the \emph{geometric median} of this sample; 
@@
 is $C$'s \emph{center}. The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
-The point $l$ which is furthest from the center is designated the \emph{left pole}, and the point $r$ which is furthest
+The point $l$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}, and the point $r$ which is furthest
 from $l$ is designated the \emph{right pole}. We then partition the cluster into a \emph{left child} and \emph{right child}, where the 
@@
 \begin{algorithm} % enter the algorithm environment
-\caption{Partition} % give the algorithm a caption
+\caption{Cluster Partition} % give the algorithm a caption
 \label{alg:partition} % and a label for \ref{} commands later in the document
@@
 
-CLAM-CAKES also reorders the dataset to improve the tree-building process 
-used in CHESS. In CHESS, each cluster stored a vector of indices of its points. 
-We use these indices during search to retrieve a cluster's points from the dataset. 
+CAKES also improves upon CHESS by reordering the dataset. In CHESS, each cluster stored a list of indices into the dataset, 
+which we used during search to retrieve a cluster's points from the dataset. 
 Though this approach allowed us to retrieve the points in constant 
-time, its memory cost was prohibitively high; since each cluster stores indices of all of its
+time, its memory cost was prohibitively high; since each cluster stores indices for each of its
 points, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
@@
 
-A later improvement reduced the memory cost to $\mathcal{O}(n)$ by storing only the indices of the points in the leaves.
-This approach, however, introduced an $\mathcal{O}(n)$ time cost every time we needed to find the indices for 
-a non-leaf cluster, since this required a traversal of the subtree rooted at that cluster.
+One may think to improve this memory cost to $\mathcal{O}(n)$ by storing only the indices of points at 
+the leaf cluster level.
+This approach, however, introduces an $\mathcal{O}(n)$ time cost whenever we need to find the indices for 
+a non-leaf cluster, since it a traversal of the subtree rooted at that cluster.
 
 In this work, we introduce a new approach, wherein after building the cluster tree, we reorder the dataset 
-so that points are stored in a depth-first order based on the tree. Then, within each cluster, we need only 
-store the cluster's cardinality, and an \emph{offset} to access the cluster's points from the dataset.
-For left leaf clusters, the offset is the index of the first point in the cluster. Right leaves have an offset 
-equal to their sibling's offset plus the their sibling's cardinality. Parents have the same offset as their left child.
-With $\mathcal{O}(n)$ memory cost and $\mathcal{O}(1)$ time cost for retrieving a points during search, 
-dataset depth-first reordering offers significant improvements over the previous approaches.
+so that points are stored in a depth-first order. Then, within each cluster, we need only 
+store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset. Parent clusters 
+have the same offset as their left child, and right child clusters have an offset equal to their sibling's offset
+plus their siblings's cardinality. With no additional memory cost nor time cost for retrieving a points during search, 
+dataset depth-first reordering offers significant improvements over the approach used in CHESS.
 
-\subsubsection {Guaranteed Decrease in Cluster Radii}
+\subsubsection {Scaling Behavior of Cluster Radii}
 \label{subsubsec:methods:guaranteed-decrease-in-cluster-radii}
+While it may be tempting to assume that cluster radii decrease with each application of Cluster Partition (TODO: ADD SECTION), unfortunately, this assumption is incorrect. 
+Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii; in particular, we can guarantee that cluster radii will stop increasing after at most 
+$d$ partitions, where $d$ is the dimensionality of the dataset.  In the remainder of this subsection, we HJFHFJ.
 
-TODO: This needs some doctoring.
+We can describe this a $d$-dimensional distribution by choosing some set of $d$ mutually orthogonal axes.
+Let $2R$ be the maximum pairwise distance among the instances in the dataset. 
+We choose the axes such that the two points that are $2R$ apart lie along one of the axes. 
+Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
+In the worst case, e.g. with a distribution that fills the $d$-sphere, our axes will be such that $2R$ is the maximum pairwise distance along every axis. 
+Such a distribution would also produce a balanced clustering.
 
-We show that cluster radius is guaranteed to decrease after at most $d$ partitions.
+Partition will select a maximally distant pair of points to use as poles, i.e., it will choose one of the $d$ axes along
+which to split the cluster into two children. 
+After one application of Cluster Partition, the maximum pairwise distance along that axis will be
+bounded above by $R$. 
+The next recursive Partition will select another of the $d$ axes. 
+Thus, after at most $d$ applications of Cluster Partition, the
+maximum pairwise distance along each axis will be bounded above by $R$. 
+The overall (i.e., not restricted to be along one axis) maximum pairwise distance, 
+will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes.
 
-Assume that we have a dataset which follows a $d$-dimensional distribution embedded in $D$-dimensional space.
-We can describe this distribution choosing some set of $d$ mutually orthogonal axes.
-Let $m$ denote the maximum distance between any pair of points along any axis from any choice of mutually orthogonal axes for our $d$-dimensional distribution.
-In the worst case scenario, the distribution of our data is a $d$-sphere and $m$ is the maximum distance between two points along every axis for every choice of axes.
-In this scenario, the cluster radius is $r = \frac{m}{2}$.
+Thus, starting with a cluster $C$ of radius $R$, after at most $d$ Partitions, the descendants of $C$ will each have radius
+bounded above by $\frac{r}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at 
+most $d$ applications of Cluster Partition.
 
-Given a cluster $C$, we choose the left and right poles, 
-and the axis defined by those poles will be the axis for partitioning  $C$'s points into two child clusters.
-For points in each child cluster, the maximum distance between any pair of points along the \emph{axis of partitioning} will have been reduced to at most $\frac{m}{2}$.
-Though this bound holds for points on the axis of partitioning, it does not serve as a bound for the maximum distance between \emph{any} pair of points in  $C$.
-To see this, consider the two dimensional scenario where the points in  $C$ are distributed in a circle of radius $r$. Clearly, the maximum possible distance between 
-any two points in  $C$ is $2r$, and occurs when the points fall on opposite ends of one of its diameters. Now consider a horizontal diameter of the circle, 
-and let $l$ denote the leftmost point on this diameter. Next, consider the orthogonal vertical diameter, and let $x$ denote the uppermost point on this diameter. When we 
-partition along the horizontal diameter, $l$ and $x$ will be assigned to the left child (as mentioned in Section 2, we assign equidistant points to the left child). 
-This reduces the maximum possible distance between any two points on the axis of partition --the horizontal diameter-- to $r$. However, by considering the 
-right triangle formed by $l$, $x$, and the center of the circle, we can see that the distance between 
-$l$ and $x$ is $\sqrt{2}{r}$.
+Note that, in practice, we never see a balanced clustering. Cluster Partition  produces unbalanced trees due to the varying density of the sampling 
+in different regions of the manifold and the low dimensional "shape" of the manifold. Further, the cluster radii decrease by a factor much larger than 
+$\frac{\sqrt{2}}{2}$ in practice, and often this decrease starts occurring after only one or two applications of Cluster Partition, rather than the 
+upper bound of $d$ applications.
 
-If $d > 1$ then the child clusters will have a maximal chord $m_{child}$ such that $m_{child} \leq m$.
-By partitioning the cluster in this way, we have taken one axis and reduced the maximum distance along that axis from $m$ to $\frac{m}{2}$.
-However, as per the distribution of data as discussed earlier, there may be other axes along which the maximum pairwise distance is still $m$.
-We, thus, recursively partition each child cluster.
-
-With each recursive application of Partition, we consume one axis along which the maximum distance was $m$ and reduce the maximum distance along that axis to $\frac{m}{2}$ in the child clusters.
-After $d$ recursive applications of partition, we will have exhausted the $d$ axes with the maximum distance of $m$.
-After these $d$ partitions, the points in each child cluster will have a maximum pairwise distance of $\frac{m}{2}$ along any axis.
-Thus the radii of those child clusters will be at most $\frac{m}{4}$.
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=2.5in]{images/geometry/sketchometry_final.png}
+    \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk}
+    \label{fig:methods:scaling_behavior}
+\end{figure}
 
@@
 \label{subsubsec:methods:clustering:complexity}
+%A lot of this is lifted from the CHESS paper and could probably get cut out. %
 
-Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
-Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
+To analyze the computational complexity of hierarchical
+clustering, we start by considering the top level of the hierarchy. 
+For a data set with cardinality $n$, we have $n$ data points at this 
+level, from which we randomly choose a sample of $\sqrt{n}$ points. 
+We compute distances between every pair of seeds in this sample, 
+thus making $n$ distance comparisons. Based on these distances 
+we compute the center and left and right poles. Once the poles 
+have been chosen, every data point except for the two poles is 
+compared to both poles, for an additional 2$n$ comparisons.
 
-\subsection {Index-Building}
-\begin{itemize}
-    \item sharded search
-    \item index building 
-    \item optimization of number of shards 
-\end{itemize}
+In general, at depth $d$ of
+the hierarchical tree, we have $\frac{n}{2^d}$ points per cluster (assuming a
+perfectly balanced clustering), $\sqrt{\frac{n}{2^d}}$ seeds, $\frac{n}{2^d}$
+pairwise seed comparisons, and $\frac{2n}{2^{d}}$ additional comparisons when 
+each other data point is compared to the two poles, for a total of $\frac{3n}{2^d}$
+distance comparisons. At each level, there are $2^d$ clusters,
+so the total number of distance comparisons per level is
+3$n$. Thus, the asymptotic complexity in terms of distance
+comparisons for a cluster tree of depth $d$ is $\mathcal{O}(dn)$
+
+By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of 
+partitioning and $\mathcal{O}(n \log n)$ cost of building the tree. This is a significant improvement over 
+exact partitioning and tree-building, which have $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ costs respectively.
+
+\subsection {Index-Building and Sharded Search}
+In this work, we also introduce the method of \emph{sharding} for $k$-nearest neighbors search.
+With this approach, we aim to leverage the fact that $\rho$-nearest neighbors search is typically 
+much faster than $k$-nearest neighbors search.
+
+We begin by randomly partitioning the dataset into $s$ shards, where $s$ is determined by [TODO: how did 
+we decide we're going to select $s$ for this version of the paper?]. On the first shard, we perform 
+$k$-nearest neighbors search using one of the four algorithms described in Section [ADD SECTION: $k$-Nearest Neighbors Search].
+The result of this search will give us the $k$ closest neighbors to the query in the first shard. We then take the furthest of 
+those $k$ neighbors and use its distance from the query as the radius for $\rho$-nearest neighbors search (see section [ADD SECTION]) 
+on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
+from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
+
+Note that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others,
+since the first shard is the only one for which we perform $k$-nearest neighbors search. However, making this first shard too small
+creates the risk of significantly overestimating the radius, resulting in much slower $\rho$-nearest neighbors search on the remaining shards.
+Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well 
+as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the 
+actual $k$-th closest neighbor, are topics for future work.
 
@@
 
-[TODO: reorder so rho nn is first, then sieves, then Greedy Sieve]
-In this section, we present four novel algorithms for $k$-nearest neighbors search: Greedy Search, $k$-NN by Repeated $\rho$-NN, Sieve Search I, and Sieve Search II. 
+In this section, we present four novel algorithms for $k$-nearest neighbors search: $k$-NN by Repeated $\rho$-NN, Sieve Search, and Sieve Search with Separate Centers, 
+and  Greedy Sieve. 
 We also have an implementation of linear search. We use a preliminary auto-tuning function to predict the variant which will perform 
@@
 
-[TODO: tighten up or omit those opening paragraphs which roughly explain the algorithm]
-
-[TODO: call queues of points H]
-\subsubsection{Greedy Search}
-\label{subsubsec:methods:knn-search:greedy-search}
-Greedy Search repeatedly partitions the cluster with the smallest $\delta_{min}$ into its children 
-until the cluster with the smallest $\delta_{min}$ is a leaf. We then remove it and add its points to a fixed size priority queue of points. 
-Search terminates when the furthest hit in the queue of points has $\delta$ greater than the $\delta_{min}$ of the closest remaining cluster.
-
-Formally, let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
-Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.
-
-[TODO: replace with pop and push language and start with initial pop]
-While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
-than $\delta_{min}$ of the top priority element in $Q$, do the following:
-\begin{enumerate}
-\item While the top priority element $q$ in $Q$ is not a leaf, replace $q$ with its children.
-\item Remove $q$ from $Q$ and add all of its points to $P$. 
-\item Remove points from $P$ until $P$ has $k$ points. 
-\end{enumerate}
-After this process, the points left in $P$ are the $k$ nearest neighbors of $q$.
-
-[TODO: fix to use priority queue language]
-
-\begin{algorithm} 
-\caption{GreedySearch(\emph{tree, query, k})} 
-\label{alg:greedy_search} 
-\begin{algorithmic}[3]
-    \STATE $Q \leftarrow$ priority queue
-    \STATE $Q.push(tree.root)$
-    \STATE $P \leftarrow$ priority queue
-    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
-        \WHILE{$\neg Q.min.isLeaf$}
-            \STATE $[l, r] \leftarrow Q.extractMin().children$
-            \STATE $Q.push([l, r])$
-        \ENDWHILE
-        \STATE $leaf \leftarrow Q.extractMin()$
-        \STATE $P.push(leaf)$
-        \WHILE{$|P| > k$}
-            \STATE $P.extractMax()$
-        \ENDWHILE
-    \ENDWHILE
-    \STATE Return $P$
-\end{algorithmic}
-\end{algorithm}
+In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far.
 
@@
 In this algorithm, we perform
-$\rho$-nearest neighbors search at [TODO: add radius], repeatedly increasing the search radius until $k$ neighbors
+$\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
+the cardinality of the dataset, repeatedly increasing the search radius until $k$ neighbors
 have been found.
 
-[TODO: change P to H]
-Let $P$ be the set of nearest neighbors found thus far.
-Search starts with a radius $m$ equal to the radius of the cluster tree divided by
+Let $H$ be the set of nearest neighbors found thus far.
+Search starts with a radius $m$ equal to the radius of the root cluster divided by
 the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
@@
 2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
-until $|P| > 0$.
+until $|H| > 0$.
 
-[TODO: Add number next to all equations and fix factor so that it is correct (k should be raised too)]
-Once $|P| > 0$, we continue to perform $\rho$-NN search, but instead of 
+Once $|H| > 0$, we continue to perform $\rho$-NN search, but instead of 
 increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
@@
 we increase the radius by a factor of 
-$$\text{min}\left(2, \frac{k}{|P|^{\frac{1}{\mu}}}\right)$$
-where $\mu$ is the mean local fractal dimension (abbreviated lfd in algorithm below) of the clusters containing the points found so far.
-Intuitively, the factor by which we increase the radius should be \emph{inversely} proportional to the number of points found so far. 
-Additionally, when the local fractal dimension near the query and found neighbors is low, this suggests that the data 
-are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} proportional to the 
-local fractal dimension. Since the local fractal dimension is memoized at clustering time, this poses no additional cost at search time.
-Once $|P| >= k$, we sort the points in $P$ by $\delta$ and return the $k$ closest points.
+\begin{equation} \text{argmin}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
+where $\mu$ is the local fractal dimension (abbreviated lfd in algorithm below) in the region around the query point.
+Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
+Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data 
+are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} related to the 
+local fractal dimension. Once $|H| >= k$, we return the $k$ closest points.
 
@@
     \begin{algorithmic}[4] % enter the algorithmic environment
-        \STATE $P \leftarrow$ $\emptyset$
-        \STATE $m \leftarrow$ $\frac{tree.radius}{tree.cardinality}$
-        \WHILE {$|P| = 0$}
-            \STATE $P.push(\rho$-NN$(tree, query, m)$)
+        \STATE $H \leftarrow$ $\emptyset$
+        \STATE $m \leftarrow$ $\frac{root.radius}{root.cardinality}$
+        \WHILE {$|H| = 0$}
+            \STATE $H.push(\rho$-NN$(tree, query, m)$)
             \STATE $m \leftarrow 2m$
         \ENDWHILE
-        \WHILE {$|P| < k$}
-            \STATE $clusters \leftarrow \{ C: \exists p \in P \land p \in C \}$
+        \WHILE {$|H| < k$}
+            \STATE $clusters \leftarrow \{ C: \exists p \in H \land p \in C \}$
             \STATE $\mu \leftarrow \frac{1}{|clusters|} \sum_{C \in clusters} C.lfd$
-            \STATE $m \leftarrow \text{min}\left(2, \frac{k}{|P|^{\frac{1}{\mu}}}\right)$
-            \STATE $P.push(\rho$-NN$(tree, query, m)$)
+            \STATE $m \leftarrow \text{min}\left(2, \frac{k}{|H|^{\frac{1}{\mu}}}\right)$
+            \STATE $H.push(\rho$-NN$(tree, query, m)$)
         \ENDWHILE
-        \STATE $P.sort()$
-        \STATE Return $P[0..k]$
+        \STATE $H.sort()$
+        \STATE Return $H[0], H[1], \cdots H[k-1]$
     \end{algorithmic}
@@
 
-[TODO: change sieve II to sieev with separate centers]
-\subsubsection{Sieve Search I}
+\subsubsection{Sieve Search}
 \label{subsubsec:methods:knn-search:sieve}
-[TODO: make writeup here tighter and more consistent with what's in the algorithm; talk 
-about enum variants here]
-With Sieve Search I, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
-can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.
-
-We start by letting $Q$ be a vector initialized with the root cluster. This vector will maintain the 
+With Sieve Search, we begin by letting $Q$ be a list initialized with the root cluster. This list will maintain the 
 clusters and points which are still in contention for being one of the nearest neighbors.
-While $Q$ contains at least one cluster (i.e., $Q$ is not a vector of only points), we repeat the process described in 
+While $Q$ contains at least one cluster (i.e., $Q$ is not a list of only points), we repeat the process described in 
 the following paragraphs.
@@
 nearest neighbor. Since this algorithm will also elements on the correct side of 
-the index of the $k$th nearest neighbor, the first $k$ elements in $Q$ a are the $k$-nearest neighbors.
+the index of the $k$th nearest neighbor, the first $k$ elements in $Q$ are the $k$-nearest neighbors.
 
@@
                 \IF {$q.\delta_{min} > \tau$}
-                    \STATE $Q.remove(q)$
+                    \STATE $Q.pop(q)$
                 \ENDIF
@@
                 \ENDIF
-                \STATE $Q.remove(q)$
+                \STATE $Q.pop(q)$
             \ENDFOR 
@@
 
-\subsubsection{Sieve Search II}
+\subsubsection{Sieve Search with Separate Centers}
 \label{subsubsec:methods:knn-search:sieve2}
-Sieve Search II is exactly the same as Sieve Search I, but with the following modification: clusters 
+Sieve Search with Separate Centers is the same as Sieve Search, but with the following modification: clusters 
 in $Q$ are represented twice-- once as their center and once as the rest of their points. 
@@
 
+\subsubsection{Greedy Sieve}
+\label{subsubsec:methods:knn-search:greedy-search}
+
+With Greedy Sieve, we first let $Q$ be a min queue of clusters by non-increasing $\delta_{min}$. We initialize $Q$ with the root cluster.
+Let $H$ be a fixed-size ($k$) max queue of points by $\delta$. This queue starts empty.
+
+While $H$ has fewer than $k$ clusters or $\delta$ of the top priority element in $H$ is greater 
+than $\delta_{min}$ of the top priority element in $Q$, do the following:
+\begin{enumerate}
+\item While the top priority element $q$ in $Q$ is not a leaf, pop $q$ from the queue and push its children.
+\item For each point $p \in q$, push $p$ to $H$. 
+\item Pop points from $H$ until $H$ has $k$ points. 
+\end{enumerate}
+After this process, the points left in $H$ are the $k$ nearest neighbors of $q$.
+
+\begin{algorithm} 
+\caption{GreedySearch(\emph{tree, query, k})} 
+\label{alg:greedy_search} 
+\begin{algorithmic}[3]
+    \STATE $Q \leftarrow$ priority queue
+    \STATE $Q.push(tree.root)$
+    \STATE $H \leftarrow$ priority queue
+    \WHILE{$|H| < k \lor H.max.\delta > Q.min.\delta_{min}$}
+        \WHILE{$\neg Q.min.isLeaf$}
+            \STATE $[l, r] \leftarrow Q.pop().children$
+            \STATE $Q.push([l, r])$
+        \ENDWHILE
+        \STATE $leaf \leftarrow Q.pop()$
+        \STATE $H.push(leaf)$
+        \WHILE{$|H| > k$}
+            \STATE $H.pop()$
+        \ENDWHILE
+    \ENDWHILE
+    \STATE Return $H$
+\end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Complexity of $k$-NN Search}
 \label{paragraph:methods:knn-complexity}
-Add asymptotic complexity of various $k$-NN algorithms here.
+To establish the asymptotic complexity of $k$-NN search, we adopt 
+the terminology used to describe the asymptotic complexity of $\rho$-NN search 
+in \cite{yu2015entropy}
+and address \emph{coarse search} and \emph{fine search} separately. 
+Coarse search refers to the process of identifying clusters
+which are likely to contain one of the $k$ nearest neighbors. Fine search refers to the process
+of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
+
+Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 
+as is the case in $\rho$-NN search. Still, this distinction is useful in analyzing 
+the asymptotic complexity of $k$-NN search. In particular, we can think of the complexity of 
+coarse search as an estimate of how many clusters we examine during search, and 
+the complexity of fine search as an estimate of how many clusters we must examine all the points 
+from during search, even if we do not actually perform these as two distinct phases.
+
+Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
+\cite{yu2015entropy}: if call the number of points within a given 
+radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
+are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
+we assume that for any point $p$, query $q$, and radius 
+$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}
+
+For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
+if this still holds for knn].
+
+To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
+number of clusters we must examine all the points from during search. With $\rho$-NN 
+search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
+where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
+as in \cite{yu2015entropy}. With $k$-NN search, however, 
+the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
+is query-dependent. Thus, for $k$-NN search, we  adjust the definition of 
+$F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
+where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.
+
+By the triangle inequality, we have that $F \subseteq B(q, r' + 2r_c)$, 
+where $B(q, r' + r_c)$ is the ball of radius $r' + r_c$ centered at $q$.
+Hence, $|F| \leq |B(q, r' + 2r_c)|$.
+
+Now we have that \begin{align*} \mathbb{E}_q[|B(q, r' + 2r_c)|] &\leq \gamma|B(p, r' + 2r_c)| \\ %because of the assumption that density at different locations on the manifold differ by only a constant factor%
+    &\leq \gamma|B(p, r')|\left(\frac{r' + 2r_c}{r'}\right)^d \label{5}\\ %because d = lfd tells us how the number of additional hits reached scales with a change in radius, and our ratio between radii is (r' + 2r_c)/r'%
+    &\leq \gamma^2 \mathbb{E}[|B(q, r')|]\left(\frac{r'+2r_c}{r'}\right)^d \\ %using density bound again to go from a statement about an arbitrary point back to a statement about the query%
+    &\leq \gamma^2 k\left(\frac{r'+ 2r_c}{r'}\right)^d  \\   %using the fact that for k-nn, the number points in the query ball is k%
+\end{align*}
+
+It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
+points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
+average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
+We rely on this fact in our estimation of $r'$.
+
+We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
+$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
+and the cardinality of the ball centered at the root cluster's center with half its radius). While, ordinarily we compute local fractal dimension 
+by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
+a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.
+
+If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
+\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}
+
+Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
+\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

### 2023-08-31 sections/3-datasets.tex
getting there

@@
 
-Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history.
-
-TODO: Add details about why this dataset is a good candidate for CAKES.
-TODO: Talk about distance functions used with this dataset: euclidean, cosine, Wasserstein-1d
+Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history. This dataset 
+is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure [TODO: ADD FIGURE NUMBER] in Section~\ref{section:methods},
+this dataset is a good candidate for CAKES because it exhibits low fractal dimension.
 
@@
 
-Like APOGEE, MaNGA helps astronomers understand the history of galaxies throughout their entire lifespan.
+Like APOGEE, MaNGA helps astronomers understand the history of galaxies throughout their entire lifespan. This 
+dataset is typically used with Euclidean (L2) distance, cosine distance, and 2d-Wasserstein distance.
 
 TODO: Add details about why this dataset is a good candidate for CAKES.
-TODO: Talk about distance functions used with this dataset: euclidean, cosine, Wasserstein-2d
 
@@
 
-TODO: Which Silva dataset exactly are we using and what is its cardinality and dimensionality?
+Silva 18S contains ribosomal DNA sequences of approximately 2.25 
+million genomes with aligned length of 50,000 letters.
+
+These sequences help biologists better understand the evolutionary relationships between organisms. 
+This dataset is typically used with Hamming and Levenshtein distances.
 
 TODO: Add details about why this dataset is a good candidate for CAKES.
-TODO: Talk about distance functions used with this dataset: hamming, levenshtein
 

### 2023-08-31 sections/2-methods.tex
a lot of citations and other stuff

@@
 The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
-The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}.
+The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}~\cite{fefferman2016testing}.
 
@@
 We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
-similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with the following 
+similar recursive procedure as outlined in CHESS, but with the following 
 improvements: better selection of poles for partitioning and depth-first dataset reordering 
-(see Section ADD SECTION: Dataset Reordering).
+(see Section ~\ref{subsubsec:methods:dataset-depth-first-reordering}).
 
@@
 \subsubsection {Dataset Depth-First Reordering}
+\label{subsubsec:methods:dataset-depth-first-reordering}
 
@@
 \label{subsubsec:methods:guaranteed-decrease-in-cluster-radii}
-While it may be tempting to assume that cluster radii decrease with each application of Cluster Partition (TODO: ADD SECTION), unfortunately, this assumption is incorrect. 
+While it may be tempting to assume that cluster radii decrease with each application of Cluster Partition (refer to Algorithm \ref{alg:partition}), unfortunately, this assumption is incorrect. 
 Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii; in particular, we can guarantee that cluster radii will stop increasing after at most 
-$d$ partitions, where $d$ is the dimensionality of the dataset.  In the remainder of this subsection, we HJFHFJ.
+$d$ partitions, where $d$ is the dimensionality of the dataset.  In the remainder of this subsection, we prove this guarantee.
 
-We can describe this a $d$-dimensional distribution by choosing some set of $d$ mutually orthogonal axes.
+We can describe a $d$-dimensional distribution by choosing some set of $d$ mutually orthogonal axes.
 Let $2R$ be the maximum pairwise distance among the instances in the dataset. 
@@
 Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
-In the worst case, e.g. with a distribution that fills the $d$-sphere, our axes will be such that $2R$ is the maximum pairwise distance along every axis. 
+In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance along every axis. 
 Such a distribution would also produce a balanced clustering.
@@
 which to split the cluster into two children. 
-After one application of Cluster Partition, the maximum pairwise distance along that axis will be
+After one application of Algorithm \ref{alg:partition}, the maximum pairwise distance along that axis will be
 bounded above by $R$. 
 The next recursive Partition will select another of the $d$ axes. 
-Thus, after at most $d$ applications of Cluster Partition, the
+Thus, after at most $d$ applications of Algorithm \ref{alg:partition}, the
 maximum pairwise distance along each axis will be bounded above by $R$. 
-The overall (i.e., not restricted to be along one axis) maximum pairwise distance, 
+The overall (i.e., not restricted to be along one axis) maximum pairwise distance 
 will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes.
@@
 Thus, starting with a cluster $C$ of radius $R$, after at most $d$ Partitions, the descendants of $C$ will each have radius
-bounded above by $\frac{r}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at 
-most $d$ applications of Cluster Partition.
+bounded above by $\frac{R}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at 
+most $d$ applications of Algorithm \ref{alg:partition}.
 
@@
 in different regions of the manifold and the low dimensional "shape" of the manifold. Further, the cluster radii decrease by a factor much larger than 
-$\frac{\sqrt{2}}{2}$ in practice, and often this decrease starts occurring after only one or two applications of Cluster Partition, rather than the 
-upper bound of $d$ applications.
+$\frac{\sqrt{2}}{2}$ in practice, and the upper bound of $d$ applications of Algorithm \ref{alg:partition} is seldom realized.
 
@@
 we decide we're going to select $s$ for this version of the paper?]. On the first shard, we perform 
-$k$-nearest neighbors search using one of the four algorithms described in Section [ADD SECTION: $k$-Nearest Neighbors Search].
+$k$-nearest neighbors search using one of the four algorithms described in Section ~\ref{subsec:methods:knn-search}.
 The result of this search will give us the $k$ closest neighbors to the query in the first shard. We then take the furthest of 
-those $k$ neighbors and use its distance from the query as the radius for $\rho$-nearest neighbors search (see section [ADD SECTION]) 
+those $k$ neighbors and use its distance from the query as the radius for $\rho$-nearest neighbors search (see Section ~\ref{subsec:methods:rnn-search}) 
 on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
@@
 \label{subsubsec:methods:knn-search:repeated-rnn}
-In this algorithm, we perform
-$\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
+
+In this algorithm, we perform $\rho$-nearest neighbors search at a radius equal to the radius of the root cluster divided by
 the cardinality of the dataset, repeatedly increasing the search radius until $k$ neighbors
@@
 To find this element, 
-we use a process similar to the Partition algorithm used in Quicksort (CITE: original Quicksort paper and reference wherever it is 
-in our paper), adjusted to account for the varying cardinalities of elements in $Q$. Our modified algorithm finds the smallest index $i$ such 
+we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of elements in $Q$.
+Our modified algorithm finds the smallest index $i$ such
 that all elements in $Q$ with $\delta_{max}$ closer to or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater
@@
 \label{paragraph:methods:knn-complexity}
+
 To establish the asymptotic complexity of $k$-NN search, we adopt 
@@
 $d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
-and the cardinality of the ball centered at the root cluster's center with half its radius). While, ordinarily we compute local fractal dimension 
+and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
+er's center with half its radius). While, ordinarily we compute local fractal dimension 
 by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 

### 2023-08-31 sections/1-introduction.tex

@@
 This is especially true in the field of astronomy. 
-The Sloan Digital Sky Survey (SDSS), for example, in its APOGEE dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions~\cite{alam2015eleventh}. 
+The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
 Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$. [add sentence about big ANN competition]
 
-Biology has also seen an influx of large, high-dimensional datasets. Silva 18S contains ribosomal DNA sequences of approximately 2.25 
+Biology has also seen an influx of large, high-dimensional datasets. Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 
 million genomes with aligned length of 50,000 letters.
 
-% (TODO: Do we need to talk about NR99; it seems like more of the same but 
-% with removed redundancy; not sure if that is helpful to us...) 
-% No we don't... I'll explain why if you want.
+More recently, Big-ANN launched NeurIPS'23: Billion Scale Approximate Nearest Neighbor Search Challenge, 
+which aims to encourage the development of indexing data structures and search algorithms for practical variants of Approximate Nearest Neighbor (ANN) or Vector search.
+This challenge tests submissions' performances with the YFCC 100M~\cite{johnson2019billion} dataset, the Yandex Text-to-Image 10M~\cite{diskann-github}, cross-modal dataset, 
+the MSMARCO passage retrieval dataset~\cite{Bruch2023AnAA}, and a 10M slice of the MS Turing data set~\cite{Singh2021FreshDiskANNAF}. These datasets 
+are 192, 200, 30,000, and 100 dimensional respectively.
 
-In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{moore1965cramming}, indicating 
+In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating 
 that computer architecture improvements will not "catch up" to our computational needs in the near future. As a result, analysis of these large datasets 
@@
 Similarity search enables a variety of applications, including recommendation and classification systems. 
-Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
+Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, similarity search can become a useful tool for 
 understanding and verifying the validity of model outputs. 
@@
 
-Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
-clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
+Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, 
+clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
 Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in big datasets.

### 2023-08-31 sections/5-discussion.tex
forgot to push a bunch of stuff so here we go

--- 2021-03-09 sections/4-discussion.tex
@@
 
-GPU-acceleration for the more computationally intensive distance functions like wasserstein.
+For future work, we plan to investigate compression by representing differences at each level of the
+binary tree, particularly for string or genomic data, where all differences are discrete.
+When we descend a level in the cluster tree, it takes fewer bits to encode each point in terms of the center. 
+However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
+This presents the following tradeoff: \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
+In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
 
-Live-updates as new points are added to the dataset.
-For example, build tree with a subsample of the full data and then add the rest of the instances to simulate live-updates.
+We also plan to use CAKES with anomaly detection. Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses 
+a variety of methods in an ensemble to determine an anomaly score for a point. We would like to explore the effectiveness of 
+distance to the $k$-th nearest neighbor as an additional method in our ensemble.
 
-More distance functions. For example, maximal-common-subgraph for molecular structures.
+Additionally, we would like to implement additional distance functions to improve the genericity of CAKES. 
+For example, we would like to implement Tanimoto Distance using maximal-common-subgraph for molecular structures. Ultimately, 
+we would also like to implement Wasserstein distance for dimensions higher than 2. In order to make this possible, 
+we plan to implement GPU-acceleration for the more computationally-intensive distance functions like Wasserstein.
 
-TODO: Provide link to source code on github.
+Finally, we would like to add the ability to view live-updates as new points are added to a dataset. For example, 
+we could build the cluster tree with a subsample of the full data set and then add the rest of the instances to 
+simulate live updates.

### 2023-08-31 sections/2-methods.tex
changes to rnn

@@
 
-[TODO: delete this and talk about overlapping children improvement instead]
-We conduct $\rho$-nearest neighbors search as described in CHESS~\cite{ishaq2019clustered}, but 
-with some implementation improvements including identification of candidate clusters which are leaves and 
-conducting exhaustive search over those leaves.
+We conduct $\rho$-nearest neighbors search as described in \cite{ishaq2019clustered}, but 
+with the following improvement: when a cluster overlaps with the query ball, instead of  
+always proceeding with both of its children, we only proceed with those children which 
+also overlap with the query ball.
 
@@
         \STATE $results \leftarrow \emptyset$
-        \IF{$clusters.left$}
+        \IF{$clusters.left \land clusters.left \cap B_X(query, r) \neq \emptyset$}
             \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
@@
         \ENDIF
-        \IF{$clusters.right$}
+        \IF{$clusters.right \land clusters.right \cap B_X(query, r) \neq \emptyset$}
             \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
@@
                 \IF{$p.\delta \leq r$}
-                    \STATE $results \leftarrow r$
+                    \STATE $results.push((p, r))$
                 \ENDIF
@@
 
-The asymptotic complexity of $\rho$-nearest neighbors is the same as in CHESS~\cite{ishaq2019clustered}.
+The asymptotic complexity of $\rho$-nearest neighbors is the same as in ~\cite{ishaq2019clustered}.
 
@@
 
-In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far.
+In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and
+$Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the nearest neighbors.
 
@@
 we increase the radius by a factor of 
-\begin{equation} \text{argmin}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
+\begin{equation} \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
 where $\mu$ is the local fractal dimension (abbreviated lfd in algorithm below) in the region around the query point.
@@
     \begin{algorithmic}[4] % enter the algorithmic environment
-        \STATE $H \leftarrow$ $\emptyset$
-        \STATE $m \leftarrow$ $\frac{root.radius}{root.cardinality}$
+        \STATE $H \Leftarrow$ $\emptyset$
+        \STATE $m \Leftarrow$ $\frac{root.radius}{root.cardinality}$
         \WHILE {$|H| = 0$}
             \STATE $H.push(\rho$-NN$(tree, query, m)$)
-            \STATE $m \leftarrow 2m$
+            \STATE $m \Leftarrow 2m$
         \ENDWHILE
         \WHILE {$|H| < k$}
-            \STATE $clusters \leftarrow \{ C: \exists p \in H \land p \in C \}$
-            \STATE $\mu \leftarrow \frac{1}{|clusters|} \sum_{C \in clusters} C.lfd$
-            \STATE $m \leftarrow \text{min}\left(2, \frac{k}{|H|^{\frac{1}{\mu}}}\right)$
+            \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
+            \STATE $\mu \leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
+            \STATE $m \leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
             \STATE $H.push(\rho$-NN$(tree, query, m)$)

### 2023-09-01 sections/2-methods.tex
switch to new figure
still losing it

@@
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in X \setminus S_q$, $f(q, x) < \max\{f(y, q): y \in S_q \}$; that is,
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in X \setminus S_q$, $f(q, x) \leq \max\{f(y, q): y \in S_q \}$; that is,
 for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
@@
     \REQUIRE $cluster$
-    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
-    \STATE $seeds \leftarrow m$ random points from $cluster.points$
-    \STATE $c \leftarrow$ geometric median of $seeds$
-    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
-    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
-    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
-    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
+    \STATE $m \Leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
+    \STATE $seeds \Leftarrow m$ random points from $cluster.points$
+    \STATE $c \Leftarrow$ geometric median of $seeds$
+    \STATE $l \Leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
+    \STATE $r \Leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
+    \STATE $left \Leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
+    \STATE $right \Leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
     \IF{$|left| > 1$}
@@
     \centering
-    \includegraphics[width=2.5in]{images/geometry/sketchometry_final.png}
+    \includegraphics[width=2.5in]{images/geometry/geometry.pdf}
     \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk}
@@
         \REQUIRE $clusters \neq \emptyset$
-        \STATE $results \leftarrow \emptyset$
+        \STATE $results \Leftarrow \emptyset$
         \IF{$clusters.left \land clusters.left \cap B_X(query, r) \neq \emptyset$}
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{Repeated$\rho$-NN(\emph{tree, query, k})} % give the algorithm a caption
+    \caption{Repeated$\rho$-NN(\emph{root, query, k})} % give the algorithm a caption
     \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
@@
         \STATE $H \Leftarrow$ $\emptyset$
-        \STATE $m \Leftarrow$ $\frac{root.radius}{root.cardinality}$
+        \STATE $m \Leftarrow$ $\frac{root.radius}{|root|}$
         \WHILE {$|H| = 0$}
-            \STATE $H.push(\rho$-NN$(tree, query, m)$)
+            \STATE $H.push(\rho$-NN$(root, query, m)$)
             \STATE $m \Leftarrow 2m$
@@
             \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
-            \STATE $\mu \leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
-            \STATE $m \leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
-            \STATE $H.push(\rho$-NN$(tree, query, m)$)
+            \STATE $\mu \Leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
+            \STATE $m \Leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
+            \STATE $H.push(\rho$-NN$(root, query, m)$)
         \ENDWHILE
@@
             \STATE $Q.swap(pivot, r)$
-            \STATE $a \leftarrow l$
-            \STATE $b \leftarrow l$
+            \STATE $a \Leftarrow l$
+            \STATE $b \Leftarrow l$
             \WHILE {$b < r$}
@@
                     \STATE $Q.swap(a, b)$
-                    \STATE $a \leftarrow a + 1$
+                    \STATE $a \Leftarrow a + 1$
                 \ENDIF
-                \STATE $b \leftarrow b + 1$
+                \STATE $b \Leftarrow b + 1$
             \ENDWHILE
             \STATE $Q.swap(a, r)$
-            \STATE $p \leftarrow r$
-            \STATE $g \leftarrow \sum_{i=0}^{p} Q[i].cardinality$
+            \STATE $p \Leftarrow r$
+            \STATE $g \Leftarrow \sum_{i=0}^{p} |Q[i]|$
             \IF {$g = k$}
@@
             \ELSE 
-                \IF {$p > 0 \land g > k + Q[p - 1].cardinality$}
+                \IF {$p > 0 \land g > k + |Q[p - 1]|$}
                     \STATE $QuicksortPartition(Q, k, l, p - 1)$
-                \ELSIF {$p > 0 \land g = k + Q[p - 1].cardinality$}
+                \ELSIF {$p > 0 \land g = k + |Q[p - 1]|$}
                     \STATE Return $p - 1$
@@
         \REQUIRE $tree$, $query$, $k$
-        \STATE $Q \leftarrow$ [$tree.root$]
+        \STATE $Q \Leftarrow$ [$tree.root$]
         \WHILE{$|Q| > k$}
-            \STATE $i \leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
-            \STATE $\tau \leftarrow Q[i].\delta_{max}$
+            \STATE $i \Leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
+            \STATE $\tau \Leftarrow Q[i].\delta_{max}$
             \FOR {$q \in Q$}
@@
             \FOR {$q \in Q$}
-                \IF {$q.isLeaf \lor |q.cardinality| \leq k$}
+                \IF {$q.isLeaf \lor |q| \leq k$}
                     \STATE $Q.push(q.points)$
                 \ELSE
-                    \STATE $[l, r] \leftarrow q.children$
+                    \STATE $[l, r] \Leftarrow q.children$
                     \STATE $Q.push([l, r])$   
@@
 \begin{algorithm} 
-\caption{GreedySearch(\emph{tree, query, k})} 
+\caption{GreedySearch(\emph{root, query, k})} 
 \label{alg:greedy_search} 
 \begin{algorithmic}[3]
-    \STATE $Q \leftarrow$ priority queue
-    \STATE $Q.push(tree.root)$
-    \STATE $H \leftarrow$ priority queue
-    \WHILE{$|H| < k \lor H.max.\delta > Q.min.\delta_{min}$}
-        \WHILE{$\neg Q.min.isLeaf$}
-            \STATE $[l, r] \leftarrow Q.pop().children$
+    \STATE $Q \Leftarrow$ priority queue
+    \STATE $Q.push(root)$
+    \STATE $H \Leftarrow$ priority queue
+    \WHILE{$|H| < k \lor H.pop.\delta > Q.pop.\delta_{min}$}
+        \WHILE{$\neg Q.pop.isLeaf$}
+            \STATE $[l, r] \Leftarrow Q.pop().children$
             \STATE $Q.push([l, r])$
         \ENDWHILE
-        \STATE $leaf \leftarrow Q.pop()$
+        \STATE $leaf \Leftarrow Q.pop()$
         \STATE $H.push(leaf)$
@@
 
-To establish the asymptotic complexity of $k$-NN search, we adopt 
-the terminology used to describe the asymptotic complexity of $\rho$-NN search 
-in \cite{yu2015entropy}
-and address \emph{coarse search} and \emph{fine search} separately. 
+To bound the complexity of $k$-NN search, we group all of the "Sieve" algorithms (Sieve, 
+Sieve with Separate Centers, and Greedy Sieve) together, and handle $k$-NN by Repeated $\rho$-NN separately.
+
+We start by considering the complexity of $k$-NN by Repeated $\rho$-NN, as this  
+is a natural extension of the complexity bounds for $\rho$-NN search described in 
+\cite{ishaq2019clustered}. For this algorithm, we adopt the terminology used in 
+\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
 Coarse search refers to the process of identifying clusters
-which are likely to contain one of the $k$ nearest neighbors. Fine search refers to the process
+which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
+Fine search refers to the process
 of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
 
-Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 
-as is the case in $\rho$-NN search. Still, this distinction is useful in analyzing 
-the asymptotic complexity of $k$-NN search. In particular, we can think of the complexity of 
-coarse search as an estimate of how many clusters we examine during search, and 
-the complexity of fine search as an estimate of how many clusters we must examine all the points 
-from during search, even if we do not actually perform these as two distinct phases.
+% Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 
+% as is the case in $\rho$-NN search. Still, this distinction is useful in analyzing 
+% the asymptotic complexity of $k$-NN search. In particular, we can think of the complexity of 
+% coarse search as an estimate of how many clusters we examine during search, and 
+% the complexity of fine search as an estimate of how many clusters we must examine all the points 
+% from during search, even if we do not actually perform these as two distinct phases.
 
-Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
-\cite{yu2015entropy}: if call the number of points within a given 
-radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
-are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
-we assume that for any point $p$, query $q$, and radius 
-$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}
+% Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
+% \cite{yu2015entropy}: if call the number of points within a given 
+% radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
+% are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
+% we assume that for any point $p$, query $q$, and radius 
+% $\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}
 
-For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
-if this still holds for knn].
+% For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
+% if this still holds for knn].
 
-To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
-number of clusters we must examine all the points from during search. With $\rho$-NN 
-search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
-where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
-as in \cite{yu2015entropy}. With $k$-NN search, however, 
-the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
-is query-dependent. Thus, for $k$-NN search, we  adjust the definition of 
-$F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
-where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.
+% To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
+% number of clusters we must examine all the points from during search. With $\rho$-NN 
+% search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
+% where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
+% as in \cite{yu2015entropy}. With $k$-NN search, however, 
+% the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
+% is query-dependent. Thus, for $k$-NN search, we  adjust the definition of 
+% $F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
+% where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.
 
-By the triangle inequality, we have that $F \subseteq B(q, r' + 2r_c)$, 
-where $B(q, r' + r_c)$ is the ball of radius $r' + r_c$ centered at $q$.
-Hence, $|F| \leq |B(q, r' + 2r_c)|$.
+% By the triangle inequality, we have that $F \subseteq B(q, r' + 2r_c)$, 
+% where $B(q, r' + r_c)$ is the ball of radius $r' + r_c$ centered at $q$.
+% Hence, $|F| \leq |B(q, r' + 2r_c)|$.
 
-Now we have that \begin{align*} \mathbb{E}_q[|B(q, r' + 2r_c)|] &\leq \gamma|B(p, r' + 2r_c)| \\ %because of the assumption that density at different locations on the manifold differ by only a constant factor%
-    &\leq \gamma|B(p, r')|\left(\frac{r' + 2r_c}{r'}\right)^d \label{5}\\ %because d = lfd tells us how the number of additional hits reached scales with a change in radius, and our ratio between radii is (r' + 2r_c)/r'%
-    &\leq \gamma^2 \mathbb{E}[|B(q, r')|]\left(\frac{r'+2r_c}{r'}\right)^d \\ %using density bound again to go from a statement about an arbitrary point back to a statement about the query%
-    &\leq \gamma^2 k\left(\frac{r'+ 2r_c}{r'}\right)^d  \\   %using the fact that for k-nn, the number points in the query ball is k%
-\end{align*}
+% Now we have that \begin{align*} \mathbb{E}_q[|B(q, r' + 2r_c)|] &\leq \gamma|B(p, r' + 2r_c)| \\ %because of the assumption that density at different locations on the manifold differ by only a constant factor%
+%     &\leq \gamma|B(p, r')|\left(\frac{r' + 2r_c}{r'}\right)^d \label{5}\\ %because d = lfd tells us how the number of additional hits reached scales with a change in radius, and our ratio between radii is (r' + 2r_c)/r'%
+%     &\leq \gamma^2 \mathbb{E}[|B(q, r')|]\left(\frac{r'+2r_c}{r'}\right)^d \\ %using density bound again to go from a statement about an arbitrary point back to a statement about the query%
+%     &\leq \gamma^2 k\left(\frac{r'+ 2r_c}{r'}\right)^d  \\   %using the fact that for k-nn, the number points in the query ball is k%
+% \end{align*}
 
-It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
-points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
-average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
-We rely on this fact in our estimation of $r'$.
+% It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
+% points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
+% average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
+% We rely on this fact in our estimation of $r'$.
 
-We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
-$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
-and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
-er's center with half its radius). While, ordinarily we compute local fractal dimension 
-by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
-a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.
+% We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
+% $d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
+% and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
+% er's center with half its radius). While, ordinarily we compute local fractal dimension 
+% by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
+% a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.
 
-If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
-\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}
+% If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
+% \begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}
 
-Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
-\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}
+% Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
+% \begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

### 2023-09-01 sections/3-datasets.tex
still losing it

@@
-\section{Datasets and Distance Functions}
+\section{Datasets}
 \label{sec:datasets-and-distance-functions}
-
-Provide details on the datasets we use for benchmarks.
-
-Add interesting t-SNE or UMAP embeddings of datasets.
-For example~\ref{fig:discussion:umap-annthyroid-euclidean}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=2.5in]{images/umaps/annthyroid-euclidean-umap2d.png}
-    \caption{UMAP embedding of Annthyroid data with the Euclidean distance function.}
-    \label{fig:discussion:umap-annthyroid-euclidean}
-\end{figure}
 
@@
 Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history. This dataset 
-is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure [TODO: ADD FIGURE NUMBER] in Section~\ref{section:methods},
+is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure 1 in Section 2,
 this dataset is a good candidate for CAKES because it exhibits low fractal dimension.
@@
 
-deep1b: cosine
+We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
+which we list below, along with the distance functions with which they are typically used.
 
-fashion-mnist: euclidean
-
-gist: euclidean
-
-glove: cosine
-
-kosarak: jaccard
-
-mnist: euclidean
-
-nytimes: cosine
-
-sift: euclidean
-
-last.fm: cosine
-
-\dots
+\begin{itemize} 
+    \item DEEP1B (Cosine): ~1 billion 96-dimensional deep descriptors to 
+    \item Fashion-MNIST (Euclidean): 60,000 784-dimensional vector representations of images of clothing items
+    \item GIST (Euclidean): 1 million 960-dimensional image descriptors 
+    \item GloVe (Cosine): ~1.2 million 25-dimensional vector representations of words
+    \item Kosarak (Jaccard): ~75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
+    \item MNIST (Euclidean): 60,000 784-dimensional vector representations of images of handwritten digits
+    \item NYTimes (Cosine): 290,000 256-dimensional vector representations of words
+    \item SIFT (Euclidean): 1 million 128-dimensional image descriptors
+    \item Last.fm (Cosine): ~292, 000 65-dimensional vector representations of songs
+\end{itemize}

### 2023-09-01 sections/4-results.tex
still losing it

@@
 
-Write about results \dots
+Figure [ADD FIGURE NUMBER] is a violin plot which shows the performance of our algorithms
+on [ADD DATASET] under [ADD DISTANCE FUNCTION] distance, using [ADD NUMBER OF SHARDS] shards.
+On the horizontal axis, we have each of the $k$-NN algorithms described in Section 2.4, as well as  
+an additional "algorithm" which we call "Oracle $\rho$-NN" [REPLACE WITH ACTUAL NAME].  With "Oracle $\rho$-NN", we perform $\rho$-NN search using the 
+distance to the $k$-th nearest neighbor as the radius. Of course, as it is impossible to
+to know the distance to the $k$-th nearest neighbor \emph{a priori}, this algorithm 
+is useful only for benchmarking purposes.
+
+The vertical axis for Figure [ADD FIGURE NUMBER] is the time query in seconds on a logarithmic scale. 
+As shown in the key, different colors represent different values of $k$. 
+The plot shows the distribution of query times for each algorithm for each value of $k$. 
+Because of the logarithmic scale of the vertical axis, it is unsurprising that we 
+see longer lower tails than upper tails for many of these algorithms.
 
@@
 
-Provide link to source code on GitHub.
+https://github.com/URI-ABD/clam
 

### 2023-09-01 sections/1-introduction.tex
changes for clarity

@@
 The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
-Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$. [add sentence about big ANN competition]
+Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.
 
 Biology has also seen an influx of large, high-dimensional datasets. Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 
-million genomes with aligned length of 50,000 letters.
+million genomes with aligned length of 50,000 letters; 
+the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences.
 
-More recently, Big-ANN launched NeurIPS'23: Billion Scale Approximate Nearest Neighbor Search Challenge, 
-which aims to encourage the development of indexing data structures and search algorithms for practical variants of Approximate Nearest Neighbor (ANN) or Vector search.
-This challenge tests submissions' performances with the YFCC 100M~\cite{johnson2019billion} dataset, the Yandex Text-to-Image 10M~\cite{diskann-github}, cross-modal dataset, 
+More recently, Big-ANN's NeurIPS'23: Billion Scale Approximate Nearest Neighbor Search Challenge has 
+solicits development of data structures and search algorithms for Approximate Nearest Neighbor (ANN) search
+on four massive datasets: YFCC 100M~\cite{johnson2019billion} dataset, the Yandex Text-to-Image 10M~\cite{diskann-github}, cross-modal dataset, 
 the MSMARCO passage retrieval dataset~\cite{Bruch2023AnAA}, and a 10M slice of the MS Turing data set~\cite{Singh2021FreshDiskANNAF}. These datasets 
-are 192, 200, 30,000, and 100 dimensional respectively.
+are 192, 200, 30,000, and 100 dimensional respectively, and submissions are evaluated with 
+a ten million point sample of each.
 
 In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating 
-that computer architecture improvements will not "catch up" to our computational needs in the near future. As a result, analysis of these large datasets 
+that computer architecture will not "catch up" to our computational needs in the near future. As a result, analysis of these large datasets 
 requires better algorithms.
@@
 
-In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}; 
-naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
+In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
+Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
 prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
-exploit the geometric and topological structure inherent in big datasets. Approximate search is sufficient for some applications, but the 
+exploit the geometric and topological structure inherent in large datasets. Approximate search is sufficient for some applications, but the 
 need for efficient, \emph{exact} search remains.
@@
 Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform
-$k$-nearest neighbors search in addition to $\rho$-nearest neighbors search (also known as ranged nearest neighbors search).
+$k$-nearest neighbors search.
 
-This paper focuses on four novel accelerated $k$-nearest neighbors search algorithms. We explain some 
-implementation improvements to CHESS, as well as improved genericity across distance functions. 
-We also include proof of some geometric guarantees for clusters. Finally, we provide a comparison of 
+This paper focuses on four novel exact $k$-nearest neighbors search algorithms. We also explain some 
+implementation improvements to the clustering and $\rho$ nearest neighbors search algorithms in CHESS, 
+as well as improved genericity across distance functions. Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
+Finally, we provide a comparison of 
 CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} on several datasets 
-from the ann and big ann benchmarks suites.
+from the ANN and Big ANN benchmarks suites.

### 2023-09-01 sections/2-methods.tex
changes for clarity

@@
 
-We define a \emph{distance function} $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
+We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
 given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
@@
 
-Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point in the dataset, 
+Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point 
+and length scale in the dataset, 
 which we define as: 
@@
 The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
-The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold is known as the \emph{manifold hypothesis}~\cite{fefferman2016testing}.
+The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space is known as the \emph{manifold hypothesis}~\cite{fefferman2016testing}.
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=2.5in]{images/lfd_plots/lfd_plot_apogee.png}
-    \caption{Mean fractal dimension of APOGEE clusters as a function of
-    depth when clustered based on L2 norm. Each plot line represents a distinct
-    decile of fractal dimension. Beyond a depth of 56, no clusters are further
-    divided due to the minimum-cluster-cardinality stopping criteria.}
-    \label{fig:methods:lfd_apogee}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=2.5in]{images/lfd_plots/lfd_plot_greengenes.png}
-    \caption{Mean fractal dimension of GreenGenes clusters as a function of
-    depth when clustered based on Hamming distance. Each plot line represents
-    a distinct decile of fractal dimension.}
-    \label{fig:methods:lfd_greengenes}
-\end{figure}
-
-Figure~\ref{fig:methods:lfd_apogee} shows the mean local fractal dimension of the APOGEE dataset at each level of hierarchical clustering, 
-under the L2 (Euclidean) distance metric. Each plot line represents a decile of fractal dimension; note that, other than the most extreme 10\% of clusters,
-virtually all clusters have a local fractal dimension of less than $2 \ll 8,575$. This suggests that APOGEE is a good candidate for use with CAKES.
-
-Figure~\ref{fig:methods:lfd_greengenes} shows the mean local fractal dimension of the GreenGenes dataset at each level of hierarchical clustering,
-under the Hamming distance metric. Each plot line represents a decile of fractal dimension; once again, other than the most extreme 10\% of clusters,
-virtually all clusters have a local fractal dimension of less than $2 \ll 2,250$. This suggests that GreenGenes is a good candidate for use with CAKES.
+Figures 2 and 3 in~\cite{ishaq2019clustered} illustrate the low fractal dimensions of two datasets used with $\rho$-NN search in CHESS. 
+We observe that for both APOGEE (Figure 2 in~\cite{ishaq2019clustered}) and GreenGenes (Figure 3 in~\cite{ishaq2019clustered}), other than the most extreme 10\% of clusters, virtually all clusters have a local fractal dimension of less than $2 \ll 8,575$
+and $2 \ll 2,250$ respectively.
+This suggests that APOGEE and GreenGenes are good datasets for use with entropy-scaling search algorithms like CHESS and CAKES.
 
@@
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in X \setminus S_q$, $f(q, x) \leq \max\{f(y, q): y \in S_q \}$; that is,
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $f(q, x) \leq \max\{f(y, q): y \in S_q \}$; that is,
 for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.

### 2023-09-01 sections/2-methods.tex
cut some stuff out, made some changes for clarity

@@
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $f(q, x) \leq \max\{f(y, q): y \in S_q \}$; that is,
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $f(x, q) \leq \max\{f(y, q): y \in S_q \}$; that is,
 for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
@@
 many instances of the same point (i.e., have cardinality greater than 1 but contain only one unique point). Singletons clearly 
-have zero radius, and so $\delta_{max} = \delta_{min} = \delta$ for these clusters. Hence, we sometimes overload the above 
+have zero radius, and so $\delta_{max}  = \delta = \delta_{min}$ for these clusters. Hence, we sometimes overload the above 
 notation to refer to distance from a query to a point; in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
@@
 those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
-repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
+repeat this procedure until each leaf contains only one datum or some other user-specified stopping criterion 
 is met.
 
+This process is described in Algorithm \ref{alg:partition}.
+
 \begin{algorithm} % enter the algorithm environment
-\caption{Cluster Partition} % give the algorithm a caption
+\caption{ClusterPartition(\emph{C})} % give the algorithm a caption
 \label{alg:partition} % and a label for \ref{} commands later in the document
 \begin{algorithmic}[1] % enter the algorithmic environment
-    \REQUIRE $cluster$
-    \STATE $m \Leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
-    \STATE $seeds \Leftarrow m$ random points from $cluster.points$
+    \STATE $m \Leftarrow \lfloor \sqrt{|C|} \rfloor$
+    \STATE $seeds \Leftarrow m$ random points from $C.points$
     \STATE $c \Leftarrow$ geometric median of $seeds$
-    \STATE $l \Leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
-    \STATE $r \Leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
-    \STATE $left \Leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
-    \STATE $right \Leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
+    \STATE $l \Leftarrow \argmax d(c,x) \ \forall \ x \in C.points$
+    \STATE $r \Leftarrow \argmax d(l,x) \ \forall \ x \in C.points$
+    \STATE $left \Leftarrow \{x | x \in C.points \land d(l,x) \le d(r,x)\}$
+    \STATE $right \Leftarrow \{x | x \in C.points \land d(r,x) < d(l,x)\}$
     \IF{$|left| > 1$}
-        \STATE Partition($left$)
+        \STATE ClusterPartition($left$)
     \ENDIF
     \IF{$|right| > 1$}
-        \STATE Partition($right$)
+        \STATE ClusterPartition($right$)
     \ENDIF
@@
 CAKES also improves upon CHESS by reordering the dataset. In CHESS, each cluster stored a list of indices into the dataset, 
-which we used during search to retrieve a cluster's points from the dataset. 
+which we used during search to retrieve a cluster's points. 
 Though this approach allowed us to retrieve the points in constant 
@@
 This approach, however, introduces an $\mathcal{O}(n)$ time cost whenever we need to find the indices for 
-a non-leaf cluster, since it a traversal of the subtree rooted at that cluster.
+a non-leaf cluster, since it requires a traversal of the subtree rooted at that cluster.
 
@@
 so that points are stored in a depth-first order. Then, within each cluster, we need only 
-store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset. Parent clusters 
-have the same offset as their left child, and right child clusters have an offset equal to their sibling's offset
-plus their siblings's cardinality. With no additional memory cost nor time cost for retrieving a points during search, 
+store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset. Each parent cluster 
+have the same offset as its left child, and each right child has an offset equal to their sibling's offset
+plus their siblings's cardinality. With no additional memory cost nor time cost for retrieving points during search, 
 dataset depth-first reordering offers significant improvements over the approach used in CHESS.
@@
 While it may be tempting to assume that cluster radii decrease with each application of Cluster Partition (refer to Algorithm \ref{alg:partition}), unfortunately, this assumption is incorrect. 
-Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii; in particular, we can guarantee that cluster radii will stop increasing after at most 
-$d$ partitions, where $d$ is the dimensionality of the dataset.  In the remainder of this subsection, we prove this guarantee.
+Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii; in particular, we prove in this section that cluster radii will stop increasing after at most 
+$d$ partitions, where $d$ is the dimensionality of the dataset.
 
-We can describe a $d$-dimensional distribution by choosing some set of $d$ mutually orthogonal axes.
-Let $2R$ be the maximum pairwise distance among the instances in the dataset. 
-We choose the axes such that the two points that are $2R$ apart lie along one of the axes. 
+We can describe a $d$-dimensional distribution of data by choosing some set of $d$ mutually orthogonal axes.
+Let $2R$ be the maximum pairwise distance among the points in the dataset. 
+We choose the axes such that the two points that are $2R$ apart lie \emph{along one of the axes}. 
 Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
@@
 
-Partition will select a maximally distant pair of points to use as poles, i.e., it will choose one of the $d$ axes along
+In this case, Cluster Partition will select a maximally distant pair of points to use as poles, i.e., it will choose one of the $d$ axes along
 which to split the cluster into two children. 
@@
 bounded above by $R$. 
-The next recursive Partition will select another of the $d$ axes. 
+The next recursive Cluster Partition will select another of the $d$ axes. 
 Thus, after at most $d$ applications of Algorithm \ref{alg:partition}, the
 maximum pairwise distance along each axis will be bounded above by $R$. 
-The overall (i.e., not restricted to be along one axis) maximum pairwise distance 
+The overall (i.e., not restricted to one axis) maximum pairwise distance 
 will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes.
@@
 in different regions of the manifold and the low dimensional "shape" of the manifold. Further, the cluster radii decrease by a factor much larger than 
-$\frac{\sqrt{2}}{2}$ in practice, and the upper bound of $d$ applications of Algorithm \ref{alg:partition} is seldom realized.
+$\frac{\sqrt{2}}{2}$ in practice, and the upper bound of $d$ partitions is seldom realized.
 
@@
 \label{subsubsec:methods:clustering:complexity}
-%A lot of this is lifted from the CHESS paper and could probably get cut out. %
 
-To analyze the computational complexity of hierarchical
-clustering, we start by considering the top level of the hierarchy. 
-For a data set with cardinality $n$, we have $n$ data points at this 
-level, from which we randomly choose a sample of $\sqrt{n}$ points. 
-We compute distances between every pair of seeds in this sample, 
-thus making $n$ distance comparisons. Based on these distances 
-we compute the center and left and right poles. Once the poles 
-have been chosen, every data point except for the two poles is 
-compared to both poles, for an additional 2$n$ comparisons.
-
-In general, at depth $d$ of
-the hierarchical tree, we have $\frac{n}{2^d}$ points per cluster (assuming a
-perfectly balanced clustering), $\sqrt{\frac{n}{2^d}}$ seeds, $\frac{n}{2^d}$
-pairwise seed comparisons, and $\frac{2n}{2^{d}}$ additional comparisons when 
-each other data point is compared to the two poles, for a total of $\frac{3n}{2^d}$
-distance comparisons. At each level, there are $2^d$ clusters,
-so the total number of distance comparisons per level is
-3$n$. Thus, the asymptotic complexity in terms of distance
-comparisons for a cluster tree of depth $d$ is $\mathcal{O}(dn)$
-
+The asympotitic complexity of Partition is the same as described in~\cite{ishaq2019clustered}
 By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of 
@@
 the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
-If no points are within a distance $m$ of the query, we increase the radius by a factor of 
+If no points are within a distance $m$ of the query, we increase the radius by a factor of arg
 2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
@@
 we increase the radius by a factor of 
-\begin{equation} \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
+\begin{equation} \argmin \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
 where $\mu$ is the local fractal dimension (abbreviated lfd in algorithm below) in the region around the query point.
@@
             \STATE $\mu \Leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
-            \STATE $m \Leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
+            \STATE $m \Leftarrow \argmin \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
             \STATE $H.push(\rho$-NN$(root, query, m)$)

### 2023-09-01 sections/2-methods.tex
more clarity changes

@@
 \subsection {Index-Building and Sharded Search}
+\label{subsec:methods:index-building}
 In this work, we also introduce the method of \emph{sharding} for $k$-nearest neighbors search.
@@
 
-We begin by randomly partitioning the dataset into $s$ shards, where $s$ is determined by [TODO: how did 
-we decide we're going to select $s$ for this version of the paper?]. On the first shard, we perform 
-$k$-nearest neighbors search using one of the four algorithms described in Section ~\ref{subsec:methods:knn-search}.
-The result of this search will give us the $k$ closest neighbors to the query in the first shard. We then take the furthest of 
-those $k$ neighbors and use its distance from the query as the radius for $\rho$-nearest neighbors search (see Section ~\ref{subsec:methods:rnn-search}) 
+We begin by randomly partitioning the dataset into $s$ shards, where $s$ is determined based 
+on some naive auto-tuning. On the first shard, we perform 
+$k$-nearest neighbors search using one of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
+The result of this search will give us the $k$ closest neighbors to the query \emph{in the first shard}. We then take the furthest of 
+those $k$ neighbors and use its distance from the query as the radius for use with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) 
 on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
@@
 since the first shard is the only one for which we perform $k$-nearest neighbors search. However, making this first shard too small
-creates the risk of significantly overestimating the radius, resulting in much slower $\rho$-nearest neighbors search on the remaining shards.
+introduces the risk of significantly overestimating the radius, resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
 Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well 
@@
 with the following improvement: when a cluster overlaps with the query ball, instead of  
-always proceeding with both of its children, we only proceed with those children which 
-also overlap with the query ball.
+always proceeding with both of its children, we proceed only with those children which 
+also overlap with the query ball. This improvement is demonstrated in Algorithm \ref{alg:rnn}.
 
@@
 
-In this section, we present four novel algorithms for $k$-nearest neighbors search: $k$-NN by Repeated $\rho$-NN, Sieve Search, and Sieve Search with Separate Centers, 
+In this section, we present four novel algorithms for exact $k$-nearest neighbors search: $k$-NN by Repeated $\rho$-NN, Sieve Search, Sieve Search with Separate Centers, 
 and Greedy Sieve. 
-We also have an implementation of linear search. We use a preliminary auto-tuning function to predict the variant which will perform 
-best for a given query, dataset, and value of $k$. We then proceed with search using that variant.
+We use a preliminary auto-tuning function to predict the variant and shard size (see Section~\ref{subsec:methods:index-building}) which will perform 
+best for a given query, dataset, and value of $k$. We then proceed with search using that variant on the first shard.
 
 In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and
-$Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the nearest neighbors.
+$Q$ to refer to the data structure which stores the clusters (and, for Sieve and Sieve with Centers, points) which are still in contention for being one of the nearest neighbors.
 
@@
 
-In this algorithm, we perform $\rho$-nearest neighbors search at a radius equal to the radius of the root cluster divided by
-the cardinality of the dataset, repeatedly increasing the search radius until $k$ neighbors
+In this algorithm, we perform $\rho$-nearest neighbors search small search radius, repeatedly increasing the radius until $k$ neighbors
 have been found.
@@
 increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
-by the local fractal dimension of the clusters containing the points found so far. In particular, 
+by the local fractal dimension in the region of the dataset around the query ball. 
+In particular, 
 we increase the radius by a factor of 
 \begin{equation} \argmin \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
-where $\mu$ is the local fractal dimension (abbreviated lfd in algorithm below) in the region around the query point.
+where $\mu$ is the harmonic mean of the local fractal dimensions (abbreviated lfd in algorithm below) of 
+each cluster which contains one of the nearest neighbors found thus far. We use the harmonic mean to 
+ensure that $\mu$ is not dominated by outlier clusters with very high fractal dimension. 
 Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
@@
             \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
-            \STATE $\mu \Leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
+            \STATE $\mu \Leftarrow |Q| \sum_{C \in Q} \frac{1}{C.lfd} $
             \STATE $m \Leftarrow \argmin \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
@@
 \label{subsubsec:methods:knn-search:sieve}
-With Sieve Search, we begin by letting $Q$ be a list initialized with the root cluster. This list will maintain the 
-clusters and points which are still in contention for being one of the nearest neighbors.
+With Sieve Search, we begin by letting $Q$ be a list containing the root cluster. 
 While $Q$ contains at least one cluster (i.e., $Q$ is not a list of only points), we repeat the process described in 
@@
 we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of elements in $Q$. 
-Our modified algorithm finds the smallest index $i$ such
+Our modified version of this algorithm (see Algorithm 4) finds the smallest index $i$ such 
 that all elements in $Q$ with $\delta_{max}$ closer to or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater
-than or equal to $k$. In the process of finding this index, we alter the order of $Q$ so that all elements to left
-of index $i$ have $\delta_{max}$ less than or equal to $Q[i]$'s $\delta_{max}$ and all elements to right of index $i$
-have $\delta_{max}$ greater than $Q[i]$'s $\delta_{max}$. We consider $Q[i]$ to be our $q_{\tau}$, and our threshold 
+than or equal to $k$. We consider $Q[i]$ to be our $q_{\tau}$, and our threshold 
 $\tau$ to be $Q[i]$'s $\delta_{max}$.
@@
 We complete this process until $Q$ contains only points. At this point, we 
-just use [INSERT ref to quicksort partition in paper] to find the $k$th 
-nearest neighbor. Since this algorithm will also elements on the correct side of 
-the index of the $k$th nearest neighbor, the first $k$ elements in $Q$ are the $k$-nearest neighbors.
+just use Algorithm 4 to find the $k$th 
+nearest neighbor and return all points to the left of it.
+
+This process is described in Algorithm~\ref{alg:sieve}.
 
@@
 center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, 
-thus allowing us to eliminate some clusters from $Q$ earlier in the process.
+thus allowing us to eliminate some clusters from $Q$ earlier in the process. More discussion about 
+when this algorithm is preferable to Sieve Search is in Section 4: Results.
 
@@
 
-With Greedy Sieve, we first let $Q$ be a min queue of clusters by non-increasing $\delta_{min}$. We initialize $Q$ with the root cluster.
-Let $H$ be a fixed-size ($k$) max queue of points by $\delta$. This queue starts empty.
+With Greedy Sieve, we first let $Q$ be a min queue of clusters based on $\delta_{min}$. We initialize $Q$ with the root cluster.
+Let $H$ be a fixed-size ($k$) max queue of points based on $\delta$. This queue starts empty.
 
 While $H$ has fewer than $k$ clusters or $\delta$ of the top priority element in $H$ is greater 
-than $\delta_{min}$ of the top priority element in $Q$, do the following:
+than $\delta_{min}$ of the top priority element in $Q$, we do the following:
 \begin{enumerate}
-\item While the top priority element $q$ in $Q$ is not a leaf, pop $q$ from the queue and push its children.
+\item While the top priority element $q$ in $Q$ is not a leaf, pop $q$ from $Q$ and push its children.
 \item For each point $p \in q$, push $p$ to $H$. 
-\item Pop points from $H$ until $H$ has $k$ points. 
+\item Pop from $H$ until $H$ has $k$ points. 
 \end{enumerate}
 After this process, the points left in $H$ are the $k$ nearest neighbors of $q$.
+
+This process is described in Algorithm \ref{alg:greedy_search}.
 

### 2023-09-01 sections/3-datasets.tex
more clarity changes

@@
 Way galaxy. Each spectral datum describes a star's flux in
-Janskys ($1 Jy = 10^{-26} W \dot m^{-2} \dot Hz^{-1}$) at each of 8,575 wavelengths ranging from $\lambda$ = 1.51 µm to 1.70 µm. Thus,
+Janskys ($1 Jy = 10^{-26} W \dot m^{-2} \dot Hz^{-1}$) at each of 8,575 wavelengths ranging from $\lambda$ = 1.51 µm to 1.70 µm. 
 Thus, each datum is a real-valued vector in $\mathbb{R}_{+}^{8575}$.
@@
 Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history. This dataset 
-is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure 1 in Section 2,
-this dataset is a good candidate for CAKES because it exhibits low fractal dimension.
+is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As explained in~\cite{ishaq2019clustered},
+this dataset is a good candidate for entropy-scaling search algorithms like CAKES because it exhibits low fractal dimension.
 
@@
 
-\dots
-
 \subsection{ANN-Benchmarks suite}
@@
 We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
-which we list below, along with the distance functions with which they are typically used.
+listed below, along with the appropriate distance function:
 
 \begin{itemize} 
-    \item DEEP1B (Cosine): ~1 billion 96-dimensional deep descriptors to 
+    \item DEEP1B (Cosine): $\sim$1 billion 96-dimensional deep descriptors to 
     \item Fashion-MNIST (Euclidean): 60,000 784-dimensional vector representations of images of clothing items
     \item GIST (Euclidean): 1 million 960-dimensional image descriptors 
-    \item GloVe (Cosine): ~1.2 million 25-dimensional vector representations of words
-    \item Kosarak (Jaccard): ~75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
+    \item GloVe (Cosine): $\sim$1.2 million 25-dimensional vector representations of words
+    \item Kosarak (Jaccard): $\sim$75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
     \item MNIST (Euclidean): 60,000 784-dimensional vector representations of images of handwritten digits
@@
     \item SIFT (Euclidean): 1 million 128-dimensional image descriptors
-    \item Last.fm (Cosine): ~292, 000 65-dimensional vector representations of songs
+    \item Last.fm (Cosine): $\sim$292, 000 65-dimensional vector representations of songs
 \end{itemize}

### 2023-09-01 sections/2-methods.tex
complexity

@@
 
+\subsubsection{Complexity of $k$-NN by Repeated $\rho$-NN Search}
+\label{paragraph:methods:repeated-rnn-complexity}
+
+Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
+(i.e. arises from the same generating phenomenon) as the rest of the data. We also rely on the same technical assumption about the self-similarity 
+of the dataset made in \cite{yu2015entropy}: if call the number of points within a given 
+radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
+are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
+we assume that for any point $p$, query $q$, and radius 
+$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}
+
+We find it useful to adopt the terminology used in 
+\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
+Coarse search refers to the process of identifying clusters
+which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
+Fine search refers to the process
+of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
+
+In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search is $\mathcal{O}(\log{2}m)$, 
+where $m$ is the metric entropy of the dataset. 
+To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
+of $\rho$-NN needed to find at least $k$ neighbors.
+
+Because we assume that the query point is sampled from 
+the same distribution as the rest of the data, we can assume that the local fractal dimension around the query
+is not significantly different from the mean local fractal dimension of the dataset. If the local fractal dimension 
+were perfectly uniform throughout the entire dataset, then 
+the use of the factor from Equation \ref{2} would guarantee that it only takes 
+one iteration of $\rho$-NN to find $k$ points. In practice, [is this the 
+way I want to be going with this??]
+
+To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
+number of clusters for which we must look at all of the cluster's points. With $\rho$-NN 
+search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
+where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
+as in \cite{yu2015entropy}. With $k$-NN by Repeated $\rho$-NN, however, 
+the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
+is query-dependent. Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of 
+$F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
+where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.
+
+~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$. By definition of 
+$r'$, we have that $|B(q, r')| = k$. Thus, we have that $|F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d$.
+
+It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
+points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
+average local fractal dimension over all regions of the dataset does not differ significantly from the local fractal dimension around any particular point.
+We rely on this fact in our estimation of $r'$.
+
+We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
+$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
+and the cardinality of the ball centered at the root cluster). In this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
+er's center with half its radius. While, ordinarily we compute local fractal dimension 
+by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
+a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.
+
+If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
+\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}
+
+Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
+\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}
+
 \subsubsection{Sieve Search}
@@
 
-\subsubsection{Complexity of $k$-NN Search}
-\label{paragraph:methods:knn-complexity}
-
-To bound the complexity of $k$-NN search, we group all of the "Sieve" algorithms (Sieve, 
-Sieve with Separate Centers, and Greedy Sieve) together, and handle $k$-NN by Repeated $\rho$-NN separately.
-
-We start by considering the complexity of $k$-NN by Repeated $\rho$-NN, as this  
-is a natural extension of the complexity bounds for $\rho$-NN search described in 
-\cite{ishaq2019clustered}. For this algorithm, we adopt the terminology used in 
-\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
-Coarse search refers to the process of identifying clusters
-which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
-Fine search refers to the process
-of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
+\subsubsection{Complexity of Sieve, Sieve with Separate Centers, and Greedy Sieve}
+\label{paragraph:methods:sieve-complexity}
 
@@
 % from during search, even if we do not actually perform these as two distinct phases.
-
-% Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
-% \cite{yu2015entropy}: if call the number of points within a given 
-% radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
-% are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
-% we assume that for any point $p$, query $q$, and radius 
-% $\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}
-
-% For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
-% if this still holds for knn].
 
@@
 % \end{align*}
-
-% It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
-% points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
-% average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
-% We rely on this fact in our estimation of $r'$.
-
-% We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
-% $d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
-% and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
-% er's center with half its radius). While, ordinarily we compute local fractal dimension 
-% by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
-% a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.
-
-% If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
-% \begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}
-
-% Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
-% \begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

### 2023-09-02 sections/4-results.tex
complexity
results table

@@
 on [ADD DATASET] under [ADD DISTANCE FUNCTION] distance, using [ADD NUMBER OF SHARDS] shards.
+As indicated in the title, the dataset has cardinality [ADD cardinality] and dimensionality [ADD DIMENSIONALITY].
 On the horizontal axis, we have each of the $k$-NN algorithms described in Section 2.4, as well as  
@@
 \end{enumerate}
+
+("deep-image", "cosine"),       // 0
+("fashion-mnist", "euclidean"), // 1
+("gist", "euclidean"),          // 2
+("glove-25", "cosine"),         // 3
+("glove-50", "cosine"),         // 4
+("glove-100", "cosine"),        // 5
+("glove-200", "cosine"),        // 6
+("kosarak", "jaccard"),         // 7
+("mnist", "euclidean"),         // 8
+("nytimes", "cosine"),          // 9
+("sift", "euclidean"),          // 10
+("lastfm", "cosine"),           // 11
+
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Performance (ROC AUC) of CHAODA vs. other methods on the 18 test datasets.}
+    \label{table:results:test-performance}
+    \vskip 0.15in
+    \begin{center}
+    \begin{small}
+    \begin{sc}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
+    \hline
+    \textbf{Dataset} & \textbf{Dist.}  &\textbf{Card.}  & \textbf{Dim.}  & \textbf{k}  & \multicolumn{2}{r}\textbf{faiss-ivf} & \multicolumn{2}{r}\textbf{redisearch} & \multicolumn{2}{r}\textbf{hnsw(faiss)} & \multicolumn{2}{r}\textbf{CAKES} \\
+    &                &                  &                   &               &             Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
+    \hline
+    nytimes          & cos              & 290,000            & 256             & 10         & 0.69 &  1413.35                       & 0.90 & 616.21                            & 0.97 & 92.92                                         & 1.00 & 42 \\
+    \hline
+    nytimes          & cos              & 290,000            & 256             & 100        & 0.61 &  0.71                          & 0.50 & 0.45                              & 0.00 & 0.00                                          & 1.00 & 42 \\
+    \hline
+    fashion-mnist    & euc              & 60,000	         & 784            & 10         & 0.54 & 324.79                           & 0.99 & 444.89                                  & 0.94 & 60.89                                                    & 1.00 & 42 \\
+    \hline
+    fashion-mnist    & euc              & 60,000             & 784            & 100        & 1.00 & 111.47                           & 0.50 & 0.49                                  & 0.99 & 277.80                                                    & 1.00 & 42 \\
+    \hline 
+    gist             & euc              & 1,000,000          & 960             & 10         & 1.00 & 3.55                           & -- & --                                     & 0.95 & 149.58                                                   & 1.00 & 42 \\
+    \hline
+    gist             & euc              & 1,000,000          & 960             & 100        & 1.00 & 2.70                           & 0.68 & 0.47                                 & 0.99 & 59.71                                                    & 1.00 & 42 \\
+    \hline
+    glove-25         & cos              & 1,183,514          & 25             & 10         & 1.00 & 78.78                          & 1.00 & 37.85                               & 1.00 & 552.67                                                  & 1.00 & 42 \\
+    \hline
+    glove-25         & cos              & 1,183,514          & 25             & 100        & -- & --                           & -- & --                                & -- & --                                                    & 1.00 & 42 \\
+    \hline
+    glove-100        & cos              & 1,183,514          & 100            & 10         & 1.00 & 21.16                          & 0.99 & 133.05                                & 0.99 & 155.90                                                  & 1.00 & 42 \\
+    \hline
+    glove-100        & cos              & 1,183,514          & 100            & 100        & 0.99 &  19.04                          & -- & --                                  & 0.99 & 92.58                                                    & 1.00 & 42 \\
+    \hline
+    sift             & euc              & 1,000,000          & 128            & 10         & 1.00 &  24.52                          & 1.00 & 505.79                               & 1.00 & 275.07                                                    & 1.00 & 42 \\
+    \hline
+    sift             & euc              & 1,000,000          & 128            & 100        & 1.00 &  23.50                          & 0.50 & 0.45                                  & 0.99 & 179.14                                                    & 1.00 & 42 \\
+    \hline
+    deep-image       & cos              & 9,990,000	         & 96             & 10          & 0.00 & 0.00                           & 0.00 & 0.00                                 & 0.00 & 0.00                                                     & 1.00 & 42 \\
+    \hline
+    deep-image       & cos              & 9,990,000	         & 96             & 100         & 0.00 & 0.00                           & 0.00 & 0.00                                & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    mnist            & euc              & 60,000             & 784             & 10          & 0.00 & 0.00                          & 0.00 & 0.00                                 & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    mnist            & euc              & 60,000             & 784             & 100         & 0.00 & 0.00                          & 0.00 & 0.00                                  & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    glove-50         & cos              & 1,183,514          & 50             & 10          & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    glove-50         & cos              & 1,183,514          & 50             & 100         & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    lastfm           & cos              & 292,385            & 65             & 10          & 0.00 & 0.00                           & 0.00 & 0.00                               & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    lastfm           & cos              & 292,385            & 65             & 100         & 0.00 & 0.00                           & 0.00 & 0.00                             & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    \end{tabular}
+    \end{sc}
+    \end{small}
+    \end{center}
+    \vskip -0.1in
+    \end{table*}

### 2023-09-02 sections/4-results.tex
bruteforce-blass benchmarks

@@
     \hline
-    \textbf{Dataset} & \textbf{Dist.}  &\textbf{Card.}  & \textbf{Dim.}  & \textbf{k}  & \multicolumn{2}{r}\textbf{faiss-ivf} & \multicolumn{2}{r}\textbf{redisearch} & \multicolumn{2}{r}\textbf{hnsw(faiss)} & \multicolumn{2}{r}\textbf{CAKES} \\
+    \textbf{Dataset} & \textbf{Dist.}  &\textbf{Card.}  & \textbf{Dim.}  & \textbf{k}  & \multicolumn{2}{r}\textbf{faiss-ivf} & \multicolumn{2}{r}\textbf{bruteforce-blas} & \multicolumn{2}{r}\textbf{hnsw(faiss)} & \multicolumn{2}{r}\textbf{CAKES} \\
     &                &                  &                   &               &             Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
     \hline
-    nytimes          & cos              & 290,000            & 256             & 10         & 0.69 &  1413.35                       & 0.90 & 616.21                            & 0.97 & 92.92                                         & 1.00 & 42 \\
+    nytimes          & cos              & 290,000            & 256             & 10         & 0.69 &  1413.35                       & 0.99 & 26.66                            & 0.97 & 92.92                                         & 1.00 & 42 \\
     \hline
-    nytimes          & cos              & 290,000            & 256             & 100        & 0.61 &  0.71                          & 0.50 & 0.45                              & 0.00 & 0.00                                          & 1.00 & 42 \\
+    nytimes          & cos              & 290,000            & 256             & 100        & 0.61 &  0.71                          & 0.99 & 10.71                              & 0.00 & 0.00                                          & 1.00 & 42 \\
     \hline
-    fashion-mnist    & euc              & 60,000	         & 784            & 10         & 0.54 & 324.79                           & 0.99 & 444.89                                  & 0.94 & 60.89                                                    & 1.00 & 42 \\
+    fashion-mnist    & euc              & 60,000	         & 784            & 10         & 0.54 & 324.79                           & 1.00 & 52.91                                  & 0.94 & 60.89                                                    & 1.00 & 42 \\
     \hline
-    fashion-mnist    & euc              & 60,000             & 784            & 100        & 1.00 & 111.47                           & 0.50 & 0.49                                  & 0.99 & 277.80                                                    & 1.00 & 42 \\
+    fashion-mnist    & euc              & 60,000             & 784            & 100        & 1.00 & 111.47                           & 0.99 & 16.02                                  & 0.99 & 277.80                                                    & 1.00 & 42 \\
     \hline 
-    gist             & euc              & 1,000,000          & 960             & 10         & 1.00 & 3.55                           & -- & --                                     & 0.95 & 149.58                                                   & 1.00 & 42 \\
+    gist             & euc              & 1,000,000          & 960             & 10         & 1.00 & 3.55                           & 1.00 & 2.63                                     & 0.95 & 149.58                                                   & 1.00 & 42 \\
     \hline
-    gist             & euc              & 1,000,000          & 960             & 100        & 1.00 & 2.70                           & 0.68 & 0.47                                 & 0.99 & 59.71                                                    & 1.00 & 42 \\
+    gist             & euc              & 1,000,000          & 960             & 100        & 1.00 & 2.70                           & 0.99 & 0.80                                 & 0.99 & 59.71                                                    & 1.00 & 42 \\
     \hline
-    glove-25         & cos              & 1,183,514          & 25             & 10         & 1.00 & 78.78                          & 1.00 & 37.85                               & 1.00 & 552.67                                                  & 1.00 & 42 \\
+    glove-25         & cos              & 1,183,514          & 25             & 10         & 1.00 & 78.78                          & 1.00 & 33.52                              & 1.00 & 552.67                                                  & 1.00 & 42 \\
     \hline
@@
     \hline
-    glove-100        & cos              & 1,183,514          & 100            & 10         & 1.00 & 21.16                          & 0.99 & 133.05                                & 0.99 & 155.90                                                  & 1.00 & 42 \\
+    glove-100        & cos              & 1,183,514          & 100            & 10         & 1.00 & 21.16                          & 1.00 & 16.52                                & 0.99 & 155.90                                                  & 1.00 & 42 \\
     \hline
-    glove-100        & cos              & 1,183,514          & 100            & 100        & 0.99 &  19.04                          & -- & --                                  & 0.99 & 92.58                                                    & 1.00 & 42 \\
+    glove-100        & cos              & 1,183,514          & 100            & 100        & 0.99 &  7.52                          & 0.99 & 3.11                                  & 0.99 & 92.58                                                    & 1.00 & 42 \\
     \hline
-    sift             & euc              & 1,000,000          & 128            & 10         & 1.00 &  24.52                          & 1.00 & 505.79                               & 1.00 & 275.07                                                    & 1.00 & 42 \\
+    sift             & euc              & 1,000,000          & 128            & 10         & 1.00 &  24.52                          & 1.00 & 16.40                               & 1.00 & 275.07                                                    & 1.00 & 42 \\
     \hline
-    sift             & euc              & 1,000,000          & 128            & 100        & 1.00 &  23.50                          & 0.50 & 0.45                                  & 0.99 & 179.14                                                    & 1.00 & 42 \\
+    sift             & euc              & 1,000,000          & 128            & 100        & 1.00 &  23.50                          & 1.00 & 6.51                                  & 0.99 & 179.14                                                    & 1.00 & 42 \\
     \hline

### 2023-09-02 sections/2-methods.tex
removed quicksort partition

@@
 we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of elements in $Q$. 
-Our modified version of this algorithm (see Algorithm 4) finds the smallest index $i$ such 
+Our modified version of this algorithm finds the smallest index $i$ such 
 that all elements in $Q$ with $\delta_{max}$ closer to or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater
@@
 This process is described in Algorithm~\ref{alg:sieve}.
-
-\begin{algorithm} % enter the algorithm environment
-    \caption{QuicksortPartition(\emph{Q, k, l, r})} % give the algorithm a caption
-    \label{alg:quicksort-partition} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[5] % enter the algorithmic environment
-        \IF{$l \geq r$}
-            \STATE Return $min(l, r)$
-        \ELSE 
-            \STATE $pivot = l + (r - l) / 2$
-            \STATE $Q.swap(pivot, r)$
-            \STATE $a \Leftarrow l$
-            \STATE $b \Leftarrow l$
-            \WHILE {$b < r$}
-                \IF {$Q[b].\delta_{max} < Q[r].\delta_{max}$}
-                    \STATE $Q.swap(a, b)$
-                    \STATE $a \Leftarrow a + 1$
-                \ENDIF
-                \STATE $b \Leftarrow b + 1$
-            \ENDWHILE
-            \STATE $Q.swap(a, r)$
-            \STATE $p \Leftarrow r$
-            \STATE $g \Leftarrow \sum_{i=0}^{p} |Q[i]|$
-            \IF {$g = k$}
-                \STATE Return $p$
-            \ELSIF {$g < k$}
-                \STATE $QuicksortPartition(Q, k, p + 1, r)$
-            \ELSE 
-                \IF {$p > 0 \land g > k + |Q[p - 1]|$}
-                    \STATE $QuicksortPartition(Q, k, l, p - 1)$
-                \ELSIF {$p > 0 \land g = k + |Q[p - 1]|$}
-                    \STATE Return $p - 1$
-                \ELSE
-                    \STATE Return $p$
-                \ENDIF
-            \ENDIF
-        \ENDIF
-    \end{algorithmic}
-    \end{algorithm}
 
@@
 
-\subsubsection{Complexity of Sieve, Sieve with Separate Centers, and Greedy Sieve}
+\subsubsection{Complexity of Sieve Algorithms}
 \label{paragraph:methods:sieve-complexity}

### 2023-09-02 sections/4-results.tex
removed quicksort partition

@@
 \end{enumerate}
-
-("deep-image", "cosine"),       // 0
-("fashion-mnist", "euclidean"), // 1
-("gist", "euclidean"),          // 2
-("glove-25", "cosine"),         // 3
-("glove-50", "cosine"),         // 4
-("glove-100", "cosine"),        // 5
-("glove-200", "cosine"),        // 6
-("kosarak", "jaccard"),         // 7
-("mnist", "euclidean"),         // 8
-("nytimes", "cosine"),          // 9
-("sift", "euclidean"),          // 10
-("lastfm", "cosine"),           // 11
 

### 2023-09-02 sections/2-methods.tex
references fixes and edits to rnn complexity

@@
 Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
-(i.e. arises from the same generating phenomenon) as the rest of the data. We also rely on the same technical assumption about the self-similarity 
-of the dataset made in \cite{yu2015entropy}: if call the number of points within a given 
-radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
-are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
-we assume that for any point $p$, query $q$, and radius 
-$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}
+(i.e. arises from the same generating phenomenon) as the rest of the data. From this assumption, we can infer that the local fractal dimension
+at the query point is not significantly different from the mean local fractal dimension of clusters nearby the query.
 
@@
 
-In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search is $\mathcal{O}(\log{2}m)$, 
+In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log{2}m)$, 
 where $m$ is the metric entropy of the dataset. 
@@
 
-Because we assume that the query point is sampled from 
-the same distribution as the rest of the data, we can assume that the local fractal dimension around the query
-is not significantly different from the mean local fractal dimension of the dataset. If the local fractal dimension 
-were perfectly uniform throughout the entire dataset, then 
-the use of the factor from Equation \ref{2} would guarantee that it only takes 
-one iteration of $\rho$-NN to find $k$ points. In practice, [is this the 
-way I want to be going with this??]
+Based on the assumption that the local fractal dimension at the query point is not significantly different than that of 
+nearby clusters, the factor from Equation \ref{2} suggests that in the expected case, we need only two iterations of 
+$\rho$-NN to find $k$ neighbors: one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
+Since this is a constant factor, it does not affect the asymptotic complexity of coarse search, and thus we have that 
+the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2{m})$.
 
@@
 
-~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$. By definition of 
+~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$, where $\gamma$ is a constant. 
+By definition of 
 $r'$, we have that $|B(q, r')| = k$. Thus, we have that $|F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d$.
 
-It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
-points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
-average local fractal dimension over all regions of the dataset does not differ significantly from the local fractal dimension around any particular point.
-We rely on this fact in our estimation of $r'$.
+It remains to provide an estimate for $r'$. 
+To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, 
+and thus the local fractal dimension at the query point is not significantly different than that of nearby clusters.
 
-We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
-$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
-and the cardinality of the ball centered at the root cluster). In this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
-er's center with half its radius. While, ordinarily we compute local fractal dimension 
-by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
-a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.
+We let $d_i$ be the local fractal dimension of some cluster $C_i$ near the query point (i.e., some cluster identified during coarse search).
+While ordinarily we compute local fractal dimension 
+by comparing cardinalities of two balls centered at \emph{the same} point at two different radii, in order to estimate $r'$, 
+we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of cluster $C_i$} at its radius $r_i$.
+We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could "drag" the query to 
+the center of $C_i$ without significantly changing the local fractal dimension at the query point.
 
-If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
-\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}
+By ~\ref{2}, we then have that \begin{equation} d_i = \frac{\log{}\frac{|C_i|}{k}}{\log{}\frac{r_i}{r'}}. \label{4} \end{equation}
 
-Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
-\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}
+Since $d_i$, $|C_i|$, $r_i$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
+\begin{equation} r' = r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}. \label{5} \end{equation}
+
+Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has 
+$\mathcal{O}\left(\log_2{m} + k\left(\frac{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}+ 2r_c}{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}}\right)^d\right)$ 
+time complexity, where $m$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
+near the query ball, $d_i$ is the local fractal dimension of $C_i$, $r_c$ is the radius of leaf clusters, and $k$ is the number of nearest neighbors.
 

### 2023-09-02 sections/3-datasets.tex
removed quicksort partition
references fixes and edits to rnn complexity

@@
 \label{subsec:datasets:apogee-2}
-The Sloan Digital Sky Survey (SDSS) [3] Apache Point Observatory Galaxy Evolution
-Experiment (APOGEE) [4] contains near-infrared
+The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss} Apache Point Observatory Galaxy Evolution
+Experiment (APOGEE)~\cite{alam2015eleventh} contains near-infrared
 spectra of approximately 130,000 stellar objects in the Milky
@@
 \label{subsec:datasets:manga}
-SDSS's Mapping Nearby Galaxies at APO (MaNGA) [CITE] contains spectral measurements across the face of each of 
+SDSS's Mapping Nearby Galaxies at APO (MaNGA)~\cite{Talbot2022SDSSIVMA} contains spectral measurements across the face of each of 
 about 10,000 galaxies across 2,700 deg$^2$.
@@
 dataset is typically used with Euclidean (L2) distance, cosine distance, and 2d-Wasserstein distance.
-
-TODO: Add details about why this dataset is a good candidate for CAKES.
 
@@
 This dataset is typically used with Hamming and Levenshtein distances.
-
-TODO: Add details about why this dataset is a good candidate for CAKES.
 

### 2023-09-02 sections/2-methods.tex
weak stab at caption

@@
     \includegraphics[width=2.5in]{images/geometry/geometry.pdf}
-    \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk}
+    \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk. $R$ 
+    denotes the radius of the root cluster $C$, and $o$ denotes its center. The right 
+    triangle formed by $o$, $o'$, and $y+$ in $C'$ shows that $R \leq R'$, where $R'$ is the radius of $C'$.
+    Hence, the radius of a child cluster can be larger than its parent. $C''$ 
+    illustrates that after we have consumed all $d$ axes, the radius of a child cluster
+    is bounded above by $\frac{R}{\sqrt{2}}$.}
     \label{fig:methods:scaling_behavior}

### 2023-09-02 sections/2-methods.tex
changes after Sat AM call. Table headers TBD.

@@
 \label{alg:partition} % and a label for \ref{} commands later in the document
-\begin{algorithmic}[1] % enter the algorithmic environment
+\begin{algorithmic} % enter the algorithmic environment
     \STATE $m \Leftarrow \lfloor \sqrt{|C|} \rfloor$
@@
     \centering
-    \includegraphics[width=2.5in]{images/geometry/geometry.pdf}
+    \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
     \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk. $R$ 
@@
     \label{alg:rnn} 
-    \begin{algorithmic}[2]
+    \begin{algorithmic}
         \REQUIRE $r \geq 0$
@@
 
-The asymptotic complexity of $\rho$-nearest neighbors is the same as in ~\cite{ishaq2019clustered}.
+The asymptotic complexity of $\rho$-nearest neighbors is the same as in ~\cite{ishaq2019clustered}, namely:
+
+\begin{gather}
+    \mathcal{O}\Bigg(
+    \underbrace{\log_2 \mathcal{N}_{\hat{r_c}}(D)}_{\textrm{metric entropy}} +
+    \overbrace{\left|B_D(q,r)\right|}^{\textrm{output size}}
+    \underbrace{\left(\frac{r+2\hat{r_c}}{r}\right)^d}_{\textrm{scaling factor}}
+     \Bigg)
+     \label{hierarchical-complexity}
+\end{gather}
+
+where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy at that radius, $B_D(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the factal dimension around the query in a region of radius $r$.
 
@@
     \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[4] % enter the algorithmic environment
+    \begin{algorithmic} % enter the algorithmic environment
         \STATE $H \Leftarrow$ $\emptyset$
@@
     \label{alg:sieve} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[6] % enter the algorithmic environment
+    \begin{algorithmic} % enter the algorithmic environment
         \REQUIRE $tree$, $query$, $k$
@@
 \label{alg:greedy_search} 
-\begin{algorithmic}[3]
+\begin{algorithmic}
     \STATE $Q \Leftarrow$ priority queue

### 2023-09-02 sections/3-datasets.tex
changes after Sat AM call. Table headers TBD.

--- 2023-09-01 sections/3-datasets.tex
@@
 \label{sec:datasets-and-distance-functions}
-
-\subsection{APOGEE-2}
-\label{subsec:datasets:apogee-2}
-The Sloan Digital Sky Survey (SDSS) [3] Apache Point Observatory Galaxy Evolution
-Experiment (APOGEE) [4] contains near-infrared
-spectra of approximately 130,000 stellar objects in the Milky
-Way galaxy. Each spectral datum describes a star's flux in
-Janskys ($1 Jy = 10^{-26} W \dot m^{-2} \dot Hz^{-1}$) at each of 8,575 wavelengths ranging from $\lambda$ = 1.51 µm to 1.70 µm. Thus,
-Thus, each datum is a real-valued vector in $\mathbb{R}_{+}^{8575}$.
-
-Astronomers use these near-infrared spectra to achieve a better understanding of our galaxy's history. This dataset 
-is often used with Euclidean (L2) distance, cosine distance, and 1d-Wasserstein distance. As illustrated by Figure 1 in Section 2,
-this dataset is a good candidate for CAKES because it exhibits low fractal dimension.
-
-\subsection{MaNGA}
-\label{subsec:datasets:manga}
-SDSS's Mapping Nearby Galaxies at APO (MaNGA) [CITE] contains spectral measurements across the face of each of 
-about 10,000 galaxies across 2,700 deg$^2$.
-
-Like APOGEE, MaNGA helps astronomers understand the history of galaxies throughout their entire lifespan. This 
-dataset is typically used with Euclidean (L2) distance, cosine distance, and 2d-Wasserstein distance.
-
-TODO: Add details about why this dataset is a good candidate for CAKES.
-
-\subsection{Silva 18S}
-\label{subsec:datasets:silva-18s}
-
-Silva 18S contains ribosomal DNA sequences of approximately 2.25 
-million genomes with aligned length of 50,000 letters.
-
-These sequences help biologists better understand the evolutionary relationships between organisms. 
-This dataset is typically used with Hamming and Levenshtein distances.
-
-TODO: Add details about why this dataset is a good candidate for CAKES.
-
-\dots
 
@@
 We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
-which we list below, along with the distance functions with which they are typically used.
+listed below, along with the appropriate distance function.
+Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
 
-\begin{itemize} 
-    \item DEEP1B (Cosine): ~1 billion 96-dimensional deep descriptors to 
-    \item Fashion-MNIST (Euclidean): 60,000 784-dimensional vector representations of images of clothing items
-    \item GIST (Euclidean): 1 million 960-dimensional image descriptors 
-    \item GloVe (Cosine): ~1.2 million 25-dimensional vector representations of words
-    \item Kosarak (Jaccard): ~75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
-    \item MNIST (Euclidean): 60,000 784-dimensional vector representations of images of handwritten digits
-    \item NYTimes (Cosine): 290,000 256-dimensional vector representations of words
-    \item SIFT (Euclidean): 1 million 128-dimensional image descriptors
-    \item Last.fm (Cosine): ~292, 000 65-dimensional vector representations of songs
-\end{itemize}
+% \begin{itemize} 
+%     \item DEEP1B (Cosine): $\sim$1 billion 96-dimensional deep descriptors to 
+%     \item Fashion-MNIST (Euclidean): 60,000 784-dimensional vector representations of images of clothing items
+%     \item GIST (Euclidean): 1 million 960-dimensional image descriptors 
+%     \item GloVe (Cosine): $\sim$1.2 million 25-dimensional vector representations of words
+%     \item Kosarak (Jaccard): $\sim$75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
+%     \item MNIST (Euclidean): 60,000 784-dimensional vector representations of images of handwritten digits
+%     \item NYTimes (Cosine): 290,000 256-dimensional vector representations of words
+%     \item SIFT (Euclidean): 1 million 128-dimensional image descriptors
+%     \item Last.fm (Cosine): $\sim$292, 000 65-dimensional vector representations of songs
+% \end{itemize}
+
+\begin{table}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Datasets used in benchmarks.}
+    \label{table:datasets:summary}
+    \vskip 0.15in
+    \begin{center}
+    \begin{small}
+    \begin{sc}
+    \begin{tabular}{|l|l|l|l|}
+    \hline
+    \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dim.}  \\
+    &                &                  &                         \\
+    \hline
+    nytimes          & cosine              & 290,000            & 256       \\
+    \hline
+    fashion-mnist    & euclidean              & 60,000             & 784       \\
+    \hline 
+    gist             & euclidean              & 1,000,000          & 960       \\
+    \hline
+    glove-25         & cosine              & 1,183,514          & 25        \\
+    \hline
+    glove-100        & cosine              & 1,183,514          & 100       \\
+    \hline
+    sift             & euclidean              & 1,000,000          & 128       \\
+    \hline
+    deep-image       & cosine              & 9,990,000          & 96        \\
+    \hline
+    mnist            & euclidean              & 60,000             & 784       \\
+    \hline
+    glove-50         & cosine              & 1,183,514          & 50        \\
+    \hline
+    lastfm           & cosine              & 292,385            & 65        \\
+    \hline
+    \end{tabular}
+    \end{sc}
+    \end{small}
+    \end{center}
+    \vskip -0.1in
+    \end{table}

### 2023-09-02 sections/4-results.tex
changes after Sat AM call. Table headers TBD.

@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Performance (ROC AUC) of CHAODA vs. other methods on the 18 test datasets.}
-    \label{table:results:test-performance}
+    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=10$}
+    \label{table:results:ann-10}
     \vskip 0.15in
@@
     \begin{sc}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
+    \begin{tabular}{|l|c|c|c|c|c|c|c|c|}
+    \textbf{Dataset}  & \multicolumn{2}{|c|}\textbf{faiss-ivf} & \multicolumn{2}{|c|}\textbf{bruteforce-blas} & \multicolumn{2}{|c|}\textbf{hnsw(faiss)} & \multicolumn{2}{|c|}\textbf{CAKES} \\
+    \cline{2-9}
+    &                    Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
     \hline
-    \textbf{Dataset} & \textbf{Dist.}  &\textbf{Card.}  & \textbf{Dim.}  & \textbf{k}  & \multicolumn{2}{r}\textbf{faiss-ivf} & \multicolumn{2}{r}\textbf{bruteforce-blas} & \multicolumn{2}{r}\textbf{hnsw(faiss)} & \multicolumn{2}{r}\textbf{CAKES} \\
-    &                &                  &                   &               &             Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
+    nytimes                & 0.69 &  1413.35                       & 0.99 & 26.66                            & 0.97 & 92.92                                         & 1.00 & 42 \\
     \hline
-    nytimes          & cos              & 290,000            & 256             & 10         & 0.69 &  1413.35                       & 0.99 & 26.66                            & 0.97 & 92.92                                         & 1.00 & 42 \\
+    fashion-mnist      & 0.54 & 324.79                           & 1.00 & 52.91                                  & 0.94 & 60.89                                                    & 1.00 & 42 \\
     \hline
-    nytimes          & cos              & 290,000            & 256             & 100        & 0.61 &  0.71                          & 0.99 & 10.71                              & 0.00 & 0.00                                          & 1.00 & 42 \\
+    gist                   & 1.00 & 3.55                           & 1.00 & 2.63                                     & 0.95 & 149.58                                                   & 1.00 & 42 \\
     \hline
-    fashion-mnist    & euc              & 60,000	         & 784            & 10         & 0.54 & 324.79                           & 1.00 & 52.91                                  & 0.94 & 60.89                                                    & 1.00 & 42 \\
+    glove-25              & 1.00 & 78.78                          & 1.00 & 33.52                              & 1.00 & 552.67                                                  & 1.00 & 42 \\
     \hline
-    fashion-mnist    & euc              & 60,000             & 784            & 100        & 1.00 & 111.47                           & 0.99 & 16.02                                  & 0.99 & 277.80                                                    & 1.00 & 42 \\
+    glove-100             & 1.00 & 21.16                          & 1.00 & 16.52                                & 0.99 & 155.90                                                  & 1.00 & 42 \\
     \hline
-    gist             & euc              & 1,000,000          & 960             & 10         & 1.00 & 3.55                           & 1.00 & 2.63                                     & 0.95 & 149.58                                                   & 1.00 & 42 \\
+    sift                  & 1.00 &  24.52                          & 1.00 & 16.40                               & 1.00 & 275.07                                                    & 1.00 & 42 \\
     \hline
-    gist             & euc              & 1,000,000          & 960             & 100        & 1.00 & 2.70                           & 0.99 & 0.80                                 & 0.99 & 59.71                                                    & 1.00 & 42 \\
+    deep-image             & 0.00 & 0.00                           & 0.00 & 0.00                                 & 0.00 & 0.00                                                     & 1.00 & 42 \\
     \hline
-    glove-25         & cos              & 1,183,514          & 25             & 10         & 1.00 & 78.78                          & 1.00 & 33.52                              & 1.00 & 552.67                                                  & 1.00 & 42 \\
+    mnist                   & 0.00 & 0.00                          & 0.00 & 0.00                                 & 0.00 & 0.00                                                    & 1.00 & 42 \\
     \hline
-    glove-25         & cos              & 1,183,514          & 25             & 100        & -- & --                                & -- & --                                & -- & --                                                    & 1.00 & 42 \\
+    glove-50               & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
     \hline
-    glove-100        & cos              & 1,183,514          & 100            & 10         & 1.00 & 21.16                          & 1.00 & 16.52                                & 0.99 & 155.90                                                  & 1.00 & 42 \\
-    \hline
-    glove-100        & cos              & 1,183,514          & 100            & 100        & 0.99 &  7.52                          & 0.99 & 3.11                                  & 0.99 & 92.58                                                    & 1.00 & 42 \\
-    \hline
-    sift             & euc              & 1,000,000          & 128            & 10         & 1.00 &  24.52                          & 1.00 & 16.40                               & 1.00 & 275.07                                                    & 1.00 & 42 \\
-    \hline
-    sift             & euc              & 1,000,000          & 128            & 100        & 1.00 &  23.50                          & 1.00 & 6.51                                  & 0.99 & 179.14                                                    & 1.00 & 42 \\
-    \hline
-    deep-image       & cos              & 9,990,000	         & 96             & 10          & 0.00 & 0.00                           & 0.00 & 0.00                                 & 0.00 & 0.00                                                     & 1.00 & 42 \\
-    \hline
-    deep-image       & cos              & 9,990,000	         & 96             & 100         & 0.00 & 0.00                           & 0.00 & 0.00                                & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    mnist            & euc              & 60,000             & 784             & 10          & 0.00 & 0.00                          & 0.00 & 0.00                                 & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    mnist            & euc              & 60,000             & 784             & 100         & 0.00 & 0.00                          & 0.00 & 0.00                                  & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    glove-50         & cos              & 1,183,514          & 50             & 10          & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    glove-50         & cos              & 1,183,514          & 50             & 100         & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    lastfm           & cos              & 292,385            & 65             & 10          & 0.00 & 0.00                           & 0.00 & 0.00                               & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    lastfm           & cos              & 292,385            & 65             & 100         & 0.00 & 0.00                           & 0.00 & 0.00                             & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00                               & 0.00 & 0.00                                                    & 1.00 & 42 \\
     \hline
@@
     \end{table*}
+
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$}
+    \label{table:results:ann-100}
+    \vskip 0.15in
+    \begin{center}
+    \begin{small}
+    \begin{sc}
+    \begin{tabular}{|l|c|c|c|c|c|c|c|c|}
+    \textbf{Dataset}  & \multicolumn{2}{|c|}\textbf{faiss-ivf} & \multicolumn{2}{|c|}\textbf{bruteforce-blas} & \multicolumn{2}{|c|}\textbf{hnsw(faiss)} & \multicolumn{2}{|c|}\textbf{CAKES} \\
+    &                    Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
+    \hline
+    nytimes                & 0.61 &  0.71                          & 0.99 & 10.71                              & 0.00 & 0.00                                          & 1.00 & 42 \\
+    \hline
+    fashion-mnist         & 1.00 & 111.47                           & 0.99 & 16.02                                  & 0.99 & 277.80                                                    & 1.00 & 42 \\ 
+    \hline
+    gist                   & 1.00 & 2.70                           & 0.99 & 0.80                                 & 0.99 & 59.71                                                    & 1.00 & 42 \\
+    \hline
+    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & 1.00 & 42 \\
+    \hline
+    glove-100             & 0.99 &  7.52                          & 0.99 & 3.11                                  & 0.99 & 92.58                                                    & 1.00 & 42 \\
+    \hline
+    sift                  & 1.00 &  23.50                          & 1.00 & 6.51                                  & 0.99 & 179.14                                                    & 1.00 & 42 \\
+    \hline
+    deep-image             & 0.00 & 0.00                           & 0.00 & 0.00                                & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    mnist                   & 0.00 & 0.00                          & 0.00 & 0.00                                  & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    glove-50               & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00                             & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    \hline
+    \end{tabular}
+    \end{sc}
+    \end{small}
+    \end{center}
+    \vskip -0.1in
+    \end{table*}

### 2023-09-02 sections/4-results.tex
update hw specs

@@
 
-Ark: a 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM and CentOS 7 Linux with kernel 5.15.50-1-MANJARO \#1 SMP PREEMPT. Rust 1.62, Python 3.10.5
+Ark: a 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
 
-Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 12.1. Rust 1.62, Python 3.9.4
+Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 12.1. Rust 1.70, Python 3.9.4
+
+Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12
 

### 2023-09-02 sections/2-methods.tex
minor explanation

@@
 Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data 
-are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} related to the 
+are relatively concentrated in that region; in other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radius increase is needed.
+Thus, in this case, the factor of radius increase should be \emph{directly} related to the
 local fractal dimension. Once $|H| >= k$, we return the $k$ closest points.

### 2023-09-02 sections/2-methods.tex
consistency of notation

@@
 the cluster radius. Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in
-a binary tree has two child nodes. We define the \emph{metric entropy} of a data set under a hierarchical clustering scheme as a refinement of [7], where
+a binary tree has two child nodes. We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(D)$ of a data set $D$ under a hierarchical clustering scheme as a refinement of [7], where
 metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a hierarchical, divisive clustering 
@@
 
-In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log{2}m)$, 
-where $m$ is the metric entropy of the dataset. 
+In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log{2}$\mathcal{N}_{\hat{r_c}}(D)$)$, 
+where $m$\mathcal{N}_{\hat{r_c}}(D)$$ is the metric entropy of the dataset $D$. 
 To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
@@
 Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has 
-$\mathcal{O}\left(\log_2{m} + k\left(\frac{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}+ 2r_c}{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}}\right)^d\right)$ 
-time complexity, where $m$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
+$\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(\frac{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}+ 2r_c}{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}}\right)^d\right)$ 
+time complexity, where $\mathcal{N}_{\hat{r_c}}(D) is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
 near the query ball, $d_i$ is the local fractal dimension of $C_i$, $r_c$ is the radius of leaf clusters, and $k$ is the number of nearest neighbors.

### 2023-09-02 sections/2-methods.tex
fix algorithm ref
temp

@@
 
-The asympotitic complexity of Partition is the same as described in~\cite{ishaq2019clustered}
+The asymptotic complexity of Cluster Partition is the same as described in~\cite{ishaq2019clustered}
 By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of 
@@
 We complete this process until $Q$ contains only points. At this point, we 
-just use Algorithm 4 to find the $k$th 
+just use the Partition algorithm to find the $k$th 
 nearest neighbor and return all points to the left of it.
@@
 
+We group the Sieve-based algorithms together, as their complexity analyses are similar.
+
+The complexity of Sieve Search is dominated by the complexity of our modified Quicksort Partition algorithm,
+which, like the original Quicksort partition algorithm described in~\cite{10.1093/comjnl/5.1.10}, has $\mathcal{O}(n)$ complexity, 
+where $n$ is the length of the input list. While the cardinality of clusters is greater than $k$, 2 is bounded 
+above by 2, as $Q$ will contain at most two clusters (assuming the worst case, which is a balanced clustering). 
+Once the cardinality of clusters in $Q$ is less than or equal to $k$, the cardinality of $Q$ is bounded above 
+by $2k$, since, in the worst case, every cluster is a non-leaf and
+
 % Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 

### 2023-09-02 sections/2-methods.tex
misc changes

@@
 
+\subsection{Clustering}
+\label{subsec:methods:clustering}
+
 We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
@@
 leaf clusters.
-
-With these concepts defined, we can now pose the $k$- and $\rho$- nearest neighbors search problems.
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $f(x, q) \leq \max\{f(y, q): y \in S_q \}$; that is,
-for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
-We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
-find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
-
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following cluster 
-properties:
-\begin{itemize}
-    \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
-    \item $\delta_{max} = \delta + r$ is the distance from the query to the theoretically farthest possible instance in $C$.
-    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest possible instance in $C$.
-\end{itemize}
-
-We define \emph{singletons} as clusters which either contain a single point (i.e., have cardinality 1) or which contain 
-many instances of the same point (i.e., have cardinality greater than 1 but contain only one unique point). Singletons clearly 
-have zero radius, and so $\delta_{max}  = \delta = \delta_{min}$ for these clusters. Hence, we sometimes overload the above 
-notation to refer to distance from a query to a point; in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
-
-\subsection{Clustering}
-\label{subsec:methods:clustering}
 
@@
 repeat this procedure until each leaf contains only one datum or some other user-specified stopping criterion 
-is met.
-
-This process is described in Algorithm \ref{alg:partition}.
+is met. This process is described in Algorithm \ref{alg:partition}.
 
@@
 will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes. 
+See Figure 1 for a concrete example.
 
@@
 $k$-nearest neighbors search using one of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
-The result of this search will give us the $k$ closest neighbors to the query \emph{in the first shard}. We then take the furthest of 
-those $k$ neighbors and use its distance from the query as the radius for use with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) 
+The result of this search will give us the $k$ closest neighbors to the query \emph{in the first shard}. We take the furthest of 
+those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) 
 on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
@@
 since the first shard is the only one for which we perform $k$-nearest neighbors search. However, making this first shard too small
-introduces the risk of significantly overestimating the radius, resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
+introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th ), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
 Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well 
@@
 actual $k$-th closest neighbor, are topics for future work.
+
+\subsection{The Search Problem}
+\label{subsec:methods:general-search}
+Having explained the clustering process, we can now pose the $k$- and $\rho$- nearest neighbors search problems.
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $f(x, q) \leq \max\{f(y, q): y \in S_q \}$; that is,
+for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
+We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
+find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
+
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following cluster 
+properties:
+\begin{itemize}
+    \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
+    \item $\delta_{max} = \delta + r$ is the distance from the query to the theoretically farthest possible instance in $C$.
+    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest possible instance in $C$.
+\end{itemize}
+
+We define a \emph{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains 
+multiple instances of the same point (i.e., has cardinality greater than 1 but contains only one unique point). A singleton clearly 
+has zero radius, and so $\delta_{max}  = \delta = \delta_{min}$ for these clusters. Hence, we sometimes overload the above 
+notation to refer to distance from a query to a point; in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
 
@@
 always proceeding with both of its children, we proceed only with those children which 
-also overlap with the query ball. This improvement is demonstrated in Algorithm \ref{alg:rnn}.
+also overlap with the query ball. This improvement is described in Algorithm \ref{alg:rnn}.
 
@@
 the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
-If no points are within a distance $m$ of the query, we increase the radius by a factor of arg
+If no points are within a distance $m$ of the query, we increase the radius by a factor of
 2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
@@
 we increase the radius by a factor of 
-\begin{equation} \argmin \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
+\begin{equation} \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
 where $\mu$ is the harmonic mean of the local fractal dimensions (abbreviated lfd in algorithm below) of 
@@
 Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data 
-are relatively concentrated in that region; in other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radius increase is needed.
-Thus, in this case, the factor of radius increase should be \emph{directly} related to the
+are relatively concentrated in that region; in other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radial increase is needed.
+Thus, the factor of radius increase should be \emph{directly} related to the
 local fractal dimension. Once $|H| >= k$, we return the $k$ closest points.
@@
             \STATE $\mu \Leftarrow |Q| \sum_{C \in Q} \frac{1}{C.lfd} $
-            \STATE $m \Leftarrow \argmin \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
+            \STATE $m \Leftarrow \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
             \STATE $H.push(\rho$-NN$(root, query, m)$)
@@
 Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
-(i.e. arises from the same generating phenomenon) as the rest of the data. From this assumption, we can infer that the local fractal dimension
-at the query point is not significantly different from the mean local fractal dimension of clusters nearby the query.
+(i.e. arises from the same generating phenomenon) as the rest of the data. Given the uses of $k$-NN search in practice, 
+this assumption is not unreasonable. From this assumption, we can infer that the local fractal dimension
+at the query point does not differ significantly from the mean local fractal dimension of clusters nearby the query.
 
@@
 
-In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log{2}$\mathcal{N}_{\hat{r_c}}(D)$)$, 
-where $m$\mathcal{N}_{\hat{r_c}}(D)$$ is the metric entropy of the dataset $D$. 
+In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log_2\mathcal{N}_{\hat{r_c}}(D))$, 
+where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset $D$. 
 To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
@@
 Since this is a constant factor, it does not affect the asymptotic complexity of coarse search, and thus we have that 
-the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2{m})$.
+the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2\mathcal{N}_{\hat{r_c}}(D))$.
 
@@
 
-~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$, where $\gamma$ is a constant. 
+The work in~\cite{yu2015entropy} showed that \begin{equation} |F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d \label{3} \end{equation} where $\gamma$ is a constant. 
 By definition of 
-$r'$, we have that $|B(q, r')| = k$. Thus, we have that $|F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d$.
-
-It remains to provide an estimate for $r'$. 
+$r'$, we have that $|B(q, r')| = k$. Thus, we have that \begin{equation} |F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d. \label{4} \end{equation} It remains to provide an estimate for $r'$. 
 To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, 
@@
 
-By ~\ref{2}, we then have that \begin{equation} d_i = \frac{\log{}\frac{|C_i|}{k}}{\log{}\frac{r_i}{r'}}. \label{4} \end{equation}
+By ~\ref{2}, we then have that \begin{equation} d_i = \frac{\log{}\frac{|C_i|}{k}}{\log{}\frac{r_i}{r'}}. \label{5} \end{equation}
 
@@
 $\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(\frac{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}+ 2r_c}{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}}\right)^d\right)$ 
-time complexity, where $\mathcal{N}_{\hat{r_c}}(D) is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
+time complexity, where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
 near the query ball, $d_i$ is the local fractal dimension of $C_i$, $r_c$ is the radius of leaf clusters, and $k$ is the number of nearest neighbors.
@@
 \label{subsubsec:methods:knn-search:sieve}
-With Sieve Search, we begin by letting $Q$ be a list containing the root cluster. 
+In this algorithm, we begin by letting $Q$ be a list containing only the root cluster. 
 While $Q$ contains at least one cluster (i.e., $Q$ is not a list of only points), we repeat the process described in 
@@
 We then remove from $Q$ any element whose $\delta_{min}$ is greater than $\tau$. Then, for each element left 
-in $Q$, if it is a leaf or contains $k$ or fewer points, we remove it and all of its points to $Q$. Finally, we
+in $Q$, if it is a leaf or contains $k$ or fewer points, we remove it and add all of its points to $Q$. Finally, we
 replace all remaining non-leaf clusters in $Q$ with their children.
@@
 \label{subsubsec:methods:knn-search:sieve2}
-Sieve Search with Separate Centers is the same as Sieve Search, but with the following modification: clusters 
+This algorithm is the same as Sieve Search, but with the following modification: clusters 
 in $Q$ are represented twice-- once as their center and once as the rest of their points. 
@@
 
-With Greedy Sieve, we first let $Q$ be a min queue of clusters based on $\delta_{min}$. We initialize $Q$ with the root cluster.
-Let $H$ be a fixed-size ($k$) max queue of points based on $\delta$. This queue starts empty.
+With Greedy Sieve, we first let $Q$ be a min queue of clusters prioritized by $\delta_{min}$. We initialize $Q$ with the root cluster.
+Let $H$ be a fixed-size ($k$) max queue of points prioritized by $\delta$. This queue starts empty.
 
@@
 \begin{enumerate}
-\item While the top priority element $q$ in $Q$ is not a leaf, pop $q$ from $Q$ and push its children.
+\item While the top priority element $q$ in $Q$ is not a leaf, pop $q$ from $Q$ and push its children to $Q$.
 \item For each point $p \in q$, push $p$ to $H$. 
@@
 \end{enumerate}
-After this process, the points left in $H$ are the $k$ nearest neighbors of $q$.
-
-This process is described in Algorithm \ref{alg:greedy_search}.
+After this process, the points left in $H$ are the $k$ nearest neighbors of $q$. This algorithm is described in Algorithm \ref{alg:greedy_search}.
 
 \begin{algorithm} 
-\caption{GreedySearch(\emph{root, query, k})} 
+\caption{GreedySieve(\emph{root, query, k})} 
 \label{alg:greedy_search} 
@@
     \STATE $H \Leftarrow$ priority queue
-    \WHILE{$|H| < k \lor H.pop.\delta > Q.pop.\delta_{min}$}
-        \WHILE{$\neg Q.pop.isLeaf$}
+    \WHILE{$|H| < k \lor H.peek.\delta > Q.peek.\delta_{min}$}
+        \WHILE{$\neg Q.peek.isLeaf$}
             \STATE $[l, r] \Leftarrow Q.pop().children$
@@
 We group the Sieve-based algorithms together, as their complexity analyses are similar.
-
-The complexity of Sieve Search is dominated by the complexity of our modified Quicksort Partition algorithm,
-which, like the original Quicksort partition algorithm described in~\cite{10.1093/comjnl/5.1.10}, has $\mathcal{O}(n)$ complexity, 
-where $n$ is the length of the input list. While the cardinality of clusters is greater than $k$, 2 is bounded 
-above by 2, as $Q$ will contain at most two clusters (assuming the worst case, which is a balanced clustering). 
-Once the cardinality of clusters in $Q$ is less than or equal to $k$, the cardinality of $Q$ is bounded above 
-by $2k$, since, in the worst case, every cluster is a non-leaf and
 

### 2023-09-02 sections/2-methods.tex
updates to rho nn complexity

@@
     \overbrace{\left|B_D(q,r)\right|}^{\textrm{output size}}
-    \underbrace{\left(\frac{r+2\hat{r_c}}{r}\right)^d}_{\textrm{scaling factor}}
+    \underbrace{\left(1+\frac{2\hat{r_c}}{r}\right)^d}_{\textrm{scaling factor}}
      \Bigg)
@@
 number of clusters for which we must look at all of the cluster's points. With $\rho$-NN 
-search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
-where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
+search, $F$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query,
+where $r$ is the search radius and $\hat{r_c}$ is the mean radius of leaf clusters,
 as in \cite{yu2015entropy}. With $k$-NN by Repeated $\rho$-NN, however, 
@@
 is query-dependent. Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of 
-$F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
-where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.
+$F$ to be the union of clusters which are within a distance of $r' + \hat{r_c}$ from the query, 
+where $r'$ is the distance to the $k$th nearest neighbor and $\hat{r_c}$ is defined as before.
 
-The work in~\cite{yu2015entropy} showed that \begin{equation} |F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d \label{3} \end{equation} where $\gamma$ is a constant. 
+The work in~\cite{yu2015entropy} showed that \begin{equation} |F| \leq \gamma |B(q, r')|\left(1+ \frac{2\hat{r_c}}{r'}\right)^d \label{3} \end{equation} where $\gamma$ is a constant. 
 By definition of 
-$r'$, we have that $|B(q, r')| = k$. Thus, we have that \begin{equation} |F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d. \label{4} \end{equation} It remains to provide an estimate for $r'$. 
+$r'$, we have that $|B(q, r')| = k$. Thus, we have that \begin{equation} |F| \leq \gamma k\left(1+ \frac{2\hat{r_c}}{r'}\right)^d. \label{4} \end{equation} It remains to provide an estimate for $r'$. 
 To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, 
@@
 
-We let $d_i$ be the local fractal dimension of some cluster $C_i$ near the query point (i.e., some cluster identified during coarse search).
+We let $\hat{d}$ be the mean of the local fractal dimensions of the clusters near the query point (i.e., the cluster identified during coarse search).
 While ordinarily we compute local fractal dimension 
 by comparing cardinalities of two balls centered at \emph{the same} point at two different radii, in order to estimate $r'$, 
-we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of cluster $C_i$} at its radius $r_i$.
-We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could "drag" the query to 
-the center of $C_i$ without significantly changing the local fractal dimension at the query point.
+we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the mean cardinality $\hat{C}$ of clusters near the query 
+at a radius equal to the mean of their radii $\hat{r}$.
+We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we replace the query with 
+the center of one of the clusters near it without significantly changing the local fractal dimension at the query point.
 
-By ~\ref{2}, we then have that \begin{equation} d_i = \frac{\log{}\frac{|C_i|}{k}}{\log{}\frac{r_i}{r'}}. \label{5} \end{equation}
+By ~\ref{2}, we then have that \begin{equation} \hat{d} = \frac{\log{}\frac{\hat{C}}{k}}{\log{}\frac{\hat{r}}{r'}}. \label{5} \end{equation}
 
-Since $d_i$, $|C_i|$, $r_i$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
-\begin{equation} r' = r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}. \label{5} \end{equation}
+Since $\hat{D}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
+\begin{equation} r' = \hat{r}\left(\frac{k}{\hat{C}}\right)^{\frac{1}{\hat{d}}}. \label{6} \end{equation}
 
 Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has 
-$\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(\frac{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}+ 2r_c}{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}}\right)^d\right)$ 
-time complexity, where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
-near the query ball, $d_i$ is the local fractal dimension of $C_i$, $r_c$ is the radius of leaf clusters, and $k$ is the number of nearest neighbors.
+$\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(1+ \frac{2\hat{r_c}}{\hat{r}}\left(\frac{k}{\hat{C}}\right)^{\frac{-1}{\hat{d}}}\right)^d\right) = 
+\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(\frac{2\hat{r_c}}{\hat{r}}\right)^d\left(\frac{k}{\hat{C}}\right)^{\frac{-d}{\hat{d}}}\right)$ 
+time complexity, where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, and $k$ is the number of nearest neighbors.
 

### 2023-09-03 sections/2-methods.tex
new geometry caption

@@
     \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
-    \caption{Scaling Behavior of Radii in a Two-Dimensional Uniform-Density Disk. $R$ 
-    denotes the radius of the root cluster $C$, and $o$ denotes its center. The right 
-    triangle formed by $o$, $o'$, and $y+$ in $C'$ shows that $R \leq R'$, where $R'$ is the radius of $C'$.
-    Hence, the radius of a child cluster can be larger than its parent. $C''$ 
-    illustrates that after we have consumed all $d$ axes, the radius of a child cluster
-    is bounded above by $\frac{R}{\sqrt{2}}$.}
+    \caption{Scaling behavior of radii in the worst case two-dimensional scenario, 
+    a uniform-density disk. $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
+    After one application of Algorithm~\ref{1}, we have $C_1$, with radius $R_1$ and center $o_1$.
+    The right triangle formed by $o_0$, $o_1$, and $y_+$ in $C_1$ shows that $R_0 \leq R_1$.
+    Hence, the radius of a child cluster can be larger than its parent. However, 
+    after another application of Algorithm~\ref{1}, we have consumed all $d$ orthogonal axes, 
+    as shown in $C_2$. Now, clearly $R_2 \leq R_0$. In fact, for each subsequent application of Algorithm~\ref{1},
+    the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.}
     \label{fig:methods:scaling_behavior}

### 2023-09-03 sections/2-methods.tex
kill smallcaps

@@
     \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
-    \caption{Scaling behavior of radii in the worst case two-dimensional scenario, 
-    a uniform-density disk. $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
+    \caption{Scaling behavior of radii on a uniform-density disk, which in two dimensions represents the worst-case scenario. 
+    $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
     After one application of Algorithm~\ref{1}, we have $C_1$, with radius $R_1$ and center $o_1$.
@@
     after another application of Algorithm~\ref{1}, we have consumed all $d$ orthogonal axes, 
-    as shown in $C_2$. Now, clearly $R_2 \leq R_0$. In fact, for each subsequent application of Algorithm~\ref{1},
+    as shown in $C_2$. 
+    Now, clearly $R_2 \leq R_0$. In fact, for each subsequent application of Algorithm~\ref{1},
     the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.}
@@
 search, $F$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query,
-where $r$ is the search radius and $\hat{r_c}$ is the mean radius of leaf clusters,
+where $r$ is the search radius, and $\hat{r_c}$ is the mean radius of leaf clusters,
 as in \cite{yu2015entropy}. With $k$-NN by Repeated $\rho$-NN, however, 

### 2023-09-03 sections/2-methods.tex
results

@@
 
-\subsection {Index-Building and Sharded Search}
+\subsection {Sharding and Auto-Tuning}
 \label{subsec:methods:index-building}
-In this work, we also introduce the method of \emph{sharding} for $k$-nearest neighbors search.
-With this approach, we aim to leverage the fact that $\rho$-nearest neighbors search is typically 
-much faster than $k$-nearest neighbors search.
+In this work, we also introduce the method of \emph{sharding} for datasets that cannot 
+fit in memory. With this approach, we begin by
 
-We begin by randomly partitioning the dataset into $s$ shards, where $s$ is determined based 
-on some naive auto-tuning. On the first shard, we perform 
-$k$-nearest neighbors search using one of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
-The result of this search will give us the $k$ closest neighbors to the query \emph{in the first shard}. We take the furthest of 
-those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) 
-on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
-from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
+% In this work, we also introduce the method of \emph{sharding} for $k$-nearest neighbors search.
+% With this approach, we aim to leverage the fact that $\rho$-nearest neighbors search is typically 
+% much faster than $k$-nearest neighbors search.
 
-Note that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others,
-since the first shard is the only one for which we perform $k$-nearest neighbors search. However, making this first shard too small
-introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th ), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
-Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well 
-as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the 
-actual $k$-th closest neighbor, are topics for future work.
+% We begin by randomly partitioning the dataset into $s$ shards, where $s$ is determined based 
+% on some naive auto-tuning. On the first shard, we perform 
+% $k$-nearest neighbors search using one of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
+% The result of this search will give us the $k$ closest neighbors to the query \emph{in the first shard}. We take the furthest of 
+% those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) 
+% on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
+% from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
+
+% Note that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others,
+% since the first shard is the only one for which we perform $k$-nearest neighbors search. However, making this first shard too small
+% introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th ), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
+% Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well 
+% as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the 
+% actual $k$-th closest neighbor, are topics for future work.
 

### 2023-09-03 sections/4-results.tex
misc changes
results
scientific notation is for nerds.
fix table alignment

@@
 \label{sec:results}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{images/result_plots/deep-image_1.png}
+    \caption{Performance of CAKES algorithms on deep-image dataset.}
+    \label{fig:methods:deep-image}
+\end{figure}
 
@@
     \begin{sc}
-    \begin{tabular}{|l|c|c|c|c|c|c|c|c|}
-    \textbf{Dataset}  & \multicolumn{2}{|c|}\textbf{faiss-ivf} & \multicolumn{2}{|c|}\textbf{bruteforce-blas} & \multicolumn{2}{|c|}\textbf{hnsw(faiss)} & \multicolumn{2}{|c|}\textbf{CAKES} \\
+    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
     \cline{2-9}
@@
     \hline
-    nytimes                & 0.69 &  1413.35                       & 0.99 & 26.66                            & 0.97 & 92.92                                         & 1.00 & 42 \\
+    fashion-mnist      & 0.54 & 324.79                           & 1.00 & 52.91                                  & 0.94 & 60.89                                                    & 1.00 & 16,220 \\
     \hline
-    fashion-mnist      & 0.54 & 324.79                           & 1.00 & 52.91                                  & 0.94 & 60.89                                                    & 1.00 & 42 \\
+    gist                   & 1.00 & 3.55                           & 1.00 & 2.63                                     & 0.95 & 149.58                                                   & 1.00 & 10,600 \\
     \hline
-    gist                   & 1.00 & 3.55                           & 1.00 & 2.63                                     & 0.95 & 149.58                                                   & 1.00 & 42 \\
+    glove-25              & 1.00 & 78.78                          & 1.00 & 33.52                              & 1.00 & 552.67                                                  & -- & 19,520 \\
     \hline
-    glove-25              & 1.00 & 78.78                          & 1.00 & 33.52                              & 1.00 & 552.67                                                  & 1.00 & 42 \\
+    glove-100             & 1.00 & 21.16                          & 1.00 & 16.52                                & 0.99 & 155.90                                                  & -- & 6,935 \\
     \hline
-    glove-100             & 1.00 & 21.16                          & 1.00 & 16.52                                & 0.99 & 155.90                                                  & 1.00 & 42 \\
-    \hline
-    sift                  & 1.00 &  24.52                          & 1.00 & 16.40                               & 1.00 & 275.07                                                    & 1.00 & 42 \\
-    \hline
-    deep-image             & 0.00 & 0.00                           & 0.00 & 0.00                                 & 0.00 & 0.00                                                     & 1.00 & 42 \\
-    \hline
-    mnist                   & 0.00 & 0.00                          & 0.00 & 0.00                                 & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    glove-50               & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00                               & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    sift                  & 1.00 &  24.52                          & 1.00 & 16.40                               & 1.00 & 275.07                                                    & 1.00 & 31,800 \\
     \hline
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$}
+    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$. Other methods did not report results for glove-25.}
     \label{table:results:ann-100}
@@
     \begin{sc}
-    \begin{tabular}{|l|c|c|c|c|c|c|c|c|}
-    \textbf{Dataset}  & \multicolumn{2}{|c|}\textbf{faiss-ivf} & \multicolumn{2}{|c|}\textbf{bruteforce-blas} & \multicolumn{2}{|c|}\textbf{hnsw(faiss)} & \multicolumn{2}{|c|}\textbf{CAKES} \\
+    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
     &                    Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
     \hline
-    nytimes                & 0.61 &  0.71                          & 0.99 & 10.71                              & 0.00 & 0.00                                          & 1.00 & 42 \\
+    fashion-mnist         & 1.00 & 111.47                           & 0.99 & 16.02                                  & 0.99 & 277.80                                                    & 1.00 & 1,549 \\
     \hline
-    fashion-mnist         & 1.00 & 111.47                           & 0.99 & 16.02                                  & 0.99 & 277.80                                                    & 1.00 & 42 \\ 
+    gist                   & 1.00 & 2.70                           & 0.99 & 0.80                                 & 0.99 & 59.71                                                    & 1.00 & 753.9 \\
     \hline
-    gist                   & 1.00 & 2.70                           & 0.99 & 0.80                                 & 0.99 & 59.71                                                    & 1.00 & 42 \\
+    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & -- & 1,471 \\
     \hline
-    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & 1.00 & 42 \\
+    glove-100             & 0.99 &  7.52                          & 0.99 & 3.11                                  & 0.99 & 92.58                                                    & -- & 396.0 \\
     \hline
-    glove-100             & 0.99 &  7.52                          & 0.99 & 3.11                                  & 0.99 & 92.58                                                    & 1.00 & 42 \\
-    \hline
-    sift                  & 1.00 &  23.50                          & 1.00 & 6.51                                  & 0.99 & 179.14                                                    & 1.00 & 42 \\
-    \hline
-    deep-image             & 0.00 & 0.00                           & 0.00 & 0.00                                & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    mnist                   & 0.00 & 0.00                          & 0.00 & 0.00                                  & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    glove-50               & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
-    \hline
-    lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00                             & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    sift                  & 1.00 &  23.50                          & 1.00 & 6.51                                  & 0.99 & 179.14                                                    & 1.00 & 2,852 \\
     \hline
@@
     \end{table*}
+
+    % \hline
+    % deep-image             & 0.00 & 0.00                           & 0.00 & 0.00                                 & 0.00 & 0.00                                                     & 1.00 & 42 \\
+    % \hline
+    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00                                 & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    % \hline
+    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    % \hline
+    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00                               & 0.00 & 0.00

### 2023-09-03 sections/3-datasets.tex
move benchmarking specs

@@
-\section{Datasets}
+\section{Datasets And Benchmarking}
 \label{sec:datasets-and-distance-functions}
@@
 We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
-listed below, along with the appropriate distance function.
+along with the appropriate distance function.
 Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
+
+Our systems used for benchmarking were as follows:
+
+\emph{(I don't think we have any macbook benchmarks, right?)}
+
+Ark: 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
+
+Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 12.1. Rust 1.70, Python 3.9.4
+
+Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12
 

### 2023-09-03 sections/4-results.tex
move benchmarking specs

@@
 see longer lower tails than upper tails for many of these algorithms.
-
-Specs of Ark, and maybe M1 MacBook Air and Dell XPS.
-
-Ark: a 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
-
-Macbook Air: 8-core (4 performance, 4 efficiency) Apple M1 processor with 16GB RAM, macOS 12.1. Rust 1.70, Python 3.9.4
-
-Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12
 

### 2023-09-03 sections/2-methods.tex
updates to sharding and auto-tuning

--- 2023-09-03 sections/2-methods.tex
@@
 
-\subsection {Index-Building and Sharded Search}
+\subsection {Sharding and Auto-Tuning}
 \label{subsec:methods:index-building}
-In this work, we also introduce the method of \emph{sharding} for $k$-nearest neighbors search.
-With this approach, we aim to leverage the fact that $\rho$-nearest neighbors search is typically 
-much faster than $k$-nearest neighbors search.
+In this work, we also introduce the method of \emph{sharding} for datasets that cannot fit in memory. 
+This approach also leverages the fact that $\rho$-nearest neighbors search is typically much faster than $k$-nearest neighbors search.
+In addition to sharding, we perform some naive auto-tuning to select the optimal $k$-NN algorithm to be used with a given dataset.
 
-We begin by randomly partitioning the dataset into $s$ shards, where $s$ is determined based 
-on some naive auto-tuning. On the first shard, we perform 
-$k$-nearest neighbors search using one of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
-The result of this search will give us the $k$ closest neighbors to the query \emph{in the first shard}. We take the furthest of 
-those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) 
-on each remaining shard. We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors
-from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
+With this approach, we begin by taking a random sample of up to 10\% of the data. 
+The exact proportion of the dataset in this random sample depends on the dimensionality of the dataset and a user-specified memory limit. 
+FIrst, we build a cluster tree on this sample, and 
+then consider all clusters at some user-specified depth in the tree. Using the cluster centers 
+of these clusters as queries, and a user-specified value of $k$, we perform $k$-NN search using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}. 
+We record the 
+time taken for each algorithm and select the one which had the lowest total search time over all the queries. 
+We then split the rest of the dataset into some user-specified number of shards, and use the algorithm identified in the previous step to perform $k$-NN search on the first shard. 
+This search finds the $k$ closest neighbors to the query \emph{in the first shard}. 
+We take the furthest of those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) on each remaining shard.
+We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
 
-Note that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others,
-since the first shard is the only one for which we perform $k$-nearest neighbors search. However, making this first shard too small
-introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th ), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
-Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well 
-as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the 
-actual $k$-th closest neighbor, are topics for future work.
+Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we use the selected algorithm to perform search on the given dataset for any value of $k$.
+Note also that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search.
+However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th ), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
+Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
+
+For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but with only one shard.
 
@@
 
-where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy at that radius, $B_D(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the factal dimension around the query in a region of radius $r$.
+where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy at that radius, $B_D(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the fractal dimension around the query in a region of radius $r$.
 
@@
 
-\subsubsection{$k$-NN by Repeated $\rho$-NN}
+\subsubsection{\texorpdfstring{$k$}{k}-NN by Repeated \texorpdfstring{$\rho$}{p}-NN}
 \label{subsubsec:methods:knn-search:repeated-rnn}
@@
 
-\subsubsection{Complexity of $k$-NN by Repeated $\rho$-NN Search}
-\label{paragraph:methods:repeated-rnn-complexity}
+\subsubsection{Complexity of \texorpdfstring{$k$}{k}-NN by Repeated \texorpdfstring{$\rho$}{p}-NN Search}
+\label{subsubsec:methods:repeated-rnn-complexity}
 
@@
 center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, 
-thus allowing us to eliminate some clusters from $Q$ earlier in the process. More discussion about 
-when this algorithm is preferable to Sieve Search is in Section 4: Results.
+thus allowing us to eliminate some clusters from $Q$ earlier in the process.
 

### 2023-09-03 sections/2-methods.tex
adding complexity for knn sieves

@@
 
-% Our $k$-NN algorithms do not use two completely distinct phases of coarse and fine search, 
-% as is the case in $\rho$-NN search. Still, this distinction is useful in analyzing 
-% the asymptotic complexity of $k$-NN search. In particular, we can think of the complexity of 
-% coarse search as an estimate of how many clusters we examine during search, and 
-% the complexity of fine search as an estimate of how many clusters we must examine all the points 
-% from during search, even if we do not actually perform these as two distinct phases.
+The complexity of the Sieve-based algorithms relies on the assumption that when the cardinalities of clusters is sufficiently close to $k$, the distribution of clusters will mimic the distribution of points in the dataset.
+From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query point.
+In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, as at each end of each of the $d$ axes, we could have a cluster which overlaps the query ball.
 
-% To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
-% number of clusters we must examine all the points from during search. With $\rho$-NN 
-% search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
-% where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
-% as in \cite{yu2015entropy}. With $k$-NN search, however, 
-% the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
-% is query-dependent. Thus, for $k$-NN search, we  adjust the definition of 
-% $F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
-% where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.
-
-% By the triangle inequality, we have that $F \subseteq B(q, r' + 2r_c)$, 
-% where $B(q, r' + r_c)$ is the ball of radius $r' + r_c$ centered at $q$.
-% Hence, $|F| \leq |B(q, r' + 2r_c)|$.
-
-% Now we have that \begin{align*} \mathbb{E}_q[|B(q, r' + 2r_c)|] &\leq \gamma|B(p, r' + 2r_c)| \\ %because of the assumption that density at different locations on the manifold differ by only a constant factor%
-%     &\leq \gamma|B(p, r')|\left(\frac{r' + 2r_c}{r'}\right)^d \label{5}\\ %because d = lfd tells us how the number of additional hits reached scales with a change in radius, and our ratio between radii is (r' + 2r_c)/r'%
-%     &\leq \gamma^2 \mathbb{E}[|B(q, r')|]\left(\frac{r'+2r_c}{r'}\right)^d \\ %using density bound again to go from a statement about an arbitrary point back to a statement about the query%
-%     &\leq \gamma^2 k\left(\frac{r'+ 2r_c}{r'}\right)^d  \\   %using the fact that for k-nn, the number points in the query ball is k%
-% \end{align*}
+In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log_2{\frac{n}{k}}$ iterations of Sieve to reduce the cardinality of clusters to at most $k$.

### 2023-09-03 sections/5-discussion.tex
changes to rnn
comment about poor recall on cosine

@@
 Exactness of search and metric vs non-metric.
+
+While CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
+This bears further investigation, but likely stems from how heavily the knn and clustering algorithms depend on the triangle inequality.
 
@@
 For example, we would like to implement Tanimoto Distance using maximal-common-subgraph for molecular structures. Ultimately, 
-we would also like to implement Wasserstein distance for dimensions higher than 2. In order to make this possible, 
+we would also like to implement Wasserstein distance for dimensions higher than two. In order to make this possible, 
 we plan to implement GPU-acceleration for the more computationally-intensive distance functions like Wasserstein.

### 2023-09-03 sections/2-methods.tex
bad recalls
explain frac dim fraction
fix to sharding

@@
 
-With this approach, we begin by taking a random sample of up to 10\% of the data. 
+With this approach, we begin by taking a random sample of up to 10\% of the data. This sample 
+is considered the first shard. 
 The exact proportion of the dataset in this random sample depends on the dimensionality of the dataset and a user-specified memory limit. 
@@
 time taken for each algorithm and select the one which had the lowest total search time over all the queries. 
-We then split the rest of the dataset into some user-specified number of shards, and use the algorithm identified in the previous step to perform $k$-NN search on the first shard. 
+We then split the rest of the dataset into shards, based on the user-defined memory limit, and use the algorithm identified in the previous step to perform $k$-NN search on the first shard. 
 This search finds the $k$ closest neighbors to the query \emph{in the first shard}. 
 We take the furthest of those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) on each remaining shard.
+
 We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
@@
 
-For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but with only one shard.
+For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but, since the entire remainder of the dataset fits in memory, we use everything 
+but that initial 10\% sample as the remaining shard for $\rho$-nearest neighbors search.
 
@@
 time complexity, where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, and $k$ is the number of nearest neighbors.
+We note that the fraction $\frac{-d}{\hat{d}}$ should, unless fractal dimension is highly variable (that is, local $d$ is quite different from $\hat{d}$), be relatively close to 1.
 

### 2023-09-03 sections/1-introduction.tex
merge conflict

@@
 million genomes with aligned length of 50,000 letters; 
-the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences.
+the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences. [tTODO: add dimensionality here]
 
-More recently, Big-ANN's NeurIPS'23: Billion Scale Approximate Nearest Neighbor Search Challenge has 
-solicits development of data structures and search algorithms for Approximate Nearest Neighbor (ANN) search
-on four massive datasets: YFCC 100M~\cite{johnson2019billion} dataset, the Yandex Text-to-Image 10M~\cite{diskann-github}, cross-modal dataset, 
-the MSMARCO passage retrieval dataset~\cite{Bruch2023AnAA}, and a 10M slice of the MS Turing data set~\cite{Singh2021FreshDiskANNAF}. These datasets 
-are 192, 200, 30,000, and 100 dimensional respectively, and submissions are evaluated with 
-a ten million point sample of each.
+% More recently, Big-ANN's NeurIPS'23: Billion Scale Approximate Nearest Neighbor Search Challenge has 
+% solicits development of data structures and search algorithms for Approximate Nearest Neighbor (ANN) search
+% on four massive datasets: YFCC 100M~\cite{johnson2019billion} dataset, the Yandex Text-to-Image 10M~\cite{diskann-github}, cross-modal dataset, 
+% the MSMARCO passage retrieval dataset~\cite{Bruch2023AnAA}, and a 10M slice of the MS Turing data set~\cite{Singh2021FreshDiskANNAF}. These datasets 
+% are 192, 200, 30,000, and 100 dimensional respectively, and submissions are evaluated with 
+% a ten million point sample of each.
 
 In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating 
-that computer architecture will not "catch up" to our computational needs in the near future. As a result, analysis of these large datasets 
+that computer architecture will not ``catch up'' to our computational needs in the near future. As a result, analysis of these large datasets 
 requires better algorithms.
 
-Many researchers are especially interested in performing similarity search on these datasets. 
+Many researchers are especially interested in similarity search on these datasets. 
 Similarity search enables a variety of applications, including recommendation and classification systems. 
-Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, similarity search can become a useful tool for 
-understanding and verifying the validity of model outputs. 
+Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4, and LLAMA-2, [CITE] similarity search can become a useful tool for 
+understanding and verifying the validity of model embeddings. 
 As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
-even state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.
+even state-of-the-art algorithms exhibit a steep recall versus throughput tradeoff~\cite{ishaq2019clustered}.
 
@@
 
+[TODO: talk about need for exact search in a medical setting; talk to Noah about this, good strains of bacteria, drug response curves]
+
 Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, 
-clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
-Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in big datasets.
+clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks [CITE], and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
 Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
@@
 
-This paper focuses on four novel exact $k$-nearest neighbors search algorithms. We also explain some 
-implementation improvements to the clustering and $\rho$ nearest neighbors search algorithms in CHESS, 
+This paper focuses on four novel algorithms for exact $k$-NN search. We also present some 
+improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, 
 as well as improved genericity across distance functions. Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
 Finally, we provide a comparison of 
-CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} on several datasets 
-from the ANN and Big ANN benchmarks suites.
+CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets 
+from the ANN benchmarks suite [CITE].

### 2023-09-03 sections/2-methods.tex
typo
merge conflict

--- 2023-09-03 sections/2-methods.tex
@@
 
-In this manuscript, we are primarily concerned with $k$-nearest neighbors in a finite-dimensional 
-vector space. Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of 
-an organism, the vector representation of a single image, or any entity on which we can define a \emph{distance function}).
+In this manuscript, we are primarily concerned with $k$-nearest neighbors in a finite-dimensional space.
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of
+an organism, the neural embedding of a single image, or any entity on which we can define a \emph{distance function}).
 
@@
 be symmetric (i.e., $f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
-When, in addition to these constraints, the distance function obeys
-the triangle inequality, it is a \emph{distance metric}, and in such cases we can guarantee that our search algorithms have perfect recall. 
+In addition to these constraints, if the distance function obeys
+the triangle inequality, i.e., it is a \emph{distance metric}, then we guarantee that CAKES has perfect recall. 
 Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
@@
 Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point 
-and length scale in the dataset, 
-which we define as: 
+and length scale in the dataset. 
+We define local fractal dimension (LFD) as: 
 \begin{equation} \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) } \label{1} \end{equation}
@@
 manifold within the 8,575-dimensional embedding space.
-The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold at a given point, for some length scale.
-The notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space is known as the \emph{manifold hypothesis}~\cite{fefferman2016testing}.
+The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
+This notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space is known as the \emph{manifold hypothesis}~\cite{fefferman2016testing}.
 
 Figures 2 and 3 in~\cite{ishaq2019clustered} illustrate the low fractal dimensions of two datasets used with $\rho$-NN search in CHESS. 
-We observe that for both APOGEE (Figure 2 in~\cite{ishaq2019clustered}) and GreenGenes (Figure 3 in~\cite{ishaq2019clustered}), other than the most extreme 10\% of clusters, virtually all clusters have a local fractal dimension of less than $2 \ll 8,575$
-and $2 \ll 2,250$ respectively.
+We observe that for both APOGEE (Figure 2 in~\cite{ishaq2019clustered}) and GreenGenes (Figure 3 in~\cite{ishaq2019clustered}), other than the most extreme 10\% of clusters, virtually all clusters have a local fractal dimension of less than 2.
 This suggests that APOGEE and GreenGenes are good datasets for use with entropy-scaling search algorithms like CHESS and CAKES.
@@
 
-CLAM assumes the manifold hypothesis. 
+CAKES assumes the manifold hypothesis. 
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy 
@@
 properties of clusters and of the manifold, CLAM does not rely on such notions; 
-they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.
+they serve merely as a convenient, intuitive way to talk about the underlying mathematics.
 
+[TODO: change to just partition because I took about quicksort partition]
 \subsubsection {Cluster Partitioning}
@@
 is $C$'s \emph{center}. The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
-The point $l$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}, and the point $r$ which is furthest
+The point $l$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}. The point $r$ which is furthest
 from $l$ is designated the \emph{right pole}. We then partition the cluster into a \emph{left child} and \emph{right child}, where the 
@@
 
-\subsubsection {Dataset Depth-First Reordering}
+\subsubsection {Depth-First Reordering}
 \label{subsubsec:methods:dataset-depth-first-reordering}
 
-CAKES also improves upon CHESS by reordering the dataset. In CHESS, each cluster stored a list of indices into the dataset, 
+CAKES also improves upon CHESS by reordering the dataset in depth-first order of traversal of the tree.
+In CHESS, each cluster stored a list of indices into the dataset, 
 which we used during search to retrieve a cluster's points. 
@@
 points, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
-With $\log{}n$ levels in the tree, this approach had $\mathcal{O}(n\log{}n)$ memory cost.
+Assuming $\log{}n$ levels in the tree, this approach had $\mathcal{O}(n\log{}n)$ memory cost. 
+[TODO: talk to Noah about how to phrase this, how to get at that balanced tree is the worst case]
 
-One may think to improve this memory cost to $\mathcal{O}(n)$ by storing only the indices of points at 
+One may improve this memory cost to $\mathcal{O}(n)$ by storing only these indices at 
 the leaf cluster level.
@@
 so that points are stored in a depth-first order. Then, within each cluster, we need only 
-store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset. Each parent cluster 
-have the same offset as its left child, and each right child has an offset equal to their sibling's offset
-plus their siblings's cardinality. With no additional memory cost nor time cost for retrieving points during search, 
+store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset. A left child has 
+the same offset as its parent cluster, and the corresponding right child has an offset equal to the left child's offset
+plus the left child's cardinality. With no additional memory cost nor time cost for retrieving points during search, 
 dataset depth-first reordering offers significant improvements over the approach used in CHESS.
@@
 \label{subsubsec:methods:guaranteed-decrease-in-cluster-radii}
-While it may be tempting to assume that cluster radii decrease with each application of Cluster Partition (refer to Algorithm \ref{alg:partition}), unfortunately, this assumption is incorrect. 
+While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:partition}), unfortunately, this assumption is incorrect. 
 Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii; in particular, we prove in this section that cluster radii will stop increasing after at most 
-$d$ partitions, where $d$ is the dimensionality of the dataset.
+$d$ partitions, where $d$ is the fractal dimensionality of the dataset.
 
@@
 Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
-In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance along every axis. 
-Such a distribution would also produce a balanced clustering.
+In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance \emph{along every axis}. 
+Such a distribution would also produce a balanced cluster tree.
 
-In this case, Cluster Partition will select a maximally distant pair of points to use as poles, i.e., it will choose one of the $d$ axes along
-which to split the cluster into two children. 
-After one application of Algorithm \ref{alg:partition}, the maximum pairwise distance along that axis will be
+In this case, Partition will select a maximally distant pair of points to use as poles; without loss of generality, 
+say it chooses two points at the extrema one of the $d$ axes. 
+After one partition, the maximum pairwise distance along that axis will be
 bounded above by $R$. 
-The next recursive Cluster Partition will select another of the $d$ axes. 
-Thus, after at most $d$ applications of Algorithm \ref{alg:partition}, the
+The next recursive Partition will select another of the $d$ axes, and once again, the distance along that axis will be bounded above by $R$
+Thus, after at most $d$ partitions, the
 maximum pairwise distance along each axis will be bounded above by $R$. 
@@
 will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes. 
-See Figure 1 for a concrete example.
+See Figure~\ref{fig:methods:scaling_behavior} for an example.
 
@@
 bounded above by $\frac{R}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at 
-most $d$ applications of Algorithm \ref{alg:partition}.
+most $d$ partitions.
 
 Note that, in practice, we never see a balanced clustering. Cluster Partition  produces unbalanced trees due to the varying density of the sampling 
-in different regions of the manifold and the low dimensional "shape" of the manifold. Further, the cluster radii decrease by a factor much larger than 
-$\frac{\sqrt{2}}{2}$ in practice, and the upper bound of $d$ partitions is seldom realized.
+in different regions of the manifold and the low dimensional ``shape'' of the manifold. Further, in practice, the cluster radii decrease by a factor much larger than 
+$\frac{\sqrt{2}}{2}$ with every partition; the upper bound of $d$ partitions is seldom realized.
 
@@
     \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
-    \caption{Scaling behavior of radii on a uniform-density disk, which in two dimensions represents the worst-case scenario. 
+    \caption{Scaling behavior of radii on a uniform-density disk, which represents the worst-case scenario for a two-dimensional distribution. 
     $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
@@
 
-The asymptotic complexity of Cluster Partition is the same as described in~\cite{ishaq2019clustered}
+The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}
 By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of 
@@
 \label{subsec:methods:index-building}
-In this work, we also introduce the method of \emph{sharding} for datasets that cannot fit in memory. 
-This approach also leverages the fact that $\rho$-nearest neighbors search is typically much faster than $k$-nearest neighbors search.
+In this work, we also introduce a method of \emph{sharding} for datasets that cannot fit in memory. 
+This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor 
+is typically much faster than $k$-nearest neighbors search.
 In addition to sharding, we perform some naive auto-tuning to select the optimal $k$-NN algorithm to be used with a given dataset.
 
-With this approach, we begin by taking a random sample of up to 10\% of the data. 
+With this approach, we begin by taking a random sample of no more than 10\% of the data. This sample 
+is considered the first shard. 
 The exact proportion of the dataset in this random sample depends on the dimensionality of the dataset and a user-specified memory limit. 
-FIrst, we build a cluster tree on this sample, and 
-then consider all clusters at some user-specified depth in the tree. Using the cluster centers 
-of these clusters as queries, and a user-specified value of $k$, we perform $k$-NN search using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}. 
-We record the 
-time taken for each algorithm and select the one which had the lowest total search time over all the queries. 
-We then split the rest of the dataset into some user-specified number of shards, and use the algorithm identified in the previous step to perform $k$-NN search on the first shard. 
-This search finds the $k$ closest neighbors to the query \emph{in the first shard}. 
-We take the furthest of those $k$ neighbors and use its distance from the query as the radius to be used with $\rho$-nearest neighbors search (see Section~\ref{subsec:methods:rnn-search}) on each remaining shard.
-We then take the union of the results of these $\rho$-nearest neighbors searches, along with the $k$ neighbors from the first shard, and then linear search over those points to find the actual $k$ nearest neighbors.
+First, we build a cluster tree on this sample, and 
+then consider all clusters at some user-specified depth in the tree, with a default depth of 10.
+Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
+We select the algorithm which had the lowest total search time over all the queries.
+We then split the rest of the dataset into shards based on the user-defined memory limit.
 
-Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we use the selected algorithm to perform search on the given dataset for any value of $k$.
+Sharded search then proceeds as follows. We use previously determined $k$ algorithm on the first shard. 
+We use the distance from the $k$-th farthest neighbor to perform $\rho$-NN search on each subsequent shard. 
+After searching each shard, we keep only the $k$-nearest neighbors so far, potentially decreasing the distance 
+from the $k$-th farthest neighbor for search on the next shard.
+
+Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
 Note also that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search.
-However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th ), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
+However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th neighbor), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
 Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
-
-For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but with only one shard.
+For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but, since the entire dataset fits in memory, there is only one shard.
 
@@
 time complexity, where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, and $k$ is the number of nearest neighbors.
+We note that the fraction $\frac{-d}{\hat{d}}$ should, unless fractal dimension is highly variable (that is, $\hat{d}$ differs significantly from $d$), be relatively close to 1.
 

### 2023-09-03 sections/4-results.tex
fixed precision issues
updates to sharding and auto-tuning
bad recalls
merge conflict

@@
 \label{sec:results}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{images/result_plots/deep-image_1.png}
-    \caption{Performance of CAKES algorithms on deep-image dataset.}
-    \label{fig:methods:deep-image}
-\end{figure}
 
@@
 
-For each pair $(X, f)$ and a small selection of depths, measure:
-\begin{enumerate}[1.]
-    \item linear search vs accelerated search:
-    \begin{enumerate}[i.]
-        \item runtime performance,
-        \item speedup factor, and
-        \item number of distance comparisons.
-    \end{enumerate}
-\end{enumerate}
-
 \begin{table*}[!t]
@@
     \hline
-    fashion-mnist      & 0.54 & 324.79                           & 1.00 & 52.91                                  & 0.94 & 60.89                                                    & 1.00 & 16,220 \\
+    fashion-mnist      & 1.000 & 188.03                           & 1.000 & 52.91                                  & 1.000* & 339.74                                                    & 1.000 & 16,220 \\
     \hline
-    gist                   & 1.00 & 3.55                           & 1.00 & 2.63                                     & 0.95 & 149.58                                                   & 1.00 & 10,600 \\
+    gist                   & 1.000 & 3.55                           & 1.000 & 2.63                                     & 0.953 & 149.58                                                   & 1.000 & 10,600 \\
     \hline
-    glove-25              & 1.00 & 78.78                          & 1.00 & 33.52                              & 1.00 & 552.67                                                  & -- & 19,520 \\
+    glove-25              & 1.000 & 78.78                          & 1.000 & 33.52                              & 1.000 & 552.67                                                  & 0.798 & 19,520 \\
     \hline
-    glove-100             & 1.00 & 21.16                          & 1.00 & 16.52                                & 0.99 & 155.90                                                  & -- & 6,935 \\
+    glove-100             & 1.000 & 21.16                          & 1.000 & 16.52                                & 0.998 & 155.90                                                  & 0.904 & 6,935 \\
     \hline
-    sift                  & 1.00 &  24.52                          & 1.00 & 16.40                               & 1.00 & 275.07                                                    & 1.00 & 31,800 \\
+    sift                  & 1.000 &  24.52                          & 1.000 & 16.40                               & 1.000 & 275.07                                                    & 1.000 & 31,800 \\
     \hline
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$. Other methods did not report results for glove-25.}
+    \caption{Runtime performance (queries per second) of CAKES vs.
+other methods, $k=100$.
+Other methods did not report results for glove-25.
+Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
     \label{table:results:ann-100}
@@
     \hline
-    fashion-mnist         & 1.00 & 111.47                           & 0.99 & 16.02                                  & 0.99 & 277.80                                                    & 1.00 & 1,549 \\
+    fashion-mnist         & 1.000 & 111.47                           & 1.000* & 16.02                                  & 1.000* & 277.80                                                    & 1.000 & 1,549 \\
     \hline
-    gist                   & 1.00 & 2.70                           & 0.99 & 0.80                                 & 0.99 & 59.71                                                    & 1.00 & 753.9 \\
+    gist                   & 1.000 & 2.70                           & 1.000* & 0.80                                 & 0.995 & 59.71                                                    & 1.000 & 753.9 \\
     \hline
-    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & -- & 1,471 \\
+    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & 0.851 & 1,471 \\
     \hline
-    glove-100             & 0.99 &  7.52                          & 0.99 & 3.11                                  & 0.99 & 92.58                                                    & -- & 396.0 \\
+    glove-100             & 1.000* &  19.04                          & 1.000* & 7.52                                 & 0.987 & 92.58                                                    & 0.905 & 396.0 \\
     \hline
-    sift                  & 1.00 &  23.50                          & 1.00 & 6.51                                  & 0.99 & 179.14                                                    & 1.00 & 2,852 \\
+    sift                  & 1.000 &  23.50                          & 1.000 & 6.51                                  & 1.000* & 179.14                                                    & 1.000 & 2,852 \\
     \hline
@@
 
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) of CAKES vs.
+other methods, $k=100$.
+Other methods did not report results for glove-25.
+Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
+    \label{table:results:ann-100}
+    \vskip 0.15in
+    \begin{center}
+    \begin{small}
+    \begin{sc}
+    \begin{tabular}{|l|p{1cm}|p{1cm}|}
+    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
     % \hline
-    % deep-image             & 0.00 & 0.00                           & 0.00 & 0.00                                 & 0.00 & 0.00                                                     & 1.00 & 42 \\
+    % deep-image             & 0.00 & 0.00                           & 0.00 & 0.00      \\
     % \hline
-    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00                                 & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00      \\
     % \hline
-    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00                              & 0.00 & 0.00                                                    & 1.00 & 42 \\
+    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00      \\
     % \hline 
-    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00                               & 0.00 & 0.00
+    % glove-200              & 0.00 & 0.00                           & 0.   % \hline
+    % \hline
+    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00      \\
+    % \hline
+    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00      \\
+    % \hline 
+    % glove-200              & 0.00 & 0.00                           & 0.00 & 0.00      \\
+    % \hline
+    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00      \\00 & 0.00      \\
+    % \hline
+    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00      \\
+
+    \hline
+    \end{tabular}
+    \end{sc}
+    \end{small}
+    \end{center}
+    \vskip -0.1in
+    \end{table*}

### 2023-09-03 sections/2-methods.tex
table and knn complexity

@@
 We group the Sieve-based algorithms together, as their complexity analyses are similar.
+The complexity of Sieve and Sieve with Separate Centers is dominated by the expected 
+cost of Quicksort-inspired partition algorithm we use to calculate thresholds. This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.
 
-The complexity of the Sieve-based algorithms relies on the assumption that when the cardinalities of clusters is sufficiently close to $k$, the distribution of clusters will mimic the distribution of points in the dataset.
+We rely on the assumption that when the cardinalities of clusters is within a low factor of $k$, say $2k$, the distribution of clusters will mimic the distribution of points in the dataset. 
 From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query point.
-In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, as at each end of each of the $d$ axes, we could have a cluster which overlaps the query ball.
+In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, as we could have a cluster overlapping the query ball at each end of each of the $d$ axes.
 
-In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log_2{\frac{n}{k}}$ iterations of Sieve to reduce the cardinality of clusters to at most $k$.
+In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log_2{\frac{n}{2k}}$ iterations of Sieve to partition clusters enough that their cardinalities are at most $2k$.
+However, since
+high cardinality clusters occur only at low depths in the tree, the length of $Q$ for 
+those iterations is relatively short.
+
+After the first $\log_2{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.

### 2023-09-03 sections/1-introduction.tex
misc changes

@@
 million genomes with aligned length of 50,000 letters; 
-the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences. [tTODO: add dimensionality here]
+the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length.
 
@@
 Similarity search enables a variety of applications, including recommendation and classification systems. 
-Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4, and LLAMA-2, [CITE] similarity search can become a useful tool for 
+Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, similarity search can become a useful tool for
 understanding and verifying the validity of model embeddings. 

### 2023-09-03 sections/2-methods.tex
misc changes

@@
 
-[TODO: change to just partition because I took about quicksort partition]
 \subsubsection {Cluster Partitioning}
@@
 \begin{algorithm} % enter the algorithm environment
-\caption{ClusterPartition(\emph{C})} % give the algorithm a caption
+\caption{Partition(\emph{C})} % give the algorithm a caption
 \label{alg:partition} % and a label for \ref{} commands later in the document
@@
     \IF{$|left| > 1$}
-        \STATE ClusterPartition($left$)
+        \STATE Partition($left$)
     \ENDIF
     \IF{$|right| > 1$}
-        \STATE ClusterPartition($right$)
+        \STATE Partition($right$)
     \ENDIF
@@
 
-Note that, in practice, we never see a balanced clustering. Cluster Partition  produces unbalanced trees due to the varying density of the sampling 
+Note that, in practice, we never see a balanced clustering. Partition  produces unbalanced trees due to the varying density of the sampling 
 in different regions of the manifold and the low dimensional ``shape'' of the manifold. Further, in practice, the cluster radii decrease by a factor much larger than 

### 2023-09-03 sections/1-introduction.tex
speedup factors

@@
 Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, 
-clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks [CITE], and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.

### 2023-09-03 sections/4-results.tex
table and knn complexity
more tables
speedup factors

@@
 
-Figure [ADD FIGURE NUMBER] is a violin plot which shows the performance of our algorithms
-on [ADD DATASET] under [ADD DISTANCE FUNCTION] distance, using [ADD NUMBER OF SHARDS] shards.
-As indicated in the title, the dataset has cardinality [ADD cardinality] and dimensionality [ADD DIMENSIONALITY].
-On the horizontal axis, we have each of the $k$-NN algorithms described in Section 2.4, as well as  
-an additional "algorithm" which we call "Oracle $\rho$-NN" [REPLACE WITH ACTUAL NAME].  With "Oracle $\rho$-NN", we perform $\rho$-NN search using the 
-distance to the $k$-th nearest neighbor as the radius. Of course, as it is impossible to
-to know the distance to the $k$-th nearest neighbor \emph{a priori}, this algorithm 
-is useful only for benchmarking purposes.
+For $k= 10$, we report a mean $989\times$ (median $248\times$) speedup over faiss-ivf, a mean $1456\times$ (median $582\times$) speedup over bruteforce-blas, 
+and a mean $63\times$ (median $49\times$) speedup over HNSW.
+For $k=100$, we report a mean $109\times$ (median $71\times$) speedup over faiss-ivf, a mean $382\times$ (median $267\times$) speedup over bruteforce-blas, and a mean $10\times$ (median $9\times$) speedup over HNSW.
 
-The vertical axis for Figure [ADD FIGURE NUMBER] is the time query in seconds on a logarithmic scale. 
-As shown in the key, different colors represent different values of $k$. 
-The plot shows the distribution of query times for each algorithm for each value of $k$. 
-Because of the logarithmic scale of the vertical axis, it is unsurprising that we 
-see longer lower tails than upper tails for many of these algorithms.
-
-https://github.com/URI-ABD/clam
+The source code for CAKES is available under an MIT
+license at https://github.com/URI-ABD/clam.
 
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs.
-other methods, $k=100$.
-Other methods did not report results for glove-25.
-Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
-    \label{table:results:ann-100}
+    \caption{Runtime performance (queries per second) of CAKES vs. linear, $k=10$}
+    \label{table:results:ann-alt-10}
     \vskip 0.15in
@@
     \begin{sc}
-    \begin{tabular}{|l|p{1cm}|p{1cm}|}
+    \begin{tabular}{|l|l|l|l|l|}
     \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
-    % \hline
-    % deep-image             & 0.00 & 0.00                           & 0.00 & 0.00      \\
-    % \hline
-    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00      \\
-    % \hline
-    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00      \\
-    % \hline 
-    % glove-200              & 0.00 & 0.00                           & 0.   % \hline
-    % \hline
-    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00      \\
-    % \hline
-    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00      \\
-    % \hline 
-    % glove-200              & 0.00 & 0.00                           & 0.00 & 0.00      \\
-    % \hline
-    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00      \\00 & 0.00      \\
-    % \hline
-    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00      \\
-
+    &                    Recall & QPS                           & Recall & QPS                           \\
+    \hline
+    deep-image             & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    \hline
+    mnist                   & 1.000 & 0.00                          & 1.000 & 0.00      \\
+    \hline
+    glove-50               & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    \hline 
+    glove-200              & 1.000 & 0.00                           & 0.00 & 0.00     \\
+    \hline
+    lastfm                 & 1.000 & 0.00                           & 0.00 & 0.00           \\
     \hline
@@
     \end{table*}
+
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) of CAKES vs. linear, $k=100$}
+    \label{table:results:ann-alt-100}
+    \vskip 0.15in
+    \begin{center}
+    \begin{small}
+    \begin{sc}
+    \begin{tabular}{|l|l|l|l|l|}
+    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
+    &                    Recall & QPS                           & Recall & QPS                           \\
+    \hline
+    deep-image             & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    \hline
+    mnist                   & 1.000 & 0.00                          & 1.000 & 0.00      \\
+    \hline
+    glove-50               & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    \hline 
+    glove-200              & 1.000 & 0.00                           & 0.00 & 0.00     \\
+    \hline
+    lastfm                 & 1.000 & 0.00                           & 0.00 & 0.00           \\
+    \hline
+    \end{tabular}
+    \end{sc}
+    \end{small}
+    \end{center}
+    \vskip -0.1in
+    \end{table*}

### 2023-09-03 sections/1-introduction.tex
some intro and lots of discussion work

--- 2023-09-03 sections/1-introduction.tex
@@
 need for efficient, \emph{exact} search remains.
-In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$.
+In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$, and small values of $k$ may be necessary when classes are not well-separated.
 
-[TODO: talk about need for exact search in a medical setting; talk to Noah about this, good strains of bacteria, drug response curves]
+For the sake of clarity, it is important to be precise about what we mean by \emph{exact} $k$-NN, and exact search in general.
+In the context of $\rho$-NN search, exact search typically means returning only those data points that are identical to the query.
+There are many algorithms that perform well here, such as FM-Index~\cite{ferragina2000opportunistic}, which are less applicable to $\rho$-NN search.
+In contrast, in the context of $k$-NN search, while all approaches return exactly $k$ elements (ignoring how distance ties are handled), some approaches may miss some of the $k$ nearest elements, and erroneously include some of the elements outside the truly nearest $k$.
+In any case, $k$-NN is typically used for inference; the \emph{prediction} of a class for a query is obviously not guaranteed to be correct.
+When we refer to \emph{exact} $k$-NN, we refer to precisely the $k$ nearest neighbors being returned; no closer points are elided and no further points are included.
+This has implications for the accuracy of classification, but does not guarantee perfect classification accuracy.
+
+There is some evidence that non-metrics such as cosine distance perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance}; this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could be exhibit suboptimal accuracy.
 

### 2023-09-03 sections/2-methods.tex
some intro and lots of discussion work

--- 2023-09-03 sections/2-methods.tex
@@
 points, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
-Assuming $\log{}n$ levels in the tree, this approach had $\mathcal{O}(n\log{}n)$ memory cost. 
-[TODO: talk to Noah about how to phrase this, how to get at that balanced tree is the worst case]
+Assuming a balanced tree and thus $\mathcal{O}\log_2(n)$ depth, this approach had $\mathcal{O}(n\log_2(n)$ memory cost.
+We note that while a balanced tree is classically a best-case scenario (e.g.
+binary search), it represents a worst-case scenario for space complexity in this application; the space complexity of a completely unbalanced (linear) tree would be $\mathcal{O}(n)$.
 
@@
 After the first $\log_2{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.
+Therefore, the asymptotic complexity of the sieve algorithms is $\mathcal{O}(log_2(\frac{n}{d})$.

### 2023-09-03 sections/5-discussion.tex
some intro and lots of discussion work

@@
-\section{Discussion}
+\section{Discussion and Future Work}
 \label{sec:discussion}
 
-Discuss results and future work \dots
+We have presented CAKES, a method for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
+CAKES provides exact $k$-NN when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
+CAKES is most effective when datasets are not uniform and do not obey a ``simple'' distribution such as Gaussian, but rather adhere to ``interesting'' low-dimensional manifolds even when embedded in a high-dimensional space; thus, CAKES is best when the manifold hypothesis is upheld.
 
-CLAM-CAKES.
-This is most effective when datasets exhibit low metric entropy and low local fractal dimension.
+Unlike locality-sensitive hashing (LSH) approaches, which require a novel hashing function for each distance metric in use, CAKES is extensible to any user-provided distance function.
+Also unlike LSH approaches, CAKES guarantees perfect precision and recall as long as the distance function in use is a \emph{metric}; that is, it obeys the properties of identity, symmetry, and the triangle inequality.
+As mentioned in Section~\ref{sec:introduction}, this should not be conflated with having $k$-NN itself provide perfect \emph{inference} accuracy.
 
-Unlike LSH, this is extensible to any user provided distance function.
+In comparison to other methods, While CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
+This bears further investigation, but likely stems from how heavily the knn and clustering algorithms depend on the triangle inequality.
+Preliminary results suggest that when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases.
 
-Exactness of search and metric vs non-metric.
+Some of the questions raised by this study suggest that future work is justified; a comparison across more datasets and larger datasets is clearly in order, as is further analysis with distance functions that existing methods such as FAISS do not support, such as Wasserstein distance~\cite{vallender1974calculation} (particularly for arbitrary dimensionality), as Tanimoto distance~\cite{bajusz2015tanimoto}, which is particularly applicable to computing distances based on maximal common subgraphs of molecular structures.
+Incorporating these or other distance functions in CAKES only requires providing a Rust implementation of the distance function.
 
-While CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
-This bears further investigation, but likely stems from how heavily the knn and clustering algorithms depend on the triangle inequality.
-
-\subsection{Future Works}
-\label{subsec:results:future-works}
-
-For future work, we plan to investigate compression by representing differences at each level of the
+In future work, we plan to investigate compression by representing differences at each level of the
 binary tree, particularly for string or genomic data, where all differences are discrete.
@@
 
-Additionally, we would like to implement additional distance functions to improve the genericity of CAKES. 
-For example, we would like to implement Tanimoto Distance using maximal-common-subgraph for molecular structures. Ultimately, 
-we would also like to implement Wasserstein distance for dimensions higher than two. In order to make this possible, 
-we plan to implement GPU-acceleration for the more computationally-intensive distance functions like Wasserstein.
-
 Finally, we would like to add the ability to view live-updates as new points are added to a dataset. For example, 
@@
 simulate live updates.
+
+% since this online update takes advantage of CAKES' fast search, it allows us to have our cake and eat it, too.

### 2023-09-03 sections/3-datasets.tex

--- 2023-09-02 sections/3-datasets.tex
@@
 We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
-listed below, along with the appropriate distance function.
+along with the appropriate distance function.
 Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
+
+The specifications for the machines used in benchmarking are as follows:
+
+\begin{itemize}
+\item Ark: 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
+\item Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12
+\end{itemize}
 

### 2023-09-03 sections/3-datasets.tex
more tweaks

@@
 \label{sec:datasets-and-distance-functions}
-
-\subsection{ANN-Benchmarks suite}
-\label{subsec:datasets:ann-benchmarks-suite}
 
@@
 
+We compared against FAISS 1.7.4~\cite{johnson2019billion} (IVF and HNSW) and a BLAS-accelerated implementation of na\"ive search as reported in~\cite{johnson2019billion}.
+
 The specifications for the machines used in benchmarking are as follows:
@@
 \begin{itemize}
-\item Ark: 28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
+\item Comparisons to FAISS, Bruteforce28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
 \item Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12

### 2023-09-03 sections/4-results.tex
more tweaks

@@
 For $k=100$, we report a mean $109\times$ (median $71\times$) speedup over faiss-ivf, a mean $382\times$ (median $267\times$) speedup over bruteforce-blas, and a mean $10\times$ (median $9\times$) speedup over HNSW.
+For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
 
 The source code for CAKES is available under an MIT
-license at https://github.com/URI-ABD/clam.
+license at \url{https://github.com/URI-ABD/clam}.
 

### 2023-09-03 sections/5-discussion.tex
more tweaks

@@
 Incorporating these or other distance functions in CAKES only requires providing a Rust implementation of the distance function.
+For all datasets benchmarked, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
+However, in limited testing on synthetic datasets obeying simple distributions, this was not always the case.
+We seek to evaluate these approaches on many more datasets and distance metrics; it could be that repeated $\rho$-NN is always best (and thus the autotuning is unnecessary), but such a determination is premature.
 

### 2023-09-03 sections/5-discussion.tex
merge conflict

@@
 In comparison to other methods, While CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
-This bears further investigation, but likely stems from how heavily the knn and clustering algorithms depend on the triangle inequality.
+This bears further investigation, but likely stems from how heavily the $k$-NN and clustering algorithms depend on the triangle inequality.
 Preliminary results suggest that when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases.
@@
 
+The source code for CAKES is available under an MIT
+license at https://github.com/URI-ABD/clam.
+
 % since this online update takes advantage of CAKES' fast search, it allows us to have our cake and eat it, too.

### 2023-09-03 sections/4-results.tex
merge conflict
updated cakes vs linear tables

@@
 \label{sec:results}
+
+We did not run any of the competitor’s algorithms.
+We lifted the reported numbers from the ann-benhmarks site and their interactive plots.
+We made sure to use the same aws instance as they did so its a fair comparison
+
+Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarking against FAISS, bruteforce-blas, and HNSW.
+Our throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
+We
+
+We report the highest throughput value for each algorithm at a recall value of 1.0.
+We report the recall value for each algorithm at the highest throughput value.
+We report the mean and median speedup factor of CAKES over each algorithm.
 
@@
 For $k=100$, we report a mean $109\times$ (median $71\times$) speedup over faiss-ivf, a mean $382\times$ (median $267\times$) speedup over bruteforce-blas, and a mean $10\times$ (median $9\times$) speedup over HNSW.
+
 For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
 
-The source code for CAKES is available under an MIT
-license at \url{https://github.com/URI-ABD/clam}.
+We notice that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent on the dataset; in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
+As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP} , CAKES's strong performance with this dataset is particularly encouraging.
+Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.
 
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs. linear, $k=10$}
+    \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=10$}
     \label{table:results:ann-alt-10}
@@
     \begin{tabular}{|l|l|l|l|l|}
-    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
-    &                    Recall & QPS                           & Recall & QPS                           \\
+    \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
     \hline
-    deep-image             & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    deep-image            & 232.1 & 37.2      \\
     \hline
-    mnist                   & 1.000 & 0.00                          & 1.000 & 0.00      \\
+    mnist                   & 129,300 & 101.9      \\
     \hline
-    glove-50               & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    glove-50              & 10,950 & 12.28      \\
     \hline 
-    glove-200              & 1.000 & 0.00                           & 0.00 & 0.00     \\
+    glove-200              & 2,290 & 8.55     \\
     \hline
-    lastfm                 & 1.000 & 0.00                           & 0.00 & 0.00           \\
+    lastfm               & 13,950 & 4.55           \\
     \hline
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs. linear, $k=100$}
+    \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=100$}
     \label{table:results:ann-alt-100}
@@
     \begin{sc}
-    \begin{tabular}{|l|l|l|l|l|}
-    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
-    &                    Recall & QPS                           & Recall & QPS                           \\
+    \begin{tabular}{|l|l|l|l|l|l|}
+    \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
     \hline
-    deep-image             & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    deep-image             &157.3 &25.74             \\
     \hline
-    mnist                   & 1.000 & 0.00                          & 1.000 & 0.00      \\
+    mnist                 &71,090 &56.63      \\
     \hline
-    glove-50               & 1.000 & 0.00                           & 0.00 & 0.00      \\
+    glove-50               & 8,867 & 9.84      \\
     \hline 
-    glove-200              & 1.000 & 0.00                           & 0.00 & 0.00     \\
+    glove-200              & 2,264 & 8.49    \\
     \hline
-    lastfm                 & 1.000 & 0.00                           & 0.00 & 0.00           \\
+    lastfm                 & 13,950     &4.55                               \\
     \hline

### 2023-09-04 sections/3-datasets.tex
more tweaks
submitted paper

@@
 
-The specifications for the machines used in benchmarking are as follows:
-
-\begin{itemize}
-\item Comparisons to FAISS, Bruteforce28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
-\item Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12
-\end{itemize}
+All benchmarks were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
+This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}
+The OS kernel was Ubuntu 22.04.3-Ubuntu SMP. 
+The Rust compiler was Rust 1.71.1, and the Python interpreter version was 3.10.12.
 

### 2023-09-04 sections/4-results.tex
submitted paper

@@
 We report the mean and median speedup factor of CAKES over each algorithm.
-
-For $k= 10$, we report a mean $989\times$ (median $248\times$) speedup over faiss-ivf, a mean $1456\times$ (median $582\times$) speedup over bruteforce-blas, 
-and a mean $63\times$ (median $49\times$) speedup over HNSW.
-For $k=100$, we report a mean $109\times$ (median $71\times$) speedup over faiss-ivf, a mean $382\times$ (median $267\times$) speedup over bruteforce-blas, and a mean $10\times$ (median $9\times$) speedup over HNSW.
+For $k= 10$, we report a mean $7654\times$ (median $1200\times$) speedup over faiss-ivf, a mean $11325\times$ (median $2821\times$) speedup over bruteforce-blas, 
+and a mean $485\times$ (median $417\times$) speedup over HNSW. 
+For $k=100$, we report a mean $4479\times$ (median $3456\times$) speedup over faiss-ivf, a mean $16010\times$ (median $13548\times$) speedup over bruteforce-blas, and a mean $403\times$ (median $372\times$) speedup over HNSW.
 
@@
     \hline
-    fashion-mnist      & 1.000 & 188.03                           & 1.000 & 52.91                                  & 1.000* & 339.74                                                    & 1.000 & 16,220 \\
+    fashion-mnist & 1.000  & 188.03    & 1.000 & 52.91  & 1.000* & 339.74 & 1.000 & 141,800 \\
     \hline
-    gist                   & 1.000 & 3.55                           & 1.000 & 2.63                                     & 0.953 & 149.58                                                   & 1.000 & 10,600 \\
+    gist          & 1.000 & 3.55  & 1.000 & 2.63        & 0.953 & 149.58  & 1.000 & 76,420 \\
     \hline
-    glove-25              & 1.000 & 78.78                          & 1.000 & 33.52                              & 1.000 & 552.67                                                  & 0.798 & 19,520 \\
+    glove-25      & 1.000 & 78.78  & 1.000 & 33.52      & 1.000 & 552.67  & 0.798 & 94,580 \\
     \hline
-    glove-100             & 1.000 & 21.16                          & 1.000 & 16.52                                & 0.998 & 155.90                                                  & 0.904 & 6,935 \\
+    glove-100     & 1.000 & 21.16  & 1.000 & 16.52      & 0.998 & 155.90  & 0.904 & 4,566 \\
     \hline
-    sift                  & 1.000 &  24.52                          & 1.000 & 16.40                               & 1.000 & 275.07                                                    & 1.000 & 31,800 \\
+    sift          & 1.000 &  24.52 & 1.000 & 16.40      & 1.000 & 275.07  & 1.000 & 357,400 \\
     \hline
@@
     \hline
-    fashion-mnist         & 1.000 & 111.47                           & 1.000* & 16.02                                  & 1.000* & 277.80                                                    & 1.000 & 1,549 \\
+    fashion-mnist  & 1.000 & 111.47    & 1.000* & 16.02   & 1.000* & 277.80   & 1.000 & 71,370 \\ 
     \hline
-    gist                   & 1.000 & 2.70                           & 1.000* & 0.80                                 & 0.995 & 59.71                                                    & 1.000 & 753.9 \\
+    gist           & 1.000 & 2.70      & 1.000* & 0.80    & 0.995 & 59.71     & 1.000 & 29,090 \\
     \hline
-    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & 0.851 & 1,471 \\
+    glove-25       & -- & --           & --     & --      & -- & --           & 0.851 & 49,350 \\
     \hline
-    glove-100             & 1.000* &  19.04                          & 1.000* & 7.52                                 & 0.987 & 92.58                                                    & 0.905 & 396.0 \\
+    glove-100      & 1.000* &  19.04   & 1.000* & 7.52    & 0.987 & 92.58     & 0.905 & 4,384 \\
     \hline
-    sift                  & 1.000 &  23.50                          & 1.000 & 6.51                                  & 1.000* & 179.14                                                    & 1.000 & 2,852 \\
+    sift           & 1.000 &  23.50    & 1.000  & 6.51    & 1.000* & 179.14   & 1.000 & 147,400 \\                                                 
     \hline

### 2023-09-04 sections/5-discussion.tex
submitted paper

@@
 In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
+Along with compression, we would like to add the ability to perform ``online'' updates as new points are added to a dataset.
 
 We also plan to use CAKES with anomaly detection. Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses 
-a variety of methods in an ensemble to determine an anomaly score for a point. We would like to explore the effectiveness of 
+a variety of methods in an ensemble to determine an anomaly score for a point. We would like to explore the effectiveness of using
 distance to the $k$-th nearest neighbor as an additional method in our ensemble.
-
-Finally, we would like to add the ability to view live-updates as new points are added to a dataset. For example, 
-we could build the cluster tree with a subsample of the full data set and then add the rest of the instances to 
-simulate live updates.
 

### 2023-09-05 sections/1-introduction.tex
fix
fixed unfortunate typos

@@
 \label{sec:introduction}
-Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
+
+Often dubbed ``the Big Data explosion,'' rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
 This is especially true in the field of astronomy. 
@@
 
-Biology has also seen an influx of large, high-dimensional datasets. Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 
-million genomes with aligned length of 50,000 letters; 
-the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length.
+Biology has also seen an influx of large, high-dimensional datasets.
+The GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length;
+Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with aligned length of 50,000 letters.
 
@@
 
-In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating 
-that computer architecture will not ``catch up'' to our computational needs in the near future. As a result, analysis of these large datasets 
-requires better algorithms.
+In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating that computer architecture will not ``catch up'' to our computational needs in the near future.
+As a result, analysis of these large datasets requires better algorithms.
 
@@
 Similarity search enables a variety of applications, including recommendation and classification systems. 
-Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, similarity search can become a useful tool for
-understanding and verifying the validity of model embeddings. 
+Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, similarity search can become a useful tool for understanding and verifying the validity of model embeddings.
 As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
@@
 In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
-Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, 
-prove prohibitively slow for large datasets. While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not 
-exploit the geometric and topological structure inherent in large datasets. Approximate search is sufficient for some applications, but the 
-need for efficient, \emph{exact} search remains.
+Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
+While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not exploit the geometric and topological structure inherent in large datasets.
+Approximate search is sufficient for some applications, but the need for efficient, \emph{exact} search remains.
 In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$, and small values of $k$ may be necessary when classes are not well-separated.
@@
 In the context of $\rho$-NN search, exact search typically means returning only those data points that are identical to the query.
-There are many algorithms that perform well here, such as FM-Index~\cite{ferragina2000opportunistic}, which are less applicable to $\rho$-NN search.
+There are many algorithms that perform well here, such as FM-Index~\cite{ferragina2000opportunistic}, which are less applicable to approximate search.
 In contrast, in the context of $k$-NN search, while all approaches return exactly $k$ elements (ignoring how distance ties are handled), some approaches may miss some of the $k$ nearest elements, and erroneously include some of the elements outside the truly nearest $k$.
-In any case, $k$-NN is typically used for inference; the \emph{prediction} of a class for a query is obviously not guaranteed to be correct.
-When we refer to \emph{exact} $k$-NN, we refer to precisely the $k$ nearest neighbors being returned; no closer points are elided and no further points are included.
+In any case, $k$-NN is commonly used for inference;
+the \emph{prediction} of a class for a query is obviously not guaranteed to be correct.
+When we refer to \emph{exact} $k$-NN, we refer to precisely the $k$ nearest neighbors being returned;
+no closer points are elided and no further points are included.
 This has implications for the accuracy of classification, but does not guarantee perfect classification accuracy.
 
-There is some evidence that non-metrics such as cosine distance perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance}; this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could be exhibit suboptimal accuracy.
+There is some evidence that non-metrics such as cosine distance perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
+this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could exhibit suboptimal accuracy.
 
-Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, 
-clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
-\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
-rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
-neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor 
-Entropy-scaling Search).
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset, rather than its cardinality.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest neighbors search to a hierarchical clustering approach.
+In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor Entropy-scaling Search).
 CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm. 
-Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform
-$k$-nearest neighbors search.
+Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform $k$-nearest neighbors search.
 
-This paper focuses on four novel algorithms for exact $k$-NN search. We also present some 
-improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, 
-as well as improved genericity across distance functions. Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
-Finally, we provide a comparison of 
-CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets 
-from the ANN benchmarks suite [CITE].
+This paper focuses on four novel algorithms for exact $k$-NN search.
+We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
+Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
+Finally, we provide a comparison of CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.

### 2023-09-05 sections/2-methods.tex
fixed unfortunate typos

--- 2023-09-03 sections/2-methods.tex
@@
 
-In this manuscript, we are primarily concerned with $k$-nearest neighbors in a finite-dimensional space.
+In this manuscript, we are primarily concerned with $k$-nearest neighbors search in a finite-dimensional space.
 Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of 
@@
 
-We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \cup \{0\}$ as a function which, 
-given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number. We require that the distance function 
-be symmetric (i.e., $f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
-In addition to these constraints, if the distance function obeys
-the triangle inequality, i.e., it is a \emph{distance metric}, then we guarantee that CAKES has perfect recall. 
-Choice of distance function varies by type of data. For example, with electromagnetic spectra, we use both 
-Euclidean (L2) and Cosine distances. With biological or string data, Hamming and Levenshtein distances are more appropriate.
+We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ as a function which, 
+given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
+We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
+In addition to these constraints, if the distance function obeys the triangle inequality, i.e., it is a \emph{distance metric}, then we guarantee that CAKES has perfect recall. 
+Choice of distance function varies by type of data.
+For example, with electromagnetic spectra, one could both Euclidean (L2) and Cosine distances.
+With biological or string data, Hamming and Levenshtein distances are more appropriate.
 
-Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point 
-and length scale in the dataset. 
+Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point and length scale in the dataset.
 We define local fractal dimension (LFD) as:
-\begin{equation} \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) } \label{1} \end{equation}
-where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ 
-centered at a point $q$ in the dataset $\textbf{X}$; in this example, we compute fractal dimension for some radius $r_1$ and a smaller radius $r_2 = \frac{r_1}{2}$.
-We stress that this concept differs from the \emph{embedding dimension} of a dataset. To illustrate the difference,
-consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575. Hence, the \emph{embedding dimension} of this dataset is 8,575. 
-However, due to physical constraints (namely, the laws that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional 
-manifold within the 8,575-dimensional embedding space.
+
+\begin{equation}
+    \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) }
+    \label{eq:methods:lfd}
+\end{equation}
+
+where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
+We stress that this concept differs from the \emph{embedding dimension} of a dataset.
+To illustrate the difference, consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575.
+Hence, the \emph{embedding dimension} of this dataset is 8,575. 
+However, due to physical constraints (namely, the laws of physics that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional manifold within the 8,575-dimensional embedding space.
 The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
@@
 We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
-The \emph{center} is the geometric median of a sample of points in the \emph{cluster}, and so it is a real data point. The \emph{radius} is the
-maximum distance from a point in the cluster to its \emph{center}. We estimate \emph{local fractal dimension} at the cluster radius and half
-the cluster radius. Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in
-a binary tree has two child nodes. We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(D)$ of a data set $D$ under a hierarchical clustering scheme as a refinement of [7], where
-metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data. Here, we use a hierarchical, divisive clustering 
-approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g. a specified cardinality or radius) are not further split. 
-Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of
-leaf clusters.
+The \emph{center} is the geometric median of a sample of points in the \emph{cluster}, and so it is a real data point.
+The \emph{radius} is the maximum distance from any point in the cluster to the \emph{center}.
+We estimate \emph{local fractal dimension} at the scale of the cluster radius and half the cluster radius.
+Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in a binary tree has two child nodes.
+We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(D)$ of a data set $D$ under a hierarchical clustering scheme as a refinement of~\cite{yu2015entropy}, where metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data.
+Here, we use a hierarchical, divisive clustering approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g.
+a specified cardinality or radius) are not further split.
+Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of leaf clusters.
 
-We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
-similar recursive procedure as outlined in CHESS, but with the following 
-improvements: better selection of poles for partitioning and depth-first dataset reordering 
-(see Section ~\ref{subsubsec:methods:dataset-depth-first-reordering}).
+We start by building a divisive hierarchical clustering of the dataset with CLAM, using a similar recursive procedure as outlined in CHESS, but with the following improvements:
+better selection of poles for partitioning and depth-first dataset reordering (see Section ~\ref{subsubsec:methods:clustering:dataset-depth-first-reordering}).
 
 CAKES assumes the manifold hypothesis. 
-In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy 
-a $d$-dimensional manifold, where $d \ll D$. 
-While we sometimes use Euclidean notions, such as voids and volumes, to talk about geometric and topological 
-properties of clusters and of the manifold, CLAM does not rely on such notions; 
+In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. 
+While we sometimes use Euclidean notions, such as voids and volumes, to talk about geometric and topological properties of clusters and of the manifold, CLAM does not rely on such notions; 
 they serve merely as a convenient, intuitive way to talk about the underlying mathematics.
@@
 \subsubsection {Cluster Partitioning}
+\label{subsubsec:methods:cluster-partitioning}
 
-For a cluster $C$ with $|C|$ points, we begin by computing a 
-random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances 
-between all points in the sample. Based on those pairwise distances, we compute the \emph{geometric median} of this sample; 
-that is, the point which minimizes the sum of distances to all other points in the sample. This geometric median 
-is $C$'s \emph{center}. The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
-The point $l$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}. The point $r$ which is furthest
-from $l$ is designated the \emph{right pole}. We then partition the cluster into a \emph{left child} and \emph{right child}, where the 
-left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
-points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
-those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
-repeat this procedure until each leaf contains only one datum or some other user-specified stopping criterion 
-is met. This process is described in Algorithm \ref{alg:partition}.
+For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in the sample.
+Based on those pairwise distances, we compute the \emph{geometric median} of this sample; 
+i.e., the point which minimizes the sum of distances to all other points in the sample.
+This geometric median is $C$'s \emph{center}.
+The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
+The point $l$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}.
+The point $r$ which is furthest from $l$ is designated the \emph{right pole}.
+We then partition the cluster into a \emph{left child} and \emph{right child}, where the left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all points in the cluster which are closer to $r$ than to $l$.
+Without loss of generality, we assign to the left child those points which are equidistant from $l$ and $r$.
+Starting from a root-cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum or some other user-specified stopping criterion is met.
+This process is described in Algorithm \ref{alg:methods:clustering:partition}.
 
@@
 \caption{Partition(\emph{C})} % give the algorithm a caption
-\label{alg:partition} % and a label for \ref{} commands later in the document
+\label{alg:methods:clustering:partition} % and a label for \ref{} commands later in the document
 \begin{algorithmic} % enter the algorithmic environment
@@
 \subsubsection {Depth-First Reordering}
-\label{subsubsec:methods:dataset-depth-first-reordering}
+\label{subsubsec:methods:clustering:dataset-depth-first-reordering}
 
 CAKES also improves upon CHESS by reordering the dataset in depth-first order of traversal of the tree.
-In CHESS, each cluster stored a list of indices into the dataset, 
-which we used during search to retrieve a cluster's points. 
-Though this approach allowed us to retrieve the points in constant 
-time, its memory cost was prohibitively high; since each cluster stores indices for each of its
-points, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
-Assuming $\log{}n$ levels in the tree, this approach had $\mathcal{O}(n\log{}n)$ memory cost. 
-[TODO: talk to Noah about how to phrase this, how to get at that balanced tree is the worst case]
+In CHESS, each cluster stored a list of indices into the dataset, which we used during search to retrieve a cluster's points. 
+Though this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high;
+since each cluster stores indices for each of its points and, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
+Assuming a balanced tree and thus $\mathcal{O}\log(n)$ depth, this approach had $\mathcal{O}(n\log(n))$ memory cost.
 
-One may improve this memory cost to $\mathcal{O}(n)$ by storing only these indices at 
-the leaf cluster level.
-This approach, however, introduces an $\mathcal{O}(n)$ time cost whenever we need to find the indices for 
+One may improve this memory cost to $\mathcal{O}(n)$ by only storing these indices at the leaf cluster level.
+This approach, however, introduces an $\mathcal{O}(n\log(n))$ time cost whenever we need to find the indices for 
 a non-leaf cluster, since it requires a traversal of the subtree rooted at that cluster.
 
-In this work, we introduce a new approach, wherein after building the cluster tree, we reorder the dataset 
-so that points are stored in a depth-first order. Then, within each cluster, we need only 
-store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset. A left child has 
-the same offset as its parent cluster, and the corresponding right child has an offset equal to the left child's offset
-plus the left child's cardinality. With no additional memory cost nor time cost for retrieving points during search, 
-dataset depth-first reordering offers significant improvements over the approach used in CHESS.
+In this work, we introduce a new approach, wherein after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
+Then, within each cluster, we need only store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset.
+The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
+A left child has the same offset as that of its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
+With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers significant improvements over the approach used in CHESS.
 
 \subsubsection {Scaling Behavior of Cluster Radii}
-\label{subsubsec:methods:guaranteed-decrease-in-cluster-radii}
-While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:partition}), unfortunately, this assumption is incorrect. 
-Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii; in particular, we prove in this section that cluster radii will stop increasing after at most 
-$d$ partitions, where $d$ is the fractal dimensionality of the dataset.
+\label{subsubsec:methods:clustering:guaranteed-decrease-in-cluster-radii}
+
+While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:clustering:partition}), unfortunately, this assumption is incorrect.
+Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii;
+in particular, we prove in this section that cluster radii will stop increasing after at most $d$ partitions, where $d$ is the fractal dimensionality of the dataset.
 
@@
 
-In this case, Partition will select a maximally distant pair of points to use as poles; without loss of generality, 
-say it chooses two points at the extrema one of the $d$ axes. 
-After one partition, the maximum pairwise distance along that axis will be
-bounded above by $R$. 
+In this case, Partition will select a maximally distant pair of points to use as poles;
+without loss of generality, say it chooses two points at the extrema one of the $d$ axes. 
+After one partition, the maximum pairwise distance along that axis will be halved, i.e. bounded above by $R$.
 The next recursive Partition will select another of the $d$ axes, and once again, the distance along that axis will be bounded above by $R$
-Thus, after at most $d$ partitions, the
-maximum pairwise distance along each axis will be bounded above by $R$. 
+Thus, after at most $d$ partitions, the maximum pairwise distance along each axis will be bounded above by $R$. 
 The overall (i.e., not restricted to one axis) maximum pairwise distance 
 will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes. 
-See Figure~\ref{fig:methods:scaling_behavior} for an example.
+See Figure~\ref{fig:methods:scaling-behavior} for an example.
 
@@
 
-Note that, in practice, we never see a balanced clustering. Partition  produces unbalanced trees due to the varying density of the sampling 
-in different regions of the manifold and the low dimensional ``shape'' of the manifold. Further, in practice, the cluster radii decrease by a factor much larger than 
-$\frac{\sqrt{2}}{2}$ with every partition; the upper bound of $d$ partitions is seldom realized.
+Note that, in practice, we never see a balanced clustering.
+Partition produces unbalanced trees due to the varying density of the sampling in different regions of the manifold and the low dimensional ``shape'' of the manifold.
+Further, in practice, the cluster radii decrease by a factor much larger than $\frac{\sqrt{2}}{2}$ with almost every partition;
+the upper bound of $d$ partitions is seldom realized.
 
@@
     \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
-    \caption{Scaling behavior of radii on a uniform-density disk, which represents the worst-case scenario for a two-dimensional distribution. 
+    \caption{
+        Scaling behavior of radii on a uniform-density disk, which represents the worst-case scenario for a two-dimensional distribution. 
         $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
-    After one application of Algorithm~\ref{1}, we have $C_1$, with radius $R_1$ and center $o_1$.
-    The right triangle formed by $o_0$, $o_1$, and $y_+$ in $C_1$ shows that $R_0 \leq R_1$.
-    Hence, the radius of a child cluster can be larger than its parent. However, 
-    after another application of Algorithm~\ref{1}, we have consumed all $d$ orthogonal axes, 
+        After one application of Algorithm~\ref{alg:methods:clustering:partition}, we have $C_1$, with radius $R_1$ and center $o_1$.
+        The right triangle formed by $o_0$, $o_1$, and $y_+$ in $C_1$ shows that $R_0 < R_1$.
+        Hence, the radius of a child cluster can be larger than its parent.
+        However, after another application of Algorithm~\ref{alg:methods:clustering:partition}, we have consumed all $d$ orthogonal axes, 
         as shown in $C_2$.
-    Now, clearly $R_2 \leq R_0$. In fact, for each subsequent application of Algorithm~\ref{1},
-    the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.}
-    \label{fig:methods:scaling_behavior}
+        Now, clearly $R_2 < R_0$.
+        In fact, for each subsequent application of Algorithm~\ref{alg:methods:clustering:partition}, the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.
+    }
+    \label{fig:methods:scaling-behavior}
 \end{figure}
@@
 \subsubsection {Complexity}
-\label{subsubsec:methods:clustering:complexity}
+\label{subsubsec:methods:clustering:clustering:complexity}
 
 The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}
-By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of 
-partitioning and $\mathcal{O}(n \log n)$ cost of building the tree. This is a significant improvement over 
-exact partitioning and tree-building, which have $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ costs respectively.
+By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of partitioning and $\mathcal{O}(n \log n)$ cost of building the tree.
+This is a significant improvement over exact partitioning and tree-building, which have $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ costs respectively.
 
 \subsection {Sharding and Auto-Tuning}
-\label{subsec:methods:index-building}
+\label{subsec:methods:sharding-and-auto-tuning}
+
 In this work, we also introduce a method of \emph{sharding} for datasets that cannot fit in memory.
-This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor 
-is typically much faster than $k$-nearest neighbors search.
+This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor is typically much faster than $k$-nearest neighbors search.
 In addition to sharding, we perform some naive auto-tuning to select the optimal $k$-NN algorithm to be used with a given dataset.
 
-With this approach, we begin by taking a random sample of no more than 10\% of the data. This sample 
-is considered the first shard. 
+With this approach, we begin by taking a random sample of no more than 10\% of the data.
+This sample is considered the first shard.
 The exact proportion of the dataset in this random sample depends on the dimensionality of the dataset and a user-specified memory limit. 
-First, we build a cluster tree on this sample, and 
-then consider all clusters at some user-specified depth in the tree, with a default depth of 10.
-Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
+We then split the rest of the dataset into more shards based on the user-defined memory limit.
+For each shard, we build a cluster tree using Algorithm~\ref{alg:methods:clustering:partition}.
+From the first shard, we consider all clusters at some user-specified depth in the tree, with a default depth of 10.
+Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
 We select the algorithm which had the lowest total search time over all the queries.
-We then split the rest of the dataset into shards based on the user-defined memory limit.
 
-Sharded search then proceeds as follows. We use previously determined $k$ algorithm on the first shard. 
-We use the distance from the $k$-th farthest neighbor to perform $\rho$-NN search on each subsequent shard. 
-After searching each shard, we keep only the $k$-nearest neighbors so far, potentially decreasing the distance 
-from the $k$-th farthest neighbor for search on the next shard.
+Sharded search then proceeds as follows.
+We use previously determined $k$-NN algorithm on the first shard. 
+We use the distance to the $k$-th farthest neighbor to perform $\rho$-NN search on each subsequent shard. 
+After searching each shard, we keep only the $k$-nearest neighbors so far, potentially decreasing the distance to the $k$-th farthest neighbor for search on the next shard.
 
 Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
-Note also that the size of the shards is not necessarily consistent; it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search.
+Note also that the size of the shards is not necessarily consistent;
+it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search. 
 However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th neighbor), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
-Finding the precise optimal ratio between the size of the first shard and the size of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
+Finding the optimal ratio between the size of the first shard and the sizes of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
 For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but, since the entire dataset fits in memory, there is only one shard.
@@
 \subsection{The Search Problem}
-\label{subsec:methods:general-search}
-Having explained the clustering process, we can now pose the $k$- and $\rho$- nearest neighbors search problems.
+\label{subsec:methods:the-search-problem}
+Having explained the clustering process, we can now pose the $k$-NN and $\rho$-NN search problems.
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $f(x, q) \leq \max\{f(y, q): y \in S_q \}$; that is,
-for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
-We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$; that is, 
-find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \ \setminus \ S_q$, $f(x, q) > \max\{f(y, q): y \in S_q \}$;
+that is, for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
+We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
+that is, find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
 
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following cluster 
-properties:
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following properties:
+
 \begin{itemize}
@@
 We define a \emph{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains 
-multiple instances of the same point (i.e., has cardinality greater than 1 but contains only one unique point). A singleton clearly 
-has zero radius, and so $\delta_{max}  = \delta = \delta_{min}$ for these clusters. Hence, we sometimes overload the above 
-notation to refer to distance from a query to a point; in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
+duplicates of the same point (i.e., has cardinality greater than 1 but contains only one unique point).
+A singleton clearly has zero radius, and so $\delta_{max} = \delta = \delta_{min}$ for these clusters.
+Hence, we sometimes overload the above notation to also refer to distance from a query to a point;
+in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
 
@@
 
-We conduct $\rho$-nearest neighbors search as described in \cite{ishaq2019clustered}, but 
-with the following improvement: when a cluster overlaps with the query ball, instead of  
-always proceeding with both of its children, we proceed only with those children which 
-also overlap with the query ball. This improvement is described in Algorithm \ref{alg:rnn}.
+We conduct $\rho$-nearest neighbors search as described in \cite{ishaq2019clustered}, but with the following improvement:
+when a cluster overlaps with the query ball, instead of  always proceeding with both of its children, we proceed only with those children which also overlap with the query ball.
+This improvement is described in Algorithm \ref{alg:methods:rnn-search}.
 
 \begin{algorithm} 
-    \caption{$\rho$-NN(\emph{clusters, query, r})} 
-    \label{alg:rnn} 
+    \caption{$\rho$-NN(\emph{cluster, query, r})} 
+    \label{alg:methods:rnn-search} 
     \begin{algorithmic}
         \REQUIRE $r \geq 0$
-        \REQUIRE $clusters \neq \emptyset$
+        \REQUIRE $cluster \neq \emptyset$
         \STATE $results \Leftarrow \emptyset$
-        \IF{$clusters.left \land clusters.left \cap B_X(query, r) \neq \emptyset$}
-            \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
-                \STATE $\rho$-NN($clusters.left, query, r$)
+        \IF{$cluster.left \cap B_X(query, r) \neq \emptyset$}
+            \IF{$cluster.left.\delta$ $\leq$ $r$ + $cluster.left.radius$}
+                \STATE $\rho$-NN($cluster.left, query, r$)
             \ENDIF
         \ENDIF
-        \IF{$clusters.right \land clusters.right \cap B_X(query, r) \neq \emptyset$}
-            \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
-                \STATE $\rho$-NN($clusters.right, query, r$)
+        \IF{$cluster.right \cap B_X(query, r) \neq \emptyset$}
+            \IF{$cluster.right.\delta$ $\leq$ $r$ + $cluster.right.radius$} 
+                \STATE $\rho$-NN($cluster.right, query, r$)
             \ENDIF
         \ENDIF
-        \IF{$\neg$$clusters.left$ $\land \neg$$clusters.right$}
-            \FOR{$p \in clusters.points$}
+        \IF{$cluster.isLeaf$}
+            \FOR{$p \in cluster.points$}
                 \IF{$p.\delta \leq r$}
@@
 
-The asymptotic complexity of $\rho$-nearest neighbors is the same as in ~\cite{ishaq2019clustered}, namely:
+The asymptotic complexity of $\rho$-nearest neighbors is the same as in~\cite{ishaq2019clustered}, namely:
 
@@
     \mathcal{O}\Bigg(
-    \underbrace{\log_2 \mathcal{N}_{\hat{r_c}}(D)}_{\textrm{metric entropy}} +
-    \overbrace{\left|B_D(q,r)\right|}^{\textrm{output size}}
-    \underbrace{\left(1+\frac{2\hat{r_c}}{r}\right)^d}_{\textrm{scaling factor}}
-     \Bigg)
+    \underbrace{\log_2 \mathcal{N}_{\hat{r_c}}(X)}_{\textrm{metric entropy}} +
+    \overbrace{\left|B_X(q,r)\right|}^{\textrm{output size}}
+    \underbrace{\left(1+\frac{2\hat{r_c}}{r}\right)^d}_{\textrm{scaling factor}}\Bigg)
     \label{hierarchical-complexity}
@@
 
-where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy at that radius, $B_D(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the fractal dimension around the query in a region of radius $r$.
+where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy at that radius, $B_X(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the fractal dimension around the query in a region of radius $r$.
 
@@
 
-In this section, we present four novel algorithms for exact $k$-nearest neighbors search: $k$-NN by Repeated $\rho$-NN, Sieve Search, Sieve Search with Separate Centers, 
-and Greedy Sieve. 
-We use a preliminary auto-tuning function to predict the variant and shard size (see Section~\ref{subsec:methods:index-building}) which will perform 
-best for a given query, dataset, and value of $k$. We then proceed with search using that variant on the first shard.
+In this section, we present four novel algorithms for exact $k$-nearest neighbors search: $k$-NN by Repeated $\rho$-NN, Sieve Search, Sieve Search with Separate Centers, and Greedy Sieve. 
+We use a preliminary auto-tuning function to predict the variant (see Section~\ref{subsec:methods:sharding-and-auto-tuning}) which will perform 
+best for a given dataset and value of $k$.
+We then proceed with search using that variant on the first shard.
 
 In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and
-$Q$ to refer to the data structure which stores the clusters (and, for Sieve and Sieve with Centers, points) which are still in contention for being one of the nearest neighbors.
+$Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the nearest neighbors.
 
@@
 
-In this algorithm, we perform $\rho$-nearest neighbors search small search radius, repeatedly increasing the radius until $k$ neighbors
-have been found.
+In this algorithm, we perform $\rho$-nearest neighbors search starting with a small search radius, and repeatedly increasing the radius until we find $k$ neighbors.
 
 Let $H$ be the set of nearest neighbors found thus far.
-Search starts with a radius $m$ equal to the radius of the root cluster divided by
-the cardinality of the dataset. We perform $\rho$-NN search with radius $m$. 
-If no points are within a distance $m$ of the query, we increase the radius by a factor of
-2 and perform $\rho$-NN search again, repeating until at least one point is found, i.e., 
-until $|H| > 0$.
+Search starts with a radius $m$ equal to the radius of the root cluster divided by the cardinality of the dataset.
+We perform $\rho$-NN search with radius $m$. 
+If no points are within a distance $m$ of the query, we increase the radius by a factor of 2 and perform $\rho$-NN search again, repeating until at least one point is found.
 
-Once $|H| > 0$, we continue to perform $\rho$-NN search, but instead of 
-increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
-by the local fractal dimension in the region of the dataset around the query ball. 
-In particular, 
-we increase the radius by a factor of 
-\begin{equation} \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
-where $\mu$ is the harmonic mean of the local fractal dimensions (abbreviated lfd in algorithm below) of 
-each cluster which contains one of the nearest neighbors found thus far. We use the harmonic mean to 
-ensure that $\mu$ is not dominated by outlier clusters with very high fractal dimension. 
+Once $|H| > 0$, we continue to perform $\rho$-NN search, but instead of increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined by the local fractal dimension in the region of the dataset around the query ball.
+In particular, we increase the radius by a factor of
+
+\begin{equation}
+    \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)
+    \label{eq:repeated-rnn-factor}
+\end{equation}
+
+where $\mu$ is the harmonic mean of the local fractal dimensions (abbreviated lfd in algorithm below) of each cluster which contains one of the nearest neighbors found thus far.
+We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high fractal dimension. 
 Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
-Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data 
-are relatively concentrated in that region; in other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radial increase is needed.
-Thus, the factor of radius increase should be \emph{directly} related to the
-local fractal dimension. Once $|H| >= k$, we return the $k$ closest points.
+Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.
+In other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radial increase is needed.
+Thus, the factor of radius increase should be \emph{directly} related to the local fractal dimension.
+Once $|H| >= k$, we sort the points by their distance to the query and return the $k$ closest points.
 
 \begin{algorithm} % enter the algorithm environment
-    \caption{Repeated$\rho$-NN(\emph{root, query, k})} % give the algorithm a caption
+    \caption{Repeated $\rho$-NN(\emph{root, query, k})} % give the algorithm a caption
     \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
@@
         \STATE $m \Leftarrow$ $\frac{root.radius}{|root|}$
-        \WHILE {$|H| = 0$}
+        \WHILE {$H = \emptyset$}
             \STATE $H.push(\rho$-NN$(root, query, m)$)
@@
             \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
-            \STATE $\mu \Leftarrow |Q| \sum_{C \in Q} \frac{1}{C.lfd} $
+            \STATE $\mu \Leftarrow \frac{|Q|}{\sum_{C \in Q} \frac{1}{C.lfd}} $
             \STATE $m \Leftarrow \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
@@
         \STATE $H.sort()$
-        \STATE Return $H[0], H[1], \cdots H[k-1]$
+        \STATE Return $H[0 \cdots k-1]$
     \end{algorithmic}
@@
 
-Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
-(i.e. arises from the same generating phenomenon) as the rest of the data. Given the uses of $k$-NN search in practice, 
-this assumption is not unreasonable. From this assumption, we can infer that the local fractal dimension
-at the query point does not differ significantly from the mean local fractal dimension of clusters nearby the query.
+Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution (i.e.
+arises from the same generating phenomenon) as the rest of the data.
+Given the uses of $k$-NN search in practice, this assumption is not unreasonable.
+From this assumption, we can infer that the local fractal dimension at the query point does not differ significantly from the mean local fractal dimension of clusters nearby the query.
 
-We find it useful to adopt the terminology used in 
-\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
-Coarse search refers to the process of identifying clusters
-which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
-Fine search refers to the process
-of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
+We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{coarse search} and \emph{fine search} separately.
+Coarse search refers to the process of identifying clusters which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
+Fine search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
 
-In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log_2\mathcal{N}_{\hat{r_c}}(D))$, 
-where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset $D$. 
-To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
-of $\rho$-NN needed to find at least $k$ neighbors.
+In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log\mathcal{N}_{\hat{r_c}}(X))$, 
+where $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy of the dataset $X$. 
+To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations of $\rho$-NN needed to find at least $k$ neighbors.
 
-Based on the assumption that the local fractal dimension at the query point is not significantly different than that of 
-nearby clusters, the factor from Equation \ref{2} suggests that in the expected case, we need only two iterations of 
-$\rho$-NN to find $k$ neighbors: one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
-Since this is a constant factor, it does not affect the asymptotic complexity of coarse search, and thus we have that 
-the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2\mathcal{N}_{\hat{r_c}}(D))$.
+Based on the assumption that the local fractal dimension at the query point is not significantly different than that of nearby clusters, the factor from Equation \ref{eq:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of $\rho$-NN to find $k$ neighbors:
+one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
+Since this is a constant factor, it does not affect the asymptotic complexity of coarse search, and thus we have that the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2\mathcal{N}_{\hat{r_c}}(X))$.
 
-To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
-number of clusters for which we must look at all of the cluster's points. With $\rho$-NN 
-search, $F$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query,
-where $r$ is the search radius, and $\hat{r_c}$ is the mean radius of leaf clusters,
-as in \cite{yu2015entropy}. With $k$-NN by Repeated $\rho$-NN, however, 
-the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
-is query-dependent. Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of 
-$F$ to be the union of clusters which are within a distance of $r' + \hat{r_c}$ from the query, 
-where $r'$ is the distance to the $k$th nearest neighbor and $\hat{r_c}$ is defined as before.
+To determine the asymptotic complexity of fine search, we must estimate $|F|$, the number of clusters for which we must look at all of the cluster's points.
+With $\rho$-NN search, $F$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query, where $r$ is the search radius, and $\hat{r_c}$ is the mean radius of leaf clusters, as in \cite{yu2015entropy}.
+With $k$-NN by Repeated $\rho$-NN, however, the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which is query-dependent.
+Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of $F$ to be the union of clusters which are within a distance of $r' + \hat{r_c}$ from the query, where $r'$ is the distance to the $k$th nearest neighbor and $\hat{r_c}$ is defined as before.
 
-The work in~\cite{yu2015entropy} showed that \begin{equation} |F| \leq \gamma |B(q, r')|\left(1+ \frac{2\hat{r_c}}{r'}\right)^d \label{3} \end{equation} where $\gamma$ is a constant. 
+The work in~\cite{yu2015entropy} showed that
+
+\begin{equation*}
+    |F| \leq \gamma |B(q, r')|\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
+\end{equation*}
+
+where $\gamma$ is a constant. 
 By definition of 
-$r'$, we have that $|B(q, r')| = k$. Thus, we have that \begin{equation} |F| \leq \gamma k\left(1+ \frac{2\hat{r_c}}{r'}\right)^d. \label{4} \end{equation} It remains to provide an estimate for $r'$. 
-To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, 
-and thus the local fractal dimension at the query point is not significantly different than that of nearby clusters.
+$r'$, we have that $|B(q, r')| = k$. Thus, we have that
+
+\begin{equation}
+    |F| \leq \gamma k\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
+    \label{eq:methods:knn-by-rnn-fine-search}
+\end{equation}
+
+It remains to provide an estimate for $r'$. 
+To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, and thus the local fractal dimension at the query point is not significantly different than that of nearby clusters.
 
 We let $\hat{d}$ be the mean of the local fractal dimensions of the clusters near the query point (i.e., the cluster identified during coarse search).
-While ordinarily we compute local fractal dimension 
-by comparing cardinalities of two balls centered at \emph{the same} point at two different radii, in order to estimate $r'$, 
-we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the mean cardinality $\hat{C}$ of clusters near the query 
-at a radius equal to the mean of their radii $\hat{r}$.
-We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we replace the query with 
-the center of one of the clusters near it without significantly changing the local fractal dimension at the query point.
+While ordinarily we compute local fractal dimension by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $r'$, we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the mean cardinality $\hat{C}$ of clusters near the query at a radius equal to the mean of their radii $\hat{r}$.
+We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could replace the query with the center of one of the clusters near it without significantly changing the local fractal dimension at the query point.
 
-By ~\ref{2}, we then have that \begin{equation} \hat{d} = \frac{\log{}\frac{\hat{C}}{k}}{\log{}\frac{\hat{r}}{r'}}. \label{5} \end{equation}
+By Equation ~\ref{eq:repeated-rnn-factor}, we then have that
 
-Since $\hat{D}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
-\begin{equation} r' = \hat{r}\left(\frac{k}{\hat{C}}\right)^{\frac{1}{\hat{d}}}. \label{6} \end{equation}
+\begin{equation*}
+    \hat{d} = \frac{\log{}\frac{\hat{C}}{k}}{\log{}\frac{\hat{r}}{r'}}
+\end{equation*}
 
-Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has 
-$\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(1+ \frac{2\hat{r_c}}{\hat{r}}\left(\frac{k}{\hat{C}}\right)^{\frac{-1}{\hat{d}}}\right)^d\right) = 
-\mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(D)} + k\left(\frac{2\hat{r_c}}{\hat{r}}\right)^d\left(\frac{k}{\hat{C}}\right)^{\frac{-d}{\hat{d}}}\right)$ 
-time complexity, where $\mathcal{N}_{\hat{r_c}}(D)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, and $k$ is the number of nearest neighbors.
-We note that the fraction $\frac{-d}{\hat{d}}$ should, unless fractal dimension is highly variable (that is, $\hat{d}$ differs significantly from $d$), be relatively close to 1.
+Since $\hat{d}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve equation~\ref{eq:methods:knn-by-rnn-fine-search} for $r'$ to get
 
-\subsubsection{Sieve Search}
+\begin{equation*}
+    r' = \hat{r}\left(\frac{k}{\hat{C}}\right)^{\frac{1}{\hat{d}}}
+\end{equation*}
+
+Simplifying the expression for fine search, we have that
+
+\begin{equation*}
+    k\left(1+ \frac{2\hat{r_c}}{\hat{r}}\left(\frac{k}{\hat{C}}\right)^{\frac{-1}{\hat{d}}}\right)^d
+\end{equation*}
+
+Since we are dealing with asymptotic complexity, we can ignore the additive factor of $1$.
+Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has a complexity of
+
+\begin{equation}
+    \mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(X)} + k\left(\frac{2\hat{r_c}}{\hat{r}}\right)^d\left(\frac{k}{\hat{C}}\right)^{\frac{-d}{\hat{d}}}\right)
+    \label{eq:methods:knn-by-rnn-complexity}
+\end{equation}
+
+where $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, and $k$ is the number of nearest neighbors.
+We note that the fraction $\frac{-d}{\hat{d}}$ should be relatively close to 1 unless fractal dimension is highly variable in a small region (that is, $\hat{d}$ differs significantly from $d$).
+
+\subsubsection{Sieve}
 \label{subsubsec:methods:knn-search:sieve}
-In this algorithm, we begin by letting $Q$ be a list containing only the root cluster. 
+
+In this algorithm, we begin by letting $Q$ be a list of clusters and points, starting with only the root cluster. 
 While $Q$ contains at least one cluster (i.e., $Q$ is not a list of only points), we repeat the process described in 
@@
 
-First, 
-we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{max}$ such that 
+First, we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{max}$ such that 
 $q_{\tau}$ and all elements in $Q$ with smaller $\delta_{max}$ collectively contain at least $k$ points.
 
-To find this element, 
-we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of elements in $Q$. 
-Our modified version of this algorithm finds the smallest index $i$ such 
-that all elements in $Q$ with $\delta_{max}$ closer to or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater
-than or equal to $k$. We consider $Q[i]$ to be our $q_{\tau}$, and our threshold 
-$\tau$ to be $Q[i]$'s $\delta_{max}$.
+To find this element, we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of clusters in $Q$ by having each cluster count a number of times equal to its cardinality.
+Our modified version of this algorithm finds the smallest index $i$ such that all elements in $Q$ with $\delta_{max}$ less than or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater than or equal to $k$.
+We consider $Q[i]$ to be our $q_{\tau}$, and our threshold $\tau$ to be $Q[i]$'s $\delta_{max}$.
 
-We then remove from $Q$ any element whose $\delta_{min}$ is greater than $\tau$. Then, for each element left 
-in $Q$, if it is a leaf or contains $k$ or fewer points, we remove it and add all of its points to $Q$. Finally, we
-replace all remaining non-leaf clusters in $Q$ with their children.
+We then remove from $Q$ any element whose $\delta_{min}$ is greater than $\tau$.
+Then we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
+Finally, we replace all non-leaf clusters in $Q$ with their children.
 
-We complete this process until $Q$ contains only points. At this point, we 
-just use the Partition algorithm to find the $k$th 
-nearest neighbor and return all points to the left of it.
+We continue this process until $Q$ contains only points.
+At this point, we just use the Quicksort Partition algorithm once more to find the $k$th nearest neighbor and return all points to the left of it.
 
-This process is described in Algorithm~\ref{alg:sieve}.
+This process is described in Algorithm~\ref{alg:methods:sieve}.
 
 \begin{algorithm} % enter the algorithm environment
-    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
-    \label{alg:sieve} % and a label for \ref{} commands later in the document
+    \caption{Sieve(\emph{root, query, k})} % give the algorithm a caption
+    \label{alg:methods:sieve} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
-        \REQUIRE $tree$, $query$, $k$
-        \STATE $Q \Leftarrow$ [$tree.root$]
+        \REQUIRE $root$, $query$, $k$
+        \REQUIRE $root.cardinality \geq k$
+        \STATE $Q \Leftarrow$ [$root$]
         \WHILE{$|Q| > k$}
-            \STATE $i \Leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
+            \STATE $i \Leftarrow QuicksortPartition(Q, k)$
             \STATE $\tau \Leftarrow Q[i].\delta_{max}$
@@
                 \IF {$q.isLeaf \lor |q| \leq k$}
-                    \STATE $Q.push(q.points)$
+                    \STATE $Q.append(q.points)$
                 \ELSE
                     \STATE $[l, r] \Leftarrow q.children$
-                    \STATE $Q.push([l, r])$   
+                    \STATE $Q.append([l, r])$   
                 \ENDIF
-                \STATE $Q.pop(q)$
+                \STATE $Q.remove(q)$
             \ENDFOR 
@@
 
-\subsubsection{Sieve Search with Separate Centers}
+\subsubsection{Sieve with Separate Centers}
 \label{subsubsec:methods:knn-search:sieve2}
-This algorithm is the same as Sieve Search, but with the following modification: clusters 
-in $Q$ are represented twice-- once as their center and once as the rest of their points. 
-Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster 
-center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, 
-thus allowing us to eliminate some clusters from $Q$ earlier in the process.
+
+This algorithm is the same as Sieve, but with the following modification:
+clusters in $Q$ are represented twice -- once as their center and once as the rest of their points.
+The clusters' multiplicities in $Q$ are adjusted to be one less than their cardinalities.
+Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, thus allowing us to eliminate some clusters from $Q$ earlier in the process.
 
@@
 
-With Greedy Sieve, we first let $Q$ be a min queue of clusters prioritized by $\delta_{min}$. We initialize $Q$ with the root cluster.
-Let $H$ be a fixed-size ($k$) max queue of points prioritized by $\delta$. This queue starts empty.
+We first let $Q$ be a min-queue of clusters prioritized by $\delta_{min}$.
+$Q$ starts containing only the root cluster.
+We also let $H$ be a fixed-size ($k$) max-queue of points prioritized by $\delta$.
+$H$ starts empty.
 
-While $H$ has fewer than $k$ clusters or $\delta$ of the top priority element in $H$ is greater 
-than $\delta_{min}$ of the top priority element in $Q$, we do the following:
+While $H$ has fewer than $k$ points or $\delta$ of the top priority element in $H$ is greater than $\delta_{min}$ of the top priority element in $Q$, we do the following:
+
 \begin{enumerate}
-\item While the top priority element $q$ in $Q$ is not a leaf, pop $q$ from $Q$ and push its children to $Q$.
+\item While the top priority element $q$ in $Q$ is not a leaf, pop it from $Q$ and push its children to $Q$.
 \item For each point $p \in q$, push $p$ to $H$. 
@@
 \end{enumerate}
-After this process, the points left in $H$ are the $k$ nearest neighbors of $q$. This algorithm is described in Algorithm \ref{alg:greedy_search}.
+
+After this process, the points left in $H$ are the $k$ nearest neighbors of $q$.
+This algorithm is described in Algorithm~\ref{alg:greedy-sieve}.
 
 \begin{algorithm} 
-\caption{GreedySieve(\emph{root, query, k})} 
-\label{alg:greedy_search} 
+\caption{Greedy Sieve(\emph{root, query, k})} 
+\label{alg:greedy-sieve} 
 \begin{algorithmic}
@@
 We group the Sieve-based algorithms together, as their complexity analyses are similar.
-The complexity of Sieve and Sieve with Separate Centers is dominated by the expected 
-cost of Quicksort-inspired partition algorithm we use to calculate thresholds. This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.
+The complexity of Sieve and Sieve with Separate Centers is dominated by the expected cost of Quicksort's partition algorithm to calculate thresholds.
+This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.
 
-We rely on the assumption that when the cardinalities of clusters is within a low factor of $k$, say $2k$, the distribution of clusters will mimic the distribution of points in the dataset. 
-From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query point.
-In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, as we could have a cluster overlapping the query ball at each end of each of the $d$ axes.
+We start by considering clusters whose cardinalities are within a small factor of $k$, say a factor of 2.
+The distribution of such clusters will mimic the distribution of points in the dataset. 
+From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query.
+In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, e.g.
+we could have a cluster overlapping the query ball at each end of each of the $d$ axes.
 
-In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log_2{\frac{n}{2k}}$ iterations of Sieve to partition clusters enough that their cardinalities are at most $2k$.
-However, since
-high cardinality clusters occur only at low depths in the tree, the length of $Q$ for 
-those iterations is relatively short.
+In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log{\frac{n}{2k}}$ iterations of Sieve to partition clusters enough that their cardinalities are at most $2k$.
+However, since high cardinality clusters occur only at low depths in the tree, the length of $Q$ for those iterations is relatively low.
 
-After the first $\log_2{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.
+After the first $\log{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.
+Therefore, the asymptotic complexity of the sieve algorithms is $\mathcal{O}\left(log(\frac{n}{d})\right)$.

### 2023-09-05 sections/3-datasets.tex
fixed unfortunate typos

@@
 \section{Datasets And Benchmarking}
-\label{sec:datasets-and-distance-functions}
+\label{sec:datasets-and-benchmarks}
 
-We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, 
-along with the appropriate distance function.
+We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, along with the appropriate distance function.
 Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
 
-We compared against FAISS 1.7.4~\cite{johnson2019billion} (IVF and HNSW) and a BLAS-accelerated implementation of na\"ive search as reported in~\cite{johnson2019billion}.
+We compared against FAISS~\cite{johnson2019billion} (IVF and HNSW) and a BLAS-accelerated implementation of na\"ive search as reported in~\cite{johnson2019billion}.
 
@@
 The Rust compiler was Rust 1.71.1, and the Python interpreter version was 3.10.12.
-
-% \begin{itemize} 
-%     \item DEEP1B (Cosine): $\sim$1 billion 96-dimensional deep descriptors to 
-%     \item Fashion-MNIST (Euclidean): 60,000 784-dimensional vector representations of images of clothing items
-%     \item GIST (Euclidean): 1 million 960-dimensional image descriptors 
-%     \item GloVe (Cosine): $\sim$1.2 million 25-dimensional vector representations of words
-%     \item Kosarak (Jaccard): $\sim$75 thousand ~28 thousand-dimensional binary vectors representing user clickstreams
-%     \item MNIST (Euclidean): 60,000 784-dimensional vector representations of images of handwritten digits
-%     \item NYTimes (Cosine): 290,000 256-dimensional vector representations of words
-%     \item SIFT (Euclidean): 1 million 128-dimensional image descriptors
-%     \item Last.fm (Cosine): $\sim$292, 000 65-dimensional vector representations of songs
-% \end{itemize}
 
@@
                     \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dim.}  \\
-    &                &                  &                         \\
-    \hline
-    nytimes          & cosine              & 290,000            & 256       \\
                     \hline
@@
                     glove-25         & cosine              & 1,183,514          & 25        \\
+                    \hline
+                    glove-50         & cosine              & 1,183,514          & 50        \\
                     \hline
@@
                     \hline
-    glove-50         & cosine              & 1,183,514          & 50        \\
-    \hline
                     lastfm           & cosine              & 292,385            & 65        \\

### 2023-09-05 sections/4-results.tex
fixed unfortunate typos

@@
 
-We did not run any of the competitor’s algorithms.
-We lifted the reported numbers from the ann-benhmarks site and their interactive plots.
-We made sure to use the same aws instance as they did so its a fair comparison
+We did not run any of the competitor's algorithms.
+We lifted the reported numbers from the ann-benchmarks site and their interactive plots.
+We made sure to use the same aws instance as they did so its a fair comparison.
 
-Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarking against FAISS, bruteforce-blas, and HNSW.
-Our throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
-We
+Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarks against FAISS, bruteforce-blas, and HNSW.
+The throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
 
 We report the highest throughput value for each algorithm at a recall value of 1.0.
-We report the recall value for each algorithm at the highest throughput value.
 We report the mean and median speedup factor of CAKES over each algorithm.
-For $k= 10$, we report a mean $7654\times$ (median $1200\times$) speedup over faiss-ivf, a mean $11325\times$ (median $2821\times$) speedup over bruteforce-blas, 
-and a mean $485\times$ (median $417\times$) speedup over HNSW. 
-For $k=100$, we report a mean $4479\times$ (median $3456\times$) speedup over faiss-ivf, a mean $16010\times$ (median $13548\times$) speedup over bruteforce-blas, and a mean $403\times$ (median $372\times$) speedup over HNSW.
+For $k= 10$, we report a mean $7,654 \times$ (median $1,200 \times$) speedup over faiss-ivf, a mean $11,325 \times$ (median $2,821 \times$) speedup over bruteforce-blas, and a mean $485 \times$ (median $417 \times$) speedup over HNSW.
+For $k=100$, we report a mean $4,479 \times$ (median $3,456 \times$) speedup over faiss-ivf, a mean $16,010 \times$ (median $13,548 \times$) speedup over bruteforce-blas, and a mean $403 \times$ (median $372 \times$) speedup over HNSW.
 
@@
 
-We notice that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent on the dataset; in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
-As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP} , CAKES's strong performance with this dataset is particularly encouraging.
+We note that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent;
+in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
+As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP}, CAKES's strong performance with this dataset is particularly encouraging.
 Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.
@@
                 \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                    \hline
                     \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
@@
 Other methods did not report results for glove-25.
-Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
+Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995.}
     \label{table:results:ann-100}
@@
                 \begin{tabular}{|l|l|l|l|l|}
+                    \hline
                     \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
@@
                 \begin{tabular}{|l|l|l|l|l|l|}
+                    \hline
                     \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
                     \hline
-    deep-image             &157.3 &25.74             \\
+                    deep-image        & 157.3                    & 25.74                               \\                            
                     \hline
-    mnist                 &71,090 &56.63      \\
+                    mnist             & 71,090                   & 56.63                               \\
                     \hline
@@
                     \hline
-    lastfm                 & 13,950     &4.55                               \\
+                    lastfm            & 13,950                   & 4.55                                \\
                     \hline

### 2023-09-05 sections/5-discussion.tex
fixed unfortunate typos

@@
 We have presented CAKES, a method for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
-CAKES provides exact $k$-NN when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
-CAKES is most effective when datasets are not uniform and do not obey a ``simple'' distribution such as Gaussian, but rather adhere to ``interesting'' low-dimensional manifolds even when embedded in a high-dimensional space; thus, CAKES is best when the manifold hypothesis is upheld.
+CAKES provides exact recall with $k$-NN when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
+CAKES is most effective when datasets are not uniform and do not obey a ``simple'' distribution such as Gaussian, but rather adhere to ``interesting'' low-dimensional manifolds even when embedded in a high-dimensional space;
+thus, CAKES works best when the manifold hypothesis is upheld.
 
-Unlike locality-sensitive hashing (LSH) approaches, which require a novel hashing function for each distance metric in use, CAKES is extensible to any user-provided distance function.
-Also unlike LSH approaches, CAKES guarantees perfect precision and recall as long as the distance function in use is a \emph{metric}; that is, it obeys the properties of identity, symmetry, and the triangle inequality.
-As mentioned in Section~\ref{sec:introduction}, this should not be conflated with having $k$-NN itself provide perfect \emph{inference} accuracy.
+In comparison to other methods, while CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
+This bears further investigation, but likely stems from how heavily the $k$-NN and clustering algorithms depend on the triangle inequality when discarding clusters from consideration.
+However, the ann-benchmarks results show that even when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases.
 
-In comparison to other methods, While CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
-This bears further investigation, but likely stems from how heavily the $k$-NN and clustering algorithms depend on the triangle inequality.
-Preliminary results suggest that when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases.
+The questions raised by this study suggest several avenues for future work.
+A comparison across more datasets and larger datasets is clearly in order, as is further analysis with distance functions that existing methods such as FAISS do not support, such as Wasserstein distance~\cite{vallender1974calculation} (particularly for arbitrary dimensionality), and Tanimoto distance~\cite{bajusz2015tanimoto}, which is particularly applicable to computing distances based on maximal common subgraphs of molecular structures.
+% TODO: Add Lev, NW, SW
+Incorporating these or other distance functions in CAKES only requires a Rust implementation of the distance function.
 
-Some of the questions raised by this study suggest that future work is justified; a comparison across more datasets and larger datasets is clearly in order, as is further analysis with distance functions that existing methods such as FAISS do not support, such as Wasserstein distance~\cite{vallender1974calculation} (particularly for arbitrary dimensionality), as Tanimoto distance~\cite{bajusz2015tanimoto}, which is particularly applicable to computing distances based on maximal common subgraphs of molecular structures.
-Incorporating these or other distance functions in CAKES only requires providing a Rust implementation of the distance function.
 For all datasets benchmarked, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
-However, in limited testing on synthetic datasets obeying simple distributions, this was not always the case.
-We seek to evaluate these approaches on many more datasets and distance metrics; it could be that repeated $\rho$-NN is always best (and thus the autotuning is unnecessary), but such a determination is premature.
+However, during limited testing on synthetic datasets obeying simple distributions, this was not always the case.
+We seek to evaluate these approaches on many more datasets and distance metrics;
+it could be that repeated $\rho$-NN is always best (and thus the auto-tuning is unnecessary), but such a determination would be premature.
 
-In future work, we plan to investigate compression by representing differences at each level of the
-binary tree, particularly for string or genomic data, where all differences are discrete.
-When we descend a level in the cluster tree, it takes fewer bits to encode each point in terms of the center. 
+In future work, we plan to investigate compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
+When we descend some levels in the cluster tree, it takes fewer bits to encode each point in terms of the center because the radii decrease.
 However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
-This presents the following tradeoff: \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
+This presents the following tradeoff:
+\emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
 In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
-Along with compression, we would like to add the ability to perform ``online'' updates as new points are added to a dataset.
 
-We also plan to use CAKES with anomaly detection. Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses 
-a variety of methods in an ensemble to determine an anomaly score for a point. We would like to explore the effectiveness of using
-distance to the $k$-th nearest neighbor as an additional method in our ensemble.
+We would like to add the ability to perform ``online'' updates to the tree as points are added to or deleted from a dataset.
+This would enable CAKES to be used in a streaming environment.
 
-The source code for CAKES is available under an MIT
-license at https://github.com/URI-ABD/clam.
+We also plan to use CAKES with anomaly detection.
+Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses a variety of methods in an ensemble to determine an anomaly score for a point.
+We would like to explore the effectiveness of using the distribution of distances among the $k$-th nearest neighbors as an additional method in our ensemble.
+
+CLAM and CAKES are implemented in Rust and their source code is available under an MIT license at https://github.com/URI-ABD/clam.
 

### 2023-09-07 sections/2-methods.tex
one more typo

@@
 Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in a binary tree has two child nodes.
-We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(D)$ of a data set $D$ under a hierarchical clustering scheme as a refinement of~\cite{yu2015entropy}, where metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data.
+We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(X)$ of a data set $X$ under a hierarchical clustering scheme as a refinement of~\cite{yu2015entropy}, where metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data.
 Here, we use a hierarchical, divisive clustering approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g.

### 2023-10-09 sections/4-results.tex
added discussion of scaling behavior and commented out things that are no longer accurate

@@
 
-We did not run any of the competitor's algorithms.
-We lifted the reported numbers from the ann-benchmarks site and their interactive plots.
-We made sure to use the same aws instance as they did so its a fair comparison.
+\subsection{Scaling Behavior}
 
-Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarks against FAISS, bruteforce-blas, and HNSW.
-The throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{images/result_plots/fashion-mnist_0.1_scaling.png}
+    \caption{
+        Fashion-mnist scaling
+    }
+    \label{fig:results:fashion-mnist-scaling}
+\end{figure}
 
-We report the highest throughput value for each algorithm at a recall value of 1.0.
-We report the mean and median speedup factor of CAKES over each algorithm.
-For $k= 10$, we report a mean $7,654 \times$ (median $1,200 \times$) speedup over faiss-ivf, a mean $11,325 \times$ (median $2,821 \times$) speedup over bruteforce-blas, and a mean $485 \times$ (median $417 \times$) speedup over HNSW.
-For $k=100$, we report a mean $4,479 \times$ (median $3,456 \times$) speedup over faiss-ivf, a mean $16,010 \times$ (median $13,548 \times$) speedup over bruteforce-blas, and a mean $403 \times$ (median $372 \times$) speedup over HNSW.
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{images/result_plots/glove-100_0.1_scaling.png}
+    \caption{
+        Glove-100 scaling
+    }
+    \label{fig:results:glove-100-scaling}
+\end{figure}
 
-For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{images/result_plots/gist_0.1_scaling.png}
+    \caption{
+        Gist scaling
+    }
+    \label{fig:results:gist-scaling}
+\end{figure}
 
-We note that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent;
-in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
-As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP}, CAKES's strong performance with this dataset is particularly encouraging.
-Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{images/result_plots/random_0.1_scaling.png}
+    \caption{
+        Random scaling
+    }
+    \label{fig:results:random-scaling}
+\end{figure}
 
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=10$}
-    \label{table:results:ann-10}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-                    \hline
-                    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-                    \cline{2-9}
-                    &               Recall & QPS      & Recall & QPS    & Recall & QPS    & Recall & QPS \\
-                    \hline
-                    fashion-mnist & 1.000  & 188.03    & 1.000 & 52.91  & 1.000* & 339.74 & 1.000 & 141,800 \\
-                    \hline
-                    gist          & 1.000 & 3.55  & 1.000 & 2.63        & 0.953 & 149.58  & 1.000 & 76,420 \\
-                    \hline
-                    glove-25      & 1.000 & 78.78  & 1.000 & 33.52      & 1.000 & 552.67  & 0.798 & 94,580 \\
-                    \hline
-                    glove-100     & 1.000 & 21.16  & 1.000 & 16.52      & 0.998 & 155.90  & 0.904 & 4,566 \\
-                    \hline
-                    sift          & 1.000 &  24.52 & 1.000 & 16.40      & 1.000 & 275.07  & 1.000 & 357,400 \\
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
+We stress that CAKES is intended for large datasets, and thus we do not expect to see sublinear performance on small datasets. 
+To explore the scaling behavior of CAKES, we augmented datasets in the ANN-benchmarks suite by adding random points within a 
+small error distance of existing points. This allowed us to isolate the effect of \emph{cardinality} on the performance of CAKES 
+from that of other factors such as dimensionality, metric, or topological structure of the dataset.
 
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES vs.
+Figures \ref{fig:results:fashion-mnist-scaling}, \ref{fig:results:glove-100-scaling}, and \ref{fig:results:gist-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-100 under cosine distance, and gist under Euclidean distance.
+Figure \ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset which does not obey any manifold structure. 
+The the horizontal axes in each figure denote the multiplier by which the cardinality of the original dataset has been increased with synthetic points.
+The vertical axes denote throughput in queries per second, on a logarithmic scale.
+
+In Figure \ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as cardinality grows, all of our algorithms exhibit significantly better throughput than linear search.
+Of our algorithms, GreedySieve consistently performs best for both $k=10$ and $k=100$.
+For $k=10$, GreedySieve exhibits sublinear time performance for all cardinality multipliers, and for $k = 100$, GreedySieve begins to exhibit sublinear time performance at a relatively low multiplier (less than 1.5).
+
+For glove-100, as shown in Figure \ref{fig:results:glove-100-scaling}, we observe that all of our algorithms exhibit superior performance compared to linear search for large cardinalities.
+However, for $k=10$, linear search outperforms all of our algorithms until about a cardinality multiplier of 4, when Sieve and Sieve with Separate Centers begin to show superior performance.
+For $k=100$, this transition happens even later, at a mutliplier of about 6, when GreedySieve begins to have better throughput.
+
+With the notably adversial dataset gist, we observe that for $k=10$, three of our algorithms--GreedySieve, Sieve, and Sieve with Separate Centers--begin to outperform linear search at cardinality multipliers of aboyt 4, 6, and 6 respectively.
+For low values of $k$, Sieve and Sieve with Separate Centers appear to the be the best-performing algorithms.
+For $k=100$, we observe that GreedySieve significantly outperforms all other algorithms after a cardinality multiplier of 4.
+At multipliers lower than this threshold value, linear search exhibits the best throughput.
+With gist, unlike with the other datasets we benchmarked on, we see that RepeatedRnn (for both values of $k$), Sieve (for $k=100$), and Sieve with Separate Centers (for $k=100$) all show performance inferior to linear search even at high cardinalies.
+
+Figure \ref{fig:results:random-scaling} displays results for a random datset of base cardinality 1000000. 
+These results illustrate the fact that absent a manifold structure in the dataset, it takes a much higher cardinality before CAKES begins to outperform linear search. 
+In particular, for $k=10$, we see that Sieve is the best-performing method as cardinality grows, but that it only begins to outperform linear search after a multiplier of about 15. 
+For $k=100$, we have that GreedySieve is the best-performing algorithm, but that it does not begin to have superior throughput until a multiplier of about 20. 
+RepeatedRnn shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
+This is unsurprising, given that RepeatedRnn relies on a low local fractal dimension around the query in order to quickly zero in on the correct radius for $k$ hits. 
+With a completely random dataset, which likely has much higher local fractal dimension, RepeatedRnn will meander for longer before finding this correct radius.
+
+Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
+This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities and consistently outperforms our other algorithms for $k=100$.
+We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
+While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
+Finally, we reiterate that the variation in performance across our algorithms based on dataset, value of $k$, and cardinality support our use of an autotuning function to select the best choice of algorithm.
+Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm differs significantly between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
+
+\subsection{Results against Competitors}
+
+% We did not run any of the competitor's algorithms.
+% We lifted the reported numbers from the ann-benchmarks site and their interactive plots.
+% We made sure to use the same AWS instance as they did so it is a fair comparison.
+
+% Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarks against FAISS, bruteforce-blas, and HNSW.
+% The throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
+
+% We report the highest throughput value for each algorithm at a recall value of 1.0.
+% We report the mean and median speedup factor of CAKES over each algorithm.
+% For $k= 10$, we report a mean $7,654 \times$ (median $1,200 \times$) speedup over faiss-ivf, a mean $11,325 \times$ (median $2,821 \times$) speedup over bruteforce-blas, and a mean $485 \times$ (median $417 \times$) speedup over HNSW.
+% For $k=100$, we report a mean $4,479 \times$ (median $3,456 \times$) speedup over faiss-ivf, a mean $16,010 \times$ (median $13,548 \times$) speedup over bruteforce-blas, and a mean $403 \times$ (median $372 \times$) speedup over HNSW.
+
+% For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
+
+% We note that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent;
+% in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
+% As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP}, CAKES's strong performance with this dataset is particularly encouraging.
+% Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.
+
+% \begin{table*}[!t]
+%     % \renewcommand{\arraystretch}{1.15}
+%     \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=10$}
+%     \label{table:results:ann-10}
+%     \vskip 0.15in
+%     \begin{center}
+%         \begin{small}
+%             \begin{sc}
+%                 \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+%                     \hline
+%                     \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+%                     \cline{2-9}
+%                     &               Recall & QPS      & Recall & QPS    & Recall & QPS    & Recall & QPS \\
+%                     \hline
+%                     fashion-mnist & 1.000  & 188.03    & 1.000 & 52.91  & 1.000* & 339.74 & 1.000 & 141,800 \\
+%                     \hline
+%                     gist          & 1.000 & 3.55  & 1.000 & 2.63        & 0.953 & 149.58  & 1.000 & 76,420 \\
+%                     \hline
+%                     glove-25      & 1.000 & 78.78  & 1.000 & 33.52      & 1.000 & 552.67  & 0.798 & 94,580 \\
+%                     \hline
+%                     glove-100     & 1.000 & 21.16  & 1.000 & 16.52      & 0.998 & 155.90  & 0.904 & 4,566 \\
+%                     \hline
+%                     sift          & 1.000 &  24.52 & 1.000 & 16.40      & 1.000 & 275.07  & 1.000 & 357,400 \\
+%                     \hline
+%                 \end{tabular}
+%             \end{sc}
+%         \end{small}
+%     \end{center}
+%     \vskip -0.1in
+% \end{table*}
+
+% \begin{table*}[!t]
+%     % \renewcommand{\arraystretch}{1.15}
+%     \caption{Runtime performance (queries per second) of CAKES vs.
 other methods, $k=100$.
@@
 Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995.}
-    \label{table:results:ann-100}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-                    \hline
-                    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-                    &                    Recall & QPS  & Recall & QPS     & Recall & QPS      & Recall & QPS \\
-                    \hline
-                    fashion-mnist  & 1.000 & 111.47    & 1.000* & 16.02   & 1.000* & 277.80   & 1.000 & 71,370 \\ 
-                    \hline
-                    gist           & 1.000 & 2.70      & 1.000* & 0.80    & 0.995 & 59.71     & 1.000 & 29,090 \\
-                    \hline
-                    glove-25       & -- & --           & --     & --      & -- & --           & 0.851 & 49,350 \\
-                    \hline
-                    glove-100      & 1.000* &  19.04   & 1.000* & 7.52    & 0.987 & 92.58     & 0.905 & 4,384 \\
-                    \hline
-                    sift           & 1.000 &  23.50    & 1.000  & 6.51    & 1.000* & 179.14   & 1.000 & 147,400 \\                                                 
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
+%     \label{table:results:ann-100}
+%     \vskip 0.15in
+%     \begin{center}
+%         \begin{small}
+%             \begin{sc}
+%                 \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+%                     \hline
+%                     \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+%                     &                    Recall & QPS  & Recall & QPS     & Recall & QPS      & Recall & QPS \\
+%                     \hline
+%                     fashion-mnist  & 1.000 & 111.47    & 1.000* & 16.02   & 1.000* & 277.80   & 1.000 & 71,370 \\ 
+%                     \hline
+%                     gist           & 1.000 & 2.70      & 1.000* & 0.80    & 0.995 & 59.71     & 1.000 & 29,090 \\
+%                     \hline
+%                     glove-25       & -- & --           & --     & --      & -- & --           & 0.851 & 49,350 \\
+%                     \hline
+%                     glove-100      & 1.000* &  19.04   & 1.000* & 7.52    & 0.987 & 92.58     & 0.905 & 4,384 \\
+%                     \hline
+%                     sift           & 1.000 &  23.50    & 1.000  & 6.51    & 1.000* & 179.14   & 1.000 & 147,400 \\                                                 
+%                     \hline
+%                 \end{tabular}
+%             \end{sc}
+%         \end{small}
+%     \end{center}
+%     \vskip -0.1in
+% \end{table*}
 
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=10$}
-    \label{table:results:ann-alt-10}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|l|l|l|l|}
-                    \hline
-                    \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
-                    \hline
-                    deep-image       & 232.1                     & 37.2      \\
-                    \hline
-                    mnist            & 129,300                   & 101.9      \\
-                    \hline
-                    glove-50         & 10,950                    & 12.28      \\
-                    \hline 
-                    glove-200        & 2,290                     & 8.55     \\
-                    \hline
-                    lastfm           & 13,950                    & 4.55           \\
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
+% \begin{table*}[!t]
+%     % \renewcommand{\arraystretch}{1.15}
+%     \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=10$}
+%     \label{table:results:ann-alt-10}
+%     \vskip 0.15in
+%     \begin{center}
+%         \begin{small}
+%             \begin{sc}
+%                 \begin{tabular}{|l|l|l|l|l|}
+%                     \hline
+%                     \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
+%                     \hline
+%                     deep-image       & 232.1                     & 37.2      \\
+%                     \hline
+%                     mnist            & 129,300                   & 101.9      \\
+%                     \hline
+%                     glove-50         & 10,950                    & 12.28      \\
+%                     \hline 
+%                     glove-200        & 2,290                     & 8.55     \\
+%                     \hline
+%                     lastfm           & 13,950                    & 4.55           \\
+%                     \hline
+%                 \end{tabular}
+%             \end{sc}
+%         \end{small}
+%     \end{center}
+%     \vskip -0.1in
+% \end{table*}
 
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=100$}
-    \label{table:results:ann-alt-100}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|l|l|l|l|l|}
-                    \hline
-                    \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
-                    \hline
-                    deep-image        & 157.3                    & 25.74                               \\                            
-                    \hline
-                    mnist             & 71,090                   & 56.63                               \\
-                    \hline
-                    glove-50          & 8,867                    & 9.84                                \\
-                    \hline 
-                    glove-200         & 2,264                    & 8.49                                \\
-                    \hline
-                    lastfm            & 13,950                   & 4.55                                \\
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
+% \begin{table*}[!t]
+%     % \renewcommand{\arraystretch}{1.15}
+%     \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=100$}
+%     \label{table:results:ann-alt-100}
+%     \vskip 0.15in
+%     \begin{center}
+%         \begin{small}
+%             \begin{sc}
+%                 \begin{tabular}{|l|l|l|l|l|l|}
+%                     \hline
+%                     \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
+%                     \hline
+%                     deep-image        & 157.3                    & 25.74                               \\                            
+%                     \hline
+%                     mnist             & 71,090                   & 56.63                               \\
+%                     \hline
+%                     glove-50          & 8,867                    & 9.84                                \\
+%                     \hline 
+%                     glove-200         & 2,264                    & 8.49                                \\
+%                     \hline
+%                     lastfm            & 13,950                   & 4.55                                \\
+%                     \hline
+%                 \end{tabular}
+%             \end{sc}
+%         \end{small}
+%     \end{center}
+%     \vskip -0.1in
+% \end{table*}

### 2023-10-09 sections/5-discussion.tex
added discussion of scaling behavior and commented out things that are no longer accurate

@@
 This bears further investigation, but likely stems from how heavily the $k$-NN and clustering algorithms depend on the triangle inequality when discarding clusters from consideration.
-However, the ann-benchmarks results show that even when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases.
+
+% Is this still true: "However, the ann-benchmarks results show that even when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases."
 
@@
 
-For all datasets benchmarked, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
-However, during limited testing on synthetic datasets obeying simple distributions, this was not always the case.
-We seek to evaluate these approaches on many more datasets and distance metrics;
-it could be that repeated $\rho$-NN is always best (and thus the auto-tuning is unnecessary), but such a determination would be premature.
+% For all datasets benchmarked, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
+% However, during limited testing on synthetic datasets obeying simple distributions, this was not always the case.
+% We seek to evaluate these approaches on many more datasets and distance metrics;
+% it could be that repeated $\rho$-NN is always best (and thus the auto-tuning is unnecessary), but such a determination would be premature.
 

### 2023-10-15 sections/2-methods.tex
reworking data augmentation explanation

@@
 After the first $\log{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.
-Therefore, the asymptotic complexity of the sieve algorithms is $\mathcal{O}\left(log(\frac{n}{d})\right)$.
+Therefore, the asymptotic complexity of the sieve algorithms is $\mathcal{O}\left(\log(\frac{n}{d})\right)$.
+
+\subsection{Synthetic Data}
+\label{subsec:methods:synthetic-data}
+
+We expect CAKES to perform well as the cardinality of datasets increases, and to explore this scaling behavior, we augmented datasets in the ANN-benchmarks suite by adding random points within a 
+small noise threshold of existing points. This allowed us to isolate the effect of \emph{cardinality} on the performance of CAKES 
+from that of other factors such as dimensionality, metric, or topological structure of the dataset.
+
+To elaborate, we start with an original dataset $X$ from the ANN-benchmark suite, a user-specified noise tolerance $\epsilon$, and a mutliplier $m$.
+For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon$ of $x$. To construct a synthetic data point $x'$, we choose $d$ random numbers 
+$r_1, r_2, \cdots r_d$ in the interval $[-\tfrac{\epsilon}{\sqrt{d}}, \tfrac{\epsilon}{\sqrt{d}}]$ and then add them componentwise to the original datapoint. 
+This ensures that any synthetic point is within a distance of $\epsilon$ from a real point in the original dataset.

### 2023-10-15 sections/3-datasets.tex
reworking data augmentation explanation

@@
 Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
+
+In addition to benchmarks on datasets from the ANN Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{subsec:methods:synthetic-data}.
+In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the multiplier $m$ of new points increases.
+While we explored the scaling behavior of CAKES with augmented versions of each dataset in \ref{table:datasets:summary}, we give an in-depth discussion of results for scaling behavior on fashion-mnist, glove-100, gist, and a purely random dataset.
 

### 2023-10-15 sections/4-results.tex
reworking data augmentation explanation

@@
 \end{figure}
-
-We stress that CAKES is intended for large datasets, and thus we do not expect to see sublinear performance on small datasets. 
-To explore the scaling behavior of CAKES, we augmented datasets in the ANN-benchmarks suite by adding random points within a 
-small error distance of existing points. This allowed us to isolate the effect of \emph{cardinality} on the performance of CAKES 
-from that of other factors such as dimensionality, metric, or topological structure of the dataset.
 

### 2023-10-15 sections/2-methods.tex
some more minor changes to synthetic data sections and results discussion

@@
 For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon$ of $x$. To construct a synthetic data point $x'$, we choose $d$ random numbers 
-$r_1, r_2, \cdots r_d$ in the interval $[-\tfrac{\epsilon}{\sqrt{d}}, \tfrac{\epsilon}{\sqrt{d}}]$ and then add them componentwise to the original datapoint. 
-This ensures that any synthetic point is within a distance of $\epsilon$ from a real point in the original dataset.
+$r_1, r_2, \cdots , r_d$ in the interval $[-\tfrac{\epsilon}{\sqrt{d}}, \tfrac{\epsilon}{\sqrt{d}}]$ and add them componentwise to the original datapoint. 
+Since $\sqrt{d \cdot (\tfrac{\epsilon}{\sqrt{d}})^2 } = \sqrt{d \cdot \tfrac{\epsilon^2}{d}}  = \sqrt{\epsilon^2} = \epsilon$, this ensures that any synthetic point is within a distance of $\epsilon$ from a real point in the original dataset.

### 2023-10-15 sections/3-datasets.tex
some more minor changes to synthetic data sections and results discussion

@@
 In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the multiplier $m$ of new points increases.
-While we explored the scaling behavior of CAKES with augmented versions of each dataset in \ref{table:datasets:summary}, we give an in-depth discussion of results for scaling behavior on fashion-mnist, glove-100, gist, and a purely random dataset.
+While we explored the scaling behavior of CAKES with augmented versions of each dataset in \ref{table:datasets:summary}, we give an in-depth discussion in \ref{sec:results} of results for scaling behavior on fashion-mnist, glove-100, gist, and a purely random dataset.
 

### 2023-10-15 sections/4-results.tex
some more minor changes to synthetic data sections and results discussion

@@
 \subsection{Scaling Behavior}
+
+Figures \ref{fig:results:fashion-mnist-scaling}, \ref{fig:results:glove-100-scaling}, and \ref{fig:results:gist-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-100 under cosine distance, and gist under Euclidean distance.
+Figure \ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset which does not obey any manifold structure. 
+The the horizontal axes in each figure denote the multiplier by which the cardinality of the original dataset has been increased with synthetic points.
+The vertical axes denote throughput in queries per second, on a logarithmic scale.
+
+In Figure \ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as cardinality grows, all of our algorithms exhibit significantly better throughput than linear search.
+Of our algorithms, GreedySieve consistently performs best for both $k=10$ and $k=100$.
+For $k=10$, GreedySieve exhibits sublinear time performance for all cardinality multipliers, and for $k = 100$, GreedySieve begins to exhibit sublinear time performance at a relatively low multiplier (less than 1.5).
+
+For glove-100, as shown in Figure \ref{fig:results:glove-100-scaling}, we observe that all of our algorithms exhibit superior performance compared to linear search for large cardinalities.
+However, for $k=10$, linear search outperforms all of our algorithms until about a cardinality multiplier of 4, when Sieve and Sieve with Separate Centers begin to show superior performance.
+For $k=100$, this transition happens even later, at a mutliplier of about 6, when GreedySieve begins to have better throughput.
+
+With the notably adversial dataset gist, we observe that for $k=10$, three of our algorithms--GreedySieve, Sieve, and Sieve with Separate Centers--begin to outperform linear search at cardinality multipliers of aboyt 4, 6, and 6 respectively.
+For low values of $k$, Sieve and Sieve with Separate Centers appear to the be the best-performing algorithms.
+For $k=100$, we observe that GreedySieve significantly outperforms all other algorithms after a cardinality multiplier of 4.
+At multipliers lower than this threshold value, linear search exhibits the best throughput.
+With gist, unlike with the other datasets we benchmarked on, we see that RepeatedRnn (for both values of $k$), Sieve (for $k=100$), and Sieve with Separate Centers (for $k=100$) all show performance inferior to linear search even at high cardinalies.
+
+Figure \ref{fig:results:random-scaling} displays results for a random datset of base cardinality 1000000. 
+These results illustrate the fact that absent a manifold structure in the dataset, it takes a much higher cardinality before CAKES begins to outperform linear search. 
+In particular, for $k=10$, we see that Sieve is the best-performing method as cardinality grows, but that it only begins to outperform linear search after a multiplier of about 15. 
+For $k=100$, we have that GreedySieve is the best-performing algorithm, but that it does not begin to have superior throughput until a multiplier of about 20. 
+RepeatedRnn shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
+This is unsurprising, given that RepeatedRnn relies on a low local fractal dimension around the query in order to quickly zero in on the correct radius for $k$ hits. 
+With a completely random dataset, which likely has much higher local fractal dimension, RepeatedRnn will meander for longer before finding this correct radius.
+
+Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
+This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities and consistently outperforms our other algorithms for $k=100$.
+We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
+While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
+Finally, we reiterate that the variation in performance across our algorithms based on dataset, value of $k$, and cardinality support our use of an autotuning function to select the best choice of algorithm.
+Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
 
@@
 \end{figure}
-
-Figures \ref{fig:results:fashion-mnist-scaling}, \ref{fig:results:glove-100-scaling}, and \ref{fig:results:gist-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-100 under cosine distance, and gist under Euclidean distance.
-Figure \ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset which does not obey any manifold structure. 
-The the horizontal axes in each figure denote the multiplier by which the cardinality of the original dataset has been increased with synthetic points.
-The vertical axes denote throughput in queries per second, on a logarithmic scale.
-
-In Figure \ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as cardinality grows, all of our algorithms exhibit significantly better throughput than linear search.
-Of our algorithms, GreedySieve consistently performs best for both $k=10$ and $k=100$.
-For $k=10$, GreedySieve exhibits sublinear time performance for all cardinality multipliers, and for $k = 100$, GreedySieve begins to exhibit sublinear time performance at a relatively low multiplier (less than 1.5).
-
-For glove-100, as shown in Figure \ref{fig:results:glove-100-scaling}, we observe that all of our algorithms exhibit superior performance compared to linear search for large cardinalities.
-However, for $k=10$, linear search outperforms all of our algorithms until about a cardinality multiplier of 4, when Sieve and Sieve with Separate Centers begin to show superior performance.
-For $k=100$, this transition happens even later, at a mutliplier of about 6, when GreedySieve begins to have better throughput.
-
-With the notably adversial dataset gist, we observe that for $k=10$, three of our algorithms--GreedySieve, Sieve, and Sieve with Separate Centers--begin to outperform linear search at cardinality multipliers of aboyt 4, 6, and 6 respectively.
-For low values of $k$, Sieve and Sieve with Separate Centers appear to the be the best-performing algorithms.
-For $k=100$, we observe that GreedySieve significantly outperforms all other algorithms after a cardinality multiplier of 4.
-At multipliers lower than this threshold value, linear search exhibits the best throughput.
-With gist, unlike with the other datasets we benchmarked on, we see that RepeatedRnn (for both values of $k$), Sieve (for $k=100$), and Sieve with Separate Centers (for $k=100$) all show performance inferior to linear search even at high cardinalies.
-
-Figure \ref{fig:results:random-scaling} displays results for a random datset of base cardinality 1000000. 
-These results illustrate the fact that absent a manifold structure in the dataset, it takes a much higher cardinality before CAKES begins to outperform linear search. 
-In particular, for $k=10$, we see that Sieve is the best-performing method as cardinality grows, but that it only begins to outperform linear search after a multiplier of about 15. 
-For $k=100$, we have that GreedySieve is the best-performing algorithm, but that it does not begin to have superior throughput until a multiplier of about 20. 
-RepeatedRnn shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
-This is unsurprising, given that RepeatedRnn relies on a low local fractal dimension around the query in order to quickly zero in on the correct radius for $k$ hits. 
-With a completely random dataset, which likely has much higher local fractal dimension, RepeatedRnn will meander for longer before finding this correct radius.
-
-Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
-This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities and consistently outperforms our other algorithms for $k=100$.
-We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
-While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
-Finally, we reiterate that the variation in performance across our algorithms based on dataset, value of $k$, and cardinality support our use of an autotuning function to select the best choice of algorithm.
-Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm differs significantly between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
 

### 2023-10-29 sections/1-introduction.tex
fixed intro to include definition of knn vs rho-NN earlier and definitions of exact and approx

@@
 
-In particular, $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
-Naive implementations of kNN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
+{\color{red} We now define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN). 
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
+the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \ \setminus \ S_q$, $f(x, q) > \max\{f(y, q): y \in S_q \}$;
+that is, for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
+On the other hand, $\rho$-nearest neighbors search aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
+that is, find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.}
+
+{\color{red} 
+Previous works have used the term \emph{approximate} search to refer to $\rho$-nearest neighbor search, but in this paper, 
+we reserve the term \emph{approximate} for $k$-NN algorihtms which may miss some of the $k$ nearest elements, and erroneously include some of the elements other than the truly nearest $k$.
+In other words, we call a $k$-NN algorithm \emph{approximate} if there exists a query, dataset, distance function, or value of $k$ for which its set $H$ of returned points satisfies the condition $H \triangle S_q \neq \emptyset$, where $\triangle$ denotes a symmetric difference.
+When we refer to an \emph{exact} $k$-NN algorithm, we describe an algorithm which, no matter the query, dataset, distance function, nor value of $k$, returns precisely the $k$ nearest neighbors to that query;
+no closer points are elided and no further points are included. }
+
+$k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
+Naive implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
 While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not exploit the geometric and topological structure inherent in large datasets.
@@
 In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$, and small values of $k$ may be necessary when classes are not well-separated.
-
-For the sake of clarity, it is important to be precise about what we mean by \emph{exact} $k$-NN, and exact search in general.
-In the context of $\rho$-NN search, exact search typically means returning only those data points that are identical to the query.
-There are many algorithms that perform well here, such as FM-Index~\cite{ferragina2000opportunistic}, which are less applicable to approximate search.
-In contrast, in the context of $k$-NN search, while all approaches return exactly $k$ elements (ignoring how distance ties are handled), some approaches may miss some of the $k$ nearest elements, and erroneously include some of the elements outside the truly nearest $k$.
-In any case, $k$-NN is commonly used for inference;
-the \emph{prediction} of a class for a query is obviously not guaranteed to be correct.
-When we refer to \emph{exact} $k$-NN, we refer to precisely the $k$ nearest neighbors being returned;
-no closer points are elided and no further points are included.
-This has implications for the accuracy of classification, but does not guarantee perfect classification accuracy.
-
-There is some evidence that non-metrics such as cosine distance perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
+There is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
 this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could exhibit suboptimal accuracy.

### 2023-11-12 sections/2-methods.tex
fixing error in repeated rnn explanation
removed scaling behavior of radii section

@@
 
-\subsubsection {Scaling Behavior of Cluster Radii}
-\label{subsubsec:methods:clustering:guaranteed-decrease-in-cluster-radii}
+% % WE WILL USE THIS FOR COMPRESSION, so removing it for now
 
-While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:clustering:partition}), unfortunately, this assumption is incorrect.
-Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii;
-in particular, we prove in this section that cluster radii will stop increasing after at most $d$ partitions, where $d$ is the fractal dimensionality of the dataset.
+% \subsubsection {Scaling Behavior of Cluster Radii}
+% \label{subsubsec:methods:clustering:guaranteed-decrease-in-cluster-radii}
 
-We can describe a $d$-dimensional distribution of data by choosing some set of $d$ mutually orthogonal axes.
-Let $2R$ be the maximum pairwise distance among the points in the dataset. 
-We choose the axes such that the two points that are $2R$ apart lie \emph{along one of the axes}. 
-Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
-In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance \emph{along every axis}. 
-Such a distribution would also produce a balanced cluster tree.
+% While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:clustering:partition}), unfortunately, this assumption is incorrect.
+% Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii;
+% in particular, we prove in this section that cluster radii will stop increasing after at most $d$ partitions, where $d$ is the fractal dimensionality of the dataset.
 
-In this case, Partition will select a maximally distant pair of points to use as poles;
-without loss of generality, say it chooses two points at the extrema one of the $d$ axes. 
-After one partition, the maximum pairwise distance along that axis will be halved, i.e. bounded above by $R$.
-The next recursive Partition will select another of the $d$ axes, and once again, the distance along that axis will be bounded above by $R$
-Thus, after at most $d$ partitions, the maximum pairwise distance along each axis will be bounded above by $R$. 
-The overall (i.e., not restricted to one axis) maximum pairwise distance 
-will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes. 
-See Figure~\ref{fig:methods:scaling-behavior} for an example.
+% We can describe a $d$-dimensional distribution of data by choosing some set of $d$ mutually orthogonal axes.
+% Let $2R$ be the maximum pairwise distance among the points in the dataset. 
+% We choose the axes such that the two points that are $2R$ apart lie \emph{along one of the axes}. 
+% Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
+% In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance \emph{along every axis}. 
+% Such a distribution would also produce a balanced cluster tree.
 
-Thus, starting with a cluster $C$ of radius $R$, after at most $d$ Partitions, the descendants of $C$ will each have radius
-bounded above by $\frac{R}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at 
-most $d$ partitions.
+% In this case, Partition will select a maximally distant pair of points to use as poles;
+% without loss of generality, say it chooses two points at the extrema one of the $d$ axes. 
+% After one partition, the maximum pairwise distance along that axis will be halved, i.e. bounded above by $R$.
+% The next recursive Partition will select another of the $d$ axes, and once again, the distance along that axis will be bounded above by $R$
+% Thus, after at most $d$ partitions, the maximum pairwise distance along each axis will be bounded above by $R$. 
+% The overall (i.e., not restricted to one axis) maximum pairwise distance 
+% will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes. 
+% See Figure~\ref{fig:methods:scaling-behavior} for an example.
 
-Note that, in practice, we never see a balanced clustering.
-Partition produces unbalanced trees due to the varying density of the sampling in different regions of the manifold and the low dimensional ``shape'' of the manifold.
-Further, in practice, the cluster radii decrease by a factor much larger than $\frac{\sqrt{2}}{2}$ with almost every partition;
-the upper bound of $d$ partitions is seldom realized.
+% Thus, starting with a cluster $C$ of radius $R$, after at most $d$ Partitions, the descendants of $C$ will each have radius
+% bounded above by $\frac{R}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at 
+% most $d$ partitions.
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
-    \caption{
-        Scaling behavior of radii on a uniform-density disk, which represents the worst-case scenario for a two-dimensional distribution. 
-        $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
-        After one application of Algorithm~\ref{alg:methods:clustering:partition}, we have $C_1$, with radius $R_1$ and center $o_1$.
-        The right triangle formed by $o_0$, $o_1$, and $y_+$ in $C_1$ shows that $R_0 < R_1$.
-        Hence, the radius of a child cluster can be larger than its parent.
-        However, after another application of Algorithm~\ref{alg:methods:clustering:partition}, we have consumed all $d$ orthogonal axes, 
-        as shown in $C_2$.
-        Now, clearly $R_2 < R_0$.
-        In fact, for each subsequent application of Algorithm~\ref{alg:methods:clustering:partition}, the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.
-    }
-    \label{fig:methods:scaling-behavior}
-\end{figure}
+% Note that, in practice, we never see a balanced clustering.
+% Partition produces unbalanced trees due to the varying density of the sampling in different regions of the manifold and the low dimensional ``shape'' of the manifold.
+% Further, in practice, the cluster radii decrease by a factor much larger than $\frac{\sqrt{2}}{2}$ with almost every partition;
+% the upper bound of $d$ partitions is seldom realized.
+
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
+%     \caption{
+%         Scaling behavior of radii on a uniform-density disk, which represents the worst-case scenario for a two-dimensional distribution. 
+%         $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center. 
+%         After one application of Algorithm~\ref{alg:methods:clustering:partition}, we have $C_1$, with radius $R_1$ and center $o_1$.
+%         The right triangle formed by $o_0$, $o_1$, and $y_+$ in $C_1$ shows that $R_0 < R_1$.
+%         Hence, the radius of a child cluster can be larger than its parent.
+%         However, after another application of Algorithm~\ref{alg:methods:clustering:partition}, we have consumed all $d$ orthogonal axes, 
+%         as shown in $C_2$.
+%         Now, clearly $R_2 < R_0$.
+%         In fact, for each subsequent application of Algorithm~\ref{alg:methods:clustering:partition}, the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.
+%     }
+%     \label{fig:methods:scaling-behavior}
+% \end{figure}
 
@@
 In other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radial increase is needed.
-Thus, the factor of radius increase should be \emph{directly} related to the local fractal dimension.
+Thus, the factor of radius increase should be \emph{inversely} related to the local fractal dimension.
 Once $|H| >= k$, we sort the points by their distance to the query and return the $k$ closest points.

### 2023-11-12 sections/2-methods.tex
added new scaling plots and commentary (more to come)

@@
 This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor is typically much faster than $k$-nearest neighbors search.
-In addition to sharding, we perform some naive auto-tuning to select the optimal $k$-NN algorithm to be used with a given dataset.
+In addition to sharding, we perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to be used with a given dataset.
 
@@
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high fractal dimension. 
-Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
+Intuitively, the factor by which we increase the radius should also be \emph{inversely} related to the number of points found so far. 
 Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.

### 2023-11-12 sections/4-results.tex
added new scaling plots and commentary (more to come)

@@
 
-Figures \ref{fig:results:fashion-mnist-scaling}, \ref{fig:results:glove-100-scaling}, and \ref{fig:results:gist-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-100 under cosine distance, and gist under Euclidean distance.
-Figure \ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset which does not obey any manifold structure. 
-The the horizontal axes in each figure denote the multiplier by which the cardinality of the original dataset has been increased with synthetic points.
-The vertical axes denote throughput in queries per second, on a logarithmic scale.
+Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, deep-image under cosine distance, and glove-25 under cosine distance.
+Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset which does not obey any manifold structure. 
+The the horizontal axes in each figure denote the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
+The vertical axes denote throughput in queries per second. Both axes are on a logarithmic scale. In this section, we discuss results only for $k = 10$. 
+To see results with other values of $k$, refer to {\color{red} ADD SECTION}.
 
-In Figure \ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as cardinality grows, all of our algorithms exhibit significantly better throughput than linear search.
-Of our algorithms, GreedySieve consistently performs best for both $k=10$ and $k=100$.
-For $k=10$, GreedySieve exhibits sublinear time performance for all cardinality multipliers, and for $k = 100$, GreedySieve begins to exhibit sublinear time performance at a relatively low multiplier (less than 1.5).
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, each of CAKES's four algorithms exhibits significantly higher throughput than na\"{i}ve linear search, plotted in orange.
+Amongst them, GreedySieve, plotted in blue, consistently performs best, exhibiting sublinear time performance for nearly all cardinality multipliers. 
+We see that for multipliers around $10^3$, Greedy Sieve exhibits throughput which is nearly three orders of magnitude higher than that of linear search.
 
-For glove-100, as shown in Figure \ref{fig:results:glove-100-scaling}, we observe that all of our algorithms exhibit superior performance compared to linear search for large cardinalities.
-However, for $k=10$, linear search outperforms all of our algorithms until about a cardinality multiplier of 4, when Sieve and Sieve with Separate Centers begin to show superior performance.
-For $k=100$, this transition happens even later, at a mutliplier of about 6, when GreedySieve begins to have better throughput.
+With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of our four algorithms outperforms linear search at high cardinality multipliers. 
+On this dataset, Repeated $\rho$-NN, plotted in green, is consistently the best performing algorithm. {\color{red} Revive lfd violin plots to add explanation for this.}
 
-With the notably adversial dataset gist, we observe that for $k=10$, three of our algorithms--GreedySieve, Sieve, and Sieve with Separate Centers--begin to outperform linear search at cardinality multipliers of aboyt 4, 6, and 6 respectively.
-For low values of $k$, Sieve and Sieve with Separate Centers appear to the be the best-performing algorithms.
-For $k=100$, we observe that GreedySieve significantly outperforms all other algorithms after a cardinality multiplier of 4.
-At multipliers lower than this threshold value, linear search exhibits the best throughput.
-With gist, unlike with the other datasets we benchmarked on, we see that RepeatedRnn (for both values of $k$), Sieve (for $k=100$), and Sieve with Separate Centers (for $k=100$) all show performance inferior to linear search even at high cardinalies.
+In Figure~\ref{fig:results:sift-scaling}, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search beginning at a multiplier of approximately 7.
+For this dataset, Sieve, plotted in red, is the highest throughput algorithm for all multipliers greater than 7.
 
-Figure \ref{fig:results:random-scaling} displays results for a random datset of base cardinality 1000000. 
-These results illustrate the fact that absent a manifold structure in the dataset, it takes a much higher cardinality before CAKES begins to outperform linear search. 
-In particular, for $k=10$, we see that Sieve is the best-performing method as cardinality grows, but that it only begins to outperform linear search after a multiplier of about 15. 
-For $k=100$, we have that GreedySieve is the best-performing algorithm, but that it does not begin to have superior throughput until a multiplier of about 20. 
-RepeatedRnn shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
-This is unsurprising, given that RepeatedRnn relies on a low local fractal dimension around the query in order to quickly zero in on the correct radius for $k$ hits. 
-With a completely random dataset, which likely has much higher local fractal dimension, RepeatedRnn will meander for longer before finding this correct radius.
+Figure \ref{fig:results:random-scaling} displays results for a random datset with the same dimensionality and cardinality as Sift (i.e., 128 and 1,000,000 respectively). 
+These results illustrate the fact that \emph{ceteris paribus}, absent a manifold structure in the dataset, CAKES's algorithms do not begin to outperform linear search until the cardinality multiplier is around 10.
+Repeated $\rho$-NN shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
+This is unsurprising, given that RepeatedRnn relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
+With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, RepeatedRnn will meander for longer before finding this correct radius.
 
 Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
-This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities and consistently outperforms our other algorithms for $k=100$.
-We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
-While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
-Finally, we reiterate that the variation in performance across our algorithms based on dataset, value of $k$, and cardinality support our use of an autotuning function to select the best choice of algorithm.
-Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
+This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities.
+% We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
+% While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
+Finally, we reiterate that the variation in performance of our algorithms across different datasets and cardinalities support our use of an autotuning function to select the best choice of algorithm. 
+% Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
 
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/fashion-mnist_0.1_scaling.png}
+    \includegraphics[width=3.4in]{images/result_plots/fashion-mnist_10_scaling.png}
     \caption{
-        Fashion-mnist scaling
+        Fashion-mnist Scaling
     }
@@
 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{images/result_plots/deep-image_10_scaling.png}
+%     \caption{
+%         Deep-image Scaling
+%     }
+%     \label{fig:results:deep-image-scaling}
+% \end{figure}
+
 \begin{figure}[ht!]
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/glove-100_0.1_scaling.png}
+    \includegraphics[width=3.4in]{images/result_plots/glove-25_10_scaling.png}
     \caption{
-        Glove-100 scaling
+        Glove-25 Scaling
     }
-    \label{fig:results:glove-100-scaling}
+    \label{fig:results:glove-25-scaling}
 \end{figure}
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/gist_0.1_scaling.png}
+    \includegraphics[width=3.4in]{images/result_plots/sift_10_scaling.png}
     \caption{
-        Gist scaling
+        Sift Scaling
     }
-    \label{fig:results:gist-scaling}
+    \label{fig:results:sift-scaling}
 \end{figure}
+
+% We need to redo this plot so that we do not actually augment random datasets but instead take larger and larger 
+% random datasets
 
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/random_0.1_scaling.png}
+    \includegraphics[width=3.4in]{images/result_plots/random_placeholder_10_scaling.png}
     \caption{
-        Random scaling
+        Random Scaling
     }

### 2023-11-12 sections/4-results.tex
headers for new datasets and naming fixes

@@
 Repeated $\rho$-NN shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
-This is unsurprising, given that RepeatedRnn relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
-With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, RepeatedRnn will meander for longer before finding this correct radius.
+This is unsurprising, given that Repeated $\rho$-NN relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
+With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, Repeated $\rho$-NN will iterate for longer before finding this correct radius.
 

### 2023-11-13 sections/2-methods.tex
better presentation of repeated rnn complexity
fixing inaccuracy

@@
 
-To determine the asymptotic complexity of fine search, we must estimate $|F|$, the number of clusters for which we must look at all of the cluster's points.
-With $\rho$-NN search, $F$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query, where $r$ is the search radius, and $\hat{r_c}$ is the mean radius of leaf clusters, as in \cite{yu2015entropy}.
+To determine the asymptotic complexity of fine search, we must estimate $|H|$, the expected number of hits we examine.
+With $\rho$-NN search, $H$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query, where $r$ is the search radius, and $\hat{r_c}$ is the mean radius of leaf clusters, as in \cite{yu2015entropy}.
 With $k$-NN by Repeated $\rho$-NN, however, the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which is query-dependent.
-Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of $F$ to be the union of clusters which are within a distance of $r' + \hat{r_c}$ from the query, where $r'$ is the distance to the $k$th nearest neighbor and $\hat{r_c}$ is defined as before.
+Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of $H$ to be the union of clusters which are within a distance of $r' + \hat{r_c}$ from the query, where $r'$ is the distance to the $k$th nearest neighbor and $\hat{r_c}$ is defined as before.
 
@@
 \begin{equation*}
-    |F| \leq \gamma |B(q, r')|\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
+    |H| \leq \gamma |B(q, r')|\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
 \end{equation*}
@@
 \begin{equation}
-    |F| \leq \gamma k\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
+    |H| \leq \gamma k\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
     \label{eq:methods:knn-by-rnn-fine-search}
@@
 
-\begin{equation}
-    \mathcal{O}\left(\log_2{\mathcal{N}_{\hat{r_c}}(X)} + k\left(\frac{2\hat{r_c}}{\hat{r}}\right)^d\left(\frac{k}{\hat{C}}\right)^{\frac{-d}{\hat{d}}}\right)
+\begin{gather}
+    \mathcal{O}\Bigg(
+    \underbrace{\log_2{\mathcal{N}_{\hat{r_c}}(X)}}_{\textrm{metric entropy}} + 
+    \overbrace{k}^{\textrm{output size}}
+    \underbrace{\left(\frac{2\hat{r_c}}{\hat{r}}\right)^d\left(\frac{k}{\hat{C}}\right)^{\frac{-d}{\hat{d}}}}_{\textrm{scaling factor}}\Bigg)
     \label{eq:methods:knn-by-rnn-complexity}
-\end{equation}
+\end{gather}
 

### 2023-11-19 sections/2-methods.tex
cite simev

@@
 We expect CAKES to perform well as the cardinality of datasets increases, and to explore this scaling behavior, we augmented datasets in the ANN-benchmarks suite by adding random points within a 
-small noise threshold of existing points. This allowed us to isolate the effect of \emph{cardinality} on the performance of CAKES 
+small noise threshold of existing points; this technique has been used in a number of bioinformatics contexts~\cite{kumar2009augmented, kumar2010recognition, daniels2012smurflite, velasco2018combining, lu2020evolution}.
+This allowed us to isolate the effect of \emph{cardinality} on the performance of CAKES
 from that of other factors such as dimensionality, metric, or topological structure of the dataset.

### 2023-11-19 sections/2-methods.tex
start of rho-NN overlapping children

@@
 The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
-The point $l$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}.
-The point $r$ which is furthest from $l$ is designated the \emph{right pole}.
-We then partition the cluster into a \emph{left child} and \emph{right child}, where the left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all points in the cluster which are closer to $r$ than to $l$.
-Without loss of generality, we assign to the left child those points which are equidistant from $l$ and $r$.
+The point $\ell$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}.
+The point $r$ which is furthest from $\ell$ is designated the \emph{right pole}.
+We then partition the cluster into a \emph{left child} and \emph{right child}, where the left child contains all points in the cluster which are closer to $\ell$ than to $r$, and the right child contains all points in the cluster which are closer to $r$ than to $\ell$.
+Without loss of generality, we assign to the left child those points which are equidistant from $\ell$ and $r$.
 Starting from a root-cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum or some other user-specified stopping criterion is met.
@@
     \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
-    \item $\delta_{max} = \delta + r$ is the distance from the query to the theoretically farthest possible instance in $C$.
-    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest possible instance in $C$.
+    \item $\delta_{\max} = \delta + r$ is the distance from the query to the theoretically farthest possible instance in $C$.
+    \item $\delta_{\min} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest possible instance in $C$.
 \end{itemize}
@@
 duplicates of the same point (i.e., has cardinality greater than 1 but contains only one unique point).
-A singleton clearly has zero radius, and so $\delta_{max} = \delta = \delta_{min}$ for these clusters.
+A singleton clearly has zero radius, and so $\delta_{\max} = \delta = \delta_{\min}$ for these clusters.
 Hence, we sometimes overload the above notation to also refer to distance from a query to a point;
-in these cases, we also have that $\delta_{max} = \delta = \delta_{min}$.
+in these cases, we also have that $\delta_{\max} = \delta = \delta_{\min}$.
 
@@
 
-We conduct $\rho$-nearest neighbors search as described in \cite{ishaq2019clustered}, but with the following improvement:
-when a cluster overlaps with the query ball, instead of  always proceeding with both of its children, we proceed only with those children which also overlap with the query ball.
-This improvement is described in Algorithm \ref{alg:methods:rnn-search}.
+We conduct $\rho$-nearest neighbors search as described in~\cite{ishaq2019clustered}, but with the following improvement:
+when a cluster overlaps the query ball, instead of always proceeding with both of its children, we proceed only with those children which also overlap the query ball.
+
+To determine whether both children overlap the query ball, we consider the following geometry. 
+Let $c$ denote the center of a cluster, $\ell$ and $r$ its left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ the query, and $\rho$ the search radius, as shown in Figure [ADD FIGURE].
+Without loss of generality, we assume that $\overline{qr} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
+
+Now let $m$ be the midpoint of $\ell$ and $r$.
+Let $d$ denote the distance from $q$ to the perpendicular bisector of $\overline{lr}$, as shown in Figure [ADD FIGURE].
+Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$, (i.e., if $d$ is to the left of the perpendicular bisector of $\overline{\ell r}$),
+the left child does not overlap the query ball. In this case, we proceed only with the right child. If $d \leq \rho$, 
+then both child clusters overlap the query ball.
+
+We can do a quick check using values computed at partition time to determine whether $d \leq \rho$.
 
@@
             \IF{$cluster.left.\delta$ $\leq$ $r$ + $cluster.left.radius$}
-                \STATE $\rho$-NN($cluster.left, query, r$)
+                \STATE{$\rho$-NN($cluster.left, query, r$)}
             \ENDIF
@@
 
-First, we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{max}$ such that 
-$q_{\tau}$ and all elements in $Q$ with smaller $\delta_{max}$ collectively contain at least $k$ points.
+First, we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{\max}$ such that 
+$q_{\tau}$ and all elements in $Q$ with smaller $\delta_{\max}$ collectively contain at least $k$ points.
 
 To find this element, we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of clusters in $Q$ by having each cluster count a number of times equal to its cardinality.
-Our modified version of this algorithm finds the smallest index $i$ such that all elements in $Q$ with $\delta_{max}$ less than or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater than or equal to $k$.
-We consider $Q[i]$ to be our $q_{\tau}$, and our threshold $\tau$ to be $Q[i]$'s $\delta_{max}$.
+Our modified version of this algorithm finds the smallest index $i$ such that all elements in $Q$ with $\delta_{\max}$ less than or equal to $Q[i]$'s $\delta_{\max}$ collectively have cardinality greater than or equal to $k$.
+We consider $Q[i]$ to be our $q_{\tau}$, and our threshold $\tau$ to be $Q[i]$'s $\delta_{\max}$.
 
-We then remove from $Q$ any element whose $\delta_{min}$ is greater than $\tau$.
+We then remove from $Q$ any element whose $\delta_{\min}$ is greater than $\tau$.
 Then we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
@@
         \STATE Return $Q$
-    \end{algorithmic}
+    \end{algorithmic}How do we algorithmically determine the depth to which we
 \end{algorithm}
@@
 
-We first let $Q$ be a min-queue of clusters prioritized by $\delta_{min}$.
+We first let $Q$ be a min-queue of clusters prioritized by $\delta_{\min}$.
 $Q$ starts containing only the root cluster.
@@
 
-While $H$ has fewer than $k$ points or $\delta$ of the top priority element in $H$ is greater than $\delta_{min}$ of the top priority element in $Q$, we do the following:
+While $H$ has fewer than $k$ points or $\delta$ of the top priority element in $H$ is greater than $\delta_{\min}$ of the top priority element in $Q$, we do the following:
 
@@
 \begin{algorithmic}
-    \STATE $Q \Leftarrow$ priority queue
-    \STATE $Q.push(root)$
-    \STATE $H \Leftarrow$ priority queue
-    \WHILE{$|H| < k \lor H.peek.\delta > Q.peek.\delta_{min}$}
+    \STATE{$Q \Leftarrow$ priority queue}
+    \STATE{$Q.push(root)$}
+    \STATE{$H \Leftarrow$ priority queue}
+    \WHILE{$|H| < k \lor H.peek.\delta > Q.peek.\delta_{\min}$}
         \WHILE{$\neg Q.peek.isLeaf$}
-            \STATE $[l, r] \Leftarrow Q.pop().children$
-            \STATE $Q.push([l, r])$
+            \STATE{$[l, r] \Leftarrow Q.pop().children$}
+            \STATE{$Q.push([l, r])$}
         \ENDWHILE
-        \STATE $leaf \Leftarrow Q.pop()$
-        \STATE $H.push(leaf)$
+        \STATE{$leaf \Leftarrow Q.pop()$}
+        \STATE{$H.push(leaf)$}
         \WHILE{$|H| > k$}
-            \STATE $H.pop()$
+            \STATE{$H.pop()$}
         \ENDWHILE

### 2023-11-21 sections/1-introduction.tex
added new scaling plots and commentary (more to come)
small fixes

@@
 
-{\color{red} We now define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN). 
+We now define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN). 
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
@@
 On the other hand, $\rho$-nearest neighbors search aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
-that is, find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.}
+that is, find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
 
@@
 $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
-Naive implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
+Na\"{i}ve implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
 While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not exploit the geometric and topological structure inherent in large datasets.

### 2023-11-21 sections/2-methods.tex
small fixes

@@
 We conduct $\rho$-nearest neighbors search as described in~\cite{ishaq2019clustered}, but with the following improvement:
-when a cluster overlaps the query ball, instead of always proceeding with both of its children, we proceed only with those children which also overlap the query ball.
+when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which could contain points in the query ball.
 
-To determine whether both children overlap the query ball, we consider the following geometry. 
-Let $c$ denote the center of a cluster, $\ell$ and $r$ its left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ the query, and $\rho$ the search radius, as shown in Figure [ADD FIGURE].
+To determine whether both children can contain points in the query ball, we consider the following geometry. 
+Let $\ell$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ denote the query, and $\rho$ denote the search radius, as shown in Figure [ADD FIGURE].
 Without loss of generality, we assume that $\overline{qr} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
 
-Now let $m$ be the midpoint of $\ell$ and $r$.
+Now let $m$ be the midpoint of $\overline{\ell r}$.
 Let $d$ denote the distance from $q$ to the perpendicular bisector of $\overline{lr}$, as shown in Figure [ADD FIGURE].
-Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$, (i.e., if $d$ is to the left of the perpendicular bisector of $\overline{\ell r}$),
-the left child does not overlap the query ball. In this case, we proceed only with the right child. If $d \leq \rho$, 
-then both child clusters overlap the query ball.
+Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$,
+the left child cannot contain points in the query ball. In this case, we proceed only with the right child. If $d \leq \rho$, 
+then both child clusters could contain points in the query ball, and we proceed with both of them.
 
-We can do a quick check using values computed at partition time to determine whether $d \leq \rho$.
+We can do a quick check using values computed during partitioning (see Section~\ref{subsubsec:methods:cluster-partitioning}) to determine whether $d \leq \rho$. 
+We note that $d = ||mq'|| = ||mr||-||q'r||$.
+Since $m$ is the midpoint of $\overline{\ell r}$, we have that $||mr|| = \tfrac{||\ell r||}{2}$, so it remains to find $||q'r||$.
+To do this, we use the Law of Cosines with $\triangle \ell rq$.
+Let $\theta$ denote $\angle lrq$, as shown in Figure [ADD FIGURE].
+By the Law of Cosines, we have that
+\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{qr}||^2 + ||\overline{\ell r}||^2 - ||\overline{q\ell}||^2}{2||\overline{\ell r}||}.\end{equation}. 
+On the other hand, since $\triangle q'qr$ is a right triangle,
+
+\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{q'r}||}{||\overline{qr}||}.\end{equation}
+
+To check this, it suffices to verify that \begin{equation}(||\overline{q\ell}|| + ||\overline{qr}||)(||\overline{q\ell}||- ||\overline{qr}||) \leq 2||\overline{\ell r}||\rho. \end{equation}
 

### 2023-11-21 sections/1-introduction.tex
finished overlapping children

@@
 
-{\color{red} 
 Previous works have used the term \emph{approximate} search to refer to $\rho$-nearest neighbor search, but in this paper, 
-we reserve the term \emph{approximate} for $k$-NN algorihtms which may miss some of the $k$ nearest elements, and erroneously include some of the elements other than the truly nearest $k$.
-In other words, we call a $k$-NN algorithm \emph{approximate} if there exists a query, dataset, distance function, or value of $k$ for which its set $H$ of returned points satisfies the condition $H \triangle S_q \neq \emptyset$, where $\triangle$ denotes a symmetric difference.
+we reserve the term \emph{approximate} for $k$-NN search algorithms which may miss some of the $k$ nearest elements, and erroneously include some of the elements other than the truly nearest $k$.
+In other words, we call a $k$-NN search algorithm \emph{approximate} if there exists a query, dataset, distance function, or value of $k$ for which its set $H$ of returned points satisfies the condition $H \triangle S_q \neq \emptyset$, where $\triangle$ denotes a symmetric difference.
 When we refer to an \emph{exact} $k$-NN algorithm, we describe an algorithm which, no matter the query, dataset, distance function, nor value of $k$, returns precisely the $k$ nearest neighbors to that query;
-no closer points are elided and no further points are included. }
+no closer points are elided and no further points are included.
 

### 2023-11-21 sections/2-methods.tex
finished overlapping children

@@
 
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[scale=0.5]{images/geometry/overlapping-children-1.pdf}
+    \caption{
+        Triangle formed by query $q$, and poles $\ell$ and $r$. Query ball of radius $\rho$ centered at $q$. 
+    }
+    \label{fig:methods:overlapping-children-1}
+\end{figure}
+
 To determine whether both children can contain points in the query ball, we consider the following geometry. 
-Let $\ell$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ denote the query, and $\rho$ denote the search radius, as shown in Figure [ADD FIGURE].
-Without loss of generality, we assume that $\overline{qr} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
+Let $\ell$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ denote the query, and $\rho$ denote the search radius, as shown in Figure~\ref{fig:methods:overlapping-children-1}.
+Without loss of generality, we assume that $\overline{qr \vphantom \ell} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[scale=0.5]{images/geometry/overlapping-children-3.pdf}
+    \caption{
+        $m$ denotes the midpoint of $\overline{\ell r}$.
+$q'$ is the projection of $q$ onto $\overline{\ell r}$.
+$d$ is the distance from $q$ to the perpendiculat bisector of $\overline{\ell r}$.
+If $d \leq \rho$, both the left child and right child clusters could contain points in the query ball.
+Otherwise, only the right child cluster could contain such points.
+    }
+    \label{fig:methods:overlapping-children-3}
+\end{figure}
 
 Now let $m$ be the midpoint of $\overline{\ell r}$.
-Let $d$ denote the distance from $q$ to the perpendicular bisector of $\overline{lr}$, as shown in Figure [ADD FIGURE].
+Let $d$ denote the distance from $q$ to the perpendicular bisector of $\overline{lr}$, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
 Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$,
@@
 We can do a quick check using values computed during partitioning (see Section~\ref{subsubsec:methods:cluster-partitioning}) to determine whether $d \leq \rho$. 
-We note that $d = ||mq'|| = ||mr||-||q'r||$.
-Since $m$ is the midpoint of $\overline{\ell r}$, we have that $||mr|| = \tfrac{||\ell r||}{2}$, so it remains to find $||q'r||$.
+We note that $d = ||\overline{mq'}|| = ||\overline{mr \vphantom {q'}}||-||\overline{q'r}||$.
+Since $m$ is the midpoint of $\overline{\ell r}$, we have that $||\overline{mr}|| = \tfrac{||\overline{\ell r}||}{2}$, so \begin{equation} d = \tfrac{||\overline{\ell r}||}{2} - ||\overline{q'r}||.
+\label{eq:methods:d-expression}
+\end{equation}
+
+It remains to find $||\overline{q'r}||$.
 To do this, we use the Law of Cosines with $\triangle \ell rq$.
-Let $\theta$ denote $\angle lrq$, as shown in Figure [ADD FIGURE].
+Let $\theta$ denote $\angle lrq$, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
 By the Law of Cosines, we have that
-\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{qr}||^2 + ||\overline{\ell r}||^2 - ||\overline{q\ell}||^2}{2||\overline{\ell r}||}.\end{equation}. 
+\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{qr\vphantom \ell}||^2 + ||\overline{\ell r}||^2 - ||\overline{q\ell}||^2}{2||\overline{\ell r}||||\overline{qr \vphantom \ell}||}.\end{equation}
 On the other hand, since $\triangle q'qr$ is a right triangle,
 
-\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{q'r}||}{||\overline{qr}||}.\end{equation}
+\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{q'r}||}{||\overline{qr \vphantom \ell}||}.\end{equation}
 
-To check this, it suffices to verify that \begin{equation}(||\overline{q\ell}|| + ||\overline{qr}||)(||\overline{q\ell}||- ||\overline{qr}||) \leq 2||\overline{\ell r}||\rho. \end{equation}
+Equating the two expressions for $\text{cos}(\theta)$ and solving for $||\overline{q'r}||$, we have \begin{equation} ||\overline{q'r}|| = \tfrac{||\overline{qr\vphantom \ell}||^2 + ||\overline{\ell r}||^2 - ||\overline{q\ell}||^2}{2 ||\overline{\ell r}||}.
+\end{equation}
+
+Substituting this back in equation~\ref{eq:methods:d-expression} gives 
+\begin{equation} d = \tfrac{||\overline{q\ell}||^2 - ||\overline{qr \vphantom \ell}||^2}{2||\overline{lr}||}.\end{equation}
+
+Then to check if $d \leq \rho$, it suffices to verify that \begin{equation}(||\overline{q\ell}|| + ||\overline{qr \vphantom \ell}||)(||\overline{q\ell}||- ||\overline{qr \vphantom \ell}||) \leq 2||\overline{\ell r}||\rho.
+\end{equation}
 

### 2023-12-10 supplement.tex
supplement

\documentclass{article}

% \usepackage{arxiv} # uncomment for preprint

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% \usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}

\usepackage{placeins}
% \usepackage{algpseudocode}
\usepackage{algorithm}

% \algnewcommand\algorithmicforeach{\textbf{foreach}}
% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

\usepackage{hyperref}

\begin{document}

\renewcommand{\thetable}{S\arabic{table}}

\renewcommand{\thefigure}{S\arabic{figure}}

\section{Datasets}

\section{Results}

\end{document}

### 2023-12-10 sections/2-methods.tex
added deltas image
fix to Sieve

@@
 
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following properties:
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[scale=0.5]{images/geometry/deltas.png}
+    \caption{}
+    \label{fig:methods:deltas}
+\end{figure}
+
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following properties, as illustrated in Figure~\ref{fig:methods:deltas}.
 
@@
 
-In this algorithm, we perform $\rho$-nearest neighbors search starting with a small search radius, and repeatedly increasing the radius until we find $k$ neighbors.
+In this algorithm, we perform $\rho$-nearest neighbors search starting with a small search radius,
+\subsubsection{Sieve with Separate Centers}
+\label{subsubsec:methods:knn-search:sieve2}
+
+This algorithm is the same as Sieve, but with the following modification:
+clusters in $Q$ are represented twice -- once as their center and once as the rest of their points.
+The clusters' multiplicities in $Q$ are adjusted to be one less than their cardinalities.
+Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, thus allowing us to eliminate some clusters from $Q$ earlier in the process.
+and repeatedly increasing the radius until we find $k$ neighbors.
 
@@
 
-In this algorithm, we begin by letting $Q$ be a list of clusters and points, starting with only the root cluster. 
+In this algorithm, we begin by letting $Q$ be a list of clusters and points. 
+{ \color{red} We represent each cluster twice in $Q$ -- once as just its center and once as the rest of its points.
+$Q$ is initialized with two elements: the root cluster and the root cluster's center.
+}
+
 While $Q$ contains at least one cluster (i.e., $Q$ is not a list of only points), we repeat the process described in 
@@
         \STATE Return $Q$
-    \end{algorithmic}How do we algorithmically determine the depth to which we
+    \end{algorithmic}
 \end{algorithm}
-
-\subsubsection{Sieve with Separate Centers}
-\label{subsubsec:methods:knn-search:sieve2}
-
-This algorithm is the same as Sieve, but with the following modification:
-clusters in $Q$ are represented twice -- once as their center and once as the rest of their points.
-The clusters' multiplicities in $Q$ are adjusted to be one less than their cardinalities.
-Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, thus allowing us to eliminate some clusters from $Q$ earlier in the process.
 

### 2023-12-11 sections/4-results.tex
fix to Sieve
messing around with csv and latex
fixed issues in csv_to_latex and added table for fashion-mnist

@@
 \label{sec:results}
-
-\subsection{Scaling Behavior}
 
@@
 The vertical axes denote throughput in queries per second. Both axes are on a logarithmic scale. In this section, we discuss results only for $k = 10$. 
-To see results with other values of $k$, refer to {\color{red} ADD SECTION}.
+To see results with other values of $k$, refer to our Supplement section.
 
@@
 
-\subsection{Results against Competitors}
-
 % We did not run any of the competitor's algorithms.
@@
 
-% \begin{table*}[!t]
-%     % \renewcommand{\arraystretch}{1.15}
-%     \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=10$}
-%     \label{table:results:ann-10}
-%     \vskip 0.15in
-%     \begin{center}
-%         \begin{small}
-%             \begin{sc}
-%                 \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-%                     \hline
-%                     \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-%                     \cline{2-9}
-%                     &               Recall & QPS      & Recall & QPS    & Recall & QPS    & Recall & QPS \\
-%                     \hline
-%                     fashion-mnist & 1.000  & 188.03    & 1.000 & 52.91  & 1.000* & 339.74 & 1.000 & 141,800 \\
-%                     \hline
-%                     gist          & 1.000 & 3.55  & 1.000 & 2.63        & 0.953 & 149.58  & 1.000 & 76,420 \\
-%                     \hline
-%                     glove-25      & 1.000 & 78.78  & 1.000 & 33.52      & 1.000 & 552.67  & 0.798 & 94,580 \\
-%                     \hline
-%                     glove-100     & 1.000 & 21.16  & 1.000 & 16.52      & 0.998 & 155.90  & 0.904 & 4,566 \\
-%                     \hline
-%                     sift          & 1.000 &  24.52 & 1.000 & 16.40      & 1.000 & 275.07  & 1.000 & 357,400 \\
-%                     \hline
-%                 \end{tabular}
-%             \end{sc}
-%         \end{small}
-%     \end{center}
-%     \vskip -0.1in
-% \end{table*}
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) and recall of CAKES vs.
+other methods on the sift dataset.
+We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \label{table:results:ann-fashion}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{sc}
+                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                    \hline
+                    \textbf{Scaling Factor}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \hline
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall       \\
+                    \hline
+                    1.000 & 2191.849 & 0.950 & 212.602 & 1.000 & 2009.483 & 1.000 & 2167.365 & 1.000 \\
+                    \hline
+                    2.000 & 2118.513 & 0.927 & 106.372 & 1.000 & 939.169 & 1.000* & 1141.225 & 1.000 \\
+                    \hline
+                    4.000 & 2043.628 & 0.898 & 53.077 & 1.000 & 461.016 & 0.997 & 982.117 & 1.000 \\
+                    \hline
+                    8.000 & 1925.593 & 0.857 & 26.577 & 1.000 & 225.827 & 0.995 & 1179.813 & 1.000 \\
+                    \hline
+                    16.000 & 1841.863 & 0.862 & 13.292 & 1.000 & 116.960 & 0.991 & 1201.409 & 1.000 \\
+                    \hline
+                    32.000 & 1850.943 & 0.775 & 6.647 & 1.000 & 59.148 & 0.985 & 1157.667 & 1.000 \\
+                    \hline
+                    64.000 & 1787.142 & 0.677 & 3.324 & 1.000 & 26.068 & 0.968 & 1102.849 & 1.000 \\
+                    \hline
+                    128.000 & 1659.660 & 0.538 & 1.662 & 1.000 & 13.265 & 0.964 & 1039.607 & 1.000 \\
+                    \hline
+                    256.000 & 1597.264 & 0.592 & 0.831 & 1.000 & 6.650 & 0.962 & 1058.972 & 1.000 \\
+                    \hline
+                    512.000 & 1826.362 & 0.581 & 0.416 & 1.000 & 3.563 & 0.949 & 1035.260 & 1.000 \\
+                    \hline
+                \end{tabular}
+            \end{sc}
+        \end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
 
-% \begin{table*}[!t]
-%     % \renewcommand{\arraystretch}{1.15}
-%     \caption{Runtime performance (queries per second) of CAKES vs.
-other methods, $k=100$.
-Other methods did not report results for glove-25.
-Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995.}
-%     \label{table:results:ann-100}
-%     \vskip 0.15in
-%     \begin{center}
-%         \begin{small}
-%             \begin{sc}
-%                 \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-%                     \hline
-%                     \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-%                     &                    Recall & QPS  & Recall & QPS     & Recall & QPS      & Recall & QPS \\
-%                     \hline
-%                     fashion-mnist  & 1.000 & 111.47    & 1.000* & 16.02   & 1.000* & 277.80   & 1.000 & 71,370 \\ 
-%                     \hline
-%                     gist           & 1.000 & 2.70      & 1.000* & 0.80    & 0.995 & 59.71     & 1.000 & 29,090 \\
-%                     \hline
-%                     glove-25       & -- & --           & --     & --      & -- & --           & 0.851 & 49,350 \\
-%                     \hline
-%                     glove-100      & 1.000* &  19.04   & 1.000* & 7.52    & 0.987 & 92.58     & 0.905 & 4,384 \\
-%                     \hline
-%                     sift           & 1.000 &  23.50    & 1.000  & 6.51    & 1.000* & 179.14   & 1.000 & 147,400 \\                                                 
-%                     \hline
-%                 \end{tabular}
-%             \end{sc}
-%         \end{small}
-%     \end{center}
-%     \vskip -0.1in
-% \end{table*}
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) and recall of CAKES vs.
+other methods on the fashion-mnist dataset.
+We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \label{table:results:ann-fashion}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{sc}
+                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                    \hline
+                    \textbf{Scaling Factor}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \hline
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall       \\
+                    \hline
+                    1.000 & 2191.849 & 0.950 & 212.602 & 1.000 & 2009.483 & 1.000 & 2167.365 & 1.000 \\
+                    \hline
+                    2.000 & 2118.513 & 0.927 & 106.372 & 1.000 & 939.169 & 1.000* & 1141.225 & 1.000 \\
+                    \hline
+                    4.000 & 2043.628 & 0.898 & 53.077 & 1.000 & 461.016 & 0.997 & 982.117 & 1.000 \\
+                    \hline
+                    8.000 & 1925.593 & 0.857 & 26.577 & 1.000 & 225.827 & 0.995 & 1179.813 & 1.000 \\
+                    \hline
+                    16.000 & 1841.863 & 0.862 & 13.292 & 1.000 & 116.960 & 0.991 & 1201.409 & 1.000 \\
+                    \hline
+                    32.000 & 1850.943 & 0.775 & 6.647 & 1.000 & 59.148 & 0.985 & 1157.667 & 1.000 \\
+                    \hline
+                    64.000 & 1787.142 & 0.677 & 3.324 & 1.000 & 26.068 & 0.968 & 1102.849 & 1.000 \\
+                    \hline
+                    128.000 & 1659.660 & 0.538 & 1.662 & 1.000 & 13.265 & 0.964 & 1039.607 & 1.000 \\
+                    \hline
+                    256.000 & 1597.264 & 0.592 & 0.831 & 1.000 & 6.650 & 0.962 & 1058.972 & 1.000 \\
+                    \hline
+                    512.000 & 1826.362 & 0.581 & 0.416 & 1.000 & 3.563 & 0.949 & 1035.260 & 1.000 \\
+                    \hline
+                \end{tabular}
+            \end{sc}
+        \end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
 

### 2023-12-12 sections/4-results.tex
added tables and fixed csv to latex error

@@
     \caption{Runtime performance (queries per second) and recall of CAKES vs.
-other methods on the sift dataset.
+other methods on the fashion-mnist dataset.
 We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
                     \hline
-                    \textbf{Scaling Factor}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Scale}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall       \\
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall     & QPS & Recall    \\
                     \hline
-                    1.000 & 2191.849 & 0.950 & 212.602 & 1.000 & 2009.483 & 1.000 & 2167.365 & 1.000 \\
+                    1 & 13330.275 & 0.954 & 2191.849 & 0.950 & 212.602 & 1.000 & 2009.483 & 1.000 & 2167.365 & 1.000 \\
                     \hline
-                    2.000 & 2118.513 & 0.927 & 106.372 & 1.000 & 939.169 & 1.000* & 1141.225 & 1.000 \\
+                    2 & 13822.528 & 0.803 & 2118.513 & 0.927 & 106.372 & 1.000 & 939.169 & 1.000 & 1141.225 & 1.000 \\
                     \hline
-                    4.000 & 2043.628 & 0.898 & 53.077 & 1.000 & 461.016 & 0.997 & 982.117 & 1.000 \\
+                    4 & 16648.524 & 0.681 & 2043.628 & 0.898 & 53.077 & 1.000 & 461.016 & 0.997 & 982.117 & 1.000 \\
                     \hline
-                    8.000 & 1925.593 & 0.857 & 26.577 & 1.000 & 225.827 & 0.995 & 1179.813 & 1.000 \\
+                    8 & 16785.981 & 0.525 & 1925.593 & 0.857 & 26.577 & 1.000 & 225.827 & 0.995 & 1179.813 & 1.000 \\
                     \hline
-                    16.000 & 1841.863 & 0.862 & 13.292 & 1.000 & 116.960 & 0.991 & 1201.409 & 1.000 \\
+                    16 & 18717.175 & 0.494 & 1841.863 & 0.862 & 13.292 & 1.000 & 116.960 & 0.991 & 1201.409 & 1.000 \\
                     \hline
-                    32.000 & 1850.943 & 0.775 & 6.647 & 1.000 & 59.148 & 0.985 & 1157.667 & 1.000 \\
+                    32 & 15597.778 & 0.542 & 1850.943 & 0.775 & 6.647 & 1.000 & 59.148 & 0.985 & 1157.667 & 1.000 \\
                     \hline
-                    64.000 & 1787.142 & 0.677 & 3.324 & 1.000 & 26.068 & 0.968 & 1102.849 & 1.000 \\
+                    64 & 14997.277 & 0.378 & 1787.142 & 0.677 & 3.324 & 1.000 & 26.068 & 0.968 & 1102.849 & 1.000 \\
                     \hline
-                    128.000 & 1659.660 & 0.538 & 1.662 & 1.000 & 13.265 & 0.964 & 1039.607 & 1.000 \\
+                    128 & 14866.019 & 0.357 & 1659.660 & 0.538 & 1.662 & 1.000 & 13.265 & 0.964 & 1039.607 & 1.000 \\
                     \hline
-                    256.000 & 1597.264 & 0.592 & 0.831 & 1.000 & 6.650 & 0.962 & 1058.972 & 1.000 \\
+                    256 & -- & -- & 1597.264 & 0.592 & 0.831 & 1.000 & 6.650 & 0.962 & 1058.972 & 1.000 \\
                     \hline
-                    512.000 & 1826.362 & 0.581 & 0.416 & 1.000 & 3.563 & 0.949 & 1035.260 & 1.000 \\
+                    512 & -- & -- & 1826.362 & 0.581 & 0.416 & 1.000 & 3.563 & 0.949 & 1035.260 & 1.000 \\
                     \hline
@@
     \caption{Runtime performance (queries per second) and recall of CAKES vs.
-other methods on the fashion-mnist dataset.
+other methods on the glove-25 dataset.
 We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
                     \hline
-                    \textbf{Scaling Factor}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Scale}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall       \\
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall     & QPS & Recall    \\
                     \hline
-                    1.000 & 2191.849 & 0.950 & 212.602 & 1.000 & 2009.483 & 1.000 & 2167.365 & 1.000 \\
+                    1 & 22755.175 & 0.801 & 2829.338 & 0.835 & 278.176 & 1.000 & 2378.006 & 1.000 & 722.326 & 1.000 \\
                     \hline
-                    2.000 & 2118.513 & 0.927 & 106.372 & 1.000 & 939.169 & 1.000* & 1141.225 & 1.000 \\
+                    2 & 23792.690 & 0.607 & 2701.804 & 0.832 & 139.414 & 1.000 & 1188.398 & 1.000 & 574.516 & 1.000 \\
                     \hline
-                    4.000 & 2043.628 & 0.898 & 53.077 & 1.000 & 461.016 & 0.997 & 982.117 & 1.000 \\
+                    4 & 24990.630 & 0.443 & 2609.474 & 0.839 & 69.813 & 1.000 & 619.241 & 1.000 & 625.208 & 1.000 \\
                     \hline
-                    8.000 & 1925.593 & 0.857 & 26.577 & 1.000 & 225.827 & 0.995 & 1179.813 & 1.000 \\
+                    8 & 27843.055 & 0.294 & 2514.357 & 0.834 & 34.878 & 1.000 & 302.683 & 1.000 & 592.699 & 1.000 \\
                     \hline
-                    16.000 & 1841.863 & 0.862 & 13.292 & 1.000 & 116.960 & 0.991 & 1201.409 & 1.000 \\
+                    16 & 31114.062 & 0.213 & 2225.654 & 0.885 & 17.437 & 1.000 & 150.971 & 1.000 & 549.073 & 1.000 \\
                     \hline
-                    32.000 & 1850.943 & 0.775 & 6.647 & 1.000 & 59.148 & 0.985 & 1157.667 & 1.000 \\
+                    32 & 32423.543 & 0.178 & 2008.537 & 0.764 & 8.701 & 1.000 & 74.016 & 0.999 & 475.266 & 1.000 \\
                     \hline
-                    64.000 & 1787.142 & 0.677 & 3.324 & 1.000 & 26.068 & 0.968 & 1102.849 & 1.000 \\
+                    64 & -- & -- & 1990.317 & 0.631 & 4.362 & 1.000 & 37.735 & 0.997 & 461.066 & 1.000 \\
                     \hline
-                    128.000 & 1659.660 & 0.538 & 1.662 & 1.000 & 13.265 & 0.964 & 1039.607 & 1.000 \\
+                    128 & -- & -- & -- & -- & 2.176 & 1.000 & 18.960 & 0.998 & 440.758 & 1.000 \\
                     \hline
-                    256.000 & 1597.264 & 0.592 & 0.831 & 1.000 & 6.650 & 0.962 & 1058.972 & 1.000 \\
+                    256 & -- & -- & -- & -- & 1.067 & 1.000 & 9.472 & 0.998 & 422.591 & 1.000 \\
                     \hline
-                    512.000 & 1826.362 & 0.581 & 0.416 & 1.000 & 3.563 & 0.949 & 1035.260 & 1.000 \\
+                    % 512 & -- & -- & -- & -- & 0.544 & 1.000 & 4.499 & 0.996 & 0.000 & 0.000 \\
+                    % \hline
+                \end{tabular}
+            \end{sc}
+        \end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
+
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) and recall of CAKES vs.
+other methods on the sift dataset.
+We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \label{table:results:ann-fashion}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{sc}
+                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                    \hline
+                    \textbf{Scale}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \hline
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall    & QPS & Recall    \\
+                    \hline
+                    1 & 19330.282 & 0.782 & 3977.432 & 0.686 & 72.610 & 1.000 & 697.809 & 1.000 & 551.894 & 1.000 \\
+                    \hline
+                    2 & 20319.372 & 0.552 & 3800.401 & 0.614 & 36.237 & 1.000 & 329.950 & 1.000 & 265.615 & 1.000 \\
+                    \hline
+                    4 & 21765.007 & 0.394 & 3686.606 & 0.637 & 18.152 & 1.000 & 165.175 & 1.000 & 142.545 & 1.000 \\
+                    \hline
+                    8 & 24796.528 & 0.298 & 3582.099 & 0.710 & 9.077 & 1.000 & 77.187 & 1.000 & 79.353 & 1.000 \\
+                    \hline
+                    16 & 26760.717 & 0.210 & 3498.893 & 0.690 & 4.539 & 1.000 & 39.809 & 1.000 & 81.172 & 1.000 \\
+                    \hline
+                    32 & 27522.668 & 0.193 & 3435.058 & 0.639 & 2.269 & 1.000 & 20.910 & 0.999 & 78.124 & 1.000 \\
+                    \hline
+                    64 & -- & -- & 3391.916 & 0.678 & 1.135 & 1.000 & 8.867 & 0.997 & 74.344 & 1.000 \\
+                    \hline
+                    128 & -- & -- & 3363.965 & 0.643 & 0.566 & 1.000 & 4.786 & 0.993 & 68.005 & 1.000 \\
+                    \hline
+                \end{tabular}
+            \end{sc}
+        \end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
+
+\begin{table*}[!t]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Runtime performance (queries per second) and recall of CAKES vs.
+other methods on a random dataset.
+We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \label{table:results:ann-fashion}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{sc}
+                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                    \hline
+                    \textbf{Scale}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \hline
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall    & QPS & Recall    \\
+                    \hline
+                    1 & 11713.738 & 0.060 & 4283.182 & 0.028 & 71.575 & 1.000 & 733.821 & 1.000 & 553.948 & 1.000 \\
+                    \hline
+                    2 & 10061.864 & 0.048 & 4043.999 & 0.021 & 36.157 & 1.000 & 358.083 & 1.000 & 269.314 & 1.000 \\
+                    \hline
+                    4 & 9120.868 & 0.031 & 3639.337 & 0.014 & 18.072 & 1.000 & 189.802 & 1.000 & 137.236 & 1.000 \\
+                    \hline
+                    8 & 8352.077 & 0.022 & 3371.923 & 0.013 & 9.036 & 1.000 & 88.375 & 1.000 & 56.871 & 1.000 \\
+                    \hline
+                    16 & 8250.619 & 0.008 & 3174.156 & 0.006 & 4.531 & 1.000 & 43.613 & 1.000 & 26.136 & 1.000 \\
+                    \hline
+                    32 & -- & -- & 3013.970 & 0.007 & 2.270 & 1.000 & 17.201 & 1.000 & 13.534 & 1.000 \\
+                    % \hline
+                    % 64 & -- & -- & -- & -- & 1.135 & 1.000 & 9.939 & 1.000 & 0.000 & 0.000 \\
+                    % \hline
+                    % 128 & --& -- & -- & -- & 0.567 & 1.000 & 5.106 & 1.000 & 0.000 & 0.000 \\
                     \hline

### 2023-12-13 supplement.tex
adding plots and more edits to tables

@@
 
-\section{Datasets}
-
 \section{Results}
 
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on fashion-mnist with $k=100$. 
+    }
+    \label{fig:supplement:fashion-mnist-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/sift-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on sift with $k=100$. 
+    }
+    \label{fig:supplement:sift-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/glove-25-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on glove-25 with $k=100$. 
+    }
+    \label{fig:supplement:glove-25-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/random-1000000-128-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on random with $k=100$. 
+    }
+    \label{fig:supplement:random-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/fashion-mnist-indexing.png}
+    \caption{
+        Indexing time for algorithms on fashion-mnist dataset. 
+    }
+    \label{fig:supplement:fashion-mnist-indexing}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/glove-25-indexing.png}
+    \caption{
+        Indexing time for algorithms on glove-25 dataset. 
+    }
+    \label{fig:supplement:glove-25-indexing}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/sift-indexing.png}
+    \caption{
+        Indexing time for algorithms on sift dataset. 
+    }
+    \label{fig:supplement:sift-indexing}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/random-indexing.png}
+    \caption{
+        Indexing time for algorithms on random dataset. 
+    }
+    \label{fig:supplement:random-indexing}
+\end{figure}
+
 \end{document}

### 2023-12-13 sections/1-introduction.tex
some edits

@@
 This is especially true in the field of astronomy. 
+% Najib: We should mention datasets in this order: neural -network embeddings, biological, radio-ml, and then astronomical.
 The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
@@
 
+% Najib: Switch the next paragraph for this:
+% In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}.
+% This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
+% Thus, analysis of these large datasets requires better algorithms.
 In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating that computer architecture will not ``catch up'' to our computational needs in the near future.
@@
 Similarity search enables a variety of applications, including recommendation and classification systems. 
+% Najib: The previous sentence needs a citation.
 Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, similarity search can become a useful tool for understanding and verifying the validity of model embeddings.
+% Najib: As the cardinalities and dimensionalities ...
 As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
@@
 
+% Najib: Rework this paragraph. I don' want any mathematically notation in the introduction, particularly because this is repeated in Methods, The Search Problem.
 We now define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN). 
@@
 Previous works have used the term \emph{approximate} search to refer to $\rho$-nearest neighbor search, but in this paper, 
+% Najib: we reserve the term \emph{approximate} for search algorithms which do not exhibit perfect recall.
 we reserve the term \emph{approximate} for $k$-NN search algorithms which may miss some of the $k$ nearest elements, and erroneously include some of the elements other than the truly nearest $k$.
+% Najib: Same here with the math notation.
 In other words, we call a $k$-NN search algorithm \emph{approximate} if there exists a query, dataset, distance function, or value of $k$ for which its set $H$ of returned points satisfies the condition $H \triangle S_q \neq \emptyset$, where $\triangle$ denotes a symmetric difference.
+% Najib: When we refer to an \emph{exact} $k$-NN algorithm, we mean an algorithm which exhibits perfect recall when compared to a na\"{i}ve linear search.
 When we refer to an \emph{exact} $k$-NN algorithm, we describe an algorithm which, no matter the query, dataset, distance function, nor value of $k$, returns precisely the $k$ nearest neighbors to that query;
@@
 
+% Najib: Remove/replace the word ``pervasive'' here. It has a negative connotation.
 $k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
 Na\"{i}ve implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
+% Najib: because the dataset's cardinality is growing exponentially.
+% Najib: While algorithms for fast $k$-NN on large datasets DO exist, most of them are approximate and do not exploit the geometric and topological structure inherent in SUCH datasets.
 While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not exploit the geometric and topological structure inherent in large datasets.
+% Najib: Further, such algorithms are often approximate and while approximate search MAY be sufficient for some applications, the need for efficient AND \emph{exact} search remains.
 Approximate search is sufficient for some applications, but the need for efficient, \emph{exact} search remains.
-In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$, and small values of $k$ may be necessary when classes are not well-separated.
-There is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
+For example, for a majority voting classifier, approximate $k$-NN may agree with exact $k$-NN for large values of $k$, but may be sensitive to local perturbations with smaller values of $k$.
+This is especially true when classes are not well-separated. % TODO: Cite something which says this.
+Further, there is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
 this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could exhibit suboptimal accuracy.
@@
 We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
-Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
+% Najib: Cite HNSW here. Also, add ANNOY.
 Finally, we provide a comparison of CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.

### 2023-12-13 sections/2-methods.tex
some edits

@@
 In this manuscript, we are primarily concerned with $k$-nearest neighbors search in a finite-dimensional space.
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation (e.g., the genome of 
-an organism, the neural embedding of a single image, or any entity on which we can define a \emph{distance function}).
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the genome of an organism, the neural-network embedding of an image, etc.
 
@@
 In addition to these constraints, if the distance function obeys the triangle inequality, i.e., it is a \emph{distance metric}, then we guarantee that CAKES has perfect recall. 
-Choice of distance function varies by type of data.
-For example, with electromagnetic spectra, one could both Euclidean (L2) and Cosine distances.
-With biological or string data, Hamming and Levenshtein distances are more appropriate.
+The choice of distance function varies by dataset.
+For example, with neural-network embeddings, one could use Euclidean (L2) distance or cosine distance.
+With biological or string data, Levenshtein or Hamming distances are more appropriate.
+With time-series data, one could use dynamic time warping (DTW) distance.
 
@@
 
+% Najib: Add our own plots of local fractal dimension instead of referencing figures from the CHESS paper. Then, rewrite this paragraph.
 Figures 2 and 3 in~\cite{ishaq2019clustered} illustrate the low fractal dimensions of two datasets used with $\rho$-NN search in CHESS. 
@@
 
-We define a \emph{cluster} as a set of points with a \emph{center}, a \emph{radius}, and an approximation of the \emph{local fractal dimension}.
+Najib We define a \emph{cluster} as a set of points with a \emph{center} and a \emph{radius}.
 The \emph{center} is the geometric median of a sample of points in the \emph{cluster}, and so it is a real data point.
-The \emph{radius} is the maximum distance from any point in the cluster to the \emph{center}.
+The \emph{radius} is the maximum distance of any point in the cluster to the \emph{center}.
 We estimate \emph{local fractal dimension} at the scale of the cluster radius and half the cluster radius.
-Each cluster (unless it is a leaf cluster) has two child clusters in much the same way that a node in a binary tree has two child nodes.
+Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
+% Najib: Add an equation for local fractal dimension with the simplification from the previous sentence.
+
+% Najib: Rewrite this paragraph. Old paper defined it as ... but we define it for the hierarchical clustering scheme as ...
 We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(X)$ of a data set $X$ under a hierarchical clustering scheme as a refinement of~\cite{yu2015entropy}, where metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data.
@@
 
-We start by building a divisive hierarchical clustering of the dataset with CLAM, using a similar recursive procedure as outlined in CHESS, but with the following improvements:
+We start by building a divisive hierarchical clustering of the dataset using CLAM.
+The procedure is similar to what was outlined in CHESS, but with the following improvements:
 better selection of poles for partitioning and depth-first dataset reordering (see Section ~\ref{subsubsec:methods:clustering:dataset-depth-first-reordering}).
@@
 \label{subsec:methods:sharding-and-auto-tuning}
+% Najib: Remove Sharding from this subsection. We do not show sharding in any of our benchmarks. All benchmarks are done with datasets that fit in memory (on giant machines, but still).
 
@@
 
+% Najib: For this figure, I want the red, green and blue lines/braces to be thinner. Perhaps 1/2 to 2/3 as thick.
+% Najib: Also, add a caption.
 \begin{figure}[ht!]
@@
 when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which could contain points in the query ball. 
+% Najib: This is laid out in Algorithm 2. We should reference that algorithm here.
 
+% Najib: Merge Figures 2 and 3. There's too much redundancy between them.
 \begin{figure}[ht!]
@@
 
+% Najib: Visually, ``m'' is not centered on the line segment. It's a little bit to the right.
 \begin{figure}[ht!]
@@
 
+% Najib: I don't like the double vertical bars for norm-notion. I think it's too messy. We should just overload the notation of the single horizontal bar across the top.
+% Najib: Remove the equation numbers from this paragraph. They're not necessary.
 We can do a quick check using values computed during partitioning (see Section~\ref{subsubsec:methods:cluster-partitioning}) to determine whether $d \leq \rho$. 
@@
 
+% Najib: Move the cluster.isLeaf check to the beginning of the algorithm because it's the base case.
 \begin{algorithm} 
@@
 
-where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy at that radius, $B_X(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the fractal dimension around the query in a region of radius $r$.
+where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy at that radius, $B_X(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the local fractal dimension around the query in a region of radius $r$.
 

### 2023-12-13 sections/4-results.tex
some edits

@@
 
-Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, deep-image under cosine distance, and glove-25 under cosine distance.
-Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset which does not obey any manifold structure. 
-The the horizontal axes in each figure denote the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
-The vertical axes denote throughput in queries per second. Both axes are on a logarithmic scale. In this section, we discuss results only for $k = 10$. 
-To see results with other values of $k$, refer to our Supplement section.
+Figures~\ref{fig:plots:fashion-mnist-scaling}A, ~\ref{fig:plots:glove-25-scaling}B, and ~\ref{fig:plots:sift-scaling}C show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
+Figure~\ref{fig:plots:random-scaling}D shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
+The vertical axis denotes throughput in queries per second.
+Both axes are on a logarithmic scale.
+In this section, we discuss results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section.
+In these plots, we label the recall at each scale for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall at all scales.
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, each of CAKES's four algorithms exhibits significantly higher throughput than na\"{i}ve linear search, plotted in orange.
-Amongst them, GreedySieve, plotted in blue, consistently performs best, exhibiting sublinear time performance for nearly all cardinality multipliers. 
-We see that for multipliers around $10^3$, Greedy Sieve exhibits throughput which is nearly three orders of magnitude higher than that of linear search.
+Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively.
+When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider
+only three decimal places.
 
-With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of our four algorithms outperforms linear search at high cardinality multipliers. 
-On this dataset, Repeated $\rho$-NN, plotted in green, is consistently the best performing algorithm. {\color{red} Revive lfd violin plots to add explanation for this.}
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
+We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
+Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
+Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
+
+With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high cardinality multipliers.
+Similar to the results for fashion-mnist, we observe that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$.
+We also see that while hnsw and annoy have higher throughput than CAKES for all multipliers, CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-glove-25}.
+For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.294 and 0.857 respectively, while CAKES exhibits perfect recall.
+On this dataset, Repeated $\rho$-NN, plotted in green, is consistently the best-performing algorithm. {\color{red} Revive lfd violin plots to add explanation for this.}
 
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/fashion-mnist_10_scaling.png}
+    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-10.png}
     \caption{
-        Fashion-mnist Scaling
+        Algorithm performance by cardinality multiplier on fashion-mnist for $k=10$.
     }
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/glove-25_10_scaling.png}
+    \includegraphics[width=3.4in]{plots/glove-25-knn-10.png}
     \caption{
-        Glove-25 Scaling
-    }
+        Algorithm performance by cardinality multiplier on glove-25 for $k=10$.  }
     \label{fig:results:glove-25-scaling}
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/sift_10_scaling.png}
+    \includegraphics[width=3.4in]{plots/sift-knn-10.png}
     \caption{
-        Sift Scaling
-    }
+        Algorithm performance by cardinality multiplier on sift for $k=10$.    }
     \label{fig:results:sift-scaling}
@@
     \centering
-    \includegraphics[width=3.4in]{images/result_plots/random_placeholder_10_scaling.png}
+    \includegraphics[width=3.4in]{plots/random-1000000-128-knn-10.png}
     \caption{
-        Random Scaling
+        Algorithm performance by cardinality multiplier on a random dataset for $k=10$.
     }

### 2023-12-13 sections/4-results.tex
quad-figure in place

@@
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-10.png}
-    \caption{
-        Algorithm performance by cardinality multiplier on fashion-mnist for $k=10$.
+\begin{figure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}\\
+\subcaption{Fashion-mnist for $k=10$.}
+\label{fig:results:fashion-mnist-scaling}
+\end{subfigure}%
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}\\
+\subcaption{Glove-25 for $k=10$.  }
+\label{fig:results:glove-25-scaling}
+\end{subfigure}%
+\\
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}\\
+\subcaption{Sift for $k=10$.    }
+\label{fig:results:sift-scaling}
+\end{subfigure}%
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{plots/random-1000000-128-knn-10.png}\\
+\subcaption{A random dataset for $k=10$.
 }
-    \label{fig:results:fashion-mnist-scaling}
+\label{fig:results:random-scaling}
+\end{subfigure}%
+
+\caption{Algorithm performance across four datasets.
+In each plot, the horizontal axis represents increasing cardinality (size) of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).}
 \end{figure}
+
+% 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/fashion-mnist-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on fashion-mnist for $k=10$.
+%     }
+%     \label{fig:results:fashion-mnist-scaling}
+% \end{figure}
 
@@
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/glove-25-knn-10.png}
-    \caption{
-        Algorithm performance by cardinality multiplier on glove-25 for $k=10$.  }
-    \label{fig:results:glove-25-scaling}
-\end{figure}
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/glove-25-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on glove-25 for $k=10$.  }
+%     \label{fig:results:glove-25-scaling}
+% \end{figure}
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/sift-knn-10.png}
-    \caption{
-        Algorithm performance by cardinality multiplier on sift for $k=10$.    }
-    \label{fig:results:sift-scaling}
-\end{figure}
+% 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/sift-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on sift for $k=10$.    }
+%     \label{fig:results:sift-scaling}
+% \end{figure}
 
@@
 % random datasets
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/random-1000000-128-knn-10.png}
-    \caption{
-        Algorithm performance by cardinality multiplier on a random dataset for $k=10$.
-    }
-    \label{fig:results:random-scaling}
-\end{figure}
+% 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/random-1000000-128-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on a random dataset for $k=10$.
+%     }
+%     \label{fig:results:random-scaling}
+% \end{figure}
 

### 2023-12-13 sections/4-results.tex
added TODO

@@
 \caption{Algorithm performance across four datasets.
-In each plot, the horizontal axis represents increasing cardinality (size) of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).}
+In each plot, the horizontal axis represents increasing cardinality (size) of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+TODO say something about plots that didn't finish.}
 \end{figure}

### 2023-12-14 sections/4-results.tex
trying to deal with merge conflict

@@
 
+Figures~\ref{fig:plots:fashion-mnist-scaling}A, ~\ref{fig:plots:glove-25-scaling}B, and ~\ref{fig:plots:sift-scaling}C show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
+Figure~\ref{fig:plots:random-scaling}D shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
+The vertical axis denotes throughput in queries per second.
+Both axes are on a logarithmic scale.
+In this section, we discuss results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section.
+In these plots, we label the recall at each scale for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall at all scales.
+
+Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively.
+When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider
+only three decimal places.
 Figures~\ref{fig:plots:fashion-mnist-scaling}A, ~\ref{fig:plots:glove-25-scaling}B, and ~\ref{fig:plots:sift-scaling}C show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
@@
 Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers. 
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
+We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
+Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
+Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
 
 With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high cardinality multipliers.
-Similar to the results for fashion-mnist, we observe that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$.
+Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$.
 We also see that while hnsw and annoy have higher throughput than CAKES for all multipliers, CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-glove-25}.

### 2023-12-15 sections/1-introduction.tex
edits to introduction

--- 2023-11-21 sections/1-introduction.tex
@@
 Often dubbed ``the Big Data explosion,'' rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
-This is especially true in the field of astronomy. 
-The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
-Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.
+% Najib: We should mention datasets in this order: neural -network embeddings, biological, radio-ml, and then astronomical. DONE. (flow is a bit awakward though)
 
+With the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, neural network embeddings are becoming an important source of big data.
 Biology has also seen an influx of large, high-dimensional datasets.
@@
 Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with aligned length of 50,000 letters.
+
+Radio frequency data offers another example of massive datasets. 
+For example, the RadioML~\cite{oshea2018radioml} dataset contains 2,555,504 samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
+
+Additionally, many large, high-dimensional datasets come from the field of astronomy.
+The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
 
@@
 
-In many fields, the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}, indicating that computer architecture will not ``catch up'' to our computational needs in the near future.
-As a result, analysis of these large datasets requires better algorithms.
+In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}.
+This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
+Thus, analysis of these large datasets requires better algorithms.
 
 Many researchers are especially interested in similarity search on these datasets. 
-Similarity search enables a variety of applications, including recommendation and classification systems. 
-Additionally, with the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, similarity search can become a useful tool for understanding and verifying the validity of model embeddings.
-As the sizes and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
+Similarity search enables a variety of applications, including recommendation~\cite{annoy} and classification systems~\cite{suyanto2022knnclassifier}.
+
+As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
 even state-of-the-art algorithms exhibit a steep recall versus throughput tradeoff~\cite{ishaq2019clustered}.
@@
 We now define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN). 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \ \setminus \ S_q$, $f(x, q) > \max\{f(y, q): y \in S_q \}$;
-that is, for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
-On the other hand, $\rho$-nearest neighbors search aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
-that is, find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search 
+aims to find the $k$ closest points to $q$ in $ \textbf{X}$.
+On the other hand, $\rho$-nearest neighbors search aims to find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
 
 Previous works have used the term \emph{approximate} search to refer to $\rho$-nearest neighbor search, but in this paper, 
-we reserve the term \emph{approximate} for $k$-NN search algorithms which may miss some of the $k$ nearest elements, and erroneously include some of the elements other than the truly nearest $k$.
-In other words, we call a $k$-NN search algorithm \emph{approximate} if there exists a query, dataset, distance function, or value of $k$ for which its set $H$ of returned points satisfies the condition $H \triangle S_q \neq \emptyset$, where $\triangle$ denotes a symmetric difference.
-When we refer to an \emph{exact} $k$-NN algorithm, we describe an algorithm which, no matter the query, dataset, distance function, nor value of $k$, returns precisely the $k$ nearest neighbors to that query;
-no closer points are elided and no further points are included.
+we reserve the term \emph{approximate} for search algorithms which do not exhibit perfect recall.
+When we refer to an \emph{exact} search algorithm, we mean an algorithm which exhibits perfect recall when compared to a na\"{i}ve linear search.
 
-$k$-nearest neighbors ($k$-NN) search is one of the most pervasive classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
-Na\"{i}ve implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
-While algorithms for fast $k$-NN on large datasets exist, most of them are approximate and do not exploit the geometric and topological structure inherent in large datasets.
-Approximate search is sufficient for some applications, but the need for efficient, \emph{exact} search remains.
-In particular, while approximate $k$-NN may agree with exact $k$-NN in a majority vote for large values of $k$, it may be sensitive to local perturbations with smaller values of $k$, and small values of $k$ may be necessary when classes are not well-separated.
-There is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
+$k$-nearest neighbors ($k$-NN) search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
+Na\"{i}ve implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets because their cardinalities are growing exponentially.
+
+While algorithms for fast $k$-NN on large datasets do exist, most do not exploit the geometric and topological structure inherent in large datasets.
+Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \emph{exact} search remains.
+For example, for a majority voting classifier, approximate $k$-NN may agree with exact $k$-NN for large values of $k$, but may be sensitive to local perturbations with smaller values of $k$.
+This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
+Further, there is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
 this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could exhibit suboptimal accuracy.
@@
 We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
-Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
-Finally, we provide a comparison of CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
+Finally, we provide a comparison of CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy} on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.

### 2023-12-15 sections/3-datasets.tex
headers for new datasets and naming fixes
edits to introduction

@@
 \label{sec:datasets-and-benchmarks}
+
+\subsection{ANN Benchmark Datasets}
+\label{subsec:ann-benchmarks}
 
@@
 \end{table}
+
+\subsection{Silva}
+\label{subsec:silva}
+
+Add section about Silva.
+
+\subsection{Radio ML}
+The RadioML dataset contains samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
+Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each modulation mode and SNR level~\cite{oshea2018radioml}.
+Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
+Each sample provided in RadioML can be thought of as an array of 2048 floating point values for a single signal capture with labels for SNR and modulation mode.
+
+\label{subsec:radioml}

### 2023-12-15 sections/2-methods.tex
changes to methods

@@
     \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) }
-    \label{eq:methods:lfd}
+    \label{eq:methods:lfd_original}
 \end{equation}
@@
 The \emph{radius} is the maximum distance of any point in the cluster to the \emph{center}.
-We estimate \emph{local fractal dimension} at the scale of the cluster radius and half the cluster radius.
 Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
-% Najib: Add an equation for local fractal dimension with the simplification from the previous sentence.
+We estimate local fractal dimension at the scale of the cluster radius and half the cluster radius. 
+This allows us to rewrite Equation \ref{eq:methods:lfd_original} as simply: 
+\begin{equation} 
+\log_2 \frac{|B_X(q, r)}{|B_X(q, r/2)|}
+\end{equation}
 
-% Najib: Rewrite this paragraph. Old paper defined it as ... but we define it for the hierarchical clustering scheme as ...
-We define the \emph{metric entropy} $\mathcal{N}_{\hat{r_c}}(X)$ of a data set $X$ under a hierarchical clustering scheme as a refinement of~\cite{yu2015entropy}, where metric entropy for a given cluster radius $r_c$ was the number of clusters of radius $r_c$ needed to cover all data.
-Here, we use a hierarchical, divisive clustering approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g.
+% Najib: Rewrite this paragraph. Old paper defined it as ... but we define it for the hierarchical clustering scheme as ... DONE. (let me know if this okay)
+
+In~\cite{yu2015entropy}, metric entropy for a given cluster radius $r_c$ was defined as the number of clusters of radius $r_c$ needed to cover all data.
+In this paper, we define the metric entropy $\mathcal{N}_{\hat{r_c}}(X)$ of a data set $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree.
+
+% Here, we use a hierarchical, divisive clustering approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g.
 a specified cardinality or radius) are not further split.
-Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of leaf clusters.
+% Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of leaf clusters.
 
@@
 
-\subsection {Sharding and Auto-Tuning}
-\label{subsec:methods:sharding-and-auto-tuning}
-% Najib: Remove Sharding from this subsection. We do not show sharding in any of our benchmarks. All benchmarks are done with datasets that fit in memory (on giant machines, but still).
+\subsection {Auto-Tuning}
+\label{subsec:methods:auto-tuning}
 
-In this work, we also introduce a method of \emph{sharding} for datasets that cannot fit in memory.
-This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor is typically much faster than $k$-nearest neighbors search.
-In addition to sharding, we perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to be used with a given dataset.
+% In this work, we also introduce a method of \emph{sharding} for datasets that cannot fit in memory.
+% This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor is typically much faster than $k$-nearest neighbors search.
 
+We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
 With this approach, we begin by taking a random sample of no more than 10\% of the data.
-This sample is considered the first shard.
-The exact proportion of the dataset in this random sample depends on the dimensionality of the dataset and a user-specified memory limit. 
-We then split the rest of the dataset into more shards based on the user-defined memory limit.
-For each shard, we build a cluster tree using Algorithm~\ref{alg:methods:clustering:partition}.
-From the first shard, we consider all clusters at some user-specified depth in the tree, with a default depth of 10.
+We build a cluster tree with this sample using Algorithm~\ref{alg:methods:clustering:partition}, and select all clusters at some user-specified depth in the tree, with a default depth of 10.
 Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
 We select the algorithm which had the lowest total search time over all the queries. 
+Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
 
-Sharded search then proceeds as follows.
-We use previously determined $k$-NN algorithm on the first shard. 
-We use the distance to the $k$-th farthest neighbor to perform $\rho$-NN search on each subsequent shard. 
-After searching each shard, we keep only the $k$-nearest neighbors so far, potentially decreasing the distance to the $k$-th farthest neighbor for search on the next shard.
-
-Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
-Note also that the size of the shards is not necessarily consistent;
-it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search. 
-However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th neighbor), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
-Finding the optimal ratio between the size of the first shard and the sizes of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
-For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but, since the entire dataset fits in memory, there is only one shard.
+% Note also that the size of the shards is not necessarily consistent;
+% it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search. 
+% However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th neighbor), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
+% Finding the optimal ratio between the size of the first shard and the sizes of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
+% For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but, since the entire dataset fits in memory, there is only one shard.
 
@@
 % Najib: For this figure, I want the red, green and blue lines/braces to be thinner. Perhaps 1/2 to 2/3 as thick.
-% Najib: Also, add a caption.
 \begin{figure}[ht!]
@@
     \includegraphics[scale=0.5]{images/geometry/deltas.png}
-    \caption{}
+    \caption{$\delta$, $\delta_{\max}$, and $\delta_{\min}$ defined for cluster $C$ and query $q$.}
     \label{fig:methods:deltas}
@@
 when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which could contain points in the query ball.
-% Najib: This is laid out in Algorithm 2. We should reference that algorithm here.
+The full algorithm is described in Algorithm~\ref{alg:methods:rho-nn-search}.
 
 % Najib: Merge Figures 2 and 3. There's too much redundancy between them.
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[scale=0.5]{images/geometry/overlapping-children-1.pdf}
-    \caption{
-        Triangle formed by query $q$, and poles $\ell$ and $r$. Query ball of radius $\rho$ centered at $q$. 
-    }
-    \label{fig:methods:overlapping-children-1}
-\end{figure}
-
-To determine whether both children can contain points in the query ball, we consider the following geometry. 
-Let $\ell$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ denote the query, and $\rho$ denote the search radius, as shown in Figure~\ref{fig:methods:overlapping-children-1}.
-Without loss of generality, we assume that $\overline{qr \vphantom \ell} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[scale=0.5]{images/geometry/overlapping-children-1.pdf}
+%     \caption{
+%         Triangle formed by query $q$, and poles $\ell$ and $r$. Query ball of radius $\rho$ centered at $q$. 
+%     }
+%     \label{fig:methods:overlapping-children-1}
+% \end{figure}
 
@@
 
+To determine whether both children can contain points in the query ball, we consider the following geometry. 
+Let $\ell$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ denote the query, and $\rho$ denote the search radius, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
+Without loss of generality, we assume that $\overline{qr \vphantom \ell} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
+
 Now let $m$ be the midpoint of $\overline{\ell r}$.
@@
 
-% Najib: I don't like the double vertical bars for norm-notion. I think it's too messy. We should just overload the notation of the single horizontal bar across the top.
-% Najib: Remove the equation numbers from this paragraph. They're not necessary.
 We can do a quick check using values computed during partitioning (see Section~\ref{subsubsec:methods:cluster-partitioning}) to determine whether $d \leq \rho$. 
-We note that $d = ||\overline{mq'}|| = ||\overline{mr \vphantom {q'}}||-||\overline{q'r}||$.
-Since $m$ is the midpoint of $\overline{\ell r}$, we have that $||\overline{mr}|| = \tfrac{||\overline{\ell r}||}{2}$, so \begin{equation} d = \tfrac{||\overline{\ell r}||}{2} - ||\overline{q'r}||.
+In the caculations that follow, we overload notation a bit and use $\overline{xy}$ to mean the length of the line segment between $x$ and $y$.
+We note that $d = \overline{mq'} = \overline{mr \vphantom {q'}}-\overline{q'r}$.
+Since $m$ is the midpoint of $\overline{\ell r}$, we have that $\overline{mr} = \tfrac{\overline{\ell r}}{2}$, so \begin{equation*} d = \tfrac{\overline{\ell r}}{2} - \overline{q'r}.
 \label{eq:methods:d-expression}
-\end{equation}
+\end{equation*}
 
-It remains to find $||\overline{q'r}||$.
+It remains to find $\overline{q'r}$.
 To do this, we use the Law of Cosines with $\triangle \ell rq$.
@@
 By the Law of Cosines, we have that
-\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{qr\vphantom \ell}||^2 + ||\overline{\ell r}||^2 - ||\overline{q\ell}||^2}{2||\overline{\ell r}||||\overline{qr \vphantom \ell}||}.\end{equation}
+\begin{equation} \text{cos}(\theta) = \tfrac{\overline{qr\vphantom \ell}^2 + \overline{\ell r}^2 - \overline{q\ell}^2}{2(\overline{\ell r})(\overline{qr \vphantom \ell})}.\end{equation}
 On the other hand, since $\triangle q'qr$ is a right triangle,
 
-\begin{equation} \text{cos}(\theta) = \tfrac{||\overline{q'r}||}{||\overline{qr \vphantom \ell}||}.\end{equation}
+\begin{equation} \text{cos}(\theta) = \tfrac{\overline{q'r}}{\overline{qr \vphantom \ell}}.\end{equation}
 
-Equating the two expressions for $\text{cos}(\theta)$ and solving for $||\overline{q'r}||$, we have \begin{equation} ||\overline{q'r}|| = \tfrac{||\overline{qr\vphantom \ell}||^2 + ||\overline{\ell r}||^2 - ||\overline{q\ell}||^2}{2 ||\overline{\ell r}||}.
+Equating the two expressions for $\text{cos}(\theta)$ and solving for $\overline{q'r}$, we have \begin{equation} \overline{q'r} = \tfrac{\overline{qr\vphantom \ell}^2 + \overline{\ell r}^2 - \overline{q\ell}^2}{2 \overline{\ell r}}.
 \end{equation}
@@
 Substituting this back in equation~\ref{eq:methods:d-expression} gives 
-\begin{equation} d = \tfrac{||\overline{q\ell}||^2 - ||\overline{qr \vphantom \ell}||^2}{2||\overline{lr}||}.\end{equation}
+\begin{equation} d = \tfrac{\overline{q\ell}^2 - \overline{qr \vphantom \ell}^2}{2\overline{lr}}.\end{equation}
 
-Then to check if $d \leq \rho$, it suffices to verify that \begin{equation}(||\overline{q\ell}|| + ||\overline{qr \vphantom \ell}||)(||\overline{q\ell}||- ||\overline{qr \vphantom \ell}||) \leq 2||\overline{\ell r}||\rho.
+Then to check if $d \leq \rho$, it suffices to verify that \begin{equation}(\overline{q\ell} + \overline{qr \vphantom \ell})(\overline{q\ell}- \overline{qr \vphantom \ell}) \leq 2\overline{\ell r}\rho.
 \end{equation}
 
-% Najib: Move the cluster.isLeaf check to the beginning of the algorithm because it's the base case.
+% Najib: move is.Leaf check to the beginning of the algorithm because it's the base base case. DONE (confirm this is correct)
+
 \begin{algorithm} 
@@
         \STATE $results \Leftarrow \emptyset$
+        \IF{$cluster.isLeaf$}
+            \FOR{$p \in cluster.points$}
+                \IF{$p.\delta \leq r$}
+                    \STATE $results.push((p, r))$
+                \ENDIF
+            \ENDFOR
+            \STATE Return $results$
+        \ENDIF
         \IF{$cluster.left \cap B_X(query, r) \neq \emptyset$}
@@
         \ENDIF
-        \IF{$cluster.isLeaf$}
-            \FOR{$p \in cluster.points$}
-                \IF{$p.\delta \leq r$}
-                    \STATE $results.push((p, r))$
-                \ENDIF
-            \ENDFOR
-        \ENDIF
-        \STATE Return $results$
+
     \end{algorithmic}
+    \label{alg:methods:rho-nn-search}
 \end{algorithm}

### 2023-12-17 sections/4-results.tex
lfd figures in place, also 2 more scaling dummies

@@
 \label{sec:results}
+
+Figures~\ref{fig:results:fashion-mnist-lfd} show the local fractal dimension of.... (TODO fill this in)
+
+%% LFD plots
+
+\begin{figure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/fashion-mnist-60000-lfd.png}\\
+\subcaption{Fashion-mnist}
+\label{fig:results:fashion-mnist-lfd}
+\end{subfigure}%
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/glove-25-1183514-lfd.png}\\
+\subcaption{Glove-25}
+\label{fig:results:glove-25-lfd}
+\end{subfigure}
+\vspace{1em}
+\\
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/sift-1000000-lfd.png}\\
+\subcaption{Sift}
+\label{fig:results:sift-lfd}
+\end{subfigure}%
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-128-1000000-lfd.png}\\
+\subcaption{A random dataset}
+\label{fig:results:random-lfd}
+\end{subfigure}
+\vspace{1em}
+\\
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
+\subcaption{Silva 18S metagenomics}
+\label{fig:results:silva-lfd}
+\end{subfigure}%
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/radioml-97920-lfd.png}\\
+\subcaption{RadioML}
+\label{fig:results:radioml-lfd}
+\end{subfigure}%
+
+\caption{Local fractal dimension vs. cluster depth, grouped by decile of LFD, across six datasets.}
+\end{figure}
+%% end LFD plots
 
@@
 \begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}\\
+\includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}
 \subcaption{Fashion-mnist for $k=10$.}
@@
 \begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}\\
+\includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}
 \subcaption{Glove-25 for $k=10$.  }
@@
 \end{subfigure}%
+\vspace{1em}
 \\
 \begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}\\
+\includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}
 \subcaption{Sift for $k=10$.    }
@@
 \begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/random-1000000-128-knn-10.png}\\
+\includegraphics[width=0.95\textwidth]{plots/random-1000000-128-knn-10.png}
 \subcaption{A random dataset for $k=10$.
@@
 \end{subfigure}%
+\vspace{1em}
+\\
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{plots/silva-knn-10.png}
+\subcaption{Silva for $k=10$.    }
+\label{fig:results:silva-scaling}
+\end{subfigure}%
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=0.95\textwidth]{plots/radioml-knn-10.png}
+\subcaption{RadioML for $k=10$.
+}
+\label{fig:results:radioml-scaling}
+\end{subfigure}%
 
-\caption{Algorithm performance across four datasets.
+\caption{Algorithm performance across six datasets, including a randomly-generated dataset.
 In each plot, the horizontal axis represents increasing cardinality (size) of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).

### 2023-12-20 sections/4-results.tex
updated lfd plots

@@
 \begin{figure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{images/lfd_plots/fashion-mnist-60000-lfd.png}\\
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{images/lfd_plots/glove-25-1183514-lfd.png}\\
@@
 \\
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{images/lfd_plots/sift-1000000-lfd.png}\\
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-128-1000000-lfd.png}\\
+\begin{subfigure}[b]{0.47\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
 \subcaption{A random dataset}
@@
 \\
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=0.95\textwidth]{images/lfd_plots/radioml-97920-lfd.png}\\
+\begin{subfigure}[b]{0.47\textwidth}
+\includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
 \subcaption{RadioML}
@@
 \begin{figure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}
@@
 \\
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{plots/random-1000000-128-knn-10.png}
@@
 \\
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{plots/silva-knn-10.png}
@@
 \end{subfigure}%
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=0.95\textwidth]{plots/radioml-knn-10.png}

### 2023-12-21 sections/1-introduction.tex
updated up to greedy sieve in methods

@@
 Often dubbed ``the Big Data explosion,'' rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale.
-% Najib: We should mention datasets in this order: neural -network embeddings, biological, radio-ml, and then astronomical. DONE. (flow is a bit awakward though)
 
-With the rise of large language models such as GPT-3~\cite{2020arXiv200514165B}, GPT-4~\cite{OpenAI2023GPT4TR}, and LLAMA-2~\cite{Touvron2023Llama2O}, neural network embeddings are becoming an important source of big data.
+% Najib: Add citations for neural-embeddings of images in addition to LLMs
+With the rise of large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, neural network embeddings are becoming an important source of big data.
+
 Biology has also seen an influx of large, high-dimensional datasets.
 The GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length;
-Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with aligned length of 50,000 letters.
+Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
 
 Radio frequency data offers another example of massive datasets. 
-For example, the RadioML~\cite{oshea2018radioml} dataset contains 2,555,504 samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
+The RadioML~\cite{oshea2018radioml} dataset contains 2,555,504 samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
 
-Additionally, many large, high-dimensional datasets come from the field of astronomy.
+The field of astronomy has also seen an influx of large datasets.
 The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
-
-% More recently, Big-ANN's NeurIPS'23: Billion Scale Approximate Nearest Neighbor Search Challenge has 
-% solicits development of data structures and search algorithms for Approximate Nearest Neighbor (ANN) search
-% on four massive datasets: YFCC 100M~\cite{johnson2019billion} dataset, the Yandex Text-to-Image 10M~\cite{diskann-github}, cross-modal dataset, 
-% the MSMARCO passage retrieval dataset~\cite{Bruch2023AnAA}, and a 10M slice of the MS Turing data set~\cite{Singh2021FreshDiskANNAF}. These datasets 
-% are 192, 200, 30,000, and 100 dimensional respectively, and submissions are evaluated with 
-% a ten million point sample of each.
 
@@
 
+% Najib: We should not cite CHESS for the recall-throughput tradeoff. Cite something else.
 As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
-even state-of-the-art algorithms exhibit a steep recall versus throughput tradeoff~\cite{ishaq2019clustered}.
+even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{ishaq2019clustered}.
 
@@
 
-While algorithms for fast $k$-NN on large datasets do exist, most do not exploit the geometric and topological structure inherent in large datasets.
+While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in large datasets.
 Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \emph{exact} search remains.
-For example, for a majority voting classifier, approximate $k$-NN may agree with exact $k$-NN for large values of $k$, but may be sensitive to local perturbations with smaller values of $k$.
+For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations with smaller values of $k$.
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
-this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could exhibit suboptimal accuracy.
+Further, there is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
+this may suggest that approximate $k$-NN search could exhibit suboptimal accuracy.
 
-Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent search algorithms to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear} such as MRPT~\cite{hyvonen2016fast}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, Inverted File indexing~\cite{johnson2019billion}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+
+TODO: Paragraph summarizing LSH.
+
+TODO: Paragraph summarizing HNSW.
+
+TODO: Paragraph summarizing FAISS-IVF.
+
 Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset, rather than its cardinality.
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest neighbors search to a hierarchical clustering approach.
-In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor Entropy-scaling Search).
-CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm. 
-Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform $k$-nearest neighbors search.
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of a dataset, rather than its cardinality.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest neighbors search from a flat clustering approach to a tree-based hierarchical clustering approach.
+In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbors Entropy-scaling Search).
+CLAM (Clustered Learning of Approximate Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm.
+Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-nearest neighbors search.
 
-This paper focuses on four novel algorithms for exact $k$-NN search.
+This paper focuses on three novel algorithms for exact $k$-NN search.
 We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
-Finally, we provide a comparison of CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy} on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
+Finally, we provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy} on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.

### 2023-12-21 sections/2-methods.tex
updated up to greedy sieve in methods

@@
 
-We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ as a function which, 
-given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
-We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x)$ for all $x, y \in \textbf{X}$) and that the distance between two points $x$ and $y$ be zero if and only if $x = y$. 
+We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
+We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and have an identity (i.e. $f(x, y) = 0 \iff x = y$). 
 In addition to these constraints, if the distance function obeys the triangle inequality, i.e., it is a \emph{distance metric}, then we guarantee that CAKES has perfect recall. 
 The choice of distance function varies by dataset.
-For example, with neural-network embeddings, one could use Euclidean (L2) distance or cosine distance.
+For example, with neural-network embeddings, one could use Euclidean (L2) distance or Cosine distance.
 With biological or string data, Levenshtein or Hamming distances are more appropriate.
-With time-series data, one could use dynamic time warping (DTW) distance.
+With time-series data, one could use Dynamic Time Warping (DTW).
 
-Some of our algorithms for $\rho$- and $k$-nearest neighbors search rely on the \emph{local fractal dimension} at some point and length scale in the dataset.
-We define local fractal dimension (LFD) as:
+Some of our search algorithms rely on the \emph{local fractal dimension} (LFD) at some point and length scale in the dataset.
+We define LFD as:
 
@@
     \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) }
-    \label{eq:methods:lfd_original}
+    \label{eq:methods:lfd-original}
 \end{equation}
@@
 However, due to physical constraints (namely, the laws of physics that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional manifold within the 8,575-dimensional embedding space.
-The \emph{local fractal dimension} is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
-This notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space is known as the \emph{manifold hypothesis}~\cite{fefferman2016testing}.
-
-% Najib: Add our own plots of local fractal dimension instead of referencing figures from the CHESS paper. Then, rewrite this paragraph.
-Figures 2 and 3 in~\cite{ishaq2019clustered} illustrate the low fractal dimensions of two datasets used with $\rho$-NN search in CHESS. 
-We observe that for both APOGEE (Figure 2 in~\cite{ishaq2019clustered}) and GreenGenes (Figure 3 in~\cite{ishaq2019clustered}), other than the most extreme 10\% of clusters, virtually all clusters have a local fractal dimension of less than 2.
-This suggests that APOGEE and GreenGenes are good datasets for use with entropy-scaling search algorithms like CHESS and CAKES.
+The LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
+This is the \emph{manifold hypothesis}~\cite{fefferman2016testing}, i.e.
+high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
 
@@
 
-Najib We define a \emph{cluster} as a set of points with a \emph{center} and a \emph{radius}.
-The \emph{center} is the geometric median of a sample of points in the \emph{cluster}, and so it is a real data point.
-The \emph{radius} is the maximum distance of any point in the cluster to the \emph{center}.
+We define a \emph{cluster} as a set of points with a \emph{center} and a \emph{radius}.
+The \emph{center} is the geometric median of the points (or a smaller sample of the points) in the \emph{cluster}, and so it is a real data point.
+The \emph{radius} is the maximum distance from the \emph{center} to any point in the \emph{cluster}.
 Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
-We estimate local fractal dimension at the scale of the cluster radius and half the cluster radius. 
-This allows us to rewrite Equation \ref{eq:methods:lfd_original} as simply: 
+We estimate LFD at the scale of the cluster radius and half that radius. 
+This allows us to rewrite Equation~\ref{eq:methods:lfd-original} as:
+
 \begin{equation} 
-\log_2 \frac{|B_X(q, r)}{|B_X(q, r/2)|}
+    \log_2 \left( \frac{|B_X(q, r)}{|B_X(q, r/2)|} \right)
+    \label{eq:methods:lfd-simplified}
 \end{equation}
 
-% Najib: Rewrite this paragraph. Old paper defined it as ... but we define it for the hierarchical clustering scheme as ... DONE. (let me know if this okay)
-
-In~\cite{yu2015entropy}, metric entropy for a given cluster radius $r_c$ was defined as the number of clusters of radius $r_c$ needed to cover all data.
+In~\cite{yu2015entropy}, metric entropy for a given cluster radius $r_c$ was defined as the number of clusters of a uniform radius $r_c$ needed to cover all data.
 In this paper, we define the metric entropy $\mathcal{N}_{\hat{r_c}}(X)$ of a data set $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree.
@@
 
-We start by building a divisive hierarchical clustering of the dataset using CLAM.
+We start by performing a divisive hierarchical clustering on the dataset using CLAM.
 The procedure is similar to what was outlined in CHESS, but with the following improvements:
-better selection of poles for partitioning and depth-first dataset reordering (see Section ~\ref{subsubsec:methods:clustering:dataset-depth-first-reordering}).
+better selection of poles for partitioning (see Algorithm~\ref{alg:methods:clustering:partition}) and depth-first reordering of the dataset (see Section~\ref{subsubsec:methods:clustering:dataset-depth-first-reordering}).
 
@@
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. 
-While we sometimes use Euclidean notions, such as voids and volumes, to talk about geometric and topological properties of clusters and of the manifold, CLAM does not rely on such notions; 
-they serve merely as a convenient, intuitive way to talk about the underlying mathematics.
+While we sometimes use Euclidean notions, such as voids and volumes, to talk about the geometric and topological properties of clusters and of the manifold, CLAM does not rely on such notions; 
+they serve only as a convenient, intuitive way to talk about the underlying mathematics.
 
@@
 
-For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in the sample.
-Based on those pairwise distances, we compute the \emph{geometric median} of this sample; 
+For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
+Using these distances, we compute the \emph{geometric median} of this sample; 
 i.e., the point which minimizes the sum of distances to all other points in the sample.
-This geometric median is $C$'s \emph{center}.
-The \emph{radius} of $C$ is the maximum distance from any point in $C$ to its center.
-The point $\ell$ which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole}.
-The point $r$ which is furthest from $\ell$ is designated the \emph{right pole}.
-We then partition the cluster into a \emph{left child} and \emph{right child}, where the left child contains all points in the cluster which are closer to $\ell$ than to $r$, and the right child contains all points in the cluster which are closer to $r$ than to $\ell$.
-Without loss of generality, we assign to the left child those points which are equidistant from $\ell$ and $r$.
-Starting from a root-cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum or some other user-specified stopping criterion is met.
+This geometric median is the \emph{center} of $C$.
+
+The \emph{radius} of $C$ is the maximum distance from the \emph{center} to any other point in $C$.
+
+The point which is responsible for that radius (i.e., the furthest point from the \emph{center}) is designated the \emph{left pole}.
+The point which is furthest from \emph{left pole} is designated the \emph{right pole}.
+
+We then partition the cluster into a \emph{left child} and a \emph{right child}, where the left child contains all points in the cluster which are closer to the \emph{left pole} than to the \emph{right pole}, and the right child contains all points in the cluster which are closer to the \emph{right pole} than to the \emph{left pole}.
+Without loss of generality, we assign to the left child those points which are equidistant from the two poles.
+
+Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria.
 This process is described in Algorithm \ref{alg:methods:clustering:partition}.
+
+During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-simplified}.
 
@@
     \begin{algorithmic} % enter the algorithmic environment
-    \STATE $m \Leftarrow \lfloor \sqrt{|C|} \rfloor$
-    \STATE $seeds \Leftarrow m$ random points from $C.points$
+        \REQUIRE $C$, a cluster
+        \STATE $seeds \Leftarrow \lfloor \sqrt{|C|} \rfloor$ random points from $C$
         \STATE $c \Leftarrow$ geometric median of $seeds$
-    \STATE $l \Leftarrow \argmax d(c,x) \ \forall \ x \in C.points$
-    \STATE $r \Leftarrow \argmax d(l,x) \ \forall \ x \in C.points$
-    \STATE $left \Leftarrow \{x | x \in C.points \land d(l,x) \le d(r,x)\}$
-    \STATE $right \Leftarrow \{x | x \in C.points \land d(r,x) < d(l,x)\}$
-    \IF{$|left| > 1$}
-        \STATE Partition($left$)
+        \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+        \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+        \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+        \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+        \IF{$|L| > 1$}
+            \STATE Partition($L$)
         \ENDIF
-    \IF{$|right| > 1$}
-        \STATE Partition($right$)
+        \IF{$|R| > 1$}
+            \STATE Partition($R$)
         \ENDIF
@@
 CAKES also improves upon CHESS by reordering the dataset in depth-first order of traversal of the tree.
-In CHESS, each cluster stored a list of indices into the dataset, which we used during search to retrieve a cluster's points. 
-Though this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high;
-since each cluster stores indices for each of its points and, for a dataset with cardinality $n$, we store a total of $n$ indices at each depth in the tree.
-Assuming a balanced tree and thus $\mathcal{O}\log(n)$ depth, this approach had $\mathcal{O}(n\log(n))$ memory cost.
 
-One may improve this memory cost to $\mathcal{O}(n)$ by only storing these indices at the leaf cluster level.
-This approach, however, introduces an $\mathcal{O}(n\log(n))$ time cost whenever we need to find the indices for 
-a non-leaf cluster, since it requires a traversal of the subtree rooted at that cluster.
+In CHESS, each cluster stored a list of indices into the dataset.
+This list was used to retrieve the clusters' points during search.
+Though this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
+With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
+Assuming a balanced tree, and thus $\mathcal{O}\log(n)$ depth, this approach had a memory overhead of $\mathcal{O}(n\log(n))$.
 
-In this work, we introduce a new approach, wherein after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
-Then, within each cluster, we need only store the cluster's cardinality and an \emph{offset} to access the cluster's points from the dataset.
+One may improve this memory overhead to $\mathcal{O}(n)$ by only storing indices at the leaf clusters.
+This approach, however, introduces a time cost of $\mathcal{O}(n\log(n))$ whenever we need to find the indices for a non-leaf cluster (because it requires a traversal of the subtree rooted at that cluster).
+
+In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
+Then, within each cluster, we need only store its \emph{cardinality} and an \emph{offset} to access the its points from the dataset.
+
 The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
 A left child has the same offset as that of its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
+
 With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers significant improvements over the approach used in CHESS.
@@
 
-The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}
+The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}.
 By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of partitioning and $\mathcal{O}(n \log n)$ cost of building the tree.
-This is a significant improvement over exact partitioning and tree-building, which have $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ costs respectively.
-
-\subsection {Auto-Tuning}
-\label{subsec:methods:auto-tuning}
-
-% In this work, we also introduce a method of \emph{sharding} for datasets that cannot fit in memory.
-% This approach also leverages the fact that $\rho$-nearest neighbors search at the radius of the $k$-th farthest neighbor is typically much faster than $k$-nearest neighbors search.
-
-We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
-With this approach, we begin by taking a random sample of no more than 10\% of the data.
-We build a cluster tree with this sample using Algorithm~\ref{alg:methods:clustering:partition}, and select all clusters at some user-specified depth in the tree, with a default depth of 10.
-Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
-We select the algorithm which had the lowest total search time over all the queries. 
-Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
-
-% Note also that the size of the shards is not necessarily consistent;
-% it is often optimal to have the first shard be smaller than the others, since the first shard is the only one for which we perform $k$-nearest neighbors search. 
-% However, making this first shard too small introduces the risk of significantly overestimating the radius (i.e., the distance to the $k$-th neighbor), resulting in much slower $\rho$-nearest neighbors search on the subsequent shards.
-% Finding the optimal ratio between the size of the first shard and the sizes of the remaining shards, as well as computing the expected level of error between the distance from the $k$-th closest neighbor in the first shard and the actual $k$-th closest neighbor, are topics for future work.
-% For datasets which \emph{do} fit in memory, we use the same approach to auto-tuning, but, since the entire dataset fits in memory, there is only one shard.
+This is a significant improvement over exact partitioning and tree-building, which cost $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ respectively.
 
@@
 \label{subsec:methods:the-search-problem}
-Having explained the clustering process, we can now pose the $k$-NN and $\rho$-NN search problems.
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find 
-the set $S_q$ such that  $|S_q| = k$ and $\forall x \in \textbf{X} \ \setminus \ S_q$, $f(x, q) > \max\{f(y, q): y \in S_q \}$;
-that is, for a given $k$, find the $k$ closest points to $q$ in $ \textbf{X}$.
+
+Given Algorithm~\ref{alg:methods:clustering:partition}, we can now pose the $k$-NN and $\rho$-NN search problems.
+
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
+i.e., for a given $k$, find the $k$ closest points to $q$ in $\textbf{X}$.
 We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
-that is, find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
-
-% Najib: For this figure, I want the red, green and blue lines/braces to be thinner. Perhaps 1/2 to 2/3 as thick.
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[scale=0.5]{images/geometry/deltas.png}
-    \caption{$\delta$, $\delta_{\max}$, and $\delta_{\min}$ defined for cluster $C$ and query $q$.}
-    \label{fig:methods:deltas}
-\end{figure}
+that is, find all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$.
 
@@
     \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
-    \item $\delta_{\max} = \delta + r$ is the distance from the query to the theoretically farthest possible instance in $C$.
-    \item $\delta_{\min} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest possible instance in $C$.
+    \item $\delta^{+} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+    \item $\delta^{-} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
 \end{itemize}
 
-We define a \emph{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains 
-duplicates of the same point (i.e., has cardinality greater than 1 but contains only one unique point).
-A singleton clearly has zero radius, and so $\delta_{\max} = \delta = \delta_{\min}$ for these clusters.
-Hence, we sometimes overload the above notation to also refer to distance from a query to a point;
-in these cases, we also have that $\delta_{\max} = \delta = \delta_{\min}$.
+% Najib: For this figure, I want the red, green and blue lines/braces to be thinner.
+Perhaps 1/2 to 2/3 as thick.
+Also, I changed the notation for the deltas because it makes things easier in the Sieve and Greedy Sieve algorithms' sections.
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[scale=0.5]{images/geometry/deltas.png}
+    \caption{$\delta$, $\delta^{+}$, and $\delta^{-}$ for a cluster $C$ and a query $q$.}
+    \label{fig:methods:deltas}
+\end{figure}
+
+We define a \emph{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \emph{unique} point).
+A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
+Hence, we overload the above notation to also refer to the distance from a query to an individual point.
 
@@
 We conduct $\rho$-nearest neighbors search as described in~\cite{ishaq2019clustered}, but with the following improvement:
-when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which could contain points in the query ball.
-The full algorithm is described in Algorithm~\ref{alg:methods:rho-nn-search}.
+when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which might contain points in the query ball.
 
-% Najib: Merge Figures 2 and 3. There's too much redundancy between them.
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[scale=0.5]{images/geometry/overlapping-children-1.pdf}
-%     \caption{
-%         Triangle formed by query $q$, and poles $\ell$ and $r$. Query ball of radius $\rho$ centered at $q$. 
-%     }
-%     \label{fig:methods:overlapping-children-1}
-% \end{figure}
-
-% Najib: Visually, ``m'' is not centered on the line segment. It's a little bit to the right.
+% Najib: Visually, ``m'' is not centered on the line segment. It's a little bit to the right. Moving ``l'' to the right is probably the easiest fix.
 \begin{figure}[ht!]
@@
     \caption{
-        $m$ denotes the midpoint of $\overline{\ell r}$.
-$q'$ is the projection of $q$ onto $\overline{\ell r}$.
-$d$ is the distance from $q$ to the perpendiculat bisector of $\overline{\ell r}$.
-If $d \leq \rho$, both the left child and right child clusters could contain points in the query ball.
+        $m$ denotes the midpoint of $\overline{l r}$. $q'$ is the projection of $q$ onto $\overline{l r}$.
+        $d$ is the distance from $q'$ to $m$.
+        If $d \leq \rho$, both child clusters could contain points in the query ball.
         Otherwise, only the right child cluster could contain such points. 
@@
 
-To determine whether both children can contain points in the query ball, we consider the following geometry. 
-Let $\ell$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}), $q$ denote the query, and $\rho$ denote the search radius, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
-Without loss of generality, we assume that $\overline{qr \vphantom \ell} \leq \overline{q\ell}$ (i.e., the query is closer to the right pole than to the left pole).
+To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children-3}.
+Here, we overload the notation for $\overline{x y}$ to refer to the line segment joining points $x$ and $y$ as well as the length of that line segment.
 
-Now let $m$ be the midpoint of $\overline{\ell r}$.
-Let $d$ denote the distance from $q$ to the perpendicular bisector of $\overline{lr}$, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
-Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$,
-the left child cannot contain points in the query ball. In this case, we proceed only with the right child. If $d \leq \rho$, 
-then both child clusters could contain points in the query ball, and we proceed with both of them.
+Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}).
+Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$ (i.e., the query is closer to $r$ than to $l$).
 
-We can do a quick check using values computed during partitioning (see Section~\ref{subsubsec:methods:cluster-partitioning}) to determine whether $d \leq \rho$. 
-In the caculations that follow, we overload notation a bit and use $\overline{xy}$ to mean the length of the line segment between $x$ and $y$.
-We note that $d = \overline{mq'} = \overline{mr \vphantom {q'}}-\overline{q'r}$.
-Since $m$ is the midpoint of $\overline{\ell r}$, we have that $\overline{mr} = \tfrac{\overline{\ell r}}{2}$, so \begin{equation*} d = \tfrac{\overline{\ell r}}{2} - \overline{q'r}.
+Now let $q'$ be the projection of $q$ to $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$ and $d$ be the distance from $q'$ to $m$.
+Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$ then the left child cannot contain points inside the query ball.
+In this case, we proceed only with the right child.
+If, however, $d \leq \rho$ then both child clusters could contain points in the query ball, and so we proceed with both of them.
+
+To check whether $d \leq \rho$, we note that
+
+\begin{equation}
+    d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}.
     \label{eq:methods:d-expression}
+\end{equation}
+
+Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
+By the Law of Cosines on $\triangle l r q$, we have that
+
+\begin{equation}
+    \text{cos}(\theta) = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}.
+    \label{eq:methods:cosine-law}
+\end{equation}
+
+Since $\triangle r q q'$ is a right triangle, we also have that
+
+\begin{equation}
+    \text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}.
+    \label{eq:methods:cosine-law-right-triangle}
+\end{equation}
+
+Combining Equations~\ref{eq:methods:cosine-law} and~\ref{eq:methods:cosine-law-right-triangle}, we have that
+
+\begin{equation}
+    \overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}.
+    \label{eq:methods:q-prime-r}
+\end{equation}
+
+Substituting Equation~\ref{eq:methods:q-prime-r} into Equation~\ref{eq:methods:d-expression}, we have that
+
+\begin{equation*}
+    d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}.
 \end{equation*}
 
-It remains to find $\overline{q'r}$.
-To do this, we use the Law of Cosines with $\triangle \ell rq$.
-Let $\theta$ denote $\angle lrq$, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
-By the Law of Cosines, we have that
-\begin{equation} \text{cos}(\theta) = \tfrac{\overline{qr\vphantom \ell}^2 + \overline{\ell r}^2 - \overline{q\ell}^2}{2(\overline{\ell r})(\overline{qr \vphantom \ell})}.\end{equation}
-On the other hand, since $\triangle q'qr$ is a right triangle,
+Thus,
 
-\begin{equation} \text{cos}(\theta) = \tfrac{\overline{q'r}}{\overline{qr \vphantom \ell}}.\end{equation}
-
-Equating the two expressions for $\text{cos}(\theta)$ and solving for $\overline{q'r}$, we have \begin{equation} \overline{q'r} = \tfrac{\overline{qr\vphantom \ell}^2 + \overline{\ell r}^2 - \overline{q\ell}^2}{2 \overline{\ell r}}.
+\begin{equation}
+    d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho.
+    \label{eq:methods:check-d}
 \end{equation}
 
-Substituting this back in equation~\ref{eq:methods:d-expression} gives 
-\begin{equation} d = \tfrac{\overline{q\ell}^2 - \overline{qr \vphantom \ell}^2}{2\overline{lr}}.\end{equation}
+Note, in particular, that Equation~\ref{eq:methods:check-d} only requires distances between real points, and so it can be used with any distance function, even when $q'$ and $m$ are not real points or cannot be imputed from the data.
 
-Then to check if $d \leq \rho$, it suffices to verify that \begin{equation}(\overline{q\ell} + \overline{qr \vphantom \ell})(\overline{q\ell}- \overline{qr \vphantom \ell}) \leq 2\overline{\ell r}\rho.
-\end{equation}
-
-% Najib: move is.Leaf check to the beginning of the algorithm because it's the base base case. DONE (confirm this is correct)
+To perform $\rho$-nearest neighbors search, we first perform a coarse tree-search, as outlined in Algorithm~\ref{alg:methods:rnn:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
+Then, for all such clusters, we perform a finer-grained search, as outlined in Algorithm~\ref{alg:methods:rnn:leaf-search}, to find all points in which are within a distance $\rho$ of the query.
 
 \begin{algorithm} 
-    \caption{$\rho$-NN(\emph{cluster, query, r})} 
-    \label{alg:methods:rnn-search} 
+    \caption{tree-search($C$, $q$, $\rho$)} 
+    \label{alg:methods:rnn:tree-search} 
     \begin{algorithmic}
-        \REQUIRE $r \geq 0$
-        \REQUIRE $cluster \neq \emptyset$
-        \STATE $results \Leftarrow \emptyset$
-        \IF{$cluster.isLeaf$}
-            \FOR{$p \in cluster.points$}
-                \IF{$p.\delta \leq r$}
-                    \STATE $results.push((p, r))$
+        \REQUIRE $C$, a cluster
+        \REQUIRE $q$, a query
+        \REQUIRE $\rho$, a search radius
+
+        \STATE $c$ $\Leftarrow$ \emph{center} of $C$
+        \STATE $r$ $\Leftarrow$ \emph{radius} of $C$
+
+        \IF{$C$ is a leaf \textbf{or} $f(q, c) + r \leq \rho$}
+            \STATE \textbf{return} $\{C\}$
+        \ELSIF{Inequality~\ref{eq:methods:check-d} is true}
+            \STATE $L$ $\Leftarrow$ \emph{left child} of $C$
+            \STATE $R$ $\Leftarrow$ \emph{right child} of $C$
+            \STATE \textbf{return} tree-search($L, q, \rho$) $\cup$ tree-search($R, q, \rho$)
+        \ELSIF{$f(q, l) < f(q, r)$}
+            \STATE \textbf{return} tree-search($L, q, \rho$)
+        \ELSE
+            \STATE \textbf{return} tree-search($R, q, \rho$)
+        \ENDIF
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm} 
+    \caption{leaf-search($S$, $q$, $\rho$)} 
+    \label{alg:methods:rnn:leaf-search} 
+    \begin{algorithmic}
+        \REQUIRE $S$, a set of clusters
+        \REQUIRE $q$, a query
+        \REQUIRE $\rho$, a search radius
+
+        \STATE $H \Leftarrow \emptyset$, a set of hits
+
+        \FOR{$C \in S$}
+            \STATE $c$ $\Leftarrow$ \emph{center} of $C$
+            \STATE $r$ $\Leftarrow$ \emph{radius} of $C$
+            \IF{$f(q, c) + r \leq \rho$}
+                \STATE $H$ $\Leftarrow$ $H \cup \{C\}$
+            \ELSE
+                \FOR{$p \in C$}
+                    \IF{$f(p, q) \leq \rho$}
+                        \STATE $H$ $\Leftarrow$ $H \cup \{p\}$
                     \ENDIF
                 \ENDFOR
-            \STATE Return $results$
             \ENDIF
-        \IF{$cluster.left \cap B_X(query, r) \neq \emptyset$}
-            \IF{$cluster.left.\delta$ $\leq$ $r$ + $cluster.left.radius$}
-                \STATE{$\rho$-NN($cluster.left, query, r$)}
-            \ENDIF
-        \ENDIF
-        \IF{$cluster.right \cap B_X(query, r) \neq \emptyset$}
-            \IF{$cluster.right.\delta$ $\leq$ $r$ + $cluster.right.radius$} 
-                \STATE $\rho$-NN($cluster.right, query, r$)
-            \ENDIF
-        \ENDIF
+        \ENDFOR
 
+        \STATE \textbf{return} $H$
     \end{algorithmic}
-    \label{alg:methods:rho-nn-search}
+\end{algorithm}
+
+\begin{algorithm} 
+    \caption{$\rho$-NN-search($root$, $q$, $\rho$)} 
+    \label{alg:methods:rnn:rnn-search} 
+    \begin{algorithmic}
+        \REQUIRE $root$, the root cluster
+        \REQUIRE $q$, a query
+        \REQUIRE $\rho$, a search radius
+
+        \STATE $S$ $\Leftarrow$ tree-search($root$, $q$, $\rho$)
+        \STATE $H$ $\Leftarrow$ leaf-search($S$, $q$, $\rho$)
+        \STATE \textbf{return} $H$
+    \end{algorithmic}
 \end{algorithm}
@@
 \begin{gather}
-    \mathcal{O}\Bigg(
-    \underbrace{\log_2 \mathcal{N}_{\hat{r_c}}(X)}_{\textrm{metric entropy}} +
-    \overbrace{\left|B_X(q,r)\right|}^{\textrm{output size}}
-    \underbrace{\left(1+\frac{2\hat{r_c}}{r}\right)^d}_{\textrm{scaling factor}}\Bigg)
-    \label{hierarchical-complexity}
+    \mathcal{O}
+    \Bigg(
+        \underbrace{\log_2 \mathcal{N}_{\hat{r}}(X)}_{\textrm{metric entropy}} +
+        \overbrace{ \left| B_X(q, \rho) \right|}^{\textrm{output size}}
+        \underbrace{ \left( 1 + \frac{2 \cdot \hat{r}}{ \rho} \right) ^ d}_{\textrm{scaling factor}}
+    \Bigg)
+    \label{eq:methods:rnn:complexity}
 \end{gather}
 
-where $\hat{r_c}$ is the \textit{mean} cluster radius of leaf clusters, $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy at that radius, $B_X(q,r)$ is a ball of data points around the query $q$ at search radius $r$, and $d$ is the local fractal dimension around the query in a region of radius $r$.
+where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query in a region of radius $r$.
+From Algorithm~\ref{alg:methods:rnn:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-nearest neighbors search performance scales linearly with the size of the output set and exponentially with the local fractal dimension around the query.
+While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the local fractal dimension around the query is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.
 
@@
 
-In this section, we present four novel algorithms for exact $k$-nearest neighbors search: $k$-NN by Repeated $\rho$-NN, Sieve Search, Sieve Search with Separate Centers, and Greedy Sieve. 
-We use a preliminary auto-tuning function to predict the variant (see Section~\ref{subsec:methods:sharding-and-auto-tuning}) which will perform 
-best for a given dataset and value of $k$.
-We then proceed with search using that variant on the first shard.
+In this section, we present three novel algorithms for exact $k$-nearest neighbors search:
+Repeated $\rho$-NN, Sieve, and Greedy Sieve.
 
-In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and
-$Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the nearest neighbors.
+In these algorithms, we use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
 
-\subsubsection{\texorpdfstring{$k$}{k}-NN by Repeated \texorpdfstring{$\rho$}{p}-NN}
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
 \label{subsubsec:methods:knn-search:repeated-rnn}
 
-In this algorithm, we perform $\rho$-nearest neighbors search starting with a small search radius,
-\subsubsection{Sieve with Separate Centers}
-\label{subsubsec:methods:knn-search:sieve2}
+In this algorithm, we perform $\rho$-nearest-neighbors search starting with a small search radius, and repeatedly increasing the radius until $|H| \geq k$.
 
-This algorithm is the same as Sieve, but with the following modification:
-clusters in $Q$ are represented twice -- once as their center and once as the rest of their points.
-The clusters' multiplicities in $Q$ are adjusted to be one less than their cardinalities.
-Because we have that for any cluster $C$, $C.\delta \leq C.\delta_{max}$, representing the cluster center separately from the rest of the points causes the threshold $\tau$ to decrease more quickly, thus allowing us to eliminate some clusters from $Q$ earlier in the process.
-and repeatedly increasing the radius until we find $k$ neighbors.
+Let the search radius, $r$, be equal to the radius of the root cluster divided by the cardinality of the dataset.
+We perform tree-search with radius $r$.
+If no clusters are returned, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
+Let $S$ be the set of clusters returned by the first tree-search which returns at least one cluster.
 
-Let $H$ be the set of nearest neighbors found thus far.
-Search starts with a radius $m$ equal to the radius of the root cluster divided by the cardinality of the dataset.
-We perform $\rho$-NN search with radius $m$. 
-If no points are within a distance $m$ of the query, we increase the radius by a factor of 2 and perform $\rho$-NN search again, repeating until at least one point is found.
-
-Once $|H| > 0$, we continue to perform $\rho$-NN search, but instead of increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined by the local fractal dimension in the region of the dataset around the query ball.
+Now, so long as $\sum_{C \in S} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
 In particular, we increase the radius by a factor of
@@
 \begin{equation}
-    \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)
+    \min \left(2, \left( {\frac{k}{\sum_{C \in S} |C|}} \right)^{\mu^{-1}} \right)
     \label{eq:repeated-rnn-factor}
@@
 
-where $\mu$ is the harmonic mean of the local fractal dimensions (abbreviated lfd in algorithm below) of each cluster which contains one of the nearest neighbors found thus far.
-We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high fractal dimension. 
+where $\mu$ is the harmonic mean of the LFD of the clusters in $S$.
+We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD. 
 Intuitively, the factor by which we increase the radius should also be \emph{inversely} related to the number of points found so far. 
-Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.
-In other words, a small increase in the radius would likely encounter unoccupied space around the manifold, so a larger radial increase is needed.
-Thus, the factor of radius increase should be \emph{inversely} related to the local fractal dimension.
-Once $|H| >= k$, we sort the points by their distance to the query and return the $k$ closest points.
+Additionally, when the LFD at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.
+In such a region, a small increase in the radius would likely encounter unoccupied space, so a larger radial increase is needed.
+Thus, the factor of radius increase should be \emph{inversely} related to the LFD.
+
+Once $\sum_{C \in S} |C| >= k$, we are guaranteed to have found at least $k$ neighbors, and so we can stop increasing the radius.
+We perform $\rho$-nn-search with radius $r$ and return the $k$ nearest neighbors.
 
 \begin{algorithm} % enter the algorithm environment
-    \caption{Repeated $\rho$-NN(\emph{root, query, k})} % give the algorithm a caption
-    \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
+    \caption{repeated-$\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
+    \label{alg:knn:repeated-rnn} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
-        \STATE $H \Leftarrow$ $\emptyset$
-        \STATE $m \Leftarrow$ $\frac{root.radius}{|root|}$
-        \WHILE {$H = \emptyset$}
-            \STATE $H.push(\rho$-NN$(root, query, m)$)
-            \STATE $m \Leftarrow 2m$
+        \REQUIRE $root$, the root cluster
+        \REQUIRE $q$, a query
+        \REQUIRE $k$, the number of neighbors to find
+
+        \STATE $r \Leftarrow radius$ of the $root$ cluster
+        \STATE $r \Leftarrow$ $\frac{r}{|root|}$
+        \STATE $S \Leftarrow \emptyset$, a set of clusters
+
+        \WHILE {$S = \emptyset$}
+            \STATE $S \Leftarrow$ tree-search($root$, $q$, $r$)
+            \STATE $r \Leftarrow 2 \cdot r$
         \ENDWHILE
-        \WHILE {$|H| < k$}
-            \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
-            \STATE $\mu \Leftarrow \frac{|Q|}{\sum_{C \in Q} \frac{1}{C.lfd}} $
-            \STATE $m \Leftarrow \min \left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
-            \STATE $H.push(\rho$-NN$(root, query, m)$)
-        \ENDWHILE
-        \STATE $H.sort()$
-        \STATE Return $H[0 \cdots k-1]$
+
+        \ALOOP{}
+            \IF{$\sum_{C \in S} |C| >= k$}
+                \STATE \textbf{break}
+            \ENDIF
+            \STATE $\mu \Leftarrow \frac{|S|}{\sum_{C \in S} \frac{1}{LFD(C)}}$
+            \STATE $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in S} |C|}} \right)^{\mu^{-1}} \right)$
+            \STATE $S \Leftarrow$ tree-search($root$, $q$, $r$)
+        \ENDLOOP
+
+        \STATE $H \Leftarrow$ sort($\rho$-NN-search($root$, $q$, $r$))
+        \STATE \textbf{return} $H[.. k]$
     \end{algorithmic}
@@
 
-\subsubsection{Complexity of \texorpdfstring{$k$}{k}-NN by Repeated \texorpdfstring{$\rho$}{p}-NN Search}
+\subsubsection{Complexity of Repeated \texorpdfstring{$\rho$}{p}-NN}
 \label{subsubsec:methods:repeated-rnn-complexity}
 
-Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution (i.e.
-arises from the same generating phenomenon) as the rest of the data.
-Given the uses of $k$-NN search in practice, this assumption is not unreasonable.
-From this assumption, we can infer that the local fractal dimension at the query point does not differ significantly from the mean local fractal dimension of clusters nearby the query.
+The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data, i.e.
+it arises from the same generative process as the rest of the data.
+Given the uses of $k$-NN search in practice, this assumption is reasonable.
+From this assumption, we can infer that the LFD at the query point does not differ significantly from the (harmonic) mean LFD of clusters near the query.
 
-We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{coarse search} and \emph{fine search} separately.
-Coarse search refers to the process of identifying clusters which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
-Fine search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.
+We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{tree-search} and \emph{leaf-search} separately.
+Tree-search refers to the process of identifying clusters which have overlap with the query ball, i.e., which might contain one of the $k$ nearest neighbors. 
+Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by tree-search.
 
-In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log\mathcal{N}_{\hat{r_c}}(X))$, 
-where $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy of the dataset $X$. 
-To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations of $\rho$-NN needed to find at least $k$ neighbors.
+In ~\cite{ishaq2019clustered}, we showed that the complexity of tree-search is $\mathcal{O}(\log\mathcal{N}_{\hat{r}}(X))$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset $X$ at a radius $\hat{r}$.
+To adjust this bound for Repeated $\rho$-NN, we must estimate the number of iterations of tree-search needed to find at least $k$ neighbors.
 
-Based on the assumption that the local fractal dimension at the query point is not significantly different than that of nearby clusters, the factor from Equation \ref{eq:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of $\rho$-NN to find $k$ neighbors:
+Based on the assumption that the LFD at the query point is not significantly different from that of nearby clusters, the factor from Equation~\ref{eq:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
 one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
-Since this is a constant factor, it does not affect the asymptotic complexity of coarse search, and thus we have that the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2\mathcal{N}_{\hat{r_c}}(X))$.
+Since this is a constant factor, it does not affect the asymptotic complexity of Repeated $\rho$-NN.
 
-To determine the asymptotic complexity of fine search, we must estimate $|H|$, the expected number of hits we examine.
-With $\rho$-NN search, $H$ is the union of clusters which are within a distance of $r + \hat{r_c}$ from the query, where $r$ is the search radius, and $\hat{r_c}$ is the mean radius of leaf clusters, as in \cite{yu2015entropy}.
-With $k$-NN by Repeated $\rho$-NN, however, the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which is query-dependent.
-Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of $H$ to be the union of clusters which are within a distance of $r' + \hat{r_c}$ from the query, where $r'$ is the distance to the $k$th nearest neighbor and $\hat{r_c}$ is defined as before.
+To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in S} |C|$, the number of points we examine.
+Let $\rho_k$ be the distance from the query to the $k$th nearest neighbor.
+Then, we see that $S$ is expected to be the set of clusters which overlap with a ball of radius $\rho_k$ around the query.
+We can estimate this region as a ball of radius $\rho_k + \hat{r}$, where $\hat{r}$ is the mean radius of clusters in $S$.
 
@@
 \begin{equation*}
-    |H| \leq \gamma |B(q, r')|\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
+    \sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(1+ \frac{2\hat{r}}{\rho_k} \right)^d
 \end{equation*}
@@
 where $\gamma$ is a constant. 
-By definition of 
-$r'$, we have that $|B(q, r')| = k$. Thus, we have that
+By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k$.
+Thus,
 
 \begin{equation}
-    |H| \leq \gamma k\left(1+ \frac{2\hat{r_c}}{r'}\right)^d
-    \label{eq:methods:knn-by-rnn-fine-search}
+    \sum_{C \in S} |C| \leq \gamma k \left( 1+ \frac{2\hat{r}}{\rho_k} \right)^d
+    \label{eq:methods:repeated-rnn-leaf-search}
 \end{equation}
 
-It remains to provide an estimate for $r'$. 
-To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, and thus the local fractal dimension at the query point is not significantly different than that of nearby clusters.
+It remains to provide an estimate for $\rho_k$. 
+For this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different than that of nearby clusters.
 
-We let $\hat{d}$ be the mean of the local fractal dimensions of the clusters near the query point (i.e., the cluster identified during coarse search).
-While ordinarily we compute local fractal dimension by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $r'$, we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the mean cardinality $\hat{C}$ of clusters near the query at a radius equal to the mean of their radii $\hat{r}$.
-We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could replace the query with the center of one of the clusters near it without significantly changing the local fractal dimension at the query point.
+We let $\hat{d}$ be the mean LFD of the clusters in $S$.
+While ordinarily we compute LDS by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \emph{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{C}$, of clusters in $S$ at a radius equal to the mean of their radii, $\hat{r}$.
 
-By Equation ~\ref{eq:repeated-rnn-factor}, we then have that
+We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could replace the query with the center of one of the nearby clusters without significantly changing the LFD at the query point.
+
+By Equation~\ref{eq:repeated-rnn-factor}, we then have that
 
 \begin{equation*}
-    \hat{d} = \frac{\log{}\frac{\hat{C}}{k}}{\log{}\frac{\hat{r}}{r'}}
+    \hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}
 \end{equation*}
 
-Since $\hat{d}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve equation~\ref{eq:methods:knn-by-rnn-fine-search} for $r'$ to get
+Since $\hat{d}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve Equation~\ref{eq:methods:repeated-rnn-leaf-search} for $\rho_k$ to get
 
 \begin{equation*}
-    r' = \hat{r}\left(\frac{k}{\hat{C}}\right)^{\frac{1}{\hat{d}}}
+    \rho_k = \hat{r} \left( \frac{k}{\hat{C}} \right)^{\hat{d}^{-1}}
 \end{equation*}
 
-Simplifying the expression for fine search, we have that
+Simplifying the term for leaf search in Equation~\ref{eq:methods:rnn:complexity}, we get
 
 \begin{equation*}
-    k\left(1+ \frac{2\hat{r_c}}{\hat{r}}\left(\frac{k}{\hat{C}}\right)^{\frac{-1}{\hat{d}}}\right)^d
+    k \left( 1+ \frac{2\hat{r}}{\rho_k} \left( \frac{k}{\hat{C}} \right) ^ {\hat{d}^{-1}} \right)^d
 \end{equation*}
 
-Since we are dealing with asymptotic complexity, we can ignore the additive factor of $1$.
-Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has a complexity of
+Since we are performing an asymptotic analysis, we can ignore the additive factor of $1$.
+By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has an asymptotic complexity of
 
 \begin{gather}
-    \mathcal{O}\Bigg(
-    \underbrace{\log_2{\mathcal{N}_{\hat{r_c}}(X)}}_{\textrm{metric entropy}} + 
+    \mathcal{O}
+    \Bigg(
+        \underbrace{\log_2{\mathcal{N}_{\hat{r}}(X)}}_{\textrm{metric entropy}} \ \ \ + 
         \overbrace{k}^{\textrm{output size}}
-    \underbrace{\left(\frac{2\hat{r_c}}{\hat{r}}\right)^d\left(\frac{k}{\hat{C}}\right)^{\frac{-d}{\hat{d}}}}_{\textrm{scaling factor}}\Bigg)
+        \underbrace{
+            \left( \frac{2\hat{r}}{\rho_k} \right) ^ d \left( \frac{k}{\hat{C}} \right) ^ {d \cdot \hat{d}^{-1}}
+        }_{\textrm{scaling factor}}
+    \Bigg)
     \label{eq:methods:knn-by-rnn-complexity}
@@
 
-where $\mathcal{N}_{\hat{r_c}}(X)$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, and $k$ is the number of nearest neighbors.
-We note that the fraction $\frac{-d}{\hat{d}}$ should be relatively close to 1 unless fractal dimension is highly variable in a small region (that is, $\hat{d}$ differs significantly from $d$).
+where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
+We note that the fraction $d \cdot \hat{d}^{-1}$ should be relatively close to 1 unless fractal dimension is highly variable in a small region (that is, $\hat{d}$ differs significantly from $d$).
 
 \subsubsection{Sieve}
-\label{subsubsec:methods:knn-search:sieve}
+\label{subsubsec:methods:knn:sieve}
 
-In this algorithm, we begin by letting $Q$ be a list of clusters and points. 
-{ \color{red} We represent each cluster twice in $Q$ -- once as just its center and once as the rest of its points.
-$Q$ is initialized with two elements: the root cluster and the root cluster's center.
-}
+This algorithm performs a breadth-first traversal of the tree, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
 
-While $Q$ contains at least one cluster (i.e., $Q$ is not a list of only points), we repeat the process described in 
-the following paragraphs.
+We begin by letting $Q$ be a list of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.
 
-First, we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{\max}$ such that 
-$q_{\tau}$ and all elements in $Q$ with smaller $\delta_{\max}$ collectively contain at least $k$ points.
+We then use the QuickSelect algorithm to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$ while accounting for multiplicities, i.e.
+we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
 
-To find this element, we use a process similar to the Partition algorithm used in Quicksort~\cite{10.1093/comjnl/5.1.10}, adjusted to account for the varying cardinalities of clusters in $Q$ by having each cluster count a number of times equal to its cardinality.
-Our modified version of this algorithm finds the smallest index $i$ such that all elements in $Q$ with $\delta_{\max}$ less than or equal to $Q[i]$'s $\delta_{\max}$ collectively have cardinality greater than or equal to $k$.
-We consider $Q[i]$ to be our $q_{\tau}$, and our threshold $\tau$ to be $Q[i]$'s $\delta_{\max}$.
+We then remove from $Q$ any element for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
+Next, we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
+Finally, we replace all non-leaf clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.
 
-We then remove from $Q$ any element whose $\delta_{\min}$ is greater than $\tau$.
-Then we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
-Finally, we replace all non-leaf clusters in $Q$ with their children.
-
-We continue this process until $Q$ contains only points.
-At this point, we just use the Quicksort Partition algorithm once more to find the $k$th nearest neighbor and return all points to the left of it.
+We continue this process until $Q$ no longer contains any clusters.
+We just use the QuickSelect algorithm once last time to reorder $Q$, to find the $k^{th}$ element, and to return the $k$ nearest neighbors.
 
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{Sieve(\emph{root, query, k})} % give the algorithm a caption
+    \caption{Sieve($root$, $q$, $k$)} % give the algorithm a caption
     \label{alg:methods:sieve} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
-        \REQUIRE $root$, $query$, $k$
-        \REQUIRE $root.cardinality \geq k$
-        \STATE $Q \Leftarrow$ [$root$]
-        \WHILE{$|Q| > k$}
-            \STATE $i \Leftarrow QuicksortPartition(Q, k)$
-            \STATE $\tau \Leftarrow Q[i].\delta_{max}$
-            \FOR {$q \in Q$}
-                \IF {$q.\delta_{min} > \tau$}
-                    \STATE $Q.pop(q)$
+        \REQUIRE $root$, the root cluster
+        \REQUIRE $q$, a query
+        \REQUIRE $k$, the number of neighbors to find
+
+        \STATE $c \Leftarrow$ \emph{center} of $root$
+        \STATE $Q \Leftarrow$ [($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1)]
+
+        \ALOOP{}
+            \IF{$k = \sum_{(\_, \_, m) \in Q} m$}
+                \STATE \textbf{break}
+            \ENDIF
+            \STATE $\tau \Leftarrow$ QuickSelect($Q$, $k$)
+            \FOR{$(p, \delta^{+}_{p}, m) \in Q$}
+                \IF{$\delta^{-}_{p} > \tau$}
+                    \STATE $Q.remove((p, \delta^{+}_{p}, m))$
                 \ENDIF
             \ENDFOR
-            \FOR {$q \in Q$}
-                \IF {$q.isLeaf \lor |q| \leq k$}
-                    \STATE $Q.append(q.points)$
+            \FOR{$(p, \delta^{+}_{p}, m) \in Q$}
+                \IF{$p$ is a point}
+                    \STATE \textbf{continue}
+                \ELSIF{$p$ is a leaf}
+                    \FOR{$c \in p$}
+                        \STATE $Q.push((c, \delta_{c}, 1))$
+                    \ENDFOR
                 \ELSE
-                    \STATE $[l, r] \Leftarrow q.children$
-                    \STATE $Q.append([l, r])$   
+                    \STATE $[l, r] \Leftarrow$ children of $p$
+                    \STATE $Q.append([(l, \delta^{+}_{l}, |l| - 1), (l_c, \delta_{l}, 1)])$
+                    \STATE $Q.append([(r, \delta^{+}_{r}, |r| - 1), (r_c, \delta_{r}, 1)])$
                 \ENDIF
-                \STATE $Q.remove(q)$
+                \STATE $Q.remove((p, \delta^{+}_{p}, m))$
             \ENDFOR
-        \ENDWHILE
-        \STATE Return $Q$
+        \ENDALOOP
+
+        \STATE \textbf{return} $Q$
     \end{algorithmic}
@@
 
+\subsubsection{Complexity of Sieve}
+\label{subsubsec:methods:sieve-complexity}
+
+The asymptotic complexity of Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
+This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.
+
+We start by considering clusters whose cardinalities are within a small factor of $k$, say a factor of 2.
+The distribution of such clusters will mimic the distribution of points in the dataset. 
+From this assumption, we can infer that the number of clusters which overlap with the query ball is a function of the LFD near the query.
+In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the LFD near the query point, e.g.
+we could have a cluster overlapping the query ball at each end of each of $d$ mutually-orthogonal axes.
+
+In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log{\frac{n}{2k}}$ iterations of breadth-first traversal to reach clusters whose cardinalities are at most $2k$.
+However, since high cardinality clusters occur only at low depths in the tree, the length of $Q$, and thus the cost of QuickSelect, for those iterations is relatively low.
+
+After the first $\log{\frac{n}{2k}}$ iterations, we are left with clusters in $Q$ whose cardinalities are at most $2k$.
+At this point, based on our assumption, we have that the length of $Q$ is close to $2d$.
+Therefore, the asymptotic complexity of the Sieve Algorithm is $\mathcal{O} \left( \log(\frac{n}{d}) \right)$.
+
 \subsubsection{Greedy Sieve}
-\label{subsubsec:methods:knn-search:greedy-search}
+\label{subsubsec:methods:knn:greedy-sieve}
 
-We first let $Q$ be a min-queue of clusters prioritized by $\delta_{\min}$.
-$Q$ starts containing only the root cluster.
-We also let $H$ be a fixed-size ($k$) max-queue of points prioritized by $\delta$.
-$H$ starts empty.
+This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
 
-While $H$ has fewer than $k$ points or $\delta$ of the top priority element in $H$ is greater than $\delta_{\min}$ of the top priority element in $Q$, we do the following:
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
+$Q$ starts containing only the root cluster while $H$ starts empty.
 
-\begin{enumerate}
+So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
+
+\begin{itemize}
     \item While the top priority element $q$ in $Q$ is not a leaf, pop it from $Q$ and push its children to $Q$.
     \item For each point $p \in q$, push $p$ to $H$.
-\item Pop from $H$ until $H$ has $k$ points. 
-\end{enumerate}
+    \item If $H$ has more than $k$ points, pop from $H$ until $H$ has $k$ points.
+\end{itemize}
 
-After this process, the points left in $H$ are the $k$ nearest neighbors of $q$.
-This algorithm is described in Algorithm~\ref{alg:greedy-sieve}.
+This algorithm is described in Algorithm~\ref{alg:methods:greedy-sieve}.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.
+when the top priority element in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
 
 \begin{algorithm} 
-\caption{Greedy Sieve(\emph{root, query, k})} 
-\label{alg:greedy-sieve} 
+\caption{Greedy Sieve($root$, $q$, $k$)} 
+\label{alg:methods:greedy-sieve} 
 \begin{algorithmic}
-    \STATE{$Q \Leftarrow$ priority queue}
-    \STATE{$Q.push(root)$}
-    \STATE{$H \Leftarrow$ priority queue}
-    \WHILE{$|H| < k \lor H.peek.\delta > Q.peek.\delta_{\min}$}
-        \WHILE{$\neg Q.peek.isLeaf$}
+    \REQUIRE $root$, the root cluster
+    \REQUIRE $q$, a query
+    \REQUIRE $k$, the number of neighbors to find
+
+    \STATE{$Q \Leftarrow [root]$, a min-priority queue by $\delta^{-}$}
+    \STATE{$H \Leftarrow []$, a max-priority queue by $\delta$}
+
+    \WHILE{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
+        \WHILE{$\neg (Q.peek$ is a leaf)}
             \STATE{$[l, r] \Leftarrow Q.pop().children$}
-            \STATE{$Q.push([l, r])$}
+            \STATE{$Q.push(l)$}
+            \STATE{$Q.push(r)$}
         \ENDWHILE
         \STATE{$leaf \Leftarrow Q.pop()$}
-        \STATE{$H.push(leaf)$}
+        \FOR{$p \in leaf$}
+            \STATE{$H.push(p)$}
+        \ENDFOR
         \WHILE{$|H| > k$}
@@
 
-\subsubsection{Complexity of Sieve Algorithms}
-\label{paragraph:methods:sieve-complexity}
+\subsubsection{Complexity of Greedy Sieve}
+\label{paragraph:methods:greedy-sieve-complexity}
 
-We group the Sieve-based algorithms together, as their complexity analyses are similar.
-The complexity of Sieve and Sieve with Separate Centers is dominated by the expected cost of Quicksort's partition algorithm to calculate thresholds.
-This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.
+TODO: Analyze complexity of Greedy Sieve.
 
-We start by considering clusters whose cardinalities are within a small factor of $k$, say a factor of 2.
-The distribution of such clusters will mimic the distribution of points in the dataset. 
-From this assumption, we can infer that number of clusters which overlap with the query ball is a function of the local fractal dimension near the query.
-In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the local fractal dimension of the dataset near the query point, e.g.
-we could have a cluster overlapping the query ball at each end of each of the $d$ axes.
+\subsection {Auto-Tuning}
+\label{subsec:methods:auto-tuning}
 
-In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log{\frac{n}{2k}}$ iterations of Sieve to partition clusters enough that their cardinalities are at most $2k$.
-However, since high cardinality clusters occur only at low depths in the tree, the length of $Q$ for those iterations is relatively low.
-
-After the first $\log{\frac{n}{2k}}$ iterations, we have that the cardinalities of clusters in $Q$ are at most $2k$, at which point, based on our assumption, we have that the length of $Q$ is close to $2d$.
-Therefore, the asymptotic complexity of the sieve algorithms is $\mathcal{O}\left(\log(\frac{n}{d})\right)$.
+We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
+With this approach, we begin by taking a random sample of no more than 10\% of the data.
+We build a cluster tree with this sample using Algorithm~\ref{alg:methods:clustering:partition}, and select all clusters at some user-specified depth in the tree, with a default depth of 10.
+Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
+We select the algorithm which had the lowest total search time over all the queries. 
+Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
 

### 2023-12-21 sections/4-results.tex
updated knn plots
added radio-ml plots with knn throughput
using scientific notation for throughput in results tables
updated up to greedy sieve in methods

@@
 
-\caption{Local fractal dimension vs. cluster depth, grouped by decile of LFD, across six datasets.}
+    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of LFD and weighted by the cardinalities of the clusters.}
+    \label{fig:results:lfd-plots}
 \end{figure}
@@
 
-Figures~\ref{fig:plots:fashion-mnist-scaling}A, ~\ref{fig:plots:glove-25-scaling}B, and ~\ref{fig:plots:sift-scaling}C show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
-Figure~\ref{fig:plots:random-scaling}D shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
+Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
 The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
@@
 only three decimal places.
-Figures~\ref{fig:plots:fashion-mnist-scaling}A, ~\ref{fig:plots:glove-25-scaling}B, and ~\ref{fig:plots:sift-scaling}C show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
-Figure~\ref{fig:plots:random-scaling}D shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
+Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
 The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
@@
 \begin{subfigure}[b]{0.47\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/random-1000000-128-knn-10.png}
+\includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
 \subcaption{A random dataset for $k=10$.
@@
 \begin{subfigure}[b]{0.47\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/radioml-knn-10.png}
-\subcaption{RadioML for $k=10$.
+\includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
+\subcaption{RadioML for $k=10$ at SnR = 10dB.
 }
@@
 %     \centering
-%     \includegraphics[width=3.4in]{plots/random-1000000-128-knn-10.png}
+%     \includegraphics[width=3.4in]{plots/random-knn-10.png}
 %     \caption{
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
                     \hline
@@
                     \hline
-                    1 & 13330.275 & 0.954 & 2191.849 & 0.950 & 212.602 & 1.000 & 2009.483 & 1.000 & 2167.365 & 1.000 \\
+                    1   & 1.33~$\times10^{4}$ & 0.954 & 2.19~$\times10^{3}$ & 0.950 & 2.13~$\times10^{2}$  & 1.000 & 2.01~$\times10^{3}$ & 1.000* & 2.17~$\times10^{3}$ & 1.000 \\
                     \hline
-                    2 & 13822.528 & 0.803 & 2118.513 & 0.927 & 106.372 & 1.000 & 939.169 & 1.000 & 1141.225 & 1.000 \\
+                    2   & 1.38~$\times10^{4}$ & 0.803 & 2.12~$\times10^{3}$ & 0.927 & 1.06~$\times10^{2}$  & 1.000 & 9.39~$\times10^{2}$ & 1.000* & 1.14~$\times10^{3}$ & 1.000 \\
                     \hline
-                    4 & 16648.524 & 0.681 & 2043.628 & 0.898 & 53.077 & 1.000 & 461.016 & 0.997 & 982.117 & 1.000 \\
+                    4   & 1.66~$\times10^{4}$ & 0.681 & 2.04~$\times10^{3}$ & 0.898 & 5.31~$\times10^{1}$  & 1.000 & 4.61~$\times10^{2}$ & 0.997  & 9.82~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8 & 16785.981 & 0.525 & 1925.593 & 0.857 & 26.577 & 1.000 & 225.827 & 0.995 & 1179.813 & 1.000 \\
+                    8   & 1.68~$\times10^{4}$ & 0.525 & 1.93~$\times10^{3}$ & 0.857 & 2.66~$\times10^{1}$  & 1.000 & 2.26~$\times10^{2}$ & 0.995  & 1.18~$\times10^{3}$ & 1.000 \\
                     \hline
-                    16 & 18717.175 & 0.494 & 1841.863 & 0.862 & 13.292 & 1.000 & 116.960 & 0.991 & 1201.409 & 1.000 \\
+                    16  & 1.87~$\times10^{4}$ & 0.494 & 1.84~$\times10^{3}$ & 0.862 & 1.33~$\times10^{1}$  & 1.000 & 1.17~$\times10^{2}$ & 0.991  & 1.20~$\times10^{3}$ & 1.000 \\
                     \hline
-                    32 & 15597.778 & 0.542 & 1850.943 & 0.775 & 6.647 & 1.000 & 59.148 & 0.985 & 1157.667 & 1.000 \\
+                    32  & 1.56~$\times10^{4}$ & 0.542 & 1.85~$\times10^{3}$ & 0.775 & 6.65~$\times10^{0}$  & 1.000 & 5.91~$\times10^{1}$ & 0.985  & 1.16~$\times10^{3}$ & 1.000 \\
                     \hline
-                    64 & 14997.277 & 0.378 & 1787.142 & 0.677 & 3.324 & 1.000 & 26.068 & 0.968 & 1102.849 & 1.000 \\
+                    64  & 1.50~$\times10^{4}$ & 0.378 & 1.78~$\times10^{3}$ & 0.677 & 3.32~$\times10^{0}$  & 1.000 & 2.61~$\times10^{1}$ & 0.968  & 1.10~$\times10^{3}$ & 1.000 \\
                     \hline
-                    128 & 14866.019 & 0.357 & 1659.660 & 0.538 & 1.662 & 1.000 & 13.265 & 0.964 & 1039.607 & 1.000 \\
+                    128 & 1.49~$\times10^{4}$ & 0.357 & 1.66~$\times10^{3}$ & 0.538 & 1.66~$\times10^{0}$  & 1.000 & 1.33~$\times10^{1}$ & 0.964  & 1.04~$\times10^{3}$ & 1.000 \\
                     \hline
-                    256 & -- & -- & 1597.264 & 0.592 & 0.831 & 1.000 & 6.650 & 0.962 & 1058.972 & 1.000 \\
+                    256 & --                  & --    & 1.60~$\times10^{3}$ & 0.592 & 8.31~$\times10^{-1}$ & 1.000 & 6.65~$\times10^{0}$ & 0.962  & 1.06~$\times10^{3}$ & 1.000 \\
                     \hline
-                    512 & -- & -- & 1826.362 & 0.581 & 0.416 & 1.000 & 3.563 & 0.949 & 1035.260 & 1.000 \\
+                    512 & --                  & --    & 1.83~$\times10^{3}$ & 0.581 & 4.16~$\times10^{-1}$ & 1.000 & 3.56~$\times10^{0}$ & 0.949  & 1.04~$\times10^{3}$ & 1.000 \\
                     \hline
@@
 We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
-    \label{table:results:ann-fashion}
+    \label{table:results:ann-glove-25}
     \vskip 0.15in
@@
                     \hline
-                    1 & 22755.175 & 0.801 & 2829.338 & 0.835 & 278.176 & 1.000 & 2378.006 & 1.000 & 722.326 & 1.000 \\
+                    1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.78~$\times10^{2}$ & 1.000 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000 \\
                     \hline
-                    2 & 23792.690 & 0.607 & 2701.804 & 0.832 & 139.414 & 1.000 & 1188.398 & 1.000 & 574.516 & 1.000 \\
+                    2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.39~$\times10^{2}$ & 1.000 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000 \\
                     \hline
-                    4 & 24990.630 & 0.443 & 2609.474 & 0.839 & 69.813 & 1.000 & 619.241 & 1.000 & 625.208 & 1.000 \\
+                    4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.98~$\times10^{1}$ & 1.000 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8 & 27843.055 & 0.294 & 2514.357 & 0.834 & 34.878 & 1.000 & 302.683 & 1.000 & 592.699 & 1.000 \\
+                    8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.49~$\times10^{1}$ & 1.000 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000 \\
                     \hline
-                    16 & 31114.062 & 0.213 & 2225.654 & 0.885 & 17.437 & 1.000 & 150.971 & 1.000 & 549.073 & 1.000 \\
+                    16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.74~$\times10^{1}$ & 1.000 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000 \\
                     \hline
-                    32 & 32423.543 & 0.178 & 2008.537 & 0.764 & 8.701 & 1.000 & 74.016 & 0.999 & 475.266 & 1.000 \\
+                    32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 8.70~$\times10^{0}$ & 1.000 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000 \\
                     \hline
-                    64 & -- & -- & 1990.317 & 0.631 & 4.362 & 1.000 & 37.735 & 0.997 & 461.066 & 1.000 \\
+                    64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 4.36~$\times10^{0}$ & 1.000 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000 \\
                     \hline
-                    128 & -- & -- & -- & -- & 2.176 & 1.000 & 18.960 & 0.998 & 440.758 & 1.000 \\
+                    128 & --                  & --    & --                  & --    & 2.18~$\times10^{0}$ & 1.000 & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000 \\
                     \hline
-                    256 & -- & -- & -- & -- & 1.067 & 1.000 & 9.472 & 0.998 & 422.591 & 1.000 \\
+                    256 & --                  & --    & --                  & --    & 1.07~$\times10^{0}$ & 1.000 & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000 \\
                     \hline
@@
 We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
-    \label{table:results:ann-fashion}
+    \label{table:results:ann-sift}
     \vskip 0.15in
@@
                     \hline
-                    1 & 19330.282 & 0.782 & 3977.432 & 0.686 & 72.610 & 1.000 & 697.809 & 1.000 & 551.894 & 1.000 \\
+                    1   & 1.93~$\times10^{4}$ & 0.782 & 3.98~$\times10^{3}$ & 0.686 & 7.26~$\times10^{1}$  & 1.000 & 6.98~$\times10^{2}$ & 1.000* & 5.52~$\times10^{2}$ & 1.000 \\
                     \hline
-                    2 & 20319.372 & 0.552 & 3800.401 & 0.614 & 36.237 & 1.000 & 329.950 & 1.000 & 265.615 & 1.000 \\
+                    2   & 2.03~$\times10^{4}$ & 0.552 & 3.80~$\times10^{3}$ & 0.614 & 3.62~$\times10^{1}$  & 1.000 & 3.30~$\times10^{2}$ & 1.000* & 2.66~$\times10^{2}$ & 1.000 \\
                     \hline
-                    4 & 21765.007 & 0.394 & 3686.606 & 0.637 & 18.152 & 1.000 & 165.175 & 1.000 & 142.545 & 1.000 \\
+                    4   & 2.18~$\times10^{4}$ & 0.394 & 3.69~$\times10^{3}$ & 0.637 & 1.82~$\times10^{1}$  & 1.000 & 1.65~$\times10^{2}$ & 1.000* & 1.43~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8 & 24796.528 & 0.298 & 3582.099 & 0.710 & 9.077 & 1.000 & 77.187 & 1.000 & 79.353 & 1.000 \\
+                    8   & 2.48~$\times10^{4}$ & 0.298 & 3.58~$\times10^{3}$ & 0.710 & 9.08~$\times10^{0}$  & 1.000 & 7.72~$\times10^{1}$ & 1.000* & 7.94~$\times10^{1}$ & 1.000 \\
                     \hline
-                    16 & 26760.717 & 0.210 & 3498.893 & 0.690 & 4.539 & 1.000 & 39.809 & 1.000 & 81.172 & 1.000 \\
+                    16  & 2.68~$\times10^{4}$ & 0.210 & 3.50~$\times10^{3}$ & 0.690 & 4.54~$\times10^{0}$  & 1.000 & 3.98~$\times10^{1}$ & 1.000* & 8.12~$\times10^{1}$ & 1.000 \\
                     \hline
-                    32 & 27522.668 & 0.193 & 3435.058 & 0.639 & 2.269 & 1.000 & 20.910 & 0.999 & 78.124 & 1.000 \\
+                    32  & 2.75~$\times10^{4}$ & 0.193 & 3.44~$\times10^{3}$ & 0.639 & 2.27~$\times10^{0}$  & 1.000 & 2.09~$\times10^{1}$ & 0.999  & 7.81~$\times10^{1}$ & 1.000 \\
                     \hline
-                    64 & -- & -- & 3391.916 & 0.678 & 1.135 & 1.000 & 8.867 & 0.997 & 74.344 & 1.000 \\
+                    64  & --                  & --    & 3.39~$\times10^{3}$ & 0.678 & 1.14~$\times10^{0}$  & 1.000 & 8.87~$\times10^{0}$ & 0.997  & 7.43~$\times10^{1}$ & 1.000 \\
                     \hline
-                    128 & -- & -- & 3363.965 & 0.643 & 0.566 & 1.000 & 4.786 & 0.993 & 68.005 & 1.000 \\
+                    128 & --                  & --    & 3.36~$\times10^{3}$ & 0.643 & 5.66~$\times10^{-1}$ & 1.000 & 4.78~$\times10^{0}$ & 0.993  & 6.80~$\times10^{1}$ & 1.000 \\
                     \hline
@@
 We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
-    \label{table:results:ann-fashion}
+    \label{table:results:ann-random}
     \vskip 0.15in
@@
                     \hline
-                    1 & 11713.738 & 0.060 & 4283.182 & 0.028 & 71.575 & 1.000 & 733.821 & 1.000 & 553.948 & 1.000 \\
+                    1  & 1.17~$\times10^{4}$ & 0.060 & 4.28~$\times10^{3}$ & 0.028 & 7.16~$\times10^{1}$ & 1.000 & 7.34~$\times10^{2}$ & 1.000* & 5.54~$\times10^{2}$ & 1.000 \\
                     \hline
-                    2 & 10061.864 & 0.048 & 4043.999 & 0.021 & 36.157 & 1.000 & 358.083 & 1.000 & 269.314 & 1.000 \\
+                    2  & 1.01~$\times10^{4}$ & 0.048 & 4.04~$\times10^{3}$ & 0.021 & 3.62~$\times10^{1}$ & 1.000 & 3.58~$\times10^{2}$ & 1.000* & 2.69~$\times10^{2}$ & 1.000 \\
                     \hline
-                    4 & 9120.868 & 0.031 & 3639.337 & 0.014 & 18.072 & 1.000 & 189.802 & 1.000 & 137.236 & 1.000 \\
+                    4  & 9.12~$\times10^{3}$ & 0.031 & 3.64~$\times10^{3}$ & 0.014 & 1.81~$\times10^{1}$ & 1.000 & 1.90~$\times10^{2}$ & 1.000* & 1.37~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8 & 8352.077 & 0.022 & 3371.923 & 0.013 & 9.036 & 1.000 & 88.375 & 1.000 & 56.871 & 1.000 \\
+                    8  & 8.35~$\times10^{3}$ & 0.022 & 3.37~$\times10^{3}$ & 0.013 & 9.04~$\times10^{0}$ & 1.000 & 8.84~$\times10^{1}$ & 1.000* & 5.69~$\times10^{1}$ & 1.000 \\
                     \hline
-                    16 & 8250.619 & 0.008 & 3174.156 & 0.006 & 4.531 & 1.000 & 43.613 & 1.000 & 26.136 & 1.000 \\
+                    16 & 8.25~$\times10^{3}$ & 0.008 & 3.17~$\times10^{3}$ & 0.006 & 4.53~$\times10^{0}$ & 1.000 & 4.36~$\times10^{1}$ & 1.000* & 2.61~$\times10^{1}$ & 1.000 \\
                     \hline
-                    32 & -- & -- & 3013.970 & 0.007 & 2.270 & 1.000 & 17.201 & 1.000 & 13.534 & 1.000 \\
+                    32 & --                  & --    & 3.01~$\times10^{3}$ & 0.007 & 2.27~$\times10^{0}$ & 1.000 & 1.72~$\times10^{1}$ & 1.000* & 1.35~$\times10^{1}$ & 1.000 \\
                     % \hline

### 2023-12-21 sections/2-methods.tex
updated through the methods section

@@
 We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
-With this approach, we begin by taking a random sample of no more than 10\% of the data.
-We build a cluster tree with this sample using Algorithm~\ref{alg:methods:clustering:partition}, and select all clusters at some user-specified depth in the tree, with a default depth of 10.
-Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the four algorithms described in Section~\ref{subsec:methods:knn-search}.
-We select the algorithm which had the lowest total search time over all the queries. 
+We start by taking the center of every cluster at a low depth, e.g. 10, in the cluster tree as a query.
+This gives us a small set of queries which are representative of the dataset.
+Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{subsec:methods:knn-search}.
+We select the fastest over all the queries as the optimal algorithm for the dataset and value of $k$.
 Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
@@
 
-We expect CAKES to perform well as the cardinality of datasets increases, and to explore this scaling behavior, we augmented datasets in the ANN-benchmarks suite by adding random points within a 
-small noise threshold of existing points; this technique has been used in a number of bioinformatics contexts~\cite{kumar2009augmented, kumar2010recognition, daniels2012smurflite, velasco2018combining, lu2020evolution}.
-This allowed us to isolate the effect of \emph{cardinality} on the performance of CAKES
-from that of other factors such as dimensionality, metric, or topological structure of the dataset.
+Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and for that performance to scale sub-linearly with the cardinality of the dataset.
+To test this hypothesis, we use some datasets from the ANN-benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to generate similar datasets with much higher cardinalities.
+We do the same with a large random dataset of uniformly distributed points in a hypercube.
+We then compare the performance of CAKES to that of other algorithms on the original datasets and the synthetically augmented datasets.
 
-To elaborate, we start with an original dataset $X$ from the ANN-benchmark suite, a user-specified noise tolerance $\epsilon$, and a mutliplier $m$.
-For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon$ of $x$. To construct a synthetic data point $x'$, we choose $d$ random numbers 
-$r_1, r_2, \cdots , r_d$ in the interval $[-\tfrac{\epsilon}{\sqrt{d}}, \tfrac{\epsilon}{\sqrt{d}}]$ and add them componentwise to the original datapoint. 
-Since $\sqrt{d \cdot (\tfrac{\epsilon}{\sqrt{d}})^2 } = \sqrt{d \cdot \tfrac{\epsilon^2}{d}}  = \sqrt{\epsilon^2} = \epsilon$, this ensures that any synthetic point is within a distance of $\epsilon$ from a real point in the original dataset.
+To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
+Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level and $m$ be a user-specified integer multiplier.
+For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the euclidean distance from $x$ to the origin.
+We construct a random vector $r$ of $d$ dimensions in the hypersphere of radius $\epsilon$ centered at the origin.
+We then add $r$ to $x$ to get a new point $x'$.
+Since $||r|| = \epsilon$, we have that $||x - x'|| \leq \epsilon$, i.e. $x'$ is within a distance $\epsilon$ of $x$.
+We repeat this process $m$ times for each $x \in X$ to get a new dataset $X'$.
+
+This augmentation process preserves the topological structure of the original dataset, but increases its cardinality by a factor of $m$, allowing us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.

### 2023-12-21 sections/1-introduction.tex
added related-works subsection in intro

@@
 
-Recent search algorithms to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear} such as MRPT~\cite{hyvonen2016fast}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, Inverted File indexing~\cite{johnson2019billion}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+This paper focuses on three novel algorithms for exact $k$-NN search.
+We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
+Finally, we provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy} on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
 
-TODO: Paragraph summarizing LSH.
+\subsection{Related Works}
+\label{subsec:intoduction:related-works}
+
+Recent search algorithms to tackling the exponential growth of data include hierarchical navigable small world networks (HNSW)~\cite{Malkov2016EfficientAR}, Inverted File indexing (FAISS-IVF)~\cite{johnson2019billion}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+
+\subsubsection{HNSW}
+\label{subsubsec:introduction:related-works:hnsw}
 
@@
 
+\subsubsection{FAISS-IVF}
+\label{subsubsec:introduction:related-works:faiss-ivf}
+
 TODO: Paragraph summarizing FAISS-IVF.
+
+\subsubsection{ANNOY}
+\label{subsubsec:introduction:related-works:annoy}
+
+TODO: Paragraph summarizing ANNOY.
+
+\subsubsection{Entropy-Scaling Search}
+\label{subsubsec:introduction:related-works:entropy-scaling-search}
 
@@
 Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-nearest neighbors search.
-
-This paper focuses on three novel algorithms for exact $k$-NN search.
-We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
-Finally, we provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy} on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.

### 2023-12-21 sections/1-introduction.tex
updated intro

@@
 Radio frequency data offers another example of massive datasets. 
-The RadioML~\cite{oshea2018radioml} dataset contains 2,555,504 samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
+The RadioML~\cite{oshea2018radioml} dataset contains approximately 2.55 million samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
 
 The field of astronomy has also seen an influx of large datasets.
+% Najib: Add exact dimensions of the SDSS dataset from the CHESS paper.
 The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
 
-In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of computing performance improvements predicted by Moore's Law~\cite{brescia2012extracting}.
+In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{brescia2012extracting}.
 This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
@@
 
-Many researchers are especially interested in similarity search on these datasets. 
+Many researchers are especially interested in similarity search on such datasets. 
 Similarity search enables a variety of applications, including recommendation~\cite{annoy} and classification systems~\cite{suyanto2022knnclassifier}.
@@
 
-We now define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN). 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search 
-aims to find the $k$ closest points to $q$ in $ \textbf{X}$.
-On the other hand, $\rho$-nearest neighbors search aims to find all points in $\textbf{X}$ that are at most a distance $\rho$ from $q$.
+Given some measure of similarity between data points, e.g.
+a distance function, we define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
+$k$-NN search aims to find the $k$ most similar points to a query point, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query point.
 
-Previous works have used the term \emph{approximate} search to refer to $\rho$-nearest neighbor search, but in this paper, 
-we reserve the term \emph{approximate} for search algorithms which do not exhibit perfect recall.
-When we refer to an \emph{exact} search algorithm, we mean an algorithm which exhibits perfect recall when compared to a na\"{i}ve linear search.
+Previous works have used the term \emph{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \emph{approximate} for search algorithms which do not exhibit perfect recall when compared to a na\"{i}ve linear search.
+In contrast, an \emph{exact} search algorithm exhibits perfect recall.
 
-$k$-nearest neighbors ($k$-NN) search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}. 
+$k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
 Na\"{i}ve implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets because their cardinalities are growing exponentially.
 
-While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in large datasets.
+While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
 Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \emph{exact} search remains.
@@
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is some evidence that distance functions which are non-metrics, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
-this may suggest that approximate $k$-NN search could exhibit suboptimal accuracy.
+Further, there is some evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
+this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.
 
 This paper focuses on three novel algorithms for exact $k$-NN search.
-We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
-Finally, we provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy} on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
+We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS, as well as improved genericity across distance functions.
+We provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
+We also benchmark CAKES on a genomic dataset, the Silva 18S dataset [citation needed], and a radio frequency dataset, the RadioML dataset [citation needed].
 
@@
 Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of a dataset, rather than its cardinality.
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest neighbors search from a flat clustering approach to a tree-based hierarchical clustering approach.
-In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbors Entropy-scaling Search).
-CLAM (Clustered Learning of Approximate Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm.
-Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-nearest neighbors search.
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, rather than its cardinality.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
+In this paper, we introduce CAKES (CLAM-Accelerated $K$-nn Entropy-scaling Search).
+CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
+Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search and improves the performance of CHESS's $\rho$-NN search.

### 2023-12-21 sections/1-introduction.tex
some more edits

@@
 % Najib: Add exact dimensions of the SDSS dataset from the CHESS paper.
-The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions.
+The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 8,575 dimensions.
 

### 2023-12-21 sections/2-methods.tex
some more edits

@@
 
-In this manuscript, we are primarily concerned with $k$-nearest neighbors search in a finite-dimensional space.
+In this manuscript, we are primarily concerned with $k$-NN search in a finite-dimensional space.
 Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the genome of an organism, the neural-network embedding of an image, etc.
@@
 We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
+% Najib: I don't like the ``have an identity'' wording here. The Measure Theory A-Grader will have to prove her chops for this one.
 We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and have an identity (i.e. $f(x, y) = 0 \iff x = y$).
-In addition to these constraints, if the distance function obeys the triangle inequality, i.e., it is a \emph{distance metric}, then we guarantee that CAKES has perfect recall. 
+If, in addition to these constraints, the distance function obeys the triangle inequality (i.e.
+$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then the distance function is also a \emph{distance metric}.
+For example, Euclidean, Levenshtein and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.
+Under distance metrics, all search algorithms in CAKES are exact.
+
 The choice of distance function varies by dataset.
-For example, with neural-network embeddings, one could use Euclidean (L2) distance or Cosine distance.
-With biological or string data, Levenshtein or Hamming distances are more appropriate.
-With time-series data, one could use Dynamic Time Warping (DTW).
+For example, with neural-network embeddings, one could use Euclidean (L2) or Cosine distance.
+With genomic or proteomic data, Levenshtein or Hamming distances are more appropriate.
+With time-series data, one could use Dynamic Time Warping (DTW) distance.
 
-Some of our search algorithms rely on the \emph{local fractal dimension} (LFD) at some point and length scale in the dataset.
-We define LFD as:
+% Najib: Reword the following sentence.
+CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, i.e., high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space, and exploits this low \emph{local fractal dimension} (LFD) to improve the performance of its search.
+We define LFD at some length scale around a point in the dataset as:
 
 \begin{equation}
-    \frac{\text{log}(\frac{|B_X(q, r_1)|}{|B_X(q, r_2)|})}{\text{log}(\frac{r_1}{r_2}) }
+    \frac{\text{log} \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
     \label{eq:methods:lfd-original}
@@
 
-where $B_X(q, r)$ is the set of points contained in a ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
+where $B_X(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
 We stress that this concept differs from the \emph{embedding dimension} of a dataset.
@@
 However, due to physical constraints (namely, the laws of physics that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional manifold within the 8,575-dimensional embedding space.
-The LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
-This is the \emph{manifold hypothesis}~\cite{fefferman2016testing}, i.e.
-high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
+LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
+Figure~\ref{fig:results:lfd-plots} illustrates this concept, showing how real datasets obey the manifold hypothesis.
 
@@
 Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
-We estimate LFD at the scale of the cluster radius and half that radius. 
-This allows us to rewrite Equation~\ref{eq:methods:lfd-original} as:
+Hereafter, when we refer to LFD, it is estimated at the scale of the cluster radius and half that radius. 
+This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as:
 
 \begin{equation} 
-    \log_2 \left( \frac{|B_X(q, r)}{|B_X(q, r/2)|} \right)
+    \log_2 \left( \frac{|B_X(q, r)|}{|B_X(q, r/2)|} \right)
     \label{eq:methods:lfd-simplified}
@@
 
-In~\cite{yu2015entropy}, metric entropy for a given cluster radius $r_c$ was defined as the number of clusters of a uniform radius $r_c$ needed to cover all data.
-In this paper, we define the metric entropy $\mathcal{N}_{\hat{r_c}}(X)$ of a data set $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree.
-
-% Here, we use a hierarchical, divisive clustering approach, but with early stopping criteria; clusters which satisfy a user-specified stopping criterion (e.g.
-a specified cardinality or radius) are not further split.
-% Since we frame the asymptotic complexity of $\rho$-NN search in terms of the number of leaf clusters, the \emph{metric entropy} is best thought of in terms of the number of leaf clusters.
+In~\cite{yu2015entropy}, metric entropy for a given cluster radius $r$ was defined as the number of clusters of a uniform radius $r$ needed to cover all data.
+In this paper, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree.
 
 We start by performing a divisive hierarchical clustering on the dataset using CLAM.
-The procedure is similar to what was outlined in CHESS, but with the following improvements:
+The procedure is similar to that outlined in CHESS, but with the following improvements:
 better selection of poles for partitioning (see Algorithm~\ref{alg:methods:clustering:partition}) and depth-first reordering of the dataset (see Section~\ref{subsubsec:methods:clustering:dataset-depth-first-reordering}).
@@
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. 
-While we sometimes use Euclidean notions, such as voids and volumes, to talk about the geometric and topological properties of clusters and of the manifold, CLAM does not rely on such notions; 
-they serve only as a convenient, intuitive way to talk about the underlying mathematics.
+While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of clusters and of the manifold, CLAM does not rely on such notions;
+they serve merely as a convenient and intuitive way to discuss the underlying mathematics.
 

### 2023-12-24 sections/4-results.tex
moved Morgan's changes
shortened table captions, bolded dataset names

@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall of CAKES vs.
-other methods on the fashion-mnist dataset.
-We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \caption{Runtime performance (queries per second) and recall on the \textbf{fashion-mnist} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
     \label{table:results:ann-fashion}
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall of CAKES vs.
-other methods on the glove-25 dataset.
-We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \caption{Runtime performance (queries per second) and recall on the \textbf{glove-25} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
     \label{table:results:ann-glove-25}
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall of CAKES vs.
-other methods on the sift dataset.
-We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \caption{Runtime performance (queries per second) and recall on the \textbf{sift} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
     \label{table:results:ann-sift}
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall of CAKES vs.
-other methods on a random dataset.
-We use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.}
+    \caption{Runtime performance (queries per second) and recall on a \textbf{random} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
     \label{table:results:ann-random}
@@
 % \end{table*}
+
+Tables~\ref{table:results:fashion}, ~\ref{table:results:glove-25}, ~\ref{table:results:sift}, and ~\ref{table:results:random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively.
+When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.
+
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
+We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
+Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
+Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers. 
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
+We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
+Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
+Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
+
+With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high cardinality multipliers.
+Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$.
+We also see that while hnsw and annoy have higher throughput than CAKES for all multipliers, CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-glove-25}.
+For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.294 and 0.857 respectively, while CAKES exhibits perfect recall.
+On this dataset, Repeated $\rho$-NN, plotted in green, is consistently the best-performing algorithm. {\color{red} Revive lfd violin plots to add explanation for this.}
+
+In Figure~\ref{fig:results:sift-scaling}, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search beginning at a multiplier of approximately 7.
+For this dataset, Sieve, plotted in red, is the highest throughput algorithm for all multipliers greater than 7.
+
+Figure \ref{fig:results:random-scaling} displays results for a random datset with the same dimensionality and cardinality as Sift (i.e., 128 and 1,000,000 respectively). 
+These results illustrate the fact that \emph{ceteris paribus}, absent a manifold structure in the dataset, CAKES's algorithms do not begin to outperform linear search until the cardinality multiplier is around 10.
+Repeated $\rho$-NN shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
+This is unsurprising, given that Repeated $\rho$-NN relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
+With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, Repeated $\rho$-NN will iterate for longer before finding this correct radius.
+
+Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
+This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities.
+% We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
+% While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
+Finally, we reiterate that the variation in performance of our algorithms across different datasets and cardinalities support our use of an autotuning function to select the best choice of algorithm. 
+% Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
+
+% 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/fashion-mnist-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on fashion-mnist for $k=10$.
+%     }
+%     \label{fig:results:fashion-mnist-scaling}
+% \end{figure}
+
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{images/result_plots/deep-image_10_scaling.png}
+%     \caption{
+%         Deep-image Scaling
+%     }
+%     \label{fig:results:deep-image-scaling}
+% \end{figure}
+
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/glove-25-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on glove-25 for $k=10$.  }
+%     \label{fig:results:glove-25-scaling}
+% \end{figure}
+
+% 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/sift-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on sift for $k=10$.    }
+%     \label{fig:results:sift-scaling}
+% \end{figure}
+
+% We need to redo this plot so that we do not actually augment random datasets but instead take larger and larger 
+% random datasets
+% 
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=3.4in]{plots/random-1000000-128-knn-10.png}
+%     \caption{
+%         Algorithm performance by cardinality multiplier on a random dataset for $k=10$.
+%     }
+%     \label{fig:results:random-scaling}
+% \end{figure}
+
+% We did not run any of the competitor's algorithms.
+% We lifted the reported numbers from the ann-benchmarks site and their interactive plots.
+% We made sure to use the same AWS instance as they did so it is a fair comparison.
+
+% Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarks against FAISS, bruteforce-blas, and HNSW.
+% The throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
+
+% We report the highest throughput value for each algorithm at a recall value of 1.0.
+% We report the mean and median speedup factor of CAKES over each algorithm.
+% For $k= 10$, we report a mean $7,654 \times$ (median $1,200 \times$) speedup over faiss-ivf, a mean $11,325 \times$ (median $2,821 \times$) speedup over bruteforce-blas, and a mean $485 \times$ (median $417 \times$) speedup over HNSW.
+% For $k=100$, we report a mean $4,479 \times$ (median $3,456 \times$) speedup over faiss-ivf, a mean $16,010 \times$ (median $13,548 \times$) speedup over bruteforce-blas, and a mean $403 \times$ (median $372 \times$) speedup over HNSW.
+
+% For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
+
+% We note that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent;
+% in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
+% As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP}, CAKES's strong performance with this dataset is particularly encouraging.
+% Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.

### 2023-12-24 sections/2-methods.tex

@@
 For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
-Using these distances, we compute the \emph{geometric median} of this sample; 
-i.e., the point which minimizes the sum of distances to all other points in the sample.
+Using these distances, we compute the \emph{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
 This geometric median is the \emph{center} of $C$.
@@
 Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
-i.e., for a given $k$, find the $k$ closest points to $q$ in $\textbf{X}$.
+in other words, for a given $k$, $k$-NN finds the $k$ closest points to $q$ in $\textbf{X}$.
 We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
@@
 
-The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data, i.e.
-it arises from the same generative process as the rest of the data.
+The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data, or in other words, it arises from the same generative process as the rest of the data.
 Given the uses of $k$-NN search in practice, this assumption is reasonable.
@@
 We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{tree-search} and \emph{leaf-search} separately.
-Tree-search refers to the process of identifying clusters which have overlap with the query ball, i.e., which might contain one of the $k$ nearest neighbors. 
+Tree-search refers to the process of identifying clusters which have overlap with the query ball, or in other words, clusters which might contain one of the $k$ nearest neighbors. 
 Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by tree-search.
@@
 During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{subsec:methods:the-search-problem}, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.
 
-We then use the QuickSelect algorithm to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$ while accounting for multiplicities, i.e.
-we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
+% TODO (Morgan): I don't like "kth smallest $\delta^{+}$ while accounting for multiplicities" 
+We then use the QuickSelect algorithm to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$ while accounting for multiplicities; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
 
@@
 We continue this process until $Q$ no longer contains any clusters.
-We just use the QuickSelect algorithm once last time to reorder $Q$, to find the $k^{th}$ element, and to return the $k$ nearest neighbors.
+We then use the QuickSelect algorithm once last time to reorder $Q$, to find the $k^{th}$ element, and to return the $k$ nearest neighbors.
 
@@
 This algorithm is described in Algorithm~\ref{alg:methods:greedy-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.
-when the top priority element in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$ (i.e., when the top priority element in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$).
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
@@
 We then add $r$ to $x$ to get a new point $x'$.
-Since $||r|| = \epsilon$, we have that $||x - x'|| \leq \epsilon$, i.e. $x'$ is within a distance $\epsilon$ of $x$.
+Since $||r|| = \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
 We repeat this process $m$ times for each $x \in X$ to get a new dataset $X'$.

### 2023-12-24 sections/2-methods.tex
more minor changes to methods

@@
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD. 
-Intuitively, the factor by which we increase the radius should also be \emph{inversely} related to the number of points found so far. 
+Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
 Additionally, when the LFD at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.
 In such a region, a small increase in the radius would likely encounter unoccupied space, so a larger radial increase is needed.
-Thus, the factor of radius increase should be \emph{inversely} related to the LFD.
+Thus, the factor of radius increase should also be \emph{inversely} related to the LFD.
 
@@
 
-Based on the assumption that the LFD at the query point is not significantly different from that of nearby clusters, the factor from Equation~\ref{eq:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
+Based on the assumption that the LFD at the query point does not differ significantly from that of nearby clusters, the factor from Equation~\ref{eq:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
 one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
-Since this is a constant factor, it does not affect the asymptotic complexity of Repeated $\rho$-NN.
+Since this is a constant factor, complexity of tree search for Repeated $\rho$-NN is the same as that of $\rho$-NN.
 
-To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in S} |C|$, the number of points we examine.
+To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in S} |C|$, the number of points in the clusters returned by tree search.
+Since we must examine each of these points, time complexity of leaf-search is linear in $\sum_{C \in S} |C|$. 
 Let $\rho_k$ be the distance from the query to the $k$th nearest neighbor.
@@
 We let $\hat{d}$ be the mean LFD of the clusters in $S$.
-While ordinarily we compute LDS by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \emph{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{C}$, of clusters in $S$ at a radius equal to the mean of their radii, $\hat{r}$.
+While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \emph{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{C}$, of clusters in $S$ at a radius equal to the mean of their radii, $\hat{r}$.
 

### 2023-12-24 sections/2-methods.tex
one more minor change

@@
 During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{subsec:methods:the-search-problem}, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{subsec:methods:the-search-problem}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.
 

### 2023-12-24 sections/4-results.tex
results for fashion-mnist

@@
 
-Figures~\ref{fig:results:fashion-mnist-lfd} show the local fractal dimension of.... (TODO fill this in)
+Since the time complexity of CAKES algorithms scales with the local fractal dimension of the dataset, we examine the local fractal dimension of each dataset we benchmarked CAKES algorithms on.
 
-%% LFD plots
+The figures in~\ref{fig:results:lfd-plots} show the LFD of fashion-mnist, glove-25, sift, a random dataset, Silva 18S, and RadioML.
+The horizontal axis denotes the depth in the cluster tree, and the vertical axis denotes the LFD (as calculated by Equation~\ref{eq:methods:lfd-simplified}) of clusters at that depth.
+We plot lines for the 0th, 5th, 25th, 50th, 75th, 95th, and 100th percentiles of LFD.
+So that these plots best reflect the distribution of LFDs across the entire \emph{dataset}, we compute percentiles counting each cluster as many times as its cardinality. 
+In other words, if, for some dataset, the 95th percentile of LFD at depth 40 is 1.5, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 1.5.
+
+Figure~\ref{fig:results:fashion-mnist-lfd} shows the local fractal dimension by depth for the fashion-mnist dataset. 
+We observe that up until depth 5, the LFD at most locations is low, as the 95th percentile (orange plotted line) is less than 4, and the 50th percentile is just above 2.
+For depths 15 through 25, we observe that the LFD is higher, as the 95th percentile (orange plotted line) of LFD is slightly less than 6, and the 50th percentile (red plotted line) is near 3.
+Finally, for depths 25 through the maximum depth, we observe that the LFD is lower again, as the 95th percentile is between 3 and 4, and the 50th percentile is less than 2.
 

### 2023-12-24 sections/4-results.tex
lfd results for glove

@@
 \label{sec:results}
+
+% These are going to be boring; saving the interesting discussion for the discussion section
 
@@
 Finally, for depths 25 through the maximum depth, we observe that the LFD is lower again, as the 95th percentile is between 3 and 4, and the 50th percentile is less than 2.
+
+Figure~\ref{fig:results:glove-25-lfd} shows the local fractal dimension by depth for the glove-25 dataset. 
+Relative to fashion-mnist, percentile lines for glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
+The 95th percentile of LFD is less than 4 for all depths, and the 50th percentile is less than 2 for all depths. 
+Before depth 25, the 95th percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths. 
+The 50th percentile hovers near 1.5 for all depths.
 

### 2023-12-24 sections/4-results.tex
lfd results for sift

@@
 % These are going to be boring; saving the interesting discussion for the discussion section
+% I think I should vary the wording more but just getting down the key observations for now
 
@@
 Figure~\ref{fig:results:fashion-mnist-lfd} shows the local fractal dimension by depth for the fashion-mnist dataset. 
-We observe that up until depth 5, the LFD at most locations is low, as the 95th percentile (orange plotted line) is less than 4, and the 50th percentile is just above 2.
+We observe that until about depth 5, the LFD at most locations is low, as the 95th percentile (orange plotted line) is less than 4, and the 50th percentile is just above 2.
 For depths 15 through 25, we observe that the LFD is higher, as the 95th percentile (orange plotted line) of LFD is slightly less than 6, and the 50th percentile (red plotted line) is near 3.
@@
 The 50th percentile hovers near 1.5 for all depths.
+
+Figure~\ref{fig:results:sift-lfd} shows the local fractal dimension by depth for the sift dataset. 
+Until about depth 5, the 95th percentile is between 2 and 6. From about depth 5 through about depth 20, the 95th percentile is greater than 6, even exceeding 8 near depth 10. 
+From about depth 20 onward, the 95th percentile is less than 6, and from about depth 30 onward, it is between 3 and 4. 
+The 50th percentile reaches its peak of about 5 at around depth 15, but hovers near 2 after about depth 30.
 

### 2023-12-24 sections/3-datasets.tex
changes to datasets section
some restructuing of datasets and results

@@
 Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
-
-In addition to benchmarks on datasets from the ANN Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{subsec:methods:synthetic-data}.
-In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the multiplier $m$ of new points increases.
-While we explored the scaling behavior of CAKES with augmented versions of each dataset in \ref{table:datasets:summary}, we give an in-depth discussion in \ref{sec:results} of results for scaling behavior on fashion-mnist, glove-100, gist, and a purely random dataset.
 
@@
 
+\subsection{Random Datasets and Synthetic Augmentations}
+\label{subsec:random-datasets}
+
+In addition to benchmarks on datasets from the ANN Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{subsec:methods:synthetic-data}.
+In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (refered to as ``scale'' in Table~\ref{table:results:ann-fashion} {\color{red} TODO: merge the tables into one and confirm this reference is still correct after the merge}) increases.
+
+We also benchmarked on a purely random dataset with the same cardinality and dimensionality as sift: 1,000,000 points in 128 dimensions. 
+This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of CAKES's algorithms.
+
 \subsection{Silva}
 \label{subsec:silva}
-
-Add section about Silva.
+Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
+We use Levenshtein distance with this dataset.
+% I think this needs another sentence but I'm not sure what else to say
 
@@
 Each sample provided in RadioML can be thought of as an array of 2048 floating point values for a single signal capture with labels for SNR and modulation mode.
+We use Dynamic Time Warping distance with this dataset.
 

### 2023-12-24 sections/4-results.tex

--- 2023-12-24 sections/4-results.tex
@@
 
+% These are going to be boring; saving the interesting discussion for the discussion section
+
+\subsection{Local Fractal Dimension of Datasets}
+\label{subsec:lfd-results}
+
 Since the time complexity of CAKES algorithms scales with the local fractal dimension of the dataset, we examine the local fractal dimension of each dataset we benchmarked CAKES algorithms on.
 
-The figures in~\ref{fig:results:lfd-plots} show the LFD of fashion-mnist, glove-25, sift, a random dataset, Silva 18S, and RadioML.
+Figure~\ref{fig:results:lfd-plots} plots LFD by depth for fashion-mnist, glove-25, sift, a random dataset, Silva 18S, and RadioML.
 The horizontal axis denotes the depth in the cluster tree, and the vertical axis denotes the LFD (as calculated by Equation~\ref{eq:methods:lfd-simplified}) of clusters at that depth.
 We plot lines for the 0th, 5th, 25th, 50th, 75th, 95th, and 100th percentiles of LFD.
-So that these plots best reflect the distribution of LFDs across the entire \emph{dataset}, we compute percentiles counting each cluster as many times as its cardinality. 
-In other words, if, for some dataset, the 95th percentile of LFD at depth 40 is 1.5, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 1.5.
+In order to have the plots to best reflect the distribution of LFDs across the entire \emph{dataset}, we compute percentiles counting each cluster as many times as its cardinality. 
+In other words, if, for some dataset, the 95th percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
 
-Figure~\ref{fig:results:fashion-mnist-lfd} shows the local fractal dimension by depth for the fashion-mnist dataset. 
-We observe that up until depth 5, the LFD at most locations is low, as the 95th percentile (orange plotted line) is less than 4, and the 50th percentile is just above 2.
-For depths 15 through 25, we observe that the LFD is higher, as the 95th percentile (orange plotted line) of LFD is slightly less than 6, and the 50th percentile (red plotted line) is near 3.
-Finally, for depths 25 through the maximum depth, we observe that the LFD is lower again, as the 95th percentile is between 3 and 4, and the 50th percentile is less than 2.
+Figure~\ref{fig:results:fashion-mnist-lfd} shows the LFD by depth for fashion-mnist. 
+We observe that until about depth 5, the LFD is low, as the 95th percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
+For depths 15 through 25, we observe that the LFD increases, with the 95th percentile (orange plotted line) slightly less than 6, and the median near 3.
+Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95th percentile is between 3 and 4, and the median is less than 2.
+
+Relative to fashion-mnist, glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}. 
+Percentile lines for glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
+The 95th percentile of LFD is less than 4 for all depths, and the median is less than 2 for all depths. 
+Before depth 25, the 95th percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths. 
+The median hovers near 1.5 for all depths.
+
+Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for sift. 
+Until about depth 5, the 95th percentile is between 2 and 6. From about depth 5 through about depth 20, the 95th percentile is greater than 6, even exceeding 8 near depth 10. 
+From about depth 20 onward, the 95th percentile is less than 6, and from about depth 30 onward, it is between 3 and 4. 
+The median reaches its peak of about 5 at around depth 15, but hovers near 2 after about depth 30.
+
+In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift. 
+We observe that the LFD starts as high as 20 (for all data) at depth 0. As shown by the needlepoint shape of the plot, the LFD for all clusters 
+is about the same from depth 0 through depth 5. 
+After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth. 
+The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
+
+Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently show LFD.
+The 95th percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 130. 
+The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 130.
+
+Relative to the other datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD. 
+Notably however, for the frist approximately 15 depths, the 95th percentile and median LFD are very close to 0. 
+Both percentiles increase sharply after depth 20, nearly reaching 16. 
+LFD remains high even as depth increases, with the 95th percentile fluctuating between about 6 and 15 for depths about 20-40. 
+It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60. 
+The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.
 
@@
 %% end LFD plots
+
+\subsection{Scaling Behavior and Recall}
+\label{subsec:scaling-behavior-results}
 

### 2023-12-24 sections/4-results.tex
finished scaling results for ann and random and moved discussion-y things out of results

@@
 
-Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
-Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
+Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
 The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
@@
 Both axes are on a logarithmic scale. 
-In this section, we discuss results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section.
-In these plots, we label the recall at each scale for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall at all scales.
-
-Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively.
-When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider
-only three decimal places.
-Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
-Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
-The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
-The vertical axis denotes throughput in queries per second.
-Both axes are on a logarithmic scale.
-In this section, we discuss results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section.
+In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section. 
 In these plots, we label the recall at each scale for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall at all scales.
@@
 Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
-We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
-Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
-Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
 
@@
 Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$. 
-We also see that while hnsw and annoy have higher throughput than CAKES for all multipliers, CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-glove-25}.
-For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.294 and 0.857 respectively, while CAKES exhibits perfect recall.
-On this dataset, Repeated $\rho$-NN, plotted in green, is consistently the best-performing algorithm. {\color{red} Revive lfd violin plots to add explanation for this.}
+We also see that while HNSW and Annoy have higher throughput than CAKES for all multipliers, CAKES exhibits better recall at all multupliers, as shown in Table~\ref{table:results:ann-glove-25}.
+In particular, at a cardinality multiplier as low as two, HNSW and Annoy have recall as low as 0.607 and 0.832 respectively, while CAKES exhibits perfect recall.
+On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN, plotted in green, is consistently the best-performing CAKES algorithm.
 
-In Figure~\ref{fig:results:sift-scaling}, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search beginning at a multiplier of approximately 7.
-For this dataset, Sieve, plotted in red, is the highest throughput algorithm for all multipliers greater than 7.
+In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the sift dataset, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search at a multiplier of approximately 7.
+Again, we observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$. 
+For this dataset, Sieve, plotted in red, is the highest throughput CAKES algorithm for all multipliers greater than 7. 
+Similar to the results with fashion-mnist and glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES for all multipliers, CAKES has better recall. 
+Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and Annoy have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
 
-Figure \ref{fig:results:random-scaling} displays results for a random datset with the same dimensionality and cardinality as Sift (i.e., 128 and 1,000,000 respectively). 
-These results illustrate the fact that \emph{ceteris paribus}, absent a manifold structure in the dataset, CAKES's algorithms do not begin to outperform linear search until the cardinality multiplier is around 10.
-Repeated $\rho$-NN shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
-This is unsurprising, given that Repeated $\rho$-NN relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
-With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, Repeated $\rho$-NN will iterate for longer before finding this correct radius.
+With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibt the worst throughput out of all algorithms, at all cardinality multipliers.
+As with the real datasets, HNSW and Annoy exhibit the highest throughput of any of the algorithms. 
+As with the real datasets, CAKES exhibits perfect recall at all cardinality multipliers. 
+HNSW and Annoy exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
+in particular, with a cardinality multiplier of 1, HNSW and Annoy have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.
 
-Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
-This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities.
-% We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
-% While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
-Finally, we reiterate that the variation in performance of our algorithms across different datasets and cardinalities support our use of an autotuning function to select the best choice of algorithm. 
-% Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
+{\color{red} TODO: Add Silva and RadioML results.}
 
@@
 \end{figure}
-
-% 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/fashion-mnist-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on fashion-mnist for $k=10$.
-%     }
-%     \label{fig:results:fashion-mnist-scaling}
-% \end{figure}
-
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{images/result_plots/deep-image_10_scaling.png}
-%     \caption{
-%         Deep-image Scaling
-%     }
-%     \label{fig:results:deep-image-scaling}
-% \end{figure}
-
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/glove-25-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on glove-25 for $k=10$.  }
-%     \label{fig:results:glove-25-scaling}
-% \end{figure}
-
-% 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/sift-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on sift for $k=10$.    }
-%     \label{fig:results:sift-scaling}
-% \end{figure}
-
-% We need to redo this plot so that we do not actually augment random datasets but instead take larger and larger 
-% random datasets
-% 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/random-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on a random dataset for $k=10$.
-%     }
-%     \label{fig:results:random-scaling}
-% \end{figure}
-
-% We did not run any of the competitor's algorithms.
-% We lifted the reported numbers from the ann-benchmarks site and their interactive plots.
-% We made sure to use the same AWS instance as they did so it is a fair comparison.
-
-% Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarks against FAISS, bruteforce-blas, and HNSW.
-% The throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
-
-% We report the highest throughput value for each algorithm at a recall value of 1.0.
-% We report the mean and median speedup factor of CAKES over each algorithm.
-% For $k= 10$, we report a mean $7,654 \times$ (median $1,200 \times$) speedup over faiss-ivf, a mean $11,325 \times$ (median $2,821 \times$) speedup over bruteforce-blas, and a mean $485 \times$ (median $417 \times$) speedup over HNSW.
-% For $k=100$, we report a mean $4,479 \times$ (median $3,456 \times$) speedup over faiss-ivf, a mean $16,010 \times$ (median $13,548 \times$) speedup over bruteforce-blas, and a mean $403 \times$ (median $372 \times$) speedup over HNSW.
-
-% For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
-
-% We note that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent;
-% in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
-% As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP}, CAKES's strong performance with this dataset is particularly encouraging.
-% Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.
 
@@
 \end{table*}
-
-% \begin{table*}[!t]
-%     % \renewcommand{\arraystretch}{1.15}
-%     \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=10$}
-%     \label{table:results:ann-alt-10}
-%     \vskip 0.15in
-%     \begin{center}
-%         \begin{small}
-%             \begin{sc}
-%                 \begin{tabular}{|l|l|l|l|l|}
-%                     \hline
-%                     \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
-%                     \hline
-%                     deep-image       & 232.1                     & 37.2      \\
-%                     \hline
-%                     mnist            & 129,300                   & 101.9      \\
-%                     \hline
-%                     glove-50         & 10,950                    & 12.28      \\
-%                     \hline 
-%                     glove-200        & 2,290                     & 8.55     \\
-%                     \hline
-%                     lastfm           & 13,950                    & 4.55           \\
-%                     \hline
-%                 \end{tabular}
-%             \end{sc}
-%         \end{small}
-%     \end{center}
-%     \vskip -0.1in
-% \end{table*}
-
-% \begin{table*}[!t]
-%     % \renewcommand{\arraystretch}{1.15}
-%     \caption{Runtime performance (queries per second) of CAKES and Speedup Factor over Naive Linear Search, $k=100$}
-%     \label{table:results:ann-alt-100}
-%     \vskip 0.15in
-%     \begin{center}
-%         \begin{small}
-%             \begin{sc}
-%                 \begin{tabular}{|l|l|l|l|l|l|}
-%                     \hline
-%                     \textbf{Dataset} & \textbf{CAKES Throughput} & \textbf{Speedup Factor over Linear} \\
-%                     \hline
-%                     deep-image        & 157.3                    & 25.74                               \\                            
-%                     \hline
-%                     mnist             & 71,090                   & 56.63                               \\
-%                     \hline
-%                     glove-50          & 8,867                    & 9.84                                \\
-%                     \hline 
-%                     glove-200         & 2,264                    & 8.49                                \\
-%                     \hline
-%                     lastfm            & 13,950                   & 4.55                                \\
-%                     \hline
-%                 \end{tabular}
-%             \end{sc}
-%         \end{small}
-%     \end{center}
-%     \vskip -0.1in
-% \end{table*}
-
-Tables~\ref{table:results:fashion}, ~\ref{table:results:glove-25}, ~\ref{table:results:sift}, and ~\ref{table:results:random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively.
-When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.
-
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
-We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
-Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
-Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers. 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
-We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
-Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
-Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
-
-With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high cardinality multipliers.
-Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$.
-We also see that while hnsw and annoy have higher throughput than CAKES for all multipliers, CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-glove-25}.
-For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.294 and 0.857 respectively, while CAKES exhibits perfect recall.
-On this dataset, Repeated $\rho$-NN, plotted in green, is consistently the best-performing algorithm. {\color{red} Revive lfd violin plots to add explanation for this.}
-
-In Figure~\ref{fig:results:sift-scaling}, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search beginning at a multiplier of approximately 7.
-For this dataset, Sieve, plotted in red, is the highest throughput algorithm for all multipliers greater than 7.
-
-Figure \ref{fig:results:random-scaling} displays results for a random datset with the same dimensionality and cardinality as Sift (i.e., 128 and 1,000,000 respectively). 
-These results illustrate the fact that \emph{ceteris paribus}, absent a manifold structure in the dataset, CAKES's algorithms do not begin to outperform linear search until the cardinality multiplier is around 10.
-Repeated $\rho$-NN shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
-This is unsurprising, given that Repeated $\rho$-NN relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
-With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, Repeated $\rho$-NN will iterate for longer before finding this correct radius.
-
-Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
-This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities.
-% We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
-% While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
-Finally, we reiterate that the variation in performance of our algorithms across different datasets and cardinalities support our use of an autotuning function to select the best choice of algorithm. 
-% Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
-
-% 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/fashion-mnist-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on fashion-mnist for $k=10$.
-%     }
-%     \label{fig:results:fashion-mnist-scaling}
-% \end{figure}
-
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{images/result_plots/deep-image_10_scaling.png}
-%     \caption{
-%         Deep-image Scaling
-%     }
-%     \label{fig:results:deep-image-scaling}
-% \end{figure}
-
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/glove-25-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on glove-25 for $k=10$.  }
-%     \label{fig:results:glove-25-scaling}
-% \end{figure}
-
-% 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/sift-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on sift for $k=10$.    }
-%     \label{fig:results:sift-scaling}
-% \end{figure}
-
-% We need to redo this plot so that we do not actually augment random datasets but instead take larger and larger 
-% random datasets
-% 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{plots/random-1000000-128-knn-10.png}
-%     \caption{
-%         Algorithm performance by cardinality multiplier on a random dataset for $k=10$.
-%     }
-%     \label{fig:results:random-scaling}
-% \end{figure}
-
-% We did not run any of the competitor's algorithms.
-% We lifted the reported numbers from the ann-benchmarks site and their interactive plots.
-% We made sure to use the same AWS instance as they did so it is a fair comparison.
-
-% Tables \ref{table:results:ann-10} and \ref{table:results:ann-100} show the results of our benchmarks against FAISS, bruteforce-blas, and HNSW.
-% The throughput values for these algorithms comes from the ANN benchmarks website's interactive plots, from which we can assess the throughput of each algorithm at a given recall value.
-
-% We report the highest throughput value for each algorithm at a recall value of 1.0.
-% We report the mean and median speedup factor of CAKES over each algorithm.
-% For $k= 10$, we report a mean $7,654 \times$ (median $1,200 \times$) speedup over faiss-ivf, a mean $11,325 \times$ (median $2,821 \times$) speedup over bruteforce-blas, and a mean $485 \times$ (median $417 \times$) speedup over HNSW.
-% For $k=100$, we report a mean $4,479 \times$ (median $3,456 \times$) speedup over faiss-ivf, a mean $16,010 \times$ (median $13,548 \times$) speedup over bruteforce-blas, and a mean $403 \times$ (median $372 \times$) speedup over HNSW.
-
-% For all datasets benchmarked in this manuscript, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
-
-% We note that the mean and median speedup factors differ significantly for all methods, which suggests that the speedup factor is heavily dataset-dependent;
-% in particular, across all algorithms, and both values of $k$, CAKES exhibits particularly high speedup factors with SIFT and GIST, under Euclidean distance.
-% As GIST was designed to be a difficult dataset for classifiers~\cite{Lee2019PracticalLP}, CAKES's strong performance with this dataset is particularly encouraging.
-% Further investigation of CAKES's performance on other challenging or adversarial datasets is warranted.

### 2023-12-24 sections/5-discussion.tex
finished scaling results for ann and random and moved discussion-y things out of results

@@
 \label{sec:discussion}
+
+{\color{red} TODO: Use lfd plots to discuss repeated rnn being best for glove low lfd.}
+
+% Moving stuff out of results into discussion: 
+These results illustrate the fact that \emph{ceteris paribus}, absent a manifold structure in the dataset, CAKES's algorithms do not begin to outperform linear search until the cardinality multiplier is around 10.
+Repeated $\rho$-NN shows significantly worse performance on this dataset relative to both the other algorithms and to its own performance on real datasets. 
+This is unsurprising, given that Repeated $\rho$-NN relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
+With a completely random dataset, whose average local fractal dimension will be near its embedding dimension, Repeated $\rho$-NN will iterate for longer before finding this correct radius.
+
+Notably, we observe that on all datasets, GreedySieve's throughput stays nearly constant as cardinality increases.
+This observation warrants further investigation, but it is especially promising given that GreedySieve consistently outperforms linear search for high cardinalities.
+% We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
+% While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
+Finally, we reiterate that the variation in performance of our algorithms across different datasets and cardinalities support our use of an autotuning function to select the best choice of algorithm. 
+% Currently, our autotuning function always uses $k=10$ for its determination, but our observation that the best algorithm often differs between small and large values of $k$ suggests that more sophisticated autotuning is necessary.
 

### 2023-12-24 sections/4-results.tex
justification of random lfd

@@
 The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
+The LFD of approximately 20 at depth 0 is what we expect for a random dataset.
+To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum pairwise distance between any two points is approximately equal to maximum pairwise distance.
+As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
+A consequence of this is that with high probability, a point in the root cluster is a distance greater than half the root radius from the center; $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r) = |X|$, where $X$ is the dataset.
+Given our definition of LFD in Equation~\ref{eq:methods:lfd-simplified}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
 

### 2023-12-25 sections/1-introduction.tex
edits to intro

@@
 
-% Najib: Add citations for neural-embeddings of images in addition to LLMs
-With the rise of large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, neural network embeddings are becoming an important source of big data.
+With the rise of large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image analysis models such as {\color{red} [NAME]~[CITE] and [NAME]~[CITE]}, neural network embeddings are becoming an important source of big data.
 
@@
 The field of astronomy has also seen an influx of large datasets.
-% Najib: Add exact dimensions of the SDSS dataset from the CHESS paper.
 The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 8,575 dimensions.
+{\color{red} TODO: Should we drop mention of SDSS? Because we don't use it in our experiments.}
 
@@
 
-% Najib: We should not cite CHESS for the recall-throughput tradeoff. Cite something else.
 As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
-even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{ishaq2019clustered}.
+even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{Malkov2016EfficientAR, johnson2019billion, annoy, aumuller2020ann}.
 
@@
 Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \emph{exact} search remains.
-For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations with smaller values of $k$.
+For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
@@
 We provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
-We also benchmark CAKES on a genomic dataset, the Silva 18S dataset [citation needed], and a radio frequency dataset, the RadioML dataset [citation needed].
+We also benchmark CAKES on a genomic dataset, the Silva 18S dataset~\cite{10.1093/nar/gks1219}, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}.
 

### 2023-12-25 sections/4-results.tex
plot tweaks

@@
     \includegraphics[width=0.95\textwidth]{images/lfd_plots/fashion-mnist-60000-lfd.png}\\
-    \subcaption{Fashion-mnist}
+    % \subcaption{Fashion-mnist}
     \label{fig:results:fashion-mnist-lfd}
@@
     \includegraphics[width=0.95\textwidth]{images/lfd_plots/glove-25-1183514-lfd.png}\\
-    \subcaption{Glove-25}
+    % \subcaption{Glove-25}
     \label{fig:results:glove-25-lfd}
     \end{subfigure}
-    \vspace{1em}
+    % \vspace{1em}
     \\
@@
     \includegraphics[width=0.95\textwidth]{images/lfd_plots/sift-1000000-lfd.png}\\
-    \subcaption{Sift}
+    % \subcaption{Sift}
     \label{fig:results:sift-lfd}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
-    \subcaption{A random dataset}
-    \label{fig:results:random-lfd}
-    \end{subfigure}
-    \vspace{1em}
+    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
+    % \subcaption{RadioML}
+    \label{fig:results:radioml-lfd}
+    \end{subfigure}%
     \\
@@
     \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
-    \subcaption{Silva 18S metagenomics}
+    % \subcaption{Silva 18S metagenomics}
     \label{fig:results:silva-lfd}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
-    \subcaption{RadioML}
-    \label{fig:results:radioml-lfd}
-    \end{subfigure}%
+    \includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
+    % \subcaption{A random dataset}
+    \label{fig:results:random-lfd}
+    \end{subfigure}
+    % \vspace{1em}
 
-    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of LFD and weighted by the cardinalities of the clusters.}
+    \caption{Local fractal dimension vs.
+cluster depth across six datasets, grouped by decile of local fractal dimension and weighted by the cardinalities of the clusters.
+The last dataset is randomly generated.
+The number in the header of each plot indicates the cardinality of the dataset.}
     \label{fig:results:lfd-plots}

### 2023-12-25 sections/2-methods.tex
fig 1 tweaks
edits to methods

@@
 In this manuscript, we are primarily concerned with $k$-NN search in a finite-dimensional space.
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the genome of an organism, the neural-network embedding of an image, etc.
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the neural-network embedding of an image, the genome of an organism, a radio frequency signal, etc.
 
 We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
-% Najib: I don't like the ``have an identity'' wording here. The Measure Theory A-Grader will have to prove her chops for this one.
-We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and have an identity (i.e. $f(x, y) = 0 \iff x = y$).
+We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and {\color{red} have an identity (TODO: improve the wording for this phrase)} (i.e.
+$f(x, y) = 0 \iff x = y$).
 If, in addition to these constraints, the distance function obeys the triangle inequality (i.e.
 $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then the distance function is also a \emph{distance metric}.
+Under distance metrics, all search algorithms in CAKES are exact.
 For example, Euclidean, Levenshtein and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.
-Under distance metrics, all search algorithms in CAKES are exact.
 
-The choice of distance function varies by dataset.
+The choice of distance function varies by dataset and domain.
 For example, with neural-network embeddings, one could use Euclidean (L2) or Cosine distance.
@@
 
-% Najib: Reword the following sentence.
-CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, i.e., high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space, and exploits this low \emph{local fractal dimension} (LFD) to improve the performance of its search.
+CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, i.e., high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
+We say that such data are \emph{manifold-constrained}, and have low \emph{local fractal dimension} (LFD).
+CAKES exploits this topological property, i.e. low LFD, to accelerate search.
 We define LFD at some length scale around a point in the dataset as:
@@
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
-Figure~\ref{fig:results:lfd-plots} illustrates this concept, showing how real datasets obey the manifold hypothesis.
+Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets obey the manifold hypothesis.
 
@@
 
-In~\cite{yu2015entropy}, metric entropy for a given cluster radius $r$ was defined as the number of clusters of a uniform radius $r$ needed to cover all data.
-In this paper, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree.
+The \emph{metric entropy} ($\mathcal{N}_{r}(X)$) for a given cluster radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover all data.
+In this paper, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
 
@@
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. 
-While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of clusters and of the manifold, CLAM does not rely on such notions;
+While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of clusters and of the manifold, CLAM and CAKES do not rely on such notions;
 they serve merely as a convenient and intuitive way to discuss the underlying mathematics.
@@
 
-Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria.
+Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, a minimum cluster radius, a minimum cluster cardinality, a maximum tree depth, etc.
 This process is described in Algorithm \ref{alg:methods:clustering:partition}.
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{Partition(\emph{C})} % give the algorithm a caption
+    \caption{Partition(\emph{C, criteria})} % give the algorithm a caption
     \label{alg:methods:clustering:partition} % and a label for \ref{} commands later in the document
@@
         \REQUIRE $C$, a cluster
+        \REQUIRE $criteria$, user-specified stopping criteria
+
         \STATE $seeds \Leftarrow \lfloor \sqrt{|C|} \rfloor$ random points from $C$
@@
         \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
-        \IF{$|L| > 1$}
+        \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
             \STATE Partition($L$)
         \ENDIF
-        \IF{$|R| > 1$}
+        \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
             \STATE Partition($R$)
@@
 With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
-Assuming a balanced tree, and thus $\mathcal{O}\log(n)$ depth, this approach had a memory overhead of $\mathcal{O}(n\log(n))$.
+Assuming a balanced tree, and thus $\mathcal{O}(\log(n))$ depth, this approach had a memory overhead of $\mathcal{O}(n\log(n))$.
 
@@
 
-% % WE WILL USE THIS FOR COMPRESSION, so removing it for now
+% % WE WILL USE THIS FOR COMPRESSION. DO NOT DELETE
 
@@
 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
-in other words, for a given $k$, $k$-NN finds the $k$ closest points to $q$ in $\textbf{X}$.
-We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-NN search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
+in other words, $k$-NN finds the $k$ closest points to $q$ in $\textbf{X}$.
+We also have the $\rho$-NN search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
 that is, find all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$.
@@
     \centering
-    \includegraphics[scale=0.5]{images/geometry/deltas.png}
+    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
     \caption{$\delta$, $\delta^{+}$, and $\delta^{-}$ for a cluster $C$ and a query $q$.}
@@
 
-We conduct $\rho$-nearest neighbors search as described in~\cite{ishaq2019clustered}, but with the following improvement:
+We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
 when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which might contain points in the query ball.
@@
 
-Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles (see Section~\ref{subsubsec:methods:cluster-partitioning}).
+Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{subsubsec:methods:cluster-partitioning}).
 Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$ (i.e., the query is closer to $r$ than to $l$).
@@
 Now let $q'$ be the projection of $q$ to $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$ and $d$ be the distance from $q'$ to $m$.
-Since we assign a point in the parent cluster to the left child if it is closer to the left pole, if $d > \rho$ then the left child cannot contain points inside the query ball.
-In this case, we proceed only with the right child.
-If, however, $d \leq \rho$ then both child clusters could contain points in the query ball, and so we proceed with both of them.
+From Algorithm~\ref{alg:methods:clustering:partition}, we assign a point in the parent cluster to the left child,
+Thus, if if $q'$ is too far from $m$, i.e. $\rho < d$, then when the left child cannot contain points inside the query ball, we proceed only with the right child.
+Otherwise, we proceed with both children.
 
@@
 
-The asymptotic complexity of $\rho$-nearest neighbors is the same as in~\cite{ishaq2019clustered}, namely:
+The asymptotic complexity of $\rho$-nearest neighbors is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn:complexity}.
 
@@
 
-In this section, we present three novel algorithms for exact $k$-nearest neighbors search:
+In this section, we present three novel algorithms for exact $k$-NN search:
 Repeated $\rho$-NN, Sieve, and Greedy Sieve.
@@
 
-We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could replace the query with the center of one of the nearby clusters without significantly changing the LFD at the query point.
+We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could move from the query to the center of one of the nearby clusters without significantly changing our estimate of the LFD.
 
-By Equation~\ref{eq:repeated-rnn-factor}, we then have that
+By Equation~\ref{eq:methods:lfd-original}, we then have that
 
@@
 
-Since $\hat{d}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve Equation~\ref{eq:methods:repeated-rnn-leaf-search} for $\rho_k$ to get
+Since $\hat{d}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve this equation for $\rho_k$ to get
 
@@
 where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-We note that the fraction $d \cdot \hat{d}^{-1}$ should be relatively close to 1 unless fractal dimension is highly variable in a small region (that is, $\hat{d}$ differs significantly from $d$).
+We note that the term $d \cdot \hat{d}^{-1}$ should be relatively close to 1 unless fractal dimension is highly variable in a small region (that is, $\hat{d}$ differs significantly from $d$).
 
@@
 We continue this process until $Q$ no longer contains any clusters.
-We then use the QuickSelect algorithm once last time to reorder $Q$, to find the $k^{th}$ element, and to return the $k$ nearest neighbors.
+We then use the QuickSelect algorithm once last time to reorder $Q$, find the $k^{th}$ element, and return the $k$ nearest neighbors.
 
@@
 At this point, based on our assumption, we have that the length of $Q$ is close to $2d$.
-Therefore, the asymptotic complexity of the Sieve Algorithm is $\mathcal{O} \left( \log(\frac{n}{d}) \right)$.
+Therefore, the asymptotic complexity of the Sieve Algorithm is $\mathcal{O} \left( k \cdot \log(\frac{n}{d}) \right)$.
 
@@
 \begin{itemize}
-    \item While the top priority element $q$ in $Q$ is not a leaf, pop it from $Q$ and push its children to $Q$.
-    \item For each point $p \in q$, push $p$ to $H$.
+    \item While the top priority element $Q$ is not a leaf, pop it from $Q$ and push its children to $Q$.
+    \item Let $q$ the top priority element (a leaf) in $Q$. For each point $p \in q$, push $p$ to $H$.
     \item If $H$ has more than $k$ points, pop from $H$ until $H$ has $k$ points.
@@
 This algorithm is described in Algorithm~\ref{alg:methods:greedy-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$ (i.e., when the top priority element in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$).
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$ (i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$).
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
@@
 Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level and $m$ be a user-specified integer multiplier.
-For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the euclidean distance from $x$ to the origin.
+For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.
 We construct a random vector $r$ of $d$ dimensions in the hypersphere of radius $\epsilon$ centered at the origin.
 We then add $r$ to $x$ to get a new point $x'$.
-Since $||r|| = \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
+Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
 We repeat this process $m$ times for each $x \in X$ to get a new dataset $X'$.

### 2023-12-25 sections/3-datasets.tex
todo note
edits to datasets

@@
 
-All benchmarks were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
-This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}
+The benchmarks for Fashion-MNIST, Glove-25, Sift and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
+This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}.
 The OS kernel was Ubuntu 22.04.3-Ubuntu SMP.
-The Rust compiler was Rust 1.71.1, and the Python interpreter version was 3.10.12.
+The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.
+
+The benchmarks for Silva and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
+The OS kernel was Manjaro Linux 5.15.130-1-MANJARO.
+The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.
 
@@
                     \hline
-                    \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dim.}  \\
+                    \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dimensionality}  \\
                     \hline
-                    fashion-mnist    & euclidean              & 60,000             & 784       \\
+                    Fashion-Mnist    & Euclidean              & 60,000             & 784       \\
                     \hline
-                    gist             & euclidean              & 1,000,000          & 960       \\
+                    Glove-25         & Cosine                 & 1,183,514          & 25        \\
                     \hline
-                    glove-25         & cosine              & 1,183,514          & 25        \\
+                    Sift             & Euclidean              & 1,000,000          & 128       \\
                     \hline
-                    glove-50         & cosine              & 1,183,514          & 50        \\
+                    Random           & Euclidean              & 1,000,000          & 128       \\
                     \hline
-                    glove-100        & cosine              & 1,183,514          & 100       \\
+                    Silva            & Levenshtein            & 2,224,640          & -         \\
                     \hline
-                    sift             & euclidean              & 1,000,000          & 128       \\
-                    \hline
-                    deep-image       & cosine              & 9,990,000          & 96        \\
-                    \hline
-                    mnist            & euclidean              & 60,000             & 784       \\
-                    \hline
-                    lastfm           & cosine              & 292,385            & 65        \\
+                    RadioML          & Dynamic Time Warping   & 97,920             & 1,024     \\
                     \hline
@@
 
-We also benchmarked on a purely random dataset with the same cardinality and dimensionality as sift: 1,000,000 points in 128 dimensions. 
-This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of CAKES's algorithms.
+For the Random dataset, we used a cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset.
+This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of the CAKES' algorithms.
 
@@
 \label{subsec:silva}
-Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
-We use Levenshtein distance with this dataset.
+
+Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
+We use Levenshtein distance on unaligned sequences as the distance function for this dataset.
+{\color{red} TODO: Define and/or cite Levenshtein distance.}
 % I think this needs another sentence but I'm not sure what else to say
@@
 \subsection{Radio ML}
+\label{subsec:radioml}
+
 The RadioML dataset contains samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
@@
 Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
-Each sample provided in RadioML can be thought of as an array of 2048 floating point values for a single signal capture with labels for SNR and modulation mode.
-We use Dynamic Time Warping distance with this dataset.
-
-\label{subsec:radioml}
+Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time series of complex-valued numbers).
+We used a subset of this dataset, containing only the 97,920 samples at 10dB SNR.
+We use Dynamic Time Warping as the distance function for this dataset.
+{\color{red} TODO: Define and/or cite Dynamic Time Warping distance.}

### 2023-12-25 sections/4-results.tex
rearrange fig 4
edits in results

--- 2023-12-24 sections/4-results.tex
@@
 Percentile lines for glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-The 95th percentile of LFD is less than 4 for all depths, and the median is less than 2 for all depths. 
+The 95th percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths. 
 Before depth 25, the 95th percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
@@
 In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift.
-We observe that the LFD starts as high as 20 (for all data) at depth 0. As shown by the needlepoint shape of the plot, the LFD for all clusters 
-is about the same from depth 0 through depth 5. 
+We observe that the LFD starts as high as 20 (for all data) at depth 0.
+As shown by the needlepoint shape of the plot, the LFD for all clusters is about the same from depth 0 through depth 5.
 After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth.
 The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
-The LFD of approximately 20 at depth 0 is what we expect for a random dataset.
-To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum pairwise distance between any two points is approximately equal to maximum pairwise distance.
+The LFD of approximately 20 at depth 0 is what we expect for this random dataset.
+To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
 As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
@@
 
-Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently show LFD.
-The 95th percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 130. 
-The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 130.
+Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
+The 95th percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
+The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 128.
 
 Relative to the other datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
-Notably however, for the frist approximately 15 depths, the 95th percentile and median LFD are very close to 0. 
+Notably however, for the first approximately 15 depths, the 95th percentile and median LFD are very close to 0.
 Both percentiles increase sharply after depth 20, nearly reaching 16.
@@
     \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
+    \subcaption{RadioML}
+    \label{fig:results:radioml-lfd}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
+    \subcaption{Silva 18S}
+    \label{fig:results:silva-lfd}
+    \end{subfigure}%  
+    \begin{subfigure}[b]{0.47\textwidth}
     \includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
@@
     \vspace{1em}
-    \\
-    \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
-    \subcaption{Silva 18S metagenomics}
-    \label{fig:results:silva-lfd}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
-    \subcaption{RadioML}
-    \label{fig:results:radioml-lfd}
-    \end{subfigure}%
-
-    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of LFD and weighted by the cardinalities of the clusters.}
+    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of local fractal dimension and weighted by the cardinalities of the clusters.
+    The last dataset is randomly generated.
+    {\color{red} TODO: Remove the headers from the plots themselves and rely on subcaptions instead.}}
     \label{fig:results:lfd-plots}
 \end{figure}
-%% end LFD plots
 
@@
 Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
-The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
+The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}).
+The left-most point on each line is on the original dataset without any synthetic augmentation.
 The vertical axis denotes throughput in queries per second.
 Both axes are on a logarithmic scale.
-In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section. 
-In these plots, we label the recall at each scale for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall at all scales.
+In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in the Supplement.
+In these plots, we label the recall for algorithms which do not exhibit perfect recall;
+if no such labels are given for an algorithm, then that algorithm exhibits perfect recall.
 
-Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively.
-When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider
-only three decimal places.
+Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier for the fashion-mnist, glove-25, sift, and random datasets respectively.
+When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
-We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$.
-Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
-Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
+We also observe that CAKES outperforms FAISS-IVF (plotted in pink) for all multipliers greater than one.
+Though state-of-the-art algorithms HNSW (plotted in gray), and ANNOY (plotted in purple) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a multiplier as low as eight, HNSW and ANNOY have recall of 0.525 and 0.857 respectively,
+and this recall gets worse as the multiplier increases.
+Meanwhile, while CAKES exhibits perfect recall.
+Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance with increasing multipliers.
 
-With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high cardinality multipliers. 
-Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$. 
-We also see that while HNSW and Annoy have higher throughput than CAKES for all multipliers, CAKES exhibits better recall at all multupliers, as shown in Table~\ref{table:results:ann-glove-25}.
-In particular, at a cardinality multiplier as low as two, HNSW and Annoy have recall as low as 0.607 and 0.832 respectively, while CAKES exhibits perfect recall.
+With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high multipliers.
+Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-IVF for all multipliers greater than four.
+We see once again that HNSW and ANNOY have higher throughput than CAKES for all multipliers, CAKES exhibits better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
+In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
+In contrast, CAKES exhibits perfect recall at all multipliers.
 On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN, plotted in green, is consistently the best-performing CAKES algorithm.
 
-In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the sift dataset, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search at a multiplier of approximately 7.
-Again, we observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$. 
-For this dataset, Sieve, plotted in red, is the highest throughput CAKES algorithm for all multipliers greater than 7. 
+In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the sift dataset, we observe that while linear search performs best for low multipliers, our algorithms begin to outperform linear search at a multiplier of eight.
+Again, we observe that CAKES outperforms FAISS-IVF for all multipliers greater than four. 
+For this dataset, Sieve, plotted in red, is the highest throughput CAKES algorithm for all multipliers greater than right.
 Similar to the results with fashion-mnist and glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES for all multipliers, CAKES has better recall.
@@
 
-With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibt the worst throughput out of all algorithms, at all cardinality multipliers.
-As with the real datasets, HNSW and Annoy exhibit the highest throughput of any of the algorithms. 
-As with the real datasets, CAKES exhibits perfect recall at all cardinality multipliers. 
+With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibit the worst throughput out of all algorithms, at all multipliers. 
+As with the real datasets, HNSW and ANNOY exhibit the highest throughput of any of the algorithms and CAKES exhibits perfect recall at all multipliers. 
 HNSW and Annoy exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
-in particular, with a cardinality multiplier of 1, HNSW and Annoy have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.
+in particular, with a multiplier of 1, HNSW and Annoy have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.
 
@@
 \begin{subfigure}[b]{0.47\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
-\subcaption{A random dataset for $k=10$.
+\includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
+\subcaption{RadioML for $k=10$ at SnR = 10dB.
 }
-\label{fig:results:random-scaling}
+\label{fig:results:radioml-scaling}
 \end{subfigure}%
@@
 \begin{subfigure}[b]{0.47\textwidth}
-\includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
-\subcaption{RadioML for $k=10$ at SnR = 10dB.
+\includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
+\subcaption{A random dataset for $k=10$.
 }
-\label{fig:results:radioml-scaling}
+\label{fig:results:random-scaling}
 \end{subfigure}%

### 2023-12-26 sections/5-discussion.tex
adding to discusssion

--- 2023-10-09 sections/5-discussion.tex
@@
 
-We have presented CAKES, a method for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
-CAKES provides exact recall with $k$-NN when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
-CAKES is most effective when datasets are not uniform and do not obey a ``simple'' distribution such as Gaussian, but rather adhere to ``interesting'' low-dimensional manifolds even when embedded in a high-dimensional space;
-thus, CAKES works best when the manifold hypothesis is upheld.
+We have presented CAKES, a set of three algorithms for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
+CAKES is an exact $k$-NN search algorithm when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
+{\color{red} TODO: Mention (probably earlier than here) that we modified cosine so that it satisfies the triangle inequality and now have perfect recall on it}
+CAKES algorithms were designed to be most effective when the dataset is not uniformly distributed, but rather adheres to a low-dimensional manifold even when embedded in a high-dimensional space; 
+in other words, CAKES was designed to work best when the manifold hypothesis is upheld.
 
-In comparison to other methods, while CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
-This bears further investigation, but likely stems from how heavily the $k$-NN and clustering algorithms depend on the triangle inequality when discarding clusters from consideration.
+Our results indicate that CAKES algorithms do indeed perform well on datasets that adhere to the manifold hypothesis, and that they perform poorly on datasets that do not. 
+We use plots of the local fractal dimension of the datasets to quantify the extent to which a dataset exhibts a manifold structure; 
+the lower the local fractal dimension across the entire dataset, the more the dataset adheres to the manifold hypothesis.
+
+The results of our benchmarks on sift and on the random dataset with the same cardinality and dimensionality as sift illustrate a striking example of this difference.
+First, we note that, as discussed in Section~\ref{subsec:lfd-results}, the local fractal dimension of the random dataset is much higher than that of sift.
+This is unsurprising, given that the random dataset is, by definition, uniformly distributed, while sift is not.
+On the random dataset, the CAKES algorithms' throughputs decrease as the cardinality multiplier increases, with sift, GreedySieve and Sieve both exhibit nearly constant throughput as the cardinality multiplier increases.
+While RepeatedRnn does not exhibit constant throughput with sift, it does exhibit a much slower decrease in throughput than it does with the random dataset.
+Since these two datasets are the same size, the discrepancy in algorithm performance on each one isolates the effect of the manifold structure of the dataset on algorithm performance.
+We emphasize that though CAKES algorithms' throughputs are lower on a random dataset, their recalls remain perfect at all multipliers for both datasets. 
+The reverse of this trend occurs with HNSW and Annoy. 
+In contrast, for these two algorithms, throughput remains constant with cardinality mutliplier on both datasets, but their recall on the random dataset is orders of mangnitude worse than their recall on sift.
+These results illustrate the difference between CAKES algorithms and the state-of-the-art, as well as their suitability for the datasets for which they were designed.
+
+We also note that on different datasets which exhibit manifold structures, the best CAKES algorithm varies. 
+For example, on Glove-25, RepeatedRnn is the best-performing CAKES algorithm, despite being the worst-perfomring of the three on sift and fashion-mnist. 
+On Silva, RepeatedRnn is comparable to the other CAKES algorithms at low multipliers, but becomes the best of the three at higher multipliers. 
+When we view these results in light of the local fractal dimension plots in Section~\ref{subsec:lfd-results}, we realize that RepeatedRnn is the best-performing CAKES algorithm on datasets where a large proportion of the dataset has a very low (near 2) local fractal dimension, like Glove-25 and Silva.
+This general trend is unsurprising, given that Repeated $\rho$-NN relies on a low local fractal dimension around the query in order to quickly home in on the correct radius for $k$ hits. 
+While it varies which of the two algorithms has better throughput, we observe that on all three ANN benchmark datasets (fashion-mnist, glove-25, and sift), GreedySieve's and Sieve's throughput stays nearly constant as the cardinality multiplier increases.
+This observation warrants further investigation, but it is especially promising given that GreedySieve and Sieve consistently outperform linear search for high cardinalities. 
+Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities support our use of an autotuning function to select the best choice of algorithm for a given dataset.
+Given the trends such as that with low local fractal dimension and RepeatedRnn's performance, we suspect that a more sophisticated, less-heuristic autotuning function could be developed to select the best algorithm based on the properties of the dataset.
+
+We also stress that the CAKES algorithms are designed to work well for \emph{big} data, and our results support this claim; we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at some cardinality multuplier when the dataset exhibits a manifold structure.
+
+{\color{red} TODO: be consistent with naming (probably don't call it RepeatedRnn in text and call it Repeated $\rho$-NN instead); will deal with this later}
+
+{\color{red} TODO: what happened with the pivot point for Sieve? Or is it just less prominent in these plots? }
+% We also notice an interesting trend with Sieve and Sieve with Separate centers: for all datasets, at both values of $k$, both algorithms exhibit a sharp ``pivot point'' after which throughput starts to increase.
+% While we observe that this pivot point seems to be proportional to the value of $k$, as it often occurs at about a multiplier of 4 when $k=10$ and a mulitplier of 32 when $k=100$, the reason for this behavior requires further investigation.
 
@@
 Incorporating these or other distance functions in CAKES only requires a Rust implementation of the distance function.
-
-% For all datasets benchmarked, the auto-tuning method selected repeated $\rho$-nearest neighbor (Section~\ref{subsubsec:methods:knn-search:repeated-rnn}).
-% However, during limited testing on synthetic datasets obeying simple distributions, this was not always the case.
-% We seek to evaluate these approaches on many more datasets and distance metrics;
-% it could be that repeated $\rho$-NN is always best (and thus the auto-tuning is unnecessary), but such a determination would be premature.
 

### 2023-12-26 sections/4-results.tex
more additions to discussion as well as fixing some inconsistencies in results

@@
 In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in the Supplement.
-In these plots, we label the recall for algorithms which do not exhibit perfect recall;
-if no such labels are given for an algorithm, then that algorithm exhibits perfect recall.
+{\color{red} TODO: revisit this sentence with cosine dist plot in mind: In these plots, we label the recall for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall" }.
 
-Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier for the fashion-mnist, glove-25, sift, and random datasets respectively.
+Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality for  Fashion-mnist, Glove-25, Sift, and a random dataset respectively.
+Though the scaling plots above present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the highest throughput CAKES algorithm at that dataset and cardinality only.
+Given that we redo the auto-tuning process described in Section~\ref{subsec:methods:auto-tuning} for each augmented dataset, these values are an accurate representation of CAKES's performance on each dataset and cardinality multiplier.
 When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
-We also observe that CAKES outperforms FAISS-IVF (plotted in pink) for all multipliers greater than one.
-Though state-of-the-art algorithms HNSW (plotted in gray), and ANNOY (plotted in purple) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-mnist, we observe that as cardinality increases, the CAKES algorithms (GreedySieve in blue, Repeated $\rho$-NN in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search (plotted in orange).
+In particular, we see that GreedySieve and Sieve begin to exhibit sublinear time performance at a cardinality near $10^5$ (i.e., at a multiplier of about {\color{red} add value} times the original dataset's cardinality).
+We also observe that CAKES's algorithms outperform FAISS-IVF (plotted in pink) for all cardinalities larger than the original dataset's cardinality.
+Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and Annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
 For example, at a multiplier as low as eight, HNSW and ANNOY have recall of 0.525 and 0.857 respectively,
 and this recall gets worse as the multiplier increases.
-Meanwhile, while CAKES exhibits perfect recall.
+Meanwhile, CAKES exhibits perfect recall.
 Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance with increasing multipliers.
 
-With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high multipliers.
-Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-IVF for all multipliers greater than four.
-We see once again that HNSW and ANNOY have higher throughput than CAKES for all multipliers, CAKES exhibits better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
+With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at some cardinality.
+Specifically, all three of CAKES' algorithms outperform linear search at a cardinality just slightly more than the original dataset's cardinality, $10^6$.
+Similar to the results for Fashion-mnist, we see that CAKES's algorithms outperform FAISS-IVF for all cardinalities greater than $10^7$.
+We see once again that HNSW and Annoy have higher throughput than CAKES's algorithms for all cardinality multipliers, CAKES's algorithms exhibit better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
 In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
@@
 
-In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the sift dataset, we observe that while linear search performs best for low multipliers, our algorithms begin to outperform linear search at a multiplier of eight.
-Again, we observe that CAKES outperforms FAISS-IVF for all multipliers greater than four. 
-For this dataset, Sieve, plotted in red, is the highest throughput CAKES algorithm for all multipliers greater than right.
-Similar to the results with fashion-mnist and glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES for all multipliers, CAKES has better recall.
+In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the Sift dataset, we observe that while linear search performs best for lower cardinalities, our algorithms begin to outperform our implementation of linear search and FAISS-IVF at a cardinality near $10^7$.
+For this dataset, Sieve, plotted in red, is the highest-throughput CAKES algorithm for all cardinalities.
+Similar to the results with Fashion-mnist and Glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES's algorithms for all cardinality multipliers, CAKES has better recall.
 Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and Annoy have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
 
-With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibit the worst throughput out of all algorithms, at all multipliers. 
-As with the real datasets, HNSW and ANNOY exhibit the highest throughput of any of the algorithms and CAKES exhibits perfect recall at all multipliers. 
+With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibit the worst throughput out of all algorithms, at all cardinalities. 
+As with the real datasets, HNSW and Annoy exhibit the highest throughput of any of the algorithms, and CAKES exhibits perfect recall at all cardinalities. 
 HNSW and Annoy exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
@@
                     \hline
-                    1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.78~$\times10^{2}$ & 1.000 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000 \\
+                    1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.78~$\times10^{2}$ & 1.000 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000* \\
                     \hline
-                    2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.39~$\times10^{2}$ & 1.000 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000 \\
+                    2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.39~$\times10^{2}$ & 1.000 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000* \\
                     \hline
-                    4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.98~$\times10^{1}$ & 1.000 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000 \\
+                    4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.98~$\times10^{1}$ & 1.000 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000* \\
                     \hline
-                    8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.49~$\times10^{1}$ & 1.000 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000 \\
+                    8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.49~$\times10^{1}$ & 1.000 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000* \\
                     \hline
-                    16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.74~$\times10^{1}$ & 1.000 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000 \\
+                    16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.74~$\times10^{1}$ & 1.000 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000* \\
                     \hline
-                    32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 8.70~$\times10^{0}$ & 1.000 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000 \\
+                    32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 8.70~$\times10^{0}$ & 1.000 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000* \\
                     \hline
-                    64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 4.36~$\times10^{0}$ & 1.000 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000 \\
+                    64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 4.36~$\times10^{0}$ & 1.000 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000* \\
                     \hline
-                    128 & --                  & --    & --                  & --    & 2.18~$\times10^{0}$ & 1.000 & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000 \\
+                    128 & --                  & --    & --                  & --    & 2.18~$\times10^{0}$ & 1.000 & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000* \\
                     \hline
-                    256 & --                  & --    & --                  & --    & 1.07~$\times10^{0}$ & 1.000 & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000 \\
+                    256 & --                  & --    & --                  & --    & 1.07~$\times10^{0}$ & 1.000 & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000* \\
                     \hline

### 2023-12-26 sections/5-discussion.tex
more additions to discussion as well as fixing some inconsistencies in results

--- 2023-10-09 sections/5-discussion.tex
@@
 
-We have presented CAKES, a method for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
-CAKES provides exact recall with $k$-NN when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
-CAKES is most effective when datasets are not uniform and do not obey a ``simple'' distribution such as Gaussian, but rather adhere to ``interesting'' low-dimensional manifolds even when embedded in a high-dimensional space;
-thus, CAKES works best when the manifold hypothesis is upheld.
+We have presented CAKES, a set of three algorithms for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
+CAKES's algorithms are exact when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
+When we benchmarked the algorithms with cosine distance, a non-metric, we found that CAKES has \emph{nearly} perfect recall; at three decimal places of precision, CAKES's algorithms have recall indistinguishable from perfect.
 
-In comparison to other methods, while CAKES has perfect recall for distance metrics, it exhibits worse recall under non-metrics (specifically, cosine distance).
-This bears further investigation, but likely stems from how heavily the $k$-NN and clustering algorithms depend on the triangle inequality when discarding clusters from consideration.
+CAKES's algorithms were designed to be most effective when the dataset is not uniformly distributed, but rather adheres to a low-dimensional manifold even when embedded in a high-dimensional space; 
+in other words, CAKES was designed to work best when the manifold hypothesis is upheld.Our results indicate that CAKES algorithms do indeed perform well on datasets that adhere to the manifold hypothesis, and that they perform poorly on datasets that do not.
+We use the LFD by depth percentiles shown in Section~\ref{subsec:lfd-results} to quantify the extent to which a dataset obeys the manifold hypothesis; we say that
+the lower a dataset's LFD, the more that dataset exhibits a manifold structure.
+We observe a striking example of this difference when contrasting CAKES's algorithms' performance on Sift versus on the random dataset with the same cardinality and dimensionality as Sift.
+
+As seen in Section~\ref{subsec:lfd-results}, the local fractal dimension of the random dataset is much higher than that of Sift.
+This is unsurprising, given that the random dataset is, by definition, uniformly distributed, while Sift is not.
+On the random dataset, CAKES's algorithms' throughputs decrease as the cardinality multiplier increases, while with Sift, GreedySieve and Sieve both exhibit nearly constant throughput.
+While Repeated $\rho$-NN does not exhibit constant throughput with Sift, it does exhibit a much slower decrease in throughput than it does with the random dataset.
+Since these two datasets are the same size, the aforementioned discrepancies isolate the effect of the manifold structure of the dataset on algorithm performance.
+We emphasize that though CAKES's algorithms' throughputs are lower on a random dataset, their recalls remain perfect at all multipliers for both datasets. 
+A sort of converse of this trend occurs with HNSW and Annoy. 
+In contrast, for these two algorithms, \emph{throughput} remains constant with cardinality multiplier on both datasets, but their \emph{recall} for the random dataset is orders of magnitude worse than their recall for Sift.
+These results illustrate the difference between CAKES's algorithms and the state-of-the-art similarity search algorithms.
+They also demonstrate CAKES's algorithms suitability on the type of datasets for which they were designed.
+
+We also note that on different datasets which exhibit manifold structures, the highest-throughput CAKES algorithm varies. 
+For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-mnist. 
+On Silva,  Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the best of the three at higher multipliers. 
+When we view these results in light of the LFD plots in Section~\ref{subsec:lfd-results}, we realize that  Repeated $\rho$-NN appears to be the best-performing CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near 2), like Glove-25 and Silva.
+This general trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly home in on the correct radius for $k$ hits. 
+While it varies which of the two algorithms has better throughput, we observe that on all three ANN benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve's and Sieve's throughput stays nearly constant as the cardinality multiplier increases.
+This observation warrants further investigation, but it is especially promising given that GreedySieve and Sieve consistently outperform linear search for high cardinalities. 
+Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities support our use of an auto-tuning function (as discussed in Section~ref{subsec:methods:auto-tuning}) to select the best choice of algorithm for a given dataset.
+Given the trends such as that with low LFD and Repeated $\rho$-NN's performance, we suspect that a more sophisticated, less-heuristic auto-tuning function could be developed to select the best algorithm based on the properties of the dataset.
+
+We also stress that the CAKES's algorithms are designed to work well for \emph{big} data, and our results support this claim; we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at some cardinality multiplier for each of the ANN benchmark datasets we tested on.
+Moreover, we notice that the cardinality multiplier at which CAKES overtakes linear search is lower for datasets with lower LFD.
+For example, Sift has much higher LFD than Fashion-mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-mnist and Glove-25, respectively.
+For the truly random dataset, which had the highest LFD, CAKES's algorithms \emph{never} outperform linear search.
+These observations support our claims that CAKES's algorithms work well with big datasets, and further support our claims that they are much better suited to datasets which exhibit a manifold structure.
 

### 2023-12-26 sections/5-discussion.tex
more changes to discussion

@@
 This general trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly home in on the correct radius for $k$ hits. 
-While it varies which of the two algorithms has better throughput, we observe that on all three ANN benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve's and Sieve's throughput stays nearly constant as the cardinality multiplier increases.
+These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset. 
+Exploring this sensitivity is a potential avenue for future investigation, as it remains to determine at what LFD value Repeated $\rho$-NN transitions from being the best-throughput CAKES algorithm to being the worst-throughput CAKES algorithm, and whether this transition is sharp.
+
+While the algorithm with best throughput varies, we observe that on all three ANN benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve's and Sieve's throughput stays nearly constant as the cardinality multiplier increases.
 This observation warrants further investigation, but it is especially promising given that GreedySieve and Sieve consistently outperform linear search for high cardinalities. 
@@
 These observations support our claims that CAKES's algorithms work well with big datasets, and further support our claims that they are much better suited to datasets which exhibit a manifold structure.
-
-% Is this still true: "However, the ann-benchmarks results show that even when FAISS and HNSW are tuned for lower recall, they are still slower than CAKES in all cases."
 

### 2023-12-26 sections/4-results.tex
changes to results; new figure label

@@
 In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in the Supplement.
-{\color{red} TODO: revisit this sentence with cosine dist plot in mind: In these plots, we label the recall for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall" }.
+For HNSW and Annoy, we report the recall for each measurement in the plots.
+For algorithms which exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.
 
@@
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-mnist, we observe that as cardinality increases, the CAKES algorithms (GreedySieve in blue, Repeated $\rho$-NN in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search (plotted in orange).
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-mnist, we observe that as cardinality increases, the CAKES algorithms (GreedySieve in blue, Repeated $\rho$-NN in green, and Sieve in purple) exhibit higher throughput than na\"{i}ve linear search (plotted in orange).
 In particular, we see that GreedySieve and Sieve begin to exhibit sublinear time performance at a cardinality near $10^5$ (i.e., at a multiplier of about {\color{red} add value} times the original dataset's cardinality).
 We also observe that CAKES's algorithms outperform FAISS-IVF (plotted in pink) for all cardinalities larger than the original dataset's cardinality.
-Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and Annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a multiplier as low as eight, HNSW and ANNOY have recall of 0.525 and 0.857 respectively,
+Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in yellow) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and Annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a multiplier as low as eight, HNSW and Annoy have recall of 0.525 and 0.857 respectively,
 and this recall gets worse as the multiplier increases.
@@
 We see once again that HNSW and Annoy have higher throughput than CAKES's algorithms for all cardinality multipliers, CAKES's algorithms exhibit better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
-In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
+In particular, at a multiplier as low as two, HNSW and Annoy have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
 In contrast, CAKES exhibits perfect recall at all multipliers.
-On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN, plotted in green, is consistently the best-performing CAKES algorithm.
+On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN is consistently the best-performing CAKES algorithm.
 
 In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the Sift dataset, we observe that while linear search performs best for lower cardinalities, our algorithms begin to outperform our implementation of linear search and FAISS-IVF at a cardinality near $10^7$.
-For this dataset, Sieve, plotted in red, is the highest-throughput CAKES algorithm for all cardinalities.
+For this dataset, Sieve, is the highest-throughput CAKES algorithm for all cardinalities.
 Similar to the results with Fashion-mnist and Glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES's algorithms for all cardinality multipliers, CAKES has better recall.
@@
 
-{\color{red} TODO: Add Silva and RadioML results.}
+For Silva and RadioML, we benchmarked only CAKES's algorithms.
+Due to the massive sizes of these datasets, we took random sub-samples of the dataset with lower cardinalities for our benchmarks, rather than augmented versions of the dataset.
+With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
+Until cardinality near $10^4$, GreedySieve is the highest-throughput CAKES algorithm, but for cardinalities greater than that, Repeated $\rho$-NN is the highest-throughput CAKES algorithm.
+
+For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
 
@@
 In each plot, the horizontal axis represents increasing cardinality (size) of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-TODO say something about plots that didn't finish.}
+For some algorithms, we were not able to take measurements for each cardinality because the index-building took too much memory.
+{\color{red} Explain this better and be more specific} }
+\label{fig:results:scaling-plots}
 \end{figure}
@@
                     \hline
-                    \textbf{Scale}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
@@
                     \hline
-                    \textbf{Scale}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline

### 2023-12-26 sections/3-datasets.tex
some changes to datasets and results

@@
 
-We benchmark on a variety of datasets from the ann-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, along with the appropriate distance function.
+We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, along with the appropriate distance function.
 Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
 
-We compared against FAISS~\cite{johnson2019billion} (IVF and HNSW) and a BLAS-accelerated implementation of na\"ive search as reported in~\cite{johnson2019billion}.
-
-The benchmarks for Fashion-MNIST, Glove-25, Sift and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
+The benchmarks for Fashion-Mnist, Glove-25, and Sift (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
 This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}.
@@
 In addition to benchmarks on datasets from the ANN Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{subsec:methods:synthetic-data}.
-In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (refered to as ``scale'' in Table~\ref{table:results:ann-fashion} {\color{red} TODO: merge the tables into one and confirm this reference is still correct after the merge}) increases.
+In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``scale'' in Table~\ref{table:results:ann-fashion} {\color{red} TODO: merge the tables into one and confirm this reference is still correct after the merge}) increases.
 
-For the Random dataset, we used a cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset.
+We also benchmarked on purely random (i.e., not an augmented version of a known dataset) datasets of various cardinalities. 
+For this Random dataset, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset.
 This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of the CAKES' algorithms.
@@
 \label{subsec:silva}
-
+To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the Silva dataset.
 Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
@@
 \label{subsec:radioml}
-
+To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset.
 The RadioML dataset contains samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
@@
 {\color{red} TODO: Define and/or cite Dynamic Time Warping distance.}
+
+\subsection{Other Algorithms}
+We benchmarked CAKES's algorithms against the state-of-the-art similarity search algorithms HNSW and ANNOY, as well as on FAISS-IVF and two different implementations of linear search.
+In particular, we use FAISS-Flat, a Python implementation of linear search, to provide ground-truth values for recall of HNSW, ANNOY, and FAISS-IVF.
+We use our own Rust implementation of linear search to provide ground-truth values for recall of CAKES's algorithms.
+We plot the throughput of these two implementations of linear search as a reference point for CAKES's algorithms' throughput in Section~\ref{subsec:scaling-behavior-results}.

### 2023-12-26 sections/4-results.tex
some changes to datasets and results

@@
 In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in the Supplement.
-For HNSW and Annoy, we report the recall for each measurement in the plots.
+For HNSW and ANNOY, we report the recall for each measurement in the plots.
 For algorithms which exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.
 
-Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality for  Fashion-mnist, Glove-25, Sift, and a random dataset respectively.
-Though the scaling plots above present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the highest throughput CAKES algorithm at that dataset and cardinality only.
+Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality for  Fashion-Mnist, Glove-25, Sift, and a random dataset respectively.
+Though the scaling plots above present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
 Given that we redo the auto-tuning process described in Section~\ref{subsec:methods:auto-tuning} for each augmented dataset, these values are an accurate representation of CAKES's performance on each dataset and cardinality multiplier.
@@
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-mnist, we observe that as cardinality increases, the CAKES algorithms (GreedySieve in blue, Repeated $\rho$-NN in green, and Sieve in purple) exhibit higher throughput than na\"{i}ve linear search (plotted in orange).
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-Mnist, we observe that as cardinality increases, the CAKES algorithms (GreedySieve in blue, Repeated $\rho$-NN in green, and Sieve in purple) are faster than na\"{i}ve linear search (plotted in orange).
 In particular, we see that GreedySieve and Sieve begin to exhibit sublinear time performance at a cardinality near $10^5$ (i.e., at a multiplier of about {\color{red} add value} times the original dataset's cardinality).
 We also observe that CAKES's algorithms outperform FAISS-IVF (plotted in pink) for all cardinalities larger than the original dataset's cardinality.
-Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in yellow) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and Annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a multiplier as low as eight, HNSW and Annoy have recall of 0.525 and 0.857 respectively,
+Though state-of-the-art algorithms HNSW (plotted in gray), and ANNOY (plotted in yellow) are faster than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
+For example, at a multiplier as low as eight, HNSW and ANNOY have recall of 0.525 and 0.857 respectively,
 and this recall gets worse as the multiplier increases.
@@
 Specifically, all three of CAKES' algorithms outperform linear search at a cardinality just slightly more than the original dataset's cardinality, $10^6$.
-Similar to the results for Fashion-mnist, we see that CAKES's algorithms outperform FAISS-IVF for all cardinalities greater than $10^7$.
-We see once again that HNSW and Annoy have higher throughput than CAKES's algorithms for all cardinality multipliers, CAKES's algorithms exhibit better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
-In particular, at a multiplier as low as two, HNSW and Annoy have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
+Similar to the results for Fashion-Mnist, we see that CAKES's algorithms outperform FAISS-IVF for all cardinalities greater than $10^7$.
+We see once again that HNSW and ANNOY are faster than CAKES's algorithms for all cardinality multipliers, CAKES's algorithms exhibit better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
+In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
 In contrast, CAKES exhibits perfect recall at all multipliers.
@@
 In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the Sift dataset, we observe that while linear search performs best for lower cardinalities, our algorithms begin to outperform our implementation of linear search and FAISS-IVF at a cardinality near $10^7$.
-For this dataset, Sieve, is the highest-throughput CAKES algorithm for all cardinalities.
-Similar to the results with Fashion-mnist and Glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES's algorithms for all cardinality multipliers, CAKES has better recall.
-Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and Annoy have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
+For this dataset, Sieve, is the fastest CAKES algorithm for all cardinalities.
+Similar to the results with Fashion-Mnist and Glove-25, we observe that while HNSW and ANNOY are faster than CAKES's algorithms for all cardinality multipliers, CAKES has better recall.
+Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and ANNOY have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
 
-With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibit the worst throughput out of all algorithms, at all cardinalities. 
-As with the real datasets, HNSW and Annoy exhibit the highest throughput of any of the algorithms, and CAKES exhibits perfect recall at all cardinalities. 
-HNSW and Annoy exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
-in particular, with a multiplier of 1, HNSW and Annoy have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.
+With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms are the slowest out of all algorithms, at all cardinalities. 
+As with the real datasets, HNSW and ANNOY are the fastest of any of the algorithms, and CAKES exhibits perfect recall at all cardinalities. 
+HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
+in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.
 
@@
 With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
-Until cardinality near $10^4$, GreedySieve is the highest-throughput CAKES algorithm, but for cardinalities greater than that, Repeated $\rho$-NN is the highest-throughput CAKES algorithm.
-
+Until cardinality near $10^4$, GreedySieve is the fastest CAKES algorithm, but for cardinalities greater than that, Repeated $\rho$-NN is the fastest CAKES algorithm.
 For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
+In particular, throughput of CAKES algorithms is identical within three decimal places for all cardinalities.
 
@@
 \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}
-\subcaption{Fashion-mnist for $k=10$.}
+\subcaption{Fashion-Mnist for $k=10$.}
 \label{fig:results:fashion-mnist-scaling}

### 2023-12-26 sections/4-results.tex
added indexing time plots

@@
 \label{sec:results}
-
-% These are going to be boring; saving the interesting discussion for the discussion section
 
@@
     \label{fig:results:lfd-plots}
+\end{figure}
+
+\subsection{Indexing and Tuning Time}
+\label{subsec:indexing-time-results}
+
+For each of the ANN benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyperparameters for these indices to achieve the highest possible recall.
+
+\begin{figure}
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-indexing.png}\\
+    \subcaption{Fashion-mnist}
+    \label{fig:results:fashion-mnist-indexing}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{plots/glove-25-indexing.png}\\
+    \subcaption{Glove-25}
+    \label{fig:results:glove-25-indexing}
+    \end{subfigure}
+    \vspace{1em}
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{plots/sift-indexing.png}\\
+    \subcaption{Sift}
+    \label{fig:results:sift-indexing}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{plots/random-indexing.png}\\
+    \subcaption{Random}
+    \label{fig:results:random-indexing}
+    \end{subfigure}%
+    \\
+    \vspace{1em}
+    \caption{Indexing and tuning time with each of the ANN benchmark datasets and the Random dataset.}
+    \label{fig:results:indexing}
 \end{figure}

### 2023-12-26 sections/5-discussion.tex
updates to discussion

@@
 
-We have presented CAKES, a set of three algorithms for fast $k$-nearest-neighbors search that can adapt to a variety of distance functions.
-CAKES's algorithms are exact when the distance function in use is a metric (particularly, when it obeys the triangle inequality).
-When we benchmarked the algorithms with cosine distance, a non-metric, we found that CAKES has \emph{nearly} perfect recall; at three decimal places of precision, CAKES's algorithms have recall indistinguishable from perfect.
+We have presented CAKES, a set of three algorithms for fast $k$-NN search that are generic over a variety of distance functions.
+CAKES's algorithms are exact when the distance function in use is a metric (i.e., when it obeys the triangle inequality).
+Even under Cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).
 
-CAKES's algorithms were designed to be most effective when the dataset is not uniformly distributed, but rather adheres to a low-dimensional manifold even when embedded in a high-dimensional space; 
-in other words, CAKES was designed to work best when the manifold hypothesis is upheld.Our results indicate that CAKES algorithms do indeed perform well on datasets that adhere to the manifold hypothesis, and that they perform poorly on datasets that do not.
-We use the LFD by depth percentiles shown in Section~\ref{subsec:lfd-results} to quantify the extent to which a dataset obeys the manifold hypothesis; we say that
-the lower a dataset's LFD, the more that dataset exhibits a manifold structure.
-We observe a striking example of this difference when contrasting CAKES's algorithms' performance on Sift versus on the random dataset with the same cardinality and dimensionality as Sift.
+CAKES's algorithms are designed to be most effective when the dataset upholds the manifold hypothesis, i.e.
+when the dataset is constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
+As a consequence, these algorithms do not perform well on datasets with random distributions, because such datasets do not exhibit a manifold structure.
+In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each of the datasets we tested on exhibits a manifold structure, as quantified by the percentiles of LFD.
+As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on datasets with low LFD, and scale linearly on datasets with high LFD.
 
-As seen in Section~\ref{subsec:lfd-results}, the local fractal dimension of the random dataset is much higher than that of Sift.
+As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the random dataset is much higher than that of Sift.
 This is unsurprising, given that the random dataset is, by definition, uniformly distributed, while Sift is not.
-On the random dataset, CAKES's algorithms' throughputs decrease as the cardinality multiplier increases, while with Sift, GreedySieve and Sieve both exhibit nearly constant throughput.
+On the random dataset, CAKES's algorithms' throughputs decrease linearly as the multiplier increases, while with Sift, GreedySieve and Sieve both exhibit nearly constant throughput.
 While Repeated $\rho$-NN does not exhibit constant throughput with Sift, it does exhibit a much slower decrease in throughput than it does with the random dataset.
-Since these two datasets are the same size, the aforementioned discrepancies isolate the effect of the manifold structure of the dataset on algorithm performance.
-We emphasize that though CAKES's algorithms' throughputs are lower on a random dataset, their recalls remain perfect at all multipliers for both datasets. 
-A sort of converse of this trend occurs with HNSW and Annoy. 
-In contrast, for these two algorithms, \emph{throughput} remains constant with cardinality multiplier on both datasets, but their \emph{recall} for the random dataset is orders of magnitude worse than their recall for Sift.
-These results illustrate the difference between CAKES's algorithms and the state-of-the-art similarity search algorithms.
-They also demonstrate CAKES's algorithms suitability on the type of datasets for which they were designed.
+Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies show the effect of the manifold structure on algorithm performance.
+We emphasize that even though CAKES's algorithms are slower on the random dataset, their recalls remain perfect at all multipliers.
 
-We also note that on different datasets which exhibit manifold structures, the highest-throughput CAKES algorithm varies. 
-For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-mnist. 
-On Silva,  Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the best of the three at higher multipliers. 
-When we view these results in light of the LFD plots in Section~\ref{subsec:lfd-results}, we realize that  Repeated $\rho$-NN appears to be the best-performing CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near 2), like Glove-25 and Silva.
-This general trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly home in on the correct radius for $k$ hits. 
+Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near constant throughput as the multiplier increases, however, their recall continues to decrease as the multiplier increases.
+This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indexes for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
+CAKES's tree is orders of magnitude cheaper to build and CAKES's algorithms', while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
+
+We also note that the CAKES algorithm with the best throughput varies across different datasets.
+For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-Mnist. 
+On Silva, Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the best at the higher multipliers. 
+When we view these results in light of the LFD plots in Section~\ref{subsec:lfd-results}, we realize that  Repeated $\rho$-NN appears to be the best-performing CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near 2), such as Glove-25 and Silva.
+This trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly estimate the correct radius for $k$ hits. 
 These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
-Exploring this sensitivity is a potential avenue for future investigation, as it remains to determine at what LFD value Repeated $\rho$-NN transitions from being the best-throughput CAKES algorithm to being the worst-throughput CAKES algorithm, and whether this transition is sharp.
+Exploring this sensitivity is a potential avenue for future investigation, as it remains to determine at what LFD value Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
 
-While the algorithm with best throughput varies, we observe that on all three ANN benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve's and Sieve's throughput stays nearly constant as the cardinality multiplier increases.
+While the fastest CAKES algorithm varies by dataset, we observe that on all three ANN-benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve and Sieve show nearly constant throughput as the multiplier increases.
 This observation warrants further investigation, but it is especially promising given that GreedySieve and Sieve consistently outperform linear search for high cardinalities.
-Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities support our use of an auto-tuning function (as discussed in Section~ref{subsec:methods:auto-tuning}) to select the best choice of algorithm for a given dataset.
-Given the trends such as that with low LFD and Repeated $\rho$-NN's performance, we suspect that a more sophisticated, less-heuristic auto-tuning function could be developed to select the best algorithm based on the properties of the dataset.
+Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities support our use of an auto-tuning function (as discussed in Section~ref{subsec:methods:auto-tuning}) to select the best algorithm for a given dataset.
+Given trends such as the performance of Repeated $\rho$-NN with low LFD, we suspect that a more sophisticated auto-tuning function could be developed to select the best algorithm based on the properties of the dataset.
 
-We also stress that the CAKES's algorithms are designed to work well for \emph{big} data, and our results support this claim; we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at some cardinality multiplier for each of the ANN benchmark datasets we tested on.
-Moreover, we notice that the cardinality multiplier at which CAKES overtakes linear search is lower for datasets with lower LFD.
+We also stress that the CAKES's algorithms are designed to work well for \emph{big} data, and our results support this claim;
+we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough multiplier for each of the ANN-benchmarks' datasets we tested.
+Moreover, we observe that CAKES's algorithms outperform linear search at a lower multiplier when the dataset has a lower LFD.
 For example, Sift has much higher LFD than Fashion-mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-mnist and Glove-25, respectively.
-For the truly random dataset, which had the highest LFD, CAKES's algorithms \emph{never} outperform linear search.
-These observations support our claims that CAKES's algorithms work well with big datasets, and further support our claims that they are much better suited to datasets which exhibit a manifold structure.
+For the Random dataset, which has the highest LFD, CAKES's algorithms \emph{never} outperform linear search.
+These observations support our claim that CAKES performs well on datasets that arise from constrained generating phenomena, even as the cardinalities of these datasets grow exponentially.
 
 The questions raised by this study suggest several avenues for future work.
-A comparison across more datasets and larger datasets is clearly in order, as is further analysis with distance functions that existing methods such as FAISS do not support, such as Wasserstein distance~\cite{vallender1974calculation} (particularly for arbitrary dimensionality), and Tanimoto distance~\cite{bajusz2015tanimoto}, which is particularly applicable to computing distances based on maximal common subgraphs of molecular structures.
-% TODO: Add Lev, NW, SW
-Incorporating these or other distance functions in CAKES only requires a Rust implementation of the distance function.
+A comparison across more datasets and larger datasets is clearly in order, as is further analysis with distance functions that existing methods such as FAISS do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
+Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
 
+{\color{red} TODO: Trim this paragraph by a lot, because we never discuss compression and radii scaling in the paper and we don't want to leak too much of our thoughts on compression.}
 In future work, we plan to investigate compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
@@
 
-We would like to add the ability to perform ``online'' updates to the tree as points are added to or deleted from a dataset.
-This would enable CAKES to be used in a streaming environment.
+Another avenue for future work is to explore the use of CAKES in a streaming environment.
+This would require the ability to perform ``online'' updates to the tree as points are added to or deleted from the dataset.
+Such online updates would take advantage of the fast search algorithm provided by CAKES.  % and would allow us to have our cake and eat it, too.
 
-We also plan to use CAKES with anomaly detection.
-Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses a variety of methods in an ensemble to determine an anomaly score for a point.
-We would like to explore the effectiveness of using the distribution of distances among the $k$-th nearest neighbors as an additional method in our ensemble.
+CAKES can also be used to extend anomaly detection in CHAODA.
+We could add to CHAODA's ensemble of graph-based anomaly detection methods by using the distribution of distances among the $k$ nearest neighbors of cluster centers.
 
 CLAM and CAKES are implemented in Rust and their source code is available under an MIT license at https://github.com/URI-ABD/clam.
-
-% since this online update takes advantage of CAKES' fast search, it allows us to have our cake and eat it, too.

### 2023-12-26 sections/4-results.tex
start of talking about indexing

@@
 
-For each of the ANN benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyperparameters for these indices to achieve the highest possible recall.
+For each of the ANN benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyperparameters for these indices to achieve the highest possible recall (lower is better).
+Though CAKES has three different algorithms, indexing is the same for each of them, so we can simply report the indexing time for CAKES as a whole.
+
+The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}).
+The left-most point on each line is on the original dataset without any synthetic augmentation.
+The vertical axis denotes throughput in queries per second.
+Both axes are on a logarithmic scale.
+
+With the Fashion-Mnist dataset, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is much lower than that of HNSW, ANNOY, and FAISS-IVF, but higher than that of FAISS-Flat.
+This is not surprising, given that FAISS-Flat is a naive linear search algorithm and is not building an index.
+
+{\color{red} TODO: Add a sentence or two about the indexing time for the other datasets.}
 
@@
     \vspace{1em}
-    \caption{Indexing and tuning time with each of the ANN benchmark datasets and the Random dataset.}
+    \caption{Indexing and tuning time for each algorithm with each of the ANN benchmark datasets and the Random dataset.}
     \label{fig:results:indexing}

### 2023-12-26 sections/5-discussion.tex
more edits to discussion

@@
 As a consequence, these algorithms do not perform well on datasets with random distributions, because such datasets do not exhibit a manifold structure.
-In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each of the datasets we tested on exhibits a manifold structure, as quantified by the percentiles of LFD.
+In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each of the datasets we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
 As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on datasets with low LFD, and scale linearly on datasets with high LFD.
 
-As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the random dataset is much higher than that of Sift.
-This is unsurprising, given that the random dataset is, by definition, uniformly distributed, while Sift is not.
-On the random dataset, CAKES's algorithms' throughputs decrease linearly as the multiplier increases, while with Sift, GreedySieve and Sieve both exhibit nearly constant throughput.
-While Repeated $\rho$-NN does not exhibit constant throughput with Sift, it does exhibit a much slower decrease in throughput than it does with the random dataset.
-Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies show the effect of the manifold structure on algorithm performance.
-We emphasize that even though CAKES's algorithms are slower on the random dataset, their recalls remain perfect at all multipliers.
+As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift.
+This is unsurprising, given that the Random dataset is, by definition, uniformly distributed, while Sift is not.
+On the Random dataset, CAKES's algorithms' throughputs decrease linearly as the multiplier increases, while with Sift, GreedySieve and Sieve both exhibit nearly constant throughput.
+While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, the throughput decreases much more slowly than it does with the Random dataset.
+Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies show the effect of a manifold structure on algorithm performance.
+We emphasize that even though CAKES's algorithms are slower on the Random dataset, their recalls remain perfect at all multipliers.
 
 Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near constant throughput as the multiplier increases, however, their recall continues to decrease as the multiplier increases.
-This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indexes for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
+This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
 CAKES's tree is orders of magnitude cheaper to build and CAKES's algorithms', while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
 
-We also note that the CAKES algorithm with the best throughput varies across different datasets.
+We also note that the CAKES algorithm with the best throughput varies by dataset.
 For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-Mnist. 
-On Silva, Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the best at the higher multipliers. 
-When we view these results in light of the LFD plots in Section~\ref{subsec:lfd-results}, we realize that  Repeated $\rho$-NN appears to be the best-performing CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near 2), such as Glove-25 and Silva.
+On Silva, Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the fastest at the higher multipliers. 
+When we view these results in light of the LFD plots in Section~\ref{subsec:lfd-results}, we realize that Repeated $\rho$-NN appears to be the fastest CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near $2$), such as Glove-25 and Silva.
 This trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly estimate the correct radius for $k$ hits. 
 These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
-Exploring this sensitivity is a potential avenue for future investigation, as it remains to determine at what LFD value Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
+Exploring this sensitivity is a potential avenue for future investigation, as it remains to be determined at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
 
-While the fastest CAKES algorithm varies by dataset, we observe that on all three ANN-benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve and Sieve show nearly constant throughput as the multiplier increases.
+While the fastest CAKES algorithm varies by dataset, we observe that on all three ANN-benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve and Sieve show nearly constant throughput as the cardinality increases.
 This observation warrants further investigation, but it is especially promising given that GreedySieve and Sieve consistently outperform linear search for high cardinalities.
-Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities support our use of an auto-tuning function (as discussed in Section~ref{subsec:methods:auto-tuning}) to select the best algorithm for a given dataset.
-Given trends such as the performance of Repeated $\rho$-NN with low LFD, we suspect that a more sophisticated auto-tuning function could be developed to select the best algorithm based on the properties of the dataset.
+Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities supports our use of an auto-tuning function (as discussed in Section~\ref{subsec:methods:auto-tuning}) to select the best algorithm for a given dataset.
+Given trends such as the performance of Repeated $\rho$-NN with low LFD, we suspect that a more sophisticated auto-tuning function could be developed to select the fastest algorithm based on the properties of the dataset.
 
-We also stress that the CAKES's algorithms are designed to work well for \emph{big} data, and our results support this claim;
-we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough multiplier for each of the ANN-benchmarks' datasets we tested.
-Moreover, we observe that CAKES's algorithms outperform linear search at a lower multiplier when the dataset has a lower LFD.
+We also stress that CAKES's algorithms are designed to work well for \emph{big} data, and our results support this claim;
+we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmarks' datasets we tested.
+Moreover, we observe that CAKES's algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD.
 For example, Sift has much higher LFD than Fashion-mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-mnist and Glove-25, respectively.
@@
 The questions raised by this study suggest several avenues for future work.
-A comparison across more datasets and larger datasets is clearly in order, as is further analysis with distance functions that existing methods such as FAISS do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
+A comparison across more (and larder) datasets is clearly in order, as is further analysis with distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
 Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
@@
 This would require the ability to perform ``online'' updates to the tree as points are added to or deleted from the dataset.
-Such online updates would take advantage of the fast search algorithm provided by CAKES.  % and would allow us to have our cake and eat it, too.
+Such online updates would take advantage of the fast search algorithms provided by CAKES.  % and would allow us to have our cake and eat it, too.
 
-CAKES can also be used to extend anomaly detection in CHAODA.
+CAKES can also be used to extend anomaly detection in CHAODA~\cite{ishaq2021clustered}.
 We could add to CHAODA's ensemble of graph-based anomaly detection methods by using the distribution of distances among the $k$ nearest neighbors of cluster centers.

### 2023-12-26 sections/4-results.tex
rearranged plots dir. broke supplement

@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/fashion-mnist-60000-lfd.png}\\
+    \includegraphics[width=0.95\textwidth]{images/lfd/fashion-mnist-60000.png}\\
     \subcaption{Fashion-mnist}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/glove-25-1183514-lfd.png}\\
+    \includegraphics[width=0.95\textwidth]{images/lfd/glove-25-1183514.png}\\
     \subcaption{Glove-25}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/sift-1000000-lfd.png}\\
+    \includegraphics[width=0.95\textwidth]{images/lfd/sift-1000000.png}\\
     \subcaption{Sift}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
+    \includegraphics[width=0.95\textwidth]{images/lfd/radio-ml-97920.png}\\
     \subcaption{RadioML}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
+    \includegraphics[width=0.95\textwidth]{images/lfd/silva-2224640.png}\\
     \subcaption{Silva 18S}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
+    \includegraphics[width=0.95\textwidth]{images/lfd/random-1000000.png}\\
     \subcaption{A random dataset}

### 2023-12-26 supplement.tex
some edits
updated plots in supplement
rearranged plots dir. broke supplement

--- 2023-12-10 supplement.tex
@@
 
-\section{Datasets}
-
 \section{Results}
 
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on fashion-mnist with $k=100$. 
+    }
+    \label{fig:supplement:fashion-mnist-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/sift-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on sift with $k=100$. 
+    }
+    \label{fig:supplement:sift-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/glove-25-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on glove-25 with $k=100$. 
+    }
+    \label{fig:supplement:glove-25-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=3.4in]{plots/random-knn-100.png}
+    \caption{
+        Scaling behavior of algorithms on random with $k=100$. 
+    }
+    \label{fig:supplement:random-k-100}
+\end{figure}
+
+\begin{figure}[ht!]
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/radius/fashion-mnist-60000.png}\\
+    \subcaption{Fashion-mnist}
+    \label{fig:results:fashion-mnist-radius}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/radius/glove-25-1183514.png}\\
+    \subcaption{Glove-25}
+    \label{fig:results:glove-25-radius}
+    \end{subfigure}
+    \vspace{1em}
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/radius/sift-1000000.png}\\
+    \subcaption{Sift}
+    \label{fig:results:sift-radius}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/radius/radio-ml-97920.png}\\
+    \subcaption{RadioML}
+    \label{fig:results:radioml-radius}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/radius/silva-2224640.png}\\
+    \subcaption{Silva 18S}
+    \label{fig:results:silva-radius}
+    \end{subfigure}%  
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/radius/random-1000000.png}\\
+    \subcaption{A random dataset}
+    \label{fig:results:random-radius}
+    \end{subfigure}
+    \vspace{1em}
+    \caption{Radius vs. cluster depth across six datasets, grouped by decile of radius and weighted by the cardinalities of the clusters.
+    The last dataset is randomly generated.}
+    \label{fig:results:radius-plots}
+\end{figure}
+
+\begin{figure}[ht!]
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/fractal_density/fashion-mnist-60000.png}\\
+    \subcaption{Fashion-mnist}
+    \label{fig:results:fashion-mnist-fractal_density}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/fractal_density/glove-25-1183514.png}\\
+    \subcaption{Glove-25}
+    \label{fig:results:glove-25-fractal_density}
+    \end{subfigure}
+    \vspace{1em}
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/fractal_density/sift-1000000.png}\\
+    \subcaption{Sift}
+    \label{fig:results:sift-fractal_density}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/fractal_density/radio-ml-97920.png}\\
+    \subcaption{RadioML}
+    \label{fig:results:radioml-fractal_density}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/fractal_density/silva-2224640.png}\\
+    \subcaption{Silva 18S}
+    \label{fig:results:silva-fractal_density}
+    \end{subfigure}%  
+    \begin{subfigure}[b]{0.47\textwidth}
+    \includegraphics[width=0.95\textwidth]{images/fractal_density/random-1000000.png}\\
+    \subcaption{A random dataset}
+    \label{fig:results:random-fractal_density}
+    \end{subfigure}
+    \vspace{1em}
+    \caption{Fractal Density vs. cluster depth across six datasets, grouped by decile of fractal density and weighted by the cardinalities of the clusters.
+    The last dataset is randomly generated.
+    Fractal Density is defined as $\frac{cardinality}{radius^{LFD}}$}
+    \label{fig:results:fractal_density-plots}
+\end{figure}
+
 \end{document}

### 2023-12-27 sections/1-introduction.tex
reworked intro

@@
 
-Often dubbed ``the Big Data explosion,'' rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale.
-
-With the rise of large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image analysis models such as {\color{red} [NAME]~[CITE] and [NAME]~[CITE]}, neural network embeddings are becoming an important source of big data.
-
-Biology has also seen an influx of large, high-dimensional datasets.
-The GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length;
-Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
-
-Radio frequency data offers another example of massive datasets. 
-The RadioML~\cite{oshea2018radioml} dataset contains approximately 2.55 million samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
-
-The field of astronomy has also seen an influx of large datasets.
-The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 8,575 dimensions.
-{\color{red} TODO: Should we drop mention of SDSS? Because we don't use it in our experiments.}
-
+Researchers are collecting data at an unprecedented scale.
 In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{brescia2012extracting}.
 This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
-Thus, analysis of these large datasets requires better algorithms.
+Often dubbed ``the Big Data explosion,'' this phenomenon has led to a need for better algorithms for analyzing such datasets.
 
-Many researchers are especially interested in similarity search on such datasets. 
+Examples of large datasets include neural network embeddings, genomic databases and time-series data such as radio frequency signals.
+Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image analysis models such as {\color{red} [NAME]~[CITE] and [NAME]~[CITE]} are a common source of neural network embeddings.
+These embeddings are often high-dimensional, and the sizes of training and inference datasets for such networks are growing exponentially.
+Among Biological datasets, the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length while Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
+Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
+
+Many researchers are especially interested in similarity search on these datasets. 
 Similarity search enables a variety of applications, including recommendation~\cite{annoy} and classification systems~\cite{suyanto2022knnclassifier}. 
-
 As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
@@
 Given some measure of similarity between data points, e.g.
-a distance function, we define two types of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
+a distance function, there are two common definitions similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
 $k$-NN search aims to find the $k$ most similar points to a query point, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query point.
@@
 $k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
-Na\"{i}ve implementations of $k$-NN, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets because their cardinalities are growing exponentially.
-
+Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets because their cardinalities are growing exponentially.
 While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
 Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \emph{exact} search remains.
+
 For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
@@
 
+{\color{red} TODO: Talk about common uses of $\rho$-NN search, as we do with $k$-NN search above.}
+
 This paper focuses on three novel algorithms for exact $k$-NN search.
 We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS, as well as improved genericity across distance functions.
-We provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
-We also benchmark CAKES on a genomic dataset, the Silva 18S dataset~\cite{10.1093/nar/gks1219}, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}.
+We provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We also benchmark CAKES on a genomic dataset, the Silva 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
 
@@
 
-Recent search algorithms to tackling the exponential growth of data include hierarchical navigable small world networks (HNSW)~\cite{Malkov2016EfficientAR}, Inverted File indexing (FAISS-IVF)~\cite{johnson2019billion}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent search algorithms for tackling the exponential growth of data include hierarchical navigable small world networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{johnson2019billion}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 
@@
 
-Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, rather than its cardinality.
+This search paradigm exploits the geometric and topological structure inherent in large datasets.
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
 In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.

### 2023-12-27 sections/2-methods.tex
misc minor changes

@@
 We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
-We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and {\color{red} have an identity (TODO: improve the wording for this phrase)} (i.e.
-$f(x, y) = 0 \iff x = y$).
+We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and has identity of indiscernibles (i.e. $f(x, y) = 0 \iff x = y$).
 If, in addition to these constraints, the distance function obeys the triangle inequality (i.e.
@@
 Under distance metrics, all search algorithms in CAKES are exact.
-For example, Euclidean, Levenshtein and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.
+For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.
 
@@
 
-% Najib: For this figure, I want the red, green and blue lines/braces to be thinner.
-Perhaps 1/2 to 2/3 as thick.
-Also, I changed the notation for the deltas because it makes things easier in the Sieve and Greedy Sieve algorithms' sections.
 \begin{figure}[ht!]
@@
 
-% Najib: Visually, ``m'' is not centered on the line segment. It's a little bit to the right. Moving ``l'' to the right is probably the easiest fix.
 \begin{figure}[ht!]

### 2023-12-27 sections/3-datasets.tex
misc minor changes

@@
                     \hline
-                    Silva            & Levenshtein            & 2,224,640          & -         \\
+                    Silva            & Levenshtein            & 2,224,640          & 50,000         \\
                     \hline
@@
 Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
-We use Levenshtein distance on unaligned sequences as the distance function for this dataset.
-{\color{red} TODO: Define and/or cite Levenshtein distance.}
-% I think this needs another sentence but I'm not sure what else to say
+We use Levenshtein distance~\cite{levenshtein1966binary} on unaligned sequences as the distance function for this dataset.
 
@@
 To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset.
-The RadioML dataset contains samples of both live and synthetically generated signal captures of different modulation modes over a range of SNR levels.
+The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
 Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each modulation mode and SNR level~\cite{oshea2018radioml}.
@@
 We used a subset of this dataset, containing only the 97,920 samples at 10dB SNR.
-We use Dynamic Time Warping as the distance function for this dataset.
-{\color{red} TODO: Define and/or cite Dynamic Time Warping distance.}
+We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.
 

### 2023-12-27 sections/5-discussion.tex
misc minor changes

@@
 The questions raised by this study suggest several avenues for future work.
-A comparison across more (and larder) datasets is clearly in order, as is further analysis with distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
+A comparison across more (and larger) datasets is clearly in order, as is further analysis with distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
 Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
+We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
 
-{\color{red} TODO: Trim this paragraph by a lot, because we never discuss compression and radii scaling in the paper and we don't want to leak too much of our thoughts on compression.}
-In future work, we plan to investigate compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
-When we descend some levels in the cluster tree, it takes fewer bits to encode each point in terms of the center because the radii decrease.
-However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
-This presents the following tradeoff:
-\emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
-In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
+% When we descend some levels in the cluster tree, it takes fewer bits to encode each point in terms of the center because the radii decrease.
+% However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
+% This presents the following tradeoff:
+% \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
+% In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
 

### 2023-12-27 sections/1-introduction.tex
filled in related works

@@
 
-Recent search algorithms for tackling the exponential growth of data include hierarchical navigable small world networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{johnson2019billion}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent search algorithms for tackling the exponential growth of data include hierarchical navigable small world networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 
@@
 
-TODO: Paragraph summarizing HNSW.
+HNSW is an approximate $k$-NN search method based on navigable small world (NSW) graphs and skip lists. 
+Similar to the NSW algorithm, HNSW builds a graph of the dataset, but unlike NSW, the graph is multi-layered.
+The query point and each data point are inserted into the graph one at a time, and, upon insertion, the element is joined by an edge to the $M$ nearest nodes in the in the graph, where $M$ is a tunable parameter.
+The highest layer in which an element can be placed is determined randomly with an exponentially decaying probability distribution.
+Search starts at the highest layer and descends to the lowest layer, greedily following a path of edges to the nearest node, until reaching the query point. 
+To improve accuracy, the efSearch hyperparameter can be changed to specify the number of closest nearest neighbors to the query vector to be found at each layer.
 
@@
 
-TODO: Paragraph summarizing FAISS-IVF.
+InVerted File indexing (IVF) is a method for approximate $k$-NN search. 
+The dataset is clustered into a high dimensional Voronoi diagram, and whichever cell the query point falls into is then searched exhaustively.
+The number of cells created is governed by the $n_{list}$ parameter. 
+Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
+To mitigate accuracy issues caused by a query point falling near a cell boundary, IVF has a tunable parameter $n_{probe}$, which specifies the number of cells to search.
 
@@
 
-TODO: Paragraph summarizing ANNOY.
+Annoy is an approximate $k$-NN search method based on random projection and tree building.
+At each intermediate node of the tree, two points are sampled from the space, and the hyperplane equidistant from them is chosen to divide the space into two subspaces.
+This process is repeated multiple times to create a forest of trees, and the number of times the process is repeated is a tunable parameter.
+{\color{red} TODO: I'm still unclear what they are doing with these trees.}
 

### 2023-12-27 sections/2-methods.tex
edits in methods

@@
 In this manuscript, we are primarily concerned with $k$-NN search in a finite-dimensional space.
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the neural-network embedding of an image, the genome of an organism, a radio frequency signal, etc.
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the neural-network embedding of an image, the genome of an organism, a measurement of a radio frequency signal, etc.
 
-We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points $x, y \in \textbf{X}$, deterministically returns a non-negative real number.
-We require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$) and has identity of indiscernibles (i.e. $f(x, y) = 0 \iff x = y$).
-If, in addition to these constraints, the distance function obeys the triangle inequality (i.e.
-$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then the distance function is also a \emph{distance metric}.
+We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a non-negative real number.
+A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
+We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
+In addition to these constraints, if the distance function obeys the triangle inequality (i.e.
+$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \emph{distance metric}.
 Under distance metrics, all search algorithms in CAKES are exact.
-For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality.
+For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality (e.g.
+consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on a Cartesian plane).
 
@@
 For example, with neural-network embeddings, one could use Euclidean (L2) or Cosine distance.
-With genomic or proteomic data, Levenshtein or Hamming distances are more appropriate.
-With time-series data, one could use Dynamic Time Warping (DTW) distance.
+With genomic or proteomic data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
+With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.
 
@@
 We stress that this concept differs from the \emph{embedding dimension} of a dataset.
-To illustrate the difference, consider SDSS's APOGEE dataset, wherein each datum is a nonnegative real-valued vector of length 8,575.
-Hence, the \emph{embedding dimension} of this dataset is 8,575. 
-However, due to physical constraints (namely, the laws of physics that govern stellar fusion and spectral emission lines), the data are constrained to a lower-dimensional manifold within the 8,575-dimensional embedding space.
+To illustrate the difference, consider the Silva 18S rRNA dataset which contains which contains genomes with unaligned lengths of up to 3,718 base pairs.
+Hence, the \emph{embedding dimension} of this dataset is at least 3,718.
+However, due to physical constraints (namely, the laws that govern RNA replication and mutation), the data are constrained to a lower-dimensional manifold within this embedding space.
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
@@
 \subsection{Clustering}
-\label{subsec:methods:clustering}
+\label{sec:methods:clustering}
 
@@
 
-The \emph{metric entropy} ($\mathcal{N}_{r}(X)$) for a given cluster radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover all data.
+The \emph{metric entropy} ($\mathcal{N}_{r}(X)$) for some radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover the data.
 In this paper, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
@@
 The procedure is similar to that outlined in CHESS, but with the following improvements:
-better selection of poles for partitioning (see Algorithm~\ref{alg:methods:clustering:partition}) and depth-first reordering of the dataset (see Section~\ref{subsubsec:methods:clustering:dataset-depth-first-reordering}).
+better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}) and depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
 
@@
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
-While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of clusters and of the manifold, CLAM and CAKES do not rely on such notions;
-they serve merely as a convenient and intuitive way to discuss the underlying mathematics.
+While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
+they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
 
-\subsubsection {Cluster Partitioning}
-\label{subsubsec:methods:cluster-partitioning}
+\subsubsection{Building the Tree}
+\label{sec:methods:clustering:building-the-tree}
+
+CLAM starts by building a binary tree of clusters using a divisive hierarchical clustering algorithm.
 
@@
 The \emph{radius} of $C$ is the maximum distance from the \emph{center} to any other point in $C$.
+The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole} and the point which is furthest from left pole is designated the \emph{right pole}.
 
-The point which is responsible for that radius (i.e., the furthest point from the \emph{center}) is designated the \emph{left pole}.
-The point which is furthest from \emph{left pole} is designated the \emph{right pole}.
-
-We then partition the cluster into a \emph{left child} and a \emph{right child}, where the left child contains all points in the cluster which are closer to the \emph{left pole} than to the \emph{right pole}, and the right child contains all points in the cluster which are closer to the \emph{right pole} than to the \emph{left pole}.
+We then partition the cluster into a \emph{left child} and a \emph{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
 Without loss of generality, we assign to the left child those points which are equidistant from the two poles.
 
-Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, a minimum cluster radius, a minimum cluster cardinality, a maximum tree depth, etc.
-This process is described in Algorithm \ref{alg:methods:clustering:partition}.
+Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc.
+This process is described in Algorithm \ref{alg:methods:partition}.
 
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{Partition(\emph{C, criteria})} % give the algorithm a caption
-    \label{alg:methods:clustering:partition} % and a label for \ref{} commands later in the document
+    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
+    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
@@
         \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+
         \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
-            \STATE Partition($L$)
+            \STATE Partition($L$, $criteria$)
         \ENDIF
+
         \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
-            \STATE Partition($R$)
+            \STATE Partition($R$, $criteria$)
         \ENDIF
@@
 
-\subsubsection {Depth-First Reordering}
-\label{subsubsec:methods:clustering:dataset-depth-first-reordering}
+Note that Algorithm~\ref{alg:methods:partition} does not necessarily produce a balanced tree.
+In fact, for real datasets, we would expect anything but a balanced tree.
+The imbalance in the tree would be due to the varying density of the sampling in different regions of the manifold and the low dimensional ``shape'' of the manifold itself.
+The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g. in a $d$-dimensional hyper-cube.
+In this sense, the imbalance in the tree is a feature, not a bug, as it reflects the underlying structure of the data.
 
-CAKES also improves upon CHESS by reordering the dataset in depth-first order of traversal of the tree.
+\subsubsection{Depth-First Reordering}
+\label{sec:methods:clustering:depth-first-reordering}
+
+After building the tree, we reorder the dataset so that the points are stored in the order that they would be visited during a depth-first traversal of the tree.
 
@@
 This list was used to retrieve the clusters' points during search.
-Though this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
+Although this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
 With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
@@
 Then, within each cluster, we need only store its \emph{cardinality} and an \emph{offset} to access the its points from the dataset.
-
 The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
@@
 
-% \subsubsection {Scaling Behavior of Cluster Radii}
-% \label{subsubsec:methods:clustering:guaranteed-decrease-in-cluster-radii}
+% \subsubsection{Scaling Behavior of Cluster Radii}
+% \label{sec:methods:clustering:scaling-behabior-of-cluster-radii}
 
-% While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:clustering:partition}), unfortunately, this assumption is incorrect.
+% While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:partition}), unfortunately, this assumption is incorrect. 
 % Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii;
@@
 % will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes. 
-% See Figure~\ref{fig:methods:scaling-behavior} for an example.
+% See Figure~\ref{fig:methods:radii-scaling-behavior} for an example.
 
@@
 %     }
-%     \label{fig:methods:scaling-behavior}
+%     \label{fig:methods:radii-scaling-behavior}
 % \end{figure}
 
-\subsubsection {Complexity}
-\label{subsubsec:methods:clustering:clustering:complexity}
+\subsubsection{Complexity}
+\label{sec:methods:clustering:complexity}
 
@@
 \subsection{The Search Problem}
-\label{subsec:methods:the-search-problem}
+\label{sec:methods:the-search-problem}
 
-Given Algorithm~\ref{alg:methods:clustering:partition}, we can now pose the $k$-NN and $\rho$-NN search problems.
+Given Algorithm~\ref{alg:methods:partition}, we can now pose the $k$-NN and $\rho$-NN search problems.
 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-NN search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $\forall x \in \textbf{X} \setminus H, \ f(x, q) > \max \{ f(y, q): y \in H \}$;
 in other words, $k$-NN finds the $k$ closest points to $q$ in $\textbf{X}$.
@@
 
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $\rho$- and $k$-NN algorithms make use of the following properties, as illustrated in Figure~\ref{fig:methods:deltas}.
+Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
+Our $\rho$-NN and $k$-NN search algorithms make use of the following properties, as illustrated in Figure~\ref{fig:methods:deltas}.
 
@@
     \centering
-    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
-    \caption{$\delta$, $\delta^{+}$, and $\delta^{-}$ for a cluster $C$ and a query $q$.}
+    \includegraphics[scale=0.8]{images/geometry/deltas.pdf}
+    \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
     \label{fig:methods:deltas}
@@
 \subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
-\label{subsec:methods:rnn-search}
+\label{sec:methods:rnn-search}
 
@@
     \centering
-    \includegraphics[scale=0.5]{images/geometry/overlapping-children-3.pdf}
-    \caption{
-        $m$ denotes the midpoint of $\overline{l r}$. $q'$ is the projection of $q$ onto $\overline{l r}$.
-        $d$ is the distance from $q'$ to $m$.
-        If $d \leq \rho$, both child clusters could contain points in the query ball.
-        Otherwise, only the right child cluster could contain such points. 
-    }
-    \label{fig:methods:overlapping-children-3}
+    \includegraphics[scale=0.8]{images/geometry/overlapping-children-3.pdf}
+    \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.}
+    \label{fig:methods:overlapping-children}
 \end{figure}
 
-To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children-3}.
+To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children}.
 Here, we overload the notation for $\overline{x y}$ to refer to the line segment joining points $x$ and $y$ as well as the length of that line segment.
 
-Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{subsubsec:methods:cluster-partitioning}).
-Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$ (i.e., the query is closer to $r$ than to $l$).
-
+Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{sec:methods:clustering:building-the-tree}).
+Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
 Now let $q'$ be the projection of $q$ to $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$ and $d$ be the distance from $q'$ to $m$.
-From Algorithm~\ref{alg:methods:clustering:partition}, we assign a point in the parent cluster to the left child,
-Thus, if if $q'$ is too far from $m$, i.e. $\rho < d$, then when the left child cannot contain points inside the query ball, we proceed only with the right child.
+As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $q'$ is too far from $m$, i.e.
+$\rho < d$, then when the left child cannot contain points inside the query ball.
+In such a case we proceed only with the right child.
 Otherwise, we proceed with both children.
@@
 
-\begin{equation}
+\begin{equation*}
     d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}.
-    \label{eq:methods:d-expression}
-\end{equation}
+\end{equation*}
 
-Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children-3}.
+Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
 By the Law of Cosines on $\triangle l r q$, we have that
 
-\begin{equation}
-    \text{cos}(\theta) = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}.
-    \label{eq:methods:cosine-law}
-\end{equation}
+\begin{equation*}
+    \text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}.
+\end{equation*}
 
@@
 
-\begin{equation}
+\begin{equation*}
     \text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}.
-    \label{eq:methods:cosine-law-right-triangle}
-\end{equation}
+\end{equation*}
 
-Combining Equations~\ref{eq:methods:cosine-law} and~\ref{eq:methods:cosine-law-right-triangle}, we have that
+Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that
 
-\begin{equation}
+\begin{equation*}
     \overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}.
-    \label{eq:methods:q-prime-r}
-\end{equation}
+\end{equation*}
 
-Substituting Equation~\ref{eq:methods:q-prime-r} into Equation~\ref{eq:methods:d-expression}, we have that
+Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that
 
@@
     d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho.
-    \label{eq:methods:check-d}
+    \label{eq:methods:overlapping-children}
 \end{equation}
 
-Note, in particular, that Equation~\ref{eq:methods:check-d} only requires distances between real points, and so it can be used with any distance function, even when $q'$ and $m$ are not real points or cannot be imputed from the data.
+Note, in particular, that Equation~\ref{eq:methods:overlapping-children} only requires distances between real points, and so it can be used with any distance function, even when $q'$ and $m$ are not real points or cannot be imputed from the data.
 
-To perform $\rho$-nearest neighbors search, we first perform a coarse tree-search, as outlined in Algorithm~\ref{alg:methods:rnn:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
-Then, for all such clusters, we perform a finer-grained search, as outlined in Algorithm~\ref{alg:methods:rnn:leaf-search}, to find all points in which are within a distance $\rho$ of the query.
+To perform $\rho$-NN search, we first perform a coarse \emph{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
+Then, for all such clusters, we perform a finer-grained \emph{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points in which are no more than a distance $\rho$ from the query.
 
@@
     \caption{tree-search($C$, $q$, $\rho$)} 
-    \label{alg:methods:rnn:tree-search} 
+    \label{alg:methods:rnn-search:tree-search} 
     \begin{algorithmic}
@@
 
-        \STATE $c$ $\Leftarrow$ \emph{center} of $C$
-        \STATE $r$ $\Leftarrow$ \emph{radius} of $C$
+        \IF{$C$ is a leaf \textbf{or} $\delta^+_C \leq \rho$}
+            \STATE \textbf{return} $\{C\}$
+        \ELSE
 
-        \IF{$C$ is a leaf \textbf{or} $f(q, c) + r \leq \rho$}
-            \STATE \textbf{return} $\{C\}$
-        \ELSIF{Inequality~\ref{eq:methods:check-d} is true}
-            \STATE $L$ $\Leftarrow$ \emph{left child} of $C$
-            \STATE $R$ $\Leftarrow$ \emph{right child} of $C$
+            \STATE $[l, r]$ $\Leftarrow$ \emph{pole}s of $C$
+            \STATE $[L, R]$ $\Leftarrow$ \emph{children} of $C$
+
+            \IF{$\overline{q l} < \overline{q r \vphantom{l}}$}
+                \STATE $[r, l] \Leftarrow [l, r]$
+                \STATE $[R, L] \Leftarrow [L, R]$
+            \ENDIF
+
+            \IF{$(\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$}
                 \STATE \textbf{return} tree-search($L, q, \rho$) $\cup$ tree-search($R, q, \rho$)
-        \ELSIF{$f(q, l) < f(q, r)$}
-            \STATE \textbf{return} tree-search($L, q, \rho$)
             \ELSE
                 \STATE \textbf{return} tree-search($R, q, \rho$)
+            \ENDIF
         \ENDIF
@@
     \caption{leaf-search($S$, $q$, $\rho$)} 
-    \label{alg:methods:rnn:leaf-search} 
+    \label{alg:methods:rnn-search:leaf-search} 
     \begin{algorithmic}
@@
         \FOR{$C \in S$}
-            \STATE $c$ $\Leftarrow$ \emph{center} of $C$
-            \STATE $r$ $\Leftarrow$ \emph{radius} of $C$
-            \IF{$f(q, c) + r \leq \rho$}
+            \IF{$\delta^+_C \leq \rho$}
                 \STATE $H$ $\Leftarrow$ $H \cup \{C\}$
@@
     \caption{$\rho$-NN-search($root$, $q$, $\rho$)} 
-    \label{alg:methods:rnn:rnn-search} 
+    \label{alg:methods:rnn-search} 
     \begin{algorithmic}
@@
 
-The asymptotic complexity of $\rho$-nearest neighbors is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn:complexity}.
+The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
 
@@
     \Bigg)
-    \label{eq:methods:rnn:complexity}
+    \label{eq:methods:rnn-search-complexity}
 \end{gather}
 
-where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query in a region of radius $r$.
-From Algorithm~\ref{alg:methods:rnn:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-nearest neighbors search performance scales linearly with the size of the output set and exponentially with the local fractal dimension around the query.
-While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the local fractal dimension around the query is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.
+Here, $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query in a region of radius $\rho + \hat{r}$.
+From Algorithm~\ref{alg:methods:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-NN search performance scales linearly with the size of the output set and exponentially with the LFD around the query.
+While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the LFD of real-life datasets is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.
 
 \subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
-\label{subsec:methods:knn-search}
+\label{sec:methods:knn-search}
 
 In this section, we present three novel algorithms for exact $k$-NN search:
-Repeated $\rho$-NN, Sieve, and Greedy Sieve.
+Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
 
-In these algorithms, we use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
+In these algorithms, we use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far, and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
 
 \subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{subsubsec:methods:knn-search:repeated-rnn}
+\label{sec:methods:knn-search:repeated-rnn}
 
@@
     \min \left(2, \left( {\frac{k}{\sum_{C \in S} |C|}} \right)^{\mu^{-1}} \right)
-    \label{eq:repeated-rnn-factor}
+    \label{eq:methods:repeated-rnn-factor}
 \end{equation}
@@
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
+We allow a maximum increase of 2 to ensure that we do not increase the radius too quickly in any single iteration.
+
 Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
 Additionally, when the LFD at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.
-In such a region, a small increase in the radius would likely encounter unoccupied space, so a larger radial increase is needed.
+In such a region, a small increase in the radius would likely encounter vacant space, so a larger radial increase is needed.
 Thus, the factor of radius increase should also be \emph{inversely} related to the LFD.
 
-Once $\sum_{C \in S} |C| >= k$, we are guaranteed to have found at least $k$ neighbors, and so we can stop increasing the radius.
-We perform $\rho$-nn-search with radius $r$ and return the $k$ nearest neighbors.
+Once $\sum_{C \in S} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we can stop increasing the radius.
+We perform $\rho$-NN search with this radius and return the $k$ nearest neighbors.
 
@@
     \caption{repeated-$\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
-    \label{alg:knn:repeated-rnn} % and a label for \ref{} commands later in the document
+    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
@@
         \STATE $r \Leftarrow$ $\frac{r}{|root|}$
-        \STATE $S \Leftarrow \emptyset$, a set of clusters
+        \STATE $Q \Leftarrow \emptyset$, a set of clusters
 
-        \WHILE {$S = \emptyset$}
-            \STATE $S \Leftarrow$ tree-search($root$, $q$, $r$)
+        \WHILE {$Q = \emptyset$}
+            \STATE $Q \Leftarrow$ tree-search($root$, $q$, $r$)
             \STATE $r \Leftarrow 2 \cdot r$
@@
         \ALOOP{}
-            \IF{$\sum_{C \in S} |C| >= k$}
+            \IF{$\sum_{C \in Q} |C| >= k$}
                 \STATE \textbf{break}
             \ENDIF
-            \STATE $\mu \Leftarrow \frac{|S|}{\sum_{C \in S} \frac{1}{LFD(C)}}$
-            \STATE $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in S} |C|}} \right)^{\mu^{-1}} \right)$
+            \STATE $\mu \Leftarrow \frac{|S|}{\sum_{C \in Q} \frac{1}{LFD(C)}}$
+            \STATE $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)$
             \STATE $S \Leftarrow$ tree-search($root$, $q$, $r$)
@@
 \subsubsection{Complexity of Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{subsubsec:methods:repeated-rnn-complexity}
+\label{sec:methods:knn-search:repeated-rnn-complexity}
 
-The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data, or in other words, it arises from the same generative process as the rest of the data.
+The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data or, in other words, that it arises from the same generative process as the rest of the dataset.
 Given the uses of $k$-NN search in practice, this assumption is reasonable.
-From this assumption, we can infer that the LFD at the query point does not differ significantly from the (harmonic) mean LFD of clusters near the query.
+From this assumption, we can infer that the LFD near the query does not differ significantly from the (harmonic) mean of the LFDs of clusters near the query at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
 
@@
 Tree-search refers to the process of identifying clusters which have overlap with the query ball, or in other words, clusters which might contain one of the $k$ nearest neighbors. 
-Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by tree-search.
+Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in the clusters identified by tree-search.
 
 In ~\cite{ishaq2019clustered}, we showed that the complexity of tree-search is $\mathcal{O}(\log\mathcal{N}_{\hat{r}}(X))$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset $X$ at a radius $\hat{r}$.
-To adjust this bound for Repeated $\rho$-NN, we must estimate the number of iterations of tree-search needed to find at least $k$ neighbors.
+To adjust this bound for Repeated $\rho$-NN, we must estimate the number of iterations of tree-search needed to find a radius that guarantees at least $k$ neighbors.
 
-Based on the assumption that the LFD at the query point does not differ significantly from that of nearby clusters, the factor from Equation~\ref{eq:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
+Based on the assumption that the LFD near the query does not differ significantly from that of nearby clusters, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
 one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
@@
 
-To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in S} |C|$, the number of points in the clusters returned by tree search.
-Since we must examine each of these points, time complexity of leaf-search is linear in $\sum_{C \in S} |C|$. 
+To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the number of points in the clusters returned by tree search.
+Since we must examine each point in each such cluster, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$. 
 Let $\rho_k$ be the distance from the query to the $k$th nearest neighbor.
-Then, we see that $S$ is expected to be the set of clusters which overlap with a ball of radius $\rho_k$ around the query.
-We can estimate this region as a ball of radius $\rho_k + \hat{r}$, where $\hat{r}$ is the mean radius of clusters in $S$.
+Then, we see that $Q$ is expected to be the set of clusters which overlap with a ball of radius $\rho_k$ around the query.
+We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $Q$.
 
@@
 
-\begin{equation}
+\begin{equation*}
     \sum_{C \in S} |C| \leq \gamma k \left( 1+ \frac{2\hat{r}}{\rho_k} \right)^d
-    \label{eq:methods:repeated-rnn-leaf-search}
-\end{equation}
+\end{equation*}
 
 It remains to provide an estimate for $\rho_k$. 
-For this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different than that of nearby clusters.
+For this, we once again rely on the assumption that the query is drawn from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different from the LFD of nearby clusters.
 
-We let $\hat{d}$ be the mean LFD of the clusters in $S$.
-While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \emph{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{C}$, of clusters in $S$ at a radius equal to the mean of their radii, $\hat{r}$.
+We let $\hat{d}$ be the mean LFD of the clusters in $Q$.
+While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \emph{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
 
@@
 
-Since $\hat{d}$, $\hat{C}$, $\hat{r}$, and $k$ are all known values, we can solve this equation for $\rho_k$ to get
+Since $\hat{d}$, $\hat{|C|}$, $\hat{r}$, and $k$ are all known values, we can solve this equation for $\rho_k$ to get
 
 \begin{equation*}
-    \rho_k = \hat{r} \left( \frac{k}{\hat{C}} \right)^{\hat{d}^{-1}}
+    \rho_k = \hat{r} \left( \frac{k}{\hat{|C|}} \right)^{\hat{d}^{-1}}
 \end{equation*}
 
-Simplifying the term for leaf search in Equation~\ref{eq:methods:rnn:complexity}, we get
+Using this to simplify the term for leaf search in Equation~\ref{eq:methods:rnn-search-complexity}, we get
 
 \begin{equation*}
-    k \left( 1+ \frac{2\hat{r}}{\rho_k} \left( \frac{k}{\hat{C}} \right) ^ {\hat{d}^{-1}} \right)^d
+    k \left( 1+ \frac{2\hat{r}}{\rho_k} \left( \frac{k}{\hat{|C|}} \right) ^ {\hat{d}^{-1}} \right)^d
 \end{equation*}
@@
         \underbrace{
-            \left( \frac{2\hat{r}}{\rho_k} \right) ^ d \left( \frac{k}{\hat{C}} \right) ^ {d \cdot \hat{d}^{-1}}
+            \left( \frac{2\hat{r}}{\rho_k} \right) ^ d \left( \frac{k}{\hat{|C|}} \right) ^ {d \cdot \hat{d}^{-1}}
         }_{\textrm{scaling factor}}
     \Bigg)
-    \label{eq:methods:knn-by-rnn-complexity}
+    \label{eq:methods:repeated-rnn-complexity}
 \end{gather}
@@
 
-\subsubsection{Sieve}
-\label{subsubsec:methods:knn:sieve}
+\subsubsection{Breadth-First Sieve}
+\label{sec:methods:knn-search:bredth-first-sieve}
 
@@
 
-We begin by letting $Q$ be a list of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
 During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{subsec:methods:the-search-problem}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:the-search-problem}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.
 
-% TODO (Morgan): I don't like "kth smallest $\delta^{+}$ while accounting for multiplicities" 
-We then use the QuickSelect algorithm to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$ while accounting for multiplicities; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
+We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
 
@@
 Next, we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
-Finally, we replace all non-leaf clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.
+Finally, we replace all remaining clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.
 
 We continue this process until $Q$ no longer contains any clusters.
-We then use the QuickSelect algorithm once last time to reorder $Q$, find the $k^{th}$ element, and return the $k$ nearest neighbors.
+We then use the QuickSelect algorithm once last time to reorder $Q$, find $\tau$, and return the $k$ nearest neighbors.
 
-This process is described in Algorithm~\ref{alg:methods:sieve}.
+This process is described in Algorithm~\ref{alg:methods:bredth-first-sieve}.
 
 \begin{algorithm} % enter the algorithm environment
-    \caption{Sieve($root$, $q$, $k$)} % give the algorithm a caption
-    \label{alg:methods:sieve} % and a label for \ref{} commands later in the document
+    \caption{Breadth-First Sieve($root$, $q$, $k$)} % give the algorithm a caption
+    \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
@@
         \STATE $c \Leftarrow$ \emph{center} of $root$
-        \STATE $Q \Leftarrow$ [($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1)]
+        \STATE $Q \Leftarrow$ \{ ($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1) \}
 
@@
                 \IF{$\delta^{-}_{p} > \tau$}
-                    \STATE $Q.remove((p, \delta^{+}_{p}, m))$
+                    \STATE $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
                 \ENDIF
@@
             \FOR{$(p, \delta^{+}_{p}, m) \in Q$}
+                \STATE $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
                 \IF{$p$ is a point}
@@
                     \FOR{$c \in p$}
-                        \STATE $Q.push((c, \delta_{c}, 1))$
+                        \STATE $Q \Leftarrow Q \cup \{ (c, \delta_{c}, 1) \}$
                     \ENDFOR
                 \ELSE
-                    \STATE $[l, r] \Leftarrow$ children of $p$
-                    \STATE $Q.append([(l, \delta^{+}_{l}, |l| - 1), (l_c, \delta_{l}, 1)])$
-                    \STATE $Q.append([(r, \delta^{+}_{r}, |r| - 1), (r_c, \delta_{r}, 1)])$
+                    \STATE $[L, R] \Leftarrow$ children of $p$
+                    \STATE $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (L_c, \delta_{L}, 1) \}$
+                    \STATE $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (R_c, \delta_{R}, 1) \}$
                 \ENDIF
-                \STATE $Q.remove((p, \delta^{+}_{p}, m))$
             \ENDFOR
@@
 
-\subsubsection{Complexity of Sieve}
-\label{subsubsec:methods:sieve-complexity}
+\subsubsection{Complexity of Breadth-First Sieve}
+\label{sec:methods:knn-search:complexity-of-breadth-first-sieve}
 
-The asymptotic complexity of Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-This method is linear in the size of the input list $Q$, so we must estimate the size of $Q$ on each iteration.
+The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
+Since this method is linear in the size of the input list $Q$, we must estimate the size of $Q$ on each iteration.
 
 We start by considering clusters whose cardinalities are within a small factor of $k$, say a factor of 2.
-The distribution of such clusters will mimic the distribution of points in the dataset. 
-From this assumption, we can infer that the number of clusters which overlap with the query ball is a function of the LFD near the query.
-In particular, we assume that the number of clusters which overlap the query ball is bounded above by 2$d$, where $d$ is the LFD near the query point, e.g.
+For small $k$, as is the case in real-world applications, the distribution of such clusters will mimic the distribution of points in the dataset.
+From this assumption, we can infer that the number of clusters which overlap with the query ball is a function of $d$, the LFD near the query.
+In particular, we assume that the number of clusters which overlap the query ball is bounded above by $2d$, e.g.
 we could have a cluster overlapping the query ball at each end of each of $d$ mutually-orthogonal axes.
 
-In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log{\frac{n}{2k}}$ iterations of breadth-first traversal to reach clusters whose cardinalities are at most $2k$.
-However, since high cardinality clusters occur only at low depths in the tree, the length of $Q$, and thus the cost of QuickSelect, for those iterations is relatively low.
+In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log{\frac{\mathcal{N}_{\hat{r}}(X)}{2k}}$ iterations of breadth-first traversal to reach clusters whose cardinalities are at most $2k$.
+However, since high cardinality clusters occur only at low depths in the tree and there are few such clusters, the length of $Q$, and thus the cost of QuickSelect, for those iterations is relatively low.
 
-After the first $\log{\frac{n}{2k}}$ iterations, we are left with clusters in $Q$ whose cardinalities are at most $2k$.
-At this point, based on our assumption, we have that the length of $Q$ is close to $2d$.
-Therefore, the asymptotic complexity of the Sieve Algorithm is $\mathcal{O} \left( k \cdot \log(\frac{n}{d}) \right)$.
+After the first $\log{\frac{\mathcal{N}_{\hat{r}}(X)}{2k}}$ iterations, we are left with clusters in $Q$ whose cardinalities are at most $2k$.
+At this point, based on our assumptions about LFD, we have that the length of $Q$ is close to $2d$.
+Therefore, the asymptotic complexity of the Breadth-First Sieve Algorithm is $\mathcal{O} \left( k \cdot \log(\frac{\mathcal{N}_{\hat{r}}(X)}{d}) \right)$.
 
-\subsubsection{Greedy Sieve}
-\label{subsubsec:methods:knn:greedy-sieve}
+\subsubsection{Depth-First Sieve}
+\label{sec:methods:knn-search:depth-first-sieve}
 
@@
 $Q$ starts containing only the root cluster while $H$ starts empty.
-
 So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
@@
 
-This algorithm is described in Algorithm~\ref{alg:methods:greedy-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$ (i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$).
+This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
@@
 \begin{algorithm} 
-\caption{Greedy Sieve($root$, $q$, $k$)} 
-\label{alg:methods:greedy-sieve} 
+    \caption{Depth-First Sieve($root$, $q$, $k$)} 
+    \label{alg:methods:depth-first-sieve} 
     \begin{algorithmic}
@@
 
-\subsubsection{Complexity of Greedy Sieve}
-\label{paragraph:methods:greedy-sieve-complexity}
+Note that this algorithm is not a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
+Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
 
-TODO: Analyze complexity of Greedy Sieve.
+\subsubsection{Complexity of Depth-First Sieve}
+\label{sec:methods:knn-search:depth-first-sieve-complexity}
 
-\subsection {Auto-Tuning}
-\label{subsec:methods:auto-tuning}
+TODO: Analyze complexity of Depth-First Sieve.
+
+\subsection{Auto-Tuning}
+\label{sec:methods:auto-tuning}
 
@@
 We start by taking the center of every cluster at a low depth, e.g. 10, in the cluster tree as a query.
-This gives us a small set of queries which are representative of the dataset.
-Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{subsec:methods:knn-search}.
-We select the fastest over all the queries as the optimal algorithm for the dataset and value of $k$.
+This gives us a small set of queries which is a representative sample of the dataset.
+Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{sec:methods:knn-search}.
+We select the fastest algorithm over all the queries as the optimal algorithm for that dataset and value of $k$.
 Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
@@
 \subsection{Synthetic Data}
-\label{subsec:methods:synthetic-data}
+\label{sec:methods:synthetic-data}
 
 Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and for that performance to scale sub-linearly with the cardinality of the dataset.
-To test this hypothesis, we use some datasets from the ANN-benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to generate similar datasets with much higher cardinalities.
+To test this hypothesis, we use some datasets from the ANN-benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to generate similar datasets with exponentially larger cardinalities.
 We do the same with a large random dataset of uniformly distributed points in a hypercube.
@@
 Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level and $m$ be a user-specified integer multiplier.
-For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.
-We construct a random vector $r$ of $d$ dimensions in the hypersphere of radius $\epsilon$ centered at the origin.
+For each datum $x \in X$, we create $m - 1$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.
+We construct a random vector $r$ of $d$ dimensions in the hyper-sphere of radius $\epsilon$ centered at the origin.
 We then add $r$ to $x$ to get a new point $x'$.
 Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
-We repeat this process $m$ times for each $x \in X$ to get a new dataset $X'$.
+This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.
 

### 2023-12-27 sections/3-datasets.tex
edits in daatsets and results

@@
 
-\subsection{ANN Benchmark Datasets}
-\label{subsec:ann-benchmarks}
+\subsection{ANN-Benchmark Datasets}
+\label{sec:datasets-and-benchmarks:ann-benchmark-datasets}
 
-We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, along with the appropriate distance function.
-Table~\ref{table:datasets:summary} summarizes the properties of these datasets.
+We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, using the associated distance function.
+Table~\ref{tab:datasets:summary} summarizes the properties of these datasets.
 
-The benchmarks for Fashion-Mnist, Glove-25, and Sift (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
+The benchmarks for Fashion-Mnist, Glove-25, Sift, and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
 This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}.
@@
     \caption{Datasets used in benchmarks.}
-    \label{table:datasets:summary}
+    \label{tab:datasets:summary}
     \vskip 0.15in
@@
                     \hline
-                    Silva            & Levenshtein            & 2,224,640          & 50,000         \\
+                    Silva            & Levenshtein            & 2,224,640          & 3,712 - 50,000         \\
                     \hline
@@
 \subsection{Random Datasets and Synthetic Augmentations}
-\label{subsec:random-datasets}
+\label{sec:datasets-and-benchmarks:random-datasets}
 
-In addition to benchmarks on datasets from the ANN Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{subsec:methods:synthetic-data}.
-In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``scale'' in Table~\ref{table:results:ann-fashion} {\color{red} TODO: merge the tables into one and confirm this reference is still correct after the merge}) increases.
+In addition to benchmarks on datasets from the ANN-Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{sec:methods:synthetic-data}.
+In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Tables~\ref{tab:results:fashion-mnist},~\ref{tab:results:glove-25},~\ref{tab:results:sift},~and~\ref{tab:results:random} {\color{red} TODO: merge the tables into one and confirm this reference is still correct after the merge}) increases.
 
-We also benchmarked on purely random (i.e., not an augmented version of a known dataset) datasets of various cardinalities. 
+We also benchmarked on purely random, i.e., not an augmented version of a known dataset, datasets of various cardinalities. 
 For this Random dataset, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset.
@@
 
-\subsection{Silva}
-\label{subsec:silva}
-To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the Silva dataset.
-Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
-We use Levenshtein distance~\cite{levenshtein1966binary} on unaligned sequences as the distance function for this dataset.
+\subsection{Silva 18S}
+\label{sec:datasets-and-benchmarks:silva-18s}
+
+To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the Silva 18S ribosomal DNA dataset~\cite{10.1093/nar/gks1219}.
+This dataset contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
+We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
+We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
+Since this dataset contains many \emph{redundant}, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition \emph{criterion} in Algorithm~\ref{alg:methods:partition}.
 
 \subsection{Radio ML}
-\label{subsec:radioml}
-To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset.
+\label{sec:datasets-and-benchmarks:radio-ml}
+
+To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset~\cite{oshea2018radioml}.
 The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
-Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each modulation mode and SNR level~\cite{oshea2018radioml}.
+Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
 Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
-Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time series of complex-valued numbers).
-We used a subset of this dataset, containing only the 97,920 samples at 10dB SNR.
+Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
+We used a subset of this dataset, containing 97,920 samples at 10dB SNR and used the other 8,576 samples at 10dB SNR as a hold-out set of queries.
 We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.
@@
 \subsection{Other Algorithms}
-We benchmarked CAKES's algorithms against the state-of-the-art similarity search algorithms HNSW and ANNOY, as well as on FAISS-IVF and two different implementations of linear search.
+\label{sec:datasets-and-benchmarks:other-algorithms}
+
+We benchmarked CAKES's algorithms against the state-of-the-art similarity search algorithms HNSW, ANNOY and FAISS-IVF, and two implementations of linear search.
 In particular, we use FAISS-Flat, a Python implementation of linear search, to provide ground-truth values for recall of HNSW, ANNOY, and FAISS-IVF.
-We use our own Rust implementation of linear search to provide ground-truth values for recall of CAKES's algorithms.
-We plot the throughput of these two implementations of linear search as a reference point for CAKES's algorithms' throughput in Section~\ref{subsec:scaling-behavior-results}.
+We use our own Rust implementation of linear search to provide ground-truth values for calculating recall of CAKES's algorithms.
+We plot the throughput of these two implementations of linear search as a reference point for all other algorithms in Figures~\ref{fig:results:scaling-plots}

### 2023-12-27 sections/4-results.tex
edits in daatsets and results

@@
 \subsection{Local Fractal Dimension of Datasets}
-\label{subsec:lfd-results}
+\label{sec:results:lfd-of-datasets}
 
-Since the time complexity of CAKES algorithms scales with the local fractal dimension of the dataset, we examine the local fractal dimension of each dataset we benchmarked CAKES algorithms on.
+Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
 
-Figure~\ref{fig:results:lfd-plots} plots LFD by depth for fashion-mnist, glove-25, sift, a random dataset, Silva 18S, and RadioML.
+Figure~\ref{fig:results:lfd-plots} the trends in LFD for (the non-augmented versions of) Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
 The horizontal axis denotes the depth in the cluster tree, and the vertical axis denotes the LFD (as calculated by Equation~\ref{eq:methods:lfd-simplified}) of clusters at that depth.
-We plot lines for the 0th, 5th, 25th, 50th, 75th, 95th, and 100th percentiles of LFD.
+We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
 In order to have the plots to best reflect the distribution of LFDs across the entire \emph{dataset}, we compute percentiles counting each cluster as many times as its cardinality.
-In other words, if, for some dataset, the 95th percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
+In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the total points in clusters at depth 40 belong to clusters whose LFD is at most 3.
 
-Figure~\ref{fig:results:fashion-mnist-lfd} shows the LFD by depth for fashion-mnist.
-We observe that until about depth 5, the LFD is low, as the 95th percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
-For depths 15 through 25, we observe that the LFD increases, with the 95th percentile (orange plotted line) slightly less than 6, and the median near 3.
-Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95th percentile is between 3 and 4, and the median is less than 2.
+Figure~\ref{fig:results:fashion-mnist-lfd} shows the LFD by depth for Fashion-Mnist.
+We observe that until about depth 5, the LFD is low, as the 95$^{th}$ percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
+For depths 15 through 25, we observe that the LFD increases, with the 95$^{th}$ percentile (orange plotted line) slightly less than 6, and the median near 3.
+Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95$^{th}$ percentile is between 3 and 4, and the median is less than 2.
 
-Relative to fashion-mnist, glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
-Percentile lines for glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-The 95th percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths. 
-Before depth 25, the 95th percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
+Relative to Fashion-Mnist, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
+Percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
+The 95$^{th}$ percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths. 
+Before depth 25, the 95$^{th}$ percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
 The median hovers near 1.5 for all depths.
 
-Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for sift.
-Until about depth 5, the 95th percentile is between 2 and 6. From about depth 5 through about depth 20, the 95th percentile is greater than 6, even exceeding 8 near depth 10.
-From about depth 20 onward, the 95th percentile is less than 6, and from about depth 30 onward, it is between 3 and 4.
+Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift.
+Until about depth 5, the 95$^{th}$ percentile is between 2 and 6.
+From about depth 5 through about depth 20, the 95$^{th}$ percentile is greater than 6, even exceeding 8 near depth 10.
+From about depth 20 onward, the 95$^{th}$ percentile is less than 6, and from about depth 30 onward, it is between 3 and 4.
 The median reaches its peak of about 5 at around depth 15, but hovers near 2 after about depth 30.
 
-In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift.
+In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as those of Sift.
 We observe that the LFD starts as high as 20 (for all data) at depth 0.
-As shown by the needlepoint shape of the plot, the LFD for all clusters is about the same from depth 0 through depth 5.
+As shown by the needlepoint shape of the plot, the LFD for all clusters has a very narrow spread from depth 0 through depth 5.
 After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth.
-The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
-The LFD of approximately 20 at depth 0 is what we expect for this random dataset.
-To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
+All percentile lines seem to decrease linearly with depth.
+The LFD of approximately 20 at depth 0 (i.e. the root cluster) is what we expect for this random dataset.
+To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.
+the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
 As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
-A consequence of this is that with high probability, a point in the root cluster is a distance greater than half the root radius from the center; $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r) = |X|$, where $X$ is the dataset.
-Given our definition of LFD in Equation~\ref{eq:methods:lfd-simplified}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
+A consequence of this is that, with high probability, every point in the root cluster is farther away from the center than half the radius of the root cluster;
+$B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r)$ contains the entire dataset.
+Given our definition of LFD in Equation~\ref{eq:methods:lfd-original}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
 
 Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-The 95th percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
+The 95$^{th}$ percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
 The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 128.
@@
 Relative to the other datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
-Notably however, for the first approximately 15 depths, the 95th percentile and median LFD are very close to 0.
+Notably however, for the first approximately 15 depths, the 95$^{th}$ percentile and median LFD are very close to 0.
 Both percentiles increase sharply after depth 20, nearly reaching 16.
-LFD remains high even as depth increases, with the 95th percentile fluctuating between about 6 and 15 for depths about 20-40.
+LFD remains high even as depth increases, with the 95$^{th}$ percentile fluctuating between about 6 and 15 for depths about 20-40.
 It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60.
@@
     \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of local fractal dimension and weighted by the cardinalities of the clusters.
-    The last dataset is randomly generated.
-    {\color{red} TODO: Remove the headers from the plots themselves and rely on subcaptions instead.}}
+    The last dataset is randomly generated.}
     \label{fig:results:lfd-plots}
@@
 \subsection{Indexing and Tuning Time}
-\label{subsec:indexing-time-results}
+\label{sec:results:indexing-and-tuning-time}
 
-For each of the ANN benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyperparameters for these indices to achieve the highest possible recall (lower is better).
-Though CAKES has three different algorithms, indexing is the same for each of them, so we can simply report the indexing time for CAKES as a whole.
+For each of the ANN-benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyper-parameters for these indices to achieve the highest possible recall.
+Though CAKES has three different algorithms, they share the same tree and so they have the same indexing time.
+Thus, we show CAKES's indexing time only once.
 
-The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}).
+The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
 The left-most point on each line is on the original dataset without any synthetic augmentation.
@@
 \subsection{Scaling Behavior and Recall}
-\label{subsec:scaling-behavior-results}
+\label{sec:results:scaling-behavior-and-recall}
 
-Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
-Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
+Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of the following ANN-benchmark datasets:
+
+\begin{itemize}
+    \item Fashion-Mnist under Euclidean distance,
+    \item Glove-25 under Cosine distance, and
+    \item Sift under Euclidean distance.
+\end{itemize}
+
+Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
 Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
-The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}).
+For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.
+
+The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{sec:methods:synthetic-data}).
 The left-most point on each line is on the original dataset without any synthetic augmentation.
@@
 Both axes are on a logarithmic scale.
-In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in the Supplement.
+In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k = 100$ can be found in the Supplement.
 For HNSW and ANNOY, we report the recall for each measurement in the plots.
@@
 
-Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality for  Fashion-Mnist, Glove-25, Sift, and a random dataset respectively.
+Tables \ref{tab:results:fashion-mnist}, \ref{tab:results:glove-25}, \ref{tab:results:sift}, and \ref{tab:results:random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random respectively.
 Though the scaling plots above present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
-Given that we redo the auto-tuning process described in Section~\ref{subsec:methods:auto-tuning} for each augmented dataset, these values are an accurate representation of CAKES's performance on each dataset and cardinality multiplier.
-When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.
+We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.
+at each cardinality), and this approach was always able to select the fastest algorithm for each dataset at each cardinality.
+Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the tuned CAKES algorithm.
+When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$ when we consider only three decimal places.
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-Mnist, we observe that as cardinality increases, the CAKES algorithms (GreedySieve in blue, Repeated $\rho$-NN in green, and Sieve in purple) are faster than na\"{i}ve linear search (plotted in orange).
-In particular, we see that GreedySieve and Sieve begin to exhibit sublinear time performance at a cardinality near $10^5$ (i.e., at a multiplier of about {\color{red} add value} times the original dataset's cardinality).
-We also observe that CAKES's algorithms outperform FAISS-IVF (plotted in pink) for all cardinalities larger than the original dataset's cardinality.
-Though state-of-the-art algorithms HNSW (plotted in gray), and ANNOY (plotted in yellow) are faster than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
-For example, at a multiplier as low as eight, HNSW and ANNOY have recall of 0.525 and 0.857 respectively,
+In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-Mnist, we observe that as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than na\"{i}ve linear search (plotted in orange).
+In particular, we see that Depth-First Sieve and Breadth-First Sieve begin to exhibit sub-linear time performance at a cardinality near $10^5$, i.e., at four times the original dataset's cardinality.
+We also observe that CAKES's algorithms outperform FAISS-IVF (plotted in pink) for all multipliers larger than four.
+Alhough algorithms HNSW (in gray), and ANNOY (in yellow) are faster than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:results:fashion-mnist} and that their recalls decrease as the multiplier increases.
+For example, at a multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively,
 and this recall gets worse as the multiplier increases.
 Meanwhile, CAKES exhibits perfect recall.
-Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance with increasing multipliers.
+Amongst the CAKES algorithms, Depth-First Sieve is consistently fastest on this dataset, exhibiting sub-linear time performance with increasing cardinality.
 
-With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at some cardinality.
-Specifically, all three of CAKES' algorithms outperform linear search at a cardinality just slightly more than the original dataset's cardinality, $10^6$.
+With Glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms quickly outperform linear search.
+Specifically, all three of CAKES' algorithms outperform linear search at double the cardinality of the original dataset.
 Similar to the results for Fashion-Mnist, we see that CAKES's algorithms outperform FAISS-IVF for all cardinalities greater than $10^7$.
-We see once again that HNSW and ANNOY are faster than CAKES's algorithms for all cardinality multipliers, CAKES's algorithms exhibit better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
+We see once again that while HNSW and ANNOY are faster than CAKES's algorithms for all multipliers, CAKES's algorithms exhibit near-perfect recall at all multipliers, as shown in Table~\ref{tab:results:glove-25}.
 In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
 In contrast, CAKES exhibits perfect recall at all multipliers.
-On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN is consistently the best-performing CAKES algorithm.
+On this dataset, unlike with Fashion-Mnist, Repeated $\rho$-NN is consistently the fastest CAKES algorithm.
 
-In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the Sift dataset, we observe that while linear search performs best for lower cardinalities, our algorithms begin to outperform our implementation of linear search and FAISS-IVF at a cardinality near $10^7$.
-For this dataset, Sieve, is the fastest CAKES algorithm for all cardinalities.
-Similar to the results with Fashion-Mnist and Glove-25, we observe that while HNSW and ANNOY are faster than CAKES's algorithms for all cardinality multipliers, CAKES has better recall.
-Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and ANNOY have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
+In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the Sift dataset, we observe that while linear search performs best at lower cardinalities, CAKES's algorithms begin to outperform linear search and FAISS-IVF at a cardinality near $10^7$.
+For this dataset, Breadth-First Sieve, is the fastest CAKES algorithm for all cardinalities.
+Similar to the results with Fashion-Mnist and Glove-25, we observe that while HNSW and ANNOY are faster than CAKES's algorithms for all multipliers, while CAKES has perfect recall.
+Table~\ref{tab:results:sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and ANNOY have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
 
-With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms are the slowest out of all algorithms, at all cardinalities. 
+With the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms are the slowest out of all algorithms, at all cardinalities.
 As with the real datasets, HNSW and ANNOY are the fastest of any of the algorithms, and CAKES exhibits perfect recall at all cardinalities.
-HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
-in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.
+HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on any of the real datasets;
+in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:random}.
 
-For Silva and RadioML, we benchmarked only CAKES's algorithms.
+For Silva and RadioML, we benchmarked only CAKES's algorithms because none HNSW, ANNOY or FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
 Due to the massive sizes of these datasets, we took random sub-samples of the dataset with lower cardinalities for our benchmarks, rather than augmented versions of the dataset.
 With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
-Until cardinality near $10^4$, GreedySieve is the fastest CAKES algorithm, but for cardinalities greater than that, Repeated $\rho$-NN is the fastest CAKES algorithm.
-For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
-In particular, throughput of CAKES algorithms is identical within three decimal places for all cardinalities.
+Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
+For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the three CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
+In particular, throughput of CAKES algorithms is identical within three significant figures for all cardinalities.
 
@@
         \includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}
-\subcaption{Glove-25 for $k=10$.  }
+        \subcaption{Glove-25 for $k=10$.}
         \label{fig:results:glove-25-scaling}
@@
         \includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}
-\subcaption{Sift for $k=10$.    }
+        \subcaption{Sift for $k=10$.}
         \label{fig:results:sift-scaling}
@@
         \includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
-\subcaption{RadioML for $k=10$ at SnR = 10dB.
-}
+        \subcaption{RadioML for $k=10$ at SnR = 10dB.}
         \label{fig:results:radioml-scaling}
@@
         \includegraphics[width=0.95\textwidth]{plots/silva-knn-10.png}
-\subcaption{Silva for $k=10$.    }
+        \subcaption{Silva for $k=10$.}
         \label{fig:results:silva-scaling}
@@
         \includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
-\subcaption{A random dataset for $k=10$.
-}
+        \subcaption{A random dataset for $k=10$.}
         \label{fig:results:random-scaling}
     \end{subfigure}%
-
-\caption{Algorithm performance across six datasets, including a randomly-generated dataset.
-In each plot, the horizontal axis represents increasing cardinality (size) of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-For some algorithms, we were not able to take measurements for each cardinality because the index-building took too much memory.
-{\color{red} Explain this better and be more specific} }
+    \caption{Throughput across six datasets, including a randomly-generated dataset.
+    In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e. more then 512GB).}
     \label{fig:results:scaling-plots}
@@
     \caption{Runtime performance (queries per second) and recall on the \textbf{fashion-mnist} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{table:results:ann-fashion}
+    \label{tab:results:fashion-mnist}
     \vskip 0.15in
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
+                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
                     \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall     & QPS & Recall    \\
+                        & QPS                 & Recall & QPS                & Recall & QPS                & Recall& QPS                 & Recall \\
                     \hline
-                    1   & 1.33~$\times10^{4}$ & 0.954 & 2.19~$\times10^{3}$ & 0.950 & 2.13~$\times10^{2}$  & 1.000 & 2.01~$\times10^{3}$ & 1.000* & 2.17~$\times10^{3}$ & 1.000 \\
+                    1   & 1.33~$\times10^{4}$ & 0.954 & 2.19~$\times10^{3}$ & 0.950 & 2.01~$\times10^{3}$ & 1.000* & 2.17~$\times10^{3}$ & 1.000 \\
                     \hline
-                    2   & 1.38~$\times10^{4}$ & 0.803 & 2.12~$\times10^{3}$ & 0.927 & 1.06~$\times10^{2}$  & 1.000 & 9.39~$\times10^{2}$ & 1.000* & 1.14~$\times10^{3}$ & 1.000 \\
+                    2   & 1.38~$\times10^{4}$ & 0.803 & 2.12~$\times10^{3}$ & 0.927 & 9.39~$\times10^{2}$ & 1.000* & 1.14~$\times10^{3}$ & 1.000 \\
                     \hline
-                    4   & 1.66~$\times10^{4}$ & 0.681 & 2.04~$\times10^{3}$ & 0.898 & 5.31~$\times10^{1}$  & 1.000 & 4.61~$\times10^{2}$ & 0.997  & 9.82~$\times10^{2}$ & 1.000 \\
+                    4   & 1.66~$\times10^{4}$ & 0.681 & 2.04~$\times10^{3}$ & 0.898 & 4.61~$\times10^{2}$ & 0.997  & 9.82~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8   & 1.68~$\times10^{4}$ & 0.525 & 1.93~$\times10^{3}$ & 0.857 & 2.66~$\times10^{1}$  & 1.000 & 2.26~$\times10^{2}$ & 0.995  & 1.18~$\times10^{3}$ & 1.000 \\
+                    8   & 1.68~$\times10^{4}$ & 0.525 & 1.93~$\times10^{3}$ & 0.857 & 2.26~$\times10^{2}$ & 0.995  & 1.18~$\times10^{3}$ & 1.000 \\
                     \hline
-                    16  & 1.87~$\times10^{4}$ & 0.494 & 1.84~$\times10^{3}$ & 0.862 & 1.33~$\times10^{1}$  & 1.000 & 1.17~$\times10^{2}$ & 0.991  & 1.20~$\times10^{3}$ & 1.000 \\
+                    16  & 1.87~$\times10^{4}$ & 0.494 & 1.84~$\times10^{3}$ & 0.862 & 1.17~$\times10^{2}$ & 0.991  & 1.20~$\times10^{3}$ & 1.000 \\
                     \hline
-                    32  & 1.56~$\times10^{4}$ & 0.542 & 1.85~$\times10^{3}$ & 0.775 & 6.65~$\times10^{0}$  & 1.000 & 5.91~$\times10^{1}$ & 0.985  & 1.16~$\times10^{3}$ & 1.000 \\
+                    32  & 1.56~$\times10^{4}$ & 0.542 & 1.85~$\times10^{3}$ & 0.775 & 5.91~$\times10^{1}$ & 0.985  & 1.16~$\times10^{3}$ & 1.000 \\
                     \hline
-                    64  & 1.50~$\times10^{4}$ & 0.378 & 1.78~$\times10^{3}$ & 0.677 & 3.32~$\times10^{0}$  & 1.000 & 2.61~$\times10^{1}$ & 0.968  & 1.10~$\times10^{3}$ & 1.000 \\
+                    64  & 1.50~$\times10^{4}$ & 0.378 & 1.78~$\times10^{3}$ & 0.677 & 2.61~$\times10^{1}$ & 0.968  & 1.10~$\times10^{3}$ & 1.000 \\
                     \hline
-                    128 & 1.49~$\times10^{4}$ & 0.357 & 1.66~$\times10^{3}$ & 0.538 & 1.66~$\times10^{0}$  & 1.000 & 1.33~$\times10^{1}$ & 0.964  & 1.04~$\times10^{3}$ & 1.000 \\
+                    128 & 1.49~$\times10^{4}$ & 0.357 & 1.66~$\times10^{3}$ & 0.538 & 1.33~$\times10^{1}$ & 0.964  & 1.04~$\times10^{3}$ & 1.000 \\
                     \hline
-                    256 & --                  & --    & 1.60~$\times10^{3}$ & 0.592 & 8.31~$\times10^{-1}$ & 1.000 & 6.65~$\times10^{0}$ & 0.962  & 1.06~$\times10^{3}$ & 1.000 \\
+                    256 & --                  & --    & 1.60~$\times10^{3}$ & 0.592 & 6.65~$\times10^{0}$ & 0.962  & 1.06~$\times10^{3}$ & 1.000 \\
                     \hline
-                    512 & --                  & --    & 1.83~$\times10^{3}$ & 0.581 & 4.16~$\times10^{-1}$ & 1.000 & 3.56~$\times10^{0}$ & 0.949  & 1.04~$\times10^{3}$ & 1.000 \\
+                    512 & --                  & --    & 1.83~$\times10^{3}$ & 0.581 & 3.56~$\times10^{0}$ & 0.949  & 1.04~$\times10^{3}$ & 1.000 \\
                     \hline
@@
     \caption{Runtime performance (queries per second) and recall on the \textbf{glove-25} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{table:results:ann-glove-25}
+    \label{tab:results:glove-25}
     \vskip 0.15in
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
+                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
                     \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall     & QPS & Recall    \\
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall     & QPS & Recall    \\
                     \hline
-                    1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.78~$\times10^{2}$ & 1.000 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000* \\
+                    1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000* \\
                     \hline
-                    2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.39~$\times10^{2}$ & 1.000 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000* \\
+                    2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000* \\
                     \hline
-                    4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.98~$\times10^{1}$ & 1.000 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000* \\
+                    4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000* \\
                     \hline
-                    8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.49~$\times10^{1}$ & 1.000 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000* \\
+                    8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000* \\
                     \hline
-                    16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.74~$\times10^{1}$ & 1.000 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000* \\
+                    16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000* \\
                     \hline
-                    32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 8.70~$\times10^{0}$ & 1.000 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000* \\
+                    32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000* \\
                     \hline
-                    64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 4.36~$\times10^{0}$ & 1.000 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000* \\
+                    64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000* \\
                     \hline
-                    128 & --                  & --    & --                  & --    & 2.18~$\times10^{0}$ & 1.000 & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000* \\
+                    128 & --                  & --    & --                  & --    & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000* \\
                     \hline
-                    256 & --                  & --    & --                  & --    & 1.07~$\times10^{0}$ & 1.000 & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000* \\
+                    256 & --                  & --    & --                  & --    & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000* \\
                     \hline
-                    % 512 & -- & -- & -- & -- & 0.544 & 1.000 & 4.499 & 0.996 & 0.000 & 0.000 \\
-                    % \hline
                 \end{tabular}
@@
     \caption{Runtime performance (queries per second) and recall on the \textbf{sift} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{table:results:ann-sift}
+    \label{tab:results:sift}
     \vskip 0.15in
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
+                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
                     \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall    & QPS & Recall    \\
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall    & QPS & Recall    \\
                     \hline
-                    1   & 1.93~$\times10^{4}$ & 0.782 & 3.98~$\times10^{3}$ & 0.686 & 7.26~$\times10^{1}$  & 1.000 & 6.98~$\times10^{2}$ & 1.000* & 5.52~$\times10^{2}$ & 1.000 \\
+                    1   & 1.93~$\times10^{4}$ & 0.782 & 3.98~$\times10^{3}$ & 0.686 & 6.98~$\times10^{2}$ & 1.000* & 5.52~$\times10^{2}$ & 1.000 \\
                     \hline
-                    2   & 2.03~$\times10^{4}$ & 0.552 & 3.80~$\times10^{3}$ & 0.614 & 3.62~$\times10^{1}$  & 1.000 & 3.30~$\times10^{2}$ & 1.000* & 2.66~$\times10^{2}$ & 1.000 \\
+                    2   & 2.03~$\times10^{4}$ & 0.552 & 3.80~$\times10^{3}$ & 0.614 & 3.30~$\times10^{2}$ & 1.000* & 2.66~$\times10^{2}$ & 1.000 \\
                     \hline
-                    4   & 2.18~$\times10^{4}$ & 0.394 & 3.69~$\times10^{3}$ & 0.637 & 1.82~$\times10^{1}$  & 1.000 & 1.65~$\times10^{2}$ & 1.000* & 1.43~$\times10^{2}$ & 1.000 \\
+                    4   & 2.18~$\times10^{4}$ & 0.394 & 3.69~$\times10^{3}$ & 0.637 & 1.65~$\times10^{2}$ & 1.000* & 1.43~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8   & 2.48~$\times10^{4}$ & 0.298 & 3.58~$\times10^{3}$ & 0.710 & 9.08~$\times10^{0}$  & 1.000 & 7.72~$\times10^{1}$ & 1.000* & 7.94~$\times10^{1}$ & 1.000 \\
+                    8   & 2.48~$\times10^{4}$ & 0.298 & 3.58~$\times10^{3}$ & 0.710 & 7.72~$\times10^{1}$ & 1.000* & 7.94~$\times10^{1}$ & 1.000 \\
                     \hline
-                    16  & 2.68~$\times10^{4}$ & 0.210 & 3.50~$\times10^{3}$ & 0.690 & 4.54~$\times10^{0}$  & 1.000 & 3.98~$\times10^{1}$ & 1.000* & 8.12~$\times10^{1}$ & 1.000 \\
+                    16  & 2.68~$\times10^{4}$ & 0.210 & 3.50~$\times10^{3}$ & 0.690 & 3.98~$\times10^{1}$ & 1.000* & 8.12~$\times10^{1}$ & 1.000 \\
                     \hline
-                    32  & 2.75~$\times10^{4}$ & 0.193 & 3.44~$\times10^{3}$ & 0.639 & 2.27~$\times10^{0}$  & 1.000 & 2.09~$\times10^{1}$ & 0.999  & 7.81~$\times10^{1}$ & 1.000 \\
+                    32  & 2.75~$\times10^{4}$ & 0.193 & 3.44~$\times10^{3}$ & 0.639 & 2.09~$\times10^{1}$ & 0.999  & 7.81~$\times10^{1}$ & 1.000 \\
                     \hline
-                    64  & --                  & --    & 3.39~$\times10^{3}$ & 0.678 & 1.14~$\times10^{0}$  & 1.000 & 8.87~$\times10^{0}$ & 0.997  & 7.43~$\times10^{1}$ & 1.000 \\
+                    64  & --                  & --    & 3.39~$\times10^{3}$ & 0.678 & 8.87~$\times10^{0}$ & 0.997  & 7.43~$\times10^{1}$ & 1.000 \\
                     \hline
-                    128 & --                  & --    & 3.36~$\times10^{3}$ & 0.643 & 5.66~$\times10^{-1}$ & 1.000 & 4.78~$\times10^{0}$ & 0.993  & 6.80~$\times10^{1}$ & 1.000 \\
+                    128 & --                  & --    & 3.36~$\times10^{3}$ & 0.643 & 4.78~$\times10^{0}$ & 0.993  & 6.80~$\times10^{1}$ & 1.000 \\
                     \hline
@@
     \caption{Runtime performance (queries per second) and recall on a \textbf{random} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{table:results:ann-random}
+    \label{tab:results:random}
     \vskip 0.15in
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
+                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
                     \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-flat}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
                     \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall       & QPS & Recall    & QPS & Recall    \\
+                    &             QPS & Recall        & QPS & Recall      & QPS & Recall    & QPS & Recall    \\
                     \hline
-                    1  & 1.17~$\times10^{4}$ & 0.060 & 4.28~$\times10^{3}$ & 0.028 & 7.16~$\times10^{1}$ & 1.000 & 7.34~$\times10^{2}$ & 1.000* & 5.54~$\times10^{2}$ & 1.000 \\
+                    1  & 1.17~$\times10^{4}$ & 0.060 & 4.28~$\times10^{3}$ & 0.028 & 7.34~$\times10^{2}$ & 1.000* & 5.54~$\times10^{2}$ & 1.000 \\
                     \hline
-                    2  & 1.01~$\times10^{4}$ & 0.048 & 4.04~$\times10^{3}$ & 0.021 & 3.62~$\times10^{1}$ & 1.000 & 3.58~$\times10^{2}$ & 1.000* & 2.69~$\times10^{2}$ & 1.000 \\
+                    2  & 1.01~$\times10^{4}$ & 0.048 & 4.04~$\times10^{3}$ & 0.021 & 3.58~$\times10^{2}$ & 1.000* & 2.69~$\times10^{2}$ & 1.000 \\
                     \hline
-                    4  & 9.12~$\times10^{3}$ & 0.031 & 3.64~$\times10^{3}$ & 0.014 & 1.81~$\times10^{1}$ & 1.000 & 1.90~$\times10^{2}$ & 1.000* & 1.37~$\times10^{2}$ & 1.000 \\
+                    4  & 9.12~$\times10^{3}$ & 0.031 & 3.64~$\times10^{3}$ & 0.014 & 1.90~$\times10^{2}$ & 1.000* & 1.37~$\times10^{2}$ & 1.000 \\
                     \hline
-                    8  & 8.35~$\times10^{3}$ & 0.022 & 3.37~$\times10^{3}$ & 0.013 & 9.04~$\times10^{0}$ & 1.000 & 8.84~$\times10^{1}$ & 1.000* & 5.69~$\times10^{1}$ & 1.000 \\
+                    8  & 8.35~$\times10^{3}$ & 0.022 & 3.37~$\times10^{3}$ & 0.013 & 8.84~$\times10^{1}$ & 1.000* & 5.69~$\times10^{1}$ & 1.000 \\
                     \hline
-                    16 & 8.25~$\times10^{3}$ & 0.008 & 3.17~$\times10^{3}$ & 0.006 & 4.53~$\times10^{0}$ & 1.000 & 4.36~$\times10^{1}$ & 1.000* & 2.61~$\times10^{1}$ & 1.000 \\
+                    16 & 8.25~$\times10^{3}$ & 0.008 & 3.17~$\times10^{3}$ & 0.006 & 4.36~$\times10^{1}$ & 1.000* & 2.61~$\times10^{1}$ & 1.000 \\
                     \hline
-                    32 & --                  & --    & 3.01~$\times10^{3}$ & 0.007 & 2.27~$\times10^{0}$ & 1.000 & 1.72~$\times10^{1}$ & 1.000* & 1.35~$\times10^{1}$ & 1.000 \\
-                    % \hline
-                    % 64 & -- & -- & -- & -- & 1.135 & 1.000 & 9.939 & 1.000 & 0.000 & 0.000 \\
-                    % \hline
-                    % 128 & --& -- & -- & -- & 0.567 & 1.000 & 5.106 & 1.000 & 0.000 & 0.000 \\
+                    32 & --                  & --    & 3.01~$\times10^{3}$ & 0.007 & 1.72~$\times10^{1}$ & 1.000* & 1.35~$\times10^{1}$ & 1.000 \\
                     \hline

### 2023-12-27 sections/5-discussion.tex
minor updates to discussion

@@
 \section{Discussion and Future Work}
-\label{sec:discussion}
+\label{sec:discussion-and-future-work}
 
@@
 This is unsurprising, given that the Random dataset is, by definition, uniformly distributed, while Sift is not.
-On the Random dataset, CAKES's algorithms' throughputs decrease linearly as the multiplier increases, while with Sift, GreedySieve and Sieve both exhibit nearly constant throughput.
+On the Random dataset, CAKES's algorithms' throughputs decrease linearly as the multiplier increases, while with Sift, Depth-First Sieve and Breadth-First Sieve both exhibit nearly constant throughput.
 While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, the throughput decreases much more slowly than it does with the Random dataset.
@@
 On Silva, Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the fastest at the higher multipliers. 
-When we view these results in light of the LFD plots in Section~\ref{subsec:lfd-results}, we realize that Repeated $\rho$-NN appears to be the fastest CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near $2$), such as Glove-25 and Silva.
+When we view these results in light of the LFD plots in Section~\ref{sec:results:lfd-of-datasets}, we realize that Repeated $\rho$-NN appears to be the fastest CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near $2$), such as Glove-25 and Silva.
 This trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly estimate the correct radius for $k$ hits. 
@@
 
-While the fastest CAKES algorithm varies by dataset, we observe that on all three ANN-benchmark datasets (Fashion-mnist, Glove-25, and Sift), GreedySieve and Sieve show nearly constant throughput as the cardinality increases.
-This observation warrants further investigation, but it is especially promising given that GreedySieve and Sieve consistently outperform linear search for high cardinalities.
-Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities supports our use of an auto-tuning function (as discussed in Section~\ref{subsec:methods:auto-tuning}) to select the best algorithm for a given dataset.
+While the fastest CAKES algorithm varies by dataset, we observe that on all three ANN-benchmark datasets (Fashion-Mnist, Glove-25, and Sift), Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as the cardinality increases.
+This observation warrants further investigation, but it is especially promising given that Depth-First Sieve and Breadth-First Sieve consistently outperform linear search for high cardinalities.
+Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities supports our use of an auto-tuning function (as discussed in Section~\ref{sec:methods:auto-tuning}) to select the best algorithm for a given dataset.
 Given trends such as the performance of Repeated $\rho$-NN with low LFD, we suspect that a more sophisticated auto-tuning function could be developed to select the fastest algorithm based on the properties of the dataset.
@@
 Moreover, we observe that CAKES's algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD.
-For example, Sift has much higher LFD than Fashion-mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-mnist and Glove-25, respectively.
+For example, Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-Mnist and Glove-25, respectively.
 For the Random dataset, which has the highest LFD, CAKES's algorithms \emph{never} outperform linear search.

### 2023-12-28 sections/1-introduction.tex
edits in methods
intro updates and citations

@@
 Examples of large datasets include neural network embeddings, genomic databases and time-series data such as radio frequency signals.
-Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image analysis models such as {\color{red} [NAME]~[CITE] and [NAME]~[CITE]} are a common source of neural network embeddings.
+Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image} are a common source of neural network embeddings.
 These embeddings are often high-dimensional, and the sizes of training and inference datasets for such networks are growing exponentially.
@@
 \subsection{Related Works}
-\label{subsec:intoduction:related-works}
+\label{sec:intoduction:related-works}
 
-Recent search algorithms for tackling the exponential growth of data include hierarchical navigable small world networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent search algorithms for tackling the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 
 \subsubsection{HNSW}
-\label{subsubsec:introduction:related-works:hnsw}
+\label{sec:introduction:related-works:hnsw}
 
-HNSW is an approximate $k$-NN search method based on navigable small world (NSW) graphs and skip lists. 
+Hierarchical Navigable Small World networks~\cite{Malkov2016EfficientAR} is an approximate $k$-NN search method based on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
 Similar to the NSW algorithm, HNSW builds a graph of the dataset, but unlike NSW, the graph is multi-layered.
@@
 Search starts at the highest layer and descends to the lowest layer, greedily following a path of edges to the nearest node, until reaching the query point. 
-To improve accuracy, the efSearch hyperparameter can be changed to specify the number of closest nearest neighbors to the query vector to be found at each layer.
+To improve accuracy, the $efSearch$ hyperparameter can be changed to specify the number of closest nearest neighbors to the query vector to be found at each layer.
 
 \subsubsection{FAISS-IVF}
-\label{subsubsec:introduction:related-works:faiss-ivf}
+\label{sec:introduction:related-works:faiss-ivf}
 
-InVerted File indexing (IVF) is a method for approximate $k$-NN search. 
+InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
 The dataset is clustered into a high dimensional Voronoi diagram, and whichever cell the query point falls into is then searched exhaustively.
@@
 \subsubsection{ANNOY}
-\label{subsubsec:introduction:related-works:annoy}
+\label{sec:introduction:related-works:annoy}
 
-Annoy is an approximate $k$-NN search method based on random projection and tree building.
-At each intermediate node of the tree, two points are sampled from the space, and the hyperplane equidistant from them is chosen to divide the space into two subspaces.
+This algorithm~\cite{annoy} is based on random projection and tree building for approximate $k$-NN search.
+At each intermediate node of the tree, two points are randomly sampled from the space, and the hyperplane equidistant from them is chosen to divide the space into two subspaces.
 This process is repeated multiple times to create a forest of trees, and the number of times the process is repeated is a tunable parameter.
-{\color{red} TODO: I'm still unclear what they are doing with these trees.}
+At search time, one can increase the number of trees to be searched to improve recall at the cost of speed.
 
 \subsubsection{Entropy-Scaling Search}
-\label{subsubsec:introduction:related-works:entropy-scaling-search}
+\label{sec:introduction:related-works:entropy-scaling-search}
 

### 2023-12-28 sections/1-introduction.tex
one more citation

@@
 Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
 In this paper, we introduce CAKES (CLAM-Accelerated $K$-nn Entropy-scaling Search).

### 2023-12-28 sections/2-methods.tex
finished reporting indexing time

@@
 
-Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and for that performance to scale sub-linearly with the cardinality of the dataset.
+Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and forS performance to scale sub-linearly with the cardinality of the dataset.
 To test this hypothesis, we use some datasets from the ANN-benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to generate similar datasets with exponentially larger cardinalities.

### 2023-12-28 sections/4-results.tex
finished reporting indexing time

@@
 
-The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
+The plots in Figure~\ref{fig:results:indexing} show the results of these benchmarks.
+The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
 The left-most point on each line is on the original dataset without any synthetic augmentation.
-The vertical axis denotes throughput in queries per second.
+The vertical axis denotes the sum of indexing and tuning time in seconds.
 Both axes are on a logarithmic scale.
+Hereafter, when we refer to the ``indexing time'' of an algorithm, we are implicitly referring to the sum of indexing and tuning time for said algorithm.
 
-With the Fashion-Mnist dataset, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is much lower than that of HNSW, ANNOY, and FAISS-IVF, but higher than that of FAISS-Flat.
+On all datasets, we observe that the indexing time for CAKES increases roughly linearly as cardinality increases.
+HNSW and ANNOY have the slowest indexing times across all the algorithms we benchmarked for each of the four datasets, at each cardinality. 
+On some datasets, HNSW and ANNOY exhibit indexing times which are orders of magnitude slower than that of CAKES.
+FAISS-Flat also exhibits the fastest indexing time on each dataset. 
 This is not surprising, given that FAISS-Flat is a naive linear search algorithm and is not building an index.
 
-{\color{red} TODO: Add a sentence or two about the indexing time for the other datasets.}
+We also highlight some differences in indexing time between different datasets.
+With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is lower than that of FAISS-IVF for all cardinalities.
+With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has lower indexing time than CAKES.
+With Sift, CAKEs's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
 

### 2023-12-28 sections/2-methods.tex
some updates to methods

@@
 
-We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a non-negative real number.
+We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
 A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
@@
 We say that such data are \emph{manifold-constrained}, and have low \emph{local fractal dimension} (LFD).
-CAKES exploits this topological property, i.e. low LFD, to accelerate search.
+In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
+While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
+they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
+
+CAKES exploits the low LFD of such datasets to accelerate search.
 We define LFD at some length scale around a point in the dataset as:
@@
 better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}) and depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
-
-CAKES assumes the manifold hypothesis.
-In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
-While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
-they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
 

### 2023-12-28 sections/1-introduction.tex
added applications for rho nn

@@
 
-{\color{red} TODO: Talk about common uses of $\rho$-NN search, as we do with $k$-NN search above.}
+$\rho$-NN search is also often used in a variety of applications, including finding evolutionarily-related organisms~\cite{budowski2010fragbag}, serving relevant advertisements to users~\cite{zhang2020privacy}, and suggesting corrections to misspelled words~\cite{ukkonen1985algorithms}.
+{\color{red} TODO: Add another sentence here.}
 

### 2023-12-28 sections/4-results.tex
merged table for results

@@
 
-Tables \ref{tab:results:fashion-mnist}, \ref{tab:results:glove-25}, \ref{tab:results:sift}, and \ref{tab:results:random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random respectively.
+Table~\ref{tab:results:qps-and-recall} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random.
 Though the scaling plots above present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
@@
 When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$ when we consider only three decimal places.
+
+{\color{red} TODO: Combine the following four paragraphs into a single, larger paragraph that summarizes the results across all datasets. There is too much repetition here.}
 
@@
 
-\begin{table*}[!t]
+\begin{table*}[h]
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall on the \textbf{fashion-mnist} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:fashion-mnist}
+    \caption{QPS (queries per second) and Recall vs naive linear search on the Fashion-Mnist, Glove-25, Sift and Random datasets.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall}
     \vskip 0.15in
@@
             \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
+                \begin{tabular}{|c|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
                     \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
+                    & \multirow{2}{*}{\textbf{Multiplier}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
+                    & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\\cline{2-10}
                     \hline
-                        & QPS                 & Recall & QPS                & Recall & QPS                & Recall& QPS                 & Recall \\
                     \hline
-                    1   & 1.33~$\times10^{4}$ & 0.954 & 2.19~$\times10^{3}$ & 0.950 & 2.01~$\times10^{3}$ & 1.000* & 2.17~$\times10^{3}$ & 1.000 \\
+                    \multirow{10}{*}{\rotatebox[origin=c]{90}{\textbf{Fashion-Mnist}}}
+                    & 1   & 1.33~$\times10^{4}$ & 0.954  & 2.19~$\times10^{3}$ & 0.950  & 2.01~$\times10^{3}$ & 1.000* & 2.17~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 2   & 1.38~$\times10^{4}$ & 0.803  & 2.12~$\times10^{3}$ & 0.927  & 9.39~$\times10^{2}$ & 1.000* & 1.14~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 4   & 1.66~$\times10^{4}$ & 0.681  & 2.04~$\times10^{3}$ & 0.898  & 4.61~$\times10^{2}$ & 0.997  & 9.82~$\times10^{2}$ & 1.000  \\\cline{2-10}
+                    & 8   & 1.68~$\times10^{4}$ & 0.525  & 1.93~$\times10^{3}$ & 0.857  & 2.26~$\times10^{2}$ & 0.995  & 1.18~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 16  & 1.87~$\times10^{4}$ & 0.494  & 1.84~$\times10^{3}$ & 0.862  & 1.17~$\times10^{2}$ & 0.991  & 1.20~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 32  & 1.56~$\times10^{4}$ & 0.542  & 1.85~$\times10^{3}$ & 0.775  & 5.91~$\times10^{1}$ & 0.985  & 1.16~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 64  & 1.50~$\times10^{4}$ & 0.378  & 1.78~$\times10^{3}$ & 0.677  & 2.61~$\times10^{1}$ & 0.968  & 1.10~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 128 & 1.49~$\times10^{4}$ & 0.357  & 1.66~$\times10^{3}$ & 0.538  & 1.33~$\times10^{1}$ & 0.964  & 1.04~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 256 & --                  & --     & 1.60~$\times10^{3}$ & 0.592  & 6.65~$\times10^{0}$ & 0.962  & 1.06~$\times10^{3}$ & 1.000  \\\cline{2-10}
+                    & 512 & --                  & --     & 1.83~$\times10^{3}$ & 0.581  & 3.56~$\times10^{0}$ & 0.949  & 1.04~$\times10^{3}$ & 1.000  \\
                     \hline
-                    2   & 1.38~$\times10^{4}$ & 0.803 & 2.12~$\times10^{3}$ & 0.927 & 9.39~$\times10^{2}$ & 1.000* & 1.14~$\times10^{3}$ & 1.000 \\
                     \hline
-                    4   & 1.66~$\times10^{4}$ & 0.681 & 2.04~$\times10^{3}$ & 0.898 & 4.61~$\times10^{2}$ & 0.997  & 9.82~$\times10^{2}$ & 1.000 \\
+                    \multirow{9}{*}{\rotatebox[origin=c]{90}{\textbf{Glove-25}}}
+                    & 1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 128 & --                  & --    & --                  & --    & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000* \\\cline{2-10}
+                    & 256 & --                  & --    & --                  & --    & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000* \\
                     \hline
-                    8   & 1.68~$\times10^{4}$ & 0.525 & 1.93~$\times10^{3}$ & 0.857 & 2.26~$\times10^{2}$ & 0.995  & 1.18~$\times10^{3}$ & 1.000 \\
                     \hline
-                    16  & 1.87~$\times10^{4}$ & 0.494 & 1.84~$\times10^{3}$ & 0.862 & 1.17~$\times10^{2}$ & 0.991  & 1.20~$\times10^{3}$ & 1.000 \\
+                    \multirow{8}{*}{\rotatebox[origin=c]{90}{\textbf{Sift}}}
+                    & 1   & 1.93~$\times10^{4}$ & 0.782 & 3.98~$\times10^{3}$ & 0.686 & 6.98~$\times10^{2}$ & 1.000* & 5.52~$\times10^{2}$ & 1.000 \\\cline{2-10}
+                    & 2   & 2.03~$\times10^{4}$ & 0.552 & 3.80~$\times10^{3}$ & 0.614 & 3.30~$\times10^{2}$ & 1.000* & 2.66~$\times10^{2}$ & 1.000 \\\cline{2-10}
+                    & 4   & 2.18~$\times10^{4}$ & 0.394 & 3.69~$\times10^{3}$ & 0.637 & 1.65~$\times10^{2}$ & 1.000* & 1.43~$\times10^{2}$ & 1.000 \\\cline{2-10}
+                    & 8   & 2.48~$\times10^{4}$ & 0.298 & 3.58~$\times10^{3}$ & 0.710 & 7.72~$\times10^{1}$ & 1.000* & 7.94~$\times10^{1}$ & 1.000 \\\cline{2-10}
+                    & 16  & 2.68~$\times10^{4}$ & 0.210 & 3.50~$\times10^{3}$ & 0.690 & 3.98~$\times10^{1}$ & 1.000* & 8.12~$\times10^{1}$ & 1.000 \\\cline{2-10}
+                    & 32  & 2.75~$\times10^{4}$ & 0.193 & 3.44~$\times10^{3}$ & 0.639 & 2.09~$\times10^{1}$ & 0.999  & 7.81~$\times10^{1}$ & 1.000 \\\cline{2-10}
+                    & 64  & --                  & --    & 3.39~$\times10^{3}$ & 0.678 & 8.87~$\times10^{0}$ & 0.997  & 7.43~$\times10^{1}$ & 1.000 \\\cline{2-10}
+                    & 128 & --                  & --    & 3.36~$\times10^{3}$ & 0.643 & 4.78~$\times10^{0}$ & 0.993  & 6.80~$\times10^{1}$ & 1.000 \\
                     \hline
-                    32  & 1.56~$\times10^{4}$ & 0.542 & 1.85~$\times10^{3}$ & 0.775 & 5.91~$\times10^{1}$ & 0.985  & 1.16~$\times10^{3}$ & 1.000 \\
                     \hline
-                    64  & 1.50~$\times10^{4}$ & 0.378 & 1.78~$\times10^{3}$ & 0.677 & 2.61~$\times10^{1}$ & 0.968  & 1.10~$\times10^{3}$ & 1.000 \\
-                    \hline
-                    128 & 1.49~$\times10^{4}$ & 0.357 & 1.66~$\times10^{3}$ & 0.538 & 1.33~$\times10^{1}$ & 0.964  & 1.04~$\times10^{3}$ & 1.000 \\
-                    \hline
-                    256 & --                  & --    & 1.60~$\times10^{3}$ & 0.592 & 6.65~$\times10^{0}$ & 0.962  & 1.06~$\times10^{3}$ & 1.000 \\
-                    \hline
-                    512 & --                  & --    & 1.83~$\times10^{3}$ & 0.581 & 3.56~$\times10^{0}$ & 0.949  & 1.04~$\times10^{3}$ & 1.000 \\
+                    \multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{Random}}}
+                    & 1  & 1.17~$\times10^{4}$ & 0.060 & 4.28~$\times10^{3}$ & 0.028 & 7.34~$\times10^{2}$ & 1.000* & 5.54~$\times10^{2}$ & 1.000 \\\cline{2-10}
+                    & 2  & 1.01~$\times10^{4}$ & 0.048 & 4.04~$\times10^{3}$ & 0.021 & 3.58~$\times10^{2}$ & 1.000* & 2.69~$\times10^{2}$ & 1.000 \\\cline{2-10}
+                    & 4  & 9.12~$\times10^{3}$ & 0.031 & 3.64~$\times10^{3}$ & 0.014 & 1.90~$\times10^{2}$ & 1.000* & 1.37~$\times10^{2}$ & 1.000 \\\cline{2-10}
+                    & 8  & 8.35~$\times10^{3}$ & 0.022 & 3.37~$\times10^{3}$ & 0.013 & 8.84~$\times10^{1}$ & 1.000* & 5.69~$\times10^{1}$ & 1.000 \\\cline{2-10}
+                    & 16 & 8.25~$\times10^{3}$ & 0.008 & 3.17~$\times10^{3}$ & 0.006 & 4.36~$\times10^{1}$ & 1.000* & 2.61~$\times10^{1}$ & 1.000 \\\cline{2-10}
+                    & 32 & --                  & --    & 3.01~$\times10^{3}$ & 0.007 & 1.72~$\times10^{1}$ & 1.000* & 1.35~$\times10^{1}$ & 1.000 \\
                     \hline
@@
 \end{table*}
-
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall on the \textbf{glove-25} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:glove-25}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
-                    \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}} & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-                    \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall     & QPS & Recall    \\
-                    \hline
-                    1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    128 & --                  & --    & --                  & --    & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000* \\
-                    \hline
-                    256 & --                  & --    & --                  & --    & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000* \\
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
-
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall on the \textbf{sift} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:sift}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
-                    \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-                    \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall    & QPS & Recall    \\
-                    \hline
-                    1   & 1.93~$\times10^{4}$ & 0.782 & 3.98~$\times10^{3}$ & 0.686 & 6.98~$\times10^{2}$ & 1.000* & 5.52~$\times10^{2}$ & 1.000 \\
-                    \hline
-                    2   & 2.03~$\times10^{4}$ & 0.552 & 3.80~$\times10^{3}$ & 0.614 & 3.30~$\times10^{2}$ & 1.000* & 2.66~$\times10^{2}$ & 1.000 \\
-                    \hline
-                    4   & 2.18~$\times10^{4}$ & 0.394 & 3.69~$\times10^{3}$ & 0.637 & 1.65~$\times10^{2}$ & 1.000* & 1.43~$\times10^{2}$ & 1.000 \\
-                    \hline
-                    8   & 2.48~$\times10^{4}$ & 0.298 & 3.58~$\times10^{3}$ & 0.710 & 7.72~$\times10^{1}$ & 1.000* & 7.94~$\times10^{1}$ & 1.000 \\
-                    \hline
-                    16  & 2.68~$\times10^{4}$ & 0.210 & 3.50~$\times10^{3}$ & 0.690 & 3.98~$\times10^{1}$ & 1.000* & 8.12~$\times10^{1}$ & 1.000 \\
-                    \hline
-                    32  & 2.75~$\times10^{4}$ & 0.193 & 3.44~$\times10^{3}$ & 0.639 & 2.09~$\times10^{1}$ & 0.999  & 7.81~$\times10^{1}$ & 1.000 \\
-                    \hline
-                    64  & --                  & --    & 3.39~$\times10^{3}$ & 0.678 & 8.87~$\times10^{0}$ & 0.997  & 7.43~$\times10^{1}$ & 1.000 \\
-                    \hline
-                    128 & --                  & --    & 3.36~$\times10^{3}$ & 0.643 & 4.78~$\times10^{0}$ & 0.993  & 6.80~$\times10^{1}$ & 1.000 \\
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
-
-\begin{table*}[!t]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Runtime performance (queries per second) and recall on a \textbf{random} dataset. 1.000* denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:random}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{sc}
-                \begin{tabular}{|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
-                    \hline
-                    \textbf{Multiplier}  & \multicolumn{2}{|c|}{\textbf{hnsw}}  & \multicolumn{2}{|c|}{\textbf{annoy}} & \multicolumn{2}{|c|}{\textbf{faiss-ivf-flat}}  & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
-                    \hline
-                    &             QPS & Recall        & QPS & Recall      & QPS & Recall    & QPS & Recall    \\
-                    \hline
-                    1  & 1.17~$\times10^{4}$ & 0.060 & 4.28~$\times10^{3}$ & 0.028 & 7.34~$\times10^{2}$ & 1.000* & 5.54~$\times10^{2}$ & 1.000 \\
-                    \hline
-                    2  & 1.01~$\times10^{4}$ & 0.048 & 4.04~$\times10^{3}$ & 0.021 & 3.58~$\times10^{2}$ & 1.000* & 2.69~$\times10^{2}$ & 1.000 \\
-                    \hline
-                    4  & 9.12~$\times10^{3}$ & 0.031 & 3.64~$\times10^{3}$ & 0.014 & 1.90~$\times10^{2}$ & 1.000* & 1.37~$\times10^{2}$ & 1.000 \\
-                    \hline
-                    8  & 8.35~$\times10^{3}$ & 0.022 & 3.37~$\times10^{3}$ & 0.013 & 8.84~$\times10^{1}$ & 1.000* & 5.69~$\times10^{1}$ & 1.000 \\
-                    \hline
-                    16 & 8.25~$\times10^{3}$ & 0.008 & 3.17~$\times10^{3}$ & 0.006 & 4.36~$\times10^{1}$ & 1.000* & 2.61~$\times10^{1}$ & 1.000 \\
-                    \hline
-                    32 & --                  & --    & 3.01~$\times10^{3}$ & 0.007 & 1.72~$\times10^{1}$ & 1.000* & 1.35~$\times10^{1}$ & 1.000 \\
-                    \hline
-                \end{tabular}
-            \end{sc}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}

### 2023-12-29 sections/3-datasets.tex
slight modifications to datasets

@@
 In addition to benchmarks on datasets from the ANN-Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{sec:methods:synthetic-data}.
-In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Tables~\ref{tab:results:fashion-mnist},~\ref{tab:results:glove-25},~\ref{tab:results:sift},~and~\ref{tab:results:random} {\color{red} TODO: merge the tables into one and confirm this reference is still correct after the merge}) increases.
+In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Table~\ref{tab:results:qps-and-recall}) increases.
 
-We also benchmarked on purely random, i.e., not an augmented version of a known dataset, datasets of various cardinalities. 
-For this Random dataset, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset.
+We also benchmarked on purely random (i.e., not an augmented version of a known dataset), datasets of various cardinalities.
+For this, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset; hereafter, we refer to the random dataset with cardinality 1,000,000 and dimensionality 128 as ``Random.''
 This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of the CAKES' algorithms.

### 2023-12-29 sections/5-discussion.tex
updates to discussion

@@
 
-CAKES's algorithms are designed to be most effective when the dataset upholds the manifold hypothesis, i.e.
-when the dataset is constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
-As a consequence, these algorithms do not perform well on datasets with random distributions, because such datasets do not exhibit a manifold structure.
+CAKES's algorithms are designed to be most effective when the data uphold the manifold hypothesis, i.e.
+when the data are constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
+As a consequence, these algorithms do not perform well on data with random distributions, because of the absence of a manifold structure.
 In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each of the datasets we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
@@
 This is unsurprising, given that the Random dataset is, by definition, uniformly distributed, while Sift is not.
-On the Random dataset, CAKES's algorithms' throughputs decrease linearly as the multiplier increases, while with Sift, Depth-First Sieve and Breadth-First Sieve both exhibit nearly constant throughput.
+On the Random dataset, CAKES's algorithms' speed decreases linearly as the multiplier increases, while with Sift, Depth-First Sieve and Breadth-First Sieve both exhibit nearly constant throughput.
 While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, the throughput decreases much more slowly than it does with the Random dataset.
@@
 On Silva, Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the fastest at the higher multipliers.
-When we view these results in light of the LFD plots in Section~\ref{sec:results:lfd-of-datasets}, we realize that Repeated $\rho$-NN appears to be the fastest CAKES algorithm on datasets where a large proportion of the dataset has a very low LFD (i.e., near $2$), such as Glove-25 and Silva.
+When we view these results in light of the LFD plots in Figure~\ref{fig:results:lfd-plots}, we realize that Repeated $\rho$-NN appears to be the fastest CAKES algorithm on datasets where a large proportion of the data have a very low LFD, such as Glove-25 and Silva.
 This trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly estimate the correct radius for $k$ hits.
@@
 Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
+
 We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
-
-% When we descend some levels in the cluster tree, it takes fewer bits to encode each point in terms of the center because the radii decrease.
-% However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
-% This presents the following tradeoff:
-% \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
-% In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
+Such an encoding would allow us to store the data in a compressed format, and to perform search on the compressed data without decompressing it.
+This would allow us to perform fast search on datasets that are too large to fit in memory.
+An exploration of compression and compressed search is an avenue for future work.
 
@@
 
-CLAM and CAKES are implemented in Rust and their source code is available under an MIT license at https://github.com/URI-ABD/clam.
+CLAM and CAKES are implemented in Rust and the source code is available under an MIT license at https://github.com/URI-ABD/clam.

### 2023-12-29 sections/4-results.tex
snark
snark x2
tightning results

@@
 
+{\color{red} TODO: remove all red TODOs.}
+
+{\color{red} TODO: Add more red TODOs.}
+
+{\color{red} TODO: Figure out if we should remove all red TODOs or add more red TODOS}
+
 On all datasets, we observe that the indexing time for CAKES increases roughly linearly as cardinality increases.
@@
 Table~\ref{tab:results:qps-and-recall} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random. 
-Though the scaling plots above present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
+Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
 We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.
@@
 
-{\color{red} TODO: Combine the following four paragraphs into a single, larger paragraph that summarizes the results across all datasets. There is too much repetition here.}
+Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
+Though we observe this trend in the algorithms' scaling behavior, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset.
+For Fashion-Mnist, CAKES begins exhibiting sub-linear performance starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
+We also observe differences in which of the three CAKES algorithms is fastest for a given dataset.
+For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve. 
+Additionally, on all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \emph{constant} in the cardinality of the dataset. 
+With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling. 
+We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, one of CAKES's algorithms becomes faster than FAISS-IVF (in pink).
+For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$. 
+On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
+While recall for CAKES's algorithms does not degrade with cardinality when used with distance functions which are metrics, recall for HNSW and ANNOY does degrade with cardinality. 
+At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
 
-In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for Fashion-Mnist, we observe that as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than na\"{i}ve linear search (plotted in orange).
-In particular, we see that Depth-First Sieve and Breadth-First Sieve begin to exhibit sub-linear time performance at a cardinality near $10^5$, i.e., at four times the original dataset's cardinality.
-We also observe that CAKES's algorithms outperform FAISS-IVF (plotted in pink) for all multipliers larger than four.
-Alhough algorithms HNSW (in gray), and ANNOY (in yellow) are faster than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:results:fashion-mnist} and that their recalls decrease as the multiplier increases.
-For example, at a multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively,
-and this recall gets worse as the multiplier increases.
-Meanwhile, CAKES exhibits perfect recall.
-Amongst the CAKES algorithms, Depth-First Sieve is consistently fastest on this dataset, exhibiting sub-linear time performance with increasing cardinality.
-
-With Glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms quickly outperform linear search.
-Specifically, all three of CAKES' algorithms outperform linear search at double the cardinality of the original dataset.
-Similar to the results for Fashion-Mnist, we see that CAKES's algorithms outperform FAISS-IVF for all cardinalities greater than $10^7$.
-We see once again that while HNSW and ANNOY are faster than CAKES's algorithms for all multipliers, CAKES's algorithms exhibit near-perfect recall at all multipliers, as shown in Table~\ref{tab:results:glove-25}.
-In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
-In contrast, CAKES exhibits perfect recall at all multipliers.
-On this dataset, unlike with Fashion-Mnist, Repeated $\rho$-NN is consistently the fastest CAKES algorithm.
-
-In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the Sift dataset, we observe that while linear search performs best at lower cardinalities, CAKES's algorithms begin to outperform linear search and FAISS-IVF at a cardinality near $10^7$.
-For this dataset, Breadth-First Sieve, is the fastest CAKES algorithm for all cardinalities.
-Similar to the results with Fashion-Mnist and Glove-25, we observe that while HNSW and ANNOY are faster than CAKES's algorithms for all multipliers, while CAKES has perfect recall.
-Table~\ref{tab:results:sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and ANNOY have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.
-
-With the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms are the slowest out of all algorithms, at all cardinalities.
-As with the real datasets, HNSW and ANNOY are the fastest of any of the algorithms, and CAKES exhibits perfect recall at all cardinalities.
-HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on any of the real datasets;
+In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms are the slowest out of all algorithms, at all cardinalities.
+As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
+HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
 in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:random}.

### 2023-12-29 sections/4-results.tex
minor tweaks to results

@@
 The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
-The left-most point on each line is on the original dataset without any synthetic augmentation.
+The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
 The vertical axis denotes the sum of indexing and tuning time in seconds.
@@
 The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{sec:methods:synthetic-data}).
-The left-most point on each line is on the original dataset without any synthetic augmentation.
+The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation..
 The vertical axis denotes throughput in queries per second.
@@
 Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
-Though we observe this trend in the algorithms' scaling behavior, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset.
+Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset.
 For Fashion-Mnist, CAKES begins exhibiting sub-linear performance starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
-We also observe differences in which of the three CAKES algorithms is fastest for a given dataset.
+Which one of the three CAKES algorithms is fastest also differs by dataset.
 For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve. 
@@
 With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling. 
-We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, one of CAKES's algorithms becomes faster than FAISS-IVF (in pink).
+We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
 For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$. 
 On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
-While recall for CAKES's algorithms does not degrade with cardinality when used with distance functions which are metrics, recall for HNSW and ANNOY does degrade with cardinality. 
+While recall for CAKES's algorithms does not degrade with cardinality, recall for HNSW and ANNOY does degrade with cardinality. 
 At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift. 
+In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (because the distance function used with these datasets is a metric), and near-perfect recall on Glove-25 (because the Cosine distance function is not a metric).
 

### 2023-12-29 sections/4-results.tex
fix reference

@@
 
+{\color{red} \sout{TODO: Eliminate all non-red TODOs.}}
+
 On all datasets, we observe that the indexing time for CAKES increases roughly linearly as cardinality increases.
@@
 HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
-in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:random}.
+in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:qps-and-recall}.
 
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+Note that for RadioML, all three algorithms performed nearly identically, so distinct plotlines are not visible.
     For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e. more then 512GB).}

### 2023-12-31 sections/1-introduction.tex
squashed all red TODOs

--- 2023-12-28 sections/1-introduction.tex
@@
 This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
-Often dubbed ``the Big Data explosion,'' this phenomenon has led to a need for better algorithms for analyzing such datasets.
+Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms for analyzing such datasets.
 
@@
 Given some measure of similarity between data points, e.g.
-a distance function, there are two common definitions similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
+a distance function, there are two common definitions of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
 $k$-NN search aims to find the $k$ most similar points to a query point, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query point.
 
-Previous works have used the term \emph{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \emph{approximate} for search algorithms which do not exhibit perfect recall when compared to a na\"{i}ve linear search.
-In contrast, an \emph{exact} search algorithm exhibits perfect recall.
+Previous works have used the term \textit{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \textit{approximate} for search algorithms which do not exhibit perfect recall when compared to a na\"{i}ve linear search.
+In contrast, an \textit{exact} search algorithm exhibits perfect recall.
 
@@
 While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
-Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \emph{exact} search remains.
+Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains.
 
@@
 
-{\color{red} TODO: Talk about common uses of $\rho$-NN search, as we do with $k$-NN search above.}
+$\rho$-NN search also has a variety of applications.
+For example, one could search for all genomes within a maximum edit distance of a query genome to find evolutionarily-related organisms~\cite{budowski2010fragbag}.
+One could also search for all words within a maximum edit distance of a misspelled word to suggest corrections~\cite{ukkonen1985algorithms}.
+Given an advertisement and a database of user profiles, one could search for all users whose profiles are ``similar enough'' to serve the advertisement to~\cite{zhang2020privacy}.
+GPS and other location-based services use $\rho$-NN search to find nearby points of interest~\cite{zhang2020privacy} to a user's location.
+In all of these cases, approximate $\rho$-NN search may be sufficient, but exact $\rho$-NN search is preferable.
 
 This paper focuses on three novel algorithms for exact $k$-NN search.
-We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS, as well as improved genericity across distance functions.
-We provide a comparison of CAKES to several state-of-the-art algorithms in similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
+We provide a comparison of CAKES to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
 We also benchmark CAKES on a genomic dataset, the Silva 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
@@
 InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
-The dataset is clustered into a high dimensional Voronoi diagram, and whichever cell the query point falls into is then searched exhaustively.
-The number of cells created is governed by the $n_{list}$ parameter. 
+The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point into is then searched exhaustively.
+The number of cells used is governed by the $n_{list}$ parameter. 
 Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
-To mitigate accuracy issues caused by a query point falling near a cell boundary, IVF has a tunable parameter $n_{probe}$, which specifies the number of cells to search.
+To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$, which specifies the number of additional adjacent or nearby cells to search.
 
@@
 This search paradigm exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \emph{metric entropy} and \emph{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
+Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \textit{metric entropy} and \textit{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
 In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.

### 2023-12-31 sections/2-methods.tex
fix reference
squashed all red TODOs

@@
 In this manuscript, we are primarily concerned with $k$-NN search in a finite-dimensional space.
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \emph{point} or \emph{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the neural-network embedding of an image, the genome of an organism, a measurement of a radio frequency signal, etc.
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \textit{point} or \textit{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the neural-network embedding of an image, the genome of an organism, a measurement of a radio frequency signal, etc.
 
-We define a \emph{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
+We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
 A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
@@
 In addition to these constraints, if the distance function obeys the triangle inequality (i.e.
-$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \emph{distance metric}.
+$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
 Under distance metrics, all search algorithms in CAKES are exact.
 For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality (e.g.
-consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on a Cartesian plane).
+consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
 
@@
 CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, i.e., high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
-We say that such data are \emph{manifold-constrained}, and have low \emph{local fractal dimension} (LFD).
+We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
-While we sometimes use Euclidean notions, such as voids, volumes and proximity, to talk about the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
+While we sometimes use Euclidean notions, such as voids, volumes and proximity to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
 they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
@@
 where $B_X(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
-We stress that this concept differs from the \emph{embedding dimension} of a dataset.
-To illustrate the difference, consider the Silva 18S rRNA dataset which contains which contains genomes with unaligned lengths of up to 3,718 base pairs.
-Hence, the \emph{embedding dimension} of this dataset is at least 3,718.
-However, due to physical constraints (namely, the laws that govern RNA replication and mutation), the data are constrained to a lower-dimensional manifold within this embedding space.
+Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases.
+
+We stress that this concept differs from the \textit{embedding dimension} of a dataset.
+To illustrate the difference, consider the Silva 18S rRNA dataset which contains which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
+Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
+However, due to physical constraints (namely, biological evolution and the chemistry of RNA replication and mutation processes), the data are constrained to a lower-dimensional manifold within this embedding space.
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
@@
 
-We define a \emph{cluster} as a set of points with a \emph{center} and a \emph{radius}.
-The \emph{center} is the geometric median of the points (or a smaller sample of the points) in the \emph{cluster}, and so it is a real data point.
-The \emph{radius} is the maximum distance from the \emph{center} to any point in the \emph{cluster}.
+We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
+The \textit{center} is the geometric median of the points (or a smaller sample of the points) in the \textit{cluster}, and so it is a real data point.
+The \textit{radius} is the maximum distance from the \textit{center} to any point in the \textit{cluster}.
 Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
-Hereafter, when we refer to LFD, it is estimated at the scale of the cluster radius and half that radius.
+Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
+As a consequence, a cluster can be a subset of the metric ball at the same center and radius, i.e. $C(c, r) \subseteq B_X(c, r)$.
+
+Hereafter, when we refer to the LFD of a \textit{cluster}, it is estimated at the length scale of the cluster radius and half that radius.
 This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as:
@@
 \begin{equation}
-    \log_2 \left( \frac{|B_X(q, r)|}{|B_X(q, r/2)|} \right)
+    \log_2 \left( \frac{|C(c, r)|}{|C(c, \frac{r}{2})|} \right)
     \label{eq:methods:lfd-simplified}
@@
 
-The \emph{metric entropy} ($\mathcal{N}_{r}(X)$) for some radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover the data.
+where $|C(c, r)|$ is the cardinality of the cluster $C$ with center $c$ and radius $r$, and $C(c, \frac{r}{2})$ is the set of points in $C$ which are no more than a distance $\frac{r}{2}$ from $c$, i.e.
+$C(c, \frac{r}{2}) = \{p : p \in C \land f(c, p) \leq \frac{r}{2}\}$.
+
+The \textit{metric entropy} $\mathcal{N}_{r}(X)$ for some radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover the data.
 In this paper, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
@@
 For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
-Using these distances, we compute the \emph{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
-This geometric median is the \emph{center} of $C$.
+Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
+This geometric median is the \textit{center} of $C$.
 
-The \emph{radius} of $C$ is the maximum distance from the \emph{center} to any other point in $C$.
-The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \emph{left pole} and the point which is furthest from left pole is designated the \emph{right pole}.
+The \textit{radius} of $C$ is the maximum distance from the \textit{center} to any other point in $C$.
+The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from left pole is designated the \textit{right pole}.
 
-We then partition the cluster into a \emph{left child} and a \emph{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
+We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
 Without loss of generality, we assign to the left child those points which are equidistant from the two poles.
@@
 
-        \STATE $seeds \Leftarrow \lfloor \sqrt{|C|} \rfloor$ random points from $C$
+        \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
         \STATE $c \Leftarrow$ geometric median of $seeds$
@@
 With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
-Assuming a balanced tree, and thus $\mathcal{O}(\log(n))$ depth, this approach had a memory overhead of $\mathcal{O}(n\log(n))$.
+Assuming a balanced tree, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.
 
 One may improve this memory overhead to $\mathcal{O}(n)$ by only storing indices at the leaf clusters.
-This approach, however, introduces a time cost of $\mathcal{O}(n\log(n))$ whenever we need to find the indices for a non-leaf cluster (because it requires a traversal of the subtree rooted at that cluster).
+This approach, however, introduces a time cost of $\mathcal{O}(n \log n)$ whenever we need to find the indices for a non-leaf cluster (because it requires a traversal of the subtree rooted at that cluster).
 
 In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
-Then, within each cluster, we need only store its \emph{cardinality} and an \emph{offset} to access the its points from the dataset.
+Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access the its points from the dataset.
 The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
@@
 
-With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers significant improvements over the approach used in CHESS.
+With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers significant improvements at search-time over the approach used in CHESS.
 
@@
 % While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:partition}), unfortunately, this assumption is incorrect. 
-% Fortunately, we \emph{can} make some guarantees about the scaling behavior of cluster radii;
+% Fortunately, we \textit{can} make some guarantees about the scaling behavior of cluster radii;
 % in particular, we prove in this section that cluster radii will stop increasing after at most $d$ partitions, where $d$ is the fractal dimensionality of the dataset.
@@
 % Let $2R$ be the maximum pairwise distance among the points in the dataset. 
-% We choose the axes such that the two points that are $2R$ apart lie \emph{along one of the axes}. 
+% We choose the axes such that the two points that are $2R$ apart lie \textit{along one of the axes}. 
 % Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset. 
-% In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance \emph{along every axis}. 
+% In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance \textit{along every axis}. 
 % Such a distribution would also produce a balanced cluster tree.
@@
 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $\forall x \in \textbf{X} \setminus H, \ f(x, q) > \max \{ f(y, q): y \in H \}$;
-in other words, $k$-NN finds the $k$ closest points to $q$ in $\textbf{X}$.
-We also have the $\rho$-NN search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
-that is, find all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$.
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, we define the $k$-NN search problem which aims to find the $k$ closest points to $q$ in $\textbf{X}$;
+i.e.
+the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
+We also have the $\rho$-NN search problem, which aims to find the all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$;
+i.e. the set $H = B_X(q, \rho) = \{x \in \textbf{X}: f(q, x) \leq \rho \}$.
 
-Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
+Given a cluster $C$, let $c$ be its center and $r$ be its radius.
 Our $\rho$-NN and $k$-NN search algorithms make use of the following properties, as illustrated in Figure~\ref{fig:methods:deltas}.
@@
 \begin{itemize}
-    \item $\delta = f(q, c)$ is the distance from the query to the cluster center $c$.
-    \item $\delta^{+} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-    \item $\delta^{-} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+    \item $\delta = f(q, c)$, the distance from the query to the cluster center $c$.
+    \item $\delta^{+} = \delta + r$, the distance from the query to the theoretically farthest point in $C$.
+    \item $\delta^{-} = \text{max}(0, \delta - r)$, the distance from the query to the theoretically closest point in $C$.
 \end{itemize}
@@
 
-We define a \emph{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \emph{unique} point).
+We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \textit{unique} point).
 A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
@@
 Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
-Now let $q'$ be the projection of $q$ to $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$ and $d$ be the distance from $q'$ to $m$.
+Now let $q'$ be the projection of $q$ to $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
 As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $q'$ is too far from $m$, i.e.
@@
 
-To perform $\rho$-NN search, we first perform a coarse \emph{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
-Then, for all such clusters, we perform a finer-grained \emph{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points in which are no more than a distance $\rho$ from the query.
+To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
+Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points in which are no more than a distance $\rho$ from the query.
 
@@
 
-            \STATE $[l, r]$ $\Leftarrow$ \emph{pole}s of $C$
-            \STATE $[L, R]$ $\Leftarrow$ \emph{children} of $C$
+            \STATE $[l, r]$ $\Leftarrow$ \textit{pole}s of $C$
+            \STATE $[L, R]$ $\Leftarrow$ \textit{children} of $C$
 
@@
     \Bigg(
-        \underbrace{\log_2 \mathcal{N}_{\hat{r}}(X)}_{\textrm{metric entropy}} +
-        \overbrace{ \left| B_X(q, \rho) \right|}^{\textrm{output size}}
-        \underbrace{ \left( 1 + \frac{2 \cdot \hat{r}}{ \rho} \right) ^ d}_{\textrm{scaling factor}}
+        \underbrace{
+            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+        }_{\textrm{tree search}}
+        \ + \ 
+        \underbrace{
+            \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
+            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
+        }_{\textrm{leaf search}}
     \Bigg)
@@
 
-Here, $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query in a region of radius $\rho + \hat{r}$.
+where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
+
 From Algorithm~\ref{alg:methods:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-NN search performance scales linearly with the size of the output set and exponentially with the LFD around the query.
@@
 
-In these algorithms, we use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far, and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
+In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure which stores the closest points to the query found so far, and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
 
@@
 
-In this algorithm, we perform $\rho$-nearest-neighbors search starting with a small search radius, and repeatedly increasing the radius until $|H| \geq k$.
+In this algorithm, we perform $\rho$-NN search starting with a small search radius, and repeatedly increasing the radius until $|H| \geq k$.
 
-Let the search radius, $r$, be equal to the radius of the root cluster divided by the cardinality of the dataset.
+Let the search radius $r$ be equal to the radius of the root cluster divided by the cardinality of the dataset.
 We perform tree-search with radius $r$.
-If no clusters are returned, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
-Let $S$ be the set of clusters returned by the first tree-search which returns at least one cluster.
+If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
+Let $Q$ be the set of clusters returned by the first tree-search which returns at least one cluster.
 
-Now, so long as $\sum_{C \in S} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
 In particular, we increase the radius by a factor of
@@
 \begin{equation}
-    \min \left(2, \left( {\frac{k}{\sum_{C \in S} |C|}} \right)^{\mu^{-1}} \right)
+    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)
     \label{eq:methods:repeated-rnn-factor}
@@
 
-Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
-Additionally, when the LFD at the radius scale from the previous iteration is low, this suggests that the data are relatively concentrated in that region.
+Intuitively, the factor by which we increase the radius should be \textit{inversely} related to the number of points found so far. 
+When the LFD at the radius scale from the previous iteration is high, this suggests that the data are densely populated in that region.
+Thus, a small increase in the radius would likely encounter many more points, so a smaller radial increase would suffice to find $k$ neighbors.
+Conversely, when the LFD at the radius scale from the previous iteration is low, this suggests that the data are sparsely populated in that region.
 In such a region, a small increase in the radius would likely encounter vacant space, so a larger radial increase is needed.
-Thus, the factor of radius increase should also be \emph{inversely} related to the LFD.
+Thus, the factor of radius increase should also be \textit{inversely} related to the LFD.
 
-Once $\sum_{C \in S} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we can stop increasing the radius.
+Once $\sum_{C \in Q} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we can stop increasing the radius.
 We perform $\rho$-NN search with this radius and return the $k$ nearest neighbors.
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{repeated-$\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
+    \caption{Repeated $\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
     \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
@@
         \STATE $r \Leftarrow$ $\frac{r}{|root|}$
-        \STATE $Q \Leftarrow \emptyset$, a set of clusters
 
-        \WHILE {$Q = \emptyset$}
+        \ALOOP{}
             \STATE $Q \Leftarrow$ tree-search($root$, $q$, $r$)
+            \IF{$Q \neq \emptyset$}
+                \STATE \textbf{break}
+            \ENDIF
             \STATE $r \Leftarrow 2 \cdot r$
-        \ENDWHILE
+        \ENDALOOP
 
@@
 
-We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{tree-search} and \emph{leaf-search} separately.
+We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \textit{tree-search} and \textit{leaf-search} separately.
 Tree-search refers to the process of identifying clusters which have overlap with the query ball, or in other words, clusters which might contain one of the $k$ nearest neighbors. 
@@
 In ~\cite{ishaq2019clustered}, we showed that the complexity of tree-search is $\mathcal{O}(\log\mathcal{N}_{\hat{r}}(X))$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset $X$ at a radius $\hat{r}$.
-To adjust this bound for Repeated $\rho$-NN, we must estimate the number of iterations of tree-search needed to find a radius that guarantees at least $k$ neighbors.
+To adjust this bound for Repeated $\rho$-NN, we must estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
 
 Based on the assumption that the LFD near the query does not differ significantly from that of nearby clusters, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
-one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
-Since this is a constant factor, complexity of tree search for Repeated $\rho$-NN is the same as that of $\rho$-NN.
+one iteration to find at least one cluster, and the one more to find enough the $k$ neighbors.
+Since this is a constant factor, complexity of tree search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e. $\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
 
-To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the number of points in the clusters returned by tree search.
-Since we must examine each point in each such cluster, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$. 
-Let $\rho_k$ be the distance from the query to the $k$th nearest neighbor.
+To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree search.
+Since we must examine every point in each such clusters, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$. 
+Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
 Then, we see that $Q$ is expected to be the set of clusters which overlap with a ball of radius $\rho_k$ around the query.
@@
 \begin{equation*}
-    \sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(1+ \frac{2\hat{r}}{\rho_k} \right)^d
+    \sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d
 \end{equation*}
@@
 \begin{equation*}
-    \sum_{C \in S} |C| \leq \gamma k \left( 1+ \frac{2\hat{r}}{\rho_k} \right)^d
+    \sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d
 \end{equation*}
@@
 
-We let $\hat{d}$ be the mean LFD of the clusters in $Q$.
-While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \emph{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \emph{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
-
+We let $\hat{d}$ be the (harmonic) mean LFD of the clusters in $Q$.
+While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
 We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could move from the query to the center of one of the nearby clusters without significantly changing our estimate of the LFD.
@@
 
-Since $\hat{d}$, $\hat{|C|}$, $\hat{r}$, and $k$ are all known values, we can solve this equation for $\rho_k$ to get
+We can rearrange this equation to get
 
 \begin{equation*}
-    \rho_k = \hat{r} \left( \frac{k}{\hat{|C|}} \right)^{\hat{d}^{-1}}
+    \frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}
 \end{equation*}
@@
 \begin{equation*}
-    k \left( 1+ \frac{2\hat{r}}{\rho_k} \left( \frac{k}{\hat{|C|}} \right) ^ {\hat{d}^{-1}} \right)^d
+    k \left( 1 + 2 \cdot \left( \frac{\hat{|C|}}{k} \right) ^ {\hat{d}^{-1}} \right)^d.
 \end{equation*}
 
-Since we are performing an asymptotic analysis, we can ignore the additive factor of $1$.
+In addition, by our assumption that the LFD at the query is not significantly different from the LFD of nearby clusters, we have that $\hat{d} \approx d$.
 By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has an asymptotic complexity of
@@
     \Bigg(
-        \underbrace{\log_2{\mathcal{N}_{\hat{r}}(X)}}_{\textrm{metric entropy}} \ \ \ + 
+        \underbrace{
+            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+        }_{\textrm{tree search}}
+        \ + \ 
+        \underbrace{
             \overbrace{k}^{\textrm{output size}}
-        \underbrace{
-            \left( \frac{2\hat{r}}{\rho_k} \right) ^ d \left( \frac{k}{\hat{|C|}} \right) ^ {d \cdot \hat{d}^{-1}}
-        }_{\textrm{scaling factor}}
+            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
+        }_{\textrm{leaf search}}
     \Bigg)
@@
 where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-We note that the term $d \cdot \hat{d}^{-1}$ should be relatively close to 1 unless fractal dimension is highly variable in a small region (that is, $\hat{d}$ differs significantly from $d$).
+We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in a small region (i.e. if $\hat{d}$ differs significantly from $d$).
 
@@
 We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
-During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:the-search-problem}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.
+During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:the-search-problem}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 
 We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
+Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot is linear-time, this version of QuickSelect may require $\mathcal{O}(|Q| \log |Q|)$ time.
 
@@
 
-        \STATE $c \Leftarrow$ \emph{center} of $root$
+        \STATE $c \Leftarrow$ \textit{center} of $root$
         \STATE $Q \Leftarrow$ \{ ($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1) \}
@@
 
-\subsubsection{Complexity of Breadth-First Sieve}
-\label{sec:methods:knn-search:complexity-of-breadth-first-sieve}
-
-The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-Since this method is linear in the size of the input list $Q$, we must estimate the size of $Q$ on each iteration.
-
-We start by considering clusters whose cardinalities are within a small factor of $k$, say a factor of 2.
-For small $k$, as is the case in real-world applications, the distribution of such clusters will mimic the distribution of points in the dataset.
-From this assumption, we can infer that the number of clusters which overlap with the query ball is a function of $d$, the LFD near the query.
-In particular, we assume that the number of clusters which overlap the query ball is bounded above by $2d$, e.g.
-we could have a cluster overlapping the query ball at each end of each of $d$ mutually-orthogonal axes.
-
-In the worst case scenario, where the data are distributed with uniform density in their embedding space, it takes at most $\log{\frac{\mathcal{N}_{\hat{r}}(X)}{2k}}$ iterations of breadth-first traversal to reach clusters whose cardinalities are at most $2k$.
-However, since high cardinality clusters occur only at low depths in the tree and there are few such clusters, the length of $Q$, and thus the cost of QuickSelect, for those iterations is relatively low.
-
-After the first $\log{\frac{\mathcal{N}_{\hat{r}}(X)}{2k}}$ iterations, we are left with clusters in $Q$ whose cardinalities are at most $2k$.
-At this point, based on our assumptions about LFD, we have that the length of $Q$ is close to $2d$.
-Therefore, the asymptotic complexity of the Breadth-First Sieve Algorithm is $\mathcal{O} \left( k \cdot \log(\frac{\mathcal{N}_{\hat{r}}(X)}{d}) \right)$.
-
 \subsubsection{Depth-First Sieve}
@@
 \begin{itemize}
-    \item While the top priority element $Q$ is not a leaf, pop it from $Q$ and push its children to $Q$.
-    \item Let $q$ the top priority element (a leaf) in $Q$. For each point $p \in q$, push $p$ to $H$.
-    \item If $H$ has more than $k$ points, pop from $H$ until $H$ has $k$ points.
+    \item While the top priority element $Q$ is not a leaf, remove it from $Q$ and add its children to $Q$.
+    \item Remove the top priority element (a leaf) from $Q$ and add all its points to $H$.
+    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
 \end{itemize}
@@
             \WHILE{$\neg (Q.peek$ is a leaf)}
-                \STATE{$[l, r] \Leftarrow Q.pop().children$}
-                \STATE{$Q.push(l)$}
-                \STATE{$Q.push(r)$}
+                \STATE{$C \Leftarrow Q.pop$, the closest cluster}
+                \STATE{$[L, R] \Leftarrow$ children of $C$}
+                \STATE{$Q.push(L)$}
+                \STATE{$Q.push(R)$}
             \ENDWHILE
-            \STATE{$leaf \Leftarrow Q.pop()$}
+            \STATE{$leaf \Leftarrow Q.pop$}
             \FOR{$p \in leaf$}
@@
             \WHILE{$|H| > k$}
-                \STATE{$H.pop()$}
+                \STATE{$H.pop$}
             \ENDWHILE
@@
 
-Note that this algorithm is not a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
+Note that this algorithm is not truly a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
 Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
 
-\subsubsection{Complexity of Depth-First Sieve}
-\label{sec:methods:knn-search:depth-first-sieve-complexity}
+\subsubsection{Complexity of Sieve Methods}
+\label{sec:methods:knn-search:complexity-of-sieve-methods}
 
-TODO: Analyze complexity of Depth-First Sieve.
+We combine the complexity analyses of the two Sieve methods because they are very similar.
+For these methods we, once again, use the terminology of tree-search and leaf-search.
+Tree-search navigates the cluster tree and adds clusters to $Q$.
+Leaf-search exhaustively searches some of the clusters in $Q$ to find the $k$ nearest neighbors.
+
+We start with the assumption that the LFD near the query does not differ significantly from that of nearby clusters, and we consider leaf clusters with cardinalities near $k$.
+Let $d$ be the LFD in this region.
+Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g. we could have a cluster overlapping the query ball at each end of each of $d$ mutually-orthogonal axes.
+In the worst case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot d \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
+Thus, the asymptotic complexity is
+
+\begin{equation*}
+    \mathcal{T} \coloneqq \mathcal{O} \big( d \cdot \log \mathcal{N}_{\hat{r}}(X) \big).
+\end{equation*}
+
+For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is
+
+\begin{equation*}
+    \mathcal{L} \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right).
+\end{equation*}
+
+The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
+Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
+
+\begin{equation}
+    \mathcal{O} \Big( ( \mathcal{T} + \mathcal{L} ) \log ( \mathcal{T} + \mathcal{L} ) \Big).
+    \label{eq:methods:breadth-first-sieve-complexity}
+\end{equation}
+
+For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the complexity of the priority queue operations.
+Thus, the asymptotic complexity is
+
+\begin{equation}
+    \mathcal{O} \Big( \mathcal{T} \log \mathcal{T} + \mathcal{L} \log k \Big).
+    \label{eq:methods:depth-first-sieve-complexity}
+\end{equation}
 
@@
 We start by taking the center of every cluster at a low depth, e.g. 10, in the cluster tree as a query.
-This gives us a small set of queries which is a representative sample of the dataset.
+This gives us a small, representative sample of the dataset.
 Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{sec:methods:knn-search}.

### 2023-12-31 sections/3-datasets.tex
squashed all red TODOs

@@
 We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
-Since this dataset contains many \emph{redundant}, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition \emph{criterion} in Algorithm~\ref{alg:methods:partition}.
+Since this dataset contains many \textit{redundant}, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition \textit{criterion} in Algorithm~\ref{alg:methods:partition}.
 

### 2023-12-31 sections/4-results.tex
brougnt back abstract changes
squashed all red TODOs

--- 2023-12-29 sections/4-results.tex
@@
 We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
-In order to have the plots to best reflect the distribution of LFDs across the entire \emph{dataset}, we compute percentiles counting each cluster as many times as its cardinality.
+In order to have the plots to best reflect the distribution of LFDs across the entire \textit{dataset}, we compute percentiles counting each cluster as many times as its cardinality.
 In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the total points in clusters at depth 40 belong to clusters whose LFD is at most 3.
@@
 
-{\color{red} TODO: remove all red TODOs.}
-
-{\color{red} TODO: Add more red TODOs.}
-
-{\color{red} TODO: Figure out if we should remove all red TODOs or add more red TODOS}
-
 On all datasets, we observe that the indexing time for CAKES increases roughly linearly as cardinality increases.
@@
 For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve. 
-Additionally, on all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \emph{constant} in the cardinality of the dataset. 
+Additionally, on all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset. 
 With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling. 
@@
 As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-HNSW and ANNOY exhibit \emph{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
-in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:random}.
+HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
+in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:qps-and-recall}.
 
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+    Note that for RadioML, all three algorithms performed nearly identically, so distinct lines are not visible.
     For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e. more then 512GB).}

### 2023-12-31 sections/5-discussion.tex
squashed all red TODOs

@@
 
-We also stress that CAKES's algorithms are designed to work well for \emph{big} data, and our results support this claim;
+We also stress that CAKES's algorithms are designed to work well for \textit{big} data, and our results support this claim;
 we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmarks' datasets we tested.
@@
 For example, Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-Mnist and Glove-25, respectively.
-For the Random dataset, which has the highest LFD, CAKES's algorithms \emph{never} outperform linear search.
+For the Random dataset, which has the highest LFD, CAKES's algorithms \textit{never} outperform linear search.
 These observations support our claim that CAKES performs well on datasets that arise from constrained generating phenomena, even as the cardinalities of these datasets grow exponentially.

### 2024-01-04 sections/1-introduction.tex
tweaks from RFD

@@
 
-Examples of large datasets include neural network embeddings, genomic databases and time-series data such as radio frequency signals.
+Examples of large datasets include genomic databases, time-series data such as radio frequency signals, and neural network embeddings.
 Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image} are a common source of neural network embeddings.
 These embeddings are often high-dimensional, and the sizes of training and inference datasets for such networks are growing exponentially.
-Among Biological datasets, the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length while Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
+Among biological datasets, the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length while SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
 Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
@@
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is some evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
+Further, there is evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
 this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.
@@
 One could also search for all words within a maximum edit distance of a misspelled word to suggest corrections~\cite{ukkonen1985algorithms}.
-Given an advertisement and a database of user profiles, one could search for all users whose profiles are ``similar enough'' to serve the advertisement to~\cite{zhang2020privacy}.
+Given an advertisement and a database of user profiles, one could search for all users whose profiles are ``similar enough'' to target the advertisement~\cite{zhang2020privacy}.
 GPS and other location-based services use $\rho$-NN search to find nearby points of interest~\cite{zhang2020privacy} to a user's location.
@@
 We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
-We provide a comparison of CAKES to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on several datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-We also benchmark CAKES on a genomic dataset, the Silva 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+We provide a comparison of CAKES to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We also benchmark CAKES on a genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
 
@@
 
-Recent search algorithms for tackling the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+However, some of these algorithms do not provide exact search.
 

### 2024-01-04 sections/2-methods.tex
tweaks from RFD

@@
 We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the Silva 18S rRNA dataset which contains which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
+To illustrate the difference, consider the SILVA 18S rRNA dataset which contains which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
 Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
@@
 
-It remains to provide an estimate for $\rho_k$. 
+An estimate for $\rho_k$ is still needed. 
 For this, we once again rely on the assumption that the query is drawn from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different from the LFD of nearby clusters.
@@
 
-By Equation~\ref{eq:methods:lfd-original}, we then have that
+By Equation~\ref{eq:methods:lfd-original},
 

### 2024-01-04 sections/3-datasets.tex
tweaks from RFD

@@
 
-The benchmarks for Silva and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
+The benchmarks for SILVA and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
 The OS kernel was Manjaro Linux 5.15.130-1-MANJARO.
@@
                     \hline
-                    Silva            & Levenshtein            & 2,224,640          & 3,712 - 50,000         \\
+                    SILVA            & Levenshtein            & 2,224,640          & 3,712 - 50,000         \\
                     \hline
@@
 
-\subsection{Silva 18S}
+\subsection{SILVA 18S}
 \label{sec:datasets-and-benchmarks:silva-18s}
 
-To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the Silva 18S ribosomal DNA dataset~\cite{10.1093/nar/gks1219}.
+To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal DNA dataset~\cite{10.1093/nar/gks1219}.
 This dataset contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.

### 2024-01-04 sections/2-methods.tex
brief complexity discussion

@@
             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-        }_{\textrm{tree search}}
+        }_{\textrm{tree-search}}
         \ + \ 
@@
 one iteration to find at least one cluster, and the one more to find enough the $k$ neighbors.
-Since this is a constant factor, complexity of tree search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e. $\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
+Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e. $\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
 
-To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree search.
+To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree-search.
 Since we must examine every point in each such clusters, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$. 
@@
 Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g. we could have a cluster overlapping the query ball at each end of each of $d$ mutually-orthogonal axes.
-In the worst case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot d \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
+In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot d \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
 Thus, the asymptotic complexity is

### 2024-01-04 sections/5-discussion.tex
brief complexity discussion

@@
 
-We also note that the CAKES algorithm with the best throughput varies by dataset.
+When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\mathcal{T} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
+We note that even though, of the three CAKES algorithms, Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve), it is not always the fastest algorithm.
+In fact, the fastest algorithm varies by dataset.
+
 For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-Mnist.

### 2024-01-04 sections/2-methods.tex
incorporate some of Carl's suggestions

@@
 
-Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and forS performance to scale sub-linearly with the cardinality of the dataset.
+Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and for its performance to scale sub-linearly with the cardinality of the dataset.
 To test this hypothesis, we use some datasets from the ANN-benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to generate similar datasets with exponentially larger cardinalities.

### 2024-01-04 sections/4-results.tex
incorporate some of Carl's suggestions

@@
 
-Figure~\ref{fig:results:lfd-plots} the trends in LFD for (the non-augmented versions of) Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
+Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for (the non-augmented versions of) Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
 The horizontal axis denotes the depth in the cluster tree, and the vertical axis denotes the LFD (as calculated by Equation~\ref{eq:methods:lfd-simplified}) of clusters at that depth.
@@
 Relative to the other datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
-Notably however, for the first approximately 15 depths, the 95$^{th}$ percentile and median LFD are very close to 0.
+Notably, however, for the first approximately 15 depths, the 95$^{th}$ percentile and median LFD are very close to 0.
 Both percentiles increase sharply after depth 20, nearly reaching 16.
@@
 The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{sec:methods:synthetic-data}).
-The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation..
+The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
 The vertical axis denotes throughput in queries per second.

### 2024-01-05 sections/1-introduction.tex
incorporate some of Carl's suggestions
addressing carl's comments

@@
 Researchers are collecting data at an unprecedented scale.
-In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{brescia2012extracting}.
+In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
 This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
@@
 InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
-The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point into is then searched exhaustively.
+The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively.
 The number of cells used is governed by the $n_{list}$ parameter. 

### 2024-01-05 sections/2-methods.tex
addressing carl's comments

@@
             \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
-        }_{\textrm{leaf search}}
+        }_{\textrm{leaf-search}}
     \Bigg)
@@
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
-We allow a maximum increase of 2 to ensure that we do not increase the radius too quickly in any single iteration.
+We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.
 
@@
 
-Using this to simplify the term for leaf search in Equation~\ref{eq:methods:rnn-search-complexity}, we get
+Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get
 
@@
             \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
-        }_{\textrm{leaf search}}
+        }_{\textrm{leaf-search}}
     \Bigg)

### 2024-01-05 sections/4-results.tex
addressing carl's comments

@@
 With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has lower indexing time than CAKES.
-With Sift, CAKEs's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
+With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
 
@@
 
-For Silva and RadioML, we benchmarked only CAKES's algorithms because none HNSW, ANNOY or FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
+For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS do not support neither the required distance functions nor, in the case of RadioML, complex-valued data.
 Due to the massive sizes of these datasets, we took random sub-samples of the dataset with lower cardinalities for our benchmarks, rather than augmented versions of the dataset.
@@
 For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the three CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
-In particular, throughput of CAKES algorithms is identical within three significant figures for all cardinalities.
+In particular, throughput of CAKES's algorithms is identical within three significant figures for all cardinalities.
 

### 2024-01-05 sections/5-discussion.tex
fixed typo in methods and added to discussion

--- 2023-12-31 sections/5-discussion.tex
@@
 We have presented CAKES, a set of three algorithms for fast $k$-NN search that are generic over a variety of distance functions.
-CAKES's algorithms are exact when the distance function in use is a metric (i.e., when it obeys the triangle inequality).
+CAKES's algorithms are exact when the distance function in use is a metric (see definition in~\ref{sec:methods}).
 Even under Cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).
 
-CAKES's algorithms are designed to be most effective when the data uphold the manifold hypothesis, i.e.
-when the data are constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
+CAKES's algorithms are designed to be most effective when the data uphold the manifold hypothesis, or in other words, when the data are constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
 As a consequence, these algorithms do not perform well on data with random distributions, because of the absence of a manifold structure.
@@
 
-As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift.
+As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though they have the same cardinality and embedding dimension.
 This is unsurprising, given that the Random dataset is, by definition, uniformly distributed, while Sift is not.
@@
 This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
-CAKES's tree is orders of magnitude cheaper to build and CAKES's algorithms', while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
+CAKES's tree is orders of magnitude cheaper to build and CAKES's algorithms, while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
 
-We also note that the CAKES algorithm with the best throughput varies by dataset.
+When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\mathcal{T} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
+Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
+We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in~\ref{eq:methods:repeated-rnn-complexity}.
+This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation. 
+For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD, 
+the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.  
+This suggests that rather than use the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may see better results by using the reciprocal of a mean which is more sensitive to high outliers, such as the geometric mean.
+We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this overshooting occurs less frequently and with less severity.
+
+As the previous paragraph suggests, the fastest CAKES algorithm varies by dataset.
 For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-Mnist.
@@
 Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities supports our use of an auto-tuning function (as discussed in Section~\ref{sec:methods:auto-tuning}) to select the best algorithm for a given dataset.
-Given trends such as the performance of Repeated $\rho$-NN with low LFD, we suspect that a more sophisticated auto-tuning function could be developed to select the fastest algorithm based on the properties of the dataset.
+Future work is needed to better characterize those datasets for which each algorithm performs best, but our results suggest that a more sophisticated auto-tuning function could be developed to select the fastest algorithm based on the properties of the dataset.
 
@@
 
-The questions raised by this study suggest several avenues for future work.
-A comparison across more (and larger) datasets is clearly in order, as is further analysis with distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
+The questions raised by this study suggest several additional avenues for future work.
+A comparison across more datasets is in order, as is further analysis with other distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
 Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
@@
 Such online updates would take advantage of the fast search algorithms provided by CAKES.  % and would allow us to have our cake and eat it, too.
-
 CAKES can also be used to extend anomaly detection in CHAODA~\cite{ishaq2021clustered}.

### 2024-01-05 sections/1-introduction.tex
general edits

@@
 This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
-Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms for analyzing such datasets.
+Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms for analyzing large datasets.
 
@@
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
+Further, there is evidence that distance functions which do not obey the triangle inequality, such as Cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
 this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.
@@
 
-This paper focuses on three novel algorithms for exact $k$-NN search.
+This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy Scaling Search), a set of three novel algorithms for exact $k$-NN search.
 We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
-We provide a comparison of CAKES to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We provide a comparison of CAKES's algorithms to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
 We also benchmark CAKES on a genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
@@
 In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
-In this paper, we introduce CAKES (CLAM-Accelerated $K$-nn Entropy-scaling Search).
 CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
+In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN. 
 Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search and improves the performance of CHESS's $\rho$-NN search.

### 2024-01-05 sections/2-methods.tex
fixed typo in methods and added to discussion
general edits

@@
 $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
-Under distance metrics, all search algorithms in CAKES are exact.
-For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics while Cosine distance is not a metric because it does not obey the triangle inequality (e.g.
-consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
+Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
+For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while Cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
 
@@
 \end{equation}
-
 where $B_X(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
@@
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
-Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets obey the manifold hypothesis.
+Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
 
@@
 Note that Algorithm~\ref{alg:methods:partition} does not necessarily produce a balanced tree.
-In fact, for real datasets, we would expect anything but a balanced tree.
-The imbalance in the tree would be due to the varying density of the sampling in different regions of the manifold and the low dimensional ``shape'' of the manifold itself.
+In fact, for real datasets, we expect anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
 The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g. in a $d$-dimensional hyper-cube.
@@
 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, we define the $k$-NN search problem which aims to find the $k$ closest points to $q$ in $\textbf{X}$;
-i.e.
-the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
-We also have the $\rho$-NN search problem, which aims to find the all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$;
-i.e. the set $H = B_X(q, \rho) = \{x \in \textbf{X}: f(q, x) \leq \rho \}$.
+Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, we define the $k$-NN search problem, which aims to find the $k$ closest points to $q$ in $\textbf{X}$;
+in other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
+We also have the $\rho$-NN search problem, which aims to find the all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$: the set $H = B_X(q, \rho) = \{x \in \textbf{X}: f(q, x) \leq \rho \}$.
 
@@
 To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
-Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points in which are no more than a distance $\rho$ from the query.
+Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points which are no more than a distance $\rho$ from the query.
 
@@
 \end{gather}
-
 where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
@@
 
-where $\mu$ is the harmonic mean of the LFD of the clusters in $S$.
+where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$.
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.

### 2024-01-05 sections/5-discussion.tex
general edits

@@
 
-When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\mathcal{T} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
+When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{d} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
 Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
@@
 the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.  
-This suggests that rather than use the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may see better results by using the reciprocal of a mean which is more sensitive to high outliers, such as the geometric mean.
+This suggests that rather than use the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may actually want a mean which is more sensitive to high outliers, such as the geometric mean.
 We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this overshooting occurs less frequently and with less severity.

### 2024-01-31 sections/2-methods.tex
typos and added a paper to bib file

@@
 We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the SILVA 18S rRNA dataset which contains which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
+To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
 Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
@@
 In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
-Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access the its points from the dataset.
+Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset.
 The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.

### 2024-04-05 sections/1-introduction.tex
paper edits up until section 4.3

@@
 This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
-Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms for analyzing large datasets.
+Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms to analyze large datasets.
 
@@
 These embeddings are often high-dimensional, and the sizes of training and inference datasets for such networks are growing exponentially.
-Among biological datasets, the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length while SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
+Among biological datasets, the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length, while SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
 Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
@@
 While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
-Further, such algorithms are often approximate and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains.
+Further, such algorithms are often approximate, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains.
 
@@
 We provide a comparison of CAKES's algorithms to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-We also benchmark CAKES on a genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+We also benchmark CAKES on a genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
 
@@
 Recent search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
-However, some of these algorithms do not provide exact search.
+However, some of these algorithms do not provide exact search (as defined in Section \ref{sec:introduction} above).
 
@@
 CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
-In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN. 
-Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search and improves the performance of CHESS's $\rho$-NN search.
+In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search. 
+Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search. We also discuss improvements to the performance of CHESS's $\rho$-NN search.

### 2024-04-05 sections/2-methods.tex
compiles, switched to algorithmicx, tables still bad
paper edits up until section 4.3

@@
 In this manuscript, we are primarily concerned with $k$-NN search in a finite-dimensional space.
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \textit{point} or \textit{datum} $x_i \in \textbf{X}$ as a singular observation, for example, the neural-network embedding of an image, the genome of an organism, a measurement of a radio frequency signal, etc.
+Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \textit{point} or \textit{datum} $x_i \in \textbf{X}$ as a singular observation.
+Examples include the neural-network embedding of an image, the genome of an organism, and a measurement of a radio frequency signal.
 
@@
 
-The choice of distance function varies by dataset and domain.
+The choice of an appropriate distance function varies by dataset and domain.
 For example, with neural-network embeddings, one could use Euclidean (L2) or Cosine distance.
@@
 
-CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, i.e., high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
+CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
 We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
@@
 Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases.
+When the vast majority of points in the dataset have low ($\ll D$) LFD, we can overload terminology to say that the dataset has low LFD.
 
@@
 Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
-As a consequence, a cluster can be a subset of the metric ball at the same center and radius, i.e. $C(c, r) \subseteq B_X(c, r)$.
+As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e. $C(c, r) \subset B_X(c, r)$.
 
@@
 The \textit{metric entropy} $\mathcal{N}_{r}(X)$ for some radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover the data.
-In this paper, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ for the hierarchical clustering scheme as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
+In this paper, where the clustering is hierarchical rather than flat, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
 
@@
 The \textit{radius} of $C$ is the maximum distance from the \textit{center} to any other point in $C$.
-The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from left pole is designated the \textit{right pole}.
+The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.
 
@@
     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-        \REQUIRE $C$, a cluster
-        \REQUIRE $criteria$, user-specified stopping criteria
+    \begin{algorithmic}[0] % enter the algorithmic environment
+        \Require $C$, a cluster
+        \Require $criteria$, user-specified stopping criteria
 
-        \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-        \STATE $c \Leftarrow$ geometric median of $seeds$
-        \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-        \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-        \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-        \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+        \State $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+        \State $c \Leftarrow$ geometric median of $seeds$
+        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+        \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+        \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
 
-        \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
-            \STATE Partition($L$, $criteria$)
-        \ENDIF
+        \If{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
+            \State Partition($L$, $criteria$)
+        \EndIf
 
-        \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
-            \STATE Partition($R$, $criteria$)
-        \ENDIF
+        \If{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
+            \State Partition($R$, $criteria$)
+        \EndIf
     \end{algorithmic}
@@
 
-Given Algorithm~\ref{alg:methods:partition}, we can now pose the $k$-NN and $\rho$-NN search problems.
+Given Algorithm~\ref{alg:methods:partition}, we can now formally pose the $k$-NN and $\rho$-NN search problems.
 
@@
 
-We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \textit{unique} point).
+We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains only duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \textit{unique} point).
 A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
@@
 We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
-when a cluster overlaps with the query ball, instead of always proceeding with both of its children, we proceed only with those children which might contain points in the query ball.
+when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children which might contain points in the query ball.
 
@@
 To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children}.
-Here, we overload the notation for $\overline{x y}$ to refer to the line segment joining points $x$ and $y$ as well as the length of that line segment.
+Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.
 
@@
 Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
-Now let $q'$ be the projection of $q$ to $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
-As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $q'$ is too far from $m$, i.e.
-$\rho < d$, then when the left child cannot contain points inside the query ball.
-In such a case we proceed only with the right child.
+Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
+As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $\rho < d$, then the left child cannot contain points inside the query ball.
+In such a case we proceed to search only the right child.
 Otherwise, we proceed with both children.
@@
     \label{alg:methods:rnn-search:tree-search} 
-    \begin{algorithmic}
-        \REQUIRE $C$, a cluster
-        \REQUIRE $q$, a query
-        \REQUIRE $\rho$, a search radius
+    \begin{algorithmic}[0]
+        \Require $C$, a cluster
+        \Require $q$, a query
+        \Require $\rho$, a search radius
 
-        \IF{$C$ is a leaf \textbf{or} $\delta^+_C \leq \rho$}
-            \STATE \textbf{return} $\{C\}$
-        \ELSE
+        \If{$C$ is a leaf \textbf{or} $\delta^+_C \leq \rho$}
+            \State \textbf{return} $\{C\}$
+        \Else
 
-            \STATE $[l, r]$ $\Leftarrow$ \textit{pole}s of $C$
-            \STATE $[L, R]$ $\Leftarrow$ \textit{children} of $C$
+            \State $[l, r]$ $\Leftarrow$ \textit{pole}s of $C$
+            \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
 
-            \IF{$\overline{q l} < \overline{q r \vphantom{l}}$}
-                \STATE $[r, l] \Leftarrow [l, r]$
-                \STATE $[R, L] \Leftarrow [L, R]$
-            \ENDIF
+            \If{$\overline{q l} < \overline{q r \vphantom{l}}$}
+                \State $[r, l] \Leftarrow [l, r]$
+                \State $[R, L] \Leftarrow [L, R]$
+            \EndIf
 
-            \IF{$(\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$}
-                \STATE \textbf{return} tree-search($L, q, \rho$) $\cup$ tree-search($R, q, \rho$)
-            \ELSE
-                \STATE \textbf{return} tree-search($R, q, \rho$)
-            \ENDIF
-        \ENDIF
+            \If{$(\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$}
+                \State \textbf{return} tree-search($L, q, \rho$) $\cup$ tree-search($R, q, \rho$)
+            \Else
+                \State \textbf{return} tree-search($R, q, \rho$)
+            \EndIf
+        \EndIf
     \end{algorithmic}
@@
     \label{alg:methods:rnn-search:leaf-search} 
-    \begin{algorithmic}
-        \REQUIRE $S$, a set of clusters
-        \REQUIRE $q$, a query
-        \REQUIRE $\rho$, a search radius
+    \begin{algorithmic}[0]
+        \Require $S$, a set of clusters
+        \Require $q$, a query
+        \Require $\rho$, a search radius
 
-        \STATE $H \Leftarrow \emptyset$, a set of hits
+        \State $H \Leftarrow \emptyset$, a set of hits
 
-        \FOR{$C \in S$}
-            \IF{$\delta^+_C \leq \rho$}
-                \STATE $H$ $\Leftarrow$ $H \cup \{C\}$
-            \ELSE
-                \FOR{$p \in C$}
-                    \IF{$f(p, q) \leq \rho$}
-                        \STATE $H$ $\Leftarrow$ $H \cup \{p\}$
-                    \ENDIF
-                \ENDFOR
-            \ENDIF
-        \ENDFOR
+        \For{$C \in S$}
+            \If{$\delta^+_C \leq \rho$}
+                \State $H$ $\Leftarrow$ $H \cup \{C\}$
+            \Else
+                \For{$p \in C$}
+                    \If{$f(p, q) \leq \rho$}
+                        \State $H$ $\Leftarrow$ $H \cup \{p\}$
+                    \EndIf
+                \EndFor
+            \EndIf
+        \EndFor
 
-        \STATE \textbf{return} $H$
+        \State \textbf{return} $H$
     \end{algorithmic}
@@
     \begin{algorithmic}
-        \REQUIRE $root$, the root cluster
-        \REQUIRE $q$, a query
-        \REQUIRE $\rho$, a search radius
+        \Require $root$, the root cluster
+        \Require $q$, a query
+        \Require $\rho$, a search radius
 
-        \STATE $S$ $\Leftarrow$ tree-search($root$, $q$, $\rho$)
-        \STATE $H$ $\Leftarrow$ leaf-search($S$, $q$, $\rho$)
-        \STATE \textbf{return} $H$
+        \State $S$ $\Leftarrow$ tree-search($root$, $q$, $\rho$)
+        \State $H$ $\Leftarrow$ leaf-search($S$, $q$, $\rho$)
+        \State \textbf{return} $H$
     \end{algorithmic}
@@
     \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-        \REQUIRE $root$, the root cluster
-        \REQUIRE $q$, a query
-        \REQUIRE $k$, the number of neighbors to find
+    \begin{algorithmic}[0] % enter the algorithmic environment
+        \Require $root$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
 
-        \STATE $r \Leftarrow radius$ of the $root$ cluster
-        \STATE $r \Leftarrow$ $\frac{r}{|root|}$
+        \State $r \Leftarrow radius$ of the $root$ cluster
+        \State $r \Leftarrow$ $\frac{r}{|root|}$
 
-        \ALOOP{}
-            \STATE $Q \Leftarrow$ tree-search($root$, $q$, $r$)
-            \IF{$Q \neq \emptyset$}
-                \STATE \textbf{break}
-            \ENDIF
-            \STATE $r \Leftarrow 2 \cdot r$
-        \ENDALOOP
+        \Loop
+            \State $Q \Leftarrow$ tree-search($root$, $q$, $r$)
+            \If{$Q \neq \emptyset$}
+                \State \textbf{break}
+            \EndIf
+            \State $r \Leftarrow 2 \cdot r$
+        \EndLoop
 
-        \ALOOP{}
-            \IF{$\sum_{C \in Q} |C| >= k$}
-                \STATE \textbf{break}
-            \ENDIF
-            \STATE $\mu \Leftarrow \frac{|S|}{\sum_{C \in Q} \frac{1}{LFD(C)}}$
-            \STATE $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)$
-            \STATE $S \Leftarrow$ tree-search($root$, $q$, $r$)
-        \ENDLOOP
+        \Loop
+            \If{$\sum_{C \in Q} |C| >= k$}
+                \State \textbf{break}
+            \EndIf
+            \State $\mu \Leftarrow \frac{|S|}{\sum_{C \in Q} \frac{1}{LFD(C)}}$
+            \State $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)$
+            \State $S \Leftarrow$ tree-search($root$, $q$, $r$)
+        \EndLoop
 
-        \STATE $H \Leftarrow$ sort($\rho$-NN-search($root$, $q$, $r$))
-        \STATE \textbf{return} $H[.. k]$
+        \State $H \Leftarrow$ sort($\rho$-NN-search($root$, $q$, $r$))
+        \State \textbf{return} $H[.. k]$
     \end{algorithmic}
@@
 where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in a small region (i.e. if $\hat{d}$ differs significantly from $d$).
+We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in the region around the query (i.e. if $\hat{d}$ differs significantly from $d$).
 
@@
 We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
-Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot is linear-time, this version of QuickSelect may require $\mathcal{O}(|Q| \log |Q|)$ time.
+Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot takes linear time in the size of the input list, this version of QuickSelect has $\mathcal{O}(|Q| \log |Q|)$ time complexity.
 
@@
 We continue this process until $Q$ no longer contains any clusters.
-We then use the QuickSelect algorithm once last time to reorder $Q$, find $\tau$, and return the $k$ nearest neighbors.
+We then use the QuickSelect algorithm one last time to reorder $Q$, find $\tau$, and return the $k$ nearest neighbors.
 
@@
     \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-        \REQUIRE $root$, the root cluster
-        \REQUIRE $q$, a query
-        \REQUIRE $k$, the number of neighbors to find
+    \begin{algorithmic}[0] % enter the algorithmic environment
+        \Require $root$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
 
-        \STATE $c \Leftarrow$ \textit{center} of $root$
-        \STATE $Q \Leftarrow$ \{ ($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1) \}
+        \State $c \Leftarrow$ \textit{center} of $root$
+        \State $Q \Leftarrow$ \{ ($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1) \}
 
-        \ALOOP{}
-            \IF{$k = \sum_{(\_, \_, m) \in Q} m$}
-                \STATE \textbf{break}
-            \ENDIF
-            \STATE $\tau \Leftarrow$ QuickSelect($Q$, $k$)
-            \FOR{$(p, \delta^{+}_{p}, m) \in Q$}
-                \IF{$\delta^{-}_{p} > \tau$}
-                    \STATE $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
-                \ENDIF
-            \ENDFOR
-            \FOR{$(p, \delta^{+}_{p}, m) \in Q$}
-                \STATE $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
-                \IF{$p$ is a point}
-                    \STATE \textbf{continue}
-                \ELSIF{$p$ is a leaf}
-                    \FOR{$c \in p$}
-                        \STATE $Q \Leftarrow Q \cup \{ (c, \delta_{c}, 1) \}$
-                    \ENDFOR
-                \ELSE
-                    \STATE $[L, R] \Leftarrow$ children of $p$
-                    \STATE $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (L_c, \delta_{L}, 1) \}$
-                    \STATE $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (R_c, \delta_{R}, 1) \}$
-                \ENDIF
-            \ENDFOR
-        \ENDALOOP
+        \Loop
+            \If{$k = \sum_{(\_, \_, m) \in Q} m$}
+                \State \textbf{break}
+            \EndIf
+            \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
+            \For{$(p, \delta^{+}_{p}, m) \in Q$}
+                \If{$\delta^{-}_{p} > \tau$}
+                    \State $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
+                \EndIf
+            \EndFor
+            \For{$(p, \delta^{+}_{p}, m) \in Q$}
+                \State $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
+                \If{$p$ is a point}
+                    \State \textbf{continue}
+                \ElsIf{$p$ is a leaf}
+                    \For{$c \in p$}
+                        \State $Q \Leftarrow Q \cup \{ (c, \delta_{c}, 1) \}$
+                    \EndFor
+                \Else
+                    \State $[L, R] \Leftarrow$ children of $p$
+                    \State $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (L_c, \delta_{L}, 1) \}$
+                    \State $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (R_c, \delta_{R}, 1) \}$
+                \EndIf
+            \EndFor
+        \EndLoop
 
-        \STATE \textbf{return} $Q$
+        \State \textbf{return} $Q$
     \end{algorithmic}
@@
     \label{alg:methods:depth-first-sieve} 
-    \begin{algorithmic}
-        \REQUIRE $root$, the root cluster
-        \REQUIRE $q$, a query
-        \REQUIRE $k$, the number of neighbors to find
+    \begin{algorithmic}[0]
+        \Require $root$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
 
-        \STATE{$Q \Leftarrow [root]$, a min-priority queue by $\delta^{-}$}
-        \STATE{$H \Leftarrow []$, a max-priority queue by $\delta$}
+        \State{$Q \Leftarrow [root]$, a min-priority queue by $\delta^{-}$}
+        \State{$H \Leftarrow []$, a max-priority queue by $\delta$}
 
-        \WHILE{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
-            \WHILE{$\neg (Q.peek$ is a leaf)}
-                \STATE{$C \Leftarrow Q.pop$, the closest cluster}
-                \STATE{$[L, R] \Leftarrow$ children of $C$}
-                \STATE{$Q.push(L)$}
-                \STATE{$Q.push(R)$}
-            \ENDWHILE
-            \STATE{$leaf \Leftarrow Q.pop$}
-            \FOR{$p \in leaf$}
-                \STATE{$H.push(p)$}
-            \ENDFOR
-            \WHILE{$|H| > k$}
-                \STATE{$H.pop$}
-            \ENDWHILE
-        \ENDWHILE
-        \STATE Return $H$
+        \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
+            \While{$\neg (Q.peek$ is a leaf)}
+                \State{$C \Leftarrow Q.pop$, the closest cluster}
+                \State{$[L, R] \Leftarrow$ children of $C$}
+                \State{$Q.push(L)$}
+                \State{$Q.push(R)$}
+            \EndWhile
+            \State{$leaf \Leftarrow Q.pop$}
+            \For{$p \in leaf$}
+                \State{$H.push(p)$}
+            \EndFor
+            \While{$|H| > k$}
+                \State{$H.pop$}
+            \EndWhile
+        \EndWhile
+        \State Return $H$
     \end{algorithmic}
@@
 
-We combine the complexity analyses of the two Sieve methods because they are very similar.
+Due to their similarity, we combine the complexity analyses of the two Sieve methods into one section.
 For these methods we, once again, use the terminology of tree-search and leaf-search.
@@
 We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
-We start by taking the center of every cluster at a low depth, e.g. 10, in the cluster tree as a query.
+We start by taking the center of every cluster at a low depth, (e.g. 10), in the cluster tree as a query.
 This gives us a small, representative sample of the dataset.
@@
 To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
-Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level and $m$ be a user-specified integer multiplier.
+Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level, and $m$ be a user-specified integer multiplier.
 For each datum $x \in X$, we create $m - 1$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.

### 2024-04-05 sections/3-datasets.tex
paper edits up until section 4.3

@@
 We use our own Rust implementation of linear search to provide ground-truth values for calculating recall of CAKES's algorithms.
-We plot the throughput of these two implementations of linear search as a reference point for all other algorithms in Figures~\ref{fig:results:scaling-plots}
+We plot the throughput of these two implementations of linear search as a reference point for all other algorithms in Figure~\ref{fig:results:scaling-plots}.

### 2024-04-05 sections/4-results.tex
fixed table
paper edits up until section 4.3

@@
 In order to have the plots to best reflect the distribution of LFDs across the entire \textit{dataset}, we compute percentiles counting each cluster as many times as its cardinality.
-In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the total points in clusters at depth 40 belong to clusters whose LFD is at most 3.
+In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to clusters whose LFD is at most 3.
 
@@
 As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
-A consequence of this is that, with high probability, every point in the root cluster is farther away from the center than half the radius of the root cluster;
+A consequence of this is that, with high probability, for every point in the root cluster, its distance from $c$ is greater $\tfrac{r}{2}$;
 $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r)$ contains the entire dataset.
@@
 On some datasets, HNSW and ANNOY exhibit indexing times which are orders of magnitude slower than that of CAKES.
-FAISS-Flat also exhibits the fastest indexing time on each dataset. 
-This is not surprising, given that FAISS-Flat is a naive linear search algorithm and is not building an index.
+FAISS-Flat exhibits the fastest indexing time on each dataset. 
+This is not surprising, however, given that FAISS-Flat is a na\"{\i}ve linear search algorithm and is not building an index.
 
 We also highlight some differences in indexing time between different datasets.
-With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is lower than that of FAISS-IVF for all cardinalities.
-With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has lower indexing time than CAKES.
+With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is faster than that of FAISS-IVF for all cardinalities.
+With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has faster indexing time than CAKES.
 With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
@@
     \begin{center}
-        \begin{small}
+        \begin{tiny}
             \begin{sc}
-                \begin{tabular}{|c|l|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|p{1.2cm}|p{0.8cm}|}
+                \begin{tabular}{|c|l|p{1.2cm}|p{0.9cm}|p{1.2cm}|p{0.9cm}|p{1.2cm}|p{0.9cm}|p{1.2cm}|p{0.9cm}|}
                     \hline
@@
             \end{sc}
-        \end{small}
+        \end{tiny}
     \end{center}

### 2024-04-14 sections/4-results.tex
paper edits 4.3-end

@@
 We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
-In order to have the plots to best reflect the distribution of LFDs across the entire \textit{dataset}, we compute percentiles counting each cluster as many times as its cardinality.
-In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to clusters whose LFD is at most 3.
+In order to have the plots best reflect the distribution of LFDs across the entire \textit{dataset}, in computing percentiles, we count each cluster as many times as its cardinality.
+In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
 
@@
 As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
-A consequence of this is that, with high probability, for every point in the root cluster, its distance from $c$ is greater $\tfrac{r}{2}$;
-$B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r)$ contains the entire dataset.
+A consequence of this is that, with high probability, for every point in the root cluster, its distance from $c$ is greater than $\tfrac{r}{2}$;
+in other words, $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r)$ contains the entire dataset.
 Given our definition of LFD in Equation~\ref{eq:methods:lfd-original}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
@@
 
-The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{sec:methods:synthetic-data}).
+The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
 The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
@@
 On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
-While recall for CAKES's algorithms does not degrade with cardinality, recall for HNSW and ANNOY does degrade with cardinality. 
+While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY does degrade with cardinality. 
 At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift. 
@@
 For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS do not support neither the required distance functions nor, in the case of RadioML, complex-valued data.
-Due to the massive sizes of these datasets, we took random sub-samples of the dataset with lower cardinalities for our benchmarks, rather than augmented versions of the dataset.
+Due to the massive sizes of these datasets, we took random sub-samples of the dataset--rather than augmented versions of the dataset--to examine how performance scales with cardinality.
 With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
 Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
-For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the three CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
+For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance for all cardinalities.
 In particular, throughput of CAKES's algorithms is identical within three significant figures for all cardinalities.

### 2024-04-14 sections/5-discussion.tex
paper edits 4.3-end

@@
 As a consequence, these algorithms do not perform well on data with random distributions, because of the absence of a manifold structure.
-In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each of the datasets we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
+In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
 As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on datasets with low LFD, and scale linearly on datasets with high LFD.
 
-As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though they have the same cardinality and embedding dimension.
+As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though both datasets have the same cardinality and embedding dimension.
 This is unsurprising, given that the Random dataset is, by definition, uniformly distributed, while Sift is not.
@@
 While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, the throughput decreases much more slowly than it does with the Random dataset.
-Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies show the effect of a manifold structure on algorithm performance.
+Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies highlight how manifold structure affects algorithm performance.
 We emphasize that even though CAKES's algorithms are slower on the Random dataset, their recalls remain perfect at all multipliers.
@@
 the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.  
-This suggests that rather than use the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may actually want a mean which is more sensitive to high outliers, such as the geometric mean.
-We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this overshooting occurs less frequently and with less severity.
+This suggests that rather than using the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
+We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
 
@@
 This would allow us to perform fast search on datasets that are too large to fit in memory.
-An exploration of compression and compressed search is an avenue for future work.
+An exploration of compression and compressed search is another avenue for future work.
 
-Another avenue for future work is to explore the use of CAKES in a streaming environment.
+We also would like to explore the use of CAKES in a streaming environment.
 This would require the ability to perform ``online'' updates to the tree as points are added to or deleted from the dataset.

### 2024-04-21 sections/1-introduction.tex
refs

@@
 While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
-Further, such algorithms are often approximate, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains.
+Further, such algorithms are often approximate~\cite{taunk2019brief}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
 

### 2024-04-21 sections/1-introduction.tex
minor edits

@@
 Researchers are collecting data at an unprecedented scale.
-In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
-This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
+In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
+This indicates that the performance computer hardware will not ``catch up'' to computational needs in the near future.
 Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms to analyze large datasets.
@@
 a distance function, there are two common definitions of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
-$k$-NN search aims to find the $k$ most similar points to a query point, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query point.
+$k$-NN search aims to find the $k$ most similar points to a query, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query.
 

### 2024-04-21 sections/1-introduction.tex
fix ref

--- 2024-04-05 sections/1-introduction.tex
@@
 While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
-Further, such algorithms are often approximate, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains.
+Further, such algorithms are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
 

### 2024-04-21 sections/1-introduction.tex

--- 2024-04-21 sections/1-introduction.tex
@@
 While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
-Further, such algorithms are often approximate~\cite{taunk2019brief}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
+Further, such algorithms are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
 

### 2024-04-21 sections/4-results.tex
move scaling plots to supp

@@
 For each of the ANN-benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyper-parameters for these indices to achieve the highest possible recall.
+For the sake of brevity, these results are reported in the supplement.
 Though CAKES has three different algorithms, they share the same tree and so they have the same indexing time.
@@
 
-The plots in Figure~\ref{fig:results:indexing} show the results of these benchmarks.
+The plots in Figure~\ref{fig:supplement:indexing} show the results of these benchmarks.
 The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
@@
 With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
-
-\begin{figure}
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-indexing.png}\\
-        \subcaption{Fashion-mnist}
-        \label{fig:results:fashion-mnist-indexing}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/glove-25-indexing.png}\\
-        \subcaption{Glove-25}
-        \label{fig:results:glove-25-indexing}
-    \end{subfigure}
-    \vspace{1em}
-    \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift-indexing.png}\\
-        \subcaption{Sift}
-        \label{fig:results:sift-indexing}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/random-indexing.png}\\
-        \subcaption{Random}
-        \label{fig:results:random-indexing}
-    \end{subfigure}%
-    \\
-    \vspace{1em}
-    \caption{Indexing and tuning time for each algorithm with each of the ANN benchmark datasets and the Random dataset.}
-    \label{fig:results:indexing}
-\end{figure}
 

### 2024-04-21 sections/4-results.tex
move more to supp

@@
 For the sake of brevity, these results are reported in the supplement.
-Though CAKES has three different algorithms, they share the same tree and so they have the same indexing time.
-Thus, we show CAKES's indexing time only once.
-
-The plots in Figure~\ref{fig:supplement:indexing} show the results of these benchmarks.
-The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
-The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
-The vertical axis denotes the sum of indexing and tuning time in seconds.
-Both axes are on a logarithmic scale.
-Hereafter, when we refer to the ``indexing time'' of an algorithm, we are implicitly referring to the sum of indexing and tuning time for said algorithm.
-
-On all datasets, we observe that the indexing time for CAKES increases roughly linearly as cardinality increases.
-HNSW and ANNOY have the slowest indexing times across all the algorithms we benchmarked for each of the four datasets, at each cardinality. 
-On some datasets, HNSW and ANNOY exhibit indexing times which are orders of magnitude slower than that of CAKES.
-FAISS-Flat exhibits the fastest indexing time on each dataset. 
-This is not surprising, however, given that FAISS-Flat is a na\"{\i}ve linear search algorithm and is not building an index.
-
-We also highlight some differences in indexing time between different datasets.
-With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is faster than that of FAISS-IVF for all cardinalities.
-With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has faster indexing time than CAKES.
-With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
 
@@
 
-Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of the following ANN-benchmark datasets:
-
-\begin{itemize}
-    \item Fashion-Mnist under Euclidean distance,
-    \item Glove-25 under Cosine distance, and
-    \item Sift under Euclidean distance.
-\end{itemize}
+Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
+Glove-25 under Cosine distance, and
+Sift under Euclidean distance.
 

### 2024-04-21 supplement.tex
fixed supplement
arxiv submission builds
supplement fixed
move scaling plots to supp
move more to supp

@@
-\documentclass{article}
+% SIAM Supplemental File Template
+\documentclass[review,supplement,onefignum,onetabnum]{siamonline220329}
 
-% \usepackage{arxiv} # uncomment for preprint
+\input{cakes_shared}
+\begin{document}
+% Optional PDF information
+\ifpdf
+\hypersetup{
+  pdftitle={Supplementary Materials: An Example Article},
+  pdfauthor={D. Doe, P. T. Frank, and J. E. Smith}
+}
+\fi
 
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-% \usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}
-\usepackage{graphicx}
-\usepackage{subfigure}
-\usepackage{amsmath}
+\externaldocument[][nocite]{paper}
 
-\usepackage{placeins}
-% \usepackage{algpseudocode}
-\usepackage{algorithm}
+\maketitle
 
-% \algnewcommand\algorithmicforeach{\textbf{foreach}}
-% \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
+\section{Supplementary Results}
 
-\usepackage{hyperref}
+\subsection{Indexing and Tuning}
 
-\begin{document}
+The plots in Figure~\ref{fig:supplement:indexing} show the results of these benchmarks.
+The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
+The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
+The vertical axis denotes the sum of indexing and tuning time in seconds.
+Both axes are on a logarithmic scale.
+Hereafter, when we refer to the ``indexing time'' of an algorithm, we are implicitly referring to the sum of indexing and tuning time for said algorithm.
 
-\renewcommand{\thetable}{S\arabic{table}}
+On all datasets, we observe that the indexing time for CAKES increases roughly linearly as cardinality increases.
+HNSW and ANNOY have the slowest indexing times across all the algorithms we benchmarked for each of the four datasets, at each cardinality. 
+On some datasets, HNSW and ANNOY exhibit indexing times which are orders of magnitude slower than that of CAKES.
+FAISS-Flat exhibits the fastest indexing time on each dataset. 
+This is not surprising, however, given that FAISS-Flat is a na\"{\i}ve linear search algorithm and is not building an index.
 
-\renewcommand{\thefigure}{S\arabic{figure}}
+We also highlight some differences in indexing time between different datasets.
+With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is faster than that of FAISS-IVF for all cardinalities.
+With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has faster indexing time than CAKES.
+With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.
 
-\section{Results}
+\begin{figure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-indexing.png}\\
+        \subcaption{Fashion-mnist}
+        \label{fig:supplement:fashion-mnist-indexing}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{plots/glove-25-indexing.png}\\
+        \subcaption{Glove-25}
+        \label{fig:supplement:glove-25-indexing}
+    \end{subfigure}
+    \vspace{1em}
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{plots/sift-indexing.png}\\
+        \subcaption{Sift}
+        \label{fig:supplement:sift-indexing}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{plots/random-indexing.png}\\
+        \subcaption{Random}
+        \label{fig:supplement:random-indexing}
+    \end{subfigure}%
+    \\
+    \vspace{1em}
+    \caption{Indexing and tuning time for each algorithm with each of the ANN benchmark datasets and the Random dataset.}
+    \label{fig:supplement:indexing}
+\end{figure}
 
@@
 \end{figure}
+
+\subsection{Clustering}
 
@@
 
+\bibliographystyle{siamplain}
+\bibliography{references}
+
 \end{document}

### 2024-04-28 supplement.tex
remove nocite
fix supplement fig labels

@@
 
-\externaldocument[][nocite]{paper}
-
 \maketitle
@@
 The plots in Figure~\ref{fig:supplement:indexing} show the results of these benchmarks.
-The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
+The horizontal axis in each subplot shows the cardinality of the dataset augmented with synthetic points.
 The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
@@
 We also highlight some differences in indexing time between different datasets.
-With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is faster than that of FAISS-IVF for all cardinalities.
-With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has faster indexing time than CAKES.
+With Fashion-Mnist, as shown in Figure~\ref{fig:supplement:fashion-mnist-indexing}, we observe that the indexing time for CAKES is faster than that of FAISS-IVF for all cardinalities.
+With Glove-25 (see Figure~\ref{fig:supplement:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has faster indexing time than CAKES.
 With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.

### 2024-04-29 sections/1-introduction.tex
careful edit

@@
 $k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
-Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets because their cardinalities are growing exponentially.
-While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
+Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets or those whose cardinalities are growing exponentially.
+While fast algorithms for $k$-NN search on large datasets exist, most do not exploit the geometric and topological structure inherent in these datasets.
 Further, such algorithms are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
@@
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is evidence that distance functions which do not obey the triangle inequality, such as Cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
+Further, there is evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
 this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.
@@
 We provide a comparison of CAKES's algorithms to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-We also benchmark CAKES on a genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+We also benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
 
@@
 InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
-The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively.
+The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively, similarly to an early precursor to our work~\cite{yu2015entropy}.
 The number of cells used is governed by the $n_{list}$ parameter. 
@@
 CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
-In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search. 
+In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language. 
 Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search. We also discuss improvements to the performance of CHESS's $\rho$-NN search.

### 2024-04-29 sections/2-methods.tex
minor edits
more minor edits
fixed figure caption
careful edit

@@
 Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
-For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while Cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
+For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
 
 The choice of an appropriate distance function varies by dataset and domain.
-For example, with neural-network embeddings, one could use Euclidean (L2) or Cosine distance.
-With genomic or proteomic data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
+For example, with neural-network embeddings, one could use Euclidean (L2) or cosine distance.
+With genomic or proteomic sequence data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
 With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.
 
-CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
+CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
 We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
@@
 Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
-However, due to physical constraints (namely, biological evolution and the chemistry of RNA replication and mutation processes), the data are constrained to a lower-dimensional manifold within this embedding space.
+However, due to physical constraints (namely, biological evolution and biochemistry), the data are constrained to a lower-dimensional manifold within this embedding space.
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
@@
 We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
-The \textit{center} is the geometric median of the points (or a smaller sample of the points) in the \textit{cluster}, and so it is a real data point.
-The \textit{radius} is the maximum distance from the \textit{center} to any point in the \textit{cluster}.
+The center is the geometric median of the points (or a smaller sample of the points) in the cluster, and so it is a real data point.
+The radius is the maximum distance from the center to any point in the cluster.
 Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
@@
 
-Hereafter, when we refer to the LFD of a \textit{cluster}, it is estimated at the length scale of the cluster radius and half that radius.
+Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius.
 This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as:
@@
 We start by performing a divisive hierarchical clustering on the dataset using CLAM.
-The procedure is similar to that outlined in CHESS, but with the following improvements:
+The procedure is similar to that outlined in CHESS~\cite{ishaq2019clustered}, but with
 better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}) and depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
@@
 
-CLAM starts by building a binary tree of clusters using a divisive hierarchical clustering algorithm.
+CLAM builds a binary tree of clusters using a divisive hierarchical clustering algorithm.
 
@@
 Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
-This geometric median is the \textit{center} of $C$.
+This geometric median is the center of $C$.
 
@@
 
-With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers significant improvements at search-time over the approach used in CHESS.
+With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS with $\mathcal{O}(n)$ space.
 
@@
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[scale=0.8]{images/geometry/deltas.pdf}
-    \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
-    \label{fig:methods:deltas}
-\end{figure}
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[scale=0.3]{images/geometry/deltas.pdf}
+%     \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
+%     \label{fig:methods:deltas}
+% \end{figure}
 
@@
 \begin{figure}[ht!]
+    \begin{subfigure}{.5\textwidth}
         \centering
-    \includegraphics[scale=0.8]{images/geometry/overlapping-children-3.pdf}
-    \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.}
+        \includegraphics[scale=0.4]{images/geometry/deltas.pdf}
+        \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
+        \label{fig:methods:deltas}
+      \end{subfigure}%
+      \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[scale=0.4]{images/geometry/overlapping-children-3.pdf}
+        \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.
+Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query.
+Other points and distances are described in the text.}
         \label{fig:methods:overlapping-children}
+      \end{subfigure}%
+    \caption{CAKES uses geometric properties of clusters.}
 \end{figure}
@@
 From Algorithm~\ref{alg:methods:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-NN search performance scales linearly with the size of the output set and exponentially with the LFD around the query.
-While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the LFD of real-life datasets is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.
+While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the LFD of real-world datasets is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.
 
@@
 
-We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
+We perform some simple auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
 We start by taking the center of every cluster at a low depth, (e.g. 10), in the cluster tree as a query.

### 2024-04-29 sections/3-datasets.tex
careful edit

@@
 
-To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal DNA dataset~\cite{10.1093/nar/gks1219}.
-This dataset contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
+To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
+This dataset contains ribosomal RNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
 We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
 We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
-Since this dataset contains many \textit{redundant}, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition \textit{criterion} in Algorithm~\ref{alg:methods:partition}.
+Since this dataset contains many redundant, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.
 
@@
 The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
-Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
+Specifically, it comprises 24 modulation modes at 26 different signal-to-noise ratio (SNR) levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
 Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.

### 2024-04-29 sections/4-results.tex
careful edit

@@
 Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
-Glove-25 under Cosine distance, and
+Glove-25 under cosine distance, and
 Sift under Euclidean distance.
 
-Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
+Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with the same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
 Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
@@
 The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
-The vertical axis denotes throughput in queries per second.
+The vertical axis denotes throughput in queries per second (higher is better).
 Both axes are on a logarithmic scale.
@@
 Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset.
-For Fashion-Mnist, CAKES begins exhibiting sub-linear performance starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
+For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
 Which one of the three CAKES algorithms is fastest also differs by dataset.
@@
 At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift. 
-In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (because the distance function used with these datasets is a metric), and near-perfect recall on Glove-25 (because the Cosine distance function is not a metric).
+In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (because the distance function used with these datasets is a metric), and near-perfect recall on Glove-25 (because the cosine distance function is not a metric).
 
@@
 
-For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS do not support neither the required distance functions nor, in the case of RadioML, complex-valued data.
-Due to the massive sizes of these datasets, we took random sub-samples of the dataset--rather than augmented versions of the dataset--to examine how performance scales with cardinality.
-With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
+For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
+Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset--rather than augmented versions of the dataset--to examine how performance scales with cardinality.
+With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
 Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{QPS (queries per second) and Recall vs naive linear search on the Fashion-Mnist, Glove-25, Sift and Random datasets.
+    \caption{Queries per second (QPS) and Recall vs naive linear search on the Fashion-Mnist, Glove-25, Sift and Random datasets.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}

### 2024-04-29 sections/5-discussion.tex
careful edit

@@
 CAKES's algorithms are exact when the distance function in use is a metric (see definition in~\ref{sec:methods}).
-Even under Cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).
+Even under cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).
 
@@
 This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
-CAKES's tree is orders of magnitude cheaper to build and CAKES's algorithms, while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
+CAKES's tree is orders of magnitude faster and less memory-intensive to build and CAKES's algorithms, while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
 
@@
 Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
-We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in~\ref{eq:methods:repeated-rnn-complexity}.
+We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
 This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation. 

### 2024-05-15 sections/2-methods.tex
added ceiling on d in complexity bounds

@@
 Let $d$ be the LFD in this region.
-Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g. we could have a cluster overlapping the query ball at each end of each of $d$ mutually-orthogonal axes.
-In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot d \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
+Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g. we could have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
+In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
 Thus, the asymptotic complexity is
@@
 \begin{equation*}
-    \mathcal{T} \coloneqq \mathcal{O} \big( d \cdot \log \mathcal{N}_{\hat{r}}(X) \big).
+    \mathcal{T} \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big).
 \end{equation*}

### 2024-05-15 sections/5-discussion.tex
added ceiling on d in complexity bounds

@@
 
-When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{d} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
+When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{\lceil d \rceil} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
 Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.

### 2024-05-15 sections/4-results.tex
RadioML not scale free
address radioml sampling

@@
 The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.
+The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or -- as this is still a synthetic dataset -- nonuniform generation of data.
+Silva, in contrast, appears to obey the manifold hypothesis at all scales.
 

### 2024-06-17 sections/2-methods.tex
shrunk and combined algs and tables, still need to remove 50 lines of text.

@@
 Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius.
-This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as:
-
-\begin{equation}
+This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as
+$
     \log_2 \left( \frac{|C(c, r)|}{|C(c, \frac{r}{2})|} \right)
     \label{eq:methods:lfd-simplified}
-\end{equation}
-
+$
 where $|C(c, r)|$ is the cardinality of the cluster $C$ with center $c$ and radius $r$, and $C(c, \frac{r}{2})$ is the set of points in $C$ which are no more than a distance $\frac{r}{2}$ from $c$, i.e.
@@
 
-To check whether $d \leq \rho$, we note that
-
-\begin{equation*}
-    d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}.
-\end{equation*}
+To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
 
 Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
-By the Law of Cosines on $\triangle l r q$, we have that
+By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
+Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
+Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
+Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.
 
-\begin{equation*}
-    \text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}.
-\end{equation*}
-
-Since $\triangle r q q'$ is a right triangle, we also have that
-
-\begin{equation*}
-    \text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}.
-\end{equation*}
-
-Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that
-
-\begin{equation*}
-    \overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}.
-\end{equation*}
-
-Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that
-
-\begin{equation*}
-    d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}.
-\end{equation*}
-
-Thus,
-
-\begin{equation}
-    d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho.
-    \label{eq:methods:overlapping-children}
-\end{equation}
-
-Note, in particular, that Equation~\ref{eq:methods:overlapping-children} only requires distances between real points, and so it can be used with any distance function, even when $q'$ and $m$ are not real points or cannot be imputed from the data.
+Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$.
+Note, in particular, that this only requires distances between real points, and so it can be used with any distance function, even when $q'$ and $m$ are not real points or cannot be imputed from the data.
 
@@
 Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points which are no more than a distance $\rho$ from the query.
+The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
 
-\begin{algorithm} 
+\begin{minipage}{0.6\textwidth}
+\begin{algorithm}[H]
     \caption{tree-search($C$, $q$, $\rho$)}
     \label{alg:methods:rnn-search:tree-search}
+    \begin{tiny}
     \begin{algorithmic}[0]
@@
     \end{algorithmic}
+    \end{tiny}
 \end{algorithm}
-
-\begin{algorithm} 
+\end{minipage}
+\hfill
+\begin{minipage}{0.35\textwidth}
+\begin{algorithm}[H]
     \caption{leaf-search($S$, $q$, $\rho$)}
     \label{alg:methods:rnn-search:leaf-search}
+    \begin{tiny}
     \begin{algorithmic}[0]
@@
     \end{algorithmic}
+    \end{tiny}
 \end{algorithm}
+\end{minipage}
 
@@
     \label{alg:methods:rnn-search}
+    \begin{tiny}
     \begin{algorithmic}
@@
     \end{algorithmic}
+    \end{tiny}
 \end{algorithm}
-
-The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
 
@@
     \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
+    \begin{tiny}
     \begin{algorithmic}[0] % enter the algorithmic environment
@@
     \end{algorithmic}
+    \end{tiny}
 \end{algorithm}
@@
 
-The work in~\cite{yu2015entropy} showed that
-
-\begin{equation*}
-    \sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d
-\end{equation*}
-
-where $\gamma$ is a constant. 
+The work in~\cite{yu2015entropy} showed that $\sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d$, where $\gamma$ is a constant.
 By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k$.
-Thus,
-
-\begin{equation*}
-    \sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d
-\end{equation*}
-
-An estimate for $\rho_k$ is still needed. 
+Thus, $\sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d$.
+However, an estimate for $\rho_k$ is still needed.
 For this, we once again rely on the assumption that the query is drawn from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different from the LFD of nearby clusters.
-
 We let $\hat{d}$ be the (harmonic) mean LFD of the clusters in $Q$.
@@
 We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could move from the query to the center of one of the nearby clusters without significantly changing our estimate of the LFD.
-
-By Equation~\ref{eq:methods:lfd-original},
-
-\begin{equation*}
-    \hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}
-\end{equation*}
-
-We can rearrange this equation to get
-
-\begin{equation*}
-    \frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}
-\end{equation*}
-
-Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get
-
-\begin{equation*}
-    k \left( 1 + 2 \cdot \left( \frac{\hat{|C|}}{k} \right) ^ {\hat{d}^{-1}} \right)^d.
-\end{equation*}
+By Equation~\ref{eq:methods:lfd-original}, $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$.
+We can rearrange this equation to get $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
+Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get $k \left( 1 + 2 \cdot \left( \frac{\hat{|C|}}{k} \right) ^ {\hat{d}^{-1}} \right)^d$.
 
@@
 
-\begin{algorithm} % enter the algorithm environment
+\begin{minipage}{0.46\textwidth}
+\begin{algorithm}[H] % enter the algorithm environment
     \caption{Breadth-First Sieve($root$, $q$, $k$)} % give the algorithm a caption
     \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
+    \begin{tiny}
     \begin{algorithmic}[0] % enter the algorithmic environment
@@
     \end{algorithmic}
+    \end{tiny}
 \end{algorithm}
-
-\subsubsection{Depth-First Sieve}
-\label{sec:methods:knn-search:depth-first-sieve}
-
-This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
-
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
-$Q$ starts containing only the root cluster while $H$ starts empty.
-So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
-
-\begin{itemize}
-    \item While the top priority element $Q$ is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top priority element (a leaf) from $Q$ and add all its points to $H$.
-    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
-\end{itemize}
-
-This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
-This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
-
-\begin{algorithm} 
+\end{minipage}
+\hfill
+\begin{minipage}{0.46\textwidth}
+\begin{algorithm}[H]
     \caption{Depth-First Sieve($root$, $q$, $k$)}
     \label{alg:methods:depth-first-sieve}
+    \begin{tiny}
     \begin{algorithmic}[0]
@@
     \end{algorithmic}
+    \end{tiny}
 \end{algorithm}
+\end{minipage}
+
+\subsubsection{Depth-First Sieve}
+\label{sec:methods:knn-search:depth-first-sieve}
+
+This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
+
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
+$Q$ starts containing only the root cluster while $H$ starts empty.
+So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
+
+\begin{itemize}
+    \item While the top priority element $Q$ is not a leaf, remove it from $Q$ and add its children to $Q$.
+    \item Remove the top priority element (a leaf) from $Q$ and add all its points to $H$.
+    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
+\end{itemize}
+
+This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
 
@@
 In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
-Thus, the asymptotic complexity is
-
-\begin{equation*}
-    \mathcal{T} \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big).
-\end{equation*}
-
-For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is
-
-\begin{equation*}
-    \mathcal{L} \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right).
-\end{equation*}
+Thus, the asymptotic complexity is $\mathcal{T} \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
+For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $\mathcal{L} \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$.
 

### 2024-06-17 sections/1-introduction.tex
shrunk to 700 lines

@@
 Researchers are collecting data at an unprecedented scale.
-In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
-This indicates that the performance computer hardware will not ``catch up'' to computational needs in the near future.
+In many fields, datasets grow exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
+This indicates that the performance of computing systems will not keep pace with the growth of data.
 Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms to analyze large datasets.
@@
 Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image} are a common source of neural network embeddings.
-These embeddings are often high-dimensional, and the sizes of training and inference datasets for such networks are growing exponentially.
-Among biological datasets, the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length, while SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
-Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
+Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
+Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated signals of different modulation modes.
 
-Many researchers are especially interested in similarity search on these datasets. 
+Many researchers are interested in similarity search on these datasets.
 Similarity search enables a variety of applications, including recommendation~\cite{annoy} and classification systems~\cite{suyanto2022knnclassifier}.
-As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become extremely challenging; 
+As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become challenging;
 even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{Malkov2016EfficientAR, johnson2019billion, annoy, aumuller2020ann}.
 
-Given some measure of similarity between data points, e.g.
-a distance function, there are two common definitions of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
+Given some measure of similarity between data points, there are two common definitions of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
 $k$-NN search aims to find the $k$ most similar points to a query, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query.
-
-Previous works have used the term \textit{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \textit{approximate} for search algorithms which do not exhibit perfect recall when compared to a na\"{i}ve linear search.
+Previous works have used the term \textit{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \textit{approximate} for algorithms which do not exhibit perfect recall.
 In contrast, an \textit{exact} search algorithm exhibits perfect recall.
@@
 $k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
-Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets or those whose cardinalities are growing exponentially.
-While fast algorithms for $k$-NN search on large datasets exist, most do not exploit the geometric and topological structure inherent in these datasets.
-Further, such algorithms are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
-
+Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
+While fast algorithms for $k$-NN search on large datasets exist, they are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
 For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
@@
 Further, there is evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
-this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.
+this suggests that approximate $k$-NN search could perform poorly in such contexts.
 
@@
 GPS and other location-based services use $\rho$-NN search to find nearby points of interest~\cite{zhang2020privacy} to a user's location.
-In all of these cases, approximate $\rho$-NN search may be sufficient, but exact $\rho$-NN search is preferable.
+%In all of these cases, approximate $\rho$-NN search may be sufficient, but exact $\rho$-NN search is preferable.
 
 This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy Scaling Search), a set of three novel algorithms for exact $k$-NN search.
-We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
-We provide a comparison of CAKES's algorithms to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-We also benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+We also present improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
+We provide a comparison to several current algorithms for similarity search; namely, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We further benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+Finally, we compare real-world datasets to an artificially-generated dataset obeying simple statistical properties.
 
@@
 
-Recent search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent $k$-nearest neighbor search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 However, some of these algorithms do not provide exact search (as defined in Section \ref{sec:introduction} above).
 
-\subsubsection{HNSW}
-\label{sec:introduction:related-works:hnsw}
-
-Hierarchical Navigable Small World networks~\cite{Malkov2016EfficientAR} is an approximate $k$-NN search method based on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
-Similar to the NSW algorithm, HNSW builds a graph of the dataset, but unlike NSW, the graph is multi-layered.
-The query point and each data point are inserted into the graph one at a time, and, upon insertion, the element is joined by an edge to the $M$ nearest nodes in the in the graph, where $M$ is a tunable parameter.
-The highest layer in which an element can be placed is determined randomly with an exponentially decaying probability distribution.
+Hierarchical Navigable Small World networks~\cite{Malkov2016EfficientAR} relies on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
+HNSW builds a multi-layered graph of the dataset.
+A query point and each data point are inserted into the graph and joined by an edge to the $M$ nearest nodes in the in the graph, where $M$ is a tunable parameter.
+The highest layer in which an element can be placed is determined randomly with an exponentially-decaying probability distribution.
 Search starts at the highest layer and descends to the lowest layer, greedily following a path of edges to the nearest node, until reaching the query point.
@@
 
-\subsubsection{FAISS-IVF}
-\label{sec:introduction:related-works:faiss-ivf}
-
-InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
-The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively, similarly to an early precursor to our work~\cite{yu2015entropy}.
+InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} clusters data into high-dimensional Voronoi cells, and whichever cell a query point falls into is then searched exhaustively, similarly to an early precursor to our work~\cite{yu2015entropy}.
 The number of cells used is governed by the $n_{list}$ parameter.
@@
 
-\subsubsection{ANNOY}
-\label{sec:introduction:related-works:annoy}
-
-This algorithm~\cite{annoy} is based on random projection and tree building for approximate $k$-NN search.
+ANNOY~\cite{annoy} relies on random projection and tree building for approximate $k$-NN search.
 At each intermediate node of the tree, two points are randomly sampled from the space, and the hyperplane equidistant from them is chosen to divide the space into two subspaces.
-This process is repeated multiple times to create a forest of trees, and the number of times the process is repeated is a tunable parameter.
+This process iterates to create a forest of trees, and the number of iterations is a tunable parameter.
 At search time, one can increase the number of trees to be searched to improve recall at the cost of speed.
 
-\subsubsection{Entropy-Scaling Search}
-\label{sec:introduction:related-works:entropy-scaling-search}
-
-This search paradigm exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \textit{metric entropy} and \textit{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
+The entropy-scaling search paradigm exploits the geometric and topological structure inherent in large datasets.
+Importantly, as suggested by their name, entropy-scaling search algorithms exhibit asymptotic complexity that scales with topological properties of the dataset, instead of its cardinality.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a hierarchical clustering approach.
 CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
 In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language.
-Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search. We also discuss improvements to the performance of CHESS's $\rho$-NN search.

### 2024-06-17 sections/2-methods.tex
shrunk to 700 lines

@@
 We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
-A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
+A distance of zero defines identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
 We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
@@
 In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
-While we sometimes use Euclidean notions, such as voids, volumes and proximity to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
+While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
 they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
-
 CAKES exploits the low LFD of such datasets to accelerate search.
@@
 Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases.
-When the vast majority of points in the dataset have low ($\ll D$) LFD, we can overload terminology to say that the dataset has low LFD.
-
+When the vast majority of points in the dataset have low ($\ll D$) LFD, we can simply say that the dataset has low LFD.
 We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
+To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomic sequences with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
 Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
-However, due to physical constraints (namely, biological evolution and biochemistry), the data are constrained to a lower-dimensional manifold within this embedding space.
+However, physical constraints (namely, biological evolution and biochemistry) constrain the data to a lower-dimensional manifold within this space.
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
@@
 CLAM builds a binary tree of clusters using a divisive hierarchical clustering algorithm.
-
 For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
@@
 The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.
-
 We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
 Without loss of generality, we assign to the left child those points which are equidistant from the two poles.
-
 Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc.
 This process is described in Algorithm \ref{alg:methods:partition}.
-
 During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-simplified}.
@@
 
-After building the tree, we reorder the dataset so that the points are stored in the order that they would be visited during a depth-first traversal of the tree.
+%After building the tree, we reorder the dataset so that the points are stored in the order that they would be visited during a depth-first traversal of the tree.
 
@@
 Assuming a balanced tree, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.
-
-One may improve this memory overhead to $\mathcal{O}(n)$ by only storing indices at the leaf clusters.
-This approach, however, introduces a time cost of $\mathcal{O}(n \log n)$ whenever we need to find the indices for a non-leaf cluster (because it requires a traversal of the subtree rooted at that cluster).
-
 In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
-Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset.
+Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset, based on a useful invariant.
 The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
-A left child has the same offset as that of its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
-
+A left child has the same offset its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
 With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS with $\mathcal{O}(n)$ space.
@@
 
-We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains only duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \textit{unique} point).
+We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains only duplicates of the same point.
 A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
@@
 To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
-
 Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
@@
 \begin{itemize}
-    \item While the top priority element $Q$ is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top priority element (a leaf) from $Q$ and add all its points to $H$.
+    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
+    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
     \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
@@
 
-Due to their similarity, we combine the complexity analyses of the two Sieve methods into one section.
-For these methods we, once again, use the terminology of tree-search and leaf-search.
+Due to their similarity, we combine the complexity analyses of both Sieve methods.
+For these methods we again use the terminology of tree-search and leaf-search.
 Tree-search navigates the cluster tree and adds clusters to $Q$.
@@
 The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
+Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is:
 
@@
 
-For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the complexity of the priority queue operations.
-Thus, the asymptotic complexity is
+For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
+Thus, the complexity is:
 
@@
 This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.
-
 This augmentation process preserves the topological structure of the original dataset, but increases its cardinality by a factor of $m$, allowing us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.

### 2024-06-17 sections/3-datasets.tex
shrunk to 700 lines

@@
 
-We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, using the associated distance function.
-Table~\ref{tab:datasets:summary} summarizes the properties of these datasets.
-
+We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}.
+Table~\ref{tab:datasets:summary} summarizes these datasets.
 The benchmarks for Fashion-Mnist, Glove-25, Sift, and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
@@
 The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.
-
 The benchmarks for SILVA and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
@@
 In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Table~\ref{tab:results:qps-and-recall}) increases.
-
-We also benchmarked on purely random (i.e., not an augmented version of a known dataset), datasets of various cardinalities.
+We also benchmarked on purely randomly-generated datasets of various cardinalities.
 For this, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset; hereafter, we refer to the random dataset with cardinality 1,000,000 and dimensionality 128 as ``Random.''
@@
 
-To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
+To demonstrate CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
 This dataset contains ribosomal RNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
@@
 
-To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset~\cite{oshea2018radioml}.
+As another example of an exotic distance function, we benchmarked CAKES on the RadioML dataset~\cite{oshea2018radioml}.
 The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.

### 2024-06-17 sections/4-results.tex
shrunk to 700 lines

@@
 Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
-
-Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for (the non-augmented versions of) Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
-The horizontal axis denotes the depth in the cluster tree, and the vertical axis denotes the LFD (as calculated by Equation~\ref{eq:methods:lfd-simplified}) of clusters at that depth.
+Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
+The horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
 We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
-In order to have the plots best reflect the distribution of LFDs across the entire \textit{dataset}, in computing percentiles, we count each cluster as many times as its cardinality.
+In order to have the plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality.
 In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
-
 Figure~\ref{fig:results:fashion-mnist-lfd} shows the LFD by depth for Fashion-Mnist.
-We observe that until about depth 5, the LFD is low, as the 95$^{th}$ percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
-For depths 15 through 25, we observe that the LFD increases, with the 95$^{th}$ percentile (orange plotted line) slightly less than 6, and the median near 3.
+We observe that until about depth 5, LFD is low, as the 95$^{th}$ percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
+For depths 15 through 25, we observe that LFD increases, with the 95$^{th}$ percentile (orange plotted line) slightly less than 6, and the median near 3.
 Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95$^{th}$ percentile is between 3 and 4, and the median is less than 2.
@@
 Sift under Euclidean distance.
-
 Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with the same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
@@
 For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.
-
 The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
-The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
-The vertical axis denotes throughput in queries per second (higher is better).
+The left-most point on each line is at the cardinality of the original dataset.
+The vertical axis denotes throughput in queries per second.
 Both axes are on a logarithmic scale.
@@
 Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the tuned CAKES algorithm.
-When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$ when we consider only three decimal places.
+When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$.
 
@@
 For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
-Additionally, on all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset. 
+On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset.
 With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
@@
 On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
-While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY does degrade with cardinality. 
+While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
 At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
-In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (because the distance function used with these datasets is a metric), and near-perfect recall on Glove-25 (because the cosine distance function is not a metric).
+In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
 
-In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms are the slowest out of all algorithms, at all cardinalities.
+In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly.
 As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
@@
 Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
-For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance for all cardinalities.
-In particular, throughput of CAKES's algorithms is identical within three significant figures for all cardinalities.
+For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance.
+In particular, throughput of CAKES's algorithms is identical within three significant figures.
 

### 2024-06-17 sections/5-discussion.tex
shrunk to 700 lines

@@
 CAKES's algorithms are exact when the distance function in use is a metric (see definition in~\ref{sec:methods}).
-Even under cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall (i.e. always greater than $0.9995$).
-
+Even under cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall.
 CAKES's algorithms are designed to be most effective when the data uphold the manifold hypothesis, or in other words, when the data are constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
@@
 
-As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though both datasets have the same cardinality and embedding dimension.
-This is unsurprising, given that the Random dataset is, by definition, uniformly distributed, while Sift is not.
-On the Random dataset, CAKES's algorithms' speed decreases linearly as the multiplier increases, while with Sift, Depth-First Sieve and Breadth-First Sieve both exhibit nearly constant throughput.
-While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, the throughput decreases much more slowly than it does with the Random dataset.
+As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though both datasets have the same cardinality and dimensionality.
+This is unsurprising, given that the Random dataset is uniformly distributed, while Sift is not.
+On the Random dataset, CAKES's speed decreases linearly as the multiplier increases, while with Sift, Depth-First Sieve and Breadth-First Sieve both exhibit nearly constant throughput.
+While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, throughput decreases much more slowly than it does with the Random dataset.
 Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies highlight how manifold structure affects algorithm performance.
-We emphasize that even though CAKES's algorithms are slower on the Random dataset, their recalls remain perfect at all multipliers.
-
-Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near constant throughput as the multiplier increases, however, their recall continues to decrease as the multiplier increases.
-This, coupled with the prohibitively increasing cost (in terms of memory and time) of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the dataset cardinalities are growing exponentially.
-CAKES's tree is orders of magnitude faster and less memory-intensive to build and CAKES's algorithms, while a little slower than HNSW and ANNOY, still show near constant scaling with cardinality in addition to having perfect recall.
+We emphasize that although CAKES's algorithms are slower on the Random dataset, their recalls remain perfect.
+Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near-constant throughput as the multiplier increases, but their recall continues to decrease as the multiplier increases.
+This, coupled with the increasing time and space cost of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the datasets are growing exponentially.
+CAKES's tree is orders of magnitude faster and less memory-intensive to build, and CAKES's algorithms, while slower than HNSW and ANNOY, still show near-constant scaling as well as perfect recall.
 
@@
 These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
-Exploring this sensitivity is a potential avenue for future investigation, as it remains to be determined at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
+Exploring this sensitivity is a potential avenue for future study, as it remains to be seen at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
 
-While the fastest CAKES algorithm varies by dataset, we observe that on all three ANN-benchmark datasets (Fashion-Mnist, Glove-25, and Sift), Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as the cardinality increases.
-This observation warrants further investigation, but it is especially promising given that Depth-First Sieve and Breadth-First Sieve consistently outperform linear search for high cardinalities.
+While the fastest CAKES algorithm varies by dataset, we note that on all three ANN-benchmark datasets (Fashion-Mnist, Glove-25, and Sift), Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as the cardinality increases.
+This observation warrants further investigation, but it is especially promising given that both algorithms consistently outperform linear search for high cardinalities.
 Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities supports our use of an auto-tuning function (as discussed in Section~\ref{sec:methods:auto-tuning}) to select the best algorithm for a given dataset.

### 2024-12-28 reviewer-response.md
response to reviewers started

Dear Dr. Becker,

We thank you and the anonymous reviewers for the thoughtful and insightful suggestions regarding our manuscript. Below, please find a point-by-point accounting of our responses to the reviews.

Associate Editor comments (Remarks to Author):
I have sent the manuscript out to two referees and received those reports back, and I have also looked at your paper myself. Both referees have identified quite a few items they would like to see addressed, but overall I'm pleased to report that I think the manuscript has merit and impact, and so I am recommending a major revision. In addition to rephrasing, reorganizing and clarifying the paper (per the referees' comments), the two major items to address are (1) details about the real world examples (cf comments from referee 1), and (2) referee 2 raises an issue about different ground truths.

Referee #2 (Remarks to the Author):

In "Let them have CAKES [...]", the authors describe a new method to perform
variants of the nearest neighbor search. The Rust implementation of this new
algorithm is relatively (comparable or somewhat slower than the competition) but
has a much better recall across many different tasks.

The description of the method is clear, although the description of the run time
costs is ambiguous. Overall, CAKES is an interesting contribution to the field
of NN-search.

# Major comments

* The big O notation by default suggest worst-case run time complexity. In this
manuscript, it seems that complexity is a mixture of expected run time
complexity and under the loose assumption that the data has particular
characteristics (e.g., low fractal dimension, the tree is balanced). These
assumptions should be made clear in the intro and when giving the run time
expressions.

For example, line 215, does the cost of partitioning of O(n log(n)) depend on
the tree being balanced? It is an important hypothesis because line 197 states
that "the imbalance in the tree is a feature, not a bug, as it reflects the
underlying structure of the data". That may look like a contradiction.

The cost on line 260 uses \hat{r}, which is a mean radius. Now that cost is in
expectation (in addition to the balanced tree hypothesis).

Another example, line 326, comes from the minimum with value 2 on line 284. If
the assumption on the LFD is not satisfied, then an extra log factor would
appear in the cost.

* The discussion on LFD after its definition line 141 is hard to understand:
there are 2 parameters in the definition of LFD (r_1 and r_2), but the
discussion seems to imply there is only 1 (a radius r). That discussion only
seems to make sense after line 163 and the rewrite of the definition 2.1.

* Smith-Waterman and Needleman-Wunsch are not distances (unlike Levenshtein and
Hamming distances). They are dynamic programming algorithms to compute
distances like the Levenshtein distance (and others depending on the exact
parameters of the algorithms).

* The explanation that the augmentation process (starting at line 418) preserves
the topological structure of the underlying data is not very convincing. This
process adds small clouds of points extending in every dimension around
existing points. That is not quite the same as generating new points that
belong to a small dimension manifold in the sample space.

* Line 467-473: Why are there multiple linear search implementations used as
ground truth? These different implementation are apparently used to evaluate
different algorithms (ANNOY, FAISS-IVF vs CAKES). This is troubling: why would
the ground truth differ based on which tool is evaluated or which language it
is written in?

Do these 2 tools return the same result? If so, why use multiple tools? If
not, then the evaluation is flawed.

# Minor comments

* Line 421: The points are created with distance \epsilon of x, not \epsilon
||x||. r is chosen in the sphere of radius \epsilon and added to x, and, as
stated correctly line 424, x' is within \epsilon of x.

* There are many small LaTeX issues: wrong space after "i.e." and "e.g." (add a
backslash), large spaces around formulas (e.g., line 342), em-dashes that look
too short (e.g., line 594).

* The figures should not be shrank so much: the labels on the axis and the keys
are not readable. Regenerate the pictures at the correct size to avoid
scaling.

* Put the keys outside of the figures, not over the data lines.

* Many horizontal and vertical lines in Table~2 should be removed to help with
readability.

* Line 585: "we observe performs quite slowly". The sentence seems incomplete.

* Many of the explicit algorithms do not serve any real purpose: they are not
more readable in pseudo-code form and are just fine in explained in the text.
(E.g., Algorithm 2.4).

* The Github repository gives very little instructions on how to compile the
code, run it or reproduce the results from this manuscript.

				Review of

	Let them have CAKES: A Cutting-Edge Algorithm for Scalable,
		Efficient, and Exact Search on Big Data

			SIMODS manuscript M166272

MAJOR ISSUES =============================================================

The CAKES manuscript proposes an entropy scaling approach to the search
problems of $k$ Nearest Neighbors ($k$-NN) and $rho$ Nearest Neighbors
($\rho-NN) in a data set $X$.  The general approach is to construct a
hierarchical data structure of clusters to assist in the searching.
While there are interesting ideas in the manuscript, there are major
issues both technically and expositionally, as well as some minor issues,
that need to be addressed before another round of reviewing.

While there is implicitly a tree that is the key data structure, the tree
itself is never named or given a notation; the same is true of the root of
the tree.  The authors should review the requirements for defining proper
terminology and notation for a data structure and apply it consistently to
their data structure.

The authors appear to use the terms similarity and distance
interchangeably, though they are certainly different.  In particular, a
similarity is a measure (nominally, non-negative) that is (ideally) larger
the closer two data points are and, in some biological contexts, has no
upper bound.  Moreover, the use of a distance measure alone is inadequate
for the described algorithms.  On line 155, there is mention of the
geometric median of the points, but there may be no such median depending
on the space (set) containing the points of the data set.  For example,
there is no obvious median for a set of biological sequences.  The authors
must identify the actual spaces from which data sets can be extracted for
CAKES to work.

On line 141, there is a function defined for the local fractal dimension
(LFD), but it is not expressed in functional notation.  In other words, it
should start with $LFD(X,r_1,r_2)$, since it is clearly a function of all
three.  This functional notation should propagate throughout the rest of
the manuscript.  Much the same is true on line 142, where it should read
$B(X,q,r)$, not $B_X(q,r)$.  If makes little sense to put $X$ as a
subscript in this context.

In the algorithms, starting with Algorithm 2.1, the pseudocode style is
ugly.  Much better would be the pseudocode style from the fourth edition
of Introduction to Algorithms, CLRS.  Also, the description of $criteria$
is as a "stopping criteria", while the logic of the code has it as a
continuing criteria.  Please fix the pseudocode throughout.

Section 2.2 is meant to be the definition of the search problem(s), but
that is really only contained in the paragraph from lines 218 to 223. In
any case, it is essential to put the problem(s) to be solved earlier, that
is, first, before any description of the solutions with ideas such as
clusters.  This movement of text will require a bit of text refactoring at
a minimum.  Also, the sentence at lines 221 to 223 should be expanded in
full detail into a separate paragraph.  Finally, one assumes that the
integer $k$ or the number $\rho$ is a parameter to the corresponding
problem and should be mentioned as such.

There are no theorems in the manuscript, yet there are arguments for
quantities like the quality of the results or the time complexity that are
given without much precision.  It is better to give precise theorems or
propositions for each thing to be proved, with appropriate assumptions,
and to give precise proofs of same.

The manuscript makes too much use of unfamiliar terminology taken from
earlier papers without defining terms.  For example, see line 233 and
Reference [18].  See also the paragraph at lines 306 to 310. All terms
used must be precisely defined in the manuscript to allow proper
assessment of the manuscript.

The Results section is a slog to read and parts of the Discussion seem
redundant.  Take lines 651 to 668 as an example.  Tightening up the
Results and the Discussion can only help the exposition.

I led a bioinformatics project that used an entropy scaling approach to
search a database of protein sequences.  See Reference [1].  As a result,
I am confused about some of the real world examples.  I know that protein
sequences are challenging to search for reasons such as their having
widely varying lengths and there being no obvious notion of a center of a
set of protein sequences.  In any case, I would like the authors to
reference [1] and discuss in more depth how they address the challenges of
the various real world examples.

MINOR ISSUES =============================================================

"Data set" is pronounced as two words and should be written as such
throughout the manuscript.

In line 22, use "rate" instead of "scale".

Throughout, the word "which" is used when "that" is the correct word.  If
"which" is used, then it is not defining and is preceded by a comma.
Change all "which"'s to "that"'s.

On line 75, there is a Subsection 1.1, but there is no Subsection 1.2.
This is not a good practice.  Just refactor the text without Section 1.1
or add a Subsection 1.2.

The use of $\forall$, $\exists$, $\Leftrightarrow$, and other similar
logical operators is not appropriate in a manuscript that is not about
logic.  Spell out the meaning in English, and be creative.

Near line 154, it is important to give an overview of your solution
strategy and, in particular, why clustering is used first.  Clustering is
a heavily overloaded word.  I would have chosen a more neutral term
instead of "cluster".

On line 155, there is the parenthetical comment about "a smaller sample of
the points".  What does that mean precisely and where does it occur later
in the manuscript?

On line 156, there is the claim "so it is a real data point".  What is a
real data point?  Please explain.

On line 161 and elsewhere, "i.e." is always preceded and followed by
commas.

On line 178, there is the phrase "find the point ... in the sample".  Can
you make the entire paragraph more precise by giving notation for the
sample?

Section 2.1.1 overall is lacking in mathematical rigor.

In Section 2.1.3, the cost of assessing "criteria' is not counted in the
time complexity, but it must be.

In line 252, there is again mention of "real points" without any
indication of the meaning of "real".

The fonts in Algorithms 2.2 through 2.7 are simply too small.  Make them
the font size of the remainder of the text.

The numbers in Table 2 are too small.  Refactor into multiple tables
perhaps.

In line 641, what is "radial increase 2.3"?

In line 642, why is "harmonic" italicized?

The References section needs a through going over for proper formatting
and capitalization.  For example, check the title capitalization in [2],
[4], and [8].  Check the journal name capitalization in [9] and [21],
among others.

Several references have repeated URLs.  Trim these up.

References [2] and [3] appear to be duplicates.

Reference [39] is incomplete and inadequate.

REFERENCE ================================================================

[1] Yoonjin Kim, Zhen Guo, Jeffrey A. Robertson, Benjamin Reidys, Ziyan
Zhang, Lenwood S. Heath, EnTrance: Exploration of Entropy Scaling Ball
Cover Search in Protein Sequences, bioRxiv
https://urldefense.com/v3/__https://www.biorxiv.org/content/10.1101/2021.05.31.446458v1__;!!Jh1S!kGffVkV6p8O_T-q3TzzRIBjJpAmGsw1kVTB7zosA7khu71pJpSxxrEqcYu6KceYMTg_VZ_MJ8B-PoPqf$ , 2021.

### 2024-12-28 reviewer-response.md
notes in reviewer response

@@
 
+*this is easy to address, and we will do so*
+
 For example, line 215, does the cost of partitioning of O(n log(n)) depend on
@@
 
+*we can address this with the plots that include a balanced tree -- expand 2.1.3, balanced vs unbalanced, dropping off points faster with an unbalanced tree?*
+
 The cost on line 260 uses \hat{r}, which is a mean radius. Now that cost is in
 expectation (in addition to the balanced tree hypothesis).
+
+*this is restating the complexity from the CHESS paper, but it's correct that it's an expected cost. Worst-case analysis isn't really useful here.*
 
@@
 appear in the cost.
+
+*write in plain english right after eq 2.3, explaining in plain english the rationale for the min(2,x) expansion: namely, we don't think the very low LFD would be representative of what we'd see if we massively increased the radius*
 
@@
 
+*Right after or in eq 2.1, immediately follow up with the simplified form where r_1 = 2r_2. Also drop subscript X, since we're dealing with only one dataset. Explain.*
+
 * Smith-Waterman and Needleman-Wunsch are not distances (unlike Levenshtein and
@@
 parameters of the algorithms).
+
+*Correct; fix in language.*
 
@@
 belong to a small dimension manifold in the sample space.
+
+*We should acknowledge this limitation, and suggest that in future work we would do something like favor augmentation along the first few PCs of the local manifold, but not in scope here.
+This goes in discussion.*
 
@@
 
+*We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results.*
+
 # Minor comments
@@
 
+*x is a vector, and epsilon is a scalar. We can clarify that x is a vector. Be sure to boldface all vectors*
+
 * There are many small LaTeX issues: wrong space after "i.e." and "e.g." (add a
@@
 too short (e.g., line 594).
+
+*easy to fix (NMD)*
 
@@
 
+*fix some font sizes and aspect ratios*
+
 * Put the keys outside of the figures, not over the data lines.
+
+*need to fix this*
 
@@
 
+*investigate this, or break up table 2 into 2a-2d (NMD)*
+
 * Line 585: "we observe performs quite slowly". The sentence seems incomplete.
+
+*will fix -> 'we observe that CAKES performes quite slowly'*
 
@@
 
+*get rid of alg 2.4*
+
 * The Github repository gives very little instructions on how to compile the
 code, run it or reproduce the results from this manuscript.
+
+*will fix (NI)*
 
@@
 
+*we fixed this in pancakes, just backport it*
+
 The authors appear to use the terms similarity and distance
@@
 
+*add def in paper for geometric median, extended from continues vector spaces into any metric space. line 43, use distance, not similarity*
+
 On line 141, there is a function defined for the local fractal dimension
@@
 
+*similar to complaint from reviewer 2*
+
 In the algorithms, starting with Algorithm 2.1, the pseudocode style is
@@
 continuing criteria.  Please fix the pseudocode throughout.
+
+*make this pseudocode style change (NMD)*
+*call it continuation criteria (NI)*
 
@@
 
+*move introduction of the three deltas to later, to start of 2.4.3, and expand definition of rnn. (NI)*
+
 There are no theorems in the manuscript, yet there are arguments for
@@
 and to give precise proofs of same.
+
+*state theorem, which is the time complexity, and formalize the argument into a proof*
 
@@
 
+*remind the reader what rnn search is, and in 306-310 reword to refer to previous section rather than external refs*
+
 The Results section is a slog to read and parts of the Discussion seem
@@
 Results and the Discussion can only help the exposition.
+
+*offload to figure captions where possible, don't spell out everything*
 
@@
 
+*fine, cite 1, and discuss (NI)*
+
 MINOR ISSUES =============================================================
@@
 
+*disagree, this appears as a term of art in existing literature, and is in the Merriam-Webster Dictionary (cite)*
+
 In line 22, use "rate" instead of "scale".
+
+*ok, will fix*
 
@@
 
+*no. maybe change some selected ones, but this is too rigid a grammatical rule.*
+
 On line 75, there is a Subsection 1.1, but there is no Subsection 1.2.
@@
 
+*make the last paragraph (entropy scaling) into 1.2*
+
 The use of $\forall$, $\exists$, $\Leftrightarrow$, and other similar
@@
 logic.  Spell out the meaning in English, and be creative.
+
+*disagree, these operators are standard in mathematics and theoretical computer science*
 
@@
 
+*we define it, and are building on prior work that relies on clustering. wontfix.*
+
 On line 155, there is the parenthetical comment about "a smaller sample of
@@
 
+*we take a root-n sample of the points (NI), justified in CHESS paper*
+
 On line 156, there is the claim "so it is a real data point".  What is a
@@
 
+*deal with by defining geometric median*
+
 On line 161 and elsewhere, "i.e." is always preceded and followed by
 commas.
+
+*fixed here*
 
@@
 
+*deal with by defining geometric median*
+
 Section 2.1.1 overall is lacking in mathematical rigor.
+
+*we will refactor 2.1.1 to be more rigorous (NI)*
 
@@
 
+*add some explanation: these can be as complex as a user wants, but for us it is a constant-time check (NI)*
+
 In line 252, there is again mention of "real points" without any
 indication of the meaning of "real".
+
+*replace with 'actual points from the dataset'*
 
@@
 
+*we will make this more legible (NMD)*
+
 The numbers in Table 2 are too small.  Refactor into multiple tables
@@
 
+*we are refactoring into multiple tables (NMD)*
+
 In line 641, what is "radial increase 2.3"?
 
+*missing word 'equation' and also put in parentheses*
+
 In line 642, why is "harmonic" italicized?
+
+*no need*
 
@@
 
+*Will go over this carefully (NMD)*
+
 REFERENCE ================================================================

### 2024-12-28 sections/2-methods.tex
notes in reviewer response

@@
 Based on the assumption that the LFD near the query does not differ significantly from that of nearby clusters, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
-one iteration to find at least one cluster, and the one more to find enough the $k$ neighbors.
+one iteration to find at least one cluster, and one more to find enough of the $k$ neighbors.
 Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e. $\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.

### 2024-12-30 sections/2-methods.tex
wip: changes by NI

@@
 \begin{equation}
-    \frac{\text{log} \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
+    \text{LFD}(q, r_1, r_2) = \frac{\text{log} \left( \frac{|B(q, r_1)|}{|B(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
     \label{eq:methods:lfd-original}
 \end{equation}
-where $B_X(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
+
+where $B(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
+
+We use a simplified version of this definition by calculating the LFD at a length scale where $r_1 = 2 \cdot r_2$.
+
+\begin{equation}
+    \text{LFD}(q, r) = \text{log}_2 \left( \frac{|B(q, r)|}{|B(q, \frac{r}{2})|} \right)
+    \label{eq:methods:lfd-half}
+\end{equation}
+
 Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases.
@@
 Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
+For real-world datasets, we expect the LFD to be locally uniform, i.e., when $r$ is small, but potentially highly variable at larger length scales, i.e., when $r$ is large.
 
@@
 We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
-The center is the geometric median of the points (or a smaller sample of the points) in the cluster, and so it is a real data point.
+The center is the geometric median of the points in the cluster, i.e., it is the point which minimizes the sum of distances to all other points in the cluster.
+In cases where the cardinality of the cluster is large, we take a random subsample of $\sqrt{|C|}$ points and compute the geometric median of that subsample~\cite{ishaq2019clustered}.
+The center, therefore, is one of the points in the cluster and is used as a representative of the cluster.
 The radius is the maximum distance from the center to any point in the cluster.
@@
 Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
-As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e. $C(c, r) \subset B_X(c, r)$.
+As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e. $C(c, r) \subset B(c, r)$.
 
-Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius.
-This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as
-$
-    \log_2 \left( \frac{|C(c, r)|}{|C(c, \frac{r}{2})|} \right)
-    \label{eq:methods:lfd-simplified}
-$
-where $|C(c, r)|$ is the cardinality of the cluster $C$ with center $c$ and radius $r$, and $C(c, \frac{r}{2})$ is the set of points in $C$ which are no more than a distance $\frac{r}{2}$ from $c$, i.e.
-$C(c, \frac{r}{2}) = \{p : p \in C \land f(c, p) \leq \frac{r}{2}\}$.
+Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius, i.e. using Equation~\ref{eq:methods:lfd-half}.
+We also only use points that are in $C(c, r)$ instead of all those in $B(c, r)$.
 
-The \textit{metric entropy} $\mathcal{N}_{r}(X)$ for some radius $r$ was defined, in~\cite{yu2015entropy}, as the minimum number of clusters of a uniform radius $r$ needed to cover the data.
-In this paper, where the clustering is hierarchical rather than flat, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ of a dataset $X$ as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
-
-We start by performing a divisive hierarchical clustering on the dataset using CLAM.
-The procedure is similar to that outlined in CHESS~\cite{ishaq2019clustered}, but with
-better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}) and depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
+The \textit{metric entropy} $\mathcal{N}_{r}(X)$ for some radius $r$ is the minimum number of clusters of a uniform radius $r$ needed to cover the data~\cite{yu2015entropy} for a flat clustering with clusters of radius $r$.
+In this paper, where the clustering is hierarchical rather than flat, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
 
@@
 
-CLAM builds a binary tree of clusters using a divisive hierarchical clustering algorithm.
+We start by performing a divisive hierarchical clustering on the dataset using CLAM.
+The procedure is almost identical to that outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}).
+After building the tree, we also perform a depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
+
 For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
@@
 This process is described in Algorithm \ref{alg:methods:partition}.
-During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-simplified}.
+During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-half}.
 
@@
         \Require $C$, a cluster
-        \Require $criteria$, user-specified stopping criteria
+        \Require $criteria$, user-specified continuation criteria
 
@@
 
-% % WE WILL USE THIS FOR COMPRESSION. DO NOT DELETE
-
-% \subsubsection{Scaling Behavior of Cluster Radii}
-% \label{sec:methods:clustering:scaling-behabior-of-cluster-radii}
-
-% While it may be tempting to assume that cluster radii decrease with each application of Partition (refer to Algorithm \ref{alg:methods:partition}), unfortunately, this assumption is incorrect.
-% Fortunately, we \textit{can} make some guarantees about the scaling behavior of cluster radii;
-% in particular, we prove in this section that cluster radii will stop increasing after at most $d$ partitions, where $d$ is the fractal dimensionality of the dataset.
-
-% We can describe a $d$-dimensional distribution of data by choosing some set of $d$ mutually orthogonal axes.
-% Let $2R$ be the maximum pairwise distance among the points in the dataset.
-% We choose the axes such that the two points that are $2R$ apart lie \textit{along one of the axes}.
-% Thus, a $d$-dimensional hyper-sphere of radius $R$ would bound the dataset.
-% In the worst case, (i.e., with a uniform-density distribution that fills the $d$-sphere), our axes will be such that $2R$ is the maximum pairwise distance \textit{along every axis}.
-% Such a distribution would also produce a balanced cluster tree.
-
-% In this case, Partition will select a maximally distant pair of points to use as poles;
-% without loss of generality, say it chooses two points at the extrema one of the $d$ axes.
-% After one partition, the maximum pairwise distance along that axis will be halved, i.e. bounded above by $R$.
-% The next recursive Partition will select another of the $d$ axes, and once again, the distance along that axis will be bounded above by $R$
-% Thus, after at most $d$ partitions, the maximum pairwise distance along each axis will be bounded above by $R$.
-% The overall (i.e., not restricted to one axis) maximum pairwise distance
-% will be bounded above by $R\sqrt{2}$ by, for example, two instances that lie at the extrema of different axes.
-% See Figure~\ref{fig:methods:radii-scaling-behavior} for an example.
-
-% Thus, starting with a cluster $C$ of radius $R$, after at most $d$ Partitions, the descendants of $C$ will each have radius
-% bounded above by $\frac{R}{\sqrt{2}}$. In other words, cluster radii are guaranteed to decrease by a multiplicative factor of $\frac{\sqrt{2}}{2}$ after at
-% most $d$ partitions.
-
-% Note that, in practice, we never see a balanced clustering.
-% Partition produces unbalanced trees due to the varying density of the sampling in different regions of the manifold and the low dimensional ``shape'' of the manifold.
-% Further, in practice, the cluster radii decrease by a factor much larger than $\frac{\sqrt{2}}{2}$ with almost every partition;
-% the upper bound of $d$ partitions is seldom realized.
-
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=3.4in]{images/geometry/geometry.pdf}
-%     \caption{
-%         Scaling behavior of radii on a uniform-density disk, which represents the worst-case scenario for a two-dimensional distribution.
-%         $R_0$ denotes the radius of the root cluster $C_0$, and $o_0$ denotes its center.
-%         After one application of Algorithm~\ref{alg:methods:clustering:partition}, we have $C_1$, with radius $R_1$ and center $o_1$.
-%         The right triangle formed by $o_0$, $o_1$, and $y_+$ in $C_1$ shows that $R_0 < R_1$.
-%         Hence, the radius of a child cluster can be larger than its parent.
-%         However, after another application of Algorithm~\ref{alg:methods:clustering:partition}, we have consumed all $d$ orthogonal axes,
-%         as shown in $C_2$.
-%         Now, clearly $R_2 < R_0$.
-%         In fact, for each subsequent application of Algorithm~\ref{alg:methods:clustering:partition}, the radius of the resulting cluster is bounded above by $\frac{\sqrt{2}R}{2}$.
-%     }
-%     \label{fig:methods:radii-scaling-behavior}
-% \end{figure}
-
 \subsubsection{Complexity}
@@
 
-\subsection{The Search Problem}
-\label{sec:methods:the-search-problem}
+This complexity analysis ignores the cost of checking the continuation criteria.
+While a user could specify criteria of any complexity (our implementation allows for this), for this paper we continue clustering until each cluster is a singleton, i.e., it contains only one point or duplicates of the same point and has a radius of zero.
+This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
 
-Given Algorithm~\ref{alg:methods:partition}, we can now formally pose the $k$-NN and $\rho$-NN search problems.
+We also note here that this $\mathcal{O}(n \log n)$ complexity assumes that the tree is balanced.
+This can realistically only happen when the data are uniformly distributed in the embedding space.
+This represents the worst-case scenario for subsequent search algorithms because the LFD of the dataset will approach or equal the embedding dimension.
+For real-world datasets with low LFD, we expect the tree to be highly unbalanced, reflecting the topological structure of the data.
+When building trees for such datasets, their depths would be larger than $\log n$, which would seem to increase the cost of building
+However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of the tree would also have fewer points, and so the cost of building each subsequent level would be lower.
+Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
+In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.
 
-Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, we define the $k$-NN search problem, which aims to find the $k$ closest points to $q$ in $\textbf{X}$;
-in other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
-We also have the $\rho$-NN search problem, which aims to find the all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$: the set $H = B_X(q, \rho) = \{x \in \textbf{X}: f(q, x) \leq \rho \}$.
+% TODO: Move most of this to the supplement
+% \subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
+% \label{sec:methods:rnn-search}
 
-Given a cluster $C$, let $c$ be its center and $r$ be its radius.
-Our $\rho$-NN and $k$-NN search algorithms make use of the following properties, as illustrated in Figure~\ref{fig:methods:deltas}.
-
-\begin{itemize}
-    \item $\delta = f(q, c)$, the distance from the query to the cluster center $c$.
-    \item $\delta^{+} = \delta + r$, the distance from the query to the theoretically farthest point in $C$.
-    \item $\delta^{-} = \text{max}(0, \delta - r)$, the distance from the query to the theoretically closest point in $C$.
-\end{itemize}
+% We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
+% when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children which might contain points in the query ball.
 
@@
 %     \centering
-%     \includegraphics[scale=0.3]{images/geometry/deltas.pdf}
-%     \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
-%     \label{fig:methods:deltas}
+%     \includegraphics[scale=0.75]{images/geometry/overlapping-children-3.pdf}
+%     \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.
+Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query.
+Other points and distances are described in the text.}
+%     \label{fig:methods:overlapping-children}
+%     \caption{CAKES uses geometric properties of clusters.}
 % \end{figure}
 
-We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains only duplicates of the same point.
-A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
-Hence, we overload the above notation to also refer to the distance from a query to an individual point.
+% To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children}.
+% Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.
 
-\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
-\label{sec:methods:rnn-search}
+% Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{sec:methods:clustering:building-the-tree}).
+% Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
+% Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
+% As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $\rho < d$, then the left child cannot contain points inside the query ball.
+% In such a case we proceed to search only the right child.
+% Otherwise, we proceed with both children.
 
-We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
-when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children which might contain points in the query ball.
+% To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
+% Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
+% By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
+% Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
+% Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
+% Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.
 
-\begin{figure}[ht!]
-    \begin{subfigure}{.5\textwidth}
+% Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$.
+Note, in particular, that this only requires distances between actual points from the dataset, and so it can be used with any distance function, even when $q'$ and $m$ are not actual points or cannot be imputed from the data.
+
+% To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
+% Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points which are no more than a distance $\rho$ from the query.
+% The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
+
+% \begin{gather}
+%     \mathcal{O}
+%     \Bigg(
+%         \underbrace{
+%             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+%         }_{\textrm{tree-search}}
+%         \ + \
+%         \underbrace{
+%             \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
+%             \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
+%         }_{\textrm{leaf-search}}
+%     \Bigg)
+%     \label{eq:methods:rnn-search-complexity}
+% \end{gather}
+% where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
+
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
+\label{sec:methods:knn-search}
+
+Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster of the tree built using Algorithm~\ref{alg:methods:partition}, we can now formally pose the $k$-NN search problem.
+
+The $k$-NN search problem, given a user specified integer $k$ and query $q$, aims to find the $k$ closest points to $q$ in $\textbf{X}$.
+In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
+
+In this section, we present three novel algorithms for exact $k$-NN search:
+Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
+
+In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure which stores the closest points to the query found so far, and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
+The two Sieve algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+
+\begin{figure}
     \centering
-        \includegraphics[scale=0.4]{images/geometry/deltas.pdf}
-        \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
+    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
+    \caption{
+        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
+        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
+        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+    }
     \label{fig:methods:deltas}
-      \end{subfigure}%
-      \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[scale=0.4]{images/geometry/overlapping-children-3.pdf}
-        \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.
-Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query.
-Other points and distances are described in the text.}
-        \label{fig:methods:overlapping-children}
-      \end{subfigure}%
-    \caption{CAKES uses geometric properties of clusters.}
-\end{figure}
+\end{figure}%
 
-To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children}.
-Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
+\label{sec:methods:knn-search:repeated-rnn}
 
-Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{sec:methods:clustering:building-the-tree}).
-Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
-Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
-As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $\rho < d$, then the left child cannot contain points inside the query ball.
-In such a case we proceed to search only the right child.
-Otherwise, we proceed with both children.
+This algorithm relies on the \textit{tree-search} and \textit{leaf-search} algorithms described in~\cite{ishaq2019clustered}.
+We reproduce the algorithms here for completeness.
 
-To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
-Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
-By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
-Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
-Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
-Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.
-
-Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$.
-Note, in particular, that this only requires distances between real points, and so it can be used with any distance function, even when $q'$ and $m$ are not real points or cannot be imputed from the data.
-
-To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
-Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points which are no more than a distance $\rho$ from the query.
-The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
-
-\begin{minipage}{0.6\textwidth}
-\begin{algorithm}[H]
-    \caption{tree-search($C$, $q$, $\rho$)}
+\begin{minipage}{.425\textwidth}
+    \begin{algorithm}[H]\small
+        \caption{tree-search($C$, $q$, $r$)}
         \label{alg:methods:rnn-search:tree-search}
-    \begin{tiny}
-    \begin{algorithmic}[0]
+        \begin{algorithmic}
             \Require $C$, a cluster
             \Require $q$, a query
-        \Require $\rho$, a search radius
+            \Require $r$, a search radius
 
-        \If{$C$ is a leaf \textbf{or} $\delta^+_C \leq \rho$}
+            \If{$C$ is a leaf \textbf{or} $\delta^+_C \leq r$}
                 \State \textbf{return} $\{C\}$
             \Else
-
                 \State $[l, r]$ $\Leftarrow$ \textit{pole}s of $C$
                 \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
+                \State \textbf{return} tree-search($L, q, r$)
 
-            \If{$\overline{q l} < \overline{q r \vphantom{l}}$}
-                \State $[r, l] \Leftarrow [l, r]$
-                \State $[R, L] \Leftarrow [L, R]$
+                \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
             \EndIf
 
-            \If{$(\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$}
-                \State \textbf{return} tree-search($L, q, \rho$) $\cup$ tree-search($R, q, \rho$)
-            \Else
-                \State \textbf{return} tree-search($R, q, \rho$)
-            \EndIf
-        \EndIf
         \end{algorithmic}
-    \end{tiny}
     \end{algorithm}
@@
 \hfill
-\begin{minipage}{0.35\textwidth}
-\begin{algorithm}[H]
-    \caption{leaf-search($S$, $q$, $\rho$)}
+\begin{minipage}{.45\textwidth}
+    \begin{algorithm}[H]\small
+        \caption{leaf-search($Q$, $q$, $r$)}
         \label{alg:methods:rnn-search:leaf-search}
-    \begin{tiny}
-    \begin{algorithmic}[0]
-        \Require $S$, a set of clusters
+        \begin{algorithmic}
+            \Require $Q$, a set of clusters
             \Require $q$, a query
-        \Require $\rho$, a search radius
+            \Require $r$, a search radius
 
-        \State $H \Leftarrow \emptyset$, a set of hits
+            \State $H \Leftarrow \emptyset$, a set of 2-tuples of points and their distances to the query
 
-        \For{$C \in S$}
-            \If{$\delta^+_C \leq \rho$}
+            \For{$C \in Q$}
+                \If{$\delta^+_C \leq r$}
                     \State $H$ $\Leftarrow$ $H \cup \{C\}$
@@
                     \For{$p \in C$}
-                    \If{$f(p, q) \leq \rho$}
-                        \State $H$ $\Leftarrow$ $H \cup \{p\}$
+                        \If{$f(p, q) \leq r$}
+                            \State $H$ $\Leftarrow$ $H \cup \{(p, f(p, q))\}$
                         \EndIf
@@
         \end{algorithmic}
-    \end{tiny}
     \end{algorithm}
@@
 
-\begin{algorithm}
-    \caption{$\rho$-NN-search($root$, $q$, $\rho$)}
-    \label{alg:methods:rnn-search}
-    \begin{tiny}
-    \begin{algorithmic}
-        \Require $root$, the root cluster
-        \Require $q$, a query
-        \Require $\rho$, a search radius
-
-        \State $S$ $\Leftarrow$ tree-search($root$, $q$, $\rho$)
-        \State $H$ $\Leftarrow$ leaf-search($S$, $q$, $\rho$)
-        \State \textbf{return} $H$
-    \end{algorithmic}
-    \end{tiny}
-\end{algorithm}
-
-\begin{gather}
-    \mathcal{O}
-    \Bigg(
-        \underbrace{
-            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-        }_{\textrm{tree-search}}
-        \ + \
-        \underbrace{
-            \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
-            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
-        }_{\textrm{leaf-search}}
-    \Bigg)
-    \label{eq:methods:rnn-search-complexity}
-\end{gather}
-where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
-
-From Algorithm~\ref{alg:methods:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-NN search performance scales linearly with the size of the output set and exponentially with the LFD around the query.
-While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the LFD of real-world datasets is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.
-
-\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
-\label{sec:methods:knn-search}
-
-In this section, we present three novel algorithms for exact $k$-NN search:
-Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-
-In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure which stores the closest points to the query found so far, and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
-
-\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{sec:methods:knn-search:repeated-rnn}
-
-In this algorithm, we perform $\rho$-NN search starting with a small search radius, and repeatedly increasing the radius until $|H| \geq k$.
-
-Let the search radius $r$ be equal to the radius of the root cluster divided by the cardinality of the dataset.
-We perform tree-search with radius $r$.
+For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
 If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
-Let $Q$ be the set of clusters returned by the first tree-search which returns at least one cluster.
 
@@
 Thus, the factor of radius increase should also be \textit{inversely} related to the LFD.
+However, we should not increase the radius too drastically with any one iteration because we assume that the LFD is only \textit{locally} uniform.
+A large increase in the radius would likely break out of the local region and potentially encounter too many new clusters, which would make the subsequent step computationally expensive.
 
-Once $\sum_{C \in Q} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we can stop increasing the radius.
-We perform $\rho$-NN search with this radius and return the $k$ nearest neighbors.
+Once $\sum_{C \in Q} |C| \geq k$, we have found at least $k$ neighbors, and so we can stop increasing the radius.
+We now perform \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors among the points in the clusters found during the previous steps.
 
-\begin{algorithm} % enter the algorithm environment
+\begin{algorithm}\small % enter the algorithm environment
     \caption{Repeated $\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
     \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
-    \begin{tiny}
-    \begin{algorithmic}[0] % enter the algorithmic environment
+    \begin{algorithmic} % enter the algorithmic environment
         \Require $root$, the root cluster
@@
             \EndIf
-            \State $\mu \Leftarrow \frac{|S|}{\sum_{C \in Q} \frac{1}{LFD(C)}}$
+            \State $\mu \Leftarrow |Q| \cdot \left( \sum_{C \in Q} LFD(C)^{-1} \right)^{-1}$
             \State $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)$
-            \State $S \Leftarrow$ tree-search($root$, $q$, $r$)
+            \State $Q \Leftarrow$ tree-search($root$, $q$, $r$)
         \EndLoop
 
-        \State $H \Leftarrow$ sort($\rho$-NN-search($root$, $q$, $r$))
-        \State \textbf{return} $H[.. k]$
+        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
+        \State \textbf{sort} $H$ by non-decreasing distance to $q$
+        \State \textbf{return} The first $k$ points in $H$
     \end{algorithmic}
-    \end{tiny}
 \end{algorithm}
@@
 
-We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \textit{tree-search} and \textit{leaf-search} separately.
+We find it useful to start by discussing the \textit{tree-search} and \textit{leaf-search} stages separately.
 Tree-search refers to the process of identifying clusters which have overlap with the query ball, or in other words, clusters which might contain one of the $k$ nearest neighbors.
@@
 
-In ~\cite{ishaq2019clustered}, we showed that the complexity of tree-search is $\mathcal{O}(\log\mathcal{N}_{\hat{r}}(X))$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset $X$ at a radius $\hat{r}$.
-To adjust this bound for Repeated $\rho$-NN, we must estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
+In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
 
-Based on the assumption that the LFD near the query does not differ significantly from that of nearby clusters, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
+\begin{gather}
+    \mathcal{O}
+    \Bigg(
+        \underbrace{
+            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+        }_{\textrm{tree-search}}
+        \ + \
+        \underbrace{
+            \overbrace{ \big| B(q, \rho) \big|}^{\textrm{output size}}
+            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
+        }_{\textrm{leaf-search}}
+    \Bigg)
+    \label{eq:methods:rnn-search-complexity}
+\end{gather}
+where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
+
+To build on this and determine the complexity of Repeated $\rho$-NN, we must first estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
+
+Based on the assumption that the LFD near the query does not differ significantly from that of the clusters near the query (the set $Q$), Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
 one iteration to find at least one cluster, and one more to find enough of the $k$ neighbors.
@@
 To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree-search.
-Since we must examine every point in each such clusters, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$.
+Since we must examine every point in each such cluster, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$.
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
@@
 By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has an asymptotic complexity of
-
 \begin{gather}
@@
 \end{gather}
-
 where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
@@
 During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:the-search-problem}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 
@@
 
-\begin{minipage}{0.46\textwidth}
-\begin{algorithm}[H] % enter the algorithm environment
+\subsubsection{Depth-First Sieve}
+\label{sec:methods:knn-search:depth-first-sieve}
+
+This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
+
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
+$Q$ starts containing only the root cluster while $H$ starts empty.
+So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
+
+\begin{itemize}
+    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
+    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
+    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
+\end{itemize}
+
+This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
+
+Note that this algorithm is not truly a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
+Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
+
+\begin{minipage}{0.5125\textwidth}
+    \begin{algorithm}[H]\small
         \caption{Breadth-First Sieve($root$, $q$, $k$)} % give the algorithm a caption
         \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
-    \begin{tiny}
-    \begin{algorithmic}[0] % enter the algorithmic environment
+        \begin{algorithmic} % enter the algorithmic environment
             \Require $root$, the root cluster
@@
 \hfill
-\begin{minipage}{0.46\textwidth}
-\begin{algorithm}[H]
+\begin{minipage}{0.425\textwidth}
+    \begin{algorithm}[H]\small
         \caption{Depth-First Sieve($root$, $q$, $k$)}
         \label{alg:methods:depth-first-sieve}
-    \begin{tiny}
-    \begin{algorithmic}[0]
+        \begin{algorithmic}
             \Require $root$, the root cluster
@@
         \end{algorithmic}
-    \end{tiny}
     \end{algorithm}
 \end{minipage}
-
-\subsubsection{Depth-First Sieve}
-\label{sec:methods:knn-search:depth-first-sieve}
-
-This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
-
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
-$Q$ starts containing only the root cluster while $H$ starts empty.
-So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
-
-\begin{itemize}
-    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
-    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
-\end{itemize}
-
-This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
-This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
-
-Note that this algorithm is not truly a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
-Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
 
@@
 This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.
-This augmentation process preserves the topological structure of the original dataset, but increases its cardinality by a factor of $m$, allowing us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.
+This augmentation process does not add to the overall topological structure of the dataset, but it does increase its cardinality by a factor of $m$.
+This allows us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.

### 2025-01-02 sections/4-results.tex
wip: updating table and plots

@@
     \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/lfd/random-1000000.png}\\
+        \subcaption{A random dataset}
+        \label{fig:results:random-lfd}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
         \includegraphics[width=0.95\textwidth]{images/lfd/radio-ml-97920.png}\\
@@
     \end{subfigure}%
-    \\
     \begin{subfigure}[b]{0.47\textwidth}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/random-1000000.png}\\
-        \subcaption{A random dataset}
-        \label{fig:results:random-lfd}
-    \end{subfigure}
     \vspace{1em}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}
+        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
         \subcaption{Fashion-Mnist for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}
+        \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
         \subcaption{Glove-25 for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}
+        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{Sift for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
+        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+        \subcaption{A random dataset for $k=10$.}
+        \label{fig:results:random-scaling}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
@@
     \vspace{1em}
-    \\
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/silva-knn-10.png}
+        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{Silva for $k=10$.}
         \label{fig:results:silva-scaling}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
-        \subcaption{A random dataset for $k=10$.}
-        \label{fig:results:random-scaling}
     \end{subfigure}%
@@
     Note that for RadioML, all three algorithms performed nearly identically, so distinct lines are not visible.
-    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e. more then 512GB).}
+    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e. more then 512GB).
+    For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}
     \label{fig:results:scaling-plots}
@@
     \begin{center}
-        \begin{tiny}
-            \begin{sc}
-                \begin{tabular}{|c|l|p{1.2cm}|p{0.9cm}|p{1.2cm}|p{0.9cm}|p{1.2cm}|p{0.9cm}|p{1.2cm}|p{0.9cm}|}
+        \begin{small}
+            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
                 \hline
-                    & \multirow{2}{*}{\textbf{Multiplier}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
+                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
                 & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\\cline{2-10}
@@
                 \multirow{10}{*}{\rotatebox[origin=c]{90}{\textbf{Fashion-Mnist}}}
-                    & 1   & 1.33~$\times10^{4}$ & 0.954  & 2.19~$\times10^{3}$ & 0.950  & 2.01~$\times10^{3}$ & 1.000* & 2.17~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 2   & 1.38~$\times10^{4}$ & 0.803  & 2.12~$\times10^{3}$ & 0.927  & 9.39~$\times10^{2}$ & 1.000* & 1.14~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 4   & 1.66~$\times10^{4}$ & 0.681  & 2.04~$\times10^{3}$ & 0.898  & 4.61~$\times10^{2}$ & 0.997  & 9.82~$\times10^{2}$ & 1.000  \\\cline{2-10}
-                    & 8   & 1.68~$\times10^{4}$ & 0.525  & 1.93~$\times10^{3}$ & 0.857  & 2.26~$\times10^{2}$ & 0.995  & 1.18~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 16  & 1.87~$\times10^{4}$ & 0.494  & 1.84~$\times10^{3}$ & 0.862  & 1.17~$\times10^{2}$ & 0.991  & 1.20~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 32  & 1.56~$\times10^{4}$ & 0.542  & 1.85~$\times10^{3}$ & 0.775  & 5.91~$\times10^{1}$ & 0.985  & 1.16~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 64  & 1.50~$\times10^{4}$ & 0.378  & 1.78~$\times10^{3}$ & 0.677  & 2.61~$\times10^{1}$ & 0.968  & 1.10~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 128 & 1.49~$\times10^{4}$ & 0.357  & 1.66~$\times10^{3}$ & 0.538  & 1.33~$\times10^{1}$ & 0.964  & 1.04~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 256 & --                  & --     & 1.60~$\times10^{3}$ & 0.592  & 6.65~$\times10^{0}$ & 0.962  & 1.06~$\times10^{3}$ & 1.000  \\\cline{2-10}
-                    & 512 & --                  & --     & 1.83~$\times10^{3}$ & 0.581  & 3.56~$\times10^{0}$ & 0.949  & 1.04~$\times10^{3}$ & 1.000  \\
+                & 1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{2-10}
+                & 2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{2-10}
+                & 4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{2-10}
+                & 8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{2-10}
+                & 16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{2-10}
+                & 32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{2-10}
+                & 64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{2-10}
+                & 128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{2-10}
+                & 256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{2-10}
+                & 512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949  & \num{2.84e3} & 1.000  \\
                 \hline
@@
                 \multirow{9}{*}{\rotatebox[origin=c]{90}{\textbf{Glove-25}}}
-                    & 1   & 2.28~$\times10^{4}$ & 0.801 & 2.83~$\times10^{3}$ & 0.835 & 2.38~$\times10^{3}$ & 1.000* & 7.22~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 2   & 2.38~$\times10^{4}$ & 0.607 & 2.70~$\times10^{3}$ & 0.832 & 1.19~$\times10^{3}$ & 1.000* & 5.75~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 4   & 2.50~$\times10^{4}$ & 0.443 & 2.61~$\times10^{3}$ & 0.839 & 6.19~$\times10^{2}$ & 1.000* & 6.25~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 8   & 2.78~$\times10^{4}$ & 0.294 & 2.51~$\times10^{3}$ & 0.834 & 3.03~$\times10^{2}$ & 1.000* & 5.93~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 16  & 3.11~$\times10^{4}$ & 0.213 & 2.23~$\times10^{3}$ & 0.885 & 1.51~$\times10^{2}$ & 1.000* & 5.49~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 32  & 3.24~$\times10^{4}$ & 0.178 & 2.01~$\times10^{3}$ & 0.764 & 7.40~$\times10^{1}$ & 0.999  & 4.75~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 64  & --                  & --    & 1.99~$\times10^{3}$ & 0.631 & 3.77~$\times10^{1}$ & 0.997  & 4.61~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 128 & --                  & --    & --                  & --    & 1.90~$\times10^{1}$ & 0.998  & 4.41~$\times10^{2}$ & 1.000* \\\cline{2-10}
-                    & 256 & --                  & --    & --                  & --    & 9.47~$\times10^{0}$ & 0.998  & 4.23~$\times10^{2}$ & 1.000* \\
+                & 1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{2-10}
+                & 2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{2-10}
+                & 4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{2-10}
+                & 8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{2-10}
+                & 16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{2-10}
+                & 32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{2-10}
+                & 64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{2-10}
+                & 128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{2-10}
+                & 256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
                 \hline
@@
                 \multirow{8}{*}{\rotatebox[origin=c]{90}{\textbf{Sift}}}
-                    & 1   & 1.93~$\times10^{4}$ & 0.782 & 3.98~$\times10^{3}$ & 0.686 & 6.98~$\times10^{2}$ & 1.000* & 5.52~$\times10^{2}$ & 1.000 \\\cline{2-10}
-                    & 2   & 2.03~$\times10^{4}$ & 0.552 & 3.80~$\times10^{3}$ & 0.614 & 3.30~$\times10^{2}$ & 1.000* & 2.66~$\times10^{2}$ & 1.000 \\\cline{2-10}
-                    & 4   & 2.18~$\times10^{4}$ & 0.394 & 3.69~$\times10^{3}$ & 0.637 & 1.65~$\times10^{2}$ & 1.000* & 1.43~$\times10^{2}$ & 1.000 \\\cline{2-10}
-                    & 8   & 2.48~$\times10^{4}$ & 0.298 & 3.58~$\times10^{3}$ & 0.710 & 7.72~$\times10^{1}$ & 1.000* & 7.94~$\times10^{1}$ & 1.000 \\\cline{2-10}
-                    & 16  & 2.68~$\times10^{4}$ & 0.210 & 3.50~$\times10^{3}$ & 0.690 & 3.98~$\times10^{1}$ & 1.000* & 8.12~$\times10^{1}$ & 1.000 \\\cline{2-10}
-                    & 32  & 2.75~$\times10^{4}$ & 0.193 & 3.44~$\times10^{3}$ & 0.639 & 2.09~$\times10^{1}$ & 0.999  & 7.81~$\times10^{1}$ & 1.000 \\\cline{2-10}
-                    & 64  & --                  & --    & 3.39~$\times10^{3}$ & 0.678 & 8.87~$\times10^{0}$ & 0.997  & 7.43~$\times10^{1}$ & 1.000 \\\cline{2-10}
-                    & 128 & --                  & --    & 3.36~$\times10^{3}$ & 0.643 & 4.78~$\times10^{0}$ & 0.993  & 6.80~$\times10^{1}$ & 1.000 \\
+                & 1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{2-10}
+                & 2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{2-10}
+                & 4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{2-10}
+                & 8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{2-10}
+                & 16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{2-10}
+                & 32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{2-10}
+                & 64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{2-10}
+                & 128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
                 \hline
@@
                 \multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{Random}}}
-                    & 1  & 1.17~$\times10^{4}$ & 0.060 & 4.28~$\times10^{3}$ & 0.028 & 7.34~$\times10^{2}$ & 1.000* & 5.54~$\times10^{2}$ & 1.000 \\\cline{2-10}
-                    & 2  & 1.01~$\times10^{4}$ & 0.048 & 4.04~$\times10^{3}$ & 0.021 & 3.58~$\times10^{2}$ & 1.000* & 2.69~$\times10^{2}$ & 1.000 \\\cline{2-10}
-                    & 4  & 9.12~$\times10^{3}$ & 0.031 & 3.64~$\times10^{3}$ & 0.014 & 1.90~$\times10^{2}$ & 1.000* & 1.37~$\times10^{2}$ & 1.000 \\\cline{2-10}
-                    & 8  & 8.35~$\times10^{3}$ & 0.022 & 3.37~$\times10^{3}$ & 0.013 & 8.84~$\times10^{1}$ & 1.000* & 5.69~$\times10^{1}$ & 1.000 \\\cline{2-10}
-                    & 16 & 8.25~$\times10^{3}$ & 0.008 & 3.17~$\times10^{3}$ & 0.006 & 4.36~$\times10^{1}$ & 1.000* & 2.61~$\times10^{1}$ & 1.000 \\\cline{2-10}
-                    & 32 & --                  & --    & 3.01~$\times10^{3}$ & 0.007 & 1.72~$\times10^{1}$ & 1.000* & 1.35~$\times10^{1}$ & 1.000 \\
+                & 1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{5.54e2} & 1.000 \\\cline{2-10}
+                & 2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.69e2} & 1.000 \\\cline{2-10}
+                & 4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.37e2} & 1.000 \\\cline{2-10}
+                & 8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{5.69e1} & 1.000 \\\cline{2-10}
+                & 16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.61e1} & 1.000 \\\cline{2-10}
+                & 32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.35e1} & 1.000 \\
                 \hline
             \end{tabular}
-            \end{sc}
-        \end{tiny}
+        \end{small}
     \end{center}

### 2025-01-02 supplement.tex
wip: updating table and plots

@@
     \centering
-    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-100.png}
+    \includegraphics[width=3.4in]{plots/fashion-mnist_PermutedBall_100_throughput.png}
     \caption{
@@
     \centering
-    \includegraphics[width=3.4in]{plots/sift-knn-100.png}
+    \includegraphics[width=3.4in]{plots/sift_PermutedBall_100_throughput.png}
     \caption{
@@
     \centering
-    \includegraphics[width=3.4in]{plots/glove-25-knn-100.png}
+    \includegraphics[width=3.4in]{plots/glove-25_PermutedBall_100_throughput.png}
     \caption{
@@
     \centering
-    \includegraphics[width=3.4in]{plots/random-knn-100.png}
+    \includegraphics[width=3.4in]{plots/sift_PermutedBall_100_throughput.png}
     \caption{

### 2025-01-05 reviewer-response.md
swapped italics so our responses are plain text and reviewer comments are italicized.

--- 2024-12-28 reviewer-response.md
@@
 
-Referee #2 (Remarks to the Author):
+**Referee #2 (Remarks to the Author):**
 
-In "Let them have CAKES [...]", the authors describe a new method to perform
+*In "Let them have CAKES [...]", the authors describe a new method to perform
 variants of the nearest neighbor search. The Rust implementation of this new
 algorithm is relatively (comparable or somewhat slower than the competition) but
-has a much better recall across many different tasks.
+has a much better recall across many different tasks.*
 
-The description of the method is clear, although the description of the run time
+*The description of the method is clear, although the description of the run time
 costs is ambiguous. Overall, CAKES is an interesting contribution to the field
-of NN-search.
+of NN-search.*
 
@@
 
-* The big O notation by default suggest worst-case run time complexity. In this
+*The big O notation by default suggest worst-case run time complexity. In this
 manuscript, it seems that complexity is a mixture of expected run time
@@
 assumptions should be made clear in the intro and when giving the run time
-expressions.
+expressions.*
 
-For example, line 215, does the cost of partitioning of O(n log(n)) depend on
+this is easy to address, and we will do so
+
+*For example, line 215, does the cost of partitioning of O(n log(n)) depend on
 the tree being balanced? It is an important hypothesis because line 197 states
 that "the imbalance in the tree is a feature, not a bug, as it reflects the
-underlying structure of the data". That may look like a contradiction.
+underlying structure of the data". That may look like a contradiction.*
 
-The cost on line 260 uses \hat{r}, which is a mean radius. Now that cost is in
-expectation (in addition to the balanced tree hypothesis).
+we can address this with the plots that include a balanced tree -- expand 2.1.3, balanced vs unbalanced, dropping off points faster with an unbalanced tree?
 
-Another example, line 326, comes from the minimum with value 2 on line 284. If
+*The cost on line 260 uses \hat{r}, which is a mean radius. Now that cost is in
+expectation (in addition to the balanced tree hypothesis).*
+
+this is restating the complexity from the CHESS paper, but it's correct that it's an expected cost. Worst-case analysis isn't really useful here.
+
+*Another example, line 326, comes from the minimum with value 2 on line 284. If
 the assumption on the LFD is not satisfied, then an extra log factor would
-appear in the cost.
+appear in the cost.*
 
-* The discussion on LFD after its definition line 141 is hard to understand:
+write in plain english right after eq 2.3, explaining in plain english the rationale for the min(2,x) expansion: namely, we don't think the very low LFD would be representative of what we'd see if we massively increased the radius
+
+*The discussion on LFD after its definition line 141 is hard to understand:
 there are 2 parameters in the definition of LFD (r_1 and r_2), but the
 discussion seems to imply there is only 1 (a radius r). That discussion only
-seems to make sense after line 163 and the rewrite of the definition 2.1.
+seems to make sense after line 163 and the rewrite of the definition 2.1.*
 
-* Smith-Waterman and Needleman-Wunsch are not distances (unlike Levenshtein and
+Right after or in eq 2.1, immediately follow up with the simplified form where $r_1 = 2r_2$. Also drop subscript $X$, since we're dealing with only one dataset. Explain.
+
+*Smith-Waterman and Needleman-Wunsch are not distances (unlike Levenshtein and
 Hamming distances). They are dynamic programming algorithms to compute
 distances like the Levenshtein distance (and others depending on the exact
-parameters of the algorithms).
+parameters of the algorithms).*
 
-* The explanation that the augmentation process (starting at line 418) preserves
+Correct; fix in language.
+
+*The explanation that the augmentation process (starting at line 418) preserves
 the topological structure of the underlying data is not very convincing. This
@@
 existing points. That is not quite the same as generating new points that
-belong to a small dimension manifold in the sample space.
+belong to a small dimension manifold in the sample space.*
 
-* Line 467-473: Why are there multiple linear search implementations used as
+We should acknowledge this limitation, and suggest that in future work we would do something like favor augmentation along the first few PCs of the local manifold, but not in scope here.
+This goes in discussion.
+
+*Line 467-473: Why are there multiple linear search implementations used as
 ground truth? These different implementation are apparently used to evaluate
@@
 the ground truth differ based on which tool is evaluated or which language it
-is written in?
+is written in?*
 
-Do these 2 tools return the same result? If so, why use multiple tools? If
-not, then the evaluation is flawed.
+*Do these 2 tools return the same result? If so, why use multiple tools? If
+not, then the evaluation is flawed.*
+
+We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results.
 
@@
 
-* Line 421: The points are created with distance \epsilon of x, not \epsilon
-||x||. r is chosen in the sphere of radius \epsilon and added to x, and, as
-stated correctly line 424, x' is within \epsilon of x.
+*Line 421: The points are created with distance $\epsilon$ of $x$, not $\epsilon
+||x||$. $r$ is chosen in the sphere of radius $\epsilon$ and added to $x$, and, as
+stated correctly line 424, $x'$ is within $\epsilon$ of $x$.*
 
-* There are many small LaTeX issues: wrong space after "i.e." and "e.g." (add a
+x is a vector, and epsilon is a scalar. We can clarify that x is a vector. Be sure to boldface all vectors
+
+*There are many small LaTeX issues: wrong space after "i.e." and "e.g." (add a
 backslash), large spaces around formulas (e.g., line 342), em-dashes that look
-too short (e.g., line 594).
+too short (e.g., line 594).*
 
-* The figures should not be shrank so much: the labels on the axis and the keys
+easy to fix (NMD)
+
+*The figures should not be shrank so much: the labels on the axis and the keys
 are not readable. Regenerate the pictures at the correct size to avoid
-scaling.
+scaling.*
 
-* Put the keys outside of the figures, not over the data lines.
+fix some font sizes and aspect ratios
 
-* Many horizontal and vertical lines in Table~2 should be removed to help with
-readability.
+*Put the keys outside of the figures, not over the data lines.*
 
-* Line 585: "we observe performs quite slowly". The sentence seems incomplete.
+need to fix this (NI)
 
-* Many of the explicit algorithms do not serve any real purpose: they are not
+*Many horizontal and vertical lines in Table~2 should be removed to help with
+readability.*
+
+investigate this, or break up table 2 into 2a-2d (NMD)
+
+*Line 585: "we observe performs quite slowly". The sentence seems incomplete.*
+
+will fix -> 'we observe that CAKES performes quite slowly'
+
+*Many of the explicit algorithms do not serve any real purpose: they are not
 more readable in pseudo-code form and are just fine in explained in the text.
-(E.g., Algorithm 2.4).
+(E.g., Algorithm 2.4).*
 
-* The Github repository gives very little instructions on how to compile the
-code, run it or reproduce the results from this manuscript.
+get rid of alg 2.4
 
-				Review of
+*The Github repository gives very little instructions on how to compile the
+code, run it or reproduce the results from this manuscript.*
 
+will fix (NI)
+
+**Review of
 Let them have CAKES: A Cutting-Edge Algorithm for Scalable,
-		Efficient, and Exact Search on Big Data
+Efficient, and Exact Search on Big Data**
 
-			SIMODS manuscript M166272
+*SIMODS manuscript M166272*
 
@@
 
-The CAKES manuscript proposes an entropy scaling approach to the search
+*The CAKES manuscript proposes an entropy scaling approach to the search
 problems of $k$ Nearest Neighbors ($k$-NN) and $rho$ Nearest Neighbors
@@
 issues both technically and expositionally, as well as some minor issues,
-that need to be addressed before another round of reviewing.
+that need to be addressed before another round of reviewing.*
 
-While there is implicitly a tree that is the key data structure, the tree
+*While there is implicitly a tree that is the key data structure, the tree
 itself is never named or given a notation; the same is true of the root of
@@
 terminology and notation for a data structure and apply it consistently to
-their data structure.
+their data structure.*
 
-The authors appear to use the terms similarity and distance
+we fixed this in pancakes, just backport it
+
+*The authors appear to use the terms similarity and distance
 interchangeably, though they are certainly different.  In particular, a
@@
 must identify the actual spaces from which data sets can be extracted for
-CAKES to work.
+CAKES to work.*
 
-On line 141, there is a function defined for the local fractal dimension
+add def in paper for geometric median, extended from continues vector spaces into any metric space. line 43, use distance, not similarity
+
+*On line 141, there is a function defined for the local fractal dimension
 (LFD), but it is not expressed in functional notation.  In other words, it
@@
 $B(X,q,r)$, not $B_X(q,r)$.  If makes little sense to put $X$ as a
-subscript in this context.
+subscript in this context.*
 
-In the algorithms, starting with Algorithm 2.1, the pseudocode style is
+similar to complaint from reviewer 2, address with same language
+
+*In the algorithms, starting with Algorithm 2.1, the pseudocode style is
 ugly.  Much better would be the pseudocode style from the fourth edition
@@
 is as a "stopping criteria", while the logic of the code has it as a
-continuing criteria.  Please fix the pseudocode throughout.
+continuing criteria.  Please fix the pseudocode throughout.*
 
-Section 2.2 is meant to be the definition of the search problem(s), but
+make this pseudocode style change (NMD)
+call it continuation criteria (NI)
+
+*Section 2.2 is meant to be the definition of the search problem(s), but
 that is really only contained in the paragraph from lines 218 to 223. In
@@
 integer $k$ or the number $\rho$ is a parameter to the corresponding
-problem and should be mentioned as such.
+problem and should be mentioned as such.*
 
-There are no theorems in the manuscript, yet there are arguments for
+move introduction of the three deltas to later, to start of 2.4.3, and expand definition of rnn. (NI)
+
+*There are no theorems in the manuscript, yet there are arguments for
 quantities like the quality of the results or the time complexity that are
@@
 propositions for each thing to be proved, with appropriate assumptions,
-and to give precise proofs of same.
+and to give precise proofs of same.*
 
-The manuscript makes too much use of unfamiliar terminology taken from
+state theorem, which is the time complexity, and formalize the argument into a proof
+
+*The manuscript makes too much use of unfamiliar terminology taken from
 earlier papers without defining terms.  For example, see line 233 and
@@
 used must be precisely defined in the manuscript to allow proper
-assessment of the manuscript.
+assessment of the manuscript.*
 
-The Results section is a slog to read and parts of the Discussion seem
+remind the reader what rnn search is, and in 306-310 reword to refer to previous section rather than external refs
+
+*The Results section is a slog to read and parts of the Discussion seem
 redundant.  Take lines 651 to 668 as an example.  Tightening up the
-Results and the Discussion can only help the exposition.
+Results and the Discussion can only help the exposition.*
 
-I led a bioinformatics project that used an entropy scaling approach to
+offload to figure captions where possible, don't spell out everything
+
+*I led a bioinformatics project that used an entropy scaling approach to
 search a database of protein sequences.  See Reference [1].  As a result,
@@
 reference [1] and discuss in more depth how they address the challenges of
-the various real world examples.
+the various real world examples.*
+
+fine, cite 1, and discuss (NI)
 
@@
 
-"Data set" is pronounced as two words and should be written as such
-throughout the manuscript.
+*"Data set" is pronounced as two words and should be written as such
+throughout the manuscript.*
 
-In line 22, use "rate" instead of "scale".
+We respectfully disagree; this appears as a term of art in existing literature, and is in the Merriam-Webster Dictionary.
+"Dataset" is listed by Merriam-Webster as the primary form, with "data set" being the less common form (https://www.merriam-webster.com/dictionary/dataset).
+Unless this is an editorial policy, we would rather keep it as "dataset."
 
-Throughout, the word "which" is used when "that" is the correct word.  If
+*In line 22, use "rate" instead of "scale".*
+
+ok, will fix
+
+*Throughout, the word "which" is used when "that" is the correct word.  If
 "which" is used, then it is not defining and is preceded by a comma.
-Change all "which"'s to "that"'s.
+Change all "which"'s to "that"'s.*
 
-On line 75, there is a Subsection 1.1, but there is no Subsection 1.2.
+We somewhat agree, but disagree that *all* instances should be changed. We have made selected changes where it seems to fit the pattern suggested by this reviewer.
+
+*On line 75, there is a Subsection 1.1, but there is no Subsection 1.2.
 This is not a good practice.  Just refactor the text without Section 1.1
-or add a Subsection 1.2.
+or add a Subsection 1.2.*
 
-The use of $\forall$, $\exists$, $\Leftrightarrow$, and other similar
+make the last paragraph (entropy scaling) into 1.2
+
+*The use of $\forall$, $\exists$, $\Leftrightarrow$, and other similar
 logical operators is not appropriate in a manuscript that is not about
-logic.  Spell out the meaning in English, and be creative.
+logic.  Spell out the meaning in English, and be creative.*
 
-Near line 154, it is important to give an overview of your solution
+We respectfully disagree.
+These operators are also standard in set theory and functional analysis, and it is in this context that we use them.
+They are also standard in the algorithms literature, including CLRS.
+Unless this is editorial policy, we would prefer to keep them.
+We believe they lend precision.
+
+*Near line 154, it is important to give an overview of your solution
 strategy and, in particular, why clustering is used first.  Clustering is
 a heavily overloaded word.  I would have chosen a more neutral term
-instead of "cluster".
+instead of "cluster".*
 
-On line 155, there is the parenthetical comment about "a smaller sample of
+We do define our clustering approach, and moreover, we build on prior work (some ours, some others) that relies on clustering. We respectfully would prefer to stay with the language we have here.
+
+*On line 155, there is the parenthetical comment about "a smaller sample of
 the points".  What does that mean precisely and where does it occur later
-in the manuscript?
+in the manuscript?*
 
-On line 156, there is the claim "so it is a real data point".  What is a
-real data point?  Please explain.
+we take a root-n sample of the points (NI), justified in CHESS paper. Make this clear.
 
-On line 161 and elsewhere, "i.e." is always preceded and followed by
-commas.
+*On line 156, there is the claim "so it is a real data point".  What is a
+real data point?  Please explain.*
 
-On line 178, there is the phrase "find the point ... in the sample".  Can
+deal with by defining geometric median
+
+*On line 161 and elsewhere, "i.e." is always preceded and followed by
+commas.*
+
+fixed here
+
+*On line 178, there is the phrase "find the point ... in the sample".  Can
 you make the entire paragraph more precise by giving notation for the
-sample?
+sample?*
 
-Section 2.1.1 overall is lacking in mathematical rigor.
+deal with by defining geometric median
 
-In Section 2.1.3, the cost of assessing "criteria' is not counted in the
-time complexity, but it must be.
+*Section 2.1.1 overall is lacking in mathematical rigor.*
 
-In line 252, there is again mention of "real points" without any
-indication of the meaning of "real".
+we will refactor 2.1.1 to be more rigorous (NI)
 
-The fonts in Algorithms 2.2 through 2.7 are simply too small.  Make them
-the font size of the remainder of the text.
+*In Section 2.1.3, the cost of assessing "criteria' is not counted in the
+time complexity, but it must be.*
 
-The numbers in Table 2 are too small.  Refactor into multiple tables
-perhaps.
+add some explanation: these can be as complex as a user wants, but for us it is a constant-time check (NI)
 
-In line 641, what is "radial increase 2.3"?
+*In line 252, there is again mention of "real points" without any
+indication of the meaning of "real".*
 
-In line 642, why is "harmonic" italicized?
+replace with 'actual points from the dataset'
 
-The References section needs a through going over for proper formatting
+*The fonts in Algorithms 2.2 through 2.7 are simply too small.  Make them
+the font size of the remainder of the text.*
+
+we will make this more legible (NMD)
+
+*The numbers in Table 2 are too small.  Refactor into multiple tables
+perhaps.*
+
+we are refactoring into multiple tables (NMD)
+
+*In line 641, what is "radial increase 2.3"?*
+
+missing word 'equation' and also put in parentheses
+
+*In line 642, why is "harmonic" italicized?*
+
+no need
+
+*The References section needs a through going over for proper formatting
 and capitalization.  For example, check the title capitalization in [2],
 [4], and [8].  Check the journal name capitalization in [9] and [21],
-among others.
+among others.*
 
-Several references have repeated URLs.  Trim these up.
+*Several references have repeated URLs.  Trim these up.*
 
-References [2] and [3] appear to be duplicates.
+*References [2] and [3] appear to be duplicates.*
 
-Reference [39] is incomplete and inadequate.
+*Reference [39] is incomplete and inadequate.*
+
+Will go over this carefully (NMD)
 
@@
 
-[1] Yoonjin Kim, Zhen Guo, Jeffrey A. Robertson, Benjamin Reidys, Ziyan
+*[1] Yoonjin Kim, Zhen Guo, Jeffrey A. Robertson, Benjamin Reidys, Ziyan
 Zhang, Lenwood S. Heath, EnTrance: Exploration of Entropy Scaling Ball
 Cover Search in Protein Sequences, bioRxiv
-https://urldefense.com/v3/__https://www.biorxiv.org/content/10.1101/2021.05.31.446458v1__;!!Jh1S!kGffVkV6p8O_T-q3TzzRIBjJpAmGsw1kVTB7zosA7khu71pJpSxxrEqcYu6KceYMTg_VZ_MJ8B-PoPqf$ , 2021.
+https://urldefense.com/v3/__https://www.biorxiv.org/content/10.1101/2021.05.31.446458v1__;!!Jh1S!kGffVkV6p8O_T-q3TzzRIBjJpAmGsw1kVTB7zosA7khu71pJpSxxrEqcYu6KceYMTg_VZ_MJ8B-PoPqf$ , 2021.*

### 2025-01-05 sections/2-methods.tex
dealt with e.g., i.e., dashes

@@
 We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
-A distance of zero defines identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
-We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
-In addition to these constraints, if the distance function obeys the triangle inequality (i.e.
-$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
+A distance of zero defines identity among points (i.e.,\,$f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
+We also require that the distance function be symmetric (i.e.,\,$f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
+In addition to these constraints, if the distance function obeys the triangle inequality (i.e.,\,$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
 Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
-For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
+For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g.,\,consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
 
@@
 Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
-For real-world datasets, we expect the LFD to be locally uniform, i.e., when $r$ is small, but potentially highly variable at larger length scales, i.e., when $r$ is large.
+For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but potentially highly variable at larger length scales, i.e.,\,when $r$ is large.
 
@@
 We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
-The center is the geometric median of the points in the cluster, i.e., it is the point which minimizes the sum of distances to all other points in the cluster.
+The center is the geometric median of the points in the cluster, i.e.,\,it is the point which minimizes the sum of distances to all other points in the cluster.
 In cases where the cardinality of the cluster is large, we take a random subsample of $\sqrt{|C|}$ points and compute the geometric median of that subsample~\cite{ishaq2019clustered}.
@@
 Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
-As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e. $C(c, r) \subset B(c, r)$.
+As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e.,\,$C(c, r) \subset B(c, r)$.
 
-Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius, i.e. using Equation~\ref{eq:methods:lfd-half}.
+Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius, i.e.,\,using Equation~\ref{eq:methods:lfd-half}.
 We also only use points that are in $C(c, r)$ instead of all those in $B(c, r)$.
@@
 The \textit{radius} of $C$ is the maximum distance from the \textit{center} to any other point in $C$.
-The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.
+The point which is responsible for that radius (i.e.,\,the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.
 We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
@@
 In fact, for real datasets, we expect anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
-The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g. in a $d$-dimensional hyper-cube.
+The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
 In this sense, the imbalance in the tree is a feature, not a bug, as it reflects the underlying structure of the data.
@@
 This complexity analysis ignores the cost of checking the continuation criteria.
-While a user could specify criteria of any complexity (our implementation allows for this), for this paper we continue clustering until each cluster is a singleton, i.e., it contains only one point or duplicates of the same point and has a radius of zero.
+While a user could specify criteria of any complexity (our implementation allows for this), for this paper we continue clustering until each cluster is a singleton, i.e.,\,it contains only one point or duplicates of the same point and has a radius of zero.
 This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
@@
 one iteration to find at least one cluster, and one more to find enough of the $k$ neighbors.
-Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e. $\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
+Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
 
@@
 where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in the region around the query (i.e. if $\hat{d}$ differs significantly from $d$).
+We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
 
@@
 This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e., the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
@@
 Let $d$ be the LFD in this region.
-Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g. we could have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
+Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g.,\,we could have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
 In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
@@
 We perform some simple auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
-We start by taking the center of every cluster at a low depth, (e.g. 10), in the cluster tree as a query.
+We start by taking the center of every cluster at a low depth, (e.g.,\,10), in the cluster tree as a query.
 This gives us a small, representative sample of the dataset.
@@
 We then add $r$ to $x$ to get a new point $x'$.
-Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
+Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e.,\,$x'$ is within a distance $\epsilon$ of $x$).
 This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.

### 2025-01-05 sections/3-datasets.tex
wip: changes by NI
dealt with e.g., i.e., dashes

@@
-\section{Datasets And Benchmarking}
+\section{Datasets And Benchmarks}
 \label{sec:datasets-and-benchmarks}
@@
 We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
-Since this dataset contains many redundant, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.
+Since this dataset contains many redundant, i.e.,\,nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.
 

### 2025-01-05 sections/4-results.tex
dealt with e.g., i.e., dashes

@@
 All percentile lines seem to decrease linearly with depth.
-The LFD of approximately 20 at depth 0 (i.e. the root cluster) is what we expect for this random dataset.
-To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.
-the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
+The LFD of approximately 20 at depth 0 (i.e.,\,the root cluster) is what we expect for this random dataset.
+To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
 As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
@@
 The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.
-The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or -- as this is still a synthetic dataset -- nonuniform generation of data.
+The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or---as this is still a synthetic dataset---nonuniform generation of data.
 Silva, in contrast, appears to obey the manifold hypothesis at all scales.
@@
 Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
-We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.
-at each cardinality), and this approach was always able to select the fastest algorithm for each dataset at each cardinality.
+We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.,\,at each cardinality), and this approach was always able to select the fastest algorithm for each dataset at each cardinality.
 Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the tuned CAKES algorithm.
@@
 For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
-Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset--rather than augmented versions of the dataset--to examine how performance scales with cardinality.
+Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
 With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
@@
     Note that for RadioML, all three algorithms performed nearly identically, so distinct lines are not visible.
-    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e. more then 512GB).
+    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
     For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}

### 2025-01-05 reviewer-response.md
dealt with e.g., i.e., dashes
table 2 broken up, but ugly black bars remain

@@
 
-easy to fix (NMD)
+We have addressed much of this, including replacing en-dashes with em-dashes where appropriate. The space around formulas seems to be something that, if accepted, the production team could address.
 
@@
 
-investigate this, or break up table 2 into 2a-2d (NMD)
+Per the suggestion of reviewer 2, we have broken this table up into four tables (one per dataset) which we helieve helps readability.
 
@@
 
-we are refactoring into multiple tables (NMD)
+We have broken this table up into four tables (one per dataset) which we helieve helps readability.
 

### 2025-01-05 sections/4-results.tex
table 2 broken up, but ugly black bars remain

@@
 HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
-in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{tab:results:qps-and-recall}.
+in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
 
@@
     % \renewcommand{\arraystretch}{1.15}
-    \caption{Queries per second (QPS) and Recall vs naive linear search on the Fashion-Mnist, Glove-25, Sift and Random datasets.
+    \caption{Queries per second (QPS) and Recall vs naive linear search on the Fashion-Mnist dataset.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:qps-and-recall}
+    \label{tab:results:qps-and-recall-fmn}
     \vskip 0.15in
     \begin{center}
-        \begin{small}
+        %\begin{small}
             \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
@@
                 & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
-                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\\cline{2-10}
-                \hline
+                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+                \cline{2-10}
                 \hline
@@
                 \hline
+            \end{tabular}
+        %\end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
+
+\begin{table*}[h]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Queries per second (QPS) and Recall vs naive linear search on the Glove-25 dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-glove}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
+                \hline
+                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
+                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+                \cline{2-10}
+
                 \hline
@@
                 \hline
+
+            \end{tabular}
+        \end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
+
+\begin{table*}[h]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Queries per second (QPS) and Recall vs naive linear search on the Sift dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-sift}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
+                \hline
+                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
+                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+                \cline{2-10}
+
                 \hline
@@
                 \hline
+
+            \end{tabular}
+        \end{small}
+    \end{center}
+    \vskip -0.1in
+\end{table*}
+
+\begin{table*}[h]
+    % \renewcommand{\arraystretch}{1.15}
+    \caption{Queries per second (QPS) and Recall vs naive linear search on the Random dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-random}
+    \vskip 0.15in
+    \begin{center}
+        \begin{small}
+            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
+                \hline
+                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
+                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+                \cline{2-10}
+
                 \hline

### 2025-01-05 sections/1-introduction.tex
wip: changes by NI
reference fixes

@@
 As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become challenging;
-even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{Malkov2016EfficientAR, johnson2019billion, annoy, aumuller2020ann}.
+even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{malkov2016hnsw, johnson2019billion, annoy, aumuller2020ann}.
 
@@
 
+\subsection{Entropy Scaling Search}
+\label{sec:intoduction:entropy-scaling-search}
+
 The entropy-scaling search paradigm exploits the geometric and topological structure inherent in large datasets.

### 2025-01-05 sections/4-results.tex
reference fixes
fashion-mnist table is now bigger

@@
     \begin{center}
-        %\begin{small}
-            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
+            \begin{tabular}{|l|p{1.55cm}|p{1.0cm}|p{1.6cm}|p{1.0cm}|p{1.6cm}|p{1.1cm}|p{1.6cm}|p{1.0cm}|}
                 \hline
-                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
-                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-                \cline{2-10}
+                \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+                & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+                \cline{1-9}
                 \hline
-                \multirow{10}{*}{\rotatebox[origin=c]{90}{\textbf{Fashion-Mnist}}}
-                & 1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{2-10}
-                & 2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{2-10}
-                & 4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{2-10}
-                & 8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{2-10}
-                & 16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{2-10}
-                & 32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{2-10}
-                & 64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{2-10}
-                & 128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{2-10}
-                & 256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{2-10}
-                & 512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
+                1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
+                2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
+                4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
+                8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
+                16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
+                32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
+                64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
+                128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
+                256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
+                512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
                 \hline
             \end{tabular}
-        %\end{small}
     \end{center}

### 2025-01-05 sections/4-results.tex
made all tables larger

@@
 
-\begin{table*}[h]
-    % \renewcommand{\arraystretch}{1.15}
+\begin{table*}
     \caption{Queries per second (QPS) and Recall vs naive linear search on the Fashion-Mnist dataset.
@@
     \label{tab:results:qps-and-recall-fmn}
-    \vskip 0.15in
     \begin{center}
-            \begin{tabular}{|l|p{1.55cm}|p{1.0cm}|p{1.6cm}|p{1.0cm}|p{1.6cm}|p{1.1cm}|p{1.6cm}|p{1.0cm}|}
+        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
             \hline
@@
     \end{center}
-    \vskip -0.1in
+    \vskip -0.2in
 \end{table*}
 
-\begin{table*}[h]
-    % \renewcommand{\arraystretch}{1.15}
+\begin{table*}
     \caption{Queries per second (QPS) and Recall vs naive linear search on the Glove-25 dataset.
@@
     \label{tab:results:qps-and-recall-glove}
-    \vskip 0.15in
     \begin{center}
-        \begin{small}
-            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
+        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
             \hline
-                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
-                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-                \cline{2-10}
-
+            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+            \cline{1-9}
             \hline
-                \multirow{9}{*}{\rotatebox[origin=c]{90}{\textbf{Glove-25}}}
-                & 1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{2-10}
-                & 2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{2-10}
-                & 4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{2-10}
-                & 8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{2-10}
-                & 16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{2-10}
-                & 32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{2-10}
-                & 64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{2-10}
-                & 128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{2-10}
-                & 256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
+            1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
+            2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
+            4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
+            8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
+            16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
+            32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
+            64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
+            128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
+            256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
             \hline
-
         \end{tabular}
-        \end{small}
     \end{center}
@@
 
-\begin{table*}[h]
-    % \renewcommand{\arraystretch}{1.15}
+\begin{table*}
     \caption{Queries per second (QPS) and Recall vs naive linear search on the Sift dataset.
@@
     \label{tab:results:qps-and-recall-sift}
-    \vskip 0.15in
     \begin{center}
-        \begin{small}
-            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
+        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
             \hline
-                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
-                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-                \cline{2-10}
+            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+            \cline{1-9}
+            \hline
+            1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
+            2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
+            4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
+            8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
+            16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
+            32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
+            64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
+            128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
+            \hline
+        \end{tabular}
+    \end{center}
+    \vskip -0.2in
+\end{table*}
 
+\begin{table*}
+    \caption{Queries per second (QPS) and Recall vs naive linear search on the Random dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-random}
+    \begin{center}
+        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
             \hline
-                \multirow{8}{*}{\rotatebox[origin=c]{90}{\textbf{Sift}}}
-                & 1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{2-10}
-                & 2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{2-10}
-                & 4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{2-10}
-                & 8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{2-10}
-                & 16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{2-10}
-                & 32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{2-10}
-                & 64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{2-10}
-                & 128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
+            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+            \cline{1-9}
             \hline
-
+            1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{5.54e2} & 1.000 \\\cline{1-9}
+            2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.69e2} & 1.000 \\\cline{1-9}
+            4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.37e2} & 1.000 \\\cline{1-9}
+            8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{5.69e1} & 1.000 \\\cline{1-9}
+            16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.61e1} & 1.000 \\\cline{1-9}
+            32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.35e1} & 1.000 \\
+            \hline
         \end{tabular}
-        \end{small}
     \end{center}
@@
 \end{table*}
-
-\begin{table*}[h]
-    % \renewcommand{\arraystretch}{1.15}
-    \caption{Queries per second (QPS) and Recall vs naive linear search on the Random dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:qps-and-recall-random}
-    \vskip 0.15in
-    \begin{center}
-        \begin{small}
-            \begin{tabular}{|c|l|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|p{1.4cm}|p{1.0cm}|}
-                \hline
-                & \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{|c}{\textbf{hnsw}} & \multicolumn{2}{|c}{\textbf{annoy}} & \multicolumn{2}{|c}{\textbf{faiss-ivf}}  & \multicolumn{2}{|c|}{\textbf{cakes}} \\\cline{3-10}
-                & & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-                \cline{2-10}
-
-                \hline
-                \multirow{6}{*}{\rotatebox[origin=c]{90}{\textbf{Random}}}
-                & 1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{5.54e2} & 1.000 \\\cline{2-10}
-                & 2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.69e2} & 1.000 \\\cline{2-10}
-                & 4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.37e2} & 1.000 \\\cline{2-10}
-                & 8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{5.69e1} & 1.000 \\\cline{2-10}
-                & 16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.61e1} & 1.000 \\\cline{2-10}
-                & 32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.35e1} & 1.000 \\
-                \hline
-            \end{tabular}
-        \end{small}
-    \end{center}
-    \vskip -0.1in
-\end{table*}

### 2025-01-05 sections/1-introduction.tex
more ref fixes

@@
 
-Recent $k$-nearest neighbor search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{Malkov2016EfficientAR}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Recent $k$-nearest neighbor search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 However, some of these algorithms do not provide exact search (as defined in Section \ref{sec:introduction} above).
 
-Hierarchical Navigable Small World networks~\cite{Malkov2016EfficientAR} relies on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
+Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
 HNSW builds a multi-layered graph of the dataset.

### 2025-01-05 sections/3-datasets.tex
reference fixes
more ref fixes

@@
 
-We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}.
+We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
 Table~\ref{tab:datasets:summary} summarizes these datasets.
 The benchmarks for Fashion-Mnist, Glove-25, Sift, and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
-This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}.
+This is the same configuration used in~\cite{aumuller2020ann}.
 The OS kernel was Ubuntu 22.04.3-Ubuntu SMP.

### 2025-01-06 reviewer-response.md
reference fixes
addressing minor comments

@@
 
-x is a vector, and epsilon is a scalar. We can clarify that x is a vector. Be sure to boldface all vectors
+Fixed and made all vectors boldfaced.
 
@@
 
-will fix -> 'we observe that CAKES performes quite slowly'
+This has been fixed.
 
@@
 
-ok, will fix
+This has been fixed.
 
@@
 
-make the last paragraph (entropy scaling) into 1.2
+We added Subsection 1.2 (Entropy-Scaling Search).
 
@@
 
-we take a root-n sample of the points (NI), justified in CHESS paper. Make this clear.
+We clarified this by changing the language about "a smaller sample of the points" to "a random subsample of $\sqrt{|C|}$ points."
 
@@
 
-deal with by defining geometric median
+We removed the language "real data point" and clarified that the center of a cluster is the point which is its geometric median.
 
@@
 
-fixed here
+Fixed-- "i.e." is now always followed by a comma.
 
@@
 
-deal with by defining geometric median
+We now refer to the sample as $S$.
 
@@
 
-replace with 'actual points from the dataset'
+We removed the section which contained this from the paper.
 
@@
 
-We have broken this table up into four tables (one per dataset) which we helieve helps readability.
+We have broken this table up into four tables (one per dataset) which we believe helps readability.
 
@@
 
-missing word 'equation' and also put in parentheses
+We added that 2.3 is an equation.
 
@@
 
-no need
+This has been fixed.
 
@@
 
+We have capitalized titles as published (confirmed for all of the references mentioned). We have corrected the journal capitalization where appropriate.
+
 *Several references have repeated URLs.  Trim these up.*
+
+We believe these are now fixed.
 
@@
 
+This has been fixed.
+
 *Reference [39] is incomplete and inadequate.*
 
-Will go over this carefully (NMD)
+This has been fixed.
 

### 2025-01-06 sections/1-introduction.tex
addressing minor comments

@@
 
-Researchers are collecting data at an unprecedented scale.
+Researchers are collecting data at an unprecedented rate.
 In many fields, datasets grow exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
@@
 $k$-NN search aims to find the $k$ most similar points to a query, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query.
-Previous works have used the term \textit{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \textit{approximate} for algorithms which do not exhibit perfect recall.
+Previous works have used the term \textit{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \textit{approximate} for algorithms that do not exhibit perfect recall.
 In contrast, an \textit{exact} search algorithm exhibits perfect recall.
@@
 This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
+Further, there is evidence that distance functions that do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
 this suggests that approximate $k$-NN search could perform poorly in such contexts.
@@
 
-This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy Scaling Search), a set of three novel algorithms for exact $k$-NN search.
+This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
 We also present improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
@@
 Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
-To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$, which specifies the number of additional adjacent or nearby cells to search.
+To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$ that specifies the number of additional adjacent or nearby cells to search.
 
@@
 
-\subsection{Entropy Scaling Search}
+\subsection{Entropy-Scaling Search}
 \label{sec:intoduction:entropy-scaling-search}

### 2025-01-06 sections/2-methods.tex
addressing minor comments

@@
 We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomic sequences with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
+To illustrate the difference, consider the SILVA 18S rRNA dataset that contains genomic sequences with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
 Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
@@
 We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
-The center is the geometric median of the points in the cluster, i.e.,\,it is the point which minimizes the sum of distances to all other points in the cluster.
+The center is the geometric median of the points in the cluster, i.e.,\,it is the point that minimizes the sum of distances to all other points in the cluster.
 In cases where the cardinality of the cluster is large, we take a random subsample of $\sqrt{|C|}$ points and compute the geometric median of that subsample~\cite{ishaq2019clustered}.
@@
 
-For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
-Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
-This geometric median is the center of $C$.
+For a cluster $C$ with $|C|$ points, we begin by taking a random subsample $S$ of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in $S$.
+Using these distances, we compute the \textit{geometric median} of $S$; in other words, we find the point that minimizes the sum of distances to all other points in $S$.
+We define the \textit{center} of $C$ to be this geometric median.
 
-The \textit{radius} of $C$ is the maximum distance from the \textit{center} to any other point in $C$.
-The point which is responsible for that radius (i.e.,\,the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.
-We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
-Without loss of generality, we assign to the left child those points which are equidistant from the two poles.
+The \textit{radius} of $C$ is the maximum distance from the center to any other point in $C$.
+The point that is responsible for that radius (i.e.,\,the furthest point from the center) is designated the \textit{left pole} and the point that is furthest from the left pole is designated the \textit{right pole}.
+We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster that are closer to the left pole than to the right pole, and the right child contains all points in the cluster that are closer to the right pole than to the left pole.
+Without loss of generality, we assign to the left child those points that are equidistant from the two poles.
 Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc.
@@
 For real-world datasets with low LFD, we expect the tree to be highly unbalanced, reflecting the topological structure of the data.
-When building trees for such datasets, their depths would be larger than $\log n$, which would seem to increase the cost of building
+When building trees for such datasets, their depths would be larger than $\log n$, which would seem to increase the cost of tree-building.
 However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of the tree would also have fewer points, and so the cost of building each subsequent level would be lower.
@@
 % We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
-% when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children which might contain points in the query ball.
+% when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children that might contain points in the query ball.
 
@@
 
-% To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters which overlap with the query ball or any clusters which lie entirely within the query ball.
-% Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points which are no more than a distance $\rho$ from the query.
+% To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters that overlap with the query ball or any clusters which lie entirely within the query ball.
+% Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points that are no more than a distance $\rho$ from the query.
 % The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
@@
 
-In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure which stores the closest points to the query found so far, and $Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the $k$ nearest neighbors.
+In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
 The two Sieve algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
@@
 We find it useful to start by discussing the \textit{tree-search} and \textit{leaf-search} stages separately.
-Tree-search refers to the process of identifying clusters which have overlap with the query ball, or in other words, clusters which might contain one of the $k$ nearest neighbors.
+Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors.
 Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in the clusters identified by tree-search.
@@
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
-Then, we see that $Q$ is expected to be the set of clusters which overlap with a ball of radius $\rho_k$ around the query.
+Then, we see that $Q$ is expected to be the set of clusters that overlap with a ball of radius $\rho_k$ around the query.
 We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $Q$.
@@
 Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level, and $m$ be a user-specified integer multiplier.
-For each datum $x \in X$, we create $m - 1$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.
-We construct a random vector $r$ of $d$ dimensions in the hyper-sphere of radius $\epsilon$ centered at the origin.
-We then add $r$ to $x$ to get a new point $x'$.
-Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e.,\,$x'$ is within a distance $\epsilon$ of $x$).
+For each datum $\mathbf{x} \in X$, we create $m - 1$ new data points within a distance $\epsilon$ of $\mathbf{x}$.
+We construct a random vector $\mathbf{r}$ of $d$ dimensions in the hyper-sphere of radius $\epsilon$ centered at the origin.
+We then add $\mathbf{r}$ to $\mathbf{x}$ to get a new point $\mathbf{x}'$.
+Since $||\mathbf{r}|| \leq \epsilon$, we have that $||\mathbf{x} - \mathbf{x}'|| \leq \epsilon$ (i.e.,\,$\mathbf{x}'$ is within a distance $\epsilon$ of $\mathbf{x}$).
 This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.

### 2025-01-06 reviewer-response.md
evaluated and removed CLRS algorithm blocks

--- 2025-01-05 reviewer-response.md
@@
 
-make this pseudocode style change (NMD)
+We evaluated the CLRS package, but did not see any aesthetic differences. We have, however, increased the font size for the algorithm environments, so that it should be more legible.
 call it continuation criteria (NI)
@@
 
-we will make this more legible (NMD)
+We have made this larger to improve readability.
 

### 2025-01-06 sections/4-results.tex

@@
 For HNSW and ANNOY, we report the recall for each measurement in the plots.
-For algorithms which exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.
+For algorithms that exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.
 
@@
 For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
-On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset.
+On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
 With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
@@
 
-In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly.
+In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly pn the Random dataset.
 As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.

### 2025-01-06 sections/5-discussion.tex

@@
 For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD,
-the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
-This suggests that rather than using the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
+the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase in (Equation)~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
+This suggests that rather than using the reciprocal of the harmonic mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
 We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.

### 2025-01-06 sections/1-introduction.tex
moved some floats for better whitespace usage

@@
 
-$\rho$-NN search also has a variety of applications.
-For example, one could search for all genomes within a maximum edit distance of a query genome to find evolutionarily-related organisms~\cite{budowski2010fragbag}.
-One could also search for all words within a maximum edit distance of a misspelled word to suggest corrections~\cite{ukkonen1985algorithms}.
-Given an advertisement and a database of user profiles, one could search for all users whose profiles are ``similar enough'' to target the advertisement~\cite{zhang2020privacy}.
-GPS and other location-based services use $\rho$-NN search to find nearby points of interest~\cite{zhang2020privacy} to a user's location.
-%In all of these cases, approximate $\rho$-NN search may be sufficient, but exact $\rho$-NN search is preferable.
-
 This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
-We also present improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
-We provide a comparison to several current algorithms for similarity search; namely, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We also compare CAKES against several current algorithms; namely, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
 We further benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.

### 2025-01-06 sections/2-methods.tex
starting to address major comments
moved some floats for better whitespace usage

@@
 For example, with neural-network embeddings, one could use Euclidean (L2) or cosine distance.
-With genomic or proteomic sequence data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
+With genomic or proteomic sequence data, Levenshtein or Hamming distance is useful.
 With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.
@@
 We define LFD at some length scale around a point in the dataset as:
-
 \begin{equation}
@@
 \end{equation}
-
 where $B(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
 
-We use a simplified version of this definition by calculating the LFD at a length scale where $r_1 = 2 \cdot r_2$.
-
+We use a simplified version of Equation~\ref{eq:methods:lfd-original} by using a length scale where $r_1 = 2 \cdot r_2$.
 \begin{equation}
@@
     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
-    \begin{algorithmic}[0] % enter the algorithmic environment
+    \begin{algorithmic} % enter the algorithmic environment
         \Require $C$, a cluster
@@
 
-% TODO: Move most of this to the supplement
-% \subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
-% \label{sec:methods:rnn-search}
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
+    \caption{
+        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
+        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
+        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+    }
+    \label{fig:methods:deltas}
+\end{figure}
 
-% We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
-% when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children that might contain points in the query ball.
-
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[scale=0.75]{images/geometry/overlapping-children-3.pdf}
-%     \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.
-Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query.
-Other points and distances are described in the text.}
-%     \label{fig:methods:overlapping-children}
-%     \caption{CAKES uses geometric properties of clusters.}
-% \end{figure}
-
-% To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children}.
-% Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.
-
-% Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{sec:methods:clustering:building-the-tree}).
-% Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
-% Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
-% As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $\rho < d$, then the left child cannot contain points inside the query ball.
-% In such a case we proceed to search only the right child.
-% Otherwise, we proceed with both children.
-
-% To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
-% Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
-% By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
-% Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
-% Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
-% Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.
-
-% Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$.
-Note, in particular, that this only requires distances between actual points from the dataset, and so it can be used with any distance function, even when $q'$ and $m$ are not actual points or cannot be imputed from the data.
-
-% To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters that overlap with the query ball or any clusters which lie entirely within the query ball.
-% Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points that are no more than a distance $\rho$ from the query.
-% The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.
-
-% \begin{gather}
-%     \mathcal{O}
-%     \Bigg(
-%         \underbrace{
-%             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-%         }_{\textrm{tree-search}}
-%         \ + \
-%         \underbrace{
-%             \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
-%             \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
-%         }_{\textrm{leaf-search}}
-%     \Bigg)
-%     \label{eq:methods:rnn-search-complexity}
-% \end{gather}
-% where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
+\begin{minipage}{.425\textwidth}
+    \begin{algorithm}[H]
+        \caption{tree-search($C$, $q$, $r$)}
+        \label{alg:methods:rnn-search:tree-search}
+        \begin{algorithmic}
+            \Require $C$, a cluster
+            \Require $q$, a query
+            \Require $r$, a search radius
+            \If{$\delta^+_C \leq r$}
+                \State \textbf{return} $\{C\}$
+            \Else
+                \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
+                \State \textbf{return} tree-search($L, q, r$) \\
+                \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
+            \EndIf
+        \end{algorithmic}
+    \end{algorithm}
+\end{minipage}
+\hfill
+\begin{minipage}{.45\textwidth}
+    \begin{algorithm}[H]
+        \caption{leaf-search($Q$, $q$, $r$)}
+        \label{alg:methods:rnn-search:leaf-search}
+        \begin{algorithmic}
+            \Require $Q$, a set of clusters
+            \Require $q$, a query
+            \Require $r$, a search radius
+            \State $H \Leftarrow \emptyset$
+            \For{$C \in Q$}
+                \If{$\delta^+_C \leq r$}
+                    \State $H$ $\Leftarrow$ $H \cup C$
+                \Else
+                    \For{$p \in C$}
+                        \If{$f(p, q) \leq r$}
+                            \State $H$ $\Leftarrow$ $H \cup \{p\}$
+                        \EndIf
+                    \EndFor
+                \EndIf
+            \EndFor
+            \State \textbf{return} $H$
+        \end{algorithmic}
+    \end{algorithm}
+\end{minipage}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
-    \caption{
-        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
-        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
-        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
-    }
-    \label{fig:methods:deltas}
-\end{figure}%
-
 \subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
@@
 
-This algorithm relies on the \textit{tree-search} and \textit{leaf-search} algorithms described in~\cite{ishaq2019clustered}.
-We reproduce the algorithms here for completeness.
+This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
 
-\begin{minipage}{.425\textwidth}
-    \begin{algorithm}[H]\small
-        \caption{tree-search($C$, $q$, $r$)}
-        \label{alg:methods:rnn-search:tree-search}
-        \begin{algorithmic}
-            \Require $C$, a cluster
-            \Require $q$, a query
-            \Require $r$, a search radius
-
-            \If{$C$ is a leaf \textbf{or} $\delta^+_C \leq r$}
-                \State \textbf{return} $\{C\}$
-            \Else
-                \State $[l, r]$ $\Leftarrow$ \textit{pole}s of $C$
-                \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
-                \State \textbf{return} tree-search($L, q, r$)
-
-                \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
-            \EndIf
-
-        \end{algorithmic}
-    \end{algorithm}
-\end{minipage}
-\hfill
-\begin{minipage}{.45\textwidth}
-    \begin{algorithm}[H]\small
-        \caption{leaf-search($Q$, $q$, $r$)}
-        \label{alg:methods:rnn-search:leaf-search}
-        \begin{algorithmic}
-            \Require $Q$, a set of clusters
-            \Require $q$, a query
-            \Require $r$, a search radius
-
-            \State $H \Leftarrow \emptyset$, a set of 2-tuples of points and their distances to the query
-
-            \For{$C \in Q$}
-                \If{$\delta^+_C \leq r$}
-                    \State $H$ $\Leftarrow$ $H \cup \{C\}$
-                \Else
-                    \For{$p \in C$}
-                        \If{$f(p, q) \leq r$}
-                            \State $H$ $\Leftarrow$ $H \cup \{(p, f(p, q))\}$
-                        \EndIf
-                    \EndFor
-                \EndIf
-            \EndFor
-
-            \State \textbf{return} $H$
-        \end{algorithmic}
-    \end{algorithm}
-\end{minipage}
-
-For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
+For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
 If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
@@
 In particular, we increase the radius by a factor of
-
 \begin{equation}
@@
 \end{equation}
-
 where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$.
@@
 Once $\sum_{C \in Q} |C| \geq k$, we have found at least $k$ neighbors, and so we can stop increasing the radius.
-We now perform \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors among the points in the clusters found during the previous steps.
+We now perform \textit{leaf-search} to find the $k$ nearest neighbors among the points in the clusters found during the previous steps.
 
-\begin{algorithm}\small % enter the algorithm environment
+\begin{algorithm} % enter the algorithm environment
     \caption{Repeated $\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
@@
 In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
-
 \begin{gather}
@@
         \underbrace{
-            \overbrace{k}^{\textrm{output size}}
+            \overbrace{k}^{\textrm{output size}} \cdot
             \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
@@
 where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
-
-\subsubsection{Breadth-First Sieve}
-\label{sec:methods:knn-search:bredth-first-sieve}
-
-This algorithm performs a breadth-first traversal of the tree, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
-
-We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
-During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
-
-We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
-Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot takes linear time in the size of the input list, this version of QuickSelect has $\mathcal{O}(|Q| \log |Q|)$ time complexity.
-
-We then remove from $Q$ any element for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
-Next, we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
-Finally, we replace all remaining clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.
-
-We continue this process until $Q$ no longer contains any clusters.
-We then use the QuickSelect algorithm one last time to reorder $Q$, find $\tau$, and return the $k$ nearest neighbors.
-
-This process is described in Algorithm~\ref{alg:methods:bredth-first-sieve}.
-
-\subsubsection{Depth-First Sieve}
-\label{sec:methods:knn-search:depth-first-sieve}
-
-This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
-
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
-$Q$ starts containing only the root cluster while $H$ starts empty.
-So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
-
-\begin{itemize}
-    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
-    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
-\end{itemize}
-
-This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
-This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
-
-Note that this algorithm is not truly a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
-Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
+Note that the scaling factor should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
 
@@
 
+\subsubsection{Breadth-First Sieve}
+\label{sec:methods:knn-search:bredth-first-sieve}
+
+This algorithm performs a breadth-first traversal of the tree, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+
+We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
+
+We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
+Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot takes linear time in the size of the input list, this version of QuickSelect has $\mathcal{O}(|Q| \log |Q|)$ time complexity.
+
+We then remove from $Q$ any element for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
+Next, we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
+Finally, we replace all remaining clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.
+
+We continue this process until $Q$ no longer contains any clusters.
+We then use the QuickSelect algorithm one last time to reorder $Q$, find $\tau$, and return the $k$ nearest neighbors.
+
+This process is described in Algorithm~\ref{alg:methods:bredth-first-sieve}.
+
+\subsubsection{Depth-First Sieve}
+\label{sec:methods:knn-search:depth-first-sieve}
+
+This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
+
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
+$Q$ starts containing only the root cluster while $H$ starts empty.
+So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
+
+\begin{itemize}
+    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
+    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
+    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
+\end{itemize}
+
+This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
+It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
+
+Note that this algorithm is not truly a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
+Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
+
 \subsubsection{Complexity of Sieve Methods}
@@
 The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is:
-
+Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
 \begin{equation}
-    \mathcal{O} \Big( ( \mathcal{T} + \mathcal{L} ) \log ( \mathcal{T} + \mathcal{L} ) \Big).
+    \mathcal{O} \Big( ( \mathcal{T} + \mathcal{L} ) \log ( \mathcal{T} + \mathcal{L} ) \Big)
     \label{eq:methods:breadth-first-sieve-complexity}
@@
 For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
-Thus, the complexity is:
-
+Thus, the complexity is
 \begin{equation}
-    \mathcal{O} \Big( \mathcal{T} \log \mathcal{T} + \mathcal{L} \log k \Big).
+    \mathcal{O} \Big( \mathcal{T} \log \mathcal{T} + \mathcal{L} \log k \Big)
     \label{eq:methods:depth-first-sieve-complexity}

### 2025-01-06 sections/2-methods.tex
some more cleanup

@@
 Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster of the tree built using Algorithm~\ref{alg:methods:partition}, we can now formally pose the $k$-NN search problem.
-
-The $k$-NN search problem, given a user specified integer $k$ and query $q$, aims to find the $k$ closest points to $q$ in $\textbf{X}$.
+Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
 In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
@@
 Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-
 In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
@@
 
-For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
+For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
 If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
@@
 
-Once $\sum_{C \in Q} |C| \geq k$, we have found at least $k$ neighbors, and so we can stop increasing the radius.
-We now perform \textit{leaf-search} to find the $k$ nearest neighbors among the points in the clusters found during the previous steps.
+Once $\sum_{C \in Q} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we perform \textit{leaf-search} to find the $k$ nearest neighbors among the points in the clusters found after the last tree-search.
 
@@
         \Require $k$, the number of neighbors to find
-
         \State $r \Leftarrow radius$ of the $root$ cluster
         \State $r \Leftarrow$ $\frac{r}{|root|}$
-
         \Loop
@@
         \EndLoop
-
         \Loop
@@
         \EndLoop
-
         \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
@@
 Based on the assumption that the LFD near the query does not differ significantly from that of the clusters near the query (the set $Q$), Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
-one iteration to find at least one cluster, and one more to find enough of the $k$ neighbors.
+one iteration to find at least one cluster, and one more to find enough clusters to guarantee $k$ neighbors.
 Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
@@
             \Require $k$, the number of neighbors to find
-
             \State $c \Leftarrow$ \textit{center} of $root$
             \State $Q \Leftarrow$ \{ ($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1) \}
-
             \Loop
@@
                         \State $[L, R] \Leftarrow$ children of $p$
-                        \State $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (L_c, \delta_{L}, 1) \}$
-                        \State $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (R_c, \delta_{R}, 1) \}$
+                        \State $l, r \Leftarrow$ centers of $L, R$
+                        \State $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
+                        \State $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
                     \EndIf
@@
             \EndLoop
-
-            \State \textbf{return} $Q$
+            \State QuickSelect($Q$, $k$)
+            \State \textbf{return} The first $k$ points in $Q$
         \end{algorithmic}
@@
             \EndWhile
-            \State Return $H$
+            \State \textbf{return} $H$
         \end{algorithmic}
@@
 
-This algorithm performs a breadth-first traversal of the tree, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
 
@@
 
-We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.
-Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot takes linear time in the size of the input list, this version of QuickSelect has $\mathcal{O}(|Q| \log |Q|)$ time complexity.
+We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$.
+Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot takes linear time in the size of the input list, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
 
@@
 We continue this process until $Q$ no longer contains any clusters.
-We then use the QuickSelect algorithm one last time to reorder $Q$, find $\tau$, and return the $k$ nearest neighbors.
-
-This process is described in Algorithm~\ref{alg:methods:bredth-first-sieve}.
+We then use the QuickSelect algorithm one last time to reorder $Q$ and return the $k$ nearest neighbors.
 
@@
 
-This algorithm performs a depth-first traversal of the tree and uses two priority queues to track clusters and hits.
+This algorithm (described in Algorithm~\ref{alg:methods:depth-first-sieve}) is quite similar to a depth-first traversal of the tree, using two priority queues to track clusters and hits, and to prioritize which branch of the tree to explore next.
 
@@
 
-This process is described in Algorithm~\ref{alg:methods:depth-first-sieve}.
-It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This process terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
@@
 Tree-search navigates the cluster tree and adds clusters to $Q$.
-Leaf-search exhaustively searches some of the clusters in $Q$ to find the $k$ nearest neighbors.
+Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.
 

### 2025-01-06 sections/3-datasets.tex
moved some floats for better whitespace usage
some more cleanup

@@
 Table~\ref{tab:datasets:summary} summarizes these datasets.
-The benchmarks for Fashion-Mnist, Glove-25, Sift, and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
-This is the same configuration used in~\cite{aumuller2020ann}.
-The OS kernel was Ubuntu 22.04.3-Ubuntu SMP.
-The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.
-The benchmarks for SILVA and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
-The OS kernel was Manjaro Linux 5.15.130-1-MANJARO.
-The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.
+All benchmarks were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
+The OS kernel was Manjaro Linux 5.15.164-1-MANJARO.
+The Rust compiler was Rust 1.83.0, and the Python interpreter version was 3.9.18.
 
-\begin{table}[!t]
-    % \renewcommand{\arraystretch}{1.15}
+\begin{table}
     \caption{Datasets used in benchmarks.}
     \label{tab:datasets:summary}
-    \vskip 0.15in
     \begin{center}
-        \begin{small}
         \begin{sc}
@@
                 \hline
-                    \textbf{Dataset} & \textbf{Distance}  &\textbf{Cardinality}  & \textbf{Dimensionality}  \\
+                \textbf{Dataset} & \textbf{Distance Function}  &\textbf{Cardinality}  & \textbf{Dimensionality}  \\
                 \hline
@@
         \end{sc}
-        \end{small}
     \end{center}
@@
 
-\subsection{Random Datasets and Synthetic Augmentations}
+\subsection{Random Datasets and Synthetic Augmentation}
 \label{sec:datasets-and-benchmarks:random-datasets}
@@
 In addition to benchmarks on datasets from the ANN-Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{sec:methods:synthetic-data}.
-In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Table~\ref{tab:results:qps-and-recall}) increases.
+In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Mult.'' in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random}) increases.
 We also benchmarked on purely randomly-generated datasets of various cardinalities.

### 2025-01-06 sections/4-results.tex
some more cleanup

@@
 
-Table~\ref{tab:results:qps-and-recall} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random.
+Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random respectively.
 Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
 We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.,\,at each cardinality), and this approach was always able to select the fastest algorithm for each dataset at each cardinality.
-Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the tuned CAKES algorithm.
+Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the fastest CAKES algorithm.
 When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$.
@@
 \begin{table*}
-    \caption{Queries per second (QPS) and Recall vs naive linear search on the Fashion-Mnist dataset.
+    \caption{Throughput (QPS) and Recall on the Fashion-Mnist dataset.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
@@
 \begin{table*}
-    \caption{Queries per second (QPS) and Recall vs naive linear search on the Glove-25 dataset.
+    \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
@@
 \begin{table*}
-    \caption{Queries per second (QPS) and Recall vs naive linear search on the Sift dataset.
+    \caption{Throughput (QPS) and Recall on the Sift dataset.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
@@
 \begin{table*}
-    \caption{Queries per second (QPS) and Recall vs naive linear search on the Random dataset.
+    \caption{Throughput (QPS) and Recall on the Random dataset.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}

### 2025-01-08 reviewer-response.md
starting to address major comments
clean up some response language
a little cleanup

@@
 
-*The cost on line 260 uses \hat{r}, which is a mean radius. Now that cost is in
+*The cost on line 260 uses $\hat{r}$, which is a mean radius. Now that cost is in
 expectation (in addition to the balanced tree hypothesis).*
@@
 
-We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results.
+We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results. **Najib, can you handle this one?**
 
@@
 
-Fixed and made all vectors boldfaced.
+We have addressed this, by clarifying in notation, particularly by making all vectors boldfaced.
 
@@
 problems of $k$ Nearest Neighbors ($k$-NN) and $rho$ Nearest Neighbors
-($\rho-NN) in a data set $X$.  The general approach is to construct a
+($\rho-NN$) in a data set $X$.  The general approach is to construct a
 hierarchical data structure of clusters to assist in the searching.

### 2025-01-09 sections/4-results.tex
wip: added plots for random dataset
updated table for results on random data
added more plots and figures

@@
 \begin{figure}
+    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/fashion-mnist-60000.png}\\
+        \includegraphics[width=0.95\textwidth]{images/lfd/fashion-mnist.png}\\
         \subcaption{Fashion-mnist}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/glove-25-1183514.png}\\
+        \includegraphics[width=0.95\textwidth]{images/lfd/glove-25.png}\\
         \subcaption{Glove-25}
@@
     \end{subfigure}
-    \vspace{1em}
     \\
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/sift-1000000.png}\\
+        \includegraphics[width=0.95\textwidth]{images/lfd/sift.png}\\
         \subcaption{Sift}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/random-1000000.png}\\
+        \includegraphics[width=0.95\textwidth]{images/lfd/random.png}\\
         \subcaption{A random dataset}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/radio-ml-97920.png}\\
+        \includegraphics[width=0.95\textwidth]{images/lfd/silva-SSU-Ref.png}\\
+        \subcaption{Silva 18S}
+        \label{fig:results:silva-lfd}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/lfd/silva-SSU-Ref.png}\\
         \subcaption{RadioML}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/silva-2224640.png}\\
-        \subcaption{Silva 18S}
-        \label{fig:results:silva-lfd}
-    \end{subfigure}%
-    \vspace{1em}
     \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of local fractal dimension and weighted by the cardinalities of the clusters.
@@
     \end{subfigure}%
-    \vspace{1em}
     \\
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
         \subcaption{A random dataset for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+        \subcaption{Silva for $k=10$.}
+        \label{fig:results:silva-scaling}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
         \label{fig:results:radioml-scaling}
-    \end{subfigure}%
-    \vspace{1em}
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
-        \subcaption{Silva for $k=10$.}
-        \label{fig:results:silva-scaling}
     \end{subfigure}%
@@
             \hline
-            1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{5.54e2} & 1.000 \\\cline{1-9}
-            2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.69e2} & 1.000 \\\cline{1-9}
-            4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.37e2} & 1.000 \\\cline{1-9}
-            8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{5.69e1} & 1.000 \\\cline{1-9}
-            16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.61e1} & 1.000 \\\cline{1-9}
-            32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.35e1} & 1.000 \\
+            1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
+            2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
+            4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
+            8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
+            16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
+            32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
             \hline

### 2025-01-10 sections/2-methods.tex
added notation for tree and root

@@
 As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e.,\,$C(c, r) \subset B(c, r)$.
+We denote the cluster tree by $\mathcal{T}$ and the root by $\mathcal{R}$ (see Section~\ref{sec:methods:clustering:building-the-tree}).
 
@@
 
-We start by performing a divisive hierarchical clustering on the dataset using CLAM.
+We start by performing a divisive hierarchical clustering on the dataset using CLAM to obtain a cluster tree $\mathcal{T}$.
 The procedure is almost identical to that outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}).
-After building the tree, we also perform a depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
+After building $\mathcal{T}$ with CLAM, we also perform a depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
 
-For a cluster $C$ with $|C|$ points, we begin by taking a random subsample $S$ of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in $S$.
+Given a cluster $C$ with $|C|$ points, we define its two children by the following process.
+We take a random subsample $S$ of $\sqrt{|C|}$ of $C$'s points, and compute pairwise distances between all points in $S$.
 Using these distances, we compute the \textit{geometric median} of $S$; in other words, we find the point that minimizes the sum of distances to all other points in $S$.
@@
 Without loss of generality, we assign to the left child those points that are equidistant from the two poles.
-Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc.
+Starting from a root cluster $\mathcal{R}$ containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria (e.g., minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc).
 This process is described in Algorithm \ref{alg:methods:partition}.
@@
 
-Note that Algorithm~\ref{alg:methods:partition} does not necessarily produce a balanced tree.
+Note that the tree $\mathcal{T}$ produced by Algorithm~\ref{alg:methods:partition} is not necessarily balanced.
 In fact, for real datasets, we expect anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
@@
 
-In CHESS, each cluster stored a list of indices into the dataset.
+In CHESS~\cite{ishaq2019clustered}, each cluster stored a list of indices into the dataset.
 This list was used to retrieve the clusters' points during search.
 Although this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
-With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
-Assuming a balanced tree, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.
-In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
+With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree $\mathcal{T}$.
+Assuming $\mathcal{T}$ is balanced, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.
+In this work, we introduce a new approach wherein, after building $\mathcal{T}$, we reorder the dataset so that points are stored in a depth-first order.
 Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset, based on a useful invariant.
-The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
+The root cluster $\mathcal{R}$ has an offset of zero and a cardinality equal to the number of points in the dataset.
 A left child has the same offset its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
@@
 The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}.
-By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of partitioning and $\mathcal{O}(n \log n)$ cost of building the tree.
+By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of partitioning and $\mathcal{O}(n \log n)$ cost of building $\mathcal{T}$.
 This is a significant improvement over exact partitioning and tree-building, which cost $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ respectively.
@@
 
-We also note here that this $\mathcal{O}(n \log n)$ complexity assumes that the tree is balanced.
+We also note here that this $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
 This can realistically only happen when the data are uniformly distributed in the embedding space.
 This represents the worst-case scenario for subsequent search algorithms because the LFD of the dataset will approach or equal the embedding dimension.
-For real-world datasets with low LFD, we expect the tree to be highly unbalanced, reflecting the topological structure of the data.
+For real-world datasets with low LFD, we expect $\mathcal{T}$ to be highly unbalanced, reflecting the topological structure of the data.
 When building trees for such datasets, their depths would be larger than $\log n$, which would seem to increase the cost of tree-building.
-However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of the tree would also have fewer points, and so the cost of building each subsequent level would be lower.
+However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
 Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
@@
 
-Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster of the tree built using Algorithm~\ref{alg:methods:partition}, we can now formally pose the $k$-NN search problem.
+Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster $\mathcal{R}$ of the tree $\mathcal{T}$ (constructed by Algorithm~\ref{alg:methods:partition}), we can now formally pose the $k$-NN search problem.
 Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
@@
 
-For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
+For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
 If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
@@
 \begin{algorithm} % enter the algorithm environment
-    \caption{Repeated $\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
+    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
     \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
-        \Require $root$, the root cluster
+        \Require $\mathcal{R}$, the root cluster
         \Require $q$, a query
         \Require $k$, the number of neighbors to find
-        \State $r \Leftarrow radius$ of the $root$ cluster
-        \State $r \Leftarrow$ $\frac{r}{|root|}$
+        \State $r \Leftarrow radius$ of $\mathcal{R}$
+        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
         \Loop
-            \State $Q \Leftarrow$ tree-search($root$, $q$, $r$)
+            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
             \If{$Q \neq \emptyset$}
@@
             \State $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)$
-            \State $Q \Leftarrow$ tree-search($root$, $q$, $r$)
+            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
         \EndLoop
@@
     \begin{algorithm}[H]\small
-        \caption{Breadth-First Sieve($root$, $q$, $k$)} % give the algorithm a caption
+        \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
         \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
         \begin{algorithmic} % enter the algorithmic environment
-            \Require $root$, the root cluster
+            \Require $\mathcal{R}$, the root cluster
             \Require $q$, a query
             \Require $k$, the number of neighbors to find
-            \State $c \Leftarrow$ \textit{center} of $root$
-            \State $Q \Leftarrow$ \{ ($root$, $\delta^{+}_{root}$, $|root| - 1$), ($c$, $\delta_{root}$, 1) \}
+            \State $c \Leftarrow$ \textit{center} of $\mathcal{R}$
+            \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
             \Loop
@@
     \begin{algorithm}[H]\small
-        \caption{Depth-First Sieve($root$, $q$, $k$)}
+        \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
         \label{alg:methods:depth-first-sieve}
         \begin{algorithmic}
-            \Require $root$, the root cluster
+            \Require $\mathcal{R}$, the root cluster
             \Require $q$, a query
@@
 
-            \State{$Q \Leftarrow [root]$, a min-priority queue by $\delta^{-}$}
+            \State{$Q \Leftarrow [\mathcal{R}]$, a min-priority queue by $\delta^{-}$}
             \State{$H \Leftarrow []$, a max-priority queue by $\delta$}
@@
 
-This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
 
@@
 
-This algorithm (described in Algorithm~\ref{alg:methods:depth-first-sieve}) is quite similar to a depth-first traversal of the tree, using two priority queues to track clusters and hits, and to prioritize which branch of the tree to explore next.
+This algorithm (described in Algorithm~\ref{alg:methods:depth-first-sieve}) is quite similar to a depth-first traversal of the tree $\mathcal{T}$, using two priority queues to track clusters and hits, and to prioritize which branch of $\mathcal{T}$ to explore next.
 
 Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
-$Q$ starts containing only the root cluster while $H$ starts empty.
+$Q$ starts containing only the root cluster $\mathcal{R}$ while $H$ starts empty.
 So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
@@
 
-Note that this algorithm is not truly a depth-first traversal of the tree in the classical sense, because we use $Q$ to prioritize which branch of the tree we descend into.
-Indeed, we expect this algorithm to often switch which branch of the tree is being explored at greater depth.
+Note that this algorithm is not truly a depth-first traversal of $\mathcal{T}$ in the classical sense, because we use $Q$ to prioritize which branch of $\mathcal{T}$ we descend into.
+Indeed, we expect this algorithm to often switch which branch of $\mathcal{T}$ is being explored at greater depth.
 
@@
 In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
-Thus, the asymptotic complexity is $\mathcal{T} \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
-For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $\mathcal{L} \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$.
+Thus, the asymptotic complexity is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
+For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$.
 
@@
 \begin{equation}
-    \mathcal{O} \Big( ( \mathcal{T} + \mathcal{L} ) \log ( \mathcal{T} + \mathcal{L} ) \Big)
+    \mathcal{O} \Big( (T + L ) \log (T + L ) \Big).
     \label{eq:methods:breadth-first-sieve-complexity}
@@
 \begin{equation}
-    \mathcal{O} \Big( \mathcal{T} \log \mathcal{T} + \mathcal{L} \log k \Big)
+    \mathcal{O} \Big( T \log T + L \log k \Big).
     \label{eq:methods:depth-first-sieve-complexity}
@@
 We perform some simple auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
-We start by taking the center of every cluster at a low depth, (e.g.,\,10), in the cluster tree as a query.
+We start by taking the center of every cluster at a low depth (e.g.,\,10) in $\mathcal{T}$ as a query.
 This gives us a small, representative sample of the dataset.

### 2025-01-10 sections/4-results.tex
added notation for tree and root

@@
 All percentile lines seem to decrease linearly with depth.
-The LFD of approximately 20 at depth 0 (i.e.,\,the root cluster) is what we expect for this random dataset.
+The LFD of approximately 20 at depth 0 (i.e.,\,the root cluster $\mathcal{R}$) is what we expect for this random dataset.
 To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
-As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
-A consequence of this is that, with high probability, for every point in the root cluster, its distance from $c$ is greater than $\tfrac{r}{2}$;
+As a result, $\mathcal{R}$'s radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
+A consequence of this is that, with high probability, for every point in $\mathcal{R}$, its distance from $c$ is greater than $\tfrac{r}{2}$;
 in other words, $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r)$ contains the entire dataset.
-Given our definition of LFD in Equation~\ref{eq:methods:lfd-original}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
+Given our definition of LFD in Equation~\ref{eq:methods:lfd-original}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
 

### 2025-01-10 sections/5-discussion.tex
added notation for tree and root

@@
 
-When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{\lceil d \rceil} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
-Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
+When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{T}{\lceil d \rceil} + L)$, where $T$ is the time complexity of tree-search and $L$ is the time complexity of leaf-search.
+Even though Repeated $\rho$-NN has the least time cost (compare to $\mathcal{O}(T\log{(T+L)} + L\log{(T+L)})$ for Breadth-First Sieve and $\mathcal{O}(T\log{T} + L\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
 We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.

### 2025-01-10 sections/4-results.tex
added most of the plots

@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/silva-SSU-Ref.png}\\
+        \includegraphics[width=0.95\textwidth]{images/lfd/radio-ml.png}\\
         \subcaption{RadioML}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-    Note that for RadioML, all three algorithms performed nearly identically, so distinct lines are not visible.
-    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
+    For HNSW and ANNOY, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
     For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}

### 2025-01-10 sections/4-results.tex
better plots

@@
     \end{subfigure}%
-    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of local fractal dimension and weighted by the cardinalities of the clusters.
-    The last dataset is randomly generated.}
+    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by percentiles of local fractal dimension and weighted by the cardinalities of the clusters.}
     \label{fig:results:lfd-plots}
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-    For HNSW and ANNOY, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
+    For HNSW and ANNOY, we were not able to take measurements for each cardinality because building the index required more RAM than was available (i.e.,\,more then 512GB).
     For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}

### 2025-01-11 sections/4-results.tex
added plots of distance counts

@@
 
+\begin{figure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
+        \subcaption{Repeated $\rho$-NN}
+        \label{fig:results:fashion-mnist-counts-throughput}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
+        \subcaption{Repeated $\rho$-NN}
+        \label{fig:results:glove-25-counts-counts}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
+        \subcaption{Breadth First Sieve}
+        \label{fig:results:sift-counts-throughput}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
+        \subcaption{Breadth First Sieve}
+        \label{fig:results:random-counts-counts}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
+        \subcaption{Depth First Sieve}
+        \label{fig:results:silva-counts-throughput}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
+        \subcaption{Depth First Sieve}
+        \label{fig:results:radioml-counts-counts}
+    \end{subfigure}%
+    \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-Mnist dataset.
+    Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
+    The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
+    The x-axis represents increasing cardinality of the dataset.}
+    \label{fig:results:distance-counts}
+\end{figure}
+
 \begin{table*}

### 2025-01-11 sections/4-results.tex
added legends

@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/fashion-mnist.png}\\
+        \includegraphics[width=0.9\textwidth]{images/lfd/fashion-mnist.png}\\
         \subcaption{Fashion-mnist}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/glove-25.png}\\
+        \includegraphics[width=0.9\textwidth]{images/lfd/glove-25.png}\\
         \subcaption{Glove-25}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/sift.png}\\
+        \includegraphics[width=0.9\textwidth]{images/lfd/sift.png}\\
         \subcaption{Sift}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/random.png}\\
+        \includegraphics[width=0.9\textwidth]{images/lfd/random.png}\\
         \subcaption{A random dataset}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/silva-SSU-Ref.png}\\
+        \includegraphics[width=0.9\textwidth]{images/lfd/silva-SSU-Ref.png}\\
         \subcaption{Silva 18S}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/lfd/radio-ml.png}\\
+        \includegraphics[width=0.9\textwidth]{images/lfd/radio-ml.png}\\
         \subcaption{RadioML}
@@
     \end{subfigure}%
-    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by percentiles of local fractal dimension and weighted by the cardinalities of the clusters.}
+    \\
+    \vskip 0.005in
+    \begin{subfigure}[b]{0.94\textwidth}
+        \centering
+        \includegraphics[width=0.7\textwidth]{images/lfd/legend.png}
+        \label{fig:results:lfd-legend}
+    \end{subfigure}%
+    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by percentiles of local fractal dimension and weighted by the cardinalities of the clusters.
+    Note that the y-axis is different for the random dataset.}
     \label{fig:results:lfd-plots}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
         \subcaption{Fashion-Mnist for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
         \subcaption{Glove-25 for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{Sift for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{plots/random_PermutedBall_10_throughput.png}
         \subcaption{A random dataset for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
         \subcaption{Silva for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{plots/radio-ml_Ball_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
         \label{fig:results:radioml-scaling}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.94\textwidth}
+        \centering
+        \includegraphics[width=0.7\textwidth]{plots/legend.png}
+        \label{fig:results:scaling-legend}
     \end{subfigure}%
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
         \subcaption{Repeated $\rho$-NN}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
+        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
         \subcaption{Repeated $\rho$-NN}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
         \subcaption{Breadth First Sieve}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
+        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
         \subcaption{Breadth First Sieve}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
+        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
         \subcaption{Depth First Sieve}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
+        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
         \subcaption{Depth First Sieve}
         \label{fig:results:radioml-counts-counts}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.94\textwidth}
+        \centering
+        \includegraphics[width=0.6\textwidth]{images/distance_counts/legend.png}
+        \label{fig:results:counts-legend}
     \end{subfigure}%

### 2025-01-11 supplement.tex
moved some floats for better whitespace usage
some supplement cleanup
updated plots in the supplement

--- 2024-04-28 supplement.tex
@@
 \hypersetup{
-  pdftitle={Supplementary Materials: An Example Article},
-  pdfauthor={D. Doe, P. T. Frank, and J. E. Smith}
+  pdftitle={Supplementary Materials: Let them have CAKES: A Cutting-Edge Algorithm for Scalable, Efficient, and Exact Search on Big DataAn Example Article},
+  pdfauthor={M. E. Prior, T. J. Howard III, O. McLaughlin, T. Ferguson, N. Ishaq, N. M. Daniels}
 }
@@
 \maketitle
+
+\section{Supplementary Methods}
+
+\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
+
+We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
+when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children that might contain points in the query ball.
+
+To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:supplement:overlapping-children}.
+Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.
+
+Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively.
+Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
+Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
+As a consequence of how we assign a point in the parent cluster to the left child in the Partition algorithm, if $\rho < d$, then the left child cannot contain points inside the query ball.
+In such a case we proceed to search only the right child.
+Otherwise, we proceed with both children.
+
+To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
+Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:supplement:overlapping-children}.
+By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
+Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
+Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
+Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.
+
+Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$.
+Note, in particular, that this only requires distances between actual points from the dataset, and so it can be used with any distance function, even when $q'$ and $m$ are not actual points or cannot be imputed from the data.
+
+To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, to find the leaf clusters that overlap with the query ball or any clusters which lie entirely within the query ball.
+Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, to find all points that are no more than a distance $\rho$ from the query.
+The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:supplement:rnn-search-complexity}.
+
+\begin{gather}
+    \mathcal{O}
+    \Bigg(
+        \underbrace{
+            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+        }_{\textrm{tree-search}}
+        \ + \
+        \underbrace{
+            \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
+            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
+        }_{\textrm{leaf-search}}
+    \Bigg)
+    \label{eq:supplement:rnn-search-complexity}
+\end{gather}
+where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.75]{images/geometry/overlapping-children-3.pdf}
+    \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.
+Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query.
+Other points and distances are described in the text.}
+    \label{fig:supplement:overlapping-children}
+    \caption{CAKES uses geometric properties of clusters.}
+\end{figure}
 
@@
 \begin{figure}
+    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-indexing.png}\\
+        \includegraphics[width=0.9\textwidth]{images/indexing/fashion-mnist-indexing.png}\\
         \subcaption{Fashion-mnist}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/glove-25-indexing.png}\\
+        \includegraphics[width=0.9\textwidth]{images/indexing/glove-25-indexing.png}\\
         \subcaption{Glove-25}
@@
     \end{subfigure}
-    \vspace{1em}
     \\
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift-indexing.png}\\
+        \includegraphics[width=0.9\textwidth]{images/indexing/sift-indexing.png}\\
         \subcaption{Sift}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/random-indexing.png}\\
+        \includegraphics[width=0.9\textwidth]{images/indexing/random-indexing.png}\\
         \subcaption{Random}
@@
     \\
-    \vspace{1em}
     \caption{Indexing and tuning time for each algorithm with each of the ANN benchmark datasets and the Random dataset.}
@@
 
-\begin{figure}[ht!]
+\begin{figure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=1.0\textwidth]{plots/fashion-mnist_PermutedBall_100_throughput.png}
+        \subcaption{Fashion-Mnist for $k=100$.}
+        \label{fig:supplement:fashion-mnist-scaling}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=1.0\textwidth]{plots/glove-25_PermutedBall_100_throughput.png}
+        \subcaption{Glove-25 for $k=100$.}
+        \label{fig:supplement:glove-25-scaling}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=1.0\textwidth]{plots/sift_PermutedBall_100_throughput.png}
+        \subcaption{Sift for $k=100$.}
+        \label{fig:supplement:sift-scaling}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=1.0\textwidth]{plots/random_PermutedBall_100_throughput.png}
+        \subcaption{A random dataset for $k=100$.}
+        \label{fig:supplement:random-scaling}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=1.0\textwidth]{plots/silva-SSU-Ref_PermutedBall_100_throughput.png}
+        \subcaption{Silva for $k=100$.}
+        \label{fig:supplement:silva-scaling}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=1.0\textwidth]{plots/radio-ml_Ball_100_throughput.png}
+        \subcaption{RadioML for $k=100$ at SnR = 10dB.}
+        \label{fig:supplement:radioml-scaling}
+    \end{subfigure}%
+    \\
+    \begin{subfigure}[b]{0.94\textwidth}
         \centering
-    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-100.png}
-    \caption{
-        Scaling behavior of algorithms on fashion-mnist with $k=100$. 
-    }
-    \label{fig:supplement:fashion-mnist-k-100}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/sift-knn-100.png}
-    \caption{
-        Scaling behavior of algorithms on sift with $k=100$. 
-    }
-    \label{fig:supplement:sift-k-100}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/glove-25-knn-100.png}
-    \caption{
-        Scaling behavior of algorithms on glove-25 with $k=100$. 
-    }
-    \label{fig:supplement:glove-25-k-100}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=3.4in]{plots/random-knn-100.png}
-    \caption{
-        Scaling behavior of algorithms on random with $k=100$. 
-    }
-    \label{fig:supplement:random-k-100}
+        \includegraphics[width=0.7\textwidth]{plots/legend.png}
+        \label{fig:supplement:scaling-legend}
+    \end{subfigure}%
+    \caption{Throughput (with $k$=100) across six datasets, including a randomly-generated dataset.
+    In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+    For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}
+    \label{fig:supplement:scaling-plots}
 \end{figure}
@@
 
-\begin{figure}[ht!]
+\begin{figure}
+    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/radius/fashion-mnist-60000.png}\\
+        \includegraphics[width=0.9\textwidth]{images/radius/fashion-mnist.png}\\
         \subcaption{Fashion-mnist}
-    \label{fig:results:fashion-mnist-radius}
+        \label{fig:supplement:fashion-mnist-radius}
     \end{subfigure}%
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/radius/glove-25-1183514.png}\\
+        \includegraphics[width=0.9\textwidth]{images/radius/glove-25.png}\\
         \subcaption{Glove-25}
-    \label{fig:results:glove-25-radius}
+        \label{fig:supplement:glove-25-radius}
     \end{subfigure}
-    \vspace{1em}
     \\
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/radius/sift-1000000.png}\\
+        \includegraphics[width=0.9\textwidth]{images/radius/sift.png}\\
         \subcaption{Sift}
-    \label{fig:results:sift-radius}
+        \label{fig:supplement:sift-radius}
     \end{subfigure}%
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/radius/radio-ml-97920.png}\\
+        \includegraphics[width=0.9\textwidth]{images/radius/random.png}\\
+        \subcaption{A random dataset}
+        \label{fig:supplement:random-radius}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.9\textwidth]{images/radius/silva-SSU-Ref.png}\\
+        \subcaption{Silva 18S}
+        \label{fig:supplement:silva-radius}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.47\textwidth}
+        \includegraphics[width=0.9\textwidth]{images/radius/radio-ml.png}\\
         \subcaption{RadioML}
-    \label{fig:results:radioml-radius}
+        \label{fig:supplement:radioml-radius}
     \end{subfigure}%
     \\
+    \vskip 0.005in
+    \begin{subfigure}[b]{0.94\textwidth}
+        \centering
+        \includegraphics[width=0.7\textwidth]{images/radius/legend.png}
+        \label{fig:supplement:radius-legend}
+    \end{subfigure}%
+    \caption{Radius vs. cluster depth across six datasets, grouped by percentile of radius and weighted by the cardinalities of the clusters.
+    In order to use the same y-axis for all plots, we divided the radii of all clusters by the maximum radius of any cluster in the dataset.
+    Note that for the Silva 18S and the RadioML datasets, we use a logarithmic scale for the y-axis.}
+    \label{fig:supplement:radius-plots}
+\end{figure}
+
+\begin{figure}
+    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/radius/silva-2224640.png}\\
-    \subcaption{Silva 18S}
-    \label{fig:results:silva-radius}
+        \includegraphics[width=0.9\textwidth]{images/fractal_density/fashion-mnist.png}\\
+        \subcaption{Fashion-mnist}
+        \label{fig:supplement:fashion-mnist-fractal_density}
     \end{subfigure}%
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/radius/random-1000000.png}\\
-    \subcaption{A random dataset}
-    \label{fig:results:random-radius}
+            \includegraphics[width=0.9\textwidth]{images/fractal_density/glove-25.png}\\
+            \subcaption{Glove-25}
+            \label{fig:supplement:glove-25-fractal_density}
     \end{subfigure}
-    \vspace{1em}
-    \caption{Radius vs. cluster depth across six datasets, grouped by decile of radius and weighted by the cardinalities of the clusters.
-    The last dataset is randomly generated.}
-    \label{fig:results:radius-plots}
-\end{figure}
-
-\begin{figure}[ht!]
+    \\
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/fractal_density/fashion-mnist-60000.png}\\
-    \subcaption{Fashion-mnist}
-    \label{fig:results:fashion-mnist-fractal_density}
+        \includegraphics[width=0.9\textwidth]{images/fractal_density/sift.png}\\
+        \subcaption{Sift}
+        \label{fig:supplement:sift-fractal_density}
     \end{subfigure}%
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/fractal_density/glove-25-1183514.png}\\
-    \subcaption{Glove-25}
-    \label{fig:results:glove-25-fractal_density}
+        \includegraphics[width=0.9\textwidth]{images/fractal_density/random.png}\\
+        \subcaption{A random dataset}
+        \label{fig:supplement:random-fractal_density}
     \end{subfigure}
-    \vspace{1em}
     \\
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/fractal_density/sift-1000000.png}\\
-    \subcaption{Sift}
-    \label{fig:results:sift-fractal_density}
+        \includegraphics[width=0.9\textwidth]{images/fractal_density/silva-SSU-Ref.png}\\
+        \subcaption{Silva 18S}
+        \label{fig:supplement:silva-fractal_density}
     \end{subfigure}%
     \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/fractal_density/radio-ml-97920.png}\\
+        \includegraphics[width=0.9\textwidth]{images/fractal_density/radio-ml.png}\\
         \subcaption{RadioML}
-    \label{fig:results:radioml-fractal_density}
+        \label{fig:supplement:radioml-fractal_density}
     \end{subfigure}%
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/fractal_density/silva-2224640.png}\\
-    \subcaption{Silva 18S}
-    \label{fig:results:silva-fractal_density}
+    \vskip 0.005in
+    \begin{subfigure}[b]{0.94\textwidth}
+        \centering
+        \includegraphics[width=0.7\textwidth]{images/fractal_density/legend.png}
+        \label{fig:supplement:fractal_density-legend}
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-    \includegraphics[width=0.95\textwidth]{images/fractal_density/random-1000000.png}\\
-    \subcaption{A random dataset}
-    \label{fig:results:random-fractal_density}
-    \end{subfigure}
-    \vspace{1em}
-    \caption{Fractal Density vs. cluster depth across six datasets, grouped by decile of fractal density and weighted by the cardinalities of the clusters.
-    The last dataset is randomly generated.
-    Fractal Density is defined as $\frac{cardinality}{radius^{LFD}}$}
-    \label{fig:results:fractal_density-plots}
+    \caption{Fractal Density vs. cluster depth across six datasets, grouped by percentile of fractal density and weighted by the cardinalities of the clusters.
+    Fractal Density is defined as $\frac{cardinality}{radius^{LFD}}$.
+    As with the radius plots, we normalized the cluster radii before calculating the Fractal Density.
+    We also excluded all clusters with normalized radii smaller than $10^{-6}$, as they distort the scale of the plots.}
+    \label{fig:supplement:fractal_density-plots}
 \end{figure}

### 2025-01-13 sections/3-datasets.tex
added most of the plots
added language for how we check recall

@@
 
+In order to calculate recall on the augmented datasets, we perform linear search in Rust and save the results to disk.
+We verified that linear search on the original dataset produced neighbors with perfect recall compared to those provided by the ANN-Benchmarks suite.
+We then use the results of linear search on the augmented datasets as ground truth for calculating recall of CAKES's algorithms in Rust and read them in Python for calculating recall of HNSW, ANNOY, and FAISS-IVF.
+
 \subsection{SILVA 18S}
@@
 Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
-We used a subset of this dataset, containing 97,920 samples at 10dB SNR and used the other 8,576 samples at 10dB SNR as a hold-out set of queries.
+We used a subset of this dataset, containing $97,304$ samples at 10dB SNR and used another $1,000$ samples at the same SNR as a hold-out set of queries.
 We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.

### 2025-01-13 reviewer-response.md
updated reviewer response

@@
 
-fix some font sizes and aspect ratios
+The figures have all been updated.
 
@@
 
-need to fix this (NI)
+The figures have all been updated.
 
@@
 
-Per the suggestion of reviewer 2, we have broken this table up into four tables (one per dataset) which we helieve helps readability.
+The table has been broken up and former sub-tables have been enlarged.
 
@@
 
-add def in paper for geometric median, extended from continues vector spaces into any metric space. line 43, use distance, not similarity
+We have added a definition of the geometric median in the paper, and have clarified how we interpret it for any metric space.
+We have also fixed the language around similarity and distance, and have made sure to use the correct term in each context.
 
@@
 
-similar to complaint from reviewer 2, address with same language
+We have dropped the X since we are only dealing with one dataset.
 
@@
 
-We evaluated the CLRS package, but did not see any aesthetic differences. We have, however, increased the font size for the algorithm environments, so that it should be more legible.
-call it continuation criteria (NI)
+We evaluated the CLRS package, but did not see any aesthetic differences.
+We have, however, increased the font size for the algorithm environments, so that it should be more legible.
+We have also fixed the name for the criteria.
 
@@
 
-move introduction of the three deltas to later, to start of 2.4.3, and expand definition of rnn. (NI)
+We have refactored the methods section to have a better flow, and have defined several terms in their proper place.
+We have also removed $\rho$-NN search from the main body of the paper in order to focus exclusively on $k$-NN search.
+The supplement now contains a section on the improvements to $\rho$-NN search.
 
@@
 
-remind the reader what rnn search is, and in 306-310 reword to refer to previous section rather than external refs
+We have addressed this by making the paper more self-contained.
 
@@
 
-We do define our clustering approach, and moreover, we build on prior work (some ours, some others) that relies on clustering. We respectfully would prefer to stay with the language we have here.
+We clearly define our clustering approach, and moreover, we build on prior work (some ours, some others) that relies on clustering. We respectfully would prefer to stay with the language we have here.
 
@@
 
-we will refactor 2.1.1 to be more rigorous (NI)
+We have updated some language in this section.
+This, combined with Algorithm 2.1, should provide a reasonable level of rigor.
 
@@
 
-add some explanation: these can be as complex as a user wants, but for us it is a constant-time check (NI)
+We have added a discussion of this in the corresponding complexity analysis.
 
@@
 
-We removed the section which contained this from the paper.
+We have fixed language around "real points" and have clarified that the points are actual data points.
 

### 2025-01-13 sections/2-methods.tex
some stylistic changes because I'm being anal about the layout

@@
     \label{fig:methods:deltas}
+    \vskip -0.5in
 \end{figure}
@@
 In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
-The two Sieve algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
 
@@
 \begin{equation}
-    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)
+    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \right)
     \label{eq:methods:repeated-rnn-factor}
 \end{equation}
-where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$.
+where $\mu$ is the multiplicative inverse of the harmonic mean of the LFD of the clusters in $Q$, i.e.,\,$\mu = |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$.
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
@@
         \Require $k$, the number of neighbors to find
+        \State $H \Leftarrow [\ ]$, a max-heap of size $k$ to contain neighbors prioritized by distance to the query
         \State $r \Leftarrow radius$ of $\mathcal{R}$
         \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
-        \Loop
         \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-            \If{$Q \neq \emptyset$}
-                \State \textbf{break}
+        \While{$\sum_{C \in Q} |C| < k$}
+            \If{$Q = \emptyset$}
+                \State $r \Leftarrow 2 \cdot r$
+            \ElsIf{$\sum_{C \in Q} |C| >= k$}
+                \State $\mu \Leftarrow |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
+                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
             \EndIf
-            \State $r \Leftarrow 2 \cdot r$
-        \EndLoop
-        \Loop
-            \If{$\sum_{C \in Q} |C| >= k$}
-                \State \textbf{break}
-            \EndIf
-            \State $\mu \Leftarrow |Q| \cdot \left( \sum_{C \in Q} LFD(C)^{-1} \right)^{-1}$
-            \State $r \Leftarrow r \cdot \min \left( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)$
             \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \EndLoop
+        \EndWhile
         \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
-        \State \textbf{sort} $H$ by non-decreasing distance to $q$
-        \State \textbf{return} The first $k$ points in $H$
+        \State \textbf{return} $H$ as a list
     \end{algorithmic}
@@
 Note that the scaling factor should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
+
+\subsubsection{Breadth-First Sieve}
+\label{sec:methods:knn-search:bredth-first-sieve}
+
+This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+
+We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 
@@
 
-\subsubsection{Breadth-First Sieve}
-\label{sec:methods:knn-search:bredth-first-sieve}
-
-This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
-
-We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
-During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
+\vskip 0.1in
 

### 2025-01-13 sections/4-results.tex
some stylistic changes because I'm being anal about the layout

@@
     \\
-    \vskip 0.005in
+    \vskip -0.1in
     \begin{subfigure}[b]{0.94\textwidth}
@@
     \end{subfigure}%
+    \vskip -0.1in
     \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by percentiles of local fractal dimension and weighted by the cardinalities of the clusters.
@@
     \label{fig:results:lfd-plots}
+    \vskip -0.4in
 \end{figure}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
         \subcaption{Fashion-Mnist for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
         \subcaption{Glove-25 for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{Sift for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{plots/random_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{plots/random_PermutedBall_10_throughput.png}
         \subcaption{A random dataset for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
         \subcaption{Silva for $k=10$.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{plots/radio-ml_Ball_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{plots/radio-ml_Ball_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
         \subcaption{Repeated $\rho$-NN}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
+        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
         \subcaption{Repeated $\rho$-NN}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
         \subcaption{Breadth First Sieve}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
+        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
         \subcaption{Breadth First Sieve}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
+        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
         \subcaption{Depth First Sieve}
@@
     \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=1.0\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
+        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
         \subcaption{Depth First Sieve}

### 2025-01-13 sections/4-results.tex
added subsection for counting distance computations

@@
 
+\begin{table}
+    \caption{Throughput (QPS) and Recall on the Fashion-Mnist dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-fmn}
+    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+        \hline
+        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \cline{1-9}
+        \hline
+        1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
+        2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
+        4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
+        8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
+        16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
+        32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
+        64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
+        128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
+        256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
+        512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
+        \hline
+    \end{tabular}
+    \vskip -0.2in
+\end{table}
+
+\begin{table}
+    \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-glove}
+    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+        \hline
+        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \cline{1-9}
+        \hline
+        1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
+        2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
+        4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
+        8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
+        16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
+        32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
+        64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
+        128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
+        256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
+        \hline
+    \end{tabular}
+    \vskip -0.2in
+\end{table}
+
+\begin{table}
+    \caption{Throughput (QPS) and Recall on the Sift dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-sift}
+    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+        \hline
+        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \cline{1-9}
+        \hline
+        1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
+        2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
+        4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
+        8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
+        16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
+        32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
+        64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
+        128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
+        \hline
+    \end{tabular}
+    \vskip -0.2in
+\end{table}
+
+\begin{table}
+    \caption{Throughput (QPS) and Recall on the Random dataset.
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    \label{tab:results:qps-and-recall-random}
+    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+        \hline
+        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \cline{1-9}
+        \hline
+        1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
+        2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
+        4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
+        8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
+        16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
+        32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
+        \hline
+    \end{tabular}
+    \vskip -0.2in
+\end{table}
+
+\subsection{Clustering Strategies and Number of Distance Computations}
+\label{sec:results:clustering-strategies-and-number-of-distance-computations}
+
+In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
+To help with this analysis, we also added some instrumentation to our implementation of CAKES to count the number of distance computations performed during search.
+These are shown in Figure~\ref{fig:results:distance-counts}.
+
 \begin{figure}
@@
 \end{figure}
-
-\begin{table*}
-    \caption{Throughput (QPS) and Recall on the Fashion-Mnist dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:qps-and-recall-fmn}
-    \begin{center}
-        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-            \hline
-            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-            \cline{1-9}
-            \hline
-            1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
-            2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
-            4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
-            8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
-            16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
-            32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
-            64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
-            128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
-            256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
-            512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
-            \hline
-        \end{tabular}
-    \end{center}
-    \vskip -0.2in
-\end{table*}
-
-\begin{table*}
-    \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:qps-and-recall-glove}
-    \begin{center}
-        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-            \hline
-            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-            \cline{1-9}
-            \hline
-            1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
-            2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
-            4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
-            8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
-            16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
-            32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
-            64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
-            128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
-            256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
-            \hline
-        \end{tabular}
-    \end{center}
-    \vskip -0.1in
-\end{table*}
-
-\begin{table*}
-    \caption{Throughput (QPS) and Recall on the Sift dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:qps-and-recall-sift}
-    \begin{center}
-        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-            \hline
-            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-            \cline{1-9}
-            \hline
-            1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
-            2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
-            4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
-            8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
-            16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
-            32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
-            64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
-            128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
-            \hline
-        \end{tabular}
-    \end{center}
-    \vskip -0.2in
-\end{table*}
-
-\begin{table*}
-    \caption{Throughput (QPS) and Recall on the Random dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
-    \label{tab:results:qps-and-recall-random}
-    \begin{center}
-        \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-            \hline
-            \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-            & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-            \cline{1-9}
-            \hline
-            1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
-            2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
-            4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
-            8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
-            16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
-            32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
-            \hline
-        \end{tabular}
-    \end{center}
-    \vskip -0.1in
-\end{table*}

### 2025-01-13 sections/2-methods.tex
moved repeated rnn alg
cleaned up some algorithm descriptions

@@
 
+\begin{algorithm} % enter the algorithm environment
+    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
+    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
+    \begin{algorithmic} % enter the algorithmic environment
+        \Require $\mathcal{R}$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
+        \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
+        \State $r \Leftarrow radius$ of $\mathcal{R}$
+        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
+        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
+        \While{$\sum_{C \in Q} |C| < k$}
+            \If{$Q = \emptyset$}
+                \State $r \Leftarrow 2 \cdot r$
+            \ElsIf{$\sum_{C \in Q} |C| >= k$}
+                \State $\mu \Leftarrow |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
+                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
+            \EndIf
+            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
+        \EndWhile
+        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
+        \State \textbf{return} $H$ as a list
+    \end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
@@
 Once $\sum_{C \in Q} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we perform \textit{leaf-search} to find the $k$ nearest neighbors among the points in the clusters found after the last tree-search.
-
-\begin{algorithm} % enter the algorithm environment
-    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
-    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-        \Require $\mathcal{R}$, the root cluster
-        \Require $q$, a query
-        \Require $k$, the number of neighbors to find
-        \State $H \Leftarrow [\ ]$, a max-heap of size $k$ to contain neighbors prioritized by distance to the query
-        \State $r \Leftarrow radius$ of $\mathcal{R}$
-        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
-        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \While{$\sum_{C \in Q} |C| < k$}
-            \If{$Q = \emptyset$}
-                \State $r \Leftarrow 2 \cdot r$
-            \ElsIf{$\sum_{C \in Q} |C| >= k$}
-                \State $\mu \Leftarrow |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
-                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
-            \EndIf
-            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \EndWhile
-        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
-        \State \textbf{return} $H$ as a list
-    \end{algorithmic}
-\end{algorithm}
 
@@
             \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
-            \Loop
-                \If{$k = \sum_{(\_, \_, m) \in Q} m$}
-                    \State \textbf{break}
+            \While{$\sum_{(\_, \_, m) \in Q} m \neq k$}
+                \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
+                \State $Q^{'} \Leftarrow \emptyset$
+                \For{$(C, \_, \_) \in Q$}
+                    \If{$\delta^{-}_{C} \leq \tau$}
+                        \If{$C$ is a point}
+                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
+                        \ElsIf{$C$ is a leaf}
+                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
+                        \Else
+                            \State $[L, R] \Leftarrow$ children of $C$
+                            \State $l, r \Leftarrow$ centers of $L, R$
+                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
+                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
                         \EndIf
-                \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
-                \For{$(p, \delta^{+}_{p}, m) \in Q$}
-                    \If{$\delta^{-}_{p} > \tau$}
-                        \State $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
                     \EndIf
                 \EndFor
-                \For{$(p, \delta^{+}_{p}, m) \in Q$}
-                    \State $Q \Leftarrow Q \setminus \{ (p, \delta^{+}_{p}, m) \}$
-                    \If{$p$ is a point}
-                        \State \textbf{continue}
-                    \ElsIf{$p$ is a leaf}
-                        \For{$c \in p$}
-                            \State $Q \Leftarrow Q \cup \{ (c, \delta_{c}, 1) \}$
-                        \EndFor
-                    \Else
-                        \State $[L, R] \Leftarrow$ children of $p$
-                        \State $l, r \Leftarrow$ centers of $L, R$
-                        \State $Q \Leftarrow Q \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
-                        \State $Q \Leftarrow Q \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
-                    \EndIf
-                \EndFor
-            \EndLoop
+                \State $Q \Leftarrow Q^{'}$
+            \EndWhile
             \State QuickSelect($Q$, $k$)
@@
             \Require $k$, the number of neighbors to find
-
-            \State{$Q \Leftarrow [\mathcal{R}]$, a min-priority queue by $\delta^{-}$}
-            \State{$H \Leftarrow []$, a max-priority queue by $\delta$}
-
+            \State{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
+            \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
             \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
-                \While{$\neg (Q.peek$ is a leaf)}
+                \While{$Q.peek$ is not a leaf}
                     \State{$C \Leftarrow Q.pop$, the closest cluster}
@@
                 \EndFor
-                \While{$|H| > k$}
-                    \State{$H.pop$}
-                \EndWhile
             \EndWhile
@@
 
-We then remove from $Q$ any element for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
-Next, we remove all leaf clusters from $Q$ and add their points to $Q$ instead.
-Finally, we replace all remaining clusters in $Q$ with the pairs of 3-tuples corresponding to their child clusters.
-
-We continue this process until $Q$ no longer contains any clusters.
+Next, we go over all items in $Q$, skipping over any for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
+If the item corresponds to a point, we keep it.
+If the item corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
+If the item corresponds to a non-leaf cluster, we to $Q$ the pairs of 3-tuples corresponding to its child clusters.
+We continue this process until the sum of multiplicities in $Q$ is exactly $k$.
 We then use the QuickSelect algorithm one last time to reorder $Q$ and return the $k$ nearest neighbors.
@@
 
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$ and the ability to maintain its maximum size) of points prioritized by $\delta$.
 $Q$ starts containing only the root cluster $\mathcal{R}$ while $H$ starts empty.
-So long as $H$ is not full or the top priority element in $H$ has $\delta$ greater than or equal to the top priority element in $Q$, we take the following steps:
+So long as $H$ is not full or the top element in $H$ has $\delta$ greater than or equal to the top element in $Q$, we take the following steps:
 
@@
     \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
-    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
 \end{itemize}
 
-This process terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This process terminates when $H$ is full and the top element in $H$ has $\delta$ less than the top element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.

### 2025-01-15 reviewer-response.md
fixing merge conflict
a few response updates

@@
 
-Associate Editor comments (Remarks to Author):
-I have sent the manuscript out to two referees and received those reports back, and I have also looked at your paper myself.
+**Associate Editor comments (Remarks to Author):**
+*I have sent the manuscript out to two referees and received those reports back, and I have also looked at your paper myself.
 Both referees have identified quite a few items they would like to see addressed, but overall I'm pleased to report that I think the manuscript has merit and impact, and so I am recommending a major revision.
-In addition to rephrasing, reorganizing and clarifying the paper (per the referees' comments), the two major items to address are (1) details about the real world examples (cf comments from referee 1), and (2) referee 2 raises an issue about different ground truths.
+In addition to rephrasing, reorganizing and clarifying the paper (per the referees' comments), the two major items to address are (1) details about the real world examples (cf comments from referee 1), and (2) referee 2 raises an issue about different ground truths.*
 
@@
 
-this is easy to address, and we will do so
+TODO We agree that because the time complexity is highly depedent on the manifold structure of the data, this was insufficiently rigorous.
+We have attempted to address this in a few ways.
+MORGAN TODO: to be addressed Wednesday.
+Changing to a theorem-style structure ...
+Also possible TODO: axe the 2.1.1 (?) description to avoid redundancy but include that we do NOT expect a balanced tree, and that's a good thing.
+Move the last section of 2.1.1 to the section on complexity analysis
 
@@
 
-we can address this with the plots that include a balanced tree -- expand 2.1.3, balanced vs unbalanced, dropping off points faster with an unbalanced tree?
+We have attempted to address this by including additional computational experiments: in section 4.?, we have included plots indicating performance where we forced a balanced tree (in general, this performs worse, as we expected).
+TODO Najib go over this paragraph.
 
@@
 
-write in plain english right after eq 2.3, explaining in plain english the rationale for the min(2,x) expansion: namely, we don't think the very low LFD would be representative of what we'd see if we massively increased the radius
+write in plain english right after eq 2.3, explaining in plain english the rationale for the min(2,x) expansion: namely, we don't think the very low LFD would be representative of what we'd see if we massively increased the radius (TODO Morgan)
 
@@
 
-We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results. **Najib, can you handle this one?**
+We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results. **Najib, can you handle this one?** (NI)
 
@@
 
-We have addressed this, by clarifying in notation, particularly by making all vectors boldfaced.
+We have addressed this, by clarifying the notation, particularly by making all vectors boldfaced.
 

### 2025-01-15 sections/2-methods.tex

@@
 
-The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data or, in other words, that it arises from the same generative process as the rest of the dataset.
-Given the uses of $k$-NN search in practice, this assumption is reasonable.
-From this assumption, we can infer that the LFD near the query does not differ significantly from the (harmonic) mean of the LFDs of clusters near the query at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
+% The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data or, in other words, that it arises from the same generative process as the rest of the dataset.
+% Given the uses of $k$-NN search in practice, this assumption is reasonable.
+% From this assumption, we can infer that the LFD near the query does not differ significantly from the (harmonic) mean of the LFDs of clusters near the query at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
 
-We find it useful to start by discussing the \textit{tree-search} and \textit{leaf-search} stages separately.
-Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors.
-Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in the clusters identified by tree-search.
+\begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
+Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
+        \mathcal{O}
+        \Bigg(
+            \underbrace{
+                \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+            }_{\textrm{tree-search}}
+            \ + \
+            \underbrace{
+                \overbrace{k}^{\textrm{output size}} \cdot
+                \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
+            }_{\textrm{leaf-search}}
+        \Bigg)
+        \label{eq:methods:repeated-rnn-complexity1}
+    \end{gather}
+    where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors. 
+\end{theorem}
 
+% \textbf{Theorem.} \textit{Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
+Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is} \begin{gather}
+%     \mathcal{O}
+%     \Bigg(
+%         \underbrace{
+%             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+%         }_{\textrm{tree-search}}
+%         \ + \
+%         \underbrace{
+%             \overbrace{k}^{\textrm{output size}} \cdot
+%             \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
+%         }_{\textrm{leaf-search}}
+%     \Bigg).
+%     \label{eq:methods:repeated-rnn-complexity1}
+% \end{gather}
+
+\begin{proof} We consider the \textit{tree-search} and \textit{leaf-search} stages of search separately.
+Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors by doing repeated iterations of the  $\rho$-NN search Algorithm~\ref{alg:methods:rnn-search:tree-search}.
 In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
@@
 
-To build on this and determine the complexity of Repeated $\rho$-NN, we must first estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
-
-Based on the assumption that the LFD near the query does not differ significantly from that of the clusters near the query (the set $Q$), Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
+To extend Equation~\ref{eq:methods:rnn-search-complexity} to Repeated $\rho$-NN, we must first estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
+Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the (harmonic) mean of the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
+Given that LFD near $q$ does not differ significantly from that of nearby clusters, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
 one iteration to find at least one cluster, and one more to find enough clusters to guarantee $k$ neighbors.
@@
 
-To determine the asymptotic complexity of leaf-search, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree-search.
-Since we must examine every point in each such cluster, time complexity of leaf-search is linear in $\sum_{C \in Q} |C|$.
+We proceed to determine the complexity of leaf-search.
+Let $Q$ be the set of clusters returned by tree-search.
+We must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree-search; since we must examine every point in each such cluster, time complexity of leaf-search is linear in this quantity.
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
@@
 Thus, $\sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d$.
-However, an estimate for $\rho_k$ is still needed.
-For this, we once again rely on the assumption that the query is drawn from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different from the LFD of nearby clusters.
+It remains to determine an estimate for $\rho_k$.
+% For this, we once again rely on the assumption that the query is drawn from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different from the LFD of nearby clusters.
 We let $\hat{d}$ be the (harmonic) mean LFD of the clusters in $Q$.
 While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
-We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could move from the query to the center of one of the nearby clusters without significantly changing our estimate of the LFD.
+Since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at the center of a cluster in $Q$.
 By Equation~\ref{eq:methods:lfd-original}, $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$.
@@
 
-In addition, by our assumption that the LFD at the query is not significantly different from the LFD of nearby clusters, we have that $\hat{d} \approx d$.
-By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has an asymptotic complexity of
-\begin{gather}
+Again using the assumption that since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
+By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has time complexity
+\begin{gather*}
     \mathcal{O}
@@
     \label{eq:methods:repeated-rnn-complexity}
-\end{gather}
-where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-Note that the scaling factor should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
+\end{gather*}
+as claimed. 
+\end{proof}
+
+We remark that the scaling factor in Equation~\ref{eq:methods:repeated-rnn-complexity1} should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
 

### 2025-01-15 sections/4-results.tex

@@
 Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
-The horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
-We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
-In order to have the plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality.
-In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
+In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD, as it seems most relevant to CAKES's performance (see Figure~\ref{fig:results:lfd-plots}).
+% The horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
+% We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
+In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD, as it seems most relevant to CAKES's performance.
+% In order to have the plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality.
+% In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
+
 Figure~\ref{fig:results:fashion-mnist-lfd} shows the LFD by depth for Fashion-Mnist.
-We observe that until about depth 5, LFD is low, as the 95$^{th}$ percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
-For depths 15 through 25, we observe that LFD increases, with the 95$^{th}$ percentile (orange plotted line) slightly less than 6, and the median near 3.
-Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95$^{th}$ percentile is between 3 and 4, and the median is less than 2.
+% We observe that until about depth 5, LFD is low, as the 95$^{th}$ percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
+% For depths 15 through 25, we observe that LFD increases, with the 95$^{th}$ percentile (orange plotted line) slightly less than 6, and the median near 3.
+% Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95$^{th}$ percentile is between 3 and 4, and the median is less than 2.
+We observe that until approximately depth 5, Fashion-Mnist's LFD is low (i.e., less than 4).
+It then starts increasing, reaching a peak of about 6 near depth 25, before decreasing again until the maximum depth.
+
+% Relative to Fashion-Mnist, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
+% Percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
+% The 95$^{th}$ percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths.
+% Before depth 25, the 95$^{th}$ percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
+% The median hovers near 1.5 for all depths.
 
@@
 Percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-The 95$^{th}$ percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths.
-Before depth 25, the 95$^{th}$ percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
-The median hovers near 1.5 for all depths.
+In particular, Glove-25's LFD is less than 3 for all depths. 
+Before depth 25, the LFD hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
 
-Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift.
-Until about depth 5, the 95$^{th}$ percentile is between 2 and 6.
-From about depth 5 through about depth 20, the 95$^{th}$ percentile is greater than 6, even exceeding 8 near depth 10.
-From about depth 20 onward, the 95$^{th}$ percentile is less than 6, and from about depth 30 onward, it is between 3 and 4.
-The median reaches its peak of about 5 at around depth 15, but hovers near 2 after about depth 30.
+% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-Mnist and Glove-25.
+% Until about depth 5, the LFD is between 2 and 6.
+% From about depth 5 through about depth 20, the 95$^{th}$ percentile is greater than 6, even exceeding 8 near depth 10.
+% From about depth 20 onward, the 95$^{th}$ percentile is less than 6, and from about depth 30 onward, it is between 3 and 4.
+% The median reaches its peak of about 5 at around depth 15, but hovers near 2 after about depth 30.
+
+Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-Mnist and Glove-25.
+Until about depth 5, the LFD is between 2 and 6. LFD is greater than 6 until about depth 20, even exceeding 8 near depth 10.
+From about depth 20 onward, LFD falls below 6, and from about depth 30 onward, it hovers between 3 and 4.
 
@@
 Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-The 95$^{th}$ percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
-The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 128.
+In particular, LFD is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
+% The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 128.
+
+% Relative to the other  datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
+% Notably, however, for the first approximately 15 depths, the 95$^{th}$ percentile and median LFD are very close to 0.
+% Both percentiles increase sharply after depth 20, nearly reaching 16.
+% LFD remains high even as depth increases, with the 95$^{th}$ percentile fluctuating between about 6 and 15 for depths about 20-40.
+% It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60.
+% The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.
+The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or---as this is still a synthetic dataset---nonuniform generation of data.
+Silva, in contrast, appears to obey the manifold hypothesis at all scales.
 
 Relative to the other  datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
-Notably, however, for the first approximately 15 depths, the 95$^{th}$ percentile and median LFD are very close to 0.
-Both percentiles increase sharply after depth 20, nearly reaching 16.
-LFD remains high even as depth increases, with the 95$^{th}$ percentile fluctuating between about 6 and 15 for depths about 20-40.
+Notably, however, for the first approximately 15 depths, LFD is very close to 0, although it increases sharply after depth 20, nearly reaching 16.
+LFD remains high even as depth increases, fluctuating between about 6 and 15 for depths about 20-40.
 It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60.
-The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.
 The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or---as this is still a synthetic dataset---nonuniform generation of data.
@@
     \vskip -0.1in
-    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by percentiles of local fractal dimension and weighted by the cardinalities of the clusters.
-    Note that the y-axis is different for the random dataset.}
+    \caption{Local fractal dimension vs.
+cluster depth across six datasets.
+The last dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset.
+In each plot, the horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
+We show lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
+So that plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality.
+For example, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.}
     \label{fig:results:lfd-plots}
@@
 Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
-Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset.
-For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
+Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset (see Figure~\ref{fig:results:scaling-plots}). 
+% For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
+
 Which one of the three CAKES algorithms is fastest also differs by dataset.
@@
 We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
-For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
+% For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
+
 On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
 While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
+% At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
 In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
@@
 As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
-in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
+% HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
+% in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
+HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets,as shown in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
 
@@
 Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
-With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
-Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
-For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance.
-In particular, throughput of CAKES's algorithms is identical within three significant figures.
+% With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
+% Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
+
+With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it levels off at higher cardinalities.
+Depth-First Sieve is the fastest CAKES algorithm for lower multipliers, but for larger cardinalities, Repeated $\rho$-NN becomes the fastest.
+For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance; in particular, throughput of CAKES's algorithms is identical within three significant figures.
 
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-    For HNSW and ANNOY, we were not able to take measurements for each cardinality because building the index required more RAM than was available (i.e.,\,more then 512GB).
-    For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}
+    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
+    For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.
+    We observe that the cardinality at which CAKES algorithms outperform linear search differs by dataset.
+For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
+    % For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
+    }
     \label{fig:results:scaling-plots}

### 2025-01-15 sections/5-discussion.tex

@@
 We have presented CAKES, a set of three algorithms for fast $k$-NN search that are generic over a variety of distance functions.
-CAKES's algorithms are exact when the distance function in use is a metric (see definition in~\ref{sec:methods}).
+CAKES's algorithms are exact when the distance function in use is a metric (see definition in Section~\ref{sec:methods}).
 Even under cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall.
@@
 
+CAKES's performance on Sift against that on the Random dataset illustrates this phenomenon well.
 As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though both datasets have the same cardinality and dimensionality.
@@
 Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies highlight how manifold structure affects algorithm performance.
+
 We emphasize that although CAKES's algorithms are slower on the Random dataset, their recalls remain perfect.
@@
 This suggests that rather than using the reciprocal of the harmonic mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
-We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
+We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly overestimates the correct radius $\rho_k$ and to improve upon the factor of radial increase in Equation~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
 
@@
 These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
-Exploring this sensitivity is a potential avenue for future study, as it remains to be seen at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
+Exploring this sensitivity is a potential avenue for future study, as it remains to learn at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
 
@@
 We also stress that CAKES's algorithms are designed to work well for \textit{big} data, and our results support this claim;
-we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmarks' datasets we tested.
+we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmark datasets we tested.
 Moreover, we observe that CAKES's algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD.
@@
 The questions raised by this study suggest several additional avenues for future work.
-A comparison across more datasets is in order, as is further analysis with other distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support, such as Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions), and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
+A comparison across more datasets is in order, as is further analysis with other distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support.
+These include Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions) and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
 Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.

### 2025-01-15 sections/2-methods.tex
complexity rewrite for sieve

@@
 
-We start with the assumption that the LFD near the query does not differ significantly from that of nearby clusters, and we consider leaf clusters with cardinalities near $k$.
-Let $d$ be the LFD in this region.
-Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g.,\,we could have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
+\begin{theorem} 
+    Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
+Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
+        \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
+        \label{eq:methods:breadth-first-sieve-complexity}
+    \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
+        \mathcal{O} \Big( T \log T + L \log k \Big).
+        \label{eq:methods:depth-first-sieve-complexity}
+    \end{equation}
+\end{theorem}
+
+\begin{proof}
+Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
+
+Consider leaf clusters with cardinalities near $k$. Let $d$ be the LFD in this region, and $Q$ denote the set of candidate points and clusters.
+Then the number of leaf-clusters in $Q$ is bounded above by $2d$, where the bound is achieved if we have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
 In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
@@
 Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
-\begin{equation}
-    \mathcal{O} \Big( (T + L ) \log (T + L ) \Big).
-    \label{eq:methods:breadth-first-sieve-complexity}
-\end{equation}
+\begin{equation*}
+    \mathcal{O} \Big( (T + L ) \log (T + L ) \Big),
+\end{equation*} as claimed.
 
@@
 Thus, the complexity is
-\begin{equation}
-    \mathcal{O} \Big( T \log T + L \log k \Big).
-    \label{eq:methods:depth-first-sieve-complexity}
-\end{equation}
+\begin{equation*}
+    \mathcal{O} \Big( T \log T + L \log k \Big), 
+\end{equation*} as claimed. 
+\end{proof}
 

### 2025-01-15 reviewer-response.md
adjust wording for specificity in response
updated todos and added data aug to discussion

@@
 
-this is restating the complexity from the CHESS paper, but it's correct that it's an expected cost. Worst-case analysis isn't really useful here.
+This is restating the complexity from the CHESS paper, but it's correct that it's an expected cost. Worst-case analysis isn't really useful here.
 
@@
 
-Right after or in eq 2.1, immediately follow up with the simplified form where $r_1 = 2r_2$. Also drop subscript $X$, since we're dealing with only one dataset. Explain.
+We have revised the LFD definitions in this section for clarity.
 
@@
 
-Correct; fix in language.
+This has been fixed.
 
@@
 
-We should acknowledge this limitation, and suggest that in future work we would do something like favor augmentation along the first few PCs of the local manifold, but not in scope here.
-This goes in discussion.
+We acknowledge this limitation of our data augmentation process; we added discussion about an improved data augmentation process as future work in our Discussion section.
 
@@
 
-get rid of alg 2.4
+We removed what was formerly Algorithm 2.4.
 
@@
 
-we fixed this in pancakes, just backport it
+We have addressed this by adjusting the notation.
 
@@
 
-state theorem, which is the time complexity, and formalize the argument into a proof
+We have expressed claimed time complexities as precise theorems with assumptions and provided proofs.
 

### 2025-01-15 sections/5-discussion.tex
updated todos and added data aug to discussion

@@
 
+Additionally, we intend to improve the data augmentation process described in Section~\ref{sec:methods:synthetic-data} to better preserve the topological structure of the underlying dataset. 
+In particular, we expect that favoring augmentation along the first principal components of the local manifold may allow for more realistic data augmentation.
+
 We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.

### 2025-01-15 reviewer-response.md
discussion of protein sequence search

@@
 
-fine, cite 1, and discuss (NI)
+The fact that protein sequences have widely varying lengths does pose problems in terms of the potential sparsity of the manifold; it suggests that many distances would be maximized.
+While this makes the notion of a 'center' somewhat arbitrary and meaningless in domain terms, there would still be a computable cluster center (as demonstrated in work we just published in IEEE Big Data, largely done after the CAKES work).
+However, the approach in [1] is quite justified.
+We have added a discussion of protein sequence search as a specific example of a difficult domain, and cited [1] among others.
 

### 2025-01-15 sections/5-discussion.tex
discussion of protein sequence search

@@
 
+We note that despite the genericity of CAKES, effectively applying it to real-world datasets may typically require domain-specific knowledge.
+As an example, consider the domain of protein sequence search; protein sequences are of highly variable length (30--1000 amino acids are typical, with some outliers) and typically comprise a 20-letter alphabet.
+There have been numerous protein sequence search algorithms~\cite{kim2021entrance, daniels2013compressive, yu2015entropy, steinegger2018clustering}, and the one thing they have in common besides the application domain is that their development required deep domain-specific knowledge.
+While simple edit distance may be a poor choice for clustering when sequence lengths exhibit such variance, other choices of distance metrics and preprocessing could be used by CAKES, such as representing each sequence as a set of $k$-mers (short substrings of length $k$) and choosing Jaccard distance as the metric; this approach was chosen by~\cite{kim2021entrance}.
+
 The questions raised by this study suggest several additional avenues for future work.

### 2025-01-15 reviewer-response.md
some more fixes

@@
 
-We have attempted to address this by including additional computational experiments: in section 4.?, we have included plots indicating performance where we forced a balanced tree (in general, this performs worse, as we expected).
-TODO Najib go over this paragraph.
+We have added Section 4.4 to compare the performance of balanced and unbalanced clustering.
+We show that balanced trees perform strictly worse during search.
 
@@
 
-We don't have a python wrapper for cakes, and the other tools were in python. There is only one ground truth. Fix language starting in Results. **Najib, can you handle this one?** (NI)
+We have fixed our approach for this and have updated the language in the paper to reflect this.
+We now perform linear search on the augmented datasets and save that as the ground-truth, using a format that can be loaded in Python for calculating the recall of HNSW, ANNOY and FAISS-IVF.
+We verified that our linear search in Rust identifies the same neighbors on the original dataset as those provided in the ANN-Benchmarks suite of datasets.
+Thus, our approach now uses a unified ground truth across all algorithms and both programming languages.
 
@@
 
-will fix (NI)
+We have updated the README to include instructions on how to compile the code, run it, and reproduce the results from the manuscript.
 

### 2025-01-15 sections/2-methods.tex
some more fixes

@@
 
-Given a cluster $C$ with $|C|$ points, we define its two children by the following process.
-We take a random subsample $S$ of $\sqrt{|C|}$ of $C$'s points, and compute pairwise distances between all points in $S$.
-Using these distances, we compute the \textit{geometric median} of $S$; in other words, we find the point that minimizes the sum of distances to all other points in $S$.
-We define the \textit{center} of $C$ to be this geometric median.
-
-The \textit{radius} of $C$ is the maximum distance from the center to any other point in $C$.
-The point that is responsible for that radius (i.e.,\,the furthest point from the center) is designated the \textit{left pole} and the point that is furthest from the left pole is designated the \textit{right pole}.
-We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster that are closer to the left pole than to the right pole, and the right child contains all points in the cluster that are closer to the right pole than to the left pole.
-Without loss of generality, we assign to the left child those points that are equidistant from the two poles.
-Starting from a root cluster $\mathcal{R}$ containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria (e.g., minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc).
-This process is described in Algorithm \ref{alg:methods:partition}.
-During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-half}.
-
 \begin{algorithm} % enter the algorithm environment
@@
 
-Note that the tree $\mathcal{T}$ produced by Algorithm~\ref{alg:methods:partition} is not necessarily balanced.
-In fact, for real datasets, we expect anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
-The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
-In this sense, the imbalance in the tree is a feature, not a bug, as it reflects the underlying structure of the data.
+Given a cluster $C$ with $|C|$ points, we define its two children by the following process.
+We take a random subsample $S$ of $\sqrt{|C|}$ of $C$'s points, and compute pairwise distances between all points in $S$.
+Using these distances, we compute the \textit{geometric median} of $S$; in other words, we find the point that minimizes the sum of distances to all other points in $S$.
+We define the \textit{center} of $C$ to be this geometric median.
+
+The \textit{radius} of $C$ is the maximum distance from the center to any other point in $C$.
+The point that is responsible for that radius (i.e.,\,the furthest point from the center) is designated the \textit{left pole} and the point that is furthest from the left pole is designated the \textit{right pole}.
+We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster that are closer to the left pole than to the right pole, and the right child contains all points in the cluster that are closer to the right pole than to the left pole.
+Without loss of generality, we assign to the left child those points that are equidistant from the two poles.
+Starting from a root cluster $\mathcal{R}$ containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria (e.g., minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc).
+This process is described in Algorithm \ref{alg:methods:partition}.
+During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-half}.
 
@@
 \label{sec:methods:clustering:depth-first-reordering}
-
-%After building the tree, we reorder the dataset so that the points are stored in the order that they would be visited during a depth-first traversal of the tree.
 
@@
 In this work, we introduce a new approach wherein, after building $\mathcal{T}$, we reorder the dataset so that points are stored in a depth-first order.
-Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset, based on a useful invariant.
+Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset.
 The root cluster $\mathcal{R}$ has an offset of zero and a cardinality equal to the number of points in the dataset.
 A left child has the same offset its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
-With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS with $\mathcal{O}(n)$ space.
+With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS but with $\mathcal{O}(n)$ memory overhead.
 
@@
 \label{sec:methods:clustering:complexity}
+
+% Note that the tree $\mathcal{T}$ produced by Algorithm~\ref{alg:methods:partition} is not necessarily balanced.
+% In fact, for real datasets, we expect anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
+% The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
+% In this sense, the imbalance in the tree is a feature, not a bug, as it reflects the underlying structure of the data.
 
@@
 We also note here that this $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
-This can realistically only happen when the data are uniformly distributed in the embedding space.
-This represents the worst-case scenario for subsequent search algorithms because the LFD of the dataset will approach or equal the embedding dimension.
-For real-world datasets with low LFD, we expect $\mathcal{T}$ to be highly unbalanced, reflecting the topological structure of the data.
-When building trees for such datasets, their depths would be larger than $\log n$, which would seem to increase the cost of tree-building.
+In practice, for real datasets, we expect Algorithm~\ref{alg:methods:partition} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
+The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
+
+When building a tree for a real dataset, the depth would be larger than $\log n$, which would seem to increase the cost of tree-building.
 However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
@@
 In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.
+
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
+\label{sec:methods:knn-search}
+
+Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster $\mathcal{R}$ of the tree $\mathcal{T}$ (constructed by Algorithm~\ref{alg:methods:partition}), we can now formally pose the $k$-NN search problem.
+Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
+In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
+
+In this section, we present three novel algorithms for exact $k$-NN search:
+Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
+In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
+These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
 
@@
     \label{fig:methods:deltas}
-    \vskip -0.5in
+    \vskip -0.25in
 \end{figure}
+
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
+\label{sec:methods:knn-search:repeated-rnn}
+
+This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
 
@@
 \hfill
-\begin{minipage}{.45\textwidth}
+\begin{minipage}{.475\textwidth}
     \begin{algorithm}[H]
@@
 
-\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
-\label{sec:methods:knn-search}
+For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
+If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
 
-Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster $\mathcal{R}$ of the tree $\mathcal{T}$ (constructed by Algorithm~\ref{alg:methods:partition}), we can now formally pose the $k$-NN search problem.
-Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
-In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
-
-In this section, we present three novel algorithms for exact $k$-NN search:
-Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
-These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+In particular, we increase the radius by a factor of
+\begin{equation}
+    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \right)
+    \label{eq:methods:repeated-rnn-factor}
+\end{equation}
+where $\mu$ is the multiplicative inverse of the harmonic mean of the LFD of the clusters in $Q$, i.e.,\,$\mu = \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$.
+We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
+We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.
 
@@
             \ElsIf{$\sum_{C \in Q} |C| >= k$}
-                \State $\mu \Leftarrow |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
+                \State $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
                 \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
@@
 \end{algorithm}
-
-\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{sec:methods:knn-search:repeated-rnn}
-
-This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
-
-For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
-If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
-
-Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of
-\begin{equation}
-    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \right)
-    \label{eq:methods:repeated-rnn-factor}
-\end{equation}
-where $\mu$ is the multiplicative inverse of the harmonic mean of the LFD of the clusters in $Q$, i.e.,\,$\mu = |Q| \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$.
-We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
-We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.
 

### 2025-01-15 sections/4-results.tex
some more fixes

@@
 A consequence of this is that, with high probability, for every point in $\mathcal{R}$, its distance from $c$ is greater than $\tfrac{r}{2}$;
-in other words, $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r)$ contains the entire dataset.
+in other words, $B(c, \tfrac{r}{2})$ contains only $c$ while $B(c, r)$ contains the entire dataset.
 Given our definition of LFD in Equation~\ref{eq:methods:lfd-original}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
@@
 
-In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly pn the Random dataset.
+In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly on the Random dataset.
 As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.

### 2025-01-15 sections/4-results.tex
updated 4.1

@@
 \label{sec:results}
+
+The algorithms presented in this paper rely heavily on the local fractal dimension (LFD) being much smaller than than the embedding dimension of the dataset.
+As such, we begin by examining our assumptions about the LFD of real-world datasets (see Section~\ref{sec:results:lfd-of-datasets} and Figure~\ref{fig:results:lfd-plots}).
+
+We continue by comparing the performance of the CAKES algorithms against HNSW, ANNOY and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift and Random datasets.
+We also compare how this performance scales to the augmented versions of these datasets.
+We also report the performance of CAKES on the Silva and RadioML datasets (see Section~\ref{sec:results:scaling-behavior-and-recall}, Figure~\ref{fig:results:scaling-plots}, and Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random}).
+
+Finally, we test our intuitions about unbalanced clustering being better for search 
+than balanced clustering (see Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} and Figure~\ref{fig:results:distance-counts}).
 
@@
 Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
-Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-Mnist, Glove-25, Sift, Random, Silva 18S, and RadioML.
-In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD, as it seems most relevant to CAKES's performance (see Figure~\ref{fig:results:lfd-plots}).
-% The horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
-% We plot lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
-In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD, as it seems most relevant to CAKES's performance.
-% In order to have the plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality.
-% In other words, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.
+Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and RadioML.
+In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
+This is because our algorithms scale exponentially in the LFD, and we are interested in the worst-case behavior of the algorithms.
 
-Figure~\ref{fig:results:fashion-mnist-lfd} shows the LFD by depth for Fashion-Mnist.
-% We observe that until about depth 5, LFD is low, as the 95$^{th}$ percentile (orange plotted line) is less than 4, and the median (red line) is just above 2.
-% For depths 15 through 25, we observe that LFD increases, with the 95$^{th}$ percentile (orange plotted line) slightly less than 6, and the median near 3.
-% Finally, for depths 25 through the maximum depth, we observe that the LFD decreases again, as the 95$^{th}$ percentile is between 3 and 4, and the median is less than 2.
-We observe that until approximately depth 5, Fashion-Mnist's LFD is low (i.e., less than 4).
-It then starts increasing, reaching a peak of about 6 near depth 25, before decreasing again until the maximum depth.
+The Fashion-MNIST dataset has an embedding dimension of 784 and uses the Euclidean distance metric.
+In Figure~\ref{fig:results:fashion-mnist-lfd} we observe that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
+It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
 
-% Relative to Fashion-Mnist, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
-% Percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-% The 95$^{th}$ percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths.
-% Before depth 25, the 95$^{th}$ percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
-% The median hovers near 1.5 for all depths.
+The Glove-25 dataset has an embedding dimension of 25 and uses the cosine distance function which, notably, is not a metric.
+Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
+All percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
+In particular, Glove-25's LFD is less than 3 for all depths.
 
-Relative to Fashion-Mnist, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
-Percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-In particular, Glove-25's LFD is less than 3 for all depths. 
-Before depth 25, the LFD hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
+The Sift dataset has an embedding dimension of 128 and uses the Euclidean distance metric.
+Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
+It increases sharply to a peak of 9 around a depth of 10.
+It then decreases smoothly until reaching the deepest leaves in the tree.
 
-% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-Mnist and Glove-25.
-% Until about depth 5, the LFD is between 2 and 6.
-% From about depth 5 through about depth 20, the 95$^{th}$ percentile is greater than 6, even exceeding 8 near depth 10.
-% From about depth 20 onward, the 95$^{th}$ percentile is less than 6, and from about depth 30 onward, it is between 3 and 4.
-% The median reaches its peak of about 5 at around depth 15, but hovers near 2 after about depth 30.
-
-Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-Mnist and Glove-25.
-Until about depth 5, the LFD is between 2 and 6. LFD is greater than 6 until about depth 20, even exceeding 8 near depth 10.
-From about depth 20 onward, LFD falls below 6, and from about depth 30 onward, it hovers between 3 and 4.
-
-In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as those of Sift.
-We observe that the LFD starts as high as 20 (for all data) at depth 0.
-As shown by the needlepoint shape of the plot, the LFD for all clusters has a very narrow spread from depth 0 through depth 5.
-After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth.
-All percentile lines seem to decrease linearly with depth.
-The LFD of approximately 20 at depth 0 (i.e.,\,the root cluster $\mathcal{R}$) is what we expect for this random dataset.
+We generated the Random dataset to have the same cardinality and dimensionality as Sift.
+We used a uniform distribution in a 128-dimensional unit-hypercube to generate the points and the Euclidean metric to measure distances among them.
+Figure~\ref{fig:results:random-lfd} shows that the character of this dataset is significantly different from the others.
+The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
+The spread in LFD starts very small for the first few clusters and increases as depth increases.
+The LFD of approximately 20 for the root cluster $\mathcal{R}$ is what we expect for this random dataset.
 To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
@@
 in other words, $B(c, \tfrac{r}{2})$ contains only $c$ while $B(c, r)$ contains the entire dataset.
-Given our definition of LFD in Equation~\ref{eq:methods:lfd-original}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
+Given our definition of LFD in Equation~\ref{eq:methods:lfd-half}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
+Theoretically, we the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
+This reflects the difference between how we empirically measure the LFD and the value we would expect.
+With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^128$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
+Unfortunately, such a large sample is practically impossible to generate.
 
-Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-In particular, LFD is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
-% The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 128.
+The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712 but were provided in an MSA of width 50,000.
+As such the embedding dimension is somewhere between 3,712 and 50,000.
+We use the Levenshtein edit distance (a metric) to measure distances between sequences.
+This dataset, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
+In particular, LFD is less than about 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
 
-% Relative to the other  datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
-% Notably, however, for the first approximately 15 depths, the 95$^{th}$ percentile and median LFD are very close to 0.
-% Both percentiles increase sharply after depth 20, nearly reaching 16.
-% LFD remains high even as depth increases, with the 95$^{th}$ percentile fluctuating between about 6 and 15 for depths about 20-40.
-% It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60.
-% The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.
-The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or---as this is still a synthetic dataset---nonuniform generation of data.
-Silva, in contrast, appears to obey the manifold hypothesis at all scales.
-
-Relative to the other  datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
-Notably, however, for the first approximately 15 depths, LFD is very close to 0, although it increases sharply after depth 20, nearly reaching 16.
-LFD remains high even as depth increases, fluctuating between about 6 and 15 for depths about 20-40.
-It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60.
-The inconsistency of LFD in the RadioML dataset suggests that it obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth; this could be due to dense sampling or---as this is still a synthetic dataset---nonuniform generation of data.
-Silva, in contrast, appears to obey the manifold hypothesis at all scales.
+The Radio-ML dataset consists of measurements of radio-frequency signals using 1,024 dimensional complex-valued vectors.
+We use the Dynamic Time Warping distance metric on this dataset.
+This dataset is synthetic~\cite{oshea2018radioml} but uses a far more elaborate generation process than our Random dataset.
+The LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
+Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
+Within each of the tree portions, this dataset has a character very similar to that of the Random dataset.
+This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth;
+This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
 
@@
     \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/lfd/fashion-mnist.png}\\
-        \subcaption{Fashion-mnist}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/lfd/fashion-mnist.png}\\
+        \subcaption{Fashion-MNIST}
         \label{fig:results:fashion-mnist-lfd}
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/lfd/glove-25.png}\\
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/lfd/glove-25.png}\\
         \subcaption{Glove-25}
@@
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/lfd/sift.png}\\
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/lfd/sift.png}\\
         \subcaption{Sift}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/lfd/random.png}\\
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/lfd/random.png}\\
         \subcaption{A random dataset}
@@
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/lfd/silva-SSU-Ref.png}\\
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/lfd/silva-SSU-Ref.png}\\
         \subcaption{Silva 18S}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/lfd/radio-ml.png}\\
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/lfd/radio-ml.png}\\
         \subcaption{RadioML}
@@
 
-Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
+Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-MNIST under Euclidean distance,
 Glove-25 under cosine distance, and
@@
 
-Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-Mnist, Glove-25, Sift, and Random respectively.
+Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-MNIST, Glove-25, Sift, and Random respectively.
 Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
@@
 
-Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
+Figure~\ref{fig:results:scaling-plots} shows that for Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
 Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset (see Figure~\ref{fig:results:scaling-plots}).
-% For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
 
 Which one of the three CAKES algorithms is fastest also differs by dataset.
-For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
+For Fashion-MNIST, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
 On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
@@
 We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
-% For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
 
@@
 While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-% At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
-In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
+In contrast, CAKES has perfect recall on Fashion-MNIST and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
 
@@
 As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-% HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
-% in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
 HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets,as shown in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
@@
 Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
-% With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
-% Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
 
@@
 \begin{figure}
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
-        \subcaption{Fashion-Mnist for $k=10$.}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
+        \subcaption{Fashion-MNIST for $k=10$.}
         \label{fig:results:fashion-mnist-scaling}
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
         \subcaption{Glove-25 for $k=10$.}
@@
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{Sift for $k=10$.}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{plots/random_PermutedBall_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{plots/random_PermutedBall_10_throughput.png}
         \subcaption{A random dataset for $k=10$.}
@@
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
         \subcaption{Silva for $k=10$.}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{plots/radio-ml_Ball_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{plots/radio-ml_Ball_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
@@
     We observe that the cardinality at which CAKES algorithms outperform linear search differs by dataset.
-For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
-    % For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
-    }
+For Fashion-MNIST, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.}
     \label{fig:results:scaling-plots}
@@
 \begin{table}
-    \caption{Throughput (QPS) and Recall on the Fashion-Mnist dataset.
+    \caption{Throughput (QPS) and Recall on the Fashion-MNIST dataset.
     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
@@
 \begin{figure}
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
         \subcaption{Repeated $\rho$-NN}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
         \subcaption{Repeated $\rho$-NN}
@@
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
         \subcaption{Breadth First Sieve}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
         \subcaption{Breadth First Sieve}
@@
     \\
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
         \subcaption{Depth First Sieve}
@@
     \end{subfigure}%
-    \begin{subfigure}[b]{0.47\textwidth}
-        \includegraphics[width=0.9\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
+    \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
         \subcaption{Depth First Sieve}
@@
     \end{subfigure}%
-    \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-Mnist dataset.
+    \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
     Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.

### 2025-01-15 sections/2-methods.tex
expanded section on clustering strategies

@@
 However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
-Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
+Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here, but we explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
 In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.

### 2025-01-15 sections/4-results.tex
expanded section on clustering strategies

--- 2025-01-15 sections/4-results.tex
@@
 In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
+These strategies are the Cartesian product of balanced vs.
+unbalanced clustering, and the presence vs.
+absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
 To help with this analysis, we also added some instrumentation to our implementation of CAKES to count the number of distance computations performed during search.
-These are shown in Figure~\ref{fig:results:distance-counts}.
+We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.
+
+These four comparisons (referred to as Ball, BalancedBall, PermutedBalancedBall, and PermutedBall, where depth-first-reordering is referred to as Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset are shown in Figure~\ref{fig:results:distance-counts}.
+Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse; the instances (such as Depth First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which is effectively PermutedBall, no longer do.
+We note that on this dataset, depth-first reordering (the Permuted plotlines) seems to have little effect.
 

### 2025-01-15 sections/4-results.tex
fixed bad merge, please proofread S4.1

--- 2025-01-15 sections/4-results.tex
@@
 Given our definition of LFD in Equation~\ref{eq:methods:lfd-half}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
-Theoretically, we the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
+Theoretically, the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
 This reflects the difference between how we empirically measure the LFD and the value we would expect.

### 2025-01-15 reviewer-response.md
went through TODOs in response.

@@
 
-write in plain english right after eq 2.3, explaining in plain english the rationale for the min(2,x) expansion: namely, we don't think the very low LFD would be representative of what we'd see if we massively increased the radius (TODO Morgan)
+write a rebuttal here, mentioning Thm 2.1 (TODO Najib)
 
@@
 
-offload to figure captions where possible, don't spell out everything
+offload to figure captions where possible, don't spell out everything (TODO MORGAN)
 

### 2025-01-16 reviewer-response.md
more offloading to figure captions

@@
 
-TODO We agree that because the time complexity is highly depedent on the manifold structure of the data, this was insufficiently rigorous.
+We agree that because the time complexity is highly dependent on the manifold structure of the data, this was insufficiently rigorous.
 We have attempted to address this in a few ways.
-MORGAN TODO: to be addressed Wednesday.
-Changing to a theorem-style structure ...
-Also possible TODO: axe the 2.1.1 (?) description to avoid redundancy but include that we do NOT expect a balanced tree, and that's a good thing.
-Move the last section of 2.1.1 to the section on complexity analysis
+First, we added the condition "when the dataset has a manifold structure" to our statement about asymptotic complexity of entropy-scaling search in Section 1.2.
+Additionally, we added a paragraph to Section 1.2 justifying our approach to complexity analyses, and restated it in Section 2.2.2 and Section 2.2.5 before giving runtime expressions.
 
@@
 
-This is restating the complexity from the CHESS paper, but it's correct that it's an expected cost. Worst-case analysis isn't really useful here.
+This is correct; we added justification for our approach to complexity analyses in Sections 1.2, 2.2.2, and 2.2.5.
 
@@
 
-offload to figure captions where possible, don't spell out everything (TODO MORGAN)
+We attempted to address this by omitting unnecessary details, offloading observations to figure captions, and eliminating redundant sections of the discussion.
 

### 2025-01-16 sections/1-introduction.tex
more offloading to figure captions

@@
 The entropy-scaling search paradigm exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms exhibit asymptotic complexity that scales with topological properties of the dataset, instead of its cardinality.
+Importantly, as suggested by their name, entropy-scaling search algorithms exhibit asymptotic complexity that scales with topological properties of the dataset, instead of its cardinality, when the dataset has a manifold structure.
 In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a hierarchical clustering approach.
@@
 In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language.
+
+We also provide a theoretical analysis of the time complexity of CAKES's algorithms in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
+These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
+Given that CAKES's algorithms are intended to be used on datasets with a manifold structure, complexity analysis assuming a uniform distribution of data would be uninformative. 
+As a result, we assume in our analyses that the dataset has a manifold structure, as quantified by low fractal dimension and metric entropy.
+For our purposes, it makes more sense to analyze the expected performance of CAKES's algorithms under these assumptions.

### 2025-01-16 sections/2-methods.tex
more offloading to figure captions

@@
 
+As stated in Section~\ref{sec:intoduction:entropy-scaling-search}, Theorem~\ref{thm:methods:rnn-complexity} assumes that the dataset $X$ has a manifold structure.
+
 % The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data or, in other words, that it arises from the same generative process as the rest of the dataset.
@@
     where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
+    \label{thm:methods:rnn-complexity}
 \end{theorem}
@@
 Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.
+As stated in Section~\ref{sec:intoduction:entropy-scaling-search}, Theorem~\ref{thm:methods:sieve-complexity} assumes that the dataset $X$ has a manifold structure.
 
@@
     Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
+Let $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$ and   $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of $X$, $d$ is the LFD of $X$, $\hat{|C|}$ is the mean cardinality of clusters overlapping the query ball, and $k$ is the number of nearest neighbors.
 Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
@@
     \end{equation}
+    \label{thm:methods:sieve-complexity}
 \end{theorem}

### 2025-01-16 sections/4-results.tex
more offloading to figure captions

@@
 In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
-This is because our algorithms scale exponentially in the LFD, and we are interested in the worst-case behavior of the algorithms.
+% This is because our algorithms scale exponentially in the LFD, a
 
@@
 For algorithms that exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.
+For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
+For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.
 
@@
 
-Figure~\ref{fig:results:scaling-plots} shows that for Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
-Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset (see Figure~\ref{fig:results:scaling-plots}).
+% Figure~\ref{fig:results:scaling-plots} shows that for Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
+% Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset (see Figure~\ref{fig:results:scaling-plots}).
 
-Which one of the three CAKES algorithms is fastest also differs by dataset.
-For Fashion-MNIST, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
-On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
-With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
-We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
+% Which one of the three CAKES algorithms is fastest also differs by dataset.
+% For Fashion-MNIST, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
+% On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
+% With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
+% We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
 
-On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
-While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-In contrast, CAKES has perfect recall on Fashion-MNIST and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
+% While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+% In contrast, CAKES has perfect recall on Fashion-MNIST and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
 
-In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly on the Random dataset.
-As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets,as shown in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
+% In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly on the Random dataset.
+% As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
+% HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets,as shown in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
 
@@
 Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
-
-With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it levels off at higher cardinalities.
-Depth-First Sieve is the fastest CAKES algorithm for lower multipliers, but for larger cardinalities, Repeated $\rho$-NN becomes the fastest.
-For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance; in particular, throughput of CAKES's algorithms is identical within three significant figures.
 
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-    For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
-    For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.
+    For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search.
     We observe that the cardinality at which CAKES algorithms outperform linear search differs by dataset. 
-For Fashion-MNIST, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.}
+    % For Fashion-MNIST, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
+    For Fashion-MNIST, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
+    On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
+    With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
+    % We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
+    With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it levels off at higher cardinalities.
+    Depth-First Sieve is the fastest CAKES algorithm for lower multipliers, but for larger cardinalities, Repeated $\rho$-NN becomes the fastest.
+    For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance; in particular, throughput of CAKES's algorithms is identical within three significant figures.
+    On all three of these datasets, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove},\ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random}.
+    }
     \label{fig:results:scaling-plots}
@@
     \caption{Throughput (QPS) and Recall on the Fashion-MNIST dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+    In contrast, CAKES has perfect recall on Fashion-MNIST, since it used with Euclidean distance, which is a metric.}
     \label{tab:results:qps-and-recall-fmn}
@@
     \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+    In contrast, CAKES has near-perfect recall on Glove-25, since it used with cosine distance, which is not a metric.}
     \label{tab:results:qps-and-recall-glove}
@@
     \caption{Throughput (QPS) and Recall on the Sift dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+    In contrast, CAKES has perfect recall on Sift, since it is used with Euclidean distance, which is a metric.}
     \label{tab:results:qps-and-recall-sift}
@@
     \caption{Throughput (QPS) and Recall on the Random dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.}
+    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.
+In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
+    As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
+    HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
+    }
     \label{tab:results:qps-and-recall-random}

### 2025-01-16 sections/5-discussion.tex
more offloading to figure captions

@@
 This trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly estimate the correct radius for $k$ hits.
-These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
-Exploring this sensitivity is a potential avenue for future study, as it remains to learn at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
+% These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
+% Exploring this sensitivity is a potential avenue for future study, as it remains to learn at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
 

### 2025-01-16 reviewer-response.md
lots more upadtes

@@
 
-write a rebuttal here, mentioning Thm 2.1 (TODO Najib)
+We have, in Theorem 2.1, explicitly stated the assumption that the query is drawn from the same distribution as the other points in the dataset.
+This should ensure that, in practical applications with real-world datasets, the LFD around a query is similar to the LFD around its nearest neighbors.
+In the theoretical worst case where the LFD around a query is extremely low, repeating tree-search with doubling radii might not encounter any additional points, we may introduce an extra factor (logarithmic in the distance between the query and the k-th nearest neighbor) the in contribution of the tree search to the overall complexity.
+However, we can easily detect and sidestep this issue in our implementation by calculating the distance from the query to the center of the root cluster and using that distance, along with the radius and LFD of the root cluster, to determine an initial radius for tree-search.
+This approach, once again, brings us to a constant factor for the contribution from tree-search.
 

### 2025-01-16 sections/4-results.tex
lots more upadtes

@@
 Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
-Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and RadioML.
+Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
 In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
@@
 
-Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-MNIST under Euclidean distance,
-Glove-25 under cosine distance, and
-Sift under Euclidean distance.
-Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with the same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
-Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
-For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.
-The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
-The left-most point on each line is at the cardinality of the original dataset.
-The vertical axis denotes throughput in queries per second.
-Both axes are on a logarithmic scale.
-In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k = 100$ can be found in the Supplement.
-For HNSW and ANNOY, we report the recall for each measurement in the plots.
-For algorithms that exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.
-For some algorithms, we were not able to take measurements for each cardinality because the index-building required more RAM than was available (i.e.,\,more then 512GB).
-For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.
+We benchmark the CAKES algorithms, na\"{i}ve linear search, HNSW, ANNOY, and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift, and Random datasets.
+We augment these datasets with synthetic points to examine how performance scales with cardinality.
+We also benchmark the CAKES algorithms on the Silva and RadioML datasets, and we subsample these datasets instead of augmenting them to examine how performance scales with cardinality.
+For each dataset, we use the corresponding distance function (see Table~\ref{tab:datasets:summary}).
+For HNSW, ANNOY, and FAISS-IVF, allow for a hyper-parameter search to tune their index for maximum recall.
+For CAKES, we build the tree and use our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) to select the fastest algorithm for each dataset and cardinality.
+We show results for $k=10$ in the main text, but results for $k=100$ can be found in the Supplement.
+Figure~\ref{fig:results:scaling-plots} shows the scaling behavior of these algorithms on these datasets while Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} report the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-MNIST, Glove-25, Sift, and Random respectively.
+
+On the Fashion-MNIST, Glove-25, and Sift datasets (in Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} respectively), we observe that as cardinality increases, the Depth-First Sieve algorithm is consistently the fastest CAKES algorithm with a throughput that is constant in the cardinality of the dataset.
+The Breadth-First Sieve algorithm is the usually the second fastest, also with a nearly constant throughput across all cardinalities.
+The Repeated $\rho$-NN algorithm, however, falls off in throughput as cardinality increases.
+All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset.
+In contrast, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities and have near constant throughput as cardinality increases, but their recall degrades quickly as cardinality increases.
+FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize.
+
+On the Random dataset, HNSW and ANNOY are still the fastest algorithms but exhibit recall values near 0.
+CAKES's algorithms show linearly decreasing throughput as cardinality increases and are also slower than na\"{i}ve linear search.
 
@@
 
-% Figure~\ref{fig:results:scaling-plots} shows that for Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
-% Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset (see Figure~\ref{fig:results:scaling-plots}).
-
-% Which one of the three CAKES algorithms is fastest also differs by dataset.
-% For Fashion-MNIST, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
-% On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
-% With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
-% We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
-
-% While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-% In contrast, CAKES has perfect recall on Fashion-MNIST and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).
-
-% In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly on the Random dataset.
-% As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-% HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets,as shown in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.
-
 For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
 Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
+For the Silva dataset (see Figure~\ref{fig:results:silva-scaling}), we observe that throughput initially decreases linearly as cardinality increases, but levels off at higher cardinalities.
+Depth-First Sieve is, once again, the fastest CAKES algorithm.
+For Radio-ML (see Figure~\ref{fig:results:radioml-scaling}), however, we see that throughput declines nearly linearly with cardinality, and that the three CAKES algorithms are, again, slower that na\"{i}ve linear search.
 
 \begin{figure}
+    \captionsetup[subfigure]{aboveskip=-5pt,belowskip=-0.5pt}
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
         \subcaption{Fashion-MNIST for $k=10$.}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
         \subcaption{Glove-25 for $k=10$.}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
         \subcaption{Sift for $k=10$.}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{plots/random_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
         \subcaption{A random dataset for $k=10$.}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
         \subcaption{Silva for $k=10$.}
@@
     \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{plots/radio-ml_Ball_10_throughput.png}
+        \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
@@
     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-    For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search.
-    We observe that the cardinality at which CAKES algorithms outperform linear search differs by dataset. 
-    % For Fashion-MNIST, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
-    For Fashion-MNIST, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
-    On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
-    With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
-    % We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
-    With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it levels off at higher cardinalities.
-    Depth-First Sieve is the fastest CAKES algorithm for lower multipliers, but for larger cardinalities, Repeated $\rho$-NN becomes the fastest.
-    For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance; in particular, throughput of CAKES's algorithms is identical within three significant figures.
-    On all three of these datasets, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove},\ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random}.
-    }
+    For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
+    For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
+    With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
+    Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
+    For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
+    HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
+    In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
     \label{fig:results:scaling-plots}
@@
 
-These four comparisons (referred to as Ball, BalancedBall, PermutedBalancedBall, and PermutedBall, where depth-first-reordering is referred to as Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset are shown in Figure~\ref{fig:results:distance-counts}.
-Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse; the instances (such as Depth First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which is effectively PermutedBall, no longer do.
+These four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is referred to as Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset are shown in Figure~\ref{fig:results:distance-counts}.
+Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
+the instances (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which is effectively PermutedBall, no longer do under a balanced clustering.
 We note that on this dataset, depth-first reordering (the Permuted plotlines) seems to have little effect.

### 2025-01-16 sections/5-discussion.tex
lots more upadtes

--- 2025-01-15 sections/5-discussion.tex
@@
 CAKES's algorithms are designed to be most effective when the data uphold the manifold hypothesis, or in other words, when the data are constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
+CAKES uses an unbalanced clustering algorithm to build a binary tree that better captures the manifold structure of the data and provides for faster search performance.
 As a consequence, these algorithms do not perform well on data with random distributions, because of the absence of a manifold structure.
 In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
-As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on datasets with low LFD, and scale linearly on datasets with high LFD.
+As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on real-world datasets with low LFD, and scale linearly on synthetic datasets with high LFD.
 
@@
 We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly overestimates the correct radius $\rho_k$ and to improve upon the factor of radial increase in Equation~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
-
-As the previous paragraph suggests, the fastest CAKES algorithm varies by dataset.
-For example, on Glove-25, Repeated $\rho$-NN exhibits the highest throughput of any CAKES algorithm, despite having the lowest throughput of the three on Sift and Fashion-Mnist.
-On Silva, Repeated $\rho$-NN is comparable to the other CAKES algorithms at low multipliers, but becomes the fastest at the higher multipliers.
-When we view these results in light of the LFD plots in Figure~\ref{fig:results:lfd-plots}, we realize that Repeated $\rho$-NN appears to be the fastest CAKES algorithm on datasets where a large proportion of the data have a very low LFD, such as Glove-25 and Silva.
-This trend is unsurprising, given that Repeated $\rho$-NN relies on a low LFD around the query in order to quickly estimate the correct radius for $k$ hits.
-These findings suggest that Repeated $\rho$-NN is the most ``sensitive'' to the manifold structure of the dataset.
-Exploring this sensitivity is a potential avenue for future study, as it remains to learn at what LFD Repeated $\rho$-NN transitions away from being the fastest CAKES algorithm, and whether this transition is sharp.
 
@@
 Additionally, we intend to improve the data augmentation process described in Section~\ref{sec:methods:synthetic-data} to better preserve the topological structure of the underlying dataset.
-In particular, we expect that favoring augmentation along the first principal components of the local manifold may allow for more realistic data augmentation.
+In particular, we expect that favoring augmentation along the first few principal components of the local manifold may allow for more realistic data augmentation.
 
 We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
-Such an encoding would allow us to store the data in a compressed format, and to perform search on the compressed data without decompressing it.
-This would allow us to perform fast search on datasets that are too large to fit in memory.
+Such an encoding would allow us to store the data in a compressed format, and to perform search on the compressed data without decompressing it in its entirety.
+This would allow us to perform fast search on datasets that are otherwise too large to fit in memory.
 An exploration of compression and compressed search is another avenue for future work.

### 2025-01-16 sections/4-results.tex
minor typos and additions

@@
 This reflects the difference between how we empirically measure the LFD and the value we would expect.
-With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^128$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
+With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^{128}$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
 Unfortunately, such a large sample is practically impossible to generate.
@@
 For each dataset, we use the corresponding distance function (see Table~\ref{tab:datasets:summary}).
-For HNSW, ANNOY, and FAISS-IVF, allow for a hyper-parameter search to tune their index for maximum recall.
+For HNSW, ANNOY, and FAISS-IVF, we allow for a hyper-parameter search to tune their index for maximum recall.
 For CAKES, we build the tree and use our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) to select the fastest algorithm for each dataset and cardinality.
@@
 The Repeated $\rho$-NN algorithm, however, falls off in throughput as cardinality increases.
-All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset.
+All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
 In contrast, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities and have near constant throughput as cardinality increases, but their recall degrades quickly as cardinality increases.
-FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize.
+FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize recall.
 
@@
 the instances (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which is effectively PermutedBall, no longer do under a balanced clustering.
-We note that on this dataset, depth-first reordering (the Permuted plotlines) seems to have little effect.
+We note that on this dataset, depth-first reordering (the Permuted plotlines) seems to have little effect; depth-first reordering is intended to improve space complexity only.
 

### 2025-01-16 sections/1-introduction.tex
more edits after another reading

@@
 Researchers are collecting data at an unprecedented rate.
-In many fields, datasets grow exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
+In many fields, datasets are growing exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
 This indicates that the performance of computing systems will not keep pace with the growth of data.

### 2025-01-16 sections/3-datasets.tex
more edits after another reading

@@
 
-We benchmarked CAKES's algorithms against the state-of-the-art similarity search algorithms HNSW, ANNOY and FAISS-IVF, and two implementations of linear search.
-In particular, we use FAISS-Flat, a Python implementation of linear search, to provide ground-truth values for recall of HNSW, ANNOY, and FAISS-IVF.
-We use our own Rust implementation of linear search to provide ground-truth values for calculating recall of CAKES's algorithms.
-We plot the throughput of these two implementations of linear search as a reference point for all other algorithms in Figure~\ref{fig:results:scaling-plots}.
+We benchmarked the three CAKES algorithms against a na\"ive linear search implementation in Rust.
+We also benchmarked against three state-of-the-art similarity search algorithms: HNSW, ANNOY, and FAISS-IVF in Python.
+We verified that our implementation of linear search produces the same neighbors as provided by the ANN-Benchmarks suite for Fashion-MNIST, Glove-25 and Sift datasets.
+We then used this linear search implementation to find and store the ground-truth for the augmented versions of the datasets.
+We used this ground-truth to calculate recall for CAKES's algorithms in Rust and for HNSW, ANNOY, and FAISS-IVF in Python.
+We plot the results of these benchmarks in Figure~\ref{fig:results:scaling-plots}.

### 2025-01-16 sections/4-results.tex
more edits after another reading

@@
 This dataset, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-In particular, LFD is less than about 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
+In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
 
@@
 Within each of the tree portions, this dataset has a character very similar to that of the Random dataset.
-This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth;
+This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
 This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
@@
 Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
-the instances (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which is effectively PermutedBall, no longer do under a balanced clustering.
-We note that on this dataset, depth-first reordering (the Permuted plotlines) seems to have little effect; depth-first reordering is intended to improve space complexity only.
+the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which shows results using PermutedBall, no longer do so under a balanced clustering.
+We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.
 

### 2025-01-16 sections/5-discussion.tex
more edits after another reading

@@
 When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{T}{\lceil d \rceil} + L)$, where $T$ is the time complexity of tree-search and $L$ is the time complexity of leaf-search.
-Even though Repeated $\rho$-NN has the least time cost (compare to $\mathcal{O}(T\log{(T+L)} + L\log{(T+L)})$ for Breadth-First Sieve and $\mathcal{O}(T\log{T} + L\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
-We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
+Even though Repeated $\rho$-NN has the least time cost (compared to $\mathcal{O}\left((T + L)\log{(T+L)}\right)$ for Breadth-First Sieve and $\mathcal{O}(T\log{T} + L\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
+We believe that some of this discrepancy can be explained by the fact that Repeated $\rho$-NN can significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search.
+This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
 This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
@@
 
-While the fastest CAKES algorithm varies by dataset, we note that on all three ANN-benchmark datasets (Fashion-Mnist, Glove-25, and Sift), Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as the cardinality increases.
+While the fastest CAKES algorithm can vary by dataset and cardinality, we note that on all three ANN-benchmark datasets (Fashion-MNIST, Glove-25, and Sift), Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as the cardinality increases.
 This observation warrants further investigation, but it is especially promising given that both algorithms consistently outperform linear search for high cardinalities.
@@
 
-We note that despite the genericity of CAKES, effectively applying it to real-world datasets may typically require domain-specific knowledge.
+We note that despite the genericity of CAKES, effectively applying it to real-world datasets may require domain-specific knowledge.
 As an example, consider the domain of protein sequence search; protein sequences are of highly variable length (30--1000 amino acids are typical, with some outliers) and typically comprise a 20-letter alphabet.
 There have been numerous protein sequence search algorithms~\cite{kim2021entrance, daniels2013compressive, yu2015entropy, steinegger2018clustering}, and the one thing they have in common besides the application domain is that their development required deep domain-specific knowledge.
-While simple edit distance may be a poor choice for clustering when sequence lengths exhibit such variance, other choices of distance metrics and preprocessing could be used by CAKES, such as representing each sequence as a set of $k$-mers (short substrings of length $k$) and choosing Jaccard distance as the metric; this approach was chosen by~\cite{kim2021entrance}.
+While simple edit distance may be a poor choice for clustering when sequence lengths exhibit such variance, other choices of distance metrics and preprocessing could be used by CAKES, such as representing each sequence as a set of $k$-mers (short substrings of length $k$) and choosing Jaccard distance on these sets as the metric; this approach was chosen by~\cite{kim2021entrance}.
+If the distance function uses a global sequence alignment algorithm, such as Needleman-Wunsch~\cite{needleman1970general}, the dataset may also be preprocessed into a number of mutually exclusive subsets based on the lengths of the sequences, and a separate CAKES tree may be built for each subset.
+At search time, the query sequence would be assigned to the appropriate tree based on its length, and the search would be performed on that tree, and on a small number of trees on adjacent lengths' bins.
 
@@
 These include Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions) and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
-Incorporating these or other distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
+Incorporating new distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
 

### 2025-01-16 sections/1-introduction.tex
broke up S5, changed Silva dim

@@
 Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image} are a common source of neural network embeddings.
-Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
+Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes; the longest of these individual sequences is 3,712 amino acids, but in a multiple sequence alignment, the length becomes 50,000 letters.
 Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated signals of different modulation modes.

### 2025-01-16 sections/2-methods.tex
broke out one equation in 2.2.2 for aesthetics
more edits after another reading
broke up S5, changed Silva dim

@@
 We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the SILVA 18S rRNA dataset that contains genomic sequences with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
-Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
+To illustrate the difference, consider the SILVA 18S rRNA dataset that contains genomic sequences with unaligned lengths of up to 3,712 base pairs and aligned length of 50,000 base pairs.
+Hence, the \textit{embedding dimension} of this dataset is at least 3,712 and at most 50,000; as used in this paper it is 3,712.
 However, physical constraints (namely, biological evolution and biochemistry) constrain the data to a lower-dimensional manifold within this space.
@@
 
-% Note that the tree $\mathcal{T}$ produced by Algorithm~\ref{alg:methods:partition} is not necessarily balanced.
-% In fact, for real datasets, we expect anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
-% The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
-% In this sense, the imbalance in the tree is a feature, not a bug, as it reflects the underlying structure of the data.
-
 The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}.
@@
 
-In this section, we present three novel algorithms for exact $k$-NN search:
-Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
-These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
-
 \begin{figure}[H]
@@
 \end{figure}
-
-\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{sec:methods:knn-search:repeated-rnn}
-
-This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
 
@@
 \end{minipage}
+
+In this section, we present three novel algorithms for exact $k$-NN search:
+Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
+In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
+These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
+\label{sec:methods:knn-search:repeated-rnn}
+
+This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
 
@@
 We can rearrange this equation to get $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
-Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get $k \left( 1 + 2 \cdot \left( \frac{\hat{|C|}}{k} \right) ^ {\hat{d}^{-1}} \right)^d$.
+Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get:
+\begin{equation*}
+    k \left( 1 + 2 \cdot \left( \frac{\hat{|C|}}{k} \right) ^ {\hat{d}^{-1}} \right)^d
+\end{equation*}
 
 Again using the assumption that since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
-By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has time complexity
-\begin{gather*}
-    \mathcal{O}
-    \Bigg(
-        \underbrace{
-            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-        }_{\textrm{tree-search}}
-        \ + \
-        \underbrace{
-            \overbrace{k}^{\textrm{output size}} \cdot
-            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
-        }_{\textrm{leaf-search}}
-    \Bigg)
-    \label{eq:methods:repeated-rnn-complexity}
-\end{gather*}
-as claimed.
+By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has time complexity claimed in Equation~\ref{eq:methods:repeated-rnn-complexity1}.
 \end{proof}
@@
 We remark that the scaling factor in Equation~\ref{eq:methods:repeated-rnn-complexity1} should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
-
-\subsubsection{Breadth-First Sieve}
-\label{sec:methods:knn-search:bredth-first-sieve}
-
-This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
-
-We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
-During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 
@@
 
-\vskip 0.1in
+\subsubsection{Breadth-First Sieve}
+\label{sec:methods:knn-search:bredth-first-sieve}
+
+This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+
+We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 

### 2025-01-16 sections/3-datasets.tex
broke up S5, changed Silva dim

@@
                 \hline
-                SILVA            & Levenshtein                 & 2,224,640          & 3,712 - 50,000         \\
+                SILVA            & Levenshtein                 & 2,224,640          & 3,712         \\
                 \hline
@@
 To demonstrate CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
-This dataset contains ribosomal RNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
+This dataset contains ribosomal RNA sequences of 2,224,640 genomes, with the longest sequence having 3,712 letters.
+We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
+We note that when put into a multiple sequence alignment, the length (and thus dimensionality) would be 50,000, but in that case one could use Hamming distance, which is computationally less expensive than Levenshtein distance.
 We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
-We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
 Since this dataset contains many redundant, i.e.,\,nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.

### 2025-01-16 sections/4-results.tex
broke up S5, changed Silva dim

@@
 
-The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712 but were provided in an MSA of width 50,000.
-As such the embedding dimension is somewhere between 3,712 and 50,000.
+The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712.
+As such the embedding dimension is 3,712, though as discussed previously, the embedding dimension would be 50,000 in a multiple sequence alignment.
 We use the Levenshtein edit distance (a metric) to measure distances between sequences.

### 2025-01-16 sections/5-discussion.tex
broke up S5, changed Silva dim

@@
 In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
+
+\subsection{Performance}
+
 As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on real-world datasets with low LFD, and scale linearly on synthetic datasets with high LFD.
-
 CAKES's performance on Sift against that on the Random dataset illustrates this phenomenon well.
@@
 We believe that some of this discrepancy can be explained by the fact that Repeated $\rho$-NN can significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search.
-This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
+This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity1}.
 This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
@@
 
+\subsection{Applicability}
+
 We note that despite the genericity of CAKES, effectively applying it to real-world datasets may require domain-specific knowledge.
@@
 At search time, the query sequence would be assigned to the appropriate tree based on its length, and the search would be performed on that tree, and on a small number of trees on adjacent lengths' bins.
+
+\subsection{Future Work}
 
@@
 
-CLAM and CAKES are implemented in Rust and the source code is available under an MIT license at https://github.com/URI-ABD/clam.
+\subsection{Availability}
+
+CLAM and CAKES are implemented in the Rust programming language, and the source code is available under an MIT license at https://github.com/URI-ABD/clam.

### 2025-01-16 sections/4-results.tex
minor tweaks

@@
 cluster depth across six datasets.
-The last dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset.
+The `random' dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset.
 In each plot, the horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.

### 2025-01-16 sections/1-introduction.tex
yet more edits

@@
 
-Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
+Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
 HNSW builds a multi-layered graph of the dataset.

### 2025-01-16 sections/2-methods.tex
minor tweaks
yet more edits

@@
 We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
-A distance of zero defines identity among points (i.e.,\,$f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
+A distance value of zero defines an identity among points (i.e.,\,$f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
 We also require that the distance function be symmetric (i.e.,\,$f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
@@
 The choice of an appropriate distance function varies by dataset and domain.
-For example, with neural-network embeddings, one could use Euclidean (L2) or cosine distance.
-With genomic or proteomic sequence data, Levenshtein or Hamming distance is useful.
+For example, with neural-network embeddings, one could use Euclidean or Cosine distance.
+With genomic or proteomic sequence data, Levenshtein and Hamming distances are useful.
 With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.
@@
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
-Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
+Section~\ref{sec:results:lfd-of-datasets} discusses this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
 For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but potentially highly variable at larger length scales, i.e.,\,when $r$ is large.
@@
     \begin{algorithmic} % enter the algorithmic environment
+        \Require $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
         \Require $C$, a cluster
@@
 
-\subsubsection{Complexity}
-\label{sec:methods:clustering:complexity}
+\subsubsection{Time Complexity}
+\label{sec:methods:clustering:time-complexity}
 
@@
 However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
-Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here, but we explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
+Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
+We do, however, explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
 In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.
 
-\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
-\label{sec:methods:knn-search}
-
-Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster $\mathcal{R}$ of the tree $\mathcal{T}$ (constructed by Algorithm~\ref{alg:methods:partition}), we can now formally pose the $k$-NN search problem.
-Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
-In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
-
 \begin{figure}[H]
+    \vskip -0.2in
     \centering
@@
     \label{fig:methods:deltas}
-    \vskip -0.25in
+    \vskip -0.5in
 \end{figure}
@@
         \begin{algorithmic}
+            \Require $f$, a distance function
             \Require $C$, a cluster
@@
         \begin{algorithmic}
+            \Require $f$, a distance function
             \Require $Q$, a set of clusters
@@
 
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
+\label{sec:methods:knn-search}
+
+Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster $\mathcal{R}$ of the tree $\mathcal{T}$ (constructed by Algorithm~\ref{alg:methods:partition}), we can now formally pose the $k$-NN search problem.
+Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
+In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
+
 In this section, we present three novel algorithms for exact $k$-NN search:
@@
 These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
-
-\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{sec:methods:knn-search:repeated-rnn}
-
-This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
-
-For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
-If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
-
-Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of
-\begin{equation}
-    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \right)
-    \label{eq:methods:repeated-rnn-factor}
-\end{equation}
-where $\mu$ is the multiplicative inverse of the harmonic mean of the LFD of the clusters in $Q$, i.e.,\,$\mu = \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$.
-We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
-We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.
 
@@
 
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
+\label{sec:methods:knn-search:repeated-rnn}
+
+This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
+
+For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
+If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
+
+Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+In particular, we increase the radius by a factor of
+\begin{equation}
+    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \right)
+    \label{eq:methods:repeated-rnn-factor}
+\end{equation}
+where $\mu$ is the multiplicative inverse of the harmonic mean of the LFD of the clusters in $Q$, i.e.,\,$\mu = \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$.
+We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
+We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.
+
 Intuitively, the factor by which we increase the radius should be \textit{inversely} related to the number of points found so far.
@@
 
-As stated in Section~\ref{sec:intoduction:entropy-scaling-search}, Theorem~\ref{thm:methods:rnn-complexity} assumes that the dataset $X$ has a manifold structure.
-
-% The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data or, in other words, that it arises from the same generative process as the rest of the dataset.
-% Given the uses of $k$-NN search in practice, this assumption is reasonable.
-% From this assumption, we can infer that the LFD near the query does not differ significantly from the (harmonic) mean of the LFDs of clusters near the query at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
-
 \begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
@@
         \Bigg)
-        \label{eq:methods:repeated-rnn-complexity1}
+        \label{eq:methods:repeated-rnn-complexity}
     \end{gather}
@@
 \end{theorem}
-
-% \textbf{Theorem.} \textit{Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
-Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is} \begin{gather}
-%     \mathcal{O}
-%     \Bigg(
-%         \underbrace{
-%             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-%         }_{\textrm{tree-search}}
-%         \ + \
-%         \underbrace{
-%             \overbrace{k}^{\textrm{output size}} \cdot
-%             \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
-%         }_{\textrm{leaf-search}}
-%     \Bigg).
-%     \label{eq:methods:repeated-rnn-complexity1}
-% \end{gather}
 
@@
 It remains to determine an estimate for $\rho_k$.
-% For this, we once again rely on the assumption that the query is drawn from the same distribution as the rest of the data, and thus the LFD at the query point is not significantly different from the LFD of nearby clusters.
 We let $\hat{d}$ be the (harmonic) mean LFD of the clusters in $Q$.
@@
 Again using the assumption that since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
-By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has time complexity claimed in Equation~\ref{eq:methods:repeated-rnn-complexity1}.
+By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has time complexity as claimed in Equation~\ref{eq:methods:repeated-rnn-complexity}.
 \end{proof}
 
-We remark that the scaling factor in Equation~\ref{eq:methods:repeated-rnn-complexity1} should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
+We remark that the scaling factor in Equation~\ref{eq:methods:repeated-rnn-complexity} should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).
+
+\subsubsection{Breadth-First Sieve}
+\label{sec:methods:knn-search:bredth-first-sieve}
+
+This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+
+We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
+Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 
@@
 Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.
-As stated in Section~\ref{sec:intoduction:entropy-scaling-search}, Theorem~\ref{thm:methods:sieve-complexity} assumes that the dataset $X$ has a manifold structure.
 
@@
 The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
-\begin{equation*}
-    \mathcal{O} \Big( (T + L ) \log (T + L ) \Big),
-\end{equation*} as claimed.
+Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is as claimed in Equation~\ref{eq:methods:breadth-first-sieve-complexity}.
 
-For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
-Thus, the complexity is
-\begin{equation*}
-    \mathcal{O} \Big( T \log T + L \log k \Big),
-\end{equation*} as claimed.
+For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is as claimed in Equation~\ref{eq:methods:depth-first-sieve-complexity}.
 \end{proof}

### 2025-01-16 sections/3-datasets.tex
yet more edits

@@
 This dataset contains ribosomal RNA sequences of 2,224,640 genomes, with the longest sequence having 3,712 letters.
+We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
 We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
-We note that when put into a multiple sequence alignment, the length (and thus dimensionality) would be 50,000, but in that case one could use Hamming distance, which is computationally less expensive than Levenshtein distance.
-We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
-Since this dataset contains many redundant, i.e.,\,nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.
+We note that the sequences in this dataset were provided in a multiple sequence alignment with a width of 50,000 characters.
+We could have used Hamming distance on the aligned sequences, but we chose to use Levenshtein distance on the unaligned sequences to help demonstrate the flexibility of CAKES in handling exotic distance functions.
+Under hamming distance, we could consider this dataset to have an embedding dimension of 50,000, but for Levenshtein distance, we state the dimensionality as the length of the longest sequence, i.e.\, 3,712.
 
@@
 
-As another example of an exotic distance function, we benchmarked CAKES on the RadioML dataset~\cite{oshea2018radioml}.
-The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
-Specifically, it comprises 24 modulation modes at 26 different signal-to-noise ratio (SNR) levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
+As another example of an exotic distance function, we benchmarked CAKES on the Radio-ML dataset~\cite{oshea2018radioml} under Dynamic Time Warping~\cite{muller2007dynamic} as the distance function.
+This dataset contains samples of synthetically generated signal captures of different modulation modes over a range of signal-to-noise ratio (SNR) levels.
+Specifically, it comprises 24 modulation modes at 26 different SNRs ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
 Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
@@
 We used a subset of this dataset, containing $97,304$ samples at 10dB SNR and used another $1,000$ samples at the same SNR as a hold-out set of queries.
-We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.
 

### 2025-01-16 sections/3-datasets.tex
fix capitalization

@@
 We could have used Hamming distance on the aligned sequences, but we chose to use Levenshtein distance on the unaligned sequences to help demonstrate the flexibility of CAKES in handling exotic distance functions.
-Under hamming distance, we could consider this dataset to have an embedding dimension of 50,000, but for Levenshtein distance, we state the dimensionality as the length of the longest sequence, i.e.\, 3,712.
+Under Hamming distance, we could consider this dataset to have an embedding dimension of 50,000, but for Levenshtein distance, we state the dimensionality as the length of the longest sequence, i.e.\, 3,712.
 

### 2025-01-24 sections/5-discussion.tex
easy switch to go between arxiv and simods

@@
 We believe that some of this discrepancy can be explained by the fact that Repeated $\rho$-NN can significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search.
-This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity1}.
+This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
 This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.

### 2025-06-24 sections/2-methods.tex
formatting fixes

@@
     \begin{algorithmic} % enter the algorithmic environment
-        \Require $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
-        \Require $C$, a cluster
-        \Require $criteria$, user-specified continuation criteria
+    \REQUIRE $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
+    \REQUIRE $C$, a cluster
+    \REQUIRE $criteria$, user-specified continuation criteria
 
-        \State $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-        \State $c \Leftarrow$ geometric median of $seeds$
-        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-        \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-        \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+    \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+    \STATE $c \Leftarrow$ geometric median of $seeds$
+    \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+    \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+    \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+    \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
 
-        \If{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
-            \State Partition($L$, $criteria$)
-        \EndIf
+    \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
+        \STATE Partition($L$, $criteria$)
+    \ENDIF
 
-        \If{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
-            \State Partition($R$, $criteria$)
-        \EndIf
+    \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
+        \STATE Partition($R$, $criteria$)
+    \ENDIF
 \end{algorithmic}
@@
 
-\begin{figure}[H]
-    \vskip -0.2in
-    \centering
-    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
-    \caption{
-        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
-        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
-        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
-    }
-    \label{fig:methods:deltas}
-    \vskip -0.5in
-\end{figure}
+% \begin{figure}[H]
+%     \vskip -0.2in
+%     \centering
+%     \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
+%     \caption{
+%         {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
+%         ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
+%         ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+%         ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+%     }
+%     \label{fig:methods:deltas}
+%     \vskip -0.5in
+% \end{figure}
 
@@
         \begin{algorithmic}
-            \Require $f$, a distance function
-            \Require $C$, a cluster
-            \Require $q$, a query
-            \Require $r$, a search radius
-            \If{$\delta^+_C \leq r$}
-                \State \textbf{return} $\{C\}$
-            \Else
-                \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
-                \State \textbf{return} tree-search($L, q, r$) \\
+            \REQUIRE $f$, a distance function
+            \REQUIRE $C$, a cluster
+            \REQUIRE $q$, a query
+            \REQUIRE $r$, a search radius
+            \IF{$\delta^+_C \leq r$}
+                \STATE \textbf{return} $\{C\}$
+            \ELSE
+                \STATE $[L, R]$ $\Leftarrow$ \textit{children} of $C$
+                \STATE \textbf{return} tree-search($L, q, r$) \\
                 \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
-            \EndIf
+            \ENDIF
         \end{algorithmic}
@@
         \begin{algorithmic}
-            \Require $f$, a distance function
-            \Require $Q$, a set of clusters
-            \Require $q$, a query
-            \Require $r$, a search radius
-            \State $H \Leftarrow \emptyset$
-            \For{$C \in Q$}
-                \If{$\delta^+_C \leq r$}
-                    \State $H$ $\Leftarrow$ $H \cup C$
-                \Else
-                    \For{$p \in C$}
-                        \If{$f(p, q) \leq r$}
-                            \State $H$ $\Leftarrow$ $H \cup \{p\}$
-                        \EndIf
-                    \EndFor
-                \EndIf
-            \EndFor
-            \State \textbf{return} $H$
+            \REQUIRE $f$, a distance function
+            \REQUIRE $Q$, a set of clusters
+            \REQUIRE $q$, a query
+            \REQUIRE $r$, a search radius
+            \STATE $H \Leftarrow \emptyset$
+            \FOR{$C \in Q$}
+                \IF{$\delta^+_C \leq r$}
+                    \STATE $H$ $\Leftarrow$ $H \cup C$
+                \ELSE
+                    \FOR{$p \in C$}
+                        \IF{$f(p, q) \leq r$}
+                            \STATE $H$ $\Leftarrow$ $H \cup \{p\}$
+                        \ENDIF
+                    \ENDFOR
+                \ENDIF
+            \ENDFOR
+            \STATE \textbf{return} $H$
         \end{algorithmic}
@@
     \begin{algorithmic} % enter the algorithmic environment
-        \Require $\mathcal{R}$, the root cluster
-        \Require $q$, a query
-        \Require $k$, the number of neighbors to find
-        \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
-        \State $r \Leftarrow radius$ of $\mathcal{R}$
-        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
-        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \While{$\sum_{C \in Q} |C| < k$}
-            \If{$Q = \emptyset$}
-                \State $r \Leftarrow 2 \cdot r$
-            \ElsIf{$\sum_{C \in Q} |C| >= k$}
-                \State $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
-                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
-            \EndIf
-            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \EndWhile
-        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
-        \State \textbf{return} $H$ as a list
+        \REQUIRE $\mathcal{R}$, the root cluster
+        \REQUIRE $q$, a query
+        \REQUIRE $k$, the number of neighbors to find
+        \STATE{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
+        \STATE $r \Leftarrow radius$ of $\mathcal{R}$
+        \STATE $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
+        \STATE $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
+        \WHILE{$\sum_{C \in Q} |C| < k$}
+            \IF{$Q = \emptyset$}
+                \STATE $r \Leftarrow 2 \cdot r$
+            \ELSIF{$\sum_{C \in Q} |C| >= k$}
+                \STATE $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
+                \STATE $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
+            \ENDIF
+            \STATE $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
+        \ENDWHILE
+        \STATE $H \Leftarrow \text{leaf-search}(Q, q, r)$
+        \STATE \textbf{return} $H$ as a list
     \end{algorithmic}
@@
 
-\begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
+% \begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
 Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
-        \mathcal{O}
-        \Bigg(
-            \underbrace{
-                \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-            }_{\textrm{tree-search}}
-            \ + \
-            \underbrace{
-                \overbrace{k}^{\textrm{output size}} \cdot
-                \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
-            }_{\textrm{leaf-search}}
-        \Bigg)
-        \label{eq:methods:repeated-rnn-complexity}
-    \end{gather}
-    where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-    \label{thm:methods:rnn-complexity}
-\end{theorem}
+%         \mathcal{O}
+%         \Bigg(
+%             \underbrace{
+%                 \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+%             }_{\textrm{tree-search}}
+%             \ + \
+%             \underbrace{
+%                 \overbrace{k}^{\textrm{output size}} \cdot
+%                 \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
+%             }_{\textrm{leaf-search}}
+%         \Bigg)
+%         \label{eq:methods:repeated-rnn-complexity}
+%     \end{gather}
+%     where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
+%     \label{thm:methods:rnn-complexity}
+% \end{theorem}
 
@@
 
-\begin{minipage}{0.5125\textwidth}
-    \begin{algorithm}[H]\small
+    \begin{algorithm}
         \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
@@
         \begin{algorithmic} % enter the algorithmic environment
-            \Require $\mathcal{R}$, the root cluster
-            \Require $q$, a query
-            \Require $k$, the number of neighbors to find
-            \State $c \Leftarrow$ \textit{center} of $\mathcal{R}$
-            \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
-            \While{$\sum_{(\_, \_, m) \in Q} m \neq k$}
-                \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
-                \State $Q^{'} \Leftarrow \emptyset$
-                \For{$(C, \_, \_) \in Q$}
-                    \If{$\delta^{-}_{C} \leq \tau$}
-                        \If{$C$ is a point}
-                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
-                        \ElsIf{$C$ is a leaf}
-                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
-                        \Else
-                            \State $[L, R] \Leftarrow$ children of $C$
-                            \State $l, r \Leftarrow$ centers of $L, R$
-                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
-                            \State $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
-                        \EndIf
-                    \EndIf
-                \EndFor
-                \State $Q \Leftarrow Q^{'}$
-            \EndWhile
-            \State QuickSelect($Q$, $k$)
-            \State \textbf{return} The first $k$ points in $Q$
+            \REQUIRE $\mathcal{R}$, the root cluster
+            \REQUIRE $q$, a query
+            \REQUIRE $k$, the number of neighbors to find
+            \STATE $c \Leftarrow$ \textit{center} of $\mathcal{R}$
+            \STATE $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
+            \WHILE{$\sum_{(\_, \_, m) \in Q} m \neq k$}
+                \STATE $\tau \Leftarrow$ QuickSelect($Q$, $k$)
+                \STATE $Q^{'} \Leftarrow \emptyset$
+                \FOR{$(C, \_, \_) \in Q$}
+                    \IF{$\delta^{-}_{C} \leq \tau$}
+                        \IF{$C$ is a point}
+                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
+                        \ELSIF{$C$ is a leaf}
+                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
+                        \ELSE
+                            \STATE $[L, R] \Leftarrow$ children of $C$
+                            \STATE $l, r \Leftarrow$ centers of $L, R$
+                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
+                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
+                        \ENDIF
+                    \ENDIF
+                \ENDFOR
+                \STATE $Q \Leftarrow Q^{'}$
+            \ENDWHILE
+            \STATE QuickSelect($Q$, $k$)
+            \STATE \textbf{return} The first $k$ points in $Q$
         \end{algorithmic}
     \end{algorithm}
-\end{minipage}
-\hfill
-\begin{minipage}{0.425\textwidth}
-    \begin{algorithm}[H]\small
+
+    \begin{algorithm}
         \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
@@
         \begin{algorithmic}
-            \Require $\mathcal{R}$, the root cluster
-            \Require $q$, a query
-            \Require $k$, the number of neighbors to find
-            \State{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
-            \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
-            \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
-                \While{$Q.peek$ is not a leaf}
-                    \State{$C \Leftarrow Q.pop$, the closest cluster}
-                    \State{$[L, R] \Leftarrow$ children of $C$}
-                    \State{$Q.push(L)$}
-                    \State{$Q.push(R)$}
-                \EndWhile
-                \State{$leaf \Leftarrow Q.pop$}
-                \For{$p \in leaf$}
-                    \State{$H.push(p)$}
-                \EndFor
-            \EndWhile
-            \State \textbf{return} $H$
+            \REQUIRE $\mathcal{R}$, the root cluster
+            \REQUIRE $q$, a query
+            \REQUIRE $k$, the number of neighbors to find
+            \STATE{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
+            \STATE{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
+            \WHILE{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
+                \WHILE{$Q.peek$ is not a leaf}
+                    \STATE{$C \Leftarrow Q.pop$, the closest cluster}
+                    \STATE{$[L, R] \Leftarrow$ children of $C$}
+                    \STATE{$Q.push(L)$}
+                    \STATE{$Q.push(R)$}
+                \ENDWHILE
+                \STATE{$leaf \Leftarrow Q.pop$}
+                \FOR{$p \in leaf$}
+                    \STATE{$H.push(p)$}
+                \ENDFOR
+            \ENDWHILE
+            \STATE \textbf{return} $H$
         \end{algorithmic}
     \end{algorithm}
-\end{minipage}
 
@@
 
-\begin{theorem}
-    Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
+% \begin{theorem}
+%     Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
 Let $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$ and   $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of $X$, $d$ is the LFD of $X$, $\hat{|C|}$ is the mean cardinality of clusters overlapping the query ball, and $k$ is the number of nearest neighbors.
 Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
-        \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
-        \label{eq:methods:breadth-first-sieve-complexity}
-    \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
-        \mathcal{O} \Big( T \log T + L \log k \Big).
-        \label{eq:methods:depth-first-sieve-complexity}
-    \end{equation}
-    \label{thm:methods:sieve-complexity}
-\end{theorem}
+%         \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
+%         \label{eq:methods:breadth-first-sieve-complexity}
+%     \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
+%         \mathcal{O} \Big( T \log T + L \log k \Big).
+%         \label{eq:methods:depth-first-sieve-complexity}
+%     \end{equation}
+%     \label{thm:methods:sieve-complexity}
+% \end{theorem}
 

### 2025-06-24 sections/2-methods.tex
reorganizing and moving to supplement

@@
 
-\begin{algorithm} % enter the algorithm environment
-    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
-    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-    \REQUIRE $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
-    \REQUIRE $C$, a cluster
-    \REQUIRE $criteria$, user-specified continuation criteria
+% \begin{algorithm} % enter the algorithm environment
+%     \caption{Partition($C$, $criteria$)} % give the algorithm a caption
+%     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
+%     \begin{algorithmic} % enter the algorithmic environment
+%     \REQUIRE $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
+%     \REQUIRE $C$, a cluster
+%     \REQUIRE $criteria$, user-specified continuation criteria
 
-    \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-    \STATE $c \Leftarrow$ geometric median of $seeds$
-    \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-    \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-    \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-    \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+%     \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+%     \STATE $c \Leftarrow$ geometric median of $seeds$
+%     \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+%     \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+%     \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+%     \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
 
-    \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
-        \STATE Partition($L$, $criteria$)
-    \ENDIF
+%     \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
+%         \STATE Partition($L$, $criteria$)
+%     \ENDIF
 
-    \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
-        \STATE Partition($R$, $criteria$)
-    \ENDIF
-\end{algorithmic}
-\end{algorithm}
+%     \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
+%         \STATE Partition($R$, $criteria$)
+%     \ENDIF
+% \end{algorithmic}
+% \end{algorithm}
 
@@
 
-When building a tree for a real dataset, the depth would be larger than $\log n$, which would seem to increase the cost of tree-building.
-However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
-Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
-We do, however, explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
-In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.
+% When building a tree for a real dataset, the depth would be larger than $\log n$, which would seem to increase the cost of tree-building.
+% However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
+% Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
+% We do, however, explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
+% In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.
 
@@
 These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
-
-\begin{algorithm} % enter the algorithm environment
-    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
-    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-        \REQUIRE $\mathcal{R}$, the root cluster
-        \REQUIRE $q$, a query
-        \REQUIRE $k$, the number of neighbors to find
-        \STATE{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
-        \STATE $r \Leftarrow radius$ of $\mathcal{R}$
-        \STATE $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
-        \STATE $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \WHILE{$\sum_{C \in Q} |C| < k$}
-            \IF{$Q = \emptyset$}
-                \STATE $r \Leftarrow 2 \cdot r$
-            \ELSIF{$\sum_{C \in Q} |C| >= k$}
-                \STATE $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
-                \STATE $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
-            \ENDIF
-            \STATE $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
-        \ENDWHILE
-        \STATE $H \Leftarrow \text{leaf-search}(Q, q, r)$
-        \STATE \textbf{return} $H$ as a list
-    \end{algorithmic}
-\end{algorithm}
 
@@
 
-The work in~\cite{yu2015entropy} showed that $\sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d$, where $\gamma$ is a constant.
-By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k$.
+The work in~\cite{yu2015entropy} showed that 
+\begin{equation} \sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d \end{equation}, where $\gamma$ is a constant.
+By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k.$
 Thus, $\sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d$.
@@
 
-    \begin{algorithm}
-        \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
-        \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
-        \begin{algorithmic} % enter the algorithmic environment
-            \REQUIRE $\mathcal{R}$, the root cluster
-            \REQUIRE $q$, a query
-            \REQUIRE $k$, the number of neighbors to find
-            \STATE $c \Leftarrow$ \textit{center} of $\mathcal{R}$
-            \STATE $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
-            \WHILE{$\sum_{(\_, \_, m) \in Q} m \neq k$}
-                \STATE $\tau \Leftarrow$ QuickSelect($Q$, $k$)
-                \STATE $Q^{'} \Leftarrow \emptyset$
-                \FOR{$(C, \_, \_) \in Q$}
-                    \IF{$\delta^{-}_{C} \leq \tau$}
-                        \IF{$C$ is a point}
-                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
-                        \ELSIF{$C$ is a leaf}
-                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
-                        \ELSE
-                            \STATE $[L, R] \Leftarrow$ children of $C$
-                            \STATE $l, r \Leftarrow$ centers of $L, R$
-                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
-                            \STATE $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
-                        \ENDIF
-                    \ENDIF
-                \ENDFOR
-                \STATE $Q \Leftarrow Q^{'}$
-            \ENDWHILE
-            \STATE QuickSelect($Q$, $k$)
-            \STATE \textbf{return} The first $k$ points in $Q$
-        \end{algorithmic}
-    \end{algorithm}
-
-    \begin{algorithm}
-        \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
-        \label{alg:methods:depth-first-sieve}
-        \begin{algorithmic}
-            \REQUIRE $\mathcal{R}$, the root cluster
-            \REQUIRE $q$, a query
-            \REQUIRE $k$, the number of neighbors to find
-            \STATE{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
-            \STATE{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
-            \WHILE{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
-                \WHILE{$Q.peek$ is not a leaf}
-                    \STATE{$C \Leftarrow Q.pop$, the closest cluster}
-                    \STATE{$[L, R] \Leftarrow$ children of $C$}
-                    \STATE{$Q.push(L)$}
-                    \STATE{$Q.push(R)$}
-                \ENDWHILE
-                \STATE{$leaf \Leftarrow Q.pop$}
-                \FOR{$p \in leaf$}
-                    \STATE{$H.push(p)$}
-                \ENDFOR
-            \ENDWHILE
-            \STATE \textbf{return} $H$
-        \end{algorithmic}
-    \end{algorithm}
-
-\vskip 0.1in
+% \vskip 0.1in
 

### 2025-06-24 sections/3-datasets.tex
formatting fixes
reorganizing and moving to supplement

@@
 We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-Table~\ref{tab:datasets:summary} summarizes these datasets.
+Table~\ref{tab:datasets:summary} summarizes the cardinality and dimensionality of each these datasets.
 All benchmarks were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
@@
                 \hline
-                \textbf{Dataset} & \textbf{Distance Function}  &\textbf{Cardinality}  & \textbf{Dimensionality}  \\
+                \textbf{Dataset} & \textbf{Dist. Function}  &\textbf{Card}  & \textbf{Dim}  \\
                 \hline
@@
 
+\subsection{Local Fractal Dimension}
+\label{sec:dayasets:lfd-of-datasets}
+
+Since the CAKES algorithms were designed to scale in time with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
+Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
+In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
+
+For the Fashion-MNIST dataset we observe in Figure~\ref{fig:results:fashion-mnist-lfd} that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
+It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
+
+Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
+In particular, Glove-25's LFD is less than 3 for all depths.
+
+Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
+It increases sharply to a peak of 9 around a depth of 10, after which it decreases smoothly until reaching the deepest leaves in the tree.
+
+We contrast the low LFD of Sift with that of the Random dataset which has the same cardinality and dimension.
+As shown in Figure~\ref{fig:results:random-lfd}, the character of this dataset is very different from the others.
+The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
+The spread in LFD starts very small for the first few clusters and increases as depth increases.
+
+The Silva dataset, shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
+In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
+
+For the Radio-ML dataset, the LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
+Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
+Within each of the tree portions, this dataset has a character similar to that of the Random dataset.
+This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
+This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
+
 \subsection{Other Algorithms}

### 2025-06-24 sections/4-results.tex
reorganizing and moving to supplement

--- 2025-01-16 sections/4-results.tex
@@
 
-The algorithms presented in this paper rely heavily on the local fractal dimension (LFD) being much smaller than than the embedding dimension of the dataset.
-As such, we begin by examining our assumptions about the LFD of real-world datasets (see Section~\ref{sec:results:lfd-of-datasets} and Figure~\ref{fig:results:lfd-plots}).
+% The algorithms presented in this paper rely heavily on the local fractal dimension (LFD) being much smaller than than the embedding dimension of the dataset.
+% As such, we begin by examining our assumptions about the LFD of real-world datasets (see Section~\ref{sec:results:lfd-of-datasets} and Figure~\ref{fig:results:lfd-plots}).
 
-We continue by comparing the performance of the CAKES algorithms against HNSW, ANNOY and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift and Random datasets.
+We compare the performance of the CAKES algorithms against HNSW, ANNOY and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift and Random datasets.
 We also compare how this performance scales to the augmented versions of these datasets.
@@
 
-Finally, we test our intuitions about unbalanced clustering being better for search
-than balanced clustering (see Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} and Figure~\ref{fig:results:distance-counts}).
+% Finally, we test our intuitions about unbalanced clustering being better for search
+% than balanced clustering (see Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} and Figure~\ref{fig:results:distance-counts}).
 
-\subsection{Local Fractal Dimension of Datasets}
-\label{sec:results:lfd-of-datasets}
+% \subsection{Local Fractal Dimension of Datasets}
+% \label{sec:results:lfd-of-datasets}
 
-Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
-Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
-In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
+% Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
+% Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
+% In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
 % This is because our algorithms scale exponentially in the LFD, a
 
-The Fashion-MNIST dataset has an embedding dimension of 784 and uses the Euclidean distance metric.
-In Figure~\ref{fig:results:fashion-mnist-lfd} we observe that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
-It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
+% The Fashion-MNIST dataset has an embedding dimension of 784 and uses the Euclidean distance metric.
+% In Figure~\ref{fig:results:fashion-mnist-lfd} we observe that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
+% It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
 
-The Glove-25 dataset has an embedding dimension of 25 and uses the cosine distance function which, notably, is not a metric.
-Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
-All percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-In particular, Glove-25's LFD is less than 3 for all depths.
+% The Glove-25 dataset has an embedding dimension of 25 and uses the cosine distance function which, notably, is not a metric.
+% Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
+% All percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
+% In particular, Glove-25's LFD is less than 3 for all depths.
 
-The Sift dataset has an embedding dimension of 128 and uses the Euclidean distance metric.
-Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
-It increases sharply to a peak of 9 around a depth of 10.
-It then decreases smoothly until reaching the deepest leaves in the tree.
+% The Sift dataset has an embedding dimension of 128 and uses the Euclidean distance metric.
+% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
+% It increases sharply to a peak of 9 around a depth of 10.
+% It then decreases smoothly until reaching the deepest leaves in the tree.
 
-We generated the Random dataset to have the same cardinality and dimensionality as Sift.
-We used a uniform distribution in a 128-dimensional unit-hypercube to generate the points and the Euclidean metric to measure distances among them.
-Figure~\ref{fig:results:random-lfd} shows that the character of this dataset is significantly different from the others.
-The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
-The spread in LFD starts very small for the first few clusters and increases as depth increases.
-The LFD of approximately 20 for the root cluster $\mathcal{R}$ is what we expect for this random dataset.
-To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
-As a result, $\mathcal{R}$'s radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
-A consequence of this is that, with high probability, for every point in $\mathcal{R}$, its distance from $c$ is greater than $\tfrac{r}{2}$;
-in other words, $B(c, \tfrac{r}{2})$ contains only $c$ while $B(c, r)$ contains the entire dataset.
-Given our definition of LFD in Equation~\ref{eq:methods:lfd-half}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
-Theoretically, the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
-This reflects the difference between how we empirically measure the LFD and the value we would expect.
-With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^{128}$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
-Unfortunately, such a large sample is practically impossible to generate.
+% We generated the Random dataset to have the same cardinality and dimensionality as Sift.
+% We used a uniform distribution in a 128-dimensional unit-hypercube to generate the points and the Euclidean metric to measure distances among them.
+% Figure~\ref{fig:results:random-lfd} shows that the character of this dataset is significantly different from the others.
+% The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
+% The spread in LFD starts very small for the first few clusters and increases as depth increases.
+% The LFD of approximately 20 for the root cluster $\mathcal{R}$ is what we expect for this random dataset.
+% To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
+% As a result, $\mathcal{R}$'s radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
+% A consequence of this is that, with high probability, for every point in $\mathcal{R}$, its distance from $c$ is greater than $\tfrac{r}{2}$;
+% in other words, $B(c, \tfrac{r}{2})$ contains only $c$ while $B(c, r)$ contains the entire dataset.
+% Given our definition of LFD in Equation~\ref{eq:methods:lfd-half}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
+% Theoretically, the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
+% This reflects the difference between how we empirically measure the LFD and the value we would expect.
+% With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^{128}$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
+% Unfortunately, such a large sample is practically impossible to generate.
 
-The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712 but were provided in an MSA of width 50,000.
-As such the embedding dimension is somewhere between 3,712 and 50,000.
-We use the Levenshtein edit distance (a metric) to measure distances between sequences.
-This dataset, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
+% The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712.
+% As such the embedding dimension is 3,712, though as discussed previously, the embedding dimension would be 50,000 in a multiple sequence alignment.
+% We use the Levenshtein edit distance (a metric) to measure distances between sequences.
+% This dataset, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
+% In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
 
-The Radio-ML dataset consists of measurements of radio-frequency signals using 1,024 dimensional complex-valued vectors.
-We use the Dynamic Time Warping distance metric on this dataset.
-This dataset is synthetic~\cite{oshea2018radioml} but uses a far more elaborate generation process than our Random dataset.
-The LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
-Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
-Within each of the tree portions, this dataset has a character very similar to that of the Random dataset.
-This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
-This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
+% The Radio-ML dataset consists of measurements of radio-frequency signals using 1,024 dimensional complex-valued vectors.
+% We use the Dynamic Time Warping distance metric on this dataset.
+% This dataset is synthetic~\cite{oshea2018radioml} but uses a far more elaborate generation process than our Random dataset.
+% The LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
+% Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
+% Within each of the tree portions, this dataset has a character very similar to that of the Random dataset.
+% This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
+% This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
 
-\begin{figure}
-    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/lfd/fashion-mnist.png}\\
-        \subcaption{Fashion-MNIST}
-        \label{fig:results:fashion-mnist-lfd}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/lfd/glove-25.png}\\
-        \subcaption{Glove-25}
-        \label{fig:results:glove-25-lfd}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/lfd/sift.png}\\
-        \subcaption{Sift}
-        \label{fig:results:sift-lfd}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/lfd/random.png}\\
-        \subcaption{A random dataset}
-        \label{fig:results:random-lfd}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/lfd/silva-SSU-Ref.png}\\
-        \subcaption{Silva 18S}
-        \label{fig:results:silva-lfd}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/lfd/radio-ml.png}\\
-        \subcaption{RadioML}
-        \label{fig:results:radioml-lfd}
-    \end{subfigure}%
-    \\
-    \vskip -0.1in
-    \begin{subfigure}[b]{0.94\textwidth}
-        \centering
-        \includegraphics[width=0.7\textwidth]{images/lfd/legend.png}
-        \label{fig:results:lfd-legend}
-    \end{subfigure}%
-    \vskip -0.1in
-    \caption{Local fractal dimension vs.
+% \begin{figure}
+%     \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/lfd/fashion-mnist.png}\\
+%         \subcaption{Fashion-MNIST}
+%         \label{fig:results:fashion-mnist-lfd}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/lfd/glove-25.png}\\
+%         \subcaption{Glove-25}
+%         \label{fig:results:glove-25-lfd}
+%     \end{subfigure}
+%     \\
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/lfd/sift.png}\\
+%         \subcaption{Sift}
+%         \label{fig:results:sift-lfd}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/lfd/random.png}\\
+%         \subcaption{A random dataset}
+%         \label{fig:results:random-lfd}
+%     \end{subfigure}
+%     \\
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/lfd/silva-SSU-Ref.png}\\
+%         \subcaption{Silva 18S}
+%         \label{fig:results:silva-lfd}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/lfd/radio-ml.png}\\
+%         \subcaption{RadioML}
+%         \label{fig:results:radioml-lfd}
+%     \end{subfigure}%
+%     \\
+%     \vskip -0.1in
+%     \begin{subfigure}[b]{0.94\textwidth}
+%         \centering
+%         \includegraphics[width=0.7\textwidth]{images/lfd/legend.png}
+%         \label{fig:results:lfd-legend}
+%     \end{subfigure}%
+%     \vskip -0.1in
+%     \caption{Local fractal dimension vs.
 cluster depth across six datasets.
-The last dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset.
+The `random' dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset.
 In each plot, the horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
@@
 For example, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.}
-    \label{fig:results:lfd-plots}
-    \vskip -0.4in
-\end{figure}
+%     \label{fig:results:lfd-plots}
+%     \vskip -0.4in
+% \end{figure}
 
@@
 
-\begin{figure}
-    \captionsetup[subfigure]{aboveskip=-5pt,belowskip=-0.5pt}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
-        \subcaption{Fashion-MNIST for $k=10$.}
-        \label{fig:results:fashion-mnist-scaling}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
-        \subcaption{Glove-25 for $k=10$.}
-        \label{fig:results:glove-25-scaling}
-    \end{subfigure}%
-    \\
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
-        \subcaption{Sift for $k=10$.}
-        \label{fig:results:sift-scaling}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
-        \subcaption{A random dataset for $k=10$.}
-        \label{fig:results:random-scaling}
-    \end{subfigure}%
-    \\
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
-        \subcaption{Silva for $k=10$.}
-        \label{fig:results:silva-scaling}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
-        \subcaption{RadioML for $k=10$ at SnR = 10dB.}
-        \label{fig:results:radioml-scaling}
-    \end{subfigure}%
-    \\
-    \begin{subfigure}[b]{0.94\textwidth}
-        \centering
-        \includegraphics[width=0.7\textwidth]{plots/legend.png}
-        \label{fig:results:scaling-legend}
-    \end{subfigure}%
-    \caption{Throughput across six datasets, including a randomly-generated dataset.
-    In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-    For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
-    For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
-    With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
-    Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
-    For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
-    HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
-    In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
-    \label{fig:results:scaling-plots}
-\end{figure}
+% \begin{figure}
+%     \captionsetup[subfigure]{aboveskip=-5pt,belowskip=-0.5pt}
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics{plots/fashion-mnist_PermutedBall_10_throughput.png}
+%         \subcaption{Fashion-MNIST for $k=10$.}
+%         \label{fig:results:fashion-mnist-scaling}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
+%         \subcaption{Glove-25 for $k=10$.}
+%         \label{fig:results:glove-25-scaling}
+%     \end{subfigure}%
+%     \\
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
+%         \subcaption{Sift for $k=10$.}
+%         \label{fig:results:sift-scaling}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
+%         \subcaption{A random dataset for $k=10$.}
+%         \label{fig:results:random-scaling}
+%     \end{subfigure}%
+%     \\
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+%         \subcaption{Silva for $k=10$.}
+%         \label{fig:results:silva-scaling}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
+%         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
+%         \label{fig:results:radioml-scaling}
+%     \end{subfigure}%
+%     \\
+%     \begin{subfigure}[b]{0.94\textwidth}
+%         \centering
+%         \includegraphics[width=0.7\textwidth]{plots/legend.png}
+%         \label{fig:results:scaling-legend}
+%     \end{subfigure}%
+%     \caption{Throughput across six datasets, including a randomly-generated dataset.
+%     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+%     For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
+%     For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
+%     With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
+%     Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
+%     For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
+%     HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
+%     In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
+%     \label{fig:results:scaling-plots}
+% \end{figure}
 
-\begin{table}
-    \caption{Throughput (QPS) and Recall on the Fashion-MNIST dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-    In contrast, CAKES has perfect recall on Fashion-MNIST, since it used with Euclidean distance, which is a metric.}
-    \label{tab:results:qps-and-recall-fmn}
-    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-        \hline
-        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-        \cline{1-9}
-        \hline
-        1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
-        2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
-        4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
-        8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
-        16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
-        32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
-        64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
-        128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
-        256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
-        512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
-        \hline
-    \end{tabular}
-    \vskip -0.2in
-\end{table}
+% \begin{table}
+%     \caption{Throughput (QPS) and Recall on the Fashion-MNIST dataset.
+%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+%     In contrast, CAKES has perfect recall on Fashion-MNIST, since it used with Euclidean distance, which is a metric.}
+%     \label{tab:results:qps-and-recall-fmn}
+%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+%         \hline
+%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+%         \cline{1-9}
+%         \hline
+%         1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
+%         2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
+%         4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
+%         8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
+%         16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
+%         32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
+%         64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
+%         128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
+%         256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
+%         512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
+%         \hline
+%     \end{tabular}
+%     \vskip -0.2in
+% \end{table}
 
-\begin{table}
-    \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-    In contrast, CAKES has near-perfect recall on Glove-25, since it used with cosine distance, which is not a metric.}
-    \label{tab:results:qps-and-recall-glove}
-    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-        \hline
-        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-        \cline{1-9}
-        \hline
-        1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
-        2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
-        4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
-        8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
-        16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
-        32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
-        64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
-        128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
-        256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
-        \hline
-    \end{tabular}
-    \vskip -0.2in
-\end{table}
+% \begin{table}
+%     \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
+%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+%     In contrast, CAKES has near-perfect recall on Glove-25, since it used with cosine distance, which is not a metric.}
+%     \label{tab:results:qps-and-recall-glove}
+%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+%         \hline
+%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+%         \cline{1-9}
+%         \hline
+%         1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
+%         2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
+%         4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
+%         8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
+%         16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
+%         32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
+%         64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
+%         128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
+%         256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
+%         \hline
+%     \end{tabular}
+%     \vskip -0.2in
+% \end{table}
 
-\begin{table}
-    \caption{Throughput (QPS) and Recall on the Sift dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-    In contrast, CAKES has perfect recall on Sift, since it is used with Euclidean distance, which is a metric.}
-    \label{tab:results:qps-and-recall-sift}
-    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-        \hline
-        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-        \cline{1-9}
-        \hline
-        1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
-        2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
-        4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
-        8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
-        16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
-        32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
-        64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
-        128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
-        \hline
-    \end{tabular}
-    \vskip -0.2in
-\end{table}
+% \begin{table}
+%     \caption{Throughput (QPS) and Recall on the Sift dataset.
+%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
+%     In contrast, CAKES has perfect recall on Sift, since it is used with Euclidean distance, which is a metric.}
+%     \label{tab:results:qps-and-recall-sift}
+%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+%         \hline
+%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+%         \cline{1-9}
+%         \hline
+%         1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
+%         2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
+%         4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
+%         8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
+%         16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
+%         32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
+%         64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
+%         128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
+%         \hline
+%     \end{tabular}
+%     \vskip -0.2in
+% \end{table}
 
-\begin{table}
-    \caption{Throughput (QPS) and Recall on the Random dataset.
-    A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.
+% \begin{table}
+%     \caption{Throughput (QPS) and Recall on the Random dataset.
+%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.
 In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
-    As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-    HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
-    }
-    \label{tab:results:qps-and-recall-random}
-    \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-        \hline
-        \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-        \cline{1-9}
-        \hline
-        1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
-        2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
-        4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
-        8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
-        16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
-        32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
-        \hline
-    \end{tabular}
-    \vskip -0.2in
-\end{table}
+%     As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
+%     HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
+%     }
+%     \label{tab:results:qps-and-recall-random}
+%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
+%         \hline
+%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
+%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+%         \cline{1-9}
+%         \hline
+%         1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
+%         2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
+%         4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
+%         8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
+%         16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
+%         32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
+%         \hline
+%     \end{tabular}
+%     \vskip -0.2in
+% \end{table}
 
-\subsection{Clustering Strategies and Number of Distance Computations}
-\label{sec:results:clustering-strategies-and-number-of-distance-computations}
-
-In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
-These strategies are the Cartesian product of balanced vs.
-unbalanced clustering, and the presence vs.
-absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
-To help with this analysis, we also added some instrumentation to our implementation of CAKES to count the number of distance computations performed during search.
-We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.
-
-These four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is referred to as Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset are shown in Figure~\ref{fig:results:distance-counts}.
-Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
-the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which shows results using PermutedBall, no longer do so under a balanced clustering.
-We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.
-
-\begin{figure}
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
-        \subcaption{Repeated $\rho$-NN}
-        \label{fig:results:fashion-mnist-counts-throughput}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
-        \subcaption{Repeated $\rho$-NN}
-        \label{fig:results:glove-25-counts-counts}
-    \end{subfigure}%
-    \\
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
-        \subcaption{Breadth First Sieve}
-        \label{fig:results:sift-counts-throughput}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
-        \subcaption{Breadth First Sieve}
-        \label{fig:results:random-counts-counts}
-    \end{subfigure}%
-    \\
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
-        \subcaption{Depth First Sieve}
-        \label{fig:results:silva-counts-throughput}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.5\textwidth}
-        \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
-        \subcaption{Depth First Sieve}
-        \label{fig:results:radioml-counts-counts}
-    \end{subfigure}%
-    \\
-    \begin{subfigure}[b]{0.94\textwidth}
-        \centering
-        \includegraphics[width=0.6\textwidth]{images/distance_counts/legend.png}
-        \label{fig:results:counts-legend}
-    \end{subfigure}%
-    \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
-    Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
-    The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
-    The x-axis represents increasing cardinality of the dataset.}
-    \label{fig:results:distance-counts}
-\end{figure}
+% \begin{figure}
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
+%         \subcaption{Repeated $\rho$-NN}
+%         \label{fig:results:fashion-mnist-counts-throughput}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
+%         \subcaption{Repeated $\rho$-NN}
+%         \label{fig:results:glove-25-counts-counts}
+%     \end{subfigure}%
+%     \\
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
+%         \subcaption{Breadth First Sieve}
+%         \label{fig:results:sift-counts-throughput}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
+%         \subcaption{Breadth First Sieve}
+%         \label{fig:results:random-counts-counts}
+%     \end{subfigure}%
+%     \\
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
+%         \subcaption{Depth First Sieve}
+%         \label{fig:results:silva-counts-throughput}
+%     \end{subfigure}%
+%     \begin{subfigure}[b]{0.5\textwidth}
+%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
+%         \subcaption{Depth First Sieve}
+%         \label{fig:results:radioml-counts-counts}
+%     \end{subfigure}%
+%     \\
+%     \begin{subfigure}[b]{0.94\textwidth}
+%         \centering
+%         \includegraphics[width=0.6\textwidth]{images/distance_counts/legend.png}
+%         \label{fig:results:counts-legend}
+%     \end{subfigure}%
+%     \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
+%     Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
+%     The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
+%     The x-axis represents increasing cardinality of the dataset.}
+%     \label{fig:results:distance-counts}
+% \end{figure}

### 2025-06-24 sections/5-discussion.tex
reorganizing and moving to supplement

@@
 
-When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{T}{\lceil d \rceil} + L)$, where $T$ is the time complexity of tree-search and $L$ is the time complexity of leaf-search.
-Even though Repeated $\rho$-NN has the least time cost (compared to $\mathcal{O}\left((T + L)\log{(T+L)}\right)$ for Breadth-First Sieve and $\mathcal{O}(T\log{T} + L\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
-We believe that some of this discrepancy can be explained by the fact that Repeated $\rho$-NN can significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search.
-This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
-This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
-For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD,
-the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase in (Equation)~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
-This suggests that rather than using the reciprocal of the harmonic mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
-We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly overestimates the correct radius $\rho_k$ and to improve upon the factor of radial increase in Equation~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
+% When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{T}{\lceil d \rceil} + L)$, where $T$ is the time complexity of tree-search and $L$ is the time complexity of leaf-search.
+% Even though Repeated $\rho$-NN has the least time cost (compared to $\mathcal{O}\left((T + L)\log{(T+L)}\right)$ for Breadth-First Sieve and $\mathcal{O}(T\log{T} + L\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
+% We believe that some of this discrepancy can be explained by the fact that Repeated $\rho$-NN can significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search.
+% This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
+% This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
+% For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD,
+% the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase in (Equation)~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
+% This suggests that rather than using the reciprocal of the harmonic mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
+% We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly overestimates the correct radius $\rho_k$ and to improve upon the factor of radial increase in Equation~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
 
@@
 
-We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
-Such an encoding would allow us to store the data in a compressed format, and to perform search on the compressed data without decompressing it in its entirety.
-This would allow us to perform fast search on datasets that are otherwise too large to fit in memory.
-An exploration of compression and compressed search is another avenue for future work.
+% We already did compression so it feels weird to have this in there:
+
+% We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
+% Such an encoding would allow us to store the data in a compressed format, and to perform search on the compressed data without decompressing it in its entirety.
+% This would allow us to perform fast search on datasets that are otherwise too large to fit in memory.
+% An exploration of compression and compressed search is another avenue for future work.
 

### 2025-06-24 supplement.tex
easy switch to go between arxiv and simods
reorganizing and moving to supplement

@@
 % SIAM Supplemental File Template
+\newif\ifarxiv
+\arxivtrue % swap this to true to build for the arxiv
+\ifarxiv
+\documentclass{article}
+\usepackage{arxiv}
+\else
 \documentclass[review,supplement,onefignum,onetabnum]{siamonline220329}
+\fi
 
@@
 % Optional PDF information
+\ifarxiv
+\else
 \ifpdf
@@
 \fi
+\fi
 
@@
 \section{Supplementary Methods}
+
+\subsection{Algorithm Details}
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
+    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
+    \begin{algorithmic} % enter the algorithmic environment
+    \Require $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
+    \Require $C$, a cluster
+    \Require $criteria$, user-specified continuation criteria
+
+    \State $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+    \State $c \Leftarrow$ geometric median of $seeds$
+    \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+    \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+    \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+    \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+
+    \If{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
+        \State Partition($L$, $criteria$)
+    \EndIf
+
+    \If{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
+        \State Partition($R$, $criteria$)
+    \EndIf
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm} % enter the algorithm environment
+    \caption{Repeated $\rho$-NN($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
+    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
+    \begin{algorithmic} % enter the algorithmic environment
+        \Require $\mathcal{R}$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
+        \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
+        \State $r \Leftarrow radius$ of $\mathcal{R}$
+        \State $r \Leftarrow$ $\frac{r}{|\mathcal{R}|}$
+        \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
+        \While{$\sum_{C \in Q} |C| < k$}
+            \If{$Q = \emptyset$}
+                \State $r \Leftarrow 2 \cdot r$
+            \ElsIf{$\sum_{C \in Q} |C| >= k$}
+                \State $\mu \Leftarrow \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$
+                \State $r \Leftarrow r \cdot \min \bigg( 2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \bigg)$
+            \EndIf
+            \State $Q \Leftarrow$ tree-search($\mathcal{R}$, $q$, $r$)
+        \EndWhile
+        \State $H \Leftarrow \text{leaf-search}(Q, q, r)$
+        \State \textbf{return} $H$ as a list
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{Breadth-First Sieve($\mathcal{R}$, $q$, $k$)} % give the algorithm a caption
+    \label{alg:methods:bredth-first-sieve} % and a label for \ref{} commands later in the document
+    \begin{algorithmic} % enter the algorithmic environment
+        \Require $\mathcal{R}$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
+        \State $c \Leftarrow$ \textit{center} of $\mathcal{R}$
+        \State $Q \Leftarrow$ \{ ($\mathcal{R}$, $\delta^{+}_{\mathcal{R}}$, $|\mathcal{R}| - 1$), ($c$, $\delta_{\mathcal{R}}$, 1) \}
+        \While{$\sum_{(\_, \_, m) \in Q} m \neq k$}
+            \State $\tau \Leftarrow$ QuickSelect($Q$, $k$)
+            \State $Q^{'} \Leftarrow \emptyset$
+            \For{$(C, \_, \_) \in Q$}
+                \If{$\delta^{-}_{C} \leq \tau$}
+                    \If{$C$ is a point}
+                        \State $Q^{'} \Leftarrow Q^{'} \cup \{ (C, \delta_{C}, 1) \}$
+                    \ElsIf{$C$ is a leaf}
+                        \State $Q^{'} \Leftarrow Q^{'} \cup \{ (p, \delta_{p}, 1)$ for $p \in C \}$
+                    \Else
+                        \State $[L, R] \Leftarrow$ children of $C$
+                        \State $l, r \Leftarrow$ centers of $L, R$
+                        \State $Q^{'} \Leftarrow Q^{'} \cup \{ (L, \delta^{+}_{L}, |L| - 1), (l, \delta_{L}, 1) \}$
+                        \State $Q^{'} \Leftarrow Q^{'} \cup \{ (R, \delta^{+}_{R}, |R| - 1), (l, \delta_{R}, 1) \}$
+                    \EndIf
+                \EndIf
+            \EndFor
+            \State $Q \Leftarrow Q^{'}$
+        \EndWhile
+        \State QuickSelect($Q$, $k$)
+        \State \textbf{return} The first $k$ points in $Q$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{Depth-First Sieve($\mathcal{R}$, $q$, $k$)}
+    \label{alg:methods:depth-first-sieve}
+    \begin{algorithmic}
+        \Require $\mathcal{R}$, the root cluster
+        \Require $q$, a query
+        \Require $k$, the number of neighbors to find
+        \State{$Q \Leftarrow [\mathcal{R}]$, a min-heap by $\delta^{-}$}
+        \State{$H \Leftarrow [\ ]$, a max-heap by $\delta$ of size $k$}
+        \While{$|H| < k$ \textbf{or} $H.peek.\delta \geq Q.peek.\delta^{-}$}
+            \While{$Q.peek$ is not a leaf}
+                \State{$C \Leftarrow Q.pop$, the closest cluster}
+                \State{$[L, R] \Leftarrow$ children of $C$}
+                \State{$Q.push(L)$}
+                \State{$Q.push(R)$}
+            \EndWhile
+            \State{$leaf \Leftarrow Q.pop$}
+            \For{$p \in leaf$}
+                \State{$H.push(p)$}
+            \EndFor
+        \EndWhile
+        \State \textbf{return} $H$
+    \end{algorithmic}
+\end{algorithm}
+
+\subsection{Algorithm Details}
+\label{sec:supplement:algorithm-details}
 
@@
 
+\subsection{Clustering Strategies and Number of Distance Computations}
+\label{sec:results:clustering-strategies-and-number-of-distance-computations}
+
+In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
+These strategies are the Cartesian product of balanced vs.
+unbalanced clustering, and the presence vs.
+absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
+To help with this analysis, we also added some instrumentation to our implementation of CAKES to count the number of distance computations performed during search.
+We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.
+
+These four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is referred to as Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset are shown in Figure~\ref{fig:results:distance-counts}.
+Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
+the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots}, which shows results using PermutedBall, no longer do so under a balanced clustering.
+We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.
+
 \bibliographystyle{siamplain}

### 2025-07-15 sections/1-introduction.tex
intro edits

--- 2025-01-16 sections/1-introduction.tex
@@
 
-Researchers are collecting data at an unprecedented rate.
-In many fields, datasets are growing exponentially, and this increase in the rate of data collection outpaces improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
-This indicates that the performance of computing systems will not keep pace with the growth of data.
-Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms to analyze large datasets.
+Researchers are collecting data at an unprecedented rate, with datasets in many fields growing exponentially, and outpacing improvements in computing performance predicted by Moore's Law~\cite{kahn2011future}.
+This ``Big Data explosion'' has created a need for algorithms that scale efficiently to large datasets.
 
-Examples of large datasets include genomic databases, time-series data such as radio frequency signals, and neural network embeddings.
+Examples of such datasets include genomic databases, time-series signals, and neural network embeddings.
 Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image} are a common source of neural network embeddings.
-Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with an aligned length of 50,000 letters.
-Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated signals of different modulation modes.
+Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} includes ribosomal DNA sequences from approximately 2.25 million genomes, reaching 50,000 letters when aligned.
+Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated radio frequency signals.
 
-Many researchers are interested in similarity search on these datasets.
-Similarity search enables a variety of applications, including recommendation~\cite{annoy} and classification systems~\cite{suyanto2022knnclassifier}.
-As the cardinalities and dimensionalities of datasets have grown, however, efficient and accurate similarity search has become challenging;
+Similarity search is a fundamental task on such datasets, enabling applications such as recommendation~\cite{annoy} and classification ~\cite{suyanto2022knnclassifier}.
+Yet as datasets grow in size and dimensionality, efficient and accurate similarity search becomes increasingly challenging;
 even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{malkov2016hnsw, johnson2019billion, annoy, aumuller2020ann}.
 
-Given some measure of similarity between data points, there are two common definitions of similarity search: $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
-$k$-NN search aims to find the $k$ most similar points to a query, while $\rho$-NN search aims to find all points within a similarity threshold $\rho$ of a query.
-Previous works have used the term \textit{approximate} search to refer to $\rho$-NN search, but in this paper, we reserve the term \textit{approximate} for algorithms that do not exhibit perfect recall.
-In contrast, an \textit{exact} search algorithm exhibits perfect recall.
+Given some measure of similarity between data points, two standard similarity search paradigms are $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
+The former retrieves the $k$ most similar points to a query, while the latter retrieves all points within a similarity threshold $\rho$ of a query.
+While prior work has used the term \textit{approximate} search to refer to $\rho$-NN, we use it more narrowly: an {approximate} search algorithm is any algorithm that yields less than perfect recall, whereas an \textit{exact} algorithm guarantees perfect recall.
 
-$k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
-Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets.
-While fast algorithms for $k$-NN search on large datasets exist, they are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
+$k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because their time complexity is linear in the dataset's cardinality.
+Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
 For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
-This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
-Further, there is evidence that distance functions that do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
-this suggests that approximate $k$-NN search could perform poorly in such contexts.
 
-This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
-We also compare CAKES against several current algorithms; namely, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-We further benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
-Finally, we compare real-world datasets to an artificially-generated dataset obeying simple statistical properties.
+For example, approximate methods may fail in majority-voting classifiers with small $k$ or poorly separated classes~\cite{zhang2022imbalanced}, and distance functions like cosine similarity, which violate the triangle inequality, have shown poor performance in biomedical applications~\cite{hu2016distance}.
+
+This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for \emph{exact} $k$-NN search.
+We benchmark CAKES against state-of-the-art methods--FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}--on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We also evaluate performance on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and on a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+To further contextualize results, we compare against a synthetic dataset with simple statistical properties.
 
@@
 
-Recent $k$-nearest neighbor search algorithms designed to scale with the exponential growth of data include Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
-However, some of these algorithms do not provide exact search (as defined in Section \ref{sec:introduction} above).
+Several algorithms have been developed to scale $k$-nearest neighbor search to large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).
 
-Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks~\cite{kleinberg2000navigation, boguna2009navigability} and skip lists.
-HNSW builds a multi-layered graph of the dataset.
-A query point and each data point are inserted into the graph and joined by an edge to the $M$ nearest nodes in the in the graph, where $M$ is a tunable parameter.
-The highest layer in which an element can be placed is determined randomly with an exponentially-decaying probability distribution.
-Search starts at the highest layer and descends to the lowest layer, greedily following a path of edges to the nearest node, until reaching the query point.
-To improve accuracy, the $efSearch$ hyperparameter can be changed to specify the number of closest nearest neighbors to the query vector to be found at each layer.
+Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networksand skip lists.
+HNSW builds a multi-layered small-world graph~\cite{kleinberg2000navigation, boguna2009navigability} of the dataset, with queries and elements placed at each layer according to an exponentially-decaying probability distribution.
+Each node connects to its $M$ nearest neighbors (a tunable parameter) in the in the graph, where $M$.
+Search proceeds top-down, using a greedy search at each layer.
+Accuracy is controlled by the $efSearch$ hyperparameter, which specifies how many candidates are explored during search.
 
-InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} clusters data into high-dimensional Voronoi cells, and whichever cell a query point falls into is then searched exhaustively, similarly to an early precursor to our work~\cite{yu2015entropy}.
-The number of cells used is governed by the $n_{list}$ parameter.
-Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
-To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$ that specifies the number of additional adjacent or nearby cells to search.
+InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches exhaustively only within cells near the query.
+The number of clusters is set via the $n_{list}$ parameter; accuracy is improved by probing additional neighboring cells, controlled by the $n_{probe}$ parameter.
+This approach resembles earlier entropy-scaling methods~\cite{yu2015entropy}.
 
-ANNOY~\cite{annoy} relies on random projection and tree building for approximate $k$-NN search.
-At each intermediate node of the tree, two points are randomly sampled from the space, and the hyperplane equidistant from them is chosen to divide the space into two subspaces.
-This process iterates to create a forest of trees, and the number of iterations is a tunable parameter.
-At search time, one can increase the number of trees to be searched to improve recall at the cost of speed.
+ANNOY~\cite{annoy} uses random projection and tree building. 
+At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.
+Multiple such trees form a forest; increasing the number of trees improves recall but increases search time.
 
@@
 
-The entropy-scaling search paradigm exploits the geometric and topological structure inherent in large datasets.
-Importantly, as suggested by their name, entropy-scaling search algorithms exhibit asymptotic complexity that scales with topological properties of the dataset, instead of its cardinality, when the dataset has a manifold structure.
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a hierarchical clustering approach.
-CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
-In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language.
+Entropy-scaling algorithms exploit the inherent structure of large datasets, achieving complexity that scales with topological properties—such as metric entropy and fractal dimension—rather than cardinality.
+This makes them especially suitable for manifold-structured data.
 
-We also provide a theoretical analysis of the time complexity of CAKES's algorithms in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended the original flat clustering approach to entropy-scaling $\rho$-NN search from a flat clustering approach~\cite{yu2015entropy} to a hierarchical one.
+CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, refined the clustering algorithm from CHESS.
+
+This paper introduces CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
+We also provide a theoretical analysis of their time complexity in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
 Given that CAKES's algorithms are intended to be used on datasets with a manifold structure, complexity analysis assuming a uniform distribution of data would be uninformative.
-As a result, we assume in our analyses that the dataset has a manifold structure, as quantified by low fractal dimension and metric entropy.
-For our purposes, it makes more sense to analyze the expected performance of CAKES's algorithms under these assumptions.
+As a result, our analysis assumes datasets with manifold structure—characterized by low fractal dimension and metric entropy—reflecting the conditions for which CAKES is designed.

### 2025-07-15 sections/1-introduction.tex
down to 10 pages

@@
 
-Examples of such datasets include genomic databases, time-series signals, and neural network embeddings.
-Large language models such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR} and LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image} are a common source of neural network embeddings.
-Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} includes ribosomal DNA sequences from approximately 2.25 million genomes, reaching 50,000 letters when aligned.
+Examples of such datasets include genomic databases, time-series signals, and neural network embeddings such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR}, LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image}.
+Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} includes ribosomal RNA sequences from approximately 2.25 million genomes, reaching 50,000 letters when aligned.
 Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated radio frequency signals.
@@
 Given some measure of similarity between data points, two standard similarity search paradigms are $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
-The former retrieves the $k$ most similar points to a query, while the latter retrieves all points within a similarity threshold $\rho$ of a query.
-While prior work has used the term \textit{approximate} search to refer to $\rho$-NN, we use it more narrowly: an {approximate} search algorithm is any algorithm that yields less than perfect recall, whereas an \textit{exact} algorithm guarantees perfect recall.
+The former retrieves the $k$ points closest to a query, while the latter retrieves all points within a similarity threshold $\rho$ of a query.
+While prior work has used the term \textit{approximate} search to refer to $\rho$-NN, we use it more narrowly: an {approximate} search algorithm is any algorithm that yields imperfect recall, whereas an \textit{exact} algorithm guarantees perfect recall.
 
-$k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because their time complexity is linear in the dataset's cardinality.
+$k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because due to linear time complexity.
 Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
-For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
-
-For example, approximate methods may fail in majority-voting classifiers with small $k$ or poorly separated classes~\cite{zhang2022imbalanced}, and distance functions like cosine similarity, which violate the triangle inequality, have shown poor performance in biomedical applications~\cite{hu2016distance}.
+In a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but be sensitive to local perturbations for smaller values of $k$.
 
@@
 
-Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networksand skip lists.
-HNSW builds a multi-layered small-world graph~\cite{kleinberg2000navigation, boguna2009navigability} of the dataset, with queries and elements placed at each layer according to an exponentially-decaying probability distribution.
-Each node connects to its $M$ nearest neighbors (a tunable parameter) in the in the graph, where $M$.
-Search proceeds top-down, using a greedy search at each layer.
-Accuracy is controlled by the $efSearch$ hyperparameter, which specifies how many candidates are explored during search.
-
+Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networks and skip lists.
 InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches exhaustively only within cells near the query.
-The number of clusters is set via the $n_{list}$ parameter; accuracy is improved by probing additional neighboring cells, controlled by the $n_{probe}$ parameter.
-This approach resembles earlier entropy-scaling methods~\cite{yu2015entropy}.
-
 ANNOY~\cite{annoy} uses random projection and tree building.

### 2025-08-26 sections/2-methods.tex
partial methods edit

@@
 
-In this manuscript, we are primarily concerned with $k$-NN search in a finite-dimensional space.
-Given a dataset $\textbf{X} = \{x_1 \dots x_n\}$ of cardinality $|\textbf{X}| = n$, we define a \textit{point} or \textit{datum} $x_i \in \textbf{X}$ as a singular observation.
-Examples include the neural-network embedding of an image, the genome of an organism, and a measurement of a radio frequency signal.
+In this manuscript, we study $k$-NN search in finite-dimensional spaces.
+Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$, we refer to each $x_i \in \textbf{X}$ as a point or datum.
+Examples include image embeddings, genomic sequences, and time-series measurements.
 
 We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
-A distance value of zero defines an identity among points (i.e.,\,$f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
-We also require that the distance function be symmetric (i.e.,\,$f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
-In addition to these constraints, if the distance function obeys the triangle inequality (i.e.,\,$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
+We require $f$ to be symmetric and identity-preserving ($f(x, y) = 0 \iff x = y$). 
+If $f$ satisfies the triangle inequality (i.e.,\,$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
+Common examples include Euclidean, Levenshtein~\cite{levenshtein1966binary}, and DTW~\cite{muller2007dynamic}; cosine distance, which violates the triangle inequality, is not a metric.
 Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
-For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g.,\,consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
 
-The choice of an appropriate distance function varies by dataset and domain.
-For example, with neural-network embeddings, one could use Euclidean or Cosine distance.
-With genomic or proteomic sequence data, Levenshtein and Hamming distances are useful.
-With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.
+% cosine distance is not a metric because it violates the triangle inequality (e.g.,\,consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
+
+The appropriate distance function depends on the data: Euclidean or cosine is often used for neural embeddings, Levenshtein or Hamming for sequences, and DTW or Wasserstein for time-series data.
 
 CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
-We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
-In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
-While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
-they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
-CAKES exploits the low LFD of such datasets to accelerate search.
-We define LFD at some length scale around a point in the dataset as:
+We define the \emph{local fractal dimension} (LFD) to quantify this property.
+
+For a point $q \in \textbf{X}$ and radii $r_1 > r_2$,
+
 \begin{equation}
@@
 \end{equation}
-where $B(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
+where $B(q, r)$ is the set of points in the metric ball of radius $r$ centered at $q$.
 
-We use a simplified version of Equation~\ref{eq:methods:lfd-original} by using a length scale where $r_1 = 2 \cdot r_2$.
+We typically use a simplified version of Equation~\ref{eq:methods:lfd-original} with $r_1 = 2 \cdot r_2$:
 \begin{equation}
-    \text{LFD}(q, r) = \text{log}_2 \left( \frac{|B(q, r)|}{|B(q, \frac{r}{2})|} \right)
+    \text{LFD}(q, r) = \text{log}_2 \left( \frac{|B(q, r)|}{|B(q, \frac{r}{2})|} \right).
     \label{eq:methods:lfd-half}
 \end{equation}
+
+% In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
+% While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
+% they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
+% CAKES exploits the low LFD of such datasets to accelerate search.
 

### 2025-08-26 sections/4-results.tex
shrunk plots

@@
 
-% \begin{figure}
-%     \captionsetup[subfigure]{aboveskip=-5pt,belowskip=-0.5pt}
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics{plots/fashion-mnist_PermutedBall_10_throughput.png}
-%         \subcaption{Fashion-MNIST for $k=10$.}
-%         \label{fig:results:fashion-mnist-scaling}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.95\textwidth]{plots/glove-25_PermutedBall_10_throughput.png}
-%         \subcaption{Glove-25 for $k=10$.}
-%         \label{fig:results:glove-25-scaling}
-%     \end{subfigure}%
-%     \\
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.95\textwidth]{plots/sift_PermutedBall_10_throughput.png}
-%         \subcaption{Sift for $k=10$.}
-%         \label{fig:results:sift-scaling}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.95\textwidth]{plots/random_PermutedBall_10_throughput.png}
-%         \subcaption{A random dataset for $k=10$.}
-%         \label{fig:results:random-scaling}
-%     \end{subfigure}%
-%     \\
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.95\textwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
-%         \subcaption{Silva for $k=10$.}
-%         \label{fig:results:silva-scaling}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.95\textwidth]{plots/radio-ml_Ball_10_throughput.png}
-%         \subcaption{RadioML for $k=10$ at SnR = 10dB.}
-%         \label{fig:results:radioml-scaling}
-%     \end{subfigure}%
-%     \\
-%     \begin{subfigure}[b]{0.94\textwidth}
-%         \centering
-%         \includegraphics[width=0.7\textwidth]{plots/legend.png}
-%         \label{fig:results:scaling-legend}
-%     \end{subfigure}%
-%     \caption{Throughput across six datasets, including a randomly-generated dataset.
-%     In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-%     For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
-%     For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
-%     With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
-%     Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
-%     For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
-%     HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
-%     In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
-%     \label{fig:results:scaling-plots}
-% \end{figure}
+\begin{figure}[t]
+    \centering
+    % Row 1
+    \subfloat[Fashion-MNIST for $k=10$.]{
+      \includegraphics[width=0.48\columnwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
+      \label{fig:results:fashion-mnist-scaling}
+    }
+    \subfloat[Glove-25 for $k=10$.]{
+      \includegraphics[width=0.48\columnwidth]{plots/glove-25_PermutedBall_10_throughput.png}
+      \label{fig:results:glove-25-scaling}
+    }
+
+    % Row 2
+    \subfloat[Sift for $k=10$.]{
+      \includegraphics[width=0.48\columnwidth]{plots/sift_PermutedBall_10_throughput.png}
+      \label{fig:results:sift-scaling}
+    }
+    \subfloat[Random dataset for $k=10$.]{
+      \includegraphics[width=0.48\columnwidth]{plots/random_PermutedBall_10_throughput.png}
+      \label{fig:results:random-scaling}
+    }
+
+    % Row 3
+    \subfloat[Silva for $k=10$.]{
+      \includegraphics[width=0.48\columnwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
+      \label{fig:results:silva-scaling}
+    }
+    \subfloat[RadioML for $k=10$ at SnR = 10dB.]{
+      \includegraphics[width=0.48\columnwidth]{plots/radio-ml_Ball_10_throughput.png}
+      \label{fig:results:radioml-scaling}
+    }
+
+    % Legend (as a plain graphic)
+    \vspace{2pt}
+    \includegraphics[width=1\columnwidth]{plots/legend.png}
+
+    \caption{Throughput across six datasets, including a randomly-generated dataset.
+        In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
+        For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
+        For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
+        With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
+        Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
+        For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
+        HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
+        In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
+    \label{fig:results:scaling-plots}
+  \end{figure}
 
@@
 %     \label{tab:results:qps-and-recall-fmn}
+%     \resizebox{\linewidth}{!}{%
 %     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
@@
 %     \end{tabular}
+% }
 %     \vskip -0.2in
@@
 %     \label{tab:results:qps-and-recall-glove}
+%     \resizebox{\linewidth}{!}{%
 %     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
@@
 %     \label{tab:results:qps-and-recall-sift}
+%     \resizebox{\linewidth}{!}{%
 %     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
@@
 %     \label{tab:results:qps-and-recall-random}
+%     \resizebox{\linewidth}{!}{%
 %     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
@@
 %     \end{tabular}
+%     }
 %     \vskip -0.2in

### 2025-08-26 sections/4-results.tex
tigtened results

@@
 
-We compare the performance of the CAKES algorithms against HNSW, ANNOY and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift and Random datasets.
-We also compare how this performance scales to the augmented versions of these datasets.
-We also report the performance of CAKES on the Silva and RadioML datasets (see Section~\ref{sec:results:scaling-behavior-and-recall}, Figure~\ref{fig:results:scaling-plots}, and Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random}).
+For each dataset, we benchmark using the distance function listed in Table~\ref{tab:datasets:summary}, and results in this section are based on k-NN with $k=10$.
+
+In Figure~\ref{fig:results:scaling-plots}, we show how the throughput of each of CAKES's algorithms scales with cardinality for each of the six datasets we examined.
+To do this, we synthetically augmented the Fashion-MNIST, Glove-25, Sift, and Random datasets to create larger datasets using the process described in Section~\ref{sec:methods:synthetic-data}.
+For Silva and RadioML, due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
+
+Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} compare the performance (throughput and recall) of the CAKES algorithms against state-of-the-art algorithms on the Fashion-MNIST, Glove-25, Sift and Random datasets.
+In particular, we examine performance of HNSW, ANNOY and FAISS-IVF on each of those datasets as well as synthetically augmented versions of the datasets to isolate the effect of dataset size on performance.
+We did not perform similar benchmarks on the Silva and RadioML datasets because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
+For HNSW, ANNOY, and FAISS-IVF, we allow for a hyper-parameter search to tune their index for maximum recall.
+For CAKES, we build the tree and use our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) to select the fastest algorithm for each dataset and cardinality.
+
+Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
 
@@
 
-\subsection{Indexing and Tuning Time}
-\label{sec:results:indexing-and-tuning-time}
+% \subsection{Indexing and Tuning Time}
+% \label{sec:results:indexing-and-tuning-time}
 
-For each of the ANN-benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyper-parameters for these indices to achieve the highest possible recall.
+% For each of the ANN-benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyper-parameters for these indices to achieve the highest possible recall.
 For the sake of brevity, these results are reported in the supplement.
 
-\subsection{Scaling Behavior and Recall}
-\label{sec:results:scaling-behavior-and-recall}
+% We benchmark the CAKES algorithms, na\"{i}ve linear search, HNSW, ANNOY, and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift, and Random datasets.
+% We augment these datasets with synthetic points to examine how performance scales with cardinality.
+% We also benchmark the CAKES algorithms on the Silva and RadioML datasets, and we subsample these datasets instead of augmenting them to examine how performance scales with cardinality.
 
-We benchmark the CAKES algorithms, na\"{i}ve linear search, HNSW, ANNOY, and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift, and Random datasets.
-We augment these datasets with synthetic points to examine how performance scales with cardinality.
-We also benchmark the CAKES algorithms on the Silva and RadioML datasets, and we subsample these datasets instead of augmenting them to examine how performance scales with cardinality.
-For each dataset, we use the corresponding distance function (see Table~\ref{tab:datasets:summary}).
-For HNSW, ANNOY, and FAISS-IVF, we allow for a hyper-parameter search to tune their index for maximum recall.
-For CAKES, we build the tree and use our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) to select the fastest algorithm for each dataset and cardinality.
-We show results for $k=10$ in the main text, but results for $k=100$ can be found in the Supplement.
-Figure~\ref{fig:results:scaling-plots} shows the scaling behavior of these algorithms on these datasets while Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} report the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-MNIST, Glove-25, Sift, and Random respectively.
+% On the Fashion-MNIST, Glove-25, and Sift datasets (in Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} respectively), we observe that as cardinality increases, the Depth-First Sieve algorithm is consistently the fastest CAKES algorithm with a throughput that is constant in the cardinality of the dataset.
+% The Breadth-First Sieve algorithm is the usually the second fastest, also with a nearly constant throughput across all cardinalities.
+% The Repeated $\rho$-NN algorithm, however, falls off in throughput as cardinality increases.
+% All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
+% In contrast, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities and have near constant throughput as cardinality increases, but their recall degrades quickly as cardinality increases.
+% FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize recall.
 
-On the Fashion-MNIST, Glove-25, and Sift datasets (in Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} respectively), we observe that as cardinality increases, the Depth-First Sieve algorithm is consistently the fastest CAKES algorithm with a throughput that is constant in the cardinality of the dataset.
-The Breadth-First Sieve algorithm is the usually the second fastest, also with a nearly constant throughput across all cardinalities.
-The Repeated $\rho$-NN algorithm, however, falls off in throughput as cardinality increases.
-All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
-In contrast, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities and have near constant throughput as cardinality increases, but their recall degrades quickly as cardinality increases.
-FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize recall.
+% On the Random dataset, HNSW and ANNOY are still the fastest algorithms but exhibit recall values near 0.
+% CAKES's algorithms show linearly decreasing throughput as cardinality increases and are also slower than na\"{i}ve linear search.
 
-On the Random dataset, HNSW and ANNOY are still the fastest algorithms but exhibit recall values near 0.
-CAKES's algorithms show linearly decreasing throughput as cardinality increases and are also slower than na\"{i}ve linear search.
-
-Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} show the throughput and recall of CAKES's algorithms at each augmented cardinality for Fashion-MNIST, Glove-25, Sift, and Random respectively.
-Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
-We used our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) for each new tree (i.e.,\,at each cardinality), and this approach was always able to select the fastest algorithm for each dataset at each cardinality.
-Since we also allow for tuning hyper-parameters for the other algorithms, and we allow for different sets of hyper-parameters at each cardinality, it is a fair comparison for these tables to only list the performance of the fastest CAKES algorithm.
-When reporting recall, we use $1.000*$ to denote that the recall is imperfect, but rounds to $1.000$.
-
-For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
-Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
-For the Silva dataset (see Figure~\ref{fig:results:silva-scaling}), we observe that throughput initially decreases linearly as cardinality increases, but levels off at higher cardinalities.
-Depth-First Sieve is, once again, the fastest CAKES algorithm.
-For Radio-ML (see Figure~\ref{fig:results:radioml-scaling}), however, we see that throughput declines nearly linearly with cardinality, and that the three CAKES algorithms are, again, slower that na\"{i}ve linear search.
-
-\begin{figure}[t]
+\begin{figure}[h]
     \centering
-    % Row 1
     \subfloat[Fashion-MNIST for $k=10$.]{
@@
 
-    % Row 2
     \subfloat[Sift for $k=10$.]{
@@
 
-    % Row 3
     \subfloat[Silva for $k=10$.]{
@@
 
-    % Legend (as a plain graphic)
     \vspace{2pt}
@@
     \caption{Throughput across six datasets, including a randomly-generated dataset.
-        In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
-        For Fashion-MNIST, Glove-25, and Sift, as cardinality increases, the CAKES algorithms become faster than linear search but the cardinality at which this occurs differs by dataset.
-        For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest, while for Sift, Repeated $\rho$-NN is the fastest for smaller cardinalities and Depth-First Sieve is the fastest for larger cardinalities.
-        With Silva, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it starts to level off at higher cardinalities.
-        Depth-First Sieve is consistently the fastest algorithm on the Silva dataset.
-        For Radio-ML and Random, we see that all three CAKES algorithms are slower than na\"{i}ve linear search, and that their throughput decreases linearly with cardinality.
-        HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets.
-        In particular, on the Random dataset, HNSW and ANNOY have near-zero recall.}
+    Each plot shows throughput (queries per second; higher is better) versus dataset cardinality.
+For Fashion-MNIST, Glove-25, and Sift, CAKES becomes faster than linear search as cardinality grows, with the crossover point varying by dataset.
+For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest.
+For Sift, Repeated $\rho$-NN is fastest at small cardinalities, while Depth-First Sieve is fastest at large cardinalities.
+For Silva, throughput for all algorithms initially decreases approximately linearly with cardinality and then levels off at higher cardinalities; Depth-First Sieve is consistently fastest.
+For Radio-ML and Random, all CAKES variants are slower than na"{i}ve linear search, and their throughput decreases linearly with cardinality.
+HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets; on Random, their recall is near zero.}
     \label{fig:results:scaling-plots}
@@
 
-% \begin{table}
-%     \caption{Throughput (QPS) and Recall on the Fashion-MNIST dataset.
-%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-%     In contrast, CAKES has perfect recall on Fashion-MNIST, since it used with Euclidean distance, which is a metric.}
-%     \label{tab:results:qps-and-recall-fmn}
-%     \resizebox{\linewidth}{!}{%
-%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-%         \hline
-%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-%         \cline{1-9}
-%         \hline
-%         1   & \num{1.33e4} & 0.954  & \num{2.19e3} & 0.950  & \num{2.01e3} & 1.000* & \num{3.46e3} & 1.000  \\\cline{1-9}
-%         2   & \num{1.38e4} & 0.803  & \num{2.12e3} & 0.927  & \num{9.39e2} & 1.000* & \num{3.68e3} & 1.000  \\\cline{1-9}
-%         4   & \num{1.66e4} & 0.681  & \num{2.04e3} & 0.898  & \num{4.61e2} & 0.997  & \num{3.44e3} & 1.000  \\\cline{1-9}
-%         8   & \num{1.68e4} & 0.525  & \num{1.93e3} & 0.857  & \num{2.26e2} & 0.995  & \num{3.30e3} & 1.000  \\\cline{1-9}
-%         16  & \num{1.87e4} & 0.494  & \num{1.84e3} & 0.862  & \num{1.17e2} & 0.991  & \num{3.34e3} & 1.000  \\\cline{1-9}
-%         32  & \num{1.56e4} & 0.542  & \num{1.85e3} & 0.775  & \num{5.91e1} & 0.985  & \num{2.96e3} & 1.000  \\\cline{1-9}
-%         64  & \num{1.50e4} & 0.378  & \num{1.78e3} & 0.677  & \num{2.61e1} & 0.968  & \num{3.25e3} & 1.000  \\\cline{1-9}
-%         128 & \num{1.49e4} & 0.357  & \num{1.66e3} & 0.538  & \num{1.33e1} & 0.964  & \num{2.96e3} & 1.000  \\\cline{1-9}
-%         256 & --           & --     & \num{1.60e3} & 0.592  & \num{6.65e0} & 0.962  & \num{2.79e3} & 1.000  \\\cline{1-9}
-%         512 & --           & --     & \num{1.83e3} & 0.581  & \num{3.56e0} & 0.949 & \num{2.84e3} & 1.000 \\
-%         \hline
-%     \end{tabular}
-% }
-%     \vskip -0.2in
-% \end{table}
+\textbf{General notes (applies to Tables \ref{tab:results:qps-and-recall-fmn},\ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}).}
+  Throughput is measured in queries per second.
+A recall value of $1.000*$ denotes imperfect recall that rounds to $1.000$.
+In each table, we observe that while recall for CAKES does \emph{not} degrade with cardinality, recall for
+  HNSW and ANNOY degrades with cardinality. CAKES exhibits perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
 
-% \begin{table}
-%     \caption{Throughput (QPS) and Recall on the Glove-25 dataset.
-%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-%     In contrast, CAKES has near-perfect recall on Glove-25, since it used with cosine distance, which is not a metric.}
-%     \label{tab:results:qps-and-recall-glove}
-%     \resizebox{\linewidth}{!}{%
-%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-%         \hline
-%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-%         \cline{1-9}
-%         \hline
-%         1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\\cline{1-9}
-%         2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\\cline{1-9}
-%         4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\\cline{1-9}
-%         8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\\cline{1-9}
-%         16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\\cline{1-9}
-%         32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\\cline{1-9}
-%         64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\\cline{1-9}
-%         128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\\cline{1-9}
-%         256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
-%         \hline
-%     \end{tabular}
-% }
-%     \vskip -0.2in
-% \end{table}
+\begin{table}[t]
+    \centering
+    \caption{Fashion-MNIST: throughput and recall. See general notes above.}
+    \label{tab:results:qps-and-recall-fmn}
+    \small
+    \setlength{\tabcolsep}{4pt}
+    \begin{adjustbox}{width=\columnwidth,center}
+    \begin{tabular}{@{} lcccccccc @{}}
+    \toprule
+    \textbf{Mult.} &
+    \multicolumn{2}{c}{\textbf{HNSW}} &
+    \multicolumn{2}{c}{\textbf{ANNOY}} &
+    \multicolumn{2}{c}{\textbf{FAISS-IVF}} &
+    \multicolumn{2}{c}{\textbf{CAKES}} \\
+    \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+    & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+    \midrule
+    1   & \num{1.33e4} & 0.954 & \num{2.19e3} & 0.950 & \num{2.01e3} & $1.000^{*}$ & \num{3.46e3} & 1.000 \\
+    2   & \num{1.38e4} & 0.803 & \num{2.12e3} & 0.927 & \num{9.39e2} & $1.000^{*}$ & \num{3.68e3} & 1.000 \\
+    4   & \num{1.66e4} & 0.681 & \num{2.04e3} & 0.898 & \num{4.61e2} & 0.997       & \num{3.44e3} & 1.000 \\
+    8   & \num{1.68e4} & 0.525 & \num{1.93e3} & 0.857 & \num{2.26e2} & 0.995       & \num{3.30e3} & 1.000 \\
+    16  & \num{1.87e4} & 0.494 & \num{1.84e3} & 0.862 & \num{1.17e2} & 0.991       & \num{3.34e3} & 1.000 \\
+    32  & \num{1.56e4} & 0.542 & \num{1.85e3} & 0.775 & \num{5.91e1} & 0.985       & \num{2.96e3} & 1.000 \\
+    64  & \num{1.50e4} & 0.378 & \num{1.78e3} & 0.677 & \num{2.61e1} & 0.968       & \num{3.25e3} & 1.000 \\
+    128 & \num{1.49e4} & 0.357 & \num{1.66e3} & 0.538 & \num{1.33e1} & 0.964       & \num{2.96e3} & 1.000 \\
+    256 & --           & --    & \num{1.60e3} & 0.592 & \num{6.65e0} & 0.962       & \num{2.79e3} & 1.000 \\
+    512 & --           & --    & \num{1.83e3} & 0.581 & \num{3.56e0} & 0.949       & \num{2.84e3} & 1.000 \\
+    \bottomrule
+    \end{tabular}
+    \end{adjustbox}
+    \end{table}
 
-% \begin{table}
-%     \caption{Throughput (QPS) and Recall on the Sift dataset.
-%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000. While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
-%     In contrast, CAKES has perfect recall on Sift, since it is used with Euclidean distance, which is a metric.}
-%     \label{tab:results:qps-and-recall-sift}
-%     \resizebox{\linewidth}{!}{%
-%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-%         \hline
-%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-%         \cline{1-9}
-%         \hline
-%         1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\\cline{1-9}
-%         2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\\cline{1-9}
-%         4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\\cline{1-9}
-%         8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\\cline{1-9}
-%         16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\\cline{1-9}
-%         32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\\cline{1-9}
-%         64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\\cline{1-9}
-%         128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
-%         \hline
-%     \end{tabular}
-% }
-%     \vskip -0.2in
-% \end{table}
+\begin{table}
+    \centering
+    \caption{Glove-25: throughput and recall. See general notes above.}
+    \label{tab:results:qps-and-recall-glove}
+    \small
+    \setlength{\tabcolsep}{4pt}
+    \begin{adjustbox}{width=\columnwidth,center}
+    \begin{tabular}{@{} lcccccccc @{}}
+        \toprule
+        \textbf{Mult.} &
+        \multicolumn{2}{c}{\textbf{HNSW}} &
+        \multicolumn{2}{c}{\textbf{ANNOY}} &
+        \multicolumn{2}{c}{\textbf{FAISS-IVF}} &
+        \multicolumn{2}{c}{\textbf{CAKES}} \\
+        \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \midrule
+        1   & \num{2.28e4} & 0.801 & \num{2.83e3} & 0.835 & \num{2.38e3} & 1.000* & \num{1.54e3} & 1.000* \\
+        2   & \num{2.38e4} & 0.607 & \num{2.70e3} & 0.832 & \num{1.19e3} & 1.000* & \num{1.49e3} & 1.000* \\
+        4   & \num{2.50e4} & 0.443 & \num{2.61e3} & 0.839 & \num{6.19e2} & 1.000* & \num{1.28e3} & 1.000* \\
+        8   & \num{2.78e4} & 0.294 & \num{2.51e3} & 0.834 & \num{3.03e2} & 1.000* & \num{1.30e3} & 1.000* \\
+        16  & \num{3.11e4} & 0.213 & \num{2.23e3} & 0.885 & \num{1.51e2} & 1.000* & \num{1.14e3} & 1.000* \\
+        32  & \num{3.24e4} & 0.178 & \num{2.01e3} & 0.764 & \num{7.40e1} & 0.999  & \num{1.05e3} & 1.000* \\
+        64  & --           & --    & \num{1.99e3} & 0.631 & \num{3.77e1} & 0.997  & \num{1.07e3} & 1.000* \\
+        128 & --           & --    & --           & --    & \num{1.90e1} & 0.998  & \num{8.92e2} & 1.000* \\
+        256 & --           & --    & --           & --    & \num{9.47e0} & 0.998  & \num{8.91e2} & 1.000* \\
+        \bottomrule
+    \end{tabular}
+    \end{adjustbox}
+    \end{table}
 
-% \begin{table}
-%     \caption{Throughput (QPS) and Recall on the Random dataset.
-%     A recall value of $1.000*$ denotes imperfect recall that rounds to 1.000.
-In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
-%     As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
-%     HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
-%     }
-%     \label{tab:results:qps-and-recall-random}
-%     \resizebox{\linewidth}{!}{%
-%     \begin{tabular}{|l|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|p{1.55cm}|p{1.1cm}|}
-%         \hline
-%         \multirow{2}{*}{\textbf{Mult.}} & \multicolumn{2}{c|}{\textbf{HNSW}} & \multicolumn{2}{c|}{\textbf{ANNOY}} & \multicolumn{2}{c|}{\textbf{FAISS-IVF}}  & \multicolumn{2}{c|}{\textbf{CAKES}} \\\cline{2-9}
-%         & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-%         \cline{1-9}
-%         \hline
-%         1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\\cline{1-9}
-%         2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\\cline{1-9}
-%         4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\\cline{1-9}
-%         8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\\cline{1-9}
-%         16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\\cline{1-9}
-%         32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
-%         \hline
-%     \end{tabular}
-%     }
-%     \vskip -0.2in
-% \end{table}
+\begin{table}
+    \centering
+    \caption{Sift: throughput and recall. See general notes above.}
+    \label{tab:results:qps-and-recall-sift}
+        \small
+        \setlength{\tabcolsep}{4pt}
+        \begin{adjustbox}{width=\columnwidth,center}
+        \begin{tabular}{@{} lcccccccc @{}}
+        \toprule
+        \textbf{Mult.} &
+        \multicolumn{2}{c}{\textbf{HNSW}} &
+        \multicolumn{2}{c}{\textbf{ANNOY}} &
+        \multicolumn{2}{c}{\textbf{FAISS-IVF}} &
+        \multicolumn{2}{c}{\textbf{CAKES}} \\
+        \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \midrule
+        1   & \num{1.93e4} & 0.782 & \num{3.98e3} & 0.686 & \num{6.98e2} & 1.000* & \num{6.20e2} & 1.000 \\
+        2   & \num{2.03e4} & 0.552 & \num{3.80e3} & 0.614 & \num{3.30e2} & 1.000* & \num{2.95e2} & 1.000 \\
+        4   & \num{2.18e4} & 0.394 & \num{3.69e3} & 0.637 & \num{1.65e2} & 1.000* & \num{1.76e2} & 1.000 \\
+        8   & \num{2.48e4} & 0.298 & \num{3.58e3} & 0.710 & \num{7.72e1} & 1.000* & \num{1.27e2} & 1.000 \\
+        16  & \num{2.68e4} & 0.210 & \num{3.50e3} & 0.690 & \num{3.98e1} & 1.000* & \num{1.47e2} & 1.000 \\
+        32  & \num{2.75e4} & 0.193 & \num{3.44e3} & 0.639 & \num{2.09e1} & 0.999  & \num{1.24e2} & 1.000 \\
+        64  & --           & --    & \num{3.39e3} & 0.678 & \num{8.87e0} & 0.997  & \num{1.34e2} & 1.000 \\
+        128 & --           & --    & \num{3.36e3} & 0.643 & \num{4.78e0} & 0.993  & \num{1.31e2} & 1.000 \\
+        \bottomrule
+    \end{tabular}
+    \end{adjustbox}
+    \end{table}
+
+\begin{table}
+    \caption{Random dataset: throughput and recall.
+    In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.CAKES exhibits perfect recall at all cardinalities, whereas HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
+    }
+    \label{tab:results:qps-and-recall-random}
+    \small
+    \setlength{\tabcolsep}{4pt}
+    \begin{adjustbox}{width=\columnwidth,center}
+    \begin{tabular}{@{} lcccccccc @{}}
+    \toprule
+    \textbf{Mult.} &
+    \multicolumn{2}{c}{\textbf{HNSW}} &
+    \multicolumn{2}{c}{\textbf{ANNOY}} &
+    \multicolumn{2}{c}{\textbf{FAISS-IVF}} &
+    \multicolumn{2}{c}{\textbf{CAKES}} \\
+    \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+        & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+        \midrule
+        1  & \num{1.17e4} & 0.060 & \num{4.28e3} & 0.028 & \num{7.342} & 1.000* & \num{6.06e2} & 1.000 \\
+        2  & \num{1.01e4} & 0.048 & \num{4.04e3} & 0.021 & \num{3.582} & 1.000* & \num{2.75e2} & 1.000 \\
+        4  & \num{9.12e3} & 0.031 & \num{3.64e3} & 0.014 & \num{1.902} & 1.000* & \num{1.35e2} & 1.000 \\
+        8  & \num{8.35e3} & 0.022 & \num{3.37e3} & 0.013 & \num{8.841} & 1.000* & \num{6.13e1} & 1.000 \\
+        16 & \num{8.25e3} & 0.008 & \num{3.17e3} & 0.006 & \num{4.361} & 1.000* & \num{2.82e1} & 1.000 \\
+        32 & --           & --    & \num{3.01e3} & 0.007 & \num{1.721} & 1.000* & \num{1.31e1} & 1.000 \\
+        \bottomrule
+    \end{tabular}
+    \end{adjustbox}
+    \end{table}
 

### 2025-08-26 sections/2-methods.tex
tightening discussion

@@
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
-Section~\ref{sec:results:lfd-of-datasets} discusses this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
+% Section~\ref{sec:results:lfd-of-datasets} discusses this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
 For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but potentially highly variable at larger length scales, i.e.,\,when $r$ is large.
@@
 
-We start by performing a divisive hierarchical clustering on the dataset using CLAM to obtain a cluster tree $\mathcal{T}$.
-The procedure is almost identical to that outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}).
-After building $\mathcal{T}$ with CLAM, we also perform a depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).
+\subsubsection{Building the Tree}
+\label{sec:methods:clustering:building-the-tree}
+
+We construct a divisive hierarchical cluster tree $\mathcal{T}$ with CLAM, following CHESS~\cite{ishaq2019clustered} but with improved pole selection, then apply a depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}).
+
+The recursive step is as follows.
+For any cluster $C$ with $|C|$ points, we draw a random subsample $S\subset C$ of $\sqrt{|C|}$ points, compute all pairwise distances in $S$, and take the geometric median of $S$—the point minimizing the sum of distances to all other points in $S$—as the \emph{center} of $C$.
+The \emph{radius} of $C$ is the maximum distance from this center to any point in $C$.
+
+Let the \emph{left pole} be the point in $C$ farthest from the center; let the \emph{right pole} be the point farthest from the left pole.
+We partition $C$ into a \emph{left child} (points closer to the left pole) and a \emph{right child} (points closer to the right pole); ties go to the left child.
+Starting from the root cluster $\mathcal{R}$ containing the entire dataset, we recurse until each leaf is a single datum or any user-specified stopping criterion is met (e.g., minimum cluster radius, minimum cardinality, or maximum tree depth).
+During partitioning, we compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
 
@@
 
-Given a cluster $C$ with $|C|$ points, we define its two children by the following process.
-We take a random subsample $S$ of $\sqrt{|C|}$ of $C$'s points, and compute pairwise distances between all points in $S$.
-Using these distances, we compute the \textit{geometric median} of $S$; in other words, we find the point that minimizes the sum of distances to all other points in $S$.
-We define the \textit{center} of $C$ to be this geometric median.
-
-The \textit{radius} of $C$ is the maximum distance from the center to any other point in $C$.
-The point that is responsible for that radius (i.e.,\,the furthest point from the center) is designated the \textit{left pole} and the point that is furthest from the left pole is designated the \textit{right pole}.
-We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster that are closer to the left pole than to the right pole, and the right child contains all points in the cluster that are closer to the right pole than to the left pole.
-Without loss of generality, we assign to the left child those points that are equidistant from the two poles.
-Starting from a root cluster $\mathcal{R}$ containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria (e.g., minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc).
-This process is described in Algorithm \ref{alg:methods:partition}.
-During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-half}.
-
 \subsubsection{Depth-First Reordering}
@@
 
-This complexity analysis ignores the cost of checking the continuation criteria.
-While a user could specify criteria of any complexity (our implementation allows for this), for this paper we continue clustering until each cluster is a singleton, i.e.,\,it contains only one point or duplicates of the same point and has a radius of zero.
-This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
-
+% This complexity analysis ignores the cost of checking the continuation criteria.
+% While a user could specify criteria of any complexity (our implementation allows for this), for this paper we continue clustering until each cluster is a singleton, i.e.,\,it contains only one point or duplicates of the same point and has a radius of zero.
+% This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
 We also note here that this $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
-In practice, for real datasets, we expect Algorithm~\ref{alg:methods:partition} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
+In practice, for real datasets, we expect the process from Section~\ref{sec:methods:clustering:building-the-tree} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
 The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
@@
 
-% \begin{figure}[H]
-%     \vskip -0.2in
-%     \centering
-%     \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
-%     \caption{
-%         {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
-%         ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
-%         ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-%         ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
-%     }
-%     \label{fig:methods:deltas}
-%     \vskip -0.5in
-% \end{figure}
+\begin{figure}[H]
+    \vskip -0.2in
+    \centering
+    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
+    \caption{
+        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
+        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
+        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+    }
+    \label{fig:methods:deltas}
+    \vskip -0.5in
+\end{figure}
 
@@
                 \STATE \textbf{return} tree-search($L, q, r$) \\
-                \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
+                 $\cup$ tree-search($R, q, r$)
             \ENDIF
@@
 
-Given a dataset $\textbf{X}$, a distance function $f$ and the root cluster $\mathcal{R}$ of the tree $\mathcal{T}$ (constructed by Algorithm~\ref{alg:methods:partition}), we can now formally pose the $k$-NN search problem.
-Given a user specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
+Let $\textbf{X}$ be a dataset, $f$ a distance function, and $\mathcal{R}$ the root cluster of a tree $\mathcal{T}$ constructed by the process in Section~\ref{sec:methods:clustering:building-the-tree}.
+Given a user-specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
 In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
@@
 
-For Repeated $\rho$-NN search (Algorithm~\ref{alg:methods:repeated-rnn}), we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
+For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
 If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
@@
 
-\begin{proof} We consider the \textit{tree-search} and \textit{leaf-search} stages of search separately.
+We consider the \textit{tree-search} and \textit{leaf-search} stages of search separately.
 Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors by doing repeated iterations of the  $\rho$-NN search Algorithm~\ref{alg:methods:rnn-search:tree-search}.
@@
 Again using the assumption that since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
-By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has time complexity as claimed in Equation~\ref{eq:methods:repeated-rnn-complexity}.
-\end{proof}
+By combining the bounds for tree-search and leaf-search, we see that time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
+            \mathcal{O}
+            \Bigg(
+                \underbrace{
+                    \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+                }_{\textrm{tree-search}}
+                \ + \
+                \underbrace{
+                    \overbrace{k}^{\textrm{output size}} \cdot
+                    \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
+                }_{\textrm{leaf-search}}
+            \Bigg)
+            \label{eq:methods:repeated-rnn-complexity}
+        \end{gather}
+        where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
 
@@
 
-This algorithm (described in Algorithm~\ref{alg:methods:bredth-first-sieve}) performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+This algorithm performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
 
@@
 
-This algorithm (described in Algorithm~\ref{alg:methods:depth-first-sieve}) is quite similar to a depth-first traversal of the tree $\mathcal{T}$, using two priority queues to track clusters and hits, and to prioritize which branch of $\mathcal{T}$ to explore next.
+This algorithm is quite similar to a depth-first traversal of the tree $\mathcal{T}$, using two priority queues to track clusters and hits, and to prioritize which branch of $\mathcal{T}$ to explore next.
 
@@
 
-\begin{proof}
 Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
@@
 The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is as claimed in Equation~\ref{eq:methods:breadth-first-sieve-complexity}.
-
-For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is as claimed in Equation~\ref{eq:methods:depth-first-sieve-complexity}.
-\end{proof}
+Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity of Breadth-First Sieve is $\mathcal{O} \Big( (T + L ) \log (T + L ) \Big) \label{eq:methods:breadth-first-sieve-complexity}$
+and the the time complexity of Depth-First Sieve search is $ \mathcal{O} \Big( T \log T + L \log k \Big).
+            \label{eq:methods:depth-first-sieve-complexity} $
 
@@
 
-Based on our asymptotic complexity analyses, we expect CAKES to perform well on datasets with low LFD, and for its performance to scale sub-linearly with the cardinality of the dataset.
-To test this hypothesis, we use some datasets from the ANN-benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to generate similar datasets with exponentially larger cardinalities.
-We do the same with a large random dataset of uniformly distributed points in a hypercube.
-We then compare the performance of CAKES to that of other algorithms on the original datasets and the synthetically augmented datasets.
+Based on our asymptotic analyses, we expect CAKES to perform well on datasets with low LFD and to scale sublinearly with dataset cardinality.
+To test this, we use datasets from the ANN-Benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to produce similar datasets with exponentially larger cardinalities.
+We perform the same procedure on a large dataset of uniformly distributed points in a hypercube, and compare CAKES with other algorithms on both the original and augmented datasets.
 
-To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
-Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level, and $m$ be a user-specified integer multiplier.
-For each datum $\mathbf{x} \in X$, we create $m - 1$ new data points within a distance $\epsilon$ of $\mathbf{x}$.
-We construct a random vector $\mathbf{r}$ of $d$ dimensions in the hyper-sphere of radius $\epsilon$ centered at the origin.
-We then add $\mathbf{r}$ to $\mathbf{x}$ to get a new point $\mathbf{x}'$.
-Since $||\mathbf{r}|| \leq \epsilon$, we have that $||\mathbf{x} - \mathbf{x}'|| \leq \epsilon$ (i.e.,\,$\mathbf{x}'$ is within a distance $\epsilon$ of $\mathbf{x}$).
-This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.
-This augmentation process does not add to the overall topological structure of the dataset, but it does increase its cardinality by a factor of $m$.
-This allows us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.
+Our augmentation procedure is as follows.
+Let $X$ be a dataset of dimension $d$, $\epsilon$ a user-specified noise level, and $m$ an integer multiplier.
+For each $\mathbf{x}\in X$, generate $m-1$ new points within distance $\epsilon$ of $\mathbf{x}$ by sampling a random vector $\mathbf{r}\in\mathbb{R}^d$ from the $\epsilon$-ball (hypersphere of radius $\epsilon$ centered at the origin) and setting $\mathbf{x}'=\mathbf{x}+\mathbf{r}$.
+Since $\lVert\mathbf{r}\rVert\le\epsilon$, we have $\lVert\mathbf{x}-\mathbf{x}'\rVert\le\epsilon$.
+The result is $X'$ with $\lvert X'\rvert=m\,\lvert X\rvert$.
+This increases cardinality by a factor of $m$ without altering the dataset’s overall topological structure, isolating the effect of cardinality from dimensionality, metric choice, and topology.

### 2025-08-26 sections/3-datasets.tex
version where we don't discuss lfd
tightening discussion

--- 2025-01-16 sections/3-datasets.tex
@@
 
-In addition to benchmarks on datasets from the ANN-Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{sec:methods:synthetic-data}.
-In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Mult.'' in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random}) increases.
-We also benchmarked on purely randomly-generated datasets of various cardinalities.
-For this, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset; hereafter, we refer to the random dataset with cardinality 1,000,000 and dimensionality 128 as ``Random.''
-This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of the CAKES' algorithms.
+In addition to ANN-Benchmarks datasets, we evaluate synthetic augmentations generated as in Sec.~\ref{sec:methods:synthetic-data}, using a noise tolerance $\epsilon=0.01$, and study scaling as the cardinality multiplier (“Mult.” in Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random}) increases.
+We also benchmark purely random datasets at various cardinalities; to match Sift we use a random dataset with base cardinality of $1{,}000{,}000$ and dimensionality $128$, and refer to this dataset as “Random.” This benchmark isolates the effect of manifold structure—expected to be absent in purely random data—on CAKES algorithms’ performance.
 
-In order to calculate recall on the augmented datasets, we perform linear search in Rust and save the results to disk.
-We verified that linear search on the original dataset produced neighbors with perfect recall compared to those provided by the ANN-Benchmarks suite.
-We then use the results of linear search on the augmented datasets as ground truth for calculating recall of CAKES's algorithms in Rust and read them in Python for calculating recall of HNSW, ANNOY, and FAISS-IVF.
+To compute recall on the augmented datasets, we run linear search in Rust and write the results to disk.
+We verified on the ANN-Benchmark datasets that this linear search attains perfect recall using the test data provided.
+We then use the linear-search results on the augmentations as ground truth to evaluate CAKES in Rust and to compute recall for HNSW, ANNOY, and FAISS-IVF in Python.
 
@@
 
-To demonstrate CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
-This dataset contains ribosomal RNA sequences of 2,224,640 genomes, with the longest sequence having 3,712 letters.
-We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
-We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
-We note that the sequences in this dataset were provided in a multiple sequence alignment with a width of 50,000 characters.
-We could have used Hamming distance on the aligned sequences, but we chose to use Levenshtein distance on the unaligned sequences to help demonstrate the flexibility of CAKES in handling exotic distance functions.
-Under Hamming distance, we could consider this dataset to have an embedding dimension of 50,000, but for Levenshtein distance, we state the dimensionality as the length of the longest sequence, i.e.\, 3,712.
+To demonstrate CAKES with an exotic distance, we benchmark the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}, which contains rRNA sequences from 2,224,640 genomes; the longest sequence has 3,712 letters.
+We hold out 1,000 random sequences as queries.
+We build the tree and perform $k$-NN using Levenshtein distance~\cite{levenshtein1966binary} on unaligned sequences.
+Although the dataset is provided as a multiple sequence alignment of width 50{,}000, we intentionally use Levenshtein rather than Hamming on the aligned strings to demonstrate CAKES’s flexibility.
+Under Hamming, the embedding dimension would be 50{,}000; under Levenshtein, we report the dimensionality as the length of the longest sequence (3{,}712).
 
@@
 
-As another example of an exotic distance function, we benchmarked CAKES on the Radio-ML dataset~\cite{oshea2018radioml} under Dynamic Time Warping~\cite{muller2007dynamic} as the distance function.
-This dataset contains samples of synthetically generated signal captures of different modulation modes over a range of signal-to-noise ratio (SNR) levels.
-Specifically, it comprises 24 modulation modes at 26 different SNRs ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
-Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
-Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
-We used a subset of this dataset, containing $97,304$ samples at 10dB SNR and used another $1,000$ samples at the same SNR as a hold-out set of queries.
+We also benchmark CAKES on the Radio-ML dataset~\cite{oshea2018radioml} using Dynamic Time Warping~\cite{muller2007dynamic} as the distance.
+The dataset spans 24 modulation modes across 26 SNR levels from $-20$ to $30$\,dB, with 4{,}096 samples per combination, totaling $24\!\cdot\!26\!\cdot\!4096=2{,}555{,}504$ samples.
+Each sample is a 1{,}024-dimensional complex-valued vector (a time series).
+For experiments, we use the 10dB SNR subset with 97{,}304 samples and a 1{,}000-sample hold-out query set at the same SNR.
+
+% \subsection{Local Fractal Dimension}
+% \label{sec:dayasets:lfd-of-datasets}
+
+% Since the CAKES algorithms were designed to scale in time with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
+% Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
+% In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
+
+% For the Fashion-MNIST dataset we observe in Figure~\ref{fig:results:fashion-mnist-lfd} that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
+% It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
+
+% Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
+% In particular, Glove-25's LFD is less than 3 for all depths.
+
+% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
+% It increases sharply to a peak of 9 around a depth of 10, after which it decreases smoothly until reaching the deepest leaves in the tree.
+
+% We contrast the low LFD of Sift with that of the Random dataset which has the same cardinality and dimension.
+As shown in Figure~\ref{fig:results:random-lfd}, the character of this dataset is very different from the others.
+% The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
+% The spread in LFD starts very small for the first few clusters and increases as depth increases.
+
+% The Silva dataset, shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
+% In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
+
+% For the Radio-ML dataset, the LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
+% Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
+% Within each of the tree portions, this dataset has a character similar to that of the Random dataset.
+% This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
+% This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
 

### 2025-08-26 sections/5-discussion.tex
tightening discussion

@@
 
-We have presented CAKES, a set of three algorithms for fast $k$-NN search that are generic over a variety of distance functions.
-CAKES's algorithms are exact when the distance function in use is a metric (see definition in Section~\ref{sec:methods}).
-Even under cosine distance, which is not a metric, CAKES's algorithms exhibit nearly perfect recall.
-CAKES's algorithms are designed to be most effective when the data uphold the manifold hypothesis, or in other words, when the data are constrained to a low-dimensional manifold even when embedded in a high-dimensional space.
-CAKES uses an unbalanced clustering algorithm to build a binary tree that better captures the manifold structure of the data and provides for faster search performance.
-As a consequence, these algorithms do not perform well on data with random distributions, because of the absence of a manifold structure.
-In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
+We have presented CAKES, a suite of three algorithms for fast $k$-NN search across diverse distance functions.
+When the distance is a metric (Sec.~\ref{sec:methods}), the algorithms are exact; under cosine distance (non-metric) they achieve nearly perfect recall.
+CAKES is most effective when data satisfy the manifold hypothesis—lying on a low-dimensional manifold within a high-dimensional space.
+It builds an unbalanced-clustering binary tree that captures manifold structure and accelerates search.
+Conversely, performance degrades on randomly distributed data, where such structure is absent.
 
@@
 
-As expected, our results from Figure~\ref{fig:results:scaling-plots} indicate that CAKES's algorithms scale sub-linearly with cardinality on real-world datasets with low LFD, and scale linearly on synthetic datasets with high LFD.
-CAKES's performance on Sift against that on the Random dataset illustrates this phenomenon well.
-As seen in Figures~\ref{fig:results:random-lfd}~and~\ref{fig:results:sift-lfd}, the LFD of the Random dataset is much higher than that of Sift, even though both datasets have the same cardinality and dimensionality.
-This is unsurprising, given that the Random dataset is uniformly distributed, while Sift is not.
-On the Random dataset, CAKES's speed decreases linearly as the multiplier increases, while with Sift, Depth-First Sieve and Breadth-First Sieve both exhibit nearly constant throughput.
-While Repeated $\rho$-NN does not exhibit near-constant throughput with Sift, throughput decreases much more slowly than it does with the Random dataset.
-Since these two datasets have the same cardinality and embedding dimension, the aforementioned discrepancies highlight how manifold structure affects algorithm performance.
+As expected, Fig.~\ref{fig:results:scaling-plots} shows CAKES scaling sublinearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD; Sift versus Random illustrates this contrast.
 
-We emphasize that although CAKES's algorithms are slower on the Random dataset, their recalls remain perfect.
-Figure~\ref{fig:results:scaling-plots} also shows that HNSW and ANNOY have near-constant throughput as the multiplier increases, but their recall continues to decrease as the multiplier increases.
-This, coupled with the increasing time and space cost of building the indices for HNSW and ANNOY, makes these algorithms unsuitable for real-world applications in which the datasets are growing exponentially.
-CAKES's tree is orders of magnitude faster and less memory-intensive to build, and CAKES's algorithms, while slower than HNSW and ANNOY, still show near-constant scaling as well as perfect recall.
+On Random, CAKES throughput decreases linearly with the multiplier, whereas on Sift, Depth-First Sieve and Breadth-First Sieve are nearly constant.
+Repeated $\rho$-NN is not near-constant on Sift but decays much more slowly than on Random.
+Because Sift and Random share cardinality and embedding dimension, this isolates the effect of manifold structure on performance.
+
+Although CAKES is slower on Random, its recall remains perfect.
+HNSW and ANNOY maintain near-constant throughput as the multiplier grows, but their recall declines, and their index build time and memory rise, making them unsuitable when datasets grow exponentially.
+CAKES’s tree builds orders of magnitude faster with far less memory; while queries are slower than HNSW/ANNOY, CAKES attains near-constant scaling with perfect recall.
+
+The fastest CAKES variant depends on dataset and cardinality.
+On Fashion-MNIST, Glove-25, and Sift, Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as cardinality increases and consistently outperform linear search at high cardinalities.
+This variability motivates auto-tuning (Sec.~\ref{sec:methods:auto-tuning}); future work will better characterize when each algorithm excels and enable more sophisticated selection from dataset properties.
+
+We stress that CAKES targets \emph{big} data: linear search is better at small sizes, but for each ANN-Benchmarks dataset tested, CAKES overtakes linear search at sufficiently large cardinality—about $10^{5}$ for Fashion-MNIST and $10^{6}$ for Glove-25.
+On the high-LFD Random dataset, CAKES never beats linear search.
+These results support that CAKES performs well on data arising from constrained generating phenomena even as cardinality grows exponentially.
 
@@
 
-While the fastest CAKES algorithm can vary by dataset and cardinality, we note that on all three ANN-benchmark datasets (Fashion-MNIST, Glove-25, and Sift), Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as the cardinality increases.
-This observation warrants further investigation, but it is especially promising given that both algorithms consistently outperform linear search for high cardinalities.
-Finally, we emphasize that the variation in performance of our algorithms across different datasets and cardinalities supports our use of an auto-tuning function (as discussed in Section~\ref{sec:methods:auto-tuning}) to select the best algorithm for a given dataset.
-Future work is needed to better characterize those datasets for which each algorithm performs best, but our results suggest that a more sophisticated auto-tuning function could be developed to select the fastest algorithm based on the properties of the dataset.
-
-We also stress that CAKES's algorithms are designed to work well for \textit{big} data, and our results support this claim;
-we observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmark datasets we tested.
-Moreover, we observe that CAKES's algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD.
-For example, Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^7$ for Sift, as opposed to about $10^5$ and $10^6$ for Fashion-Mnist and Glove-25, respectively.
-For the Random dataset, which has the highest LFD, CAKES's algorithms \textit{never} outperform linear search.
-These observations support our claim that CAKES performs well on datasets that arise from constrained generating phenomena, even as the cardinalities of these datasets grow exponentially.
-
 \subsection{Applicability}
 
-We note that despite the genericity of CAKES, effectively applying it to real-world datasets may require domain-specific knowledge.
-As an example, consider the domain of protein sequence search; protein sequences are of highly variable length (30--1000 amino acids are typical, with some outliers) and typically comprise a 20-letter alphabet.
-There have been numerous protein sequence search algorithms~\cite{kim2021entrance, daniels2013compressive, yu2015entropy, steinegger2018clustering}, and the one thing they have in common besides the application domain is that their development required deep domain-specific knowledge.
-While simple edit distance may be a poor choice for clustering when sequence lengths exhibit such variance, other choices of distance metrics and preprocessing could be used by CAKES, such as representing each sequence as a set of $k$-mers (short substrings of length $k$) and choosing Jaccard distance on these sets as the metric; this approach was chosen by~\cite{kim2021entrance}.
-If the distance function uses a global sequence alignment algorithm, such as Needleman-Wunsch~\cite{needleman1970general}, the dataset may also be preprocessed into a number of mutually exclusive subsets based on the lengths of the sequences, and a separate CAKES tree may be built for each subset.
-At search time, the query sequence would be assigned to the appropriate tree based on its length, and the search would be performed on that tree, and on a small number of trees on adjacent lengths' bins.
+Despite CAKES’s generality, effective deployment on real data can require domain knowledge.
+In protein sequence search, sequences vary widely in length (typically 30–1000 amino acids, with outliers) over a 20-letter alphabet.
+Numerous algorithms exist~\cite{kim2021entrance, daniels2013compressive, yu2015entropy, steinegger2018clustering}, and their development relied on deep domain expertise.
+
+Simple edit distance can be poor when lengths vary substantially.
+CAKES can get around that issue by instead using alternative metrics or preprocessing—for example, represent each sequence as a set of $k$-mers and use Jaccard distance, as in~\cite{kim2021entrance}.
+If the distance relies on global alignment (e.g., Needleman–Wunsch~\cite{needleman1970general}), the dataset can be partitioned into mutually exclusive length bins with a separate CAKES tree per bin; at query time, route the sequence to its bin and a small number of adjacent bins.
 
@@
 
-The questions raised by this study suggest several additional avenues for future work.
-A comparison across more datasets is in order, as is further analysis with other distance functions that existing methods, such as FAISS, HNSW and ANNOY, do not support.
-These include Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly for high dimensional distributions) and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures by their maximal common subgraphs.
-Incorporating new distance functions in CAKES requires only that a Rust implementation of the distance function be provided.
+This study suggests several directions.
+First, we want to extended our evaluation to more datasets and distance functions not supported by FAISS, HNSW, or ANNOY, including Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly high-dimensional) and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures via maximal common subgraphs.
+Adding a new distance to CAKES requires only a Rust implementation.
 
-Additionally, we intend to improve the data augmentation process described in Section~\ref{sec:methods:synthetic-data} to better preserve the topological structure of the underlying dataset.
-In particular, we expect that favoring augmentation along the first few principal components of the local manifold may allow for more realistic data augmentation.
+Second, we intend to improve the augmentation procedure (Sec.~\ref{sec:methods:synthetic-data}) to better preserve topological structure of datasets, e.g., by favoring perturbations along the leading principal components of the local manifold.
 
-% We already did compression so it feels weird to have this in there:
+Third, we plan to explore CAKES algorithms' use in a streaming environment, enabling online updates to the tree as points are inserted or deleted.
 
-% We also plan to investigate hierarchical data compression by representing differences at each level of the binary tree, particularly for string or genomic data, where all differences are discrete.
-% Such an encoding would allow us to store the data in a compressed format, and to perform search on the compressed data without decompressing it in its entirety.
-% This would allow us to perform fast search on datasets that are otherwise too large to fit in memory.
-% An exploration of compression and compressed search is another avenue for future work.
-
-We also would like to explore the use of CAKES in a streaming environment.
-This would require the ability to perform ``online'' updates to the tree as points are added to or deleted from the dataset.
-Such online updates would take advantage of the fast search algorithms provided by CAKES.  % and would allow us to have our cake and eat it, too.
-CAKES can also be used to extend anomaly detection in CHAODA~\cite{ishaq2021clustered}.
-We could add to CHAODA's ensemble of graph-based anomaly detection methods by using the distribution of distances among the $k$ nearest neighbors of cluster centers.
+Finally, extend anomaly detection in CHAODA~\cite{ishaq2021clustered} by adding graph-based methods that use the distribution of distances among the $k$ nearest neighbors of cluster centers.
 
@@
 
-CLAM and CAKES are implemented in the Rust programming language, and the source code is available under an MIT license at https://github.com/URI-ABD/clam.
+CLAM and CAKES are implemented in Rust and released under the MIT license at https://github.com/URI-ABD/clam.

### 2025-08-26 sections/1-introduction.tex
minor clarity edits

@@
 
-$k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because due to linear time complexity.
+$k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because they scale linearly with dataset size.
 Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.

### 2025-08-26 sections/2-methods.tex
minor clarity edits

@@
 Hence, the \textit{embedding dimension} of this dataset is at least 3,712 and at most 50,000; as used in this paper it is 3,712.
-However, physical constraints (namely, biological evolution and biochemistry) constrain the data to a lower-dimensional manifold within this space.
+However, physical limitations (namely, biological evolution and biochemistry) constrain the data to a lower-dimensional manifold within this space.
 LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
@@
 
-We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
-The center is the geometric median of the points in the cluster, i.e.,\,it is the point that minimizes the sum of distances to all other points in the cluster.
-In cases where the cardinality of the cluster is large, we take a random subsample of $\sqrt{|C|}$ points and compute the geometric median of that subsample~\cite{ishaq2019clustered}.
-The center, therefore, is one of the points in the cluster and is used as a representative of the cluster.
-The radius is the maximum distance from the center to any point in the cluster.
-Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
-Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
-As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e.,\,$C(c, r) \subset B(c, r)$.
-We denote the cluster tree by $\mathcal{T}$ and the root by $\mathcal{R}$ (see Section~\ref{sec:methods:clustering:building-the-tree}).
+We define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
+The center is the geometric median of the points in $C$; for large $|C|$ we take a random subsample of size $\sqrt{|C|}$ and use its geometric median~\cite{ishaq2019clustered}.
+Thus the center is one of the points in $C$ and serves as its representative.
+The radius is the maximum distance from the center to any point in $C$.
+Each non-leaf cluster has two children (as in a binary tree).
+Clusters may have overlapping volumes, but each point in an overlap is assigned to exactly one cluster; consequently $C(c,r)\subset B(c,r)$.
+We denote the cluster tree by $\mathcal{T}$ and its root by $\mathcal{R}$ (Sec.~\ref{sec:methods:clustering:building-the-tree}).
 
-Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius, i.e.,\,using Equation~\ref{eq:methods:lfd-half}.
-We also only use points that are in $C(c, r)$ instead of all those in $B(c, r)$.
+Unless stated otherwise, the LFD of a cluster is estimated at scales $r$ and $r/2$ via Eq.~\ref{eq:methods:lfd-half}, using only points in $C(c,r)$ (not all of $B(c,r)$).
 
-The \textit{metric entropy} $\mathcal{N}_{r}(X)$ for some radius $r$ is the minimum number of clusters of a uniform radius $r$ needed to cover the data~\cite{yu2015entropy} for a flat clustering with clusters of radius $r$.
-In this paper, where the clustering is hierarchical rather than flat, we define the metric entropy $\mathcal{N}_{\hat{r}}(X)$ as the number of leaf clusters in the tree where $\hat{r}$ is the mean radius of all leaf clusters.
-
-\subsubsection{Building the Tree}
-\label{sec:methods:clustering:building-the-tree}
+For a flat clustering, the \emph{metric entropy} $\mathcal{N}_r(X)$ is the minimum number of radius-$r$ clusters needed to cover $X$~\cite{yu2015entropy}.
+In our hierarchical setting, we define $\mathcal{N}_{\hat r}(X)$ as the number of leaf clusters, where $\hat r$ is the mean radius of all leaves.
 

### 2025-08-26 sections/4-results.tex
minor clarity edits

@@
 
-For each dataset, we benchmark using the distance function listed in Table~\ref{tab:datasets:summary}, and results in this section are based on k-NN with $k=10$.
+For each dataset, we benchmark using the distance function listed in Table~\ref{tab:datasets:summary}.
+As described in Section~\ref{sec:introduction}, approximate search can be more sensitive to local perturbations than exact search, especially for small values of $k$, and hence we focus on $k=10$ for our benchmarks.
 

### 2025-08-26 sections/4-results.tex
minor edits to results

@@
 For Silva and RadioML, due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
+These plots also illustrate how the CAKES algorithms compare to na\"{i}ve linear search.
 
@@
 
-\textbf{General notes (applies to Tables \ref{tab:results:qps-and-recall-fmn},\ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}).}
   Throughput is measured in queries per second.
 A recall value of $1.000*$ denotes imperfect recall that rounds to $1.000$.
-In each table, we observe that while recall for CAKES does \emph{not} degrade with cardinality, recall for
-  HNSW and ANNOY degrades with cardinality. CAKES exhibits perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
+In each of Tables \ref{tab:results:qps-and-recall-fmn},\ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, we observe that while recall for CAKES does \emph{not} degrade with cardinality, recall for
+  HNSW and ANNOY does degrade with cardinality. CAKES exhibits perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
 
@@
     \caption{Random dataset: throughput and recall.
-    In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.CAKES exhibits perfect recall at all cardinalities, whereas HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
+    In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
+CAKES exhibits perfect recall at all cardinalities, whereas HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
     }

### 2025-08-27 sections/1-introduction.tex
orphans, whitespace, lfd plots

@@
 ANNOY~\cite{annoy} uses random projection and tree building.
-At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.
-Multiple such trees form a forest; increasing the number of trees improves recall but increases search time.
+At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.  Increasing the number of such trees improves recall but increases search time.
 
@@
 
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended the original flat clustering approach to entropy-scaling $\rho$-NN search from a flat clustering approach~\cite{yu2015entropy} to a hierarchical one.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended the flat clustering approach used in~\cite{yu2015entropy} to a hierarchical one.
 CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, refined the clustering algorithm from CHESS.
@@
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
-Given that CAKES's algorithms are intended to be used on datasets with a manifold structure, complexity analysis assuming a uniform distribution of data would be uninformative.
+Given that CAKES's algorithms are intended for datasets with a manifold structure, complexity analyses assuming a uniform distribution would be uninformative.
 As a result, our analysis assumes datasets with manifold structure—characterized by low fractal dimension and metric entropy—reflecting the conditions for which CAKES is designed.

### 2025-08-27 sections/2-methods.tex
orphans, whitespace, lfd plots

@@
 
-CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
+CAKES derives advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena often occupy a low-dimensional manifold within their embedding space.
 We define the \emph{local fractal dimension} (LFD) to quantify this property.
-
 For a point $q \in \textbf{X}$ and radii $r_1 > r_2$,
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
     \vskip -0.2in
@@
     \label{fig:methods:deltas}
-    \vskip -0.5in
+    % \vskip -0.5in
 \end{figure}
+\FloatBarrier
 
-\begin{minipage}{.425\textwidth}
-    \begin{algorithm}[H]
+% \begin{minipage}{.425\textwidth}
+\FloatBarrier
+    \begin{algorithm}[!t]
         \caption{tree-search($C$, $q$, $r$)}
         \label{alg:methods:rnn-search:tree-search}
-        \begin{algorithmic}
+        \begin{algorithmic}[1]
             \REQUIRE $f$, a distance function
@@
     \end{algorithm}
-\end{minipage}
-\hfill
-\begin{minipage}{.475\textwidth}
-    \begin{algorithm}[H]
+% \end{minipage}
+
+% \begin{minipage}{.475\textwidth}
+    \begin{algorithm}[!t]
         \caption{leaf-search($Q$, $q$, $r$)}
         \label{alg:methods:rnn-search:leaf-search}
-        \begin{algorithmic}
+        \begin{algorithmic}[1]
             \REQUIRE $f$, a distance function
@@
     \end{algorithm}
-\end{minipage}
+% \end{minipage}
 

### 2025-08-27 sections/3-datasets.tex
orphans, whitespace, lfd plots

@@
 
-% \subsection{Local Fractal Dimension}
-% \label{sec:dayasets:lfd-of-datasets}
+\subsection{Local Fractal Dimension}
+\label{sec:dayasets:lfd-of-datasets}
+
+\begin{figure}[t]
+    \centering
+    % Row 1
+    \subfloat[Fashion-MNIST\label{fig:results:fashion-mnist-lfd}]{
+      \includegraphics[width=0.48\columnwidth]{images/lfd/fashion-mnist.png}}
+    \hfill
+    \subfloat[Glove-25\label{fig:results:glove-25-lfd}]{
+      \includegraphics[width=0.48\columnwidth]{images/lfd/glove-25.png}}
+
+    % Row 2
+    \par\medskip
+    \subfloat[Sift\label{fig:results:sift-lfd}]{
+      \includegraphics[width=0.48\columnwidth]{images/lfd/sift.png}}
+    \hfill
+    \subfloat[Random dataset\label{fig:results:random-lfd}]{
+      \includegraphics[width=0.48\columnwidth]{images/lfd/random.png}}
+
+    % Row 3
+    \par\medskip
+    \subfloat[Silva 18S\label{fig:results:silva-lfd}]{
+      \includegraphics[width=0.48\columnwidth]{images/lfd/silva-SSU-Ref.png}}
+    \hfill
+    \subfloat[RadioML\label{fig:results:radioml-lfd}]{
+      \includegraphics[width=0.48\columnwidth]{images/lfd/radio-ml.png}}
+
+    % Legend
+    \par\smallskip
+    \includegraphics[width=0.65\columnwidth]{images/lfd/legend.png}
+
+    \caption{Local fractal dimension vs.
+cluster depth across six datasets.
+The ‘random’ dataset is generated as in Sec.~\ref{sec:datasets-and-benchmarks:random-datasets}; note the different y-axis.
+The x-axis is tree depth; the y-axis is LFD at that depth.
+We plot the 5th, 25th, 50th, 75th, and 95th percentiles, plus min/max.
+Each cluster is counted by its cardinality so the curves reflect dataset-level distributions.}
+    \label{fig:results:lfd-plots}
+  \end{figure}
 

### 2025-08-27 sections/3-datasets.tex
added lfd discussion

@@
 \label{sec:dayasets:lfd-of-datasets}
+
+Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each
+dataset we used for benchmarks, as shown in Figure~\ref{fig:results:lfd-plots}. In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to
+the 95th percentile of LFD.
+
+In Figure~\ref{fig:results:fashion-mnist-lfd}, we observe that until approximately depth 5, Fashion-MNIST’s LFD is low (i.e., less than 4). 
+It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
+Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}. 
+In particular, Glove-25’s LFD is less than 3 for all depths.
+Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25. 
+It increases sharply to a peak of 9 around a depth of 10.
+For the Random dataset, which has the same cardinality and dimensionality as Sift, LFD starts at 20 at depth
+0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree. The spread in LFD starts very
+small for the first few clusters and increases as depth increases. For Silva, as
+shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD. In particular, LFD is less than 3 for all depths, hovering near 1 for
+clusters at depth 40 and deeper. For the Radio-ML dataset, Figure~\ref{fig:results:radioml-lfd} LFD values show three distinct peaks around an LFD
+of 12 at or near depths of 8, 25 and 50. Each peak is followed by a linear decrease until encountering a sharp spike for
+the next peak. Within each of the three portions, this dataset has a character very similar to that of the Random dataset.
 

### 2025-08-27 sections/5-discussion.tex
added lfd discussion

@@
 Conversely, performance degrades on randomly distributed data, where such structure is absent.
+In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we texted exhibits a manifold structure, as quantified by the percentiles of LFD.
 
@@
 
-We stress that CAKES targets \emph{big} data: linear search is better at small sizes, but for each ANN-Benchmarks dataset tested, CAKES overtakes linear search at sufficiently large cardinality—about $10^{5}$ for Fashion-MNIST and $10^{6}$ for Glove-25.
-On the high-LFD Random dataset, CAKES never beats linear search.
-These results support that CAKES performs well on data arising from constrained generating phenomena even as cardinality grows exponentially.
+We stress that CAKES targets \emph{big} data, and our results support this claim; we
+observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes
+linear search at a large enough cardinality for each of the ANN-benchmark datasets we tested. 
+Moreover, we observe
+that CAKES’s algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD. For example,
+Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at
+a cardinality of $10^{7}$ 
+for Sift, as opposed to about $10^{5}$
+and $10^{6}$ 
+for Fashion-Mnist and Glove-25, respectively. For the
+Random dataset, which has the highest LFD, CAKES’s algorithms \emph{never} outperform linear search.
 

### 2025-08-29 sections/1-introduction.tex
found a of space

@@
 
-Examples of such datasets include genomic databases, time-series signals, and neural network embeddings such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR}, LLAMA-2~\cite{Touvron2023Llama2O}, and image embedding models~\cite{radford2021learning, dosovitskiy2020image}.
-Among biological datasets,  SILVA 18S~\cite{10.1093/nar/gks1219} includes ribosomal RNA sequences from approximately 2.25 million genomes, reaching 50,000 letters when aligned.
+Examples of such datasets include genomic databases, time-series signals, and embeddings from neural networks such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR}, LLAMA-2~\cite{Touvron2023Llama2O}, and image analysis models~\cite{radford2021learning, dosovitskiy2020image}.
+Among biological datasets, SILVA 18S~\cite{10.1093/nar/gks1219} includes ribosomal RNA sequences from approximately 2.25 million genomes.
 Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated radio frequency signals.
@@
 
-Given some measure of similarity between data points, two standard similarity search paradigms are $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
+Given some measure of similarity between data points, two common similarity search paradigms are $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
 The former retrieves the $k$ points closest to a query, while the latter retrieves all points within a similarity threshold $\rho$ of a query.
@@
 Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
-In a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but be sensitive to local perturbations for smaller values of $k$.
 
@@
 
-Several algorithms have been developed to scale $k$-nearest neighbor search to large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Several algorithms have been developed to tackle $k$-nearest neighbor search on large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
 However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).
@@
 Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networks and skip lists.
-InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches exhaustively only within cells near the query.
+InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
 ANNOY~\cite{annoy} uses random projection and tree building.
-At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.  Increasing the number of such trees improves recall but increases search time.
+At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.
+Increasing the number of such trees improves recall but increases search time.
 
@@
 
-Entropy-scaling algorithms exploit the inherent structure of large datasets, achieving complexity that scales with topological properties—such as metric entropy and fractal dimension—rather than cardinality.
-This makes them especially suitable for manifold-structured data.
+Entropy-scaling algorithms exploit the inherent structure of large datasets, achieving complexity that scales with topological properties, such as metric entropy and fractal dimension, rather than cardinality.
+This makes them especially suitable for manifold-constrained data.
 
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended the flat clustering approach used in~\cite{yu2015entropy} to a hierarchical one.
-CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, refined the clustering algorithm from CHESS.
+In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended the flat clustering from~\cite{yu2015entropy} to a hierarchical approach.
+CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, refines the clustering algorithm from CHESS.
 
 This paper introduces CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
-We also provide a theoretical analysis of their time complexity in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
+We provide a theoretical analysis of their time complexity in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.

### 2025-08-29 sections/2-methods.tex
found a of space

@@
 In this manuscript, we study $k$-NN search in finite-dimensional spaces.
-Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$, we refer to each $x_i \in \textbf{X}$ as a point or datum.
-Examples include image embeddings, genomic sequences, and time-series measurements.
+Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$, we refer to each $x_i \in \textbf{X}$ as a point.
+Examples include neural embeddings, genomic sequences, and time-series.
 
 We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
-We require $f$ to be symmetric and identity-preserving ($f(x, y) = 0 \iff x = y$). 
-If $f$ satisfies the triangle inequality (i.e.,\,$f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
-Common examples include Euclidean, Levenshtein~\cite{levenshtein1966binary}, and DTW~\cite{muller2007dynamic}; cosine distance, which violates the triangle inequality, is not a metric.
-Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
+We require $f$ to be symmetric ($f(x, y) = f(y, x)$) and identity-preserving ($f(x, y) = 0 \iff x = y$). 
+If $f$ satisfies the triangle inequality (i.e.,\,$f(x, y) \leq f(x, z) + f(z, y)$), then it is also a \textit{distance metric}.
+Common examples include Euclidean, Levenshtein~\cite{levenshtein1966binary}, and DTW~\cite{muller2007dynamic};
+Cosine distance violates the triangle inequality and is not a metric.
+As in~\cite{yu2015entropy}, all CAKES' algorithms are exact under distance metrics.
 
@@
 
-The appropriate distance function depends on the data: Euclidean or cosine is often used for neural embeddings, Levenshtein or Hamming for sequences, and DTW or Wasserstein for time-series data.
+The appropriate distance function depends on the data: Euclidean or Cosine are often used for neural embeddings, Levenshtein or Hamming for sequences, and DTW or Wasserstein for time-series.
 
-CAKES derives advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena often occupy a low-dimensional manifold within their embedding space.
+CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, the notion that high($D$)-dimensional data collected from constrained generating phenomena often occupy a low($d$)-dimensional manifold within their embedding space ($d \ll D$).
 We define the \emph{local fractal dimension} (LFD) to quantify this property.
@@
 where $B(q, r)$ is the set of points in the metric ball of radius $r$ centered at $q$.
-
 We typically use a simplified version of Equation~\ref{eq:methods:lfd-original} with $r_1 = 2 \cdot r_2$:
@@
 Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases.
-When the vast majority of points in the dataset have low ($\ll D$) LFD, we can simply say that the dataset has low LFD.
+When the vast majority of points in the dataset have low ($d \ll D$) LFD, we simply say that the dataset has low LFD.
 We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the SILVA 18S rRNA dataset that contains genomic sequences with unaligned lengths of up to 3,712 base pairs and aligned length of 50,000 base pairs.
-Hence, the \textit{embedding dimension} of this dataset is at least 3,712 and at most 50,000; as used in this paper it is 3,712.
-However, physical limitations (namely, biological evolution and biochemistry) constrain the data to a lower-dimensional manifold within this space.
-LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
+To illustrate the difference, consider the SILVA 18S rRNA dataset that contains genomic sequences with unaligned lengths of up to 3,712 bases and aligned length of 50,000 bases.
+Hence, the \textit{embedding dimension} of this dataset is at least 3,712 and at most 50,000.
+However, physical limitations (namely, biological evolution and biochemistry) constrain the data to a low-dimensional manifold within this space.
+LFD is an approximation of the dimensionality of that low-dimensional manifold in the ``vicinity'' of a given point.
 % Section~\ref{sec:results:lfd-of-datasets} discusses this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
-For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but potentially highly variable at larger length scales, i.e.,\,when $r$ is large.
+For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but (potentially) globally variable.
 
@@
 We define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
-The center is the geometric median of the points in $C$; for large $|C|$ we take a random subsample of size $\sqrt{|C|}$ and use its geometric median~\cite{ishaq2019clustered}.
-Thus the center is one of the points in $C$ and serves as its representative.
+The center is the geometric median—the point minimizing the sum of distances to all other points—of the points in $C$;
+in practice, for large $|C|$ we find the geometric median of a random subsample of $\sqrt{|C|}$ points from $C$.
 The radius is the maximum distance from the center to any point in $C$.
 Each non-leaf cluster has two children (as in a binary tree).
-Clusters may have overlapping volumes, but each point in an overlap is assigned to exactly one cluster; consequently $C(c,r)\subset B(c,r)$.
+Clusters may have overlapping volumes, but each point in an overlap is assigned to exactly one cluster;
+consequently $C(c,r)\subset B(c,r)$.
 We denote the cluster tree by $\mathcal{T}$ and its root by $\mathcal{R}$ (Sec.~\ref{sec:methods:clustering:building-the-tree}).
 
-Unless stated otherwise, the LFD of a cluster is estimated at scales $r$ and $r/2$ via Eq.~\ref{eq:methods:lfd-half}, using only points in $C(c,r)$ (not all of $B(c,r)$).
+Unless stated otherwise, we estimate the LFD of clusters at scales $r$ and $r/2$ via Eq.~\ref{eq:methods:lfd-half}, using only the points in $C(c,r)$ (not all of $B(c,r)$).
 
-For a flat clustering, the \emph{metric entropy} $\mathcal{N}_r(X)$ is the minimum number of radius-$r$ clusters needed to cover $X$~\cite{yu2015entropy}.
+For a flat clustering, \emph{metric entropy} $\mathcal{N}_r(X)$ is the minimum number of radius-$r$ clusters needed to cover $X$~\cite{yu2015entropy}.
 In our hierarchical setting, we define $\mathcal{N}_{\hat r}(X)$ as the number of leaf clusters, where $\hat r$ is the mean radius of all leaves.
@@
 
-We construct a divisive hierarchical cluster tree $\mathcal{T}$ with CLAM, following CHESS~\cite{ishaq2019clustered} but with improved pole selection, then apply a depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}).
+We construct $\mathcal{T}$ using CLAM, following CHESS~\cite{ishaq2019clustered} but with improved pole selection, then apply a depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}) on the dataset.
 
 The recursive step is as follows.
-For any cluster $C$ with $|C|$ points, we draw a random subsample $S\subset C$ of $\sqrt{|C|}$ points, compute all pairwise distances in $S$, and take the geometric median of $S$—the point minimizing the sum of distances to all other points in $S$—as the \emph{center} of $C$.
-The \emph{radius} of $C$ is the maximum distance from this center to any point in $C$.
+For a cluster $C$ with $|C|$ points, we draw a random subsample $S\subset C$ of $\sqrt{|C|}$ points, compute all pairwise distances in $S$, and take the geometric median of $S$ as the \emph{center} of $C$.
+The point in $C$ farthest from the center is the \emph{left pole}.
+The distance from the center to the left pole if the \emph{radius}.
+The point in $C$ farthest from the left pole is the \emph{right pole}.
+We partition $C$ into \emph{left} and \emph{right child} clusters containing points closer to the left and right poles respectively, with ties going to the left child.
 
-Let the \emph{left pole} be the point in $C$ farthest from the center; let the \emph{right pole} be the point farthest from the left pole.
-We partition $C$ into a \emph{left child} (points closer to the left pole) and a \emph{right child} (points closer to the right pole); ties go to the left child.
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, we recurse until each leaf is a single datum or any user-specified stopping criterion is met (e.g., minimum cluster radius, minimum cardinality, or maximum tree depth).
+Starting from the root cluster $\mathcal{R}$ containing the entire dataset, we recurse until each leaf contains a single point or some user-specified stopping criterion is met (e.g., minimum cluster radius, maximum tree depth, etc.).
 During partitioning, we compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
@@
 
-The asymptotic complexity of Partition is the same as described in~\cite{ishaq2019clustered}.
-By using an approximate partitioning with a $\sqrt{n}$ sample, we achieve $\mathcal{O}(n)$ cost of partitioning and $\mathcal{O}(n \log n)$ cost of building $\mathcal{T}$.
-This is a significant improvement over exact partitioning and tree-building, which cost $\mathcal{O}(n^2)$ and $\mathcal{O}(n^2 \log n)$ respectively.
+The asymptotic complexity of building $\mathcal{T}$ is the same as described in~\cite{ishaq2019clustered}, i.e., $\mathcal{O}(n \log n)$.
+This is a significant improvement over the exact approach, which costs $\mathcal{O}(n^2 \log n)$.
 
@@
 % This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
-We also note here that this $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
-In practice, for real datasets, we expect the process from Section~\ref{sec:methods:clustering:building-the-tree} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
-The only case in which we would expect a balanced tree is if the dataset were uniformly distributed, e.g.,\,in a $d$-dimensional hyper-cube.
+We also note here that the $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
+For real datasets, we expect the process from Section~\ref{sec:methods:clustering:building-the-tree} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
+The only case in which we would expect a balanced tree is if the dataset were uniformly distributed in a $D$-dimensional space.
 
@@
 
-Let $\textbf{X}$ be a dataset, $f$ a distance function, and $\mathcal{R}$ the root cluster of a tree $\mathcal{T}$ constructed by the process in Section~\ref{sec:methods:clustering:building-the-tree}.
 Given a user-specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
-In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B_X(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
+In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
 
@@
 Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being one of the $k$ nearest neighbors.
-These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being among the $k$ nearest neighbors.
+These algorithms also use some terminology defined in Figure~\ref{fig:methods:deltas}.
 
@@
 For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
-If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.
+If no clusters are found, we keep doubling $r$ and repeating tree-search, until we find a a non-empty set, $Q$, of clusters.
 
-Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
 In particular, we increase the radius by a factor of
@@
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
-We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.
+We cap this factor at 2 to ensure that we do not increase the radius too quickly in any single iteration.
 
@@
 We consider the \textit{tree-search} and \textit{leaf-search} stages of search separately.
-Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors by doing repeated iterations of the  $\rho$-NN search Algorithm~\ref{alg:methods:rnn-search:tree-search}.
+Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors.
 In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
@@
         }_{\textrm{tree-search}}
-        \ + \
+        +
         \underbrace{
             \overbrace{ \big| B(q, \rho) \big|}^{\textrm{output size}}
-            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
+            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ {\text{LFD}(q, \rho)}}^{\textrm{scaling factor}}
         }_{\textrm{leaf-search}}
-    \Bigg)
+    \Bigg).
     \label{eq:methods:rnn-search-complexity}
 \end{gather}
-where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.
 
 To extend Equation~\ref{eq:methods:rnn-search-complexity} to Repeated $\rho$-NN, we must first estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
-Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the (harmonic) mean of the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
-Given that LFD near $q$ does not differ significantly from that of nearby clusters, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
+Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
+Thus, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
 one iteration to find at least one cluster, and one more to find enough clusters to guarantee $k$ neighbors.
-Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that of $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
+Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that for $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
 
@@
 Let $Q$ be the set of clusters returned by tree-search.
-We must estimate $\sum_{C \in Q} |C|$, the total cardinality of the clusters returned by tree-search; since we must examine every point in each such cluster, time complexity of leaf-search is linear in this quantity.
+We must estimate $\sum_{C \in Q} |C|$, the total cardinality of those clusters;
+since we must examine every point in each such cluster, time complexity of leaf-search is linear in this quantity.
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
@@
 
-The work in~\cite{yu2015entropy} showed that 
-\begin{equation} \sum_{C \in S} |C| \leq \gamma  \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d \end{equation}, where $\gamma$ is a constant.
-By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k.$
+The work in~\cite{yu2015entropy} showed that, for some constant $\gamma$,
+\begin{equation*}
+    \sum_{C \in S} |C| \leq \gamma \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d.
+\end{equation*}
+
+By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k$.
 Thus, $\sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d$.
-It remains to determine an estimate for $\rho_k$.
-We let $\hat{d}$ be the (harmonic) mean LFD of the clusters in $Q$.
+It remains to estimate $\rho_k$.
+
+Let $\hat{d}$ be the harmonic mean of the LFDs of the clusters in $Q$.
 While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
 Since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at the center of a cluster in $Q$.
-By Equation~\ref{eq:methods:lfd-original}, $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$.
-We can rearrange this equation to get $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
+By Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
 Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get:
 \begin{equation*}
-    k \left( 1 + 2 \cdot \left( \frac{\hat{|C|}}{k} \right) ^ {\hat{d}^{-1}} \right)^d
+    k \Bigg( 1 + 2 \cdot \bigg( \frac{\hat{|C|}}{k} \bigg) ^ {\hat{d}^{-1}} \Bigg)^d
 \end{equation*}
 
-Again using the assumption that since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
-By combining the bounds for tree-search and leaf-search, we see that time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
+Again using the assumption that $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
+By combining the bounds for tree-search and leaf-search, we see that time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is
+\begin{gather}
     \mathcal{O}
@@
         }_{\textrm{leaf-search}}
-            \Bigg)
+    \Bigg).
     \label{eq:methods:repeated-rnn-complexity}
 \end{gather}
-        where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
 
@@
 
-This algorithm performs a breadth-first traversal of the tree $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} at each level.
+This algorithm (BFS) performs a breadth-first traversal of $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} (a subroutine in QuickSort), at each level of $\mathcal{T}$.
 
@@
 
-% \vskip 0.1in
-
 We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$.
@@
 Next, we go over all items in $Q$, skipping over any for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
-If the item corresponds to a point, we keep it.
-If the item corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
-If the item corresponds to a non-leaf cluster, we to $Q$ the pairs of 3-tuples corresponding to its child clusters.
+If an item corresponds to a point, we keep it.
+If it corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
+Otherwise (since it must correspond to a non-leaf cluster), we to $Q$ the pairs of 3-tuples corresponding to its child clusters.
 We continue this process until the sum of multiplicities in $Q$ is exactly $k$.
@@
 
-This algorithm is quite similar to a depth-first traversal of the tree $\mathcal{T}$, using two priority queues to track clusters and hits, and to prioritize which branch of $\mathcal{T}$ to explore next.
+This algorithm (DFS) is similar to a depth-first traversal of $\mathcal{T}$.
+It uses a max-priority queue to track hits, and a min-priority queue to choose the next branch of $\mathcal{T}$ to explore.
 
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$ and the ability to maintain its maximum size) of points prioritized by $\delta$.
-$Q$ starts containing only the root cluster $\mathcal{R}$ while $H$ starts empty.
-So long as $H$ is not full or the top element in $H$ has $\delta$ greater than or equal to the top element in $Q$, we take the following steps:
-
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
+$Q$ starts containing only $\mathcal{R}$ while $H$ starts empty.
+So long as $H$ is not full or the top-priority elements from $H$ and $Q$ have $\delta_H \geq \delta_Q$, we repeat following:
 \begin{itemize}
@@
 \end{itemize}
-
-This process terminates when $H$ is full and the top element in $H$ has $\delta$ less than the top element in $Q$, i.e.,\,the theoretically closest point left to be considered in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
+This process terminates when $H$ is full and the top-priority elements in $H$ and $Q$ are such that $\delta_H < \delta_Q$, i.e.,\,the theoretically closest point left is farther from the query than the $k^{th}$ nearest neighbor found so far.
 This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
 
-Note that this algorithm is not truly a depth-first traversal of $\mathcal{T}$ in the classical sense, because we use $Q$ to prioritize which branch of $\mathcal{T}$ we descend into.
-Indeed, we expect this algorithm to often switch which branch of $\mathcal{T}$ is being explored at greater depth.
+Note that this algorithm is not a depth-first traversal of $\mathcal{T}$ in the classical sense, because we use $Q$ to prioritize which branch we descend into.
+Indeed, we expect this algorithm to often switch the branch being explored.
 
@@
 For these methods we again use the terminology of tree-search and leaf-search.
-Tree-search navigates the cluster tree and adds clusters to $Q$.
+Tree-search navigates the cluster tree and finds $Q$.
 Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.
@@
 Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
+Let $d$ be the LFD in this region, and consider leaf clusters with cardinalities near $k$.
+Then the number of leaf-clusters in $Q$ is bounded above by $2d$, where the bound is achieved by having a cluster overlap the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
+In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search would look at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
+Thus, the asymptotic complexity is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
+For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|C|}}{k} \big) ^ {d^{-1}} \Big)^d \Big)$.
 
-Consider leaf clusters with cardinalities near $k$. Let $d$ be the LFD in this region, and $Q$ denote the set of candidate points and clusters.
-Then the number of leaf-clusters in $Q$ is bounded above by $2d$, where the bound is achieved if we have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
-In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
-Thus, the asymptotic complexity is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
-For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$.
-
-The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
-Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity of Breadth-First Sieve is $\mathcal{O} \Big( (T + L ) \log (T + L ) \Big) \label{eq:methods:breadth-first-sieve-complexity}$
-and the the time complexity of Depth-First Sieve search is $ \mathcal{O} \Big( T \log T + L \log k \Big).
-            \label{eq:methods:depth-first-sieve-complexity} $
+The asymptotic complexity of BFS is dominated by QuickSelect.
+Since QuickSelect is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity of BFS is $\mathcal{O} \big( (T + L ) \log (T + L ) \big)$ and that of DFS is $\mathcal{O} \big( T \log T + L \log k \big)$.
 
@@
 
-Based on our asymptotic analyses, we expect CAKES to perform well on datasets with low LFD and to scale sublinearly with dataset cardinality.
-To test this, we use datasets from the ANN-Benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to produce similar datasets with exponentially larger cardinalities.
-We perform the same procedure on a large dataset of uniformly distributed points in a hypercube, and compare CAKES with other algorithms on both the original and augmented datasets.
+Based on the asymptotic analysis, we expect CAKES to perform well on datasets with low LFD and scale sublinearly with cardinality.
+To test this, we use datasets from the ANN-Benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to have exponentially larger cardinalities.
+We perform the same procedure on a dataset of uniformly distributed points in a hypercube, and compare CAKES against other algorithms on both the original and augmented datasets.
 

### 2025-08-29 sections/3-datasets.tex
found a of space

--- 2025-08-27 sections/3-datasets.tex
@@
 
-In addition to ANN-Benchmarks datasets, we evaluate synthetic augmentations generated as in Sec.~\ref{sec:methods:synthetic-data}, using a noise tolerance $\epsilon=0.01$, and study scaling as the cardinality multiplier (“Mult.” in Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random}) increases.
-We also benchmark purely random datasets at various cardinalities; to match Sift we use a random dataset with base cardinality of $1{,}000{,}000$ and dimensionality $128$, and refer to this dataset as “Random.” This benchmark isolates the effect of manifold structure—expected to be absent in purely random data—on CAKES algorithms’ performance.
-
-To compute recall on the augmented datasets, we run linear search in Rust and write the results to disk.
-We verified on the ANN-Benchmark datasets that this linear search attains perfect recall using the test data provided.
-We then use the linear-search results on the augmentations as ground truth to evaluate CAKES in Rust and to compute recall for HNSW, ANNOY, and FAISS-IVF in Python.
+In addition to the ANN-Benchmarks datasets, we evaluate synthetic augmentations generated as in Sec.~\ref{sec:methods:synthetic-data}, using $\epsilon=0.01$, and study scaling behavior as the multiplier (“Mult.” in Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random}) increases.
+We also benchmark a random dataset at various cardinalities;
+we use a random dataset with base cardinality of $10^6$ and dimensionality $128$, and refer to this dataset as ``Random.''
+This benchmark isolates the effect of manifold structure—expected to be absent in purely random data—on CAKES algorithms.
 
@@
 
-To demonstrate CAKES with an exotic distance, we benchmark the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}, which contains rRNA sequences from 2,224,640 genomes; the longest sequence has 3,712 letters.
-We hold out 1,000 random sequences as queries.
-We build the tree and perform $k$-NN using Levenshtein distance~\cite{levenshtein1966binary} on unaligned sequences.
-Although the dataset is provided as a multiple sequence alignment of width 50{,}000, we intentionally use Levenshtein rather than Hamming on the aligned strings to demonstrate CAKES’s flexibility.
-Under Hamming, the embedding dimension would be 50{,}000; under Levenshtein, we report the dimensionality as the length of the longest sequence (3{,}712).
+To demonstrate CAKES with an exotic distance function, we use the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}, which contains rRNA sequences from 2,224,640 genomes;
+the longest sequence being 3,712 letters.
+Holding out 1,000 random sequences as queries, we build the tree and perform $k$-NN using Levenshtein distance~\cite{levenshtein1966binary} on \emph{unaligned} sequences.
+Although this dataset is a multiple sequence alignment of width 50,000, we do not use Hamming distance on the aligned strings so as to demonstrate CAKES's flexibility.
+Under Hamming, the embedding dimension would be 50,000;
+under Levenshtein, the dimensionality is 3,712.
 
@@
 
-We also benchmark CAKES on the Radio-ML dataset~\cite{oshea2018radioml} using Dynamic Time Warping~\cite{muller2007dynamic} as the distance.
-The dataset spans 24 modulation modes across 26 SNR levels from $-20$ to $30$\,dB, with 4{,}096 samples per combination, totaling $24\!\cdot\!26\!\cdot\!4096=2{,}555{,}504$ samples.
-Each sample is a 1{,}024-dimensional complex-valued vector (a time series).
-For experiments, we use the 10dB SNR subset with 97{,}304 samples and a 1{,}000-sample hold-out query set at the same SNR.
+We also use the Radio-ML dataset~\cite{oshea2018radioml} under Dynamic Time Warping~\cite{muller2007dynamic} as the metric.
+This dataset contains 24 modulation modes spanning 26 SNR levels from $-20$dB to $30$dB, with 4,096 samples per combination, yielding 2,555,504 samples.
+Each sample is a 1,024-dimensional complex-valued vector (a time-series).
+For the benchmarks, we use the 10dB SNR subset with 97,304 samples and a 1,000-sample hold-out set of queries.
 
@@
 \label{sec:dayasets:lfd-of-datasets}
+
+Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks (Figure~\ref{fig:results:lfd-plots}).
+In this section, when we discuss LFD, unless otherwise noted, we are referring to the 95th percentile from Figure~\ref{fig:results:lfd-plots}.
+
+For Fashion-MNIST, Figure~\ref{fig:results:fashion-mnist-lfd}, we observe that until depth 5, LFD is less than 4. 
+It then starts increasing, reaching a peak of about 6 near depth 20.
+In comparison, Glove-25 has even lower LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}. 
+In particular, Glove-25's LFD is less than 3 for all depths.
+Figure~\ref{fig:results:sift-lfd} shows the LFD for Sift, which is higher relative to both Fashion-MNIST and Glove-25, increasing sharply to a peak of 9 around depth 10.
+For Random, which has the same cardinality and dimensionality as Sift, LFD starts at 20 the root cluster, and all percentile values decrease linearly with depth until reaching the leaves of the tree.
+The spread in LFD starts narrow for the first few depths and widens as depth increases.
+Silva, Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD;
+less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
+For the Radio-ML dataset, Figure~\ref{fig:results:radioml-lfd} LFD values show three distinct peaks around an LFD of 12 near depths of 8, 25 and 50.
+Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
+Within each of the three spans of depths, the character of Radio-ML is similar to that of Random.
 
@@
 
-We benchmarked the three CAKES algorithms against a na\"ive linear search implementation in Rust.
-We also benchmarked against three state-of-the-art similarity search algorithms: HNSW, ANNOY, and FAISS-IVF in Python.
-We verified that our implementation of linear search produces the same neighbors as provided by the ANN-Benchmarks suite for Fashion-MNIST, Glove-25 and Sift datasets.
-We then used this linear search implementation to find and store the ground-truth for the augmented versions of the datasets.
-We used this ground-truth to calculate recall for CAKES's algorithms in Rust and for HNSW, ANNOY, and FAISS-IVF in Python.
+We benchmark the three CAKES algorithms against a na\"ive linear search implementation in Rust.
+We also benchmark against three state-of-the-art similarity search algorithms: HNSW, ANNOY, and FAISS-IVF in Python.
+We verified that our linear search produces the same results as in the ground truth of the ANN-Benchmarks suite for Fashion-MNIST, Glove-25 and Sift datasets.
+We then used this implementation to find and store the ground-truth for the augmented versions of those datasets, and use it to calculate recall for all algorithms.
 We plot the results of these benchmarks in Figure~\ref{fig:results:scaling-plots}.

### 2025-08-29 sections/4-results.tex
table formatting fixes
found  a little more space

@@
 
-% The algorithms presented in this paper rely heavily on the local fractal dimension (LFD) being much smaller than than the embedding dimension of the dataset.
-% As such, we begin by examining our assumptions about the LFD of real-world datasets (see Section~\ref{sec:results:lfd-of-datasets} and Figure~\ref{fig:results:lfd-plots}).
+For every dataset and distance funciton listed in Table~\ref{tab:datasets:summary}, we benchmark $k$-NN search using $k=10$.
+Smaller values of $k$, as described in Section~\ref{sec:introduction}, would be too sensitive to local perturbations, while larger values can capture too much of the global structure.
 
-For each dataset, we benchmark using the distance function listed in Table~\ref{tab:datasets:summary}.
-As described in Section~\ref{sec:introduction}, approximate search can be more sensitive to local perturbations than exact search, especially for small values of $k$, and hence we focus on $k=10$ for our benchmarks.
-
-In Figure~\ref{fig:results:scaling-plots}, we show how the throughput of each of CAKES's algorithms scales with cardinality for each of the six datasets we examined.
+In Figure~\ref{fig:results:scaling-plots}, we show how the throughput of each algorithm scales with cardinality for each of the six datasets we examined.
 To do this, we synthetically augmented the Fashion-MNIST, Glove-25, Sift, and Random datasets to create larger datasets using the process described in Section~\ref{sec:methods:synthetic-data}.
-For Silva and RadioML, due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset---rather than augmented versions of the dataset---to examine how performance scales with cardinality.
+For Silva and RadioML, due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset to examine how performance scales with cardinality.
 These plots also illustrate how the CAKES algorithms compare to na\"{i}ve linear search.
@@
 Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
+
+Throughput is measured in queries per second.
+A recall value of $1.000*$ denotes imperfect recall that rounds to $1.000$.
+In each of Tables \ref{tab:results:qps-and-recall-fmn},\ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, we observe that while recall for CAKES does \emph{not} degrade with cardinality, recall for
+HNSW and ANNOY does degrade with cardinality. CAKES exhibits perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
 
@@
 For Silva, throughput for all algorithms initially decreases approximately linearly with cardinality and then levels off at higher cardinalities; Depth-First Sieve is consistently fastest.
-For Radio-ML and Random, all CAKES variants are slower than na"{i}ve linear search, and their throughput decreases linearly with cardinality.
+For Radio-ML and Random, all CAKES variants are slower than na\"{i}ve linear search, and their throughput decreases linearly with cardinality.
 HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets; on Random, their recall is near zero.}
@@
 
-  Throughput is measured in queries per second.
-A recall value of $1.000*$ denotes imperfect recall that rounds to $1.000$.
-In each of Tables \ref{tab:results:qps-and-recall-fmn},\ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, we observe that while recall for CAKES does \emph{not} degrade with cardinality, recall for
-  HNSW and ANNOY does degrade with cardinality. CAKES exhibits perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
+% \begin{table}[t]
+%     \centering
+%     \caption{Fashion-MNIST: throughput and recall. See general notes above.}
+%     \label{tab:results:qps-and-recall-fmn}
+%     \small
+%     \setlength{\tabcolsep}{4pt}
+%     \begin{adjustbox}{width=\columnwidth,center}
+%     \begin{tabular}{@{} lcccccccc @{}}
+%     \toprule
+%     \textbf{M} &
+%     \multicolumn{2}{c}{\textbf{HNSW}} &
+%     \multicolumn{2}{c}{\textbf{ANNOY}} &
+%     \multicolumn{2}{c}{\textbf{FAISS-IVF}} &
+%     \multicolumn{2}{c}{\textbf{CAKES}} \\
+%     \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+%     & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
+%     \midrule
+%     1   & \num{1.33e4} & 0.95 & \num{2.19e3} & 0.95 & \num{2.01e3} & $1.00^{*}$ & \num{3.46e3} & 1.00 \\
+%     2   & \num{1.38e4} & 0.80 & \num{2.12e3} & 0.93 & \num{9.39e2} & $1.00^{*}$ & \num{3.68e3} & 1.00 \\
+%     4   & \num{1.66e4} & 0.68 & \num{2.04e3} & 0.90 & \num{4.61e2} & $1.00^{*}$ & \num{3.44e3} & 1.00 \\
+%     8   & \num{1.68e4} & 0.53 & \num{1.93e3} & 0.86 & \num{2.26e2} & $1.00^{*}$ & \num{3.30e3} & 1.00 \\
+%     16  & \num{1.87e4} & 0.49 & \num{1.84e3} & 0.86 & \num{1.17e2} & 0.99       & \num{3.34e3} & 1.00 \\
+%     32  & \num{1.56e4} & 0.54 & \num{1.85e3} & 0.78 & \num{5.91e1} & 0.99       & \num{2.96e3} & 1.00 \\
+%     64  & \num{1.50e4} & 0.38 & \num{1.78e3} & 0.69 & \num{2.61e1} & 0.97       & \num{3.25e3} & 1.00 \\
+%     128 & \num{1.49e4} & 0.36 & \num{1.66e3} & 0.54 & \num{1.33e1} & 0.96       & \num{2.96e3} & 1.00 \\
+%     256 & --           & --    & \num{1.60e3} & 0.59 & \num{6.65e0} & 0.96       & \num{2.79e3} & 1.00 \\
+%     512 & --           & --    & \num{1.83e3} & 0.58 & \num{3.56e0} & 0.95       & \num{2.84e3} & 1.00 \\
+%     \bottomrule
+%     \end{tabular}
+%     \end{adjustbox}
+%     \end{table}
 
@@
   \centering
-    \caption{Fashion-MNIST: throughput and recall. See general notes above.}
+  \begin{tiny}
+  \caption{Fashion-MNIST: throughput and recall.}
+  \end{tiny}
   \label{tab:results:qps-and-recall-fmn}
@@
   \begin{adjustbox}{width=\columnwidth,center}
-    \begin{tabular}{@{} lcccccccc @{}}
+  \begin{tabular}{@{} lllllllll @{}}
     \toprule
@@
   \centering
-    \caption{Glove-25: throughput and recall. See general notes above.}
+  \caption{Glove-25: throughput and recall.}
   \label{tab:results:qps-and-recall-glove}
@@
   \centering
-    \caption{Sift: throughput and recall. See general notes above.}
+  \caption{Sift: throughput and recall.}
   \label{tab:results:qps-and-recall-sift}
@@
     \begin{adjustbox}{width=\columnwidth,center}
-        \begin{tabular}{@{} lcccccccc @{}}
+    \begin{tabular}{@{} lllllllll @{}}
     \toprule
@@
   In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
-CAKES exhibits perfect recall at all cardinalities, whereas HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.
-    }
+  CAKES exhibits perfect recall at all cardinalities, whereas HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.}
   \label{tab:results:qps-and-recall-random}
@@
     \multicolumn{2}{c}{\textbf{CAKES}} \\
-    \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
+    \cmidrule(lr){2-3}\cmidrule(lllllllll){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
     & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\

### 2025-08-29 sections/5-discussion.tex
found  a little more space

@@
 We have presented CAKES, a suite of three algorithms for fast $k$-NN search across diverse distance functions.
-When the distance is a metric (Sec.~\ref{sec:methods}), the algorithms are exact; under cosine distance (non-metric) they achieve nearly perfect recall.
-CAKES is most effective when data satisfy the manifold hypothesis—lying on a low-dimensional manifold within a high-dimensional space.
-It builds an unbalanced-clustering binary tree that captures manifold structure and accelerates search.
+When the distance is a metric (Sec.~\ref{sec:methods}), the algorithms are exact.
+Under Cosine distance, a non-metric, they achieve nearly perfect recall.
+CAKES is most effective when data satisfy the manifold hypothesis, i.e., they occupy a low-dimensional manifold within a high-dimensional space. 
+It builds an unbalanced cluster tree that captures manifold structure and helps accelerate search.
 Conversely, performance degrades on randomly distributed data, where such structure is absent.
-In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we texted exhibits a manifold structure, as quantified by the percentiles of LFD.
+In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
 
@@
 
-As expected, Fig.~\ref{fig:results:scaling-plots} shows CAKES scaling sublinearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD; Sift versus Random illustrates this contrast.
-
-On Random, CAKES throughput decreases linearly with the multiplier, whereas on Sift, Depth-First Sieve and Breadth-First Sieve are nearly constant.
+As expected, Fig.~\ref{fig:results:scaling-plots} shows CAKES scaling sublinearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD;
+Sift versus Random illustrates this contrast.
+On Random, CAKES' throughput decreases linearly with increasing cardinality, whereas on Sift, Depth-First Sieve and Breadth-First Sieve have near-constant throughput.
 Repeated $\rho$-NN is not near-constant on Sift but decays much more slowly than on Random.
 Because Sift and Random share cardinality and embedding dimension, this isolates the effect of manifold structure on performance.
+Although CAKES is slower on Random, it maintains perfect recall.
 
-Although CAKES is slower on Random, its recall remains perfect.
-HNSW and ANNOY maintain near-constant throughput as the multiplier grows, but their recall declines, and their index build time and memory rise, making them unsuitable when datasets grow exponentially.
-CAKES’s tree builds orders of magnitude faster with far less memory; while queries are slower than HNSW/ANNOY, CAKES attains near-constant scaling with perfect recall.
+HNSW and ANNOY maintain near-constant throughput as the cardinality grows, but their recalls decline and their index build time and memory rise, making them unsuitable when datasets grow exponentially.
+CAKES's tree builds orders of magnitude faster with far less memory; while queries are slower than HNSW/ANNOY, CAKES attains near-constant scaling with perfect recall.
 
-The fastest CAKES variant depends on dataset and cardinality.
-On Fashion-MNIST, Glove-25, and Sift, Depth-First Sieve and Breadth-First Sieve show nearly constant throughput as cardinality increases and consistently outperform linear search at high cardinalities.
-This variability motivates auto-tuning (Sec.~\ref{sec:methods:auto-tuning}); future work will better characterize when each algorithm excels and enable more sophisticated selection from dataset properties.
-
-We stress that CAKES targets \emph{big} data, and our results support this claim; we
-observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes
-linear search at a large enough cardinality for each of the ANN-benchmark datasets we tested. 
-Moreover, we observe
-that CAKES’s algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD. For example,
-Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at
-a cardinality of $10^{7}$ 
-for Sift, as opposed to about $10^{5}$
-and $10^{6}$ 
-for Fashion-Mnist and Glove-25, respectively. For the
-Random dataset, which has the highest LFD, CAKES’s algorithms \emph{never} outperform linear search.
+We stress that CAKES targets \emph{big} data, and our results support this claim.
+We observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmark datasets we tested.
+Moreover, we observe that CAKES's algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD.
+For example, Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^{7}$  for Sift, as opposed to about $10^{5}$ and $10^{6}$ for Fashion-Mnist and Glove-25, respectively.
+For the Random dataset, which has the highest LFD, CAKES's algorithms \emph{never} outperform linear search.
 
@@
 
-\subsection{Applicability}
+\subsection{Future Work}
 
-Despite CAKES’s generality, effective deployment on real data can require domain knowledge.
-In protein sequence search, sequences vary widely in length (typically 30–1000 amino acids, with outliers) over a 20-letter alphabet.
+This study can be extended in several directions.
+We can start by evaluating more datasets and distance functions not supported by FAISS, HNSW, or ANNOY, such as Wasserstein~\cite{vallender1974calculation} for probability distributions (particularly high-dimensional) and Tanimoto~\cite{bajusz2015tanimoto} for molecular structures.
+
+We can also improve the augmentation procedure (Sec.~\ref{sec:methods:synthetic-data}) to better preserve topological structure of datasets, e.g., by favoring perturbations along the leading principal components of the local manifold.
+This would better test the scaling behavior of various algorithms.
+
+We have noted how the fastest algorithm in CAKES varies depending on the dataset and distance function being used, as well as the cardinality of the dataset.
+This variability motivated our auto-tuning process (Sec.~\ref{sec:methods:auto-tuning}).
+Future work should examine why and when each algorithm excels and learn to use the geometric and topological properties of datasets to choose the ideal algorithm to use.
+
+CAKES algorithms can also be extended for use in a streaming environment, enabling online updates to the tree as points are inserted or deleted.
+
+Despite CAKES' generality, effective deployment on real data requires domain knowledge.
+In protein sequence search, for example, sequences vary widely in length (typically 30-1000 amino acids, with outliers) and have a 20-letter alphabet.
 Numerous algorithms exist~\cite{kim2021entrance, daniels2013compressive, yu2015entropy, steinegger2018clustering}, and their development relied on deep domain expertise.
-
 Simple edit distance can be poor when lengths vary substantially.
 CAKES can get around that issue by instead using alternative metrics or preprocessing—for example, represent each sequence as a set of $k$-mers and use Jaccard distance, as in~\cite{kim2021entrance}.
-If the distance relies on global alignment (e.g., Needleman–Wunsch~\cite{needleman1970general}), the dataset can be partitioned into mutually exclusive length bins with a separate CAKES tree per bin; at query time, route the sequence to its bin and a small number of adjacent bins.
-
-\subsection{Future Work}
-
-This study suggests several directions.
-First, we want to extended our evaluation to more datasets and distance functions not supported by FAISS, HNSW, or ANNOY, including Wasserstein distance~\cite{vallender1974calculation} for probability distributions (particularly high-dimensional) and Tanimoto distance~\cite{bajusz2015tanimoto} for comparing molecular structures via maximal common subgraphs.
-Adding a new distance to CAKES requires only a Rust implementation.
-
-Second, we intend to improve the augmentation procedure (Sec.~\ref{sec:methods:synthetic-data}) to better preserve topological structure of datasets, e.g., by favoring perturbations along the leading principal components of the local manifold.
-
-Third, we plan to explore CAKES algorithms' use in a streaming environment, enabling online updates to the tree as points are inserted or deleted.
-
-Finally, extend anomaly detection in CHAODA~\cite{ishaq2021clustered} by adding graph-based methods that use the distribution of distances among the $k$ nearest neighbors of cluster centers.
+If the distance relies on global alignment (e.g., Needleman-Wunsch~\cite{needleman1970general}), the dataset can be partitioned into mutually exclusive length bins with a separate tree per bin;
+at query time, we simply route the sequence to its bin by length, alogn with a small number of adjacent bins.
 
@@
 
-CLAM and CAKES are implemented in Rust and released under the MIT license at https://github.com/URI-ABD/clam.
+CLAM and CAKES are implemented in Rust and available under the MIT license at https://github.com/URI-ABD/clam.

### 2025-08-29 sections/4-results.tex
added balanced vs unbalanced to results

@@
 \label{sec:results}
+
+\subsection{Scaling Behavior and Recall}
+\label{sec:results:scaling-behavior-and-recall}
 
@@
 % \end{figure}
+
+\subsection{Clustering Strategies and Number of Distance Computations}
+\label{sec:results:clustering-strategies-and-number-of-distance-computations}
+
+In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
+These strategies are the Cartesian product of balanced vs.
+unbalanced clustering, and the presence vs.
+absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
+To help with this analysis, we also added instrumentation to our implementation of CAKES to count the number of distance computations during search.
+We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.
+
+Figure~\ref{fig:results:distance-counts}.
+shows these four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is called Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset.
+Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
+the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots} ( which shows results using PermutedBall), no longer do so under a balanced clustering.
+We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.
+
+\begin{figure}[h]
+  \centering
+  \subfloat[Repeated $\rho$-NN]{
+  \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
+  \label{fig:results:fashion-mnist-rnn-throughput}
+  }
+  \subfloat[Repeated $\rho$-NN]{
+    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}    
+    \label{fig:results:fashion-mnist-rnn-counts}
+  }
+
+  \subfloat[Breadth First Sieve]{
+    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
+    \label{fig:results:fashionmnist-bfs-throughput}
+    }
+  \subfloat[Breadth First Sieve]{
+    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
+    \label{fig:results:fashionmnist-bfs-counts}
+    }
+
+  \subfloat[Depth First Sieve]{
+    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
+    \label{fig:results:fashionmnist-dfs-throughput}
+    }
+  \subfloat[Depth First Sieve]{
+    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
+    \label{fig:results:fashionmnist-dfs-counts}
+  }
+
+  \vspace{2pt}
+  \includegraphics[width=1\columnwidth]{images/distance_counts/legend.png}
+
+  \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
+  Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
+  The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
+  The x-axis represents increasing cardinality of the dataset.}
+  \label{fig:results:distance-counts}
+\end{figure}

### 2025-08-29 sections/4-results.tex
acknowleged unbalanced vs balanced section in discussion

@@
 
-In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
-These strategies are the Cartesian product of balanced vs.
-unbalanced clustering, and the presence vs.
-absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
-To help with this analysis, we also added instrumentation to our implementation of CAKES to count the number of distance computations during search.
-We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.
+Beyond scaling behavior, we examine how clustering strategy affects search performance.
+We consider four strategies—the Cartesian product of balanced vs.\ unbalanced clustering and the presence vs.\ absence of depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}).
+We instrument CAKES to count distance computations during search; this adds substantial overhead and would not be used in real-world applications.
 
-Figure~\ref{fig:results:distance-counts}.
-shows these four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is called Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset.
-Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
-the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots} ( which shows results using PermutedBall), no longer do so under a balanced clustering.
-We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.
+Figure~\ref{fig:results:distance-counts} compares \textit{Ball}, \textit{BalancedBall}, \textit{PermutedBall}, and \textit{PermutedBalancedBall} (``Permuted'' denotes depth-first reordering) for all three algorithms on Fashion-MNIST.
+Balanced clustering is strictly worse in throughput and has poorer asymptotics: algorithms such as Depth-First Sieve that appear nearly constant-time in Fig.~\ref{fig:results:scaling-plots} (with \textit{PermutedBall}) no longer do so under balanced clustering.
+On this dataset, depth-first reordering has little effect on throughput; it is primarily intended to improve space complexity.
 

### 2025-08-29 sections/2-methods.tex
added reference to balanced vs unbalanced in methods

@@
 % This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
-We also note here that the $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
-For real datasets, we expect the process from Section~\ref{sec:methods:clustering:building-the-tree} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
-The only case in which we would expect a balanced tree is if the dataset were uniformly distributed in a $D$-dimensional space.
+
+The $\mathcal{O}(n\log n)$ build complexity assumes a balanced tree $\mathcal{T}$.
+In practice, Algorithm~\ref{alg:methods:partition} yields unbalanced trees on real data because of nonuniform sampling densities and the manifold’s low-dimensional structure.
+A balanced tree is expected only for uniformly distributed data (e.g., in a $d$-dimensional hypercube).
+
+Although real trees may be deeper than $\log n$, which may appear to increase the cost of tree-building, many leaves appear at shallow depths, so later levels contain fewer points and their per-level build costs decline.
+A general analysis for unbalanced trees is dataset-dependent and is not provided; instead, we report empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
+In any case, the tree is built once, so this cost is amortized over all queries and subsequent applications.
 

### 2025-08-29 sections/2-methods.tex
small mistake

@@
 The $\mathcal{O}(n\log n)$ build complexity assumes a balanced tree $\mathcal{T}$.
-In practice, Algorithm~\ref{alg:methods:partition} yields unbalanced trees on real data because of nonuniform sampling densities and the manifold’s low-dimensional structure.
+In practice, our clustering algorithm yields unbalanced trees on real data because of nonuniform sampling densities and the manifold’s low-dimensional structure.
 A balanced tree is expected only for uniformly distributed data (e.g., in a $d$-dimensional hypercube).

### 2025-08-29 sections/1-introduction.tex
back down to 10 pages with some extra material

@@
 Several algorithms have been developed to tackle $k$-nearest neighbor search on large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+
+HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
+IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
+ANNOY~\cite{annoy} uses random projection and tree building.
 However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).
-
-Hierarchical Navigable Small World networks~\cite{malkov2016hnsw} rely on navigable small world (NSW) networks and skip lists.
-InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
-ANNOY~\cite{annoy} uses random projection and tree building.
-At each intermediate node of the tree, a random hyperplane, defined by two sampled points, splits the space.
-Increasing the number of such trees improves recall but increases search time.
 
@@
 This paper introduces CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
-We provide a theoretical analysis of their time complexity in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
+We provide a theoretical complexity analysis in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
-Given that CAKES's algorithms are intended for datasets with a manifold structure, complexity analyses assuming a uniform distribution would be uninformative.
-As a result, our analysis assumes datasets with manifold structure—characterized by low fractal dimension and metric entropy—reflecting the conditions for which CAKES is designed.
+Given that CAKES's algorithms are intended for datasets with a manifold structure, an analysis assuming a uniform distribution would be uninformative.
+As a result, our analysis assumes datasets with manifold structure, reflecting the conditions for which CAKES is designed.

### 2025-08-29 sections/2-methods.tex
back down to 10 pages with some extra material

--- 2025-08-29 sections/2-methods.tex
@@
 Cosine distance violates the triangle inequality and is not a metric.
-As in~\cite{yu2015entropy}, all CAKES' algorithms are exact under distance metrics.
-
-% cosine distance is not a metric because it violates the triangle inequality (e.g.,\,consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).
-
 The appropriate distance function depends on the data: Euclidean or Cosine are often used for neural embeddings, Levenshtein or Hamming for sequences, and DTW or Wasserstein for time-series.
+CAKES provides exact search under distance metrics.
 
@@
 For a point $q \in \textbf{X}$ and radii $r_1 > r_2$,
-
 \begin{equation}
-    \text{LFD}(q, r_1, r_2) = \frac{\text{log} \left( \frac{|B(q, r_1)|}{|B(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
+    \text{LFD}(q, r_1, r_2) = \frac{\text{log} \Big( \frac{|B(q, r_1)|}{|B(q, r_2)|} \Big) }{\text{log} \big( \frac{r_1}{r_2} \big) }
     \label{eq:methods:lfd-original}
@@
 The center is the geometric median—the point minimizing the sum of distances to all other points—of the points in $C$;
-in practice, for large $|C|$ we find the geometric median of a random subsample of $\sqrt{|C|}$ points from $C$.
-The radius is the maximum distance from the center to any point in $C$.
+in practice, for large $|C|$ we only compute this for a smaller sample.
+A cluster's radius is the maximum distance from the center to any point in it.
 Each non-leaf cluster has two children (as in a binary tree).
@@
 
-Unless stated otherwise, we estimate the LFD of clusters at scales $r$ and $r/2$ via Eq.~\ref{eq:methods:lfd-half}, using only the points in $C(c,r)$ (not all of $B(c,r)$).
+We estimate the LFD of clusters at scales $r$ and $r/2$ via Eq.~\ref{eq:methods:lfd-half}, using only the points in $C(c,r)$ (not all of $B(c,r)$).
 
@@
 We construct $\mathcal{T}$ using CLAM, following CHESS~\cite{ishaq2019clustered} but with improved pole selection, then apply a depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}) on the dataset.
+The procedure is reproduced in Alg.~\ref{alg:methods:partition} for completeness.
 
-The recursive step is as follows.
-For a cluster $C$ with $|C|$ points, we draw a random subsample $S\subset C$ of $\sqrt{|C|}$ points, compute all pairwise distances in $S$, and take the geometric median of $S$ as the \emph{center} of $C$.
-The point in $C$ farthest from the center is the \emph{left pole}.
-The distance from the center to the left pole if the \emph{radius}.
-The point in $C$ farthest from the left pole is the \emph{right pole}.
-We partition $C$ into \emph{left} and \emph{right child} clusters containing points closer to the left and right poles respectively, with ties going to the left child.
+Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} gives us $\mathcal{T}$.
+During this procedure, we also compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
 
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, we recurse until each leaf contains a single point or some user-specified stopping criterion is met (e.g., minimum cluster radius, maximum tree depth, etc.).
-During partitioning, we compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
+\begin{algorithm} % enter the algorithm environment
+    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
+    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
+    \begin{algorithmic} % enter the algorithmic environment
+    \REQUIRE $f$, a distance function
+    \REQUIRE $C$, a cluster
+    \REQUIRE $criteria$, user-specified continuation criteria
 
-% \begin{algorithm} % enter the algorithm environment
-%     \caption{Partition($C$, $criteria$)} % give the algorithm a caption
-%     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
-%     \begin{algorithmic} % enter the algorithmic environment
-%     \REQUIRE $f: X \times X \mapsto \mathbb{R}^+ \cup \{0\}$, a distance function
-%     \REQUIRE $C$, a cluster
-%     \REQUIRE $criteria$, user-specified continuation criteria
+    \STATE $S \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+    \STATE $c \Leftarrow$ geometric median of $S$
+    \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+    \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+    \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+    \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
 
-%     \STATE $seeds \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-%     \STATE $c \Leftarrow$ geometric median of $seeds$
-%     \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-%     \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-%     \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-%     \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+    \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
+        \STATE Partition($L$, $criteria$)
+    \ENDIF
 
-%     \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
-%         \STATE Partition($L$, $criteria$)
-%     \ENDIF
-
-%     \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
-%         \STATE Partition($R$, $criteria$)
-%     \ENDIF
-% \end{algorithmic}
-% \end{algorithm}
+    \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
+        \STATE Partition($R$, $criteria$)
+    \ENDIF
+\end{algorithmic}
+\end{algorithm}
 
@@
 This list was used to retrieve the clusters' points during search.
-Although this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
-With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree $\mathcal{T}$.
-Assuming $\mathcal{T}$ is balanced, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.
+With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, it stored $n$ indices at each depth in $\mathcal{T}$.
+Assuming $\mathcal{T}$ is balanced, and thus $\mathcal{O}(\log n)$ depth, this approach had a prohibitively high memory overhead of $\mathcal{O}(n \log n)$.
 In this work, we introduce a new approach wherein, after building $\mathcal{T}$, we reorder the dataset so that points are stored in a depth-first order.
-Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset.
+Then, each cluster need only store its \textit{cardinality} and an \textit{offset} to access its points.
 The root cluster $\mathcal{R}$ has an offset of zero and a cardinality equal to the number of points in the dataset.
@@
 This is a significant improvement over the exact approach, which costs $\mathcal{O}(n^2 \log n)$.
+The $\mathcal{O}(n\log n)$ build complexity assumes a balanced tree $\mathcal{T}$.
+In practice, Alg.~\ref{alg:methods:partition} yields unbalanced trees on real data because of non-uniform sampling densities and the manifold's low-dimensional structure.
+A balanced tree is expected only for uniformly distributed data (e.g., in a $d$-dimensional hypercube).
 
-% This complexity analysis ignores the cost of checking the continuation criteria.
-% While a user could specify criteria of any complexity (our implementation allows for this), for this paper we continue clustering until each cluster is a singleton, i.e.,\,it contains only one point or duplicates of the same point and has a radius of zero.
-% This criterion costs $\mathcal{O}(1)$ to check per cluster, and so does it not affect the overall complexity of the algorithm as used in this paper.
-We also note here that the $\mathcal{O}(n \log n)$ complexity assumes that $\mathcal{T}$ is balanced.
-For real datasets, we expect the process from Section~\ref{sec:methods:clustering:building-the-tree} to produce anything but a balanced tree; the varying sampling density in different regions of the manifold and the low dimensional ``shape'' of the manifold itself will cause it to be unbalanced.
-The only case in which we would expect a balanced tree is if the dataset were uniformly distributed in a $D$-dimensional space.
-
-% When building a tree for a real dataset, the depth would be larger than $\log n$, which would seem to increase the cost of tree-building.
-% However, because of the large numbers of leaf clusters at shallow depths, each subsequent level of $\mathcal{T}$ would also have fewer points, and so the cost of building each subsequent level would be lower.
-% Any analysis of unbalanced trees would be highly dependent on the specific dataset, and so we do not provide one here.
-% We do, however, explore empirical results based on forcing a balanced tree in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
-% In any case, the tree is built only once, so this cost is amortized over all search queries and other applications.
+Although (unbalanced) trees are deeper than $\log n$, which may appear to increase the cost of tree-building, many leaves appear at shallow depths, so later levels contain fewer points and the per-level build cost declines.
+Any analysis for unbalanced trees would be dataset-dependent, so we only report empirical results based on balanced version of Alg.~\ref{alg:methods:partition} in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
+In any case, once the tree is built, its cost is amortized over all queries and subsequent applications.
 
@@
     \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
+    \vskip -0.1in
     \caption{
@@
                 \STATE $[L, R]$ $\Leftarrow$ \textit{children} of $C$
-                \STATE \textbf{return} tree-search($L, q, r$) \\
-                 $\cup$ tree-search($R, q, r$)
+                \STATE \textbf{return} tree-search($L, q, r$) $\cup$ tree-search($R, q, r$)
             \ENDIF
@@
 Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-In these algorithms, we use $H$, for \textit{hits}, to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being among the $k$ nearest neighbors.
-These algorithms also use some terminology defined in Figure~\ref{fig:methods:deltas}.
+In these algorithms, we use $H$ to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being among the $k$ nearest neighbors.
+These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
 
@@
 
-This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
+This algorithm relies on the \textit{tree-search} (Alg.~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Alg.~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
 
-For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by the cardinality of the dataset.
-If no clusters are found, we keep doubling $r$ and repeating tree-search, until we find a a non-empty set, $Q$, of clusters.
+We begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by $|X|$.
+If no clusters are found, we double $r$ and repeat tree-search, until we find a a non-empty set, $Q$, of clusters.
 
-Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+Now, so long as $\sum_{C \in Q} |C| < k$, we continue to repeat tree-search, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
 In particular, we increase the radius by a factor of
 \begin{equation}
-    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu} \right)
+  \bigg( {\frac{k}{\sum_{C \in Q} |C|}} \bigg)^{\mu}
   \label{eq:methods:repeated-rnn-factor}
@@
 We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
-We cap this factor at 2 to ensure that we do not increase the radius too quickly in any single iteration.
+In practice, we cap this factor at 2 to ensure that we do not increase the radius too quickly in any single iteration.
 
 Intuitively, the factor by which we increase the radius should be \textit{inversely} related to the number of points found so far.
-When the LFD at the radius scale from the previous iteration is high, this suggests that the data are densely populated in that region.
-Thus, a small increase in the radius would likely encounter many more points, so a smaller radial increase would suffice to find $k$ neighbors.
-Conversely, when the LFD at the radius scale from the previous iteration is low, this suggests that the data are sparsely populated in that region.
-In such a region, a small increase in the radius would likely encounter vacant space, so a larger radial increase is needed.
-Thus, the factor of radius increase should also be \textit{inversely} related to the LFD.
+When the LFD at a given radius scale is high, the data are dense in that region.
+Thus, a small increase in the radius would likely encounter many more points, and find $k$ neighbors.
+Conversely, when the LFD at that scale is low, the data are sparse.
+In such a region, a small increase in the radius would likely encounter vacant space, so a larger increase is needed.
+Thus, the factor of radial increase should be \textit{inversely} related to the LFD.
 However, we should not increase the radius too drastically with any one iteration because we assume that the LFD is only \textit{locally} uniform.
@@
 
-To extend Equation~\ref{eq:methods:rnn-search-complexity} to Repeated $\rho$-NN, we must first estimate the number of iterations of tree-search (Algorithm~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees at least $k$ neighbors.
+To extend Eq.~\ref{eq:methods:rnn-search-complexity} to Repeated $\rho$-NN, we must estimate the number of iterations of tree-search (Alg.~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees $k$ neighbors.
 Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
@@
         }_{\textrm{tree-search}}
-        \ + \
+        +
         \underbrace{
-            \overbrace{k}^{\textrm{output size}} \cdot
+            k \cdot
             \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
@@
 
-We begin by letting $Q$ be a set of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
+We begin by letting $Q$ be a list of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
 During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
@@
 We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$.
-Since this step may require a binary search for the correct pivot element to find $\tau$ and reordering with a new pivot takes linear time in the size of the input list, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
+Since this step may require a binary search of $Q$ to find $\tau$ and reordering $Q$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
 
-Next, we go over all items in $Q$, skipping over any for which $\delta^{-} > \tau$ because such elements cannot contain (or be) one of the $k$ nearest neighbors.
+Next, we check every item in $Q$, removing any for which $\delta^{-} > \tau$ because they cannot contain (or be) one of the $k$ nearest neighbors.
 If an item corresponds to a point, we keep it.
 If it corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
-Otherwise (since it must correspond to a non-leaf cluster), we to $Q$ the pairs of 3-tuples corresponding to its child clusters.
+Otherwise (since it must correspond to a non-leaf cluster), we add to $Q$ the pairs of 3-tuples corresponding to its child clusters.
 We continue this process until the sum of multiplicities in $Q$ is exactly $k$.
-We then use the QuickSelect algorithm one last time to reorder $Q$ and return the $k$ nearest neighbors.
+These are the $k$ nearest neighbors.
 
@@
 $Q$ starts containing only $\mathcal{R}$ while $H$ starts empty.
-So long as $H$ is not full or the top-priority elements from $H$ and $Q$ have $\delta_H \geq \delta_Q$, we repeat following:
+So long as $H$ is not full or the top-priority items from $H$ and $Q$ have $\delta_H \geq \delta_Q$, we repeat following:
 \begin{itemize}
-    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
+    \item While the top-priority cluster is not a leaf, remove it from $Q$ and add its children to $Q$.
+    \item Remove the top-priority cluster (a leaf) from $Q$ and add all its points to $H$.
 \end{itemize}
-This process terminates when $H$ is full and the top-priority elements in $H$ and $Q$ are such that $\delta_H < \delta_Q$, i.e.,\,the theoretically closest point left is farther from the query than the $k^{th}$ nearest neighbor found so far.
-This leaves $H$ containing exactly the $k$ nearest neighbors to the query.
-
-Note that this algorithm is not a depth-first traversal of $\mathcal{T}$ in the classical sense, because we use $Q$ to prioritize which branch we descend into.
-Indeed, we expect this algorithm to often switch the branch being explored.
+This process terminates when $H$ is full and the top-priority items in $H$ and $Q$ are such that $\delta_H < \delta_Q$, i.e.,\,the theoretically closest point left is farther from the query than the $k^{th}$ nearest neighbor found so far.
+Thus, $H$ contains the $k$ nearest neighbors to the query.
 
@@
 In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search would look at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
-Thus, the asymptotic complexity is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
-For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|C|}}{k} \big) ^ {d^{-1}} \Big)^d \Big)$.
+Thus, the asymptotic complexity of tree-search is
+\begin{equation*}
+    T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big).
+\end{equation*}
+For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the complexity is
+\begin{equation*}
+    L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|C|}}{k} \big) ^ {d^{-1}} \Big)^d \Big).
+\end{equation*}
 
 The asymptotic complexity of BFS is dominated by QuickSelect.
-Since QuickSelect is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity of BFS is $\mathcal{O} \big( (T + L ) \log (T + L ) \big)$ and that of DFS is $\mathcal{O} \big( T \log T + L \log k \big)$.
+Since QuickSelect is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity of BFS is
+\begin{equation}
+    \mathcal{O} \big( (T + L ) \log (T + L ) \big),
+    \label{eq:methods:bfs-complexity}
+\end{equation}
+and that of DFS is 
+\begin{equation}
+    \mathcal{O} \big( T \log T + L \log k \big).
+    \label{eq:methods:dfs-complexity}
+\end{equation}
 

### 2025-08-29 sections/4-results.tex
back down to 10 pages with some extra material

@@
 
-% \begin{figure}
-%     \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/lfd/fashion-mnist.png}\\
-%         \subcaption{Fashion-MNIST}
-%         \label{fig:results:fashion-mnist-lfd}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/lfd/glove-25.png}\\
-%         \subcaption{Glove-25}
-%         \label{fig:results:glove-25-lfd}
-%     \end{subfigure}
-%     \\
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/lfd/sift.png}\\
-%         \subcaption{Sift}
-%         \label{fig:results:sift-lfd}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/lfd/random.png}\\
-%         \subcaption{A random dataset}
-%         \label{fig:results:random-lfd}
-%     \end{subfigure}
-%     \\
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/lfd/silva-SSU-Ref.png}\\
-%         \subcaption{Silva 18S}
-%         \label{fig:results:silva-lfd}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/lfd/radio-ml.png}\\
-%         \subcaption{RadioML}
-%         \label{fig:results:radioml-lfd}
-%     \end{subfigure}%
-%     \\
-%     \vskip -0.1in
-%     \begin{subfigure}[b]{0.94\textwidth}
-%         \centering
-%         \includegraphics[width=0.7\textwidth]{images/lfd/legend.png}
-%         \label{fig:results:lfd-legend}
-%     \end{subfigure}%
-%     \vskip -0.1in
-%     \caption{Local fractal dimension vs.
-cluster depth across six datasets.
-The `random' dataset is randomly generated according to the procedure in Section~\ref{sec:datasets-and-benchmarks:random-datasets}; note that the y-axis is different for this dataset.
-In each plot, the horizontal axis denotes depth in the cluster tree, and the vertical axis denotes the LFD of clusters at that depth.
-We show lines for the 5$^{th}$, 25$^{th}$, 50th, 75$^{th}$ and 95$^{th}$ percentiles of LFD, as well as the minimum and maximum LFD at each depth.
-So that plots best reflect the distribution of LFDs across the entire \textit{dataset}, we count each cluster as many times as its cardinality.
-For example, if, for some dataset, the 95$^{th}$ percentile of LFD at depth 40 is 3, this means that 95\% of the points in clusters at depth 40 belong to a cluster whose LFD is at most 3.}
-%     \label{fig:results:lfd-plots}
-%     \vskip -0.4in
-% \end{figure}
+\subsection{Indexing and Tuning Time}
+\label{sec:results:indexing-and-tuning-time}
 
-% \subsection{Indexing and Tuning Time}
-% \label{sec:results:indexing-and-tuning-time}
-
-% For each of the ANN-benchmark datasets and the Random dataset, we report the time taken for each algorithm to build the index and to tune the hyper-parameters for these indices to achieve the highest possible recall.
-For the sake of brevity, these results are reported in the supplement.
-
-% We benchmark the CAKES algorithms, na\"{i}ve linear search, HNSW, ANNOY, and FAISS-IVF on the Fashion-MNIST, Glove-25, Sift, and Random datasets.
-% We augment these datasets with synthetic points to examine how performance scales with cardinality.
-% We also benchmark the CAKES algorithms on the Silva and RadioML datasets, and we subsample these datasets instead of augmenting them to examine how performance scales with cardinality.
-
-% On the Fashion-MNIST, Glove-25, and Sift datasets (in Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} respectively), we observe that as cardinality increases, the Depth-First Sieve algorithm is consistently the fastest CAKES algorithm with a throughput that is constant in the cardinality of the dataset.
-% The Breadth-First Sieve algorithm is the usually the second fastest, also with a nearly constant throughput across all cardinalities.
-% The Repeated $\rho$-NN algorithm, however, falls off in throughput as cardinality increases.
-% All three CAKES algorithms exhibit perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
-% In contrast, HNSW and ANNOY are faster than CAKES's algorithms for all cardinalities and have near constant throughput as cardinality increases, but their recall degrades quickly as cardinality increases.
-% FAISS-IVF exhibits linearly decreasing throughput as cardinality increases, which is expected from the algorithm given that we tune the hyper-parameters to maximize recall.
-
-% On the Random dataset, HNSW and ANNOY are still the fastest algorithms but exhibit recall values near 0.
-% CAKES's algorithms show linearly decreasing throughput as cardinality increases and are also slower than na\"{i}ve linear search.
+For each of the ANN-benchmark datasets and Random, Fig.~\ref{fig:results:indexing} reports the time taken for each algorithm to build its index and to tune any hyper-parameters to maximize recall.
 
@@
   \centering
-  \subfloat[Fashion-MNIST for $k=10$.]{
+  \vskip -0.2in
+  \subfloat[Fashion-MNIST]{
+    \includegraphics[width=0.24\textwidth]{images/indexing/fashion-mnist-indexing.png}
+    \label{fig:results:fashion-mnist-indexing}
+  }
+  \subfloat[Glove-25]{=
+    \includegraphics[width=0.24\textwidth]{images/indexing/glove-25-indexing.png}
+    \label{fig:results:glove-25-indexing}
+  }
+
+  \vskip -0.1in
+  \subfloat[Sift for $k=10$.]{
+    \includegraphics[width=0.24\textwidth]{images/indexing/sift-indexing.png}
+    \label{fig:results:sift-indexing}
+  }
+  \subfloat[Random dataset for $k=10$.]{
+    \includegraphics[width=0.24\textwidth]{images/indexing/random-indexing.png}
+    \label{fig:results:random-indexing}
+  }
+  \caption{Indexing and tuning time.
+  We report the time taken by various algorithms to build and tune its index on various datasets.
+  For CAKES, this includes the time required to build  the tree and perform auto-tuning.
+  All axes are on logarithmic scales.
+  For all datasets, we see that the indexing time for CAKES increases roughly linearly with the cardinality of the dataset.
+  HNSW and ANNOY are the slowest at indexing.
+  FAISS-Flat does not actually build an index;
+  it provides a na\"{i}ve linear search in the FAISS family of algorithms.
+  It is included only to illustrate the base cost of simply having a dataset.}
+  \label{fig:results:indexing}
+\end{figure}
+
+\begin{figure}[h]
+  \vskip -0.2in
+  \centering
+  \subfloat[Fashion-MNIST]{
     \includegraphics[width=0.48\columnwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
@@
   }
-  \subfloat[Glove-25 for $k=10$.]{
+  \subfloat[Glove-25]{
     \includegraphics[width=0.48\columnwidth]{plots/glove-25_PermutedBall_10_throughput.png}
@@
 
-  \subfloat[Sift for $k=10$.]{
+  \vskip -0.1in
+  \subfloat[Sift]{
     \includegraphics[width=0.48\columnwidth]{plots/sift_PermutedBall_10_throughput.png}
@@
   }
-  \subfloat[Random dataset for $k=10$.]{
+  \subfloat[Random dataset]{
     \includegraphics[width=0.48\columnwidth]{plots/random_PermutedBall_10_throughput.png}
@@
 
-  \subfloat[Silva for $k=10$.]{
+  \vskip -0.1in
+  \subfloat[Silva]{
     \includegraphics[width=0.48\columnwidth]{plots/silva-SSU-Ref_PermutedBall_10_throughput.png}
@@
   }
-  \subfloat[RadioML for $k=10$ at SnR = 10dB.]{
+  \subfloat[RadioML at SnR = 10dB.]{
     \includegraphics[width=0.48\columnwidth]{plots/radio-ml_Ball_10_throughput.png}
@@
 
-  \vspace{2pt}
-  \includegraphics[width=1\columnwidth]{plots/legend.png}
+  \vskip 0.05in
+  \includegraphics[width=0.75\columnwidth]{plots/legend.png}
 
@@
 
-% \begin{table}[t]
-%     \centering
-%     \caption{Fashion-MNIST: throughput and recall. See general notes above.}
-%     \label{tab:results:qps-and-recall-fmn}
-%     \small
-%     \setlength{\tabcolsep}{4pt}
-%     \begin{adjustbox}{width=\columnwidth,center}
-%     \begin{tabular}{@{} lcccccccc @{}}
-%     \toprule
-%     \textbf{M} &
-%     \multicolumn{2}{c}{\textbf{HNSW}} &
-%     \multicolumn{2}{c}{\textbf{ANNOY}} &
-%     \multicolumn{2}{c}{\textbf{FAISS-IVF}} &
-%     \multicolumn{2}{c}{\textbf{CAKES}} \\
-%     \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
-%     & QPS & Recall & QPS & Recall & QPS & Recall & QPS & Recall \\
-%     \midrule
-%     1   & \num{1.33e4} & 0.95 & \num{2.19e3} & 0.95 & \num{2.01e3} & $1.00^{*}$ & \num{3.46e3} & 1.00 \\
-%     2   & \num{1.38e4} & 0.80 & \num{2.12e3} & 0.93 & \num{9.39e2} & $1.00^{*}$ & \num{3.68e3} & 1.00 \\
-%     4   & \num{1.66e4} & 0.68 & \num{2.04e3} & 0.90 & \num{4.61e2} & $1.00^{*}$ & \num{3.44e3} & 1.00 \\
-%     8   & \num{1.68e4} & 0.53 & \num{1.93e3} & 0.86 & \num{2.26e2} & $1.00^{*}$ & \num{3.30e3} & 1.00 \\
-%     16  & \num{1.87e4} & 0.49 & \num{1.84e3} & 0.86 & \num{1.17e2} & 0.99       & \num{3.34e3} & 1.00 \\
-%     32  & \num{1.56e4} & 0.54 & \num{1.85e3} & 0.78 & \num{5.91e1} & 0.99       & \num{2.96e3} & 1.00 \\
-%     64  & \num{1.50e4} & 0.38 & \num{1.78e3} & 0.69 & \num{2.61e1} & 0.97       & \num{3.25e3} & 1.00 \\
-%     128 & \num{1.49e4} & 0.36 & \num{1.66e3} & 0.54 & \num{1.33e1} & 0.96       & \num{2.96e3} & 1.00 \\
-%     256 & --           & --    & \num{1.60e3} & 0.59 & \num{6.65e0} & 0.96       & \num{2.79e3} & 1.00 \\
-%     512 & --           & --    & \num{1.83e3} & 0.58 & \num{3.56e0} & 0.95       & \num{2.84e3} & 1.00 \\
-%     \bottomrule
-%     \end{tabular}
-%     \end{adjustbox}
-%     \end{table}
-
 \begin{table}[t]
   \centering
-  \begin{tiny}
   \caption{Fashion-MNIST: throughput and recall.}
-  \end{tiny}
+  \vskip -0.05in
   \label{tab:results:qps-and-recall-fmn}
@@
   \end{adjustbox}
+  \vskip -0.1in
 \end{table}
@@
   \caption{Glove-25: throughput and recall.}
+  \vskip -0.05in
   \label{tab:results:qps-and-recall-glove}
@@
 \begin{table}
+  \vskip -0.1in
   \centering
@@
   \label{tab:results:qps-and-recall-sift}
+  \vskip -0.05in
     \small
@@
 \begin{table}
+  \vskip -0.1in
   \caption{Random dataset: throughput and recall.
@@
   \label{tab:results:qps-and-recall-random}
+  \vskip -0.05in
   \small
@@
 
-\subsection{Clustering Strategies and Number of Distance Computations}
+\subsection{Clustering Strategies}
 \label{sec:results:clustering-strategies-and-number-of-distance-computations}
@@
 We consider four strategies—the Cartesian product of balanced vs.\ unbalanced clustering and the presence vs.\ absence of depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}).
-We instrument CAKES to count distance computations during search; this adds substantial overhead and would not be used in real-world applications.
+We instrument CAKES to count distance computations during search;
+this adds substantial overhead and would not be used in real-world applications.
 
@@
 Balanced clustering is strictly worse in throughput and has poorer asymptotics: algorithms such as Depth-First Sieve that appear nearly constant-time in Fig.~\ref{fig:results:scaling-plots} (with \textit{PermutedBall}) no longer do so under balanced clustering.
-On this dataset, depth-first reordering has little effect on throughput; it is primarily intended to improve space complexity.
+On this dataset, depth-first reordering has little effect on throughput; it is primarily intended to improve memory consumption.
 
@@
   \vspace{2pt}
-  \includegraphics[width=1\columnwidth]{images/distance_counts/legend.png}
+  \includegraphics[width=0.75\columnwidth]{images/distance_counts/legend.png}
 

### 2025-08-29 sections/5-discussion.tex
acknowleged unbalanced vs balanced section in discussion
back down to 10 pages with some extra material

@@
 It builds an unbalanced cluster tree that captures manifold structure and helps accelerate search.
+The results in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} shows that this unbalanced clustering approach is superior to balanced clustering in both absolute and asymptotic performance.
 Conversely, performance degrades on randomly distributed data, where such structure is absent.
@@
 
-We have noted how the fastest algorithm in CAKES varies depending on the dataset and distance function being used, as well as the cardinality of the dataset.
+We have noted how the fastest algorithm in CAKES changes depending on the dataset and distance function being used, as well as the cardinality of the dataset.
 This variability motivated our auto-tuning process (Sec.~\ref{sec:methods:auto-tuning}).
@@
 Simple edit distance can be poor when lengths vary substantially.
-CAKES can get around that issue by instead using alternative metrics or preprocessing—for example, represent each sequence as a set of $k$-mers and use Jaccard distance, as in~\cite{kim2021entrance}.
-If the distance relies on global alignment (e.g., Needleman-Wunsch~\cite{needleman1970general}), the dataset can be partitioned into mutually exclusive length bins with a separate tree per bin;
+CAKES can get around that issue by instead using alternative metrics or preprocessing—for example, the dataset can be partitioned into mutually exclusive length bins with a separate tree per bin;
 at query time, we simply route the sequence to its bin by length, alogn with a small number of adjacent bins.
@@
 
-CLAM and CAKES are implemented in Rust and available under the MIT license at https://github.com/URI-ABD/clam.
+CAKES is implemented in Rust and available under the MIT license at https://github.com/URI-ABD/clam.

### 2025-08-29 sections/3-datasets.tex
minor typos

@@
 Figure~\ref{fig:results:sift-lfd} shows the LFD for Sift, which is higher relative to both Fashion-MNIST and Glove-25, increasing sharply to a peak of 9 around depth 10.
-For Random, which has the same cardinality and dimensionality as Sift, LFD starts at 20 the root cluster, and all percentile values decrease linearly with depth until reaching the leaves of the tree.
+For Random, which has the same cardinality and dimensionality as Sift, LFD starts at 20 at the root cluster, and all percentile values decrease linearly with depth until reaching the leaves of the tree.
 The spread in LFD starts narrow for the first few depths and widens as depth increases.

### 2025-08-29 sections/4-results.tex
minor typos

@@
 
-Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
+Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm for that dataset and cardinality only.
 

### 2025-08-30 sections/5-discussion.tex
minor revisions

@@
 
-HNSW and ANNOY maintain near-constant throughput as the cardinality grows, but their recalls decline and their index build time and memory rise, making them unsuitable when datasets grow exponentially.
-CAKES's tree builds orders of magnitude faster with far less memory; while queries are slower than HNSW/ANNOY, CAKES attains near-constant scaling with perfect recall.
+HNSW and ANNOY maintain near-constant throughput as the cardinality grows, but their recalls decline and their index build times rise, making them unsuitable when datasets grow exponentially.
+CAKES's tree builds orders of magnitude faster; while queries are slower than HNSW/ANNOY, CAKES attains near-constant scaling with perfect recall.
 

### 2025-08-30 sections/2-methods.tex
minor revisions
one less orphan

@@
 
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} gives us $\mathcal{T}$.
-During this procedure, we also compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
+Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds $\mathcal{T}$.
+During Partition, we also compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
 

### 2025-11-05 sections/1-introduction.tex
down to 7 pages

@@
 Researchers are collecting data at an unprecedented rate, with datasets in many fields growing exponentially, and outpacing improvements in computing performance predicted by Moore's Law~\cite{kahn2011future}.
+Examples include genomic databases~\cite{10.1093/nar/gks1219}, time-series signals~\cite{oshea2018radioml}, and neural embeddings~\cite{2020arXiv200514165B, OpenAI2023GPT4TR, Touvron2023Llama2O, radford2021learning, dosovitskiy2020image}.
 This ``Big Data explosion'' has created a need for algorithms that scale efficiently to large datasets.
-
-Examples of such datasets include genomic databases, time-series signals, and embeddings from neural networks such as GPT~\cite{2020arXiv200514165B, OpenAI2023GPT4TR}, LLAMA-2~\cite{Touvron2023Llama2O}, and image analysis models~\cite{radford2021learning, dosovitskiy2020image}.
-Among biological datasets, SILVA 18S~\cite{10.1093/nar/gks1219} includes ribosomal RNA sequences from approximately 2.25 million genomes.
-Among time-series datasets, the RadioML dataset~\cite{oshea2018radioml} contains approximately 2.55 million samples of synthetically-generated radio frequency signals.
-
-Similarity search is a fundamental task on such datasets, enabling applications such as recommendation~\cite{annoy} and classification ~\cite{suyanto2022knnclassifier}.
+One such class of algorithms is similarity search algorithms, which enable applications such as recommendation~\cite{annoy} and classification ~\cite{suyanto2022knnclassifier}.
 Yet as datasets grow in size and dimensionality, efficient and accurate similarity search becomes increasingly challenging;
@@
 The former retrieves the $k$ points closest to a query, while the latter retrieves all points within a similarity threshold $\rho$ of a query.
-While prior work has used the term \textit{approximate} search to refer to $\rho$-NN, we use it more narrowly: an {approximate} search algorithm is any algorithm that yields imperfect recall, whereas an \textit{exact} algorithm guarantees perfect recall.
-
+While prior work has used the term \textit{approximate} search to refer to $\rho$-NN, we use it more narrowly: an {approximate} search algorithm yields imperfect recall, whereas an \textit{exact} algorithm guarantees perfect recall.
 $k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because they scale linearly with dataset size.
@@
 This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for \emph{exact} $k$-NN search.
-We benchmark CAKES against state-of-the-art methods--FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}--on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-We also evaluate performance on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and on a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
-To further contextualize results, we compare against a synthetic dataset with simple statistical properties.
+We benchmark CAKES against FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We also evaluate performance on SILVA 18S~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and RadioML~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
+To further contextualize results, we also use a synthetic dataset with simple statistical properties.
 
@@
 HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
+FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
 ANNOY~\cite{annoy} uses random projection and tree building.

### 2025-11-05 sections/2-methods.tex
down to 7 pages

--- 2025-08-29 sections/2-methods.tex
@@
 
-In this manuscript, we study $k$-NN search in finite-dimensional spaces.
-Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$, we refer to each $x_i \in \textbf{X}$ as a point.
-Examples include neural embeddings, genomic sequences, and time-series.
-
-We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
-We require $f$ to be symmetric ($f(x, y) = f(y, x)$) and identity-preserving ($f(x, y) = 0 \iff x = y$). 
-If $f$ satisfies the triangle inequality (i.e.,\,$f(x, y) \leq f(x, z) + f(z, y)$), then it is also a \textit{distance metric}.
-Common examples include Euclidean, Levenshtein~\cite{levenshtein1966binary}, and DTW~\cite{muller2007dynamic};
-Cosine distance violates the triangle inequality and is not a metric.
-The appropriate distance function depends on the data: Euclidean or Cosine are often used for neural embeddings, Levenshtein or Hamming for sequences, and DTW or Wasserstein for time-series.
-CAKES provides exact search under distance metrics.
+We present CAKES, a set of three novel algorithms for exact $k$-NN search in finite-dimensional spaces.
+Given a \textit{dataset} $\textbf{X} = \{x_1, \dots, x_n\}$ and a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$, $k$-NN search aims to find the $k$ points in $\textbf{X}$ closest to a given \textit{query} point $q$ according to $f$.
+If $f$ is a distance metric (i.e. it obeys the triangle inequality), then CAKES guarantees exact $k$-NN search.
 
@@
 where $B(q, r)$ is the set of points in the metric ball of radius $r$ centered at $q$.
-We typically use a simplified version of Equation~\ref{eq:methods:lfd-original} with $r_1 = 2 \cdot r_2$:
-\begin{equation}
-    \text{LFD}(q, r) = \text{log}_2 \left( \frac{|B(q, r)|}{|B(q, \frac{r}{2})|} \right).
-    \label{eq:methods:lfd-half}
-\end{equation}
 
@@
 When the vast majority of points in the dataset have low ($d \ll D$) LFD, we simply say that the dataset has low LFD.
-We stress that this concept differs from the \textit{embedding dimension} of a dataset.
-To illustrate the difference, consider the SILVA 18S rRNA dataset that contains genomic sequences with unaligned lengths of up to 3,712 bases and aligned length of 50,000 bases.
-Hence, the \textit{embedding dimension} of this dataset is at least 3,712 and at most 50,000.
-However, physical limitations (namely, biological evolution and biochemistry) constrain the data to a low-dimensional manifold within this space.
-LFD is an approximation of the dimensionality of that low-dimensional manifold in the ``vicinity'' of a given point.
-% Section~\ref{sec:results:lfd-of-datasets} discusses this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.
-For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but (potentially) globally variable.
+For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but (often) globally variable.
 
@@
 
-We define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
-The center is the geometric median—the point minimizing the sum of distances to all other points—of the points in $C$;
-in practice, for large $|C|$ we only compute this for a smaller sample.
-A cluster's radius is the maximum distance from the center to any point in it.
-Each non-leaf cluster has two children (as in a binary tree).
-Clusters may have overlapping volumes, but each point in an overlap is assigned to exactly one cluster;
-consequently $C(c,r)\subset B(c,r)$.
-We denote the cluster tree by $\mathcal{T}$ and its root by $\mathcal{R}$ (Sec.~\ref{sec:methods:clustering:building-the-tree}).
-
-We estimate the LFD of clusters at scales $r$ and $r/2$ via Eq.~\ref{eq:methods:lfd-half}, using only the points in $C(c,r)$ (not all of $B(c,r)$).
-
-For a flat clustering, \emph{metric entropy} $\mathcal{N}_r(X)$ is the minimum number of radius-$r$ clusters needed to cover $X$~\cite{yu2015entropy}.
-In our hierarchical setting, we define $\mathcal{N}_{\hat r}(X)$ as the number of leaf clusters, where $\hat r$ is the mean radius of all leaves.
-
-\subsubsection{Building the Tree}
-\label{sec:methods:clustering:building-the-tree}
-
-We construct $\mathcal{T}$ using CLAM, following CHESS~\cite{ishaq2019clustered} but with improved pole selection, then apply a depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}) on the dataset.
-The procedure is reproduced in Alg.~\ref{alg:methods:partition} for completeness.
-
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} gives us $\mathcal{T}$.
-During this procedure, we also compute and cache the LFD of each cluster using Eq.~\ref{eq:methods:lfd-half}.
+We define a \emph{cluster} $C$ as a set of points with a \emph{center} $c$ and \emph{radius} $r$.
+The center is the geometric median of the points in the cluster (i.e., the point minimizing the sum of distances to all other points in $C$) and the radius is the maximum distance from the center to any point in $C$.
+A cluster with only one point is a $leaf$, otherwise a $parent$ cluster has two $child$ clusters.
+Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds the tree $\mathcal{T}$ in $\mathcal{O}(n \log n)$ time as described in~\cite{ishaq2019clustered}.
 
 \begin{algorithm} % enter the algorithm environment
-    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
+    \caption{\textsc{Partition}($C$, $criteria$)} % give the algorithm a caption
     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
     \begin{algorithmic} % enter the algorithmic environment
-    \REQUIRE $f$, a distance function
-    \REQUIRE $C$, a cluster
-    \REQUIRE $criteria$, user-specified continuation criteria
+        \Require $f$, a distance function
+        \Require $C$, a cluster
+        \Require $criteria$, user-specified continuation criteria
 
-    \STATE $S \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-    \STATE $c \Leftarrow$ geometric median of $S$
-    \STATE $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-    \STATE $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-    \STATE $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-    \STATE $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
-
-    \IF{$|L| > 1$ \textbf{and} $L$ satisfies $criteria$}
-        \STATE Partition($L$, $criteria$)
-    \ENDIF
-
-    \IF{$|R| > 1$ \textbf{and} $R$ satisfies $criteria$}
-        \STATE Partition($R$, $criteria$)
-    \ENDIF
+        \If{$|C| > 1$ \textbf{and} $C$ satisfies $criteria$}
+            \State $S \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+            \State $c \Leftarrow$ geometric median of $S$
+            \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+            \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+            \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+            \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+            \State $L \Leftarrow \Call{Partition}{L, criteria}$
+            \State $R \Leftarrow \Call{Partition}{R, criteria}$
+            % \State $L \Leftarrow Partition(L, criteria)$
+            % \State $R \Leftarrow Partition(R, criteria)$
+        \EndIf
     \end{algorithmic}
 \end{algorithm}
-
-\subsubsection{Depth-First Reordering}
-\label{sec:methods:clustering:depth-first-reordering}
-
-In CHESS~\cite{ishaq2019clustered}, each cluster stored a list of indices into the dataset.
-This list was used to retrieve the clusters' points during search.
-With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, it stored $n$ indices at each depth in $\mathcal{T}$.
-Assuming $\mathcal{T}$ is balanced, and thus $\mathcal{O}(\log n)$ depth, this approach had a prohibitively high memory overhead of $\mathcal{O}(n \log n)$.
-In this work, we introduce a new approach wherein, after building $\mathcal{T}$, we reorder the dataset so that points are stored in a depth-first order.
-Then, each cluster need only store its \textit{cardinality} and an \textit{offset} to access its points.
-The root cluster $\mathcal{R}$ has an offset of zero and a cardinality equal to the number of points in the dataset.
-A left child has the same offset its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
-With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS but with $\mathcal{O}(n)$ memory overhead.
-
-\subsubsection{Time Complexity}
-\label{sec:methods:clustering:time-complexity}
-
-The asymptotic complexity of building $\mathcal{T}$ is the same as described in~\cite{ishaq2019clustered}, i.e., $\mathcal{O}(n \log n)$.
-This is a significant improvement over the exact approach, which costs $\mathcal{O}(n^2 \log n)$.
-The $\mathcal{O}(n\log n)$ build complexity assumes a balanced tree $\mathcal{T}$.
-In practice, Alg.~\ref{alg:methods:partition} yields unbalanced trees on real data because of non-uniform sampling densities and the manifold's low-dimensional structure.
-A balanced tree is expected only for uniformly distributed data (e.g., in a $d$-dimensional hypercube).
-
-Although (unbalanced) trees are deeper than $\log n$, which may appear to increase the cost of tree-building, many leaves appear at shallow depths, so later levels contain fewer points and the per-level build cost declines.
-Any analysis for unbalanced trees would be dataset-dependent, so we only report empirical results based on balanced version of Alg.~\ref{alg:methods:partition} in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations}.
-In any case, once the tree is built, its cost is amortized over all queries and subsequent applications.
 
@@
     \label{fig:methods:deltas}
-    % \vskip -0.5in
+    \vskip -0.2in
 \end{figure}
-\FloatBarrier
+% \FloatBarrier
 
-% \begin{minipage}{.425\textwidth}
-\FloatBarrier
+% \FloatBarrier
 \begin{algorithm}[!t]
-        \caption{tree-search($C$, $q$, $r$)}
+    \caption{\textsc{TreeSearch}($C$, $q$, $r$)}
     \label{alg:methods:rnn-search:tree-search}
-        \begin{algorithmic}[1]
-            \REQUIRE $f$, a distance function
-            \REQUIRE $C$, a cluster
-            \REQUIRE $q$, a query
-            \REQUIRE $r$, a search radius
-            \IF{$\delta^+_C \leq r$}
-                \STATE \textbf{return} $\{C\}$
-            \ELSE
-                \STATE $[L, R]$ $\Leftarrow$ \textit{children} of $C$
-                \STATE \textbf{return} tree-search($L, q, r$) $\cup$ tree-search($R, q, r$)
-            \ENDIF
+    \begin{algorithmic}
+        \Require $C$, a cluster
+        \Require $q$, a query
+        \Require $r$, a search radius
+        \If{$\delta^+_C \leq r$}
+            \State \textbf{return} $\{C\}$
+        \EndIf
+        \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
+        \State \textbf{return} \Call{TreeSearch}{$L, q, r$} $\cup$ \Call{TreeSearch}{$R, q, r$}
+        % \State \textbf{return} tree-search($L, q, r$) $\cup$ tree-search($R, q, r$)
     \end{algorithmic}
 \end{algorithm}
-% \end{minipage}
 
-% \begin{minipage}{.475\textwidth}
 \begin{algorithm}[!t]
-        \caption{leaf-search($Q$, $q$, $r$)}
+    \caption{\textsc{LeafSearch}($Q$, $q$, $r$)}
     \label{alg:methods:rnn-search:leaf-search}
-        \begin{algorithmic}[1]
-            \REQUIRE $f$, a distance function
-            \REQUIRE $Q$, a set of clusters
-            \REQUIRE $q$, a query
-            \REQUIRE $r$, a search radius
-            \STATE $H \Leftarrow \emptyset$
-            \FOR{$C \in Q$}
-                \IF{$\delta^+_C \leq r$}
-                    \STATE $H$ $\Leftarrow$ $H \cup C$
-                \ELSE
-                    \FOR{$p \in C$}
-                        \IF{$f(p, q) \leq r$}
-                            \STATE $H$ $\Leftarrow$ $H \cup \{p\}$
-                        \ENDIF
-                    \ENDFOR
-                \ENDIF
-            \ENDFOR
-            \STATE \textbf{return} $H$
+    \begin{algorithmic}
+        \Require $f$, a distance function
+        \Require $Q$, a set of clusters
+        \Require $q$, a query
+        \Require $r$, a search radius
+        \State $H \Leftarrow \emptyset$
+        \For{$C \in Q$}
+            \If{$\delta^+_C \leq r$}
+                \State $H$ $\Leftarrow$ $H \cup C$
+            \Else
+                \For{$p \in C$ \textbf{if} $f(p, q) \leq r$}
+                    \State $H$ $\Leftarrow$ $H \cup \{p\}$
+                \EndFor
+            \EndIf
+        \EndFor
+        \State \textbf{return} $H$
     \end{algorithmic}
 \end{algorithm}
-% \end{minipage}
 
@@
 
-Given a user-specified integer $k$ and query $q$, $k$-NN search aims to find the $k$ closest points to $q$ in $\textbf{X}$.
-In other words, $k$-NN search aims to find the set $H$ such that $|H| = k$ and $H = B(q, \rho_k)$ where $\rho_k = \max \left\{ f(q, p) \ \forall \ p \in H \right\}$ is the distance from $q$ to the $k^{th}$ nearest neighbor in $\textbf{X}$.
-
-In this section, we present three novel algorithms for exact $k$-NN search:
+Given a query $q$ and integer $k$, $k$-NN search finds the $k$ closest points to $q$ in $\textbf{X}$. In this section, we present three novel algorithms for exact $k$-NN search:
 Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-In these algorithms, we use $H$ to refer to the data structure that stores the closest points to the query found so far, and $Q$ to refer to the data structure that stores the clusters and points that are still in contention for being among the $k$ nearest neighbors.
+We use $H$ to refer to the data structure containing the nearest neighbors found so far, and $Q$ for the data structure containing the clusters and points that are still in contention for being among the $k$ nearest neighbors.
 These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
@@
 
-This algorithm relies on the \textit{tree-search} (Alg.~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Alg.~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.
+We begin using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
+If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
+Now, so long as $\sum_{C \in Q} |C| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{C \in Q} |C|}} \big)^\frac{1}{\mu}$
+where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
+Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
 
-We begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root $\mathcal{R}$ divided by $|X|$.
-If no clusters are found, we double $r$ and repeat tree-search, until we find a a non-empty set, $Q$, of clusters.
-
-Now, so long as $\sum_{C \in Q} |C| < k$, we continue to repeat tree-search, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of
-\begin{equation}
-  \bigg( {\frac{k}{\sum_{C \in Q} |C|}} \bigg)^{\mu}
-  \label{eq:methods:repeated-rnn-factor}
-\end{equation}
-where $\mu$ is the multiplicative inverse of the harmonic mean of the LFD of the clusters in $Q$, i.e.,\,$\mu = \frac{1}{|Q|} \cdot \sum_{C \in Q} \big( LFD(C)^{-1} \big)$.
-We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
-In practice, we cap this factor at 2 to ensure that we do not increase the radius too quickly in any single iteration.
-
-Intuitively, the factor by which we increase the radius should be \textit{inversely} related to the number of points found so far.
-When the LFD at a given radius scale is high, the data are dense in that region.
-Thus, a small increase in the radius would likely encounter many more points, and find $k$ neighbors.
-Conversely, when the LFD at that scale is low, the data are sparse.
-In such a region, a small increase in the radius would likely encounter vacant space, so a larger increase is needed.
-Thus, the factor of radial increase should be \textit{inversely} related to the LFD.
-However, we should not increase the radius too drastically with any one iteration because we assume that the LFD is only \textit{locally} uniform.
-A large increase in the radius would likely break out of the local region and potentially encounter too many new clusters, which would make the subsequent step computationally expensive.
-
-Once $\sum_{C \in Q} |C| \geq k$, we are guaranteed to have found at least $k$ neighbors, and so we perform \textit{leaf-search} to find the $k$ nearest neighbors among the points in the clusters found after the last tree-search.
+Once $\sum_{C \in Q} |C| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
 
@@
 
-% \begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
-Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
-%         \mathcal{O}
-%         \Bigg(
-%             \underbrace{
-%                 \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-%             }_{\textrm{tree-search}}
-%             \ + \
-%             \underbrace{
-%                 \overbrace{k}^{\textrm{output size}} \cdot
-%                 \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
-%             }_{\textrm{leaf-search}}
-%         \Bigg)
-%         \label{eq:methods:repeated-rnn-complexity}
-%     \end{gather}
-%     where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
-%     \label{thm:methods:rnn-complexity}
-% \end{theorem}
-
-We consider the \textit{tree-search} and \textit{leaf-search} stages of search separately.
-Tree-search refers to the process of identifying clusters that overlap with the query ball, or in other words, clusters that might contain one of the $k$ nearest neighbors.
-In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
+In CHESS~\cite{ishaq2019clustered}, we showed that the complexity of ranged search is
 \begin{gather}
@@
             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
-        }_{\textrm{tree-search}}
+        }_{\textrm{\textsc{TreeSearch}}}
         +
@@
             \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ {\text{LFD}(q, \rho)}}^{\textrm{scaling factor}}
-        }_{\textrm{leaf-search}}
+        }_{\textrm{\textsc{LeafSearch}}}
     \Bigg).
@@
 \end{gather}
+We now extend this to Repeated $\rho$-NN.
 
-To extend Eq.~\ref{eq:methods:rnn-search-complexity} to Repeated $\rho$-NN, we must estimate the number of iterations of tree-search (Alg.~\ref{alg:methods:rnn-search:tree-search}) needed to find a radius that guarantees $k$ neighbors.
-Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
-Thus, Equation~\ref{eq:methods:repeated-rnn-factor} suggests that in the expected case, we need only two iterations of tree-search to find $k$ neighbors:
-one iteration to find at least one cluster, and one more to find enough clusters to guarantee $k$ neighbors.
-Since this is a constant factor, complexity of tree-search for Repeated $\rho$-NN is the same as that for $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
+We first need to estimate the number of iterations of \textsc{TreeSearch} needed to find a radius that guarantees at least one neighbor.
+Since we expect $q$ to be from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the nearest neighbor.
+Thus, the multiplicative factor described above suggests that in the expected case, we need only one iteration of \textsc{TreeSearch} to find the first cluster, and one more to find enough clusters to guarantee $k$ neighbors.
+Since this is a constant factor, Repeated $\rho$-NN has the same complexity for \textsc{TreeSearch} as $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
 
-We proceed to determine the complexity of leaf-search.
-Let $Q$ be the set of clusters returned by tree-search.
-We must estimate $\sum_{C \in Q} |C|$, the total cardinality of those clusters;
-since we must examine every point in each such cluster, time complexity of leaf-search is linear in this quantity.
+For \textsc{LeafSearch}, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of those clusters;
+since we must examine every point in each such cluster, time complexity of \textsc{LeafSearch} is linear in this quantity.
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
-Then, we see that $Q$ is expected to be the set of clusters that overlap with a ball of radius $\rho_k$ around the query.
+Then, we see that $Q$ is the set of clusters that overlap with a ball of radius $\rho_k$ around the query.
 We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $Q$.
-
-The work in~\cite{yu2015entropy} showed that, for some constant $\gamma$,
-\begin{equation*}
-    \sum_{C \in S} |C| \leq \gamma \left| B(q, \rho_k) \right| \left(\frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^d.
-\end{equation*}
-
-By definition of $\rho_k$, we have that $|B(q, \rho_k)| = k$.
-Thus, $\sum_{C \in S} |C| \leq \gamma k \left( 1 + 2 \cdot \frac{\hat{r}}{\rho_k} \right)^d$.
+Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot \left( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of the LFDs of the clusters in $Q$.
 It remains to estimate $\rho_k$.
 
-Let $\hat{d}$ be the harmonic mean of the LFDs of the clusters in $Q$.
-While ordinarily we compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
+While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
 Since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at the center of a cluster in $Q$.
 By Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
-Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get:
-\begin{equation*}
-    k \Bigg( 1 + 2 \cdot \bigg( \frac{\hat{|C|}}{k} \bigg) ^ {\hat{d}^{-1}} \Bigg)^d
-\end{equation*}
+This simplifies the \textsc{LeafSearch} component to $k \Bigg( 1 + 2 \cdot \bigg( \frac{\hat{|C|}}{k} \bigg) ^ {\hat{d}^{-1}} \Bigg)^d$.
 
 Again using the assumption that $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
-By combining the bounds for tree-search and leaf-search, we see that time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is
+Thus, the asymptotic complexity of Repeated $\rho$-NN is
 \begin{gather}
@@
             \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
-        }_{\textrm{leaf-search}}
+        }_{\textrm{\textsc{LeafSearch}}}
     \Bigg).
@@
 
-This algorithm (BFS) performs a breadth-first traversal of $\mathcal{T}$, pruning clusters by using a modified version of the QuickSelect algorithm~\cite{hoare1961algorithm} (a subroutine in QuickSort), at each level of $\mathcal{T}$.
+This algorithm (BFS) performs a breadth-first traversal of $\mathcal{T}$, pruning clusters by using a modified version of Hoare's QuickSelect algorithm~\cite{hoare1961algorithm}.
 
@@
 During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
-Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{sec:methods:knn-search}, since $c$ is a point, $\delta_{C} = \delta_{c} = \delta^{+}_{c} = \delta^{-}_{c}$.
 
-We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$.
-Since this step may require a binary search of $Q$ to find $\tau$ and reordering $Q$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
+We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$.
+Since this step may require a binary search of $Q$ and reordering $Q$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
 
@@
 If it corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
-Otherwise (since it must correspond to a non-leaf cluster), we add to $Q$ the pairs of 3-tuples corresponding to its child clusters.
-We continue this process until the sum of multiplicities in $Q$ is exactly $k$.
+Otherwise (since it must correspond to a parent cluster), we add to $Q$ the 3-tuples corresponding to its child clusters, thus expanding the search to the next level in $T$.
+We continue until the sum of multiplicities in $Q$ is exactly $k$.
 These are the $k$ nearest neighbors.
@@
 
-This algorithm (DFS) is similar to a depth-first traversal of $\mathcal{T}$.
-It uses a max-priority queue to track hits, and a min-priority queue to choose the next branch of $\mathcal{T}$ to explore.
+This algorithm (DFS) uses a max-priority queue to track hits, and a min-priority queue to choose the next branch of $\mathcal{T}$ to explore, making it similar to a depth-first traversal.
 
@@
 \end{itemize}
-This process terminates when $H$ is full and the top-priority items in $H$ and $Q$ are such that $\delta_H < \delta_Q$, i.e.,\,the theoretically closest point left is farther from the query than the $k^{th}$ nearest neighbor found so far.
+This process terminates when $H$ is full and the top-priority items in $H$ and $Q$ are such that $\delta_H < \delta_Q$, i.e.,\,the next closest point to be visited is farther from the query than the $k^{th}$ nearest neighbor found so far.
 Thus, $H$ contains the $k$ nearest neighbors to the query.
@@
 Due to their similarity, we combine the complexity analyses of both Sieve methods.
-For these methods we again use the terminology of tree-search and leaf-search.
-Tree-search navigates the cluster tree and finds $Q$.
-Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.
-
-% \begin{theorem}
-%     Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$.
-Let $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$ and   $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of $X$, $d$ is the LFD of $X$, $\hat{|C|}$ is the mean cardinality of clusters overlapping the query ball, and $k$ is the number of nearest neighbors.
-Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
-%         \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
-%         \label{eq:methods:breadth-first-sieve-complexity}
-%     \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
-%         \mathcal{O} \Big( T \log T + L \log k \Big).
-%         \label{eq:methods:depth-first-sieve-complexity}
-%     \end{equation}
-%     \label{thm:methods:sieve-complexity}
-% \end{theorem}
-
-Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.
-Let $d$ be the LFD in this region, and consider leaf clusters with cardinalities near $k$.
-Then the number of leaf-clusters in $Q$ is bounded above by $2d$, where the bound is achieved by having a cluster overlap the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
-In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search would look at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
-Thus, the asymptotic complexity of tree-search is
-\begin{equation*}
-    T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big).
-\end{equation*}
-For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the complexity is
-\begin{equation*}
-    L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|C|}}{k} \big) ^ {d^{-1}} \Big)^d \Big).
-\end{equation*}
+For these methods we again use the terminology of \textsc{TreeSearch} and \textsc{LeafSearch}.
+Let $d$ be the LFD in the region around $q$, and consider clusters with cardinalities near $k$.
+Then the number of such clusters is bounded above by $2d$, where the bound is achieved by having a cluster overlap the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
+In the worst-case scenario for \textsc{TreeSearch}, these clusters would all come from different branches of the tree, and \textsc{TreeSearch} would look at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
+Thus, the asymptotic complexity of this component is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
+For \textsc{LeafSearch}, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the complexity is $L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|C|}}{k} \big) ^ {d^{-1}} \Big)^d \Big)$.
 
 The asymptotic complexity of BFS is dominated by QuickSelect.
-Since QuickSelect is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity of BFS is
+Since $Q$ contains the clusters from \textsc{TreeSearch} and the points from \textsc{LeafSearch}, we see that the asymptotic complexity of BFS is
 \begin{equation}
-    \mathcal{O} \big( (T + L ) \log (T + L ) \big),
+    \mathcal{O} \big( (T + L ) \log (T + L ) \big).
     \label{eq:methods:bfs-complexity}
 \end{equation}
-and that of DFS is 
+Similarly, the asymptotic complexity of DFS is dominated by the priority queue operations for $T$ clusters and $L$ points;
+it is given by
 \begin{equation}
@@
 \end{equation}
-
-\subsection{Auto-Tuning}
-\label{sec:methods:auto-tuning}
-
-We perform some simple auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
-We start by taking the center of every cluster at a low depth (e.g.,\,10) in $\mathcal{T}$ as a query.
-This gives us a small, representative sample of the dataset.
-Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{sec:methods:knn-search}.
-We select the fastest algorithm over all the queries as the optimal algorithm for that dataset and value of $k$.
-Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.
-
-\subsection{Synthetic Data}
-\label{sec:methods:synthetic-data}
-
-Based on the asymptotic analysis, we expect CAKES to perform well on datasets with low LFD and scale sublinearly with cardinality.
-To test this, we use datasets from the ANN-Benchmarks suite~\cite{aumuller2020ann} and synthetically augment them to have exponentially larger cardinalities.
-We perform the same procedure on a dataset of uniformly distributed points in a hypercube, and compare CAKES against other algorithms on both the original and augmented datasets.
-
-Our augmentation procedure is as follows.
-Let $X$ be a dataset of dimension $d$, $\epsilon$ a user-specified noise level, and $m$ an integer multiplier.
-For each $\mathbf{x}\in X$, generate $m-1$ new points within distance $\epsilon$ of $\mathbf{x}$ by sampling a random vector $\mathbf{r}\in\mathbb{R}^d$ from the $\epsilon$-ball (hypersphere of radius $\epsilon$ centered at the origin) and setting $\mathbf{x}'=\mathbf{x}+\mathbf{r}$.
-Since $\lVert\mathbf{r}\rVert\le\epsilon$, we have $\lVert\mathbf{x}-\mathbf{x}'\rVert\le\epsilon$.
-The result is $X'$ with $\lvert X'\rvert=m\,\lvert X\rvert$.
-This increases cardinality by a factor of $m$ without altering the dataset’s overall topological structure, isolating the effect of cardinality from dimensionality, metric choice, and topology.

### 2025-11-05 sections/3-datasets.tex
down to 7 pages

--- 2025-08-29 sections/3-datasets.tex
@@
 
-\subsection{ANN-Benchmark Datasets}
+\subsection{Datasets}
 \label{sec:datasets-and-benchmarks:ann-benchmark-datasets}
 
-We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
-Table~\ref{tab:datasets:summary} summarizes the cardinality and dimensionality of each these datasets.
-All benchmarks were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
-The OS kernel was Manjaro Linux 5.15.164-1-MANJARO.
-The Rust compiler was Rust 1.83.0, and the Python interpreter version was 3.9.18.
+Table~\ref{tab:datasets:summary} summarizes the cardinality and dimensionality of each dataset we use.
+These include the Fashion-MNIST, Glove-25, and Sift datasets from the ANN-Benchmarks suite~\cite{aumuller2020ann}, as well as a Random dataset we generated (using the same cardinality and dimensionality as Sift) as an anti-example to demonstrate the effect of manifold structure on CAKES performance.
+Thus, Random shows that CAKES is designed to exploit manifold structure when it exists, and should not be used on datasets that do not obey the manifold hypothesis.
+The Glove-25 dataset uses the Cosine distance function, while the other three datasets use the Euclidean distance metric.
+All benchmarks using these datasets were conducted on an r6i.16xlarge instance on Amazon Web Services (AWS), with 64 vCPUs and 512 GiB of RAM.
+The Rust compiler was Rust 1.83.0 and the Python interpreter version was 3.9.18.
+
+We also include benchmarks on the SilVA-18S rRNA and Radio-ML datasets.
+For the SilVA-18S dataset, which consists of 2,224,640 sequences of up-to 3,712 nucleotides, we use the Levenshtein distance metric~\cite{levenshtein1966binary}.
+The Radio-ML dataset consists of 1,024-dimensional complex-valued time-series samples, spanning 24 modulation modes and 26 SNR levels from $-20$dB to $30$dB.
+We use the 10dB SNR subset with 97,304 samples and use Dynamic Time Warping~\cite{muller2007dynamic} as the distance metric.
+For both datasets, we hold out 1,000 random samples as queries and build the tree on the remaining samples.
+These were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM running Manjaro Linux 5.15.164-1-MANJARO with Rust compiler version 1.83.0.
 
@@
 
-\subsection{Random Datasets and Synthetic Augmentation}
-\label{sec:datasets-and-benchmarks:random-datasets}
+\subsection{Scaling Experiments}
+\label{sec:datasets-and-benchmarks:scaling-experiments}
 
-In addition to the ANN-Benchmarks datasets, we evaluate synthetic augmentations generated as in Sec.~\ref{sec:methods:synthetic-data}, using $\epsilon=0.01$, and study scaling behavior as the multiplier (“Mult.” in Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random}) increases.
-We also benchmark a random dataset at various cardinalities;
-we use a random dataset with base cardinality of $10^6$ and dimensionality $128$, and refer to this dataset as ``Random.''
-This benchmark isolates the effect of manifold structure—expected to be absent in purely random data—on CAKES algorithms.
-
-\subsection{SILVA 18S}
-\label{sec:datasets-and-benchmarks:silva-18s}
-
-To demonstrate CAKES with an exotic distance function, we use the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}, which contains rRNA sequences from 2,224,640 genomes;
-the longest sequence being 3,712 letters.
-Holding out 1,000 random sequences as queries, we build the tree and perform $k$-NN using Levenshtein distance~\cite{levenshtein1966binary} on \emph{unaligned} sequences.
-Although this dataset is a multiple sequence alignment of width 50,000, we do not use Hamming distance on the aligned strings so as to demonstrate CAKES's flexibility.
-Under Hamming, the embedding dimension would be 50,000;
-under Levenshtein, the dimensionality is 3,712.
-
-\subsection{Radio ML}
-\label{sec:datasets-and-benchmarks:radio-ml}
-
-We also use the Radio-ML dataset~\cite{oshea2018radioml} under Dynamic Time Warping~\cite{muller2007dynamic} as the metric.
-This dataset contains 24 modulation modes spanning 26 SNR levels from $-20$dB to $30$dB, with 4,096 samples per combination, yielding 2,555,504 samples.
-Each sample is a 1,024-dimensional complex-valued vector (a time-series).
-For the benchmarks, we use the 10dB SNR subset with 97,304 samples and a 1,000-sample hold-out set of queries.
+In order to study scaling behavior as cardinality increases, we synthetically augment the ANN-Benchmarks datasets by generating new points by applying small random perturbations to existing points in the original dataset.
+Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random} show the results as the cardinality increases, with ``Mult.'' indicating the multiplier by which the cardinality was augmented.
+Since the SilVA-18S and Radio-ML datasets are already large and are not vector datasets, we do not augment them;
+instead, we subsample them to lower cardinalities for the scaling experiments.
 
 \subsection{Local Fractal Dimension}
-\label{sec:dayasets:lfd-of-datasets}
+\label{sec:datasets:lfd-of-datasets}
 
-Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks (Figure~\ref{fig:results:lfd-plots}).
-In this section, when we discuss LFD, unless otherwise noted, we are referring to the 95th percentile from Figure~\ref{fig:results:lfd-plots}.
-
-For Fashion-MNIST, Figure~\ref{fig:results:fashion-mnist-lfd}, we observe that until depth 5, LFD is less than 4. 
-It then starts increasing, reaching a peak of about 6 near depth 20.
-In comparison, Glove-25 has even lower LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}. 
-In particular, Glove-25's LFD is less than 3 for all depths.
-Figure~\ref{fig:results:sift-lfd} shows the LFD for Sift, which is higher relative to both Fashion-MNIST and Glove-25, increasing sharply to a peak of 9 around depth 10.
-For Random, which has the same cardinality and dimensionality as Sift, LFD starts at 20 the root cluster, and all percentile values decrease linearly with depth until reaching the leaves of the tree.
-The spread in LFD starts narrow for the first few depths and widens as depth increases.
-Silva, Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD;
-less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
-For the Radio-ML dataset, Figure~\ref{fig:results:radioml-lfd} LFD values show three distinct peaks around an LFD of 12 near depths of 8, 25 and 50.
-Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
-Within each of the three spans of depths, the character of Radio-ML is similar to that of Random.
+Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine, in Figure~\ref{fig:results:lfd-plots}, the LFD of each dataset we used for benchmarks.
 
@@
     \caption{Local fractal dimension vs.
-cluster depth across six datasets.
-The ‘random’ dataset is generated as in Sec.~\ref{sec:datasets-and-benchmarks:random-datasets}; note the different y-axis.
+cluster depth across six datasets; note the different y-axis.
 The x-axis is tree depth; the y-axis is LFD at that depth.
-We plot the 5th, 25th, 50th, 75th, and 95th percentiles, plus min/max.
-Each cluster is counted by its cardinality so the curves reflect dataset-level distributions.}
+We plot the 5th, 25th, 50th, 75th, and 95th percentiles, and min/max.
+The Fashion-MNIST, Glove-25, Sift and Silva datasets are real-world datasets and exhibit low LFD indicating the presence of manifold structure.
+The Random dataset has high LFD, indicating lack of manifold structure.
+The RadioML dataset is also a randomly generated dataset, but with a much more elaborate generation process that imparts some manifold structure at three distinct radial scales, indicated by the three peaks in LFD at different depths in the tree.}
     \label{fig:results:lfd-plots}
 \end{figure}
-
-% Since the CAKES algorithms were designed to scale in time with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
-% Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
-% In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
-
-% For the Fashion-MNIST dataset we observe in Figure~\ref{fig:results:fashion-mnist-lfd} that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
-% It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
-
-% Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
-% In particular, Glove-25's LFD is less than 3 for all depths.
-
-% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
-% It increases sharply to a peak of 9 around a depth of 10, after which it decreases smoothly until reaching the deepest leaves in the tree.
-
-% We contrast the low LFD of Sift with that of the Random dataset which has the same cardinality and dimension.
-As shown in Figure~\ref{fig:results:random-lfd}, the character of this dataset is very different from the others.
-% The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
-% The spread in LFD starts very small for the first few clusters and increases as depth increases.
-
-% The Silva dataset, shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-% In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
-
-% For the Radio-ML dataset, the LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
-% Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
-% Within each of the tree portions, this dataset has a character similar to that of the Random dataset.
-% This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
-% This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
 
@@
 
-We benchmark the three CAKES algorithms against a na\"ive linear search implementation in Rust.
-We also benchmark against three state-of-the-art similarity search algorithms: HNSW, ANNOY, and FAISS-IVF in Python.
-We verified that our linear search produces the same results as in the ground truth of the ANN-Benchmarks suite for Fashion-MNIST, Glove-25 and Sift datasets.
-We then used this implementation to find and store the ground-truth for the augmented versions of those datasets, and use it to calculate recall for all algorithms.
-We plot the results of these benchmarks in Figure~\ref{fig:results:scaling-plots}.
+Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, and \ref{tab:results:qps-and-recall-random} show the throughput (QPS) and recall for each algorithm on each of the vector datasets.
+Since HNSW, ANNOY, and FAISS do not support the Levenshtein or DTW distance functions, we only benchmark CAKES on the SilVA-18S and RadioML datasets and report results in Figures~\ref{fig:results:silva-scaling} and~\ref{fig:results:radioml-scaling}.
+All benchmarks were run with $k=10$, the throughput is averaged over all queries, and recall is computed using a na\"ive linear scan.
+
+HNSW, ANNOY, and FAISS have various hyper-parameters that can be tuned to trade off throughput, recall, and indexing time.
+We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the highest throughput configuration that achieves perfect recall (or as close to it as possible) for each dataset and cardinality.
+To make the comparison fair, we also allow CAKES to select the optimal algorithm among the three CAKES algorithms for each dataset and cardinality by using a small number of the centers of clusters at low depth as representative queries.

### 2025-11-05 sections/4-results.tex
down to 7 pages

--- 2025-08-29 sections/4-results.tex
@@
 
-For every dataset and distance funciton listed in Table~\ref{tab:datasets:summary}, we benchmark $k$-NN search using $k=10$.
-Smaller values of $k$, as described in Section~\ref{sec:introduction}, would be too sensitive to local perturbations, while larger values can capture too much of the global structure.
+Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, and \ref{tab:results:qps-and-recall-random} summarize the throughput and recall of CAKES against state-of-the-art $k$-NN search algorithms on the Fashion-MNIST, Glove-25, Sift, and Random datasets respectively.
+Throughput is measured in queries per second (QPS), and a recall of $1.000*$ indicates imperfect recall that rounds to $1.000$.
 
-In Figure~\ref{fig:results:scaling-plots}, we show how the throughput of each algorithm scales with cardinality for each of the six datasets we examined.
-To do this, we synthetically augmented the Fashion-MNIST, Glove-25, Sift, and Random datasets to create larger datasets using the process described in Section~\ref{sec:methods:synthetic-data}.
-For Silva and RadioML, due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset to examine how performance scales with cardinality.
-These plots also illustrate how the CAKES algorithms compare to na\"{i}ve linear search.
+We see that CAKES has perfect recall on all datasets using the Euclidean distance metric, and near-perfect recall on the Glove-25 dataset using Cosine distance.
+This is because Cosine distance violates the triangle inequality and is not a metric.
+The throughput of CAKES drops slightly for the first two multipliers, but then remains relatively constant as cardinality increases, demonstrating the scalability of CAKES on manifold-structured data.
+In contrast, the recall of HNSW and ANNOY degrades significantly as cardinality increases, even though their throughput remains high.
+FAISS-IVF keeps high recall but its throughput degrades linearly with cardinality.
+This indicates that these approximate algorithms are unable to maintain both high recall and high throughput as dataset sizes increase, even on manifold-structured data.
 
-Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift} and~\ref{tab:results:qps-and-recall-random} compare the performance (throughput and recall) of the CAKES algorithms against state-of-the-art algorithms on the Fashion-MNIST, Glove-25, Sift and Random datasets.
-In particular, we examine performance of HNSW, ANNOY and FAISS-IVF on each of those datasets as well as synthetically augmented versions of the datasets to isolate the effect of dataset size on performance.
-We did not perform similar benchmarks on the Silva and RadioML datasets because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
-For HNSW, ANNOY, and FAISS-IVF, we allow for a hyper-parameter search to tune their index for maximum recall.
-For CAKES, we build the tree and use our auto-tuning approach (see Section~\ref{sec:methods:auto-tuning}) to select the fastest algorithm for each dataset and cardinality.
+For the Random dataset, CAKES again attains perfect recall at all cardinalities, while HNSW and ANNOY exhibit near-zero recall, and FAISS-IVF attains perfect recall but with very low throughput.
 
-Though the plots in Figure~\ref{fig:results:scaling-plots} present results for each of CAKES's three algorithms separately, the results in the CAKES column in these tables represent the fastest CAKES algorithm at that dataset and cardinality only.
+On the Silva dataset (Fig.~\ref{fig:results:silva-scaling}), all algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then level off at higher cardinalities.
+CAKES's Depth-First Sieve is consistently the fastest algorithm on this dataset.
 
-Throughput is measured in queries per second.
-A recall value of $1.000*$ denotes imperfect recall that rounds to $1.000$.
-In each of Tables \ref{tab:results:qps-and-recall-fmn},\ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, we observe that while recall for CAKES does \emph{not} degrade with cardinality, recall for
-HNSW and ANNOY does degrade with cardinality. CAKES exhibits perfect recall on the Fashion-MNIST and Sift datasets, and near-perfect recall on the Glove-25 dataset (which uses cosine distance).
-
-% Finally, we test our intuitions about unbalanced clustering being better for search
-% than balanced clustering (see Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} and Figure~\ref{fig:results:distance-counts}).
-
-% \subsection{Local Fractal Dimension of Datasets}
-% \label{sec:results:lfd-of-datasets}
-
-% Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine the LFD of each dataset we used for benchmarks.
-% Figure~\ref{fig:results:lfd-plots} illustrates the trends in LFD for Fashion-MNIST, Glove-25, Sift, Random, Silva 18S, and Radio-ML.
-% In this section, when we discuss trends in LFD, unless otherwise noted, we are referring to the 95$^{th}$ percentile of LFD.
-% This is because our algorithms scale exponentially in the LFD, a
-
-% The Fashion-MNIST dataset has an embedding dimension of 784 and uses the Euclidean distance metric.
-% In Figure~\ref{fig:results:fashion-mnist-lfd} we observe that until approximately depth 5, Fashion-MNIST's LFD is low (i.e., less than 4).
-% It then starts increasing, reaching a peak of about 6 near depth 20, before decreasing to 1 at the maximum depth.
-
-% The Glove-25 dataset has an embedding dimension of 25 and uses the cosine distance function which, notably, is not a metric.
-% Relative to Fashion-MNIST, Glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
-% All percentile lines for Glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
-% In particular, Glove-25's LFD is less than 3 for all depths.
-
-% The Sift dataset has an embedding dimension of 128 and uses the Euclidean distance metric.
-% Figure~\ref{fig:results:sift-lfd} shows the LFD by depth for Sift, which has higher LFD relative to Fashion-MNIST and Glove-25.
-% It increases sharply to a peak of 9 around a depth of 10.
-% It then decreases smoothly until reaching the deepest leaves in the tree.
-
-% We generated the Random dataset to have the same cardinality and dimensionality as Sift.
-% We used a uniform distribution in a 128-dimensional unit-hypercube to generate the points and the Euclidean metric to measure distances among them.
-% Figure~\ref{fig:results:random-lfd} shows that the character of this dataset is significantly different from the others.
-% The LFD starts at 20 at depth 0 and all percentile lines decrease linearly with depth until reaching the leaves of the tree.
-% The spread in LFD starts very small for the first few clusters and increases as depth increases.
-% The LFD of approximately 20 for the root cluster $\mathcal{R}$ is what we expect for this random dataset.
-% To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, i.e.,\,the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
-% As a result, $\mathcal{R}$'s radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
-% A consequence of this is that, with high probability, for every point in $\mathcal{R}$, its distance from $c$ is greater than $\tfrac{r}{2}$;
-% in other words, $B(c, \tfrac{r}{2})$ contains only $c$ while $B(c, r)$ contains the entire dataset.
-% Given our definition of LFD in Equation~\ref{eq:methods:lfd-half}, this means that the LFD of $\mathcal{R}$ is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
-% Theoretically, the LFD of this dataset should be 128, i.e.\, it should be the same as the embedding dimension.
-% This reflects the difference between how we empirically measure the LFD and the value we would expect.
-% With sample sizes larger than 1,000,000, we would expect the LFD to approach 128 until we have sampled $2^{128}$ points, at which point the LFD would be 128 and would stay at 128 for even larger sample sizes.
-% Unfortunately, such a large sample is practically impossible to generate.
-
-% The Silva-18S dataset consists of genomic sequences whose unaligned lengths are at-most 3,712.
-% As such the embedding dimension is 3,712, though as discussed previously, the embedding dimension would be 50,000 in a multiple sequence alignment.
-% We use the Levenshtein edit distance (a metric) to measure distances between sequences.
-% This dataset, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
-% In particular, LFD is less than 3 for all depths, hovering near 1 for clusters at depth 40 and deeper.
-
-% The Radio-ML dataset consists of measurements of radio-frequency signals using 1,024 dimensional complex-valued vectors.
-% We use the Dynamic Time Warping distance metric on this dataset.
-% This dataset is synthetic~\cite{oshea2018radioml} but uses a far more elaborate generation process than our Random dataset.
-% The LFD values show three distinct peaks around an LFD of 12 at or near depths of 8, 25 and 50.
-% Each peak is followed by a linear decrease until encountering a sharp spike for the next peak.
-% Within each of the tree portions, this dataset has a character very similar to that of the Random dataset.
-% This suggests that the dataset obeys the manifold hypothesis at some scales, but that it is not ``scale free,'' as the LFD varies significantly by depth.
-% This is likely the result of a piecewise uniform sampling strategy used to generate the different modulation modes present in the dataset.
+On the Radio-ML dataset (Fig.~\ref{fig:results:radioml-scaling}), all CAKES variants are slower than na\"{i}ve linear search, and their throughput decreases linearly with cardinality.
+This is expected, as the Radio-ML dataset is a randomly generated dataset and exhibits little manifold structure under the DTW distance metric.
 
@@
   \centering
-  \subfloat[Fashion-MNIST]{
-    \includegraphics[width=0.48\columnwidth]{plots/fashion-mnist_PermutedBall_10_throughput.png}
-    \label{fig:results:fashion-mnist-scaling}
-  }
-  \subfloat[Glove-25]{
-    \includegraphics[width=0.48\columnwidth]{plots/glove-25_PermutedBall_10_throughput.png}
-    \label{fig:results:glove-25-scaling}
-  }
-
-  \vskip -0.1in
-  \subfloat[Sift]{
-    \includegraphics[width=0.48\columnwidth]{plots/sift_PermutedBall_10_throughput.png}
-    \label{fig:results:sift-scaling}
-  }
-  \subfloat[Random dataset]{
-    \includegraphics[width=0.48\columnwidth]{plots/random_PermutedBall_10_throughput.png}
-    \label{fig:results:random-scaling}
-  }
-
-  \vskip -0.1in
   \subfloat[Silva]{
@@
 
-  \caption{Throughput across six datasets, including a randomly-generated dataset.
-  Each plot shows throughput (queries per second; higher is better) versus dataset cardinality.
-For Fashion-MNIST, Glove-25, and Sift, CAKES becomes faster than linear search as cardinality grows, with the crossover point varying by dataset.
-For Fashion-MNIST and Glove-25, Depth-First Sieve is consistently fastest.
-For Sift, Repeated $\rho$-NN is fastest at small cardinalities, while Depth-First Sieve is fastest at large cardinalities.
-For Silva, throughput for all algorithms initially decreases approximately linearly with cardinality and then levels off at higher cardinalities; Depth-First Sieve is consistently fastest.
-For Radio-ML and Random, all CAKES variants are slower than na\"{i}ve linear search, and their throughput decreases linearly with cardinality.
-HNSW and ANNOY are the fastest algorithms on all four datasets we benchmarked them on, but their recall degrades quickly as cardinality increases on all datasets; on Random, their recall is near zero.}
+  \caption{Throughput vs. dataset cardinality on Silva-18S and Radio-ML.}
   \label{fig:results:scaling-plots}
@@
 \end{table}
-
-% \begin{figure}
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
-%         \subcaption{Repeated $\rho$-NN}
-%         \label{fig:results:fashion-mnist-counts-throughput}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}
-%         \subcaption{Repeated $\rho$-NN}
-%         \label{fig:results:glove-25-counts-counts}
-%     \end{subfigure}%
-%     \\
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
-%         \subcaption{Breadth First Sieve}
-%         \label{fig:results:sift-counts-throughput}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
-%         \subcaption{Breadth First Sieve}
-%         \label{fig:results:random-counts-counts}
-%     \end{subfigure}%
-%     \\
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
-%         \subcaption{Depth First Sieve}
-%         \label{fig:results:silva-counts-throughput}
-%     \end{subfigure}%
-%     \begin{subfigure}[b]{0.5\textwidth}
-%         \includegraphics[width=0.99\textwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
-%         \subcaption{Depth First Sieve}
-%         \label{fig:results:radioml-counts-counts}
-%     \end{subfigure}%
-%     \\
-%     \begin{subfigure}[b]{0.94\textwidth}
-%         \centering
-%         \includegraphics[width=0.6\textwidth]{images/distance_counts/legend.png}
-%         \label{fig:results:counts-legend}
-%     \end{subfigure}%
-%     \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
-%     Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
-%     The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
-%     The x-axis represents increasing cardinality of the dataset.}
-%     \label{fig:results:distance-counts}
-% \end{figure}
-
-\subsection{Clustering Strategies}
-\label{sec:results:clustering-strategies-and-number-of-distance-computations}
-
-Beyond scaling behavior, we examine how clustering strategy affects search performance.
-We consider four strategies—the Cartesian product of balanced vs.\ unbalanced clustering and the presence vs.\ absence of depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}).
-We instrument CAKES to count distance computations during search;
-this adds substantial overhead and would not be used in real-world applications.
-
-Figure~\ref{fig:results:distance-counts} compares \textit{Ball}, \textit{BalancedBall}, \textit{PermutedBall}, and \textit{PermutedBalancedBall} (``Permuted'' denotes depth-first reordering) for all three algorithms on Fashion-MNIST.
-Balanced clustering is strictly worse in throughput and has poorer asymptotics: algorithms such as Depth-First Sieve that appear nearly constant-time in Fig.~\ref{fig:results:scaling-plots} (with \textit{PermutedBall}) no longer do so under balanced clustering.
-On this dataset, depth-first reordering has little effect on throughput; it is primarily intended to improve memory consumption.
-
-\begin{figure}[h]
-  \centering
-  \subfloat[Repeated $\rho$-NN]{
-  \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
-  \label{fig:results:fashion-mnist-rnn-throughput}
-  }
-  \subfloat[Repeated $\rho$-NN]{
-    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}    
-    \label{fig:results:fashion-mnist-rnn-counts}
-  }
-
-  \subfloat[Breadth First Sieve]{
-    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
-    \label{fig:results:fashionmnist-bfs-throughput}
-    }
-  \subfloat[Breadth First Sieve]{
-    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
-    \label{fig:results:fashionmnist-bfs-counts}
-    }
-
-  \subfloat[Depth First Sieve]{
-    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
-    \label{fig:results:fashionmnist-dfs-throughput}
-    }
-  \subfloat[Depth First Sieve]{
-    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
-    \label{fig:results:fashionmnist-dfs-counts}
-  }
-
-  \vspace{2pt}
-  \includegraphics[width=0.75\columnwidth]{images/distance_counts/legend.png}
-
-  \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
-  Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
-  The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
-  The x-axis represents increasing cardinality of the dataset.}
-  \label{fig:results:distance-counts}
-\end{figure}

### 2025-11-05 sections/5-discussion.tex
down to 7 pages

@@
 We have presented CAKES, a suite of three algorithms for fast $k$-NN search across diverse distance functions.
-When the distance is a metric (Sec.~\ref{sec:methods}), the algorithms are exact.
-Under Cosine distance, a non-metric, they achieve nearly perfect recall.
-CAKES is most effective when data satisfy the manifold hypothesis, i.e., they occupy a low-dimensional manifold within a high-dimensional space.
-It builds an unbalanced cluster tree that captures manifold structure and helps accelerate search.
-The results in Section~\ref{sec:results:clustering-strategies-and-number-of-distance-computations} shows that this unbalanced clustering approach is superior to balanced clustering in both absolute and asymptotic performance.
-Conversely, performance degrades on randomly distributed data, where such structure is absent.
+When the distance is a metric (Sec.~\ref{sec:methods}), CAKES is exact.
+Under Cosine distance, a non-metric, CAKES achieves near-perfect recall.
+Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove} and \ref{tab:results:qps-and-recall-sift}, and Figure~\ref{fig:results:silva-scaling} show that CAKES is most effective when data satisfy the manifold hypothesis, i.e., they occupy a low-dimensional manifold within a high-dimensional space.
+Conversely, performance degrades on randomly generated data, where such structure is absent, as shown in Table~\ref{tab:results:qps-and-recall-random} and Figure~\ref{fig:results:radioml-scaling}.
+
 In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
+As expected, CAKES scales sub-linearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD;
+Sift versus Random illustrates this contrast, as does Radio-ML.
 
-\subsection{Performance}
-
-As expected, Fig.~\ref{fig:results:scaling-plots} shows CAKES scaling sublinearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD;
-Sift versus Random illustrates this contrast.
-On Random, CAKES' throughput decreases linearly with increasing cardinality, whereas on Sift, Depth-First Sieve and Breadth-First Sieve have near-constant throughput.
-Repeated $\rho$-NN is not near-constant on Sift but decays much more slowly than on Random.
-Because Sift and Random share cardinality and embedding dimension, this isolates the effect of manifold structure on performance.
-Although CAKES is slower on Random, it maintains perfect recall.
-
-HNSW and ANNOY maintain near-constant throughput as the cardinality grows, but their recalls decline and their index build times rise, making them unsuitable when datasets grow exponentially.
-CAKES's tree builds orders of magnitude faster; while queries are slower than HNSW/ANNOY, CAKES attains near-constant scaling with perfect recall.
-
-We stress that CAKES targets \emph{big} data, and our results support this claim.
-We observe that while linear search can outperform CAKES on datasets with small cardinality, CAKES always overtakes linear search at a large enough cardinality for each of the ANN-benchmark datasets we tested.
-Moreover, we observe that CAKES's algorithms outperform linear search at a lower cardinality when the dataset has a lower LFD.
-For example, Sift has much higher LFD than Fashion-Mnist and Glove-25, and we observe that CAKES overtakes linear search at a cardinality of $10^{7}$  for Sift, as opposed to about $10^{5}$ and $10^{6}$ for Fashion-Mnist and Glove-25, respectively.
-For the Random dataset, which has the highest LFD, CAKES's algorithms \emph{never} outperform linear search.
-
-% When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{T}{\lceil d \rceil} + L)$, where $T$ is the time complexity of tree-search and $L$ is the time complexity of leaf-search.
-% Even though Repeated $\rho$-NN has the least time cost (compared to $\mathcal{O}\left((T + L)\log{(T+L)}\right)$ for Breadth-First Sieve and $\mathcal{O}(T\log{T} + L\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
-% We believe that some of this discrepancy can be explained by the fact that Repeated $\rho$-NN can significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search.
-% This causes leaf-search to exhaustively search many more clusters than necessary, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
-% This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
-% For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD,
-% the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase in (Equation)~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
-% This suggests that rather than using the reciprocal of the harmonic mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
-% We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly overestimates the correct radius $\rho_k$ and to improve upon the factor of radial increase in Equation~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
-
-\subsection{Future Work}
+With HNSW, ANNOY, and FAISS-IVF failing to maintain both high recall and high throughput as cardinality grows, CAKES fills an important gap in the $k$-NN search literature.
+It is a scalable solution for exact $k$-NN search on large real-world datasets that obey the manifold hypothesis, and whose cardinalities are expected to grow exponentially over time.
+Most datasets collected from constrained physical processes (e.g., neural embeddings, genomic sequences, sensor data, etc.) are expected to satisfy the manifold hypothesis, making CAKES widely applicable.
 
 This study can be extended in several directions.
-We can start by evaluating more datasets and distance functions not supported by FAISS, HNSW, or ANNOY, such as Wasserstein~\cite{vallender1974calculation} for probability distributions (particularly high-dimensional) and Tanimoto~\cite{bajusz2015tanimoto} for molecular structures.
-
-We can also improve the augmentation procedure (Sec.~\ref{sec:methods:synthetic-data}) to better preserve topological structure of datasets, e.g., by favoring perturbations along the leading principal components of the local manifold.
-This would better test the scaling behavior of various algorithms.
-
-We have noted how the fastest algorithm in CAKES changes depending on the dataset and distance function being used, as well as the cardinality of the dataset.
-This variability motivated our auto-tuning process (Sec.~\ref{sec:methods:auto-tuning}).
-Future work should examine why and when each algorithm excels and learn to use the geometric and topological properties of datasets to choose the ideal algorithm to use.
-
-CAKES algorithms can also be extended for use in a streaming environment, enabling online updates to the tree as points are inserted or deleted.
-
-Despite CAKES' generality, effective deployment on real data requires domain knowledge.
-In protein sequence search, for example, sequences vary widely in length (typically 30-1000 amino acids, with outliers) and have a 20-letter alphabet.
-Numerous algorithms exist~\cite{kim2021entrance, daniels2013compressive, yu2015entropy, steinegger2018clustering}, and their development relied on deep domain expertise.
-Simple edit distance can be poor when lengths vary substantially.
-CAKES can get around that issue by instead using alternative metrics or preprocessing—for example, the dataset can be partitioned into mutually exclusive length bins with a separate tree per bin;
-at query time, we simply route the sequence to its bin by length, alogn with a small number of adjacent bins.
+More specialized distance functions, such as Wasserstein~\cite{vallender1974calculation} for probability distributions and Tanimoto~\cite{bajusz2015tanimoto} for molecular structures, would require only a Rust implementation of the distance function itself to be compatible with CAKES.
+An examination of the geometric and topological properties of datasets that favor one CAKES algorithm over another would also be valuable.
 

### 2025-11-05 sections/1-introduction.tex
removed another half a column

@@
 This ``Big Data explosion'' has created a need for algorithms that scale efficiently to large datasets.
-One such class of algorithms is similarity search algorithms, which enable applications such as recommendation~\cite{annoy} and classification ~\cite{suyanto2022knnclassifier}.
-Yet as datasets grow in size and dimensionality, efficient and accurate similarity search becomes increasingly challenging;
-even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{malkov2016hnsw, johnson2019billion, annoy, aumuller2020ann}.
+One such class of algorithms is similarity search;
+yet as datasets grow, fast and accurate similarity search becomes increasingly challenging.
+Even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{malkov2016hnsw,johnson2019billion,annoy,aumuller2020ann}.
 
-Given some measure of similarity between data points, two common similarity search paradigms are $k$-nearest neighbor search ($k$-NN) and $\rho$-nearest neighbor search ($\rho$-NN).
-The former retrieves the $k$ points closest to a query, while the latter retrieves all points within a similarity threshold $\rho$ of a query.
-While prior work has used the term \textit{approximate} search to refer to $\rho$-NN, we use it more narrowly: an {approximate} search algorithm yields imperfect recall, whereas an \textit{exact} algorithm guarantees perfect recall.
-$k$-NN search is one of the most widely-used methods in classification and recommendation~\cite{fix1952discriminatory, cover1967nearest}, but na\"{i}ve implementations are prohibitively slow because they scale linearly with dataset size.
-Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.
+Given some distance function to measure similarity between data points, $k$-nearest neighbors ($k$-NN) search is the task of finding the $k$ points in a dataset that are closest to a given query point.
+We use the term \textit{approximate} to refer to algorithms that may not return all true nearest neighbors, while \textit{exact} refers to algorithms that guarantee perfect recall as measured against a brute-force search.
+Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for fast \textit{and} exact search remains~\cite{ukey2023survey}.
 
-This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for \emph{exact} $k$-NN search.
-We benchmark CAKES against FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
+We benchmark these against FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
 We also evaluate performance on SILVA 18S~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and RadioML~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
@@
 
-Several algorithms have been developed to tackle $k$-nearest neighbor search on large datasets, including Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, random projection and tree building (ANNOY)~\cite{annoy}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}.
+Several algorithms have been developed to tackle $k$-NN search on large datasets, including entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}, Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, and random projection and tree building (ANNOY)~\cite{annoy}.
 
@@
 
-Entropy-scaling algorithms exploit the inherent structure of large datasets, achieving complexity that scales with topological properties, such as metric entropy and fractal dimension, rather than cardinality.
+Entropy-scaling algorithms exploit the inherent manifold structure of large datasets, achieving complexity that scales with topological properties, such as metric entropy and fractal dimension, rather than cardinality.
 This makes them especially suitable for manifold-constrained data.
 
-In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended the flat clustering from~\cite{yu2015entropy} to a hierarchical approach.
-CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, refines the clustering algorithm from CHESS.
-
-This paper introduces CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
-We provide a theoretical complexity analysis in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
-These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
+In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
+These algorithms build upon previous work in CHESS~\cite{ishaq2019clustered}, improving the clustering method and extending to $k$-NN search.
+We provide a theoretical analysis of CAKES in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
+These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a random uniform distribution.
 Given that CAKES's algorithms are intended for datasets with a manifold structure, an analysis assuming a uniform distribution would be uninformative.

### 2025-11-05 sections/2-methods.tex
removed another half a column

@@
 \end{equation}
-where $B(q, r)$ is the set of points in the metric ball of radius $r$ centered at $q$.
+where $B(q, r)$ is the set of points within radius $r$ from $q$.
 
-% In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
-% While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
-% they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
-% CAKES exploits the low LFD of such datasets to accelerate search.
-
-Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases.
-When the vast majority of points in the dataset have low ($d \ll D$) LFD, we simply say that the dataset has low LFD.
+When the majority of points in the dataset have low ($d \ll D$) LFD, we simply say that the dataset has low LFD.
 For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but (often) globally variable.
@@
 
-We define a \emph{cluster} $C$ as a set of points with a \emph{center} $c$ and \emph{radius} $r$.
-The center is the geometric median of the points in the cluster (i.e., the point minimizing the sum of distances to all other points in $C$) and the radius is the maximum distance from the center to any point in $C$.
+We define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
+The center $c$ is the geometric median of the points in the cluster (i.e., the point minimizing the sum of distances to all other points in $C$) and the radius $r$ is the maximum distance from $c$ to any point in $C$.
 A cluster with only one point is a $leaf$, otherwise a $parent$ cluster has two $child$ clusters.
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds the tree $\mathcal{T}$ in $\mathcal{O}(n \log n)$ time as described in~\cite{ishaq2019clustered}.
+Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds the tree $\mathcal{T}$ in $\mathcal{O}(n \log n)$ time.
+
+\begin{figure}[t]
+    \vskip -0.1in
+    \centering
+    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
+    \caption{
+        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
+        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
+        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+    }
+    \label{fig:methods:deltas}
+    \vskip -0.05in
+\end{figure}
 
@@
         \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
-            \State $L \Leftarrow \Call{Partition}{L, criteria}$
-            \State $R \Leftarrow \Call{Partition}{R, criteria}$
-            % \State $L \Leftarrow Partition(L, criteria)$
-            % \State $R \Leftarrow Partition(R, criteria)$
+        \State \Call{Partition}{L, criteria}
+        \State \Call{Partition}{R, criteria}
         \EndIf
@@
 
-\begin{figure}[t]
-    \vskip -0.2in
-    \centering
-    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
-    \vskip -0.1in
-    \caption{
-        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
-        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
-        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
-    }
-    \label{fig:methods:deltas}
-    \vskip -0.2in
-\end{figure}
-% \FloatBarrier
+\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
+\label{sec:methods:knn-search}
 
-% \FloatBarrier
+Given a query $q$ and integer $k$, $k$-NN search finds the $k$ closest points to $q$ in $\textbf{X}$. In this section, we present three novel algorithms for exact $k$-NN search:
+Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
+We use $Q$ for the data structure containing the clusters that are still in contention for being among the $k$ nearest neighbors, and $H$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
+These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
+\label{sec:methods:knn-search:repeated-rnn}
+
+We begin using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
+If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
+Now, so long as $\sum_{C \in Q} |C| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{C \in Q} |C|}} \big)^\frac{1}{\mu}$
+where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
+Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
+Once $\sum_{C \in Q} |C| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
+
 \begin{algorithm}[!t]
@@
         \State \textbf{return} \Call{TreeSearch}{$L, q, r$} $\cup$ \Call{TreeSearch}{$R, q, r$}
-        % \State \textbf{return} tree-search($L, q, r$) $\cup$ tree-search($R, q, r$)
     \end{algorithmic}
@@
 
-\subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
-\label{sec:methods:knn-search}
-
-Given a query $q$ and integer $k$, $k$-NN search finds the $k$ closest points to $q$ in $\textbf{X}$. In this section, we present three novel algorithms for exact $k$-NN search:
-Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-We use $H$ to refer to the data structure containing the nearest neighbors found so far, and $Q$ for the data structure containing the clusters and points that are still in contention for being among the $k$ nearest neighbors.
-These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
-
-\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{sec:methods:knn-search:repeated-rnn}
-
-We begin using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
-If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
-Now, so long as $\sum_{C \in Q} |C| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{C \in Q} |C|}} \big)^\frac{1}{\mu}$
-where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
-Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
-
-Once $\sum_{C \in Q} |C| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
-
 \subsubsection{Complexity of Repeated \texorpdfstring{$\rho$}{p}-NN}
@@
 \end{gather}
-We now extend this to Repeated $\rho$-NN.
 
-We first need to estimate the number of iterations of \textsc{TreeSearch} needed to find a radius that guarantees at least one neighbor.
+To extend this to Repeated $\rho$-NN, we first need to estimate the number of iterations of \textsc{TreeSearch} needed to find a radius that guarantees at least one neighbor.
 Since we expect $q$ to be from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the nearest neighbor.
@@
 
-For \textsc{LeafSearch}, we must estimate $\sum_{C \in Q} |C|$, the total cardinality of those clusters;
+For \textsc{LeafSearch}, we must estimate $\sum_{C \in Q} |C|$;
 since we must examine every point in each such cluster, time complexity of \textsc{LeafSearch} is linear in this quantity.
@@
 We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $Q$.
-Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot \left( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of the LFDs of the clusters in $Q$.
+Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot \left( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of LFDs in $Q$.
 It remains to estimate $\rho_k$.
@@
 
-Again using the assumption that $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at a cluster in $Q$, we have that $\hat{d} \approx d$.
+Again using the assumption that $q$ is from the same distribution as $X$, we have that $\hat{d} \approx d$.
 Thus, the asymptotic complexity of Repeated $\rho$-NN is
@@
 
-We begin by letting $Q$ be a list of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is either a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
-During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{c}, 1)$ to $Q$, where $c$ is the center of $C$.
+We begin by letting $Q$ be a list of 3-tuples $(C, \delta^{+}, m)$, where $C$ is a cluster, $\delta^{+}$ is as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $C$ in $Q$.
+$Q$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $H$ starts with the center of $\mathcal{R}$.
 
-We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$.
+We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $H$).
 Since this step may require a binary search of $Q$ and reordering $Q$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
 
-Next, we check every item in $Q$, removing any for which $\delta^{-} > \tau$ because they cannot contain (or be) one of the $k$ nearest neighbors.
-If an item corresponds to a point, we keep it.
-If it corresponds to a leaf cluster, we add all its points to $Q$ with a multiplicity of 1 each.
-Otherwise (since it must correspond to a parent cluster), we add to $Q$ the 3-tuples corresponding to its child clusters, thus expanding the search to the next level in $T$.
-We continue until the sum of multiplicities in $Q$ is exactly $k$.
-These are the $k$ nearest neighbors.
+Next, we check every cluster in $Q$, removing any for which $\delta^{-} > \tau$ because they cannot contain one of the $k$ nearest neighbors.
+For each of the remaining clusters, if it is a leaf, we remove it from $Q$ and add all its points to $H$.
+Otherwise (since it must be a parent cluster), we remove it from $Q$ and, for each of its child clusters, we compute $\delta^{+}$ and add the corresponding 3-tuple to $Q$ and the child's center to $H$, thus expanding the search to the next level in $T$.
+We continue until $Q$ is empty or the farthest point in $H$ is closer to $q$ than the $\delta^{-}$ of every cluster in $Q$.
+Thus, $H$ contains the $k$ nearest neighbors to the query.
 
@@
 
-This algorithm (DFS) uses a max-priority queue to track hits, and a min-priority queue to choose the next branch of $\mathcal{T}$ to explore, making it similar to a depth-first traversal.
+This algorithm (DFS) greedily descends into the child cluster whose center if closer to the query than that of its sibling.
 
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ and $H$ be a max-queue (with capacity $k$) of points prioritized by $\delta$.
-$Q$ starts containing only $\mathcal{R}$ while $H$ starts empty.
+Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ starting with only $\mathcal{R}$ while $H$ starts empty.
 So long as $H$ is not full or the top-priority items from $H$ and $Q$ have $\delta_H \geq \delta_Q$, we repeat following:
@@
 
-Due to their similarity, we combine the complexity analyses of both Sieve methods.
-For these methods we again use the terminology of \textsc{TreeSearch} and \textsc{LeafSearch}.
+Due to their similarity, we combine the analyses of BFS and DFS.
+
 Let $d$ be the LFD in the region around $q$, and consider clusters with cardinalities near $k$.
@@
 
-The asymptotic complexity of BFS is dominated by QuickSelect.
-Since $Q$ contains the clusters from \textsc{TreeSearch} and the points from \textsc{LeafSearch}, we see that the asymptotic complexity of BFS is
+The asymptotic complexity of BFS is dominated by our modified QuickSelect algorithm, which accounts for the clusters in $Q$ and the points in $H$, and is log-linear in their sum.
+Thus, the asymptotic complexity of BFS is
 \begin{equation}
-    \mathcal{O} \big( (T + L ) \log (T + L ) \big).
+    \mathcal{O} \big( (T + L) \log(T + L) \big).
     \label{eq:methods:bfs-complexity}
@@
 \begin{equation}
-    \mathcal{O} \big( T \log T + L \log k \big).
+    \mathcal{O} \big( T \log T + L \log L \big).
     \label{eq:methods:dfs-complexity}

### 2025-11-05 sections/3-datasets.tex
removed another half a column

@@
 These include the Fashion-MNIST, Glove-25, and Sift datasets from the ANN-Benchmarks suite~\cite{aumuller2020ann}, as well as a Random dataset we generated (using the same cardinality and dimensionality as Sift) as an anti-example to demonstrate the effect of manifold structure on CAKES performance.
-Thus, Random shows that CAKES is designed to exploit manifold structure when it exists, and should not be used on datasets that do not obey the manifold hypothesis.
 The Glove-25 dataset uses the Cosine distance function, while the other three datasets use the Euclidean distance metric.
@@
 
-We also include benchmarks on the SilVA-18S rRNA and Radio-ML datasets.
+We also include benchmarks on the SilVA-18S and Radio-ML datasets.
 For the SilVA-18S dataset, which consists of 2,224,640 sequences of up-to 3,712 nucleotides, we use the Levenshtein distance metric~\cite{levenshtein1966binary}.
 The Radio-ML dataset consists of 1,024-dimensional complex-valued time-series samples, spanning 24 modulation modes and 26 SNR levels from $-20$dB to $30$dB.
-We use the 10dB SNR subset with 97,304 samples and use Dynamic Time Warping~\cite{muller2007dynamic} as the distance metric.
+We use the 10dB SNR subset with 97,304 samples and Dynamic Time Warping~\cite{muller2007dynamic} as the distance metric.
 For both datasets, we hold out 1,000 random samples as queries and build the tree on the remaining samples.
-These were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM running Manjaro Linux 5.15.164-1-MANJARO with Rust compiler version 1.83.0.
+These were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM running Manjaro Linux 5.15.164-1-MANJARO.
 
@@
 The Random dataset has high LFD, indicating lack of manifold structure.
-The RadioML dataset is also a randomly generated dataset, but with a much more elaborate generation process that imparts some manifold structure at three distinct radial scales, indicated by the three peaks in LFD at different depths in the tree.}
+The RadioML dataset is also a randomly generated dataset, but with a much more elaborate generation process.
+It seems to have little manifold structure, if any, instead exhibiting behavior akin to the Random dataset across three separate radial scales, indicated by the three peaks in LFD at different depths in the tree.}
     \label{fig:results:lfd-plots}
@@
 Since HNSW, ANNOY, and FAISS do not support the Levenshtein or DTW distance functions, we only benchmark CAKES on the SilVA-18S and RadioML datasets and report results in Figures~\ref{fig:results:silva-scaling} and~\ref{fig:results:radioml-scaling}.
-All benchmarks were run with $k=10$, the throughput is averaged over all queries, and recall is computed using a na\"ive linear scan.
+All benchmarks were run with $k=10$, throughput is averaged over all queries, and recall is computed using a na\"ive linear scan.
 
 HNSW, ANNOY, and FAISS have various hyper-parameters that can be tuned to trade off throughput, recall, and indexing time.
-We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the highest throughput configuration that achieves perfect recall (or as close to it as possible) for each dataset and cardinality.
-To make the comparison fair, we also allow CAKES to select the optimal algorithm among the three CAKES algorithms for each dataset and cardinality by using a small number of the centers of clusters at low depth as representative queries.
+We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the configuration with the highest throughput that still achieves perfect recall (or as close to it as possible) for each dataset and cardinality.
+To make the comparison fair, we also allow CAKES to select the fastest of its algorithms for each dataset and cardinality by testing each with a small number of the centers of clusters at low depth as representative queries.

### 2025-11-05 sections/4-results.tex
removed another half a column

@@
 Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, and \ref{tab:results:qps-and-recall-random} summarize the throughput and recall of CAKES against state-of-the-art $k$-NN search algorithms on the Fashion-MNIST, Glove-25, Sift, and Random datasets respectively.
-Throughput is measured in queries per second (QPS), and a recall of $1.000*$ indicates imperfect recall that rounds to $1.000$.
+A recall of $1.000*$ indicates imperfect recall that rounds to $1.000$.
 
@@
 FAISS-IVF keeps high recall but its throughput degrades linearly with cardinality.
-This indicates that these approximate algorithms are unable to maintain both high recall and high throughput as dataset sizes increase, even on manifold-structured data.
-
 For the Random dataset, CAKES again attains perfect recall at all cardinalities, while HNSW and ANNOY exhibit near-zero recall, and FAISS-IVF attains perfect recall but with very low throughput.
 
-On the Silva dataset (Fig.~\ref{fig:results:silva-scaling}), all algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then level off at higher cardinalities.
-CAKES's Depth-First Sieve is consistently the fastest algorithm on this dataset.
+On Silva (Fig.~\ref{fig:results:silva-scaling}), all CAKES algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then level off at higher cardinalities.
+DFS is consistently the fastest algorithm on this dataset.
 
-On the Radio-ML dataset (Fig.~\ref{fig:results:radioml-scaling}), all CAKES variants are slower than na\"{i}ve linear search, and their throughput decreases linearly with cardinality.
-This is expected, as the Radio-ML dataset is a randomly generated dataset and exhibits little manifold structure under the DTW distance metric.
+On Radio-ML (Fig.~\ref{fig:results:radioml-scaling}), all CAKES variants are slower than na\"{i}ve linear search, and their throughput decreases linearly with cardinality.
 
-\subsection{Indexing and Tuning Time}
+\subsection{Indexing Time}
 \label{sec:results:indexing-and-tuning-time}
@@
   We report the time taken by various algorithms to build and tune its index on various datasets.
-  For CAKES, this includes the time required to build  the tree and perform auto-tuning.
+  For CAKES, this includes the time required to build the tree and find the fastest internal algorithm.
   All axes are on logarithmic scales.

### 2025-11-05 sections/5-discussion.tex
removed another half a column

@@
 When the distance is a metric (Sec.~\ref{sec:methods}), CAKES is exact.
-Under Cosine distance, a non-metric, CAKES achieves near-perfect recall.
+Even under a non-metric, namely Cosine distance, CAKES achieves near-perfect recall.
 Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove} and \ref{tab:results:qps-and-recall-sift}, and Figure~\ref{fig:results:silva-scaling} show that CAKES is most effective when data satisfy the manifold hypothesis, i.e., they occupy a low-dimensional manifold within a high-dimensional space.
 Conversely, performance degrades on randomly generated data, where such structure is absent, as shown in Table~\ref{tab:results:qps-and-recall-random} and Figure~\ref{fig:results:radioml-scaling}.
+HNSW and ANNOY do maintain high throughput but their recall degrades significantly as cardinality grows, and is near zero on the Random dataset.
+FAISS-IVF maintains high recall but its throughput degrades significantly as cardinality grows.
 
-In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset we tested exhibits a manifold structure, as quantified by the percentiles of LFD.
+In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset exhibits a manifold structure, as quantified by LFD.
 As expected, CAKES scales sub-linearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD;

### 2025-11-08 sections/1-introduction.tex
down to six pages

@@
 yet as datasets grow, fast and accurate similarity search becomes increasingly challenging.
-Even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{malkov2016hnsw,johnson2019billion,annoy,aumuller2020ann}.
+Even state-of-the-art algorithms exhibit a steep tradeoff between recall and throughput~\cite{malkov2016hnsw, johnson2019billion, annoy, aumuller2020ann}.
 
-Given some distance function to measure similarity between data points, $k$-nearest neighbors ($k$-NN) search is the task of finding the $k$ points in a dataset that are closest to a given query point.
-We use the term \textit{approximate} to refer to algorithms that may not return all true nearest neighbors, while \textit{exact} refers to algorithms that guarantee perfect recall as measured against a brute-force search.
+Given a distance function for the data, $k$-nearest neighbors ($k$-NN) search finds the $k$ points in a dataset that are closest to a given query point.
+We use the term \textit{approximate} to refer to algorithms that have imperfect recall, while \textit{exact} refers to algorithms that have perfect recall as measured against a brute-force search.
 Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for fast \textit{and} exact search remains~\cite{ukey2023survey}.
@@
 
-Several algorithms have been developed to tackle $k$-NN search on large datasets, including entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}, Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, and random projection and tree building (ANNOY)~\cite{annoy}.
-
+Modern k-NN search strategies include entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}, Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, and random projection and tree building (ANNOY)~\cite{annoy}.
 HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
 FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
-ANNOY~\cite{annoy} uses random projection and tree building.
 However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).
@@
 
-In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust.
-These algorithms build upon previous work in CHESS~\cite{ishaq2019clustered}, improving the clustering method and extending to $k$-NN search.
-We provide a theoretical analysis of CAKES in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}.
+In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust, extending CHESS~\cite{ishaq2019clustered} to $k$-NN search.
+Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods} provide a theoretical analysis.
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a random uniform distribution.

### 2025-11-08 sections/2-methods.tex
down to six pages

@@
 A cluster with only one point is a $leaf$, otherwise a $parent$ cluster has two $child$ clusters.
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds the tree $\mathcal{T}$ in $\mathcal{O}(n \log n)$ time.
+Using previous work in CHESS~\cite{ishaq2019clustered}, and CLAM~\cite{ishaq2021clustered}, we build a binary tree $\mathcal{T}$ of clusters by recursively partitioning the dataset $\textbf{X}$.
+% Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds the tree $\mathcal{T}$ in $\mathcal{O}(n \log n)$ time.
 
-\begin{figure}[t]
-    \vskip -0.1in
-    \centering
-    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
-    \caption{
-        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
-        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
-        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
-    }
-    \label{fig:methods:deltas}
-    \vskip -0.05in
-\end{figure}
+% \begin{figure}[t]
+%     \vskip -0.1in
+%     \centering
+%     \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
+%     \caption{
+%         {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
+%         ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
+%         ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
+%         ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
+%     }
+%     \label{fig:methods:deltas}
+%     \vskip -0.05in
+% \end{figure}
 
-\begin{algorithm} % enter the algorithm environment
-    \caption{\textsc{Partition}($C$, $criteria$)} % give the algorithm a caption
-    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
-    \begin{algorithmic} % enter the algorithmic environment
-        \Require $f$, a distance function
-        \Require $C$, a cluster
-        \Require $criteria$, user-specified continuation criteria
+% \begin{algorithm} % enter the algorithm environment
+%     \caption{\textsc{Partition}($C$, $criteria$)} % give the algorithm a caption
+%     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
+%     \begin{algorithmic} % enter the algorithmic environment
+%         \Require $f$, a distance function
+%         \Require $C$, a cluster
+%         \Require $criteria$, user-specified continuation criteria
 
-        \If{$|C| > 1$ \textbf{and} $C$ satisfies $criteria$}
-        \State $S \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-        \State $c \Leftarrow$ geometric median of $S$
-        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-        \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-        \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
-        \State \Call{Partition}{L, criteria}
-        \State \Call{Partition}{R, criteria}
-        \EndIf
-    \end{algorithmic}
-\end{algorithm}
+%         \If{$|C| > 1$ \textbf{and} $C$ satisfies $criteria$}
+%         \State $S \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
+%         \State $c \Leftarrow$ geometric median of $S$
+%         \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
+%         \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
+%         \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
+%         \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
+%         \State \Call{Partition}{L, criteria}
+%         \State \Call{Partition}{R, criteria}
+%         \EndIf
+%     \end{algorithmic}
+% \end{algorithm}
 
@@
 We use $Q$ for the data structure containing the clusters that are still in contention for being among the $k$ nearest neighbors, and $H$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
-These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+We also use $\delta$, $\delta^{+}$, and $\delta^{-}$ to denote the distances from the query to the cluster center, the theoretically farthest point in the cluster, and the theoretically closest point in the cluster, respectively.
+% These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
 
-\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
-\label{sec:methods:knn-search:repeated-rnn}
-
-We begin using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
-If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
-Now, so long as $\sum_{C \in Q} |C| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{C \in Q} |C|}} \big)^\frac{1}{\mu}$
-where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
-Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
-Once $\sum_{C \in Q} |C| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
-
-\begin{algorithm}[!t]
+\begin{algorithm}[h]
     \caption{\textsc{TreeSearch}($C$, $q$, $r$)}
@@
 \end{algorithm}
+\vspace{-0.2in}
 
-\begin{algorithm}[!t]
+\begin{algorithm}[h]
     \caption{\textsc{LeafSearch}($Q$, $q$, $r$)}
@@
 \end{algorithm}
+
+\subsubsection{Repeated \texorpdfstring{$\rho$}{p}-NN}
+\label{sec:methods:knn-search:repeated-rnn}
+
+We begin using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
+If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
+Now, so long as $\sum_{C \in Q} |C| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{C \in Q} |C|}} \big)^\frac{1}{\mu}$
+where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
+Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
+Once $\sum_{C \in Q} |C| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
 
@@
 We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $Q$.
-Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot \left( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} \right)^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of LFDs in $Q$.
+Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot ( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} )^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of LFDs in $Q$.
 It remains to estimate $\rho_k$.
@@
 Since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at the center of a cluster in $Q$.
-By Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
-This simplifies the \textsc{LeafSearch} component to $k \Bigg( 1 + 2 \cdot \bigg( \frac{\hat{|C|}}{k} \bigg) ^ {\hat{d}^{-1}} \Bigg)^d$.
+By Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|C|}}{k} )^{\hat{d}^{-1}}$.
+This simplifies the \textsc{LeafSearch} component to $k ( 1 + 2 \cdot ( \frac{\hat{|C|}}{k} ) ^ {\hat{d}^{-1}} )^d$.
 
@@
 
-We begin by letting $Q$ be a list of 3-tuples $(C, \delta^{+}, m)$, where $C$ is a cluster, $\delta^{+}$ is as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $C$ in $Q$.
+We begin by letting $Q$ be a list of 3-tuples $(C, \delta^{+}, m)$, where $m$ is the multiplicity of $C$ in $Q$.
 $Q$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $H$ starts with the center of $\mathcal{R}$.
@@
 \label{sec:methods:knn-search:depth-first-sieve}
-
-This algorithm (DFS) greedily descends into the child cluster whose center if closer to the query than that of its sibling.
 
@@
 The asymptotic complexity of BFS is dominated by our modified QuickSelect algorithm, which accounts for the clusters in $Q$ and the points in $H$, and is log-linear in their sum.
-Thus, the asymptotic complexity of BFS is
+Similarly, that of DFS is dominated by the priority queue operations for $T$ clusters and $L$ points (See Equation~\ref{eq:methods:sieve-complexity}).
 \begin{equation}
-    \mathcal{O} \big( (T + L) \log(T + L) \big).
-    \label{eq:methods:bfs-complexity}
+    \label{eq:methods:sieve-complexity}
+    \begin{split}
+    \mathcal{O} & \big( (T + L) \log(T + L) \big) \\
+    \mathcal{O} & \big( T \log T + L \log L \big)
+    \end{split}
 \end{equation}
-Similarly, the asymptotic complexity of DFS is dominated by the priority queue operations for $T$ clusters and $L$ points;
-it is given by
-\begin{equation}
-    \mathcal{O} \big( T \log T + L \log L \big).
-    \label{eq:methods:dfs-complexity}
-\end{equation}

### 2025-11-08 sections/4-results.tex
down to six pages

@@
 
-  \vskip 0.05in
+  \vskip 0.1in
   \includegraphics[width=0.75\columnwidth]{plots/legend.png}
 
+  \vskip -0.1in
   \caption{Throughput vs. dataset cardinality on Silva-18S and Radio-ML.}
@@
 
-\begin{table}[t]
+\begin{table}[h]
   \centering
@@
 
-\begin{table}
+\begin{table}[h]
   \centering
@@
   \end{adjustbox}
-  \vskip -0.1in
+  \vskip -0.075in
 \end{table}
 
-\begin{table}
-  \vskip -0.1in
+\begin{table}[h]
+  \vskip -0.075in
   \centering
@@
 
-\begin{table}
+\begin{table}[h]
   \vskip -0.1in
-  \caption{Random dataset: throughput and recall.
-  In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, we observe that CAKES's algorithms perform quite slowly.
-  CAKES exhibits perfect recall at all cardinalities, whereas HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets.}
+  \caption{Random dataset: throughput and recall.}
   \label{tab:results:qps-and-recall-random}

### 2025-11-08 sections/5-discussion.tex
down to six pages

@@
 
-This study can be extended in several directions.
-More specialized distance functions, such as Wasserstein~\cite{vallender1974calculation} for probability distributions and Tanimoto~\cite{bajusz2015tanimoto} for molecular structures, would require only a Rust implementation of the distance function itself to be compatible with CAKES.
-An examination of the geometric and topological properties of datasets that favor one CAKES algorithm over another would also be valuable.
-
-\subsection{Availability}
+This work could be extended with additional domain-specific distance functions, e.g.\ Wasserstein~\cite{vallender1974calculation} for probability distributions and Tanimoto~\cite{bajusz2015tanimoto} for molecular structures, requiring only a Rust implementation of the distance function itself to be compatible with CAKES.
+Examining the geometric and topological properties of datasets that favor one CAKES algorithm over another would also be valuable.
+% An examination of the geometric and topological properties of datasets that favor one CAKES algorithm over another would also be valuable.
 

### 2025-11-09 sections/1-introduction.tex
brought back Partition

@@
 Given a distance function for the data, $k$-nearest neighbors ($k$-NN) search finds the $k$ points in a dataset that are closest to a given query point.
-We use the term \textit{approximate} to refer to algorithms that have imperfect recall, while \textit{exact} refers to algorithms that have perfect recall as measured against a brute-force search.
-Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for fast \textit{and} exact search remains~\cite{ukey2023survey}.
+We use the term \textit{approximate} to refer to algorithms that have imperfect recall, and \textit{exact} to refer to algorithms that have perfect recall as measured against a brute-force search.
+Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for \textit{fast and exact} search remains~\cite{ukey2023survey}.
 
 This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
-We benchmark these against FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
+We benchmark CAKES against FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
 We also evaluate performance on SILVA 18S~\cite{10.1093/nar/gks1219} using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and RadioML~\cite{oshea2018radioml} using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
@@
 
-Entropy-scaling algorithms exploit the inherent manifold structure of large datasets, achieving complexity that scales with topological properties, such as metric entropy and fractal dimension, rather than cardinality.
+Entropy-scaling algorithms exploit the inherent manifold structure of large datasets, achieving performance that scales with topological properties, such as metric entropy and fractal dimension, rather than cardinality.
 This makes them especially suitable for manifold-constrained data.
 
-In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in Rust, extending CHESS~\cite{ishaq2019clustered} to $k$-NN search.
-Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods} provide a theoretical analysis.
+CAKES is a set of three algorithms for exact $k$-NN search: Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
+In Section~\ref{sec:methods}, we describe the clustering and search algorithms, and provide a theoretical analysis.
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a random uniform distribution.
-Given that CAKES's algorithms are intended for datasets with a manifold structure, an analysis assuming a uniform distribution would be uninformative.
-As a result, our analysis assumes datasets with manifold structure, reflecting the conditions for which CAKES is designed.
+Since CAKES is designed for datasets with a manifold structure, an analysis assuming a uniform distribution would be uninformative.
+Therefore, our analysis assumes datasets with manifold structure, reflecting the conditions for which CAKES is designed.

### 2025-11-09 sections/2-methods.tex
brought back Partition

--- 2025-11-05 sections/2-methods.tex
@@
 
-We present CAKES, a set of three novel algorithms for exact $k$-NN search in finite-dimensional spaces.
-Given a \textit{dataset} $\textbf{X} = \{x_1, \dots, x_n\}$ and a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$, $k$-NN search aims to find the $k$ points in $\textbf{X}$ closest to a given \textit{query} point $q$ according to $f$.
-If $f$ is a distance metric (i.e. it obeys the triangle inequality), then CAKES guarantees exact $k$-NN search.
-
-CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, the notion that high($D$)-dimensional data collected from constrained generating phenomena often occupy a low($d$)-dimensional manifold within their embedding space ($d \ll D$).
-We define the \emph{local fractal dimension} (LFD) to quantify this property.
-For a point $q \in \textbf{X}$ and radii $r_1 > r_2$,
-\begin{equation}
-    \text{LFD}(q, r_1, r_2) = \frac{\text{log} \Big( \frac{|B(q, r_1)|}{|B(q, r_2)|} \Big) }{\text{log} \big( \frac{r_1}{r_2} \big) }
-    \label{eq:methods:lfd-original}
-\end{equation}
-where $B(q, r)$ is the set of points within radius $r$ from $q$.
-
-When the majority of points in the dataset have low ($d \ll D$) LFD, we simply say that the dataset has low LFD.
-For real-world datasets, we expect the LFD to be locally uniform, i.e.,\,when $r$ is small, but (often) globally variable.
-
-\subsection{Clustering}
-\label{sec:methods:clustering}
-
-We define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
+Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$ and a distance function $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$, we define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
 The center $c$ is the geometric median of the points in the cluster (i.e., the point minimizing the sum of distances to all other points in $C$) and the radius $r$ is the maximum distance from $c$ to any point in $C$.
-A cluster with only one point is a $leaf$, otherwise a $parent$ cluster has two $child$ clusters.
-Starting from the root cluster $\mathcal{R}$ containing the entire dataset, Alg.~\ref{alg:methods:partition} recursively builds the tree $\mathcal{T}$ in $\mathcal{O}(n \log n)$ time.
-
-\begin{figure}[t]
-    \vskip -0.1in
-    \centering
-    \includegraphics[scale=0.5]{images/geometry/deltas.pdf}
-    \caption{
-        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
-        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
-        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
-        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
-    }
-    \label{fig:methods:deltas}
-    \vskip -0.05in
-\end{figure}
-
-\begin{algorithm} % enter the algorithm environment
+Using previous work in CHESS~\cite{ishaq2019clustered}, and CLAM~\cite{ishaq2021clustered}, we build a binary tree $\mathcal{T}$ of clusters by recursively partitioning the dataset $\textbf{X}$ (See Alg.~\ref{alg:methods:partition}).
+In practice, we approximate $c$ using a random sample of points from $C$.
+\begin{algorithm}[h] % enter the algorithm environment
     \caption{\textsc{Partition}($C$, $criteria$)} % give the algorithm a caption
@@
         \If{$|C| > 1$ \textbf{and} $C$ satisfies $criteria$}
-        \State $S \Leftarrow$ random sample of $\left\lceil \sqrt{|C|} \right\rceil$ points from $C$
-        \State $c \Leftarrow$ geometric median of $S$
+        \State $c \Leftarrow$ geometric median of $C$
         \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
@@
 
+The manifold hypothesis~\cite{fefferman2016testing} states that high($D$)-dimensional data collected from constrained generating phenomena occupy a low($d$)-dimensional manifold within their embedding space ($d \ll D$).
+We define the \emph{local fractal dimension} (LFD) to quantify this property:
+\begin{equation}
+    \text{LFD}(p, r_1, r_2) = \frac{\text{log} \big( \frac{|B(p, r_1)|}{|B(p, r_2)|} \big) }{\text{log} \big( \frac{r_1}{r_2} \big) }
+    \label{eq:methods:lfd-original}
+\end{equation}
+where $p \in X$ is a point and $B(p, r)$ is the set of points within radius $r$ from $p$.
+When $d \ll D$ for the majority of points, we simply say that the dataset has low LFD.
+
 \subsection{\texorpdfstring{$k$}{k}-Nearest Neighbors Search}
@@
 
-Given a query $q$ and integer $k$, $k$-NN search finds the $k$ closest points to $q$ in $\textbf{X}$. In this section, we present three novel algorithms for exact $k$-NN search:
-Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-We use $Q$ for the data structure containing the clusters that are still in contention for being among the $k$ nearest neighbors, and $H$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
-These algorithms also use some terminology defined and illustrated in Figure~\ref{fig:methods:deltas}.
+$k$-NN search in CAKES guarantees perfect recall when $f$ is a metric (i.e.\ it satisfies the triangle inequality) with performance that scales with LFD rather than $|X|$.
+
+In the following, we use $Q$ for the data structure containing clusters that are in contention for containing the $k$ nearest neighbors and $H$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
+We denote the distances from the query to the center, the theoretically farthest, and the theoretically closest points in a cluster $C$ using $\delta_C$, $\delta^{+}_C$, and $\delta^{-}_C$, respectively.
+We also assume that the query $q$ is drawn from the same distribution as $X$.
 
@@
 
-We begin using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
+We begin by using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
 If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
@@
 
-\begin{algorithm}[!t]
+\begin{algorithm}[h]
     \caption{\textsc{TreeSearch}($C$, $q$, $r$)}
@@
         \For{$C \in Q$}
-            \If{$\delta^+_C \leq r$}
-                \State $H$ $\Leftarrow$ $H \cup C$
-            \Else
             \For{$p \in C$ \textbf{if} $f(p, q) \leq r$}
@@
             \EndFor
-            \EndIf
         \EndFor
@@
 \end{algorithm}
+\vspace{-0.2in}
 
@@
 
-For \textsc{LeafSearch}, we must estimate $\sum_{C \in Q} |C|$;
-since we must examine every point in each such cluster, time complexity of \textsc{LeafSearch} is linear in this quantity.
+For \textsc{LeafSearch}, we must estimate $\sum_{C \in Q} |C|$.
+Since we must examine every point in each such cluster, the time complexity of \textsc{LeafSearch} is linear in this quantity.
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
@@
 While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
-Since $q$ is from the same distribution as $X$, the LFD at $q$ should not be significantly different from that at the center of a cluster in $Q$.
-By Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = \left( \frac{\hat{|C|}}{k} \right)^{\hat{d}^{-1}}$.
-This simplifies the \textsc{LeafSearch} component to $k \Bigg( 1 + 2 \cdot \bigg( \frac{\hat{|C|}}{k} \bigg) ^ {\hat{d}^{-1}} \Bigg)^d$.
+Assuming that $q$ is from the same distribution as $X$, by Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|C|}}{k} )^{\hat{d}^{-1}}$.
+This simplifies the \textsc{LeafSearch} component to $k ( 1 + 2 \cdot ( \frac{\hat{|C|}}{k} ) ^ {\hat{d}^{-1}} )^d$.
 
@@
     \mathcal{O}
-    \Bigg(
+    \bigg(
         \underbrace{
@@
             k \cdot
-            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
+            \overbrace{\Big( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \Big)^d}^{\textrm{scaling factor}}
         }_{\textrm{\textsc{LeafSearch}}}
-    \Bigg).
+    \bigg).
     \label{eq:methods:repeated-rnn-complexity}
@@
 
-We begin by letting $Q$ be a list of 3-tuples $(C, \delta^{+}, m)$, where $C$ is a cluster, $\delta^{+}$ is as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $C$ in $Q$.
-$Q$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $H$ starts with the center of $\mathcal{R}$.
-
-We then use the QuickSelect algorithm, modified to account for multiplicities and to reorder $Q$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $H$).
+We begin by letting $Q$ be a list of 3-tuples $(C, \delta^{+}, m)$, where $m$ is the multiplicity of $C$ in $Q$.
+$Q$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $H$ starts with the center of $\mathcal{R}$, where $\mathcal{R}$ is the root of $\mathcal{T}$.
+We then use QuickSelect, modified to account for multiplicities and to reorder $Q$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $H$).
 Since this step may require a binary search of $Q$ and reordering $Q$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
@@
 
-This algorithm (DFS) greedily descends into the child cluster whose center if closer to the query than that of its sibling.
-
 Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ starting with only $\mathcal{R}$ while $H$ starts empty.
@@
     \item While the top-priority cluster is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top-priority cluster (a leaf) from $Q$ and add all its points to $H$.
+    \item Remove the top-priority cluster (a leaf) from $Q$, add all its points to $H$, and trim $H$ to size $k$.
 \end{itemize}
-This process terminates when $H$ is full and the top-priority items in $H$ and $Q$ are such that $\delta_H < \delta_Q$, i.e.,\,the next closest point to be visited is farther from the query than the $k^{th}$ nearest neighbor found so far.
+This process terminates when $|H| = k$ and the top-priority items in $H$ and $Q$ are such that the next closest point to be visited in $Q$ is farther from the query than the $k^{th}$ nearest neighbor found so far.
 Thus, $H$ contains the $k$ nearest neighbors to the query.
@@
 The asymptotic complexity of BFS is dominated by our modified QuickSelect algorithm, which accounts for the clusters in $Q$ and the points in $H$, and is log-linear in their sum.
-Thus, the asymptotic complexity of BFS is
-\begin{equation}
-    \mathcal{O} \big( (T + L) \log(T + L) \big).
-    \label{eq:methods:bfs-complexity}
-\end{equation}
-Similarly, the asymptotic complexity of DFS is dominated by the priority queue operations for $T$ clusters and $L$ points;
-it is given by
-\begin{equation}
-    \mathcal{O} \big( T \log T + L \log L \big).
-    \label{eq:methods:dfs-complexity}
-\end{equation}
+Similarly, that of DFS is dominated by the priority queue operations for $T$ clusters and $L$ points.
+These complexities are:
+\begin{equation*}
+    % \label{eq:methods:sieve-complexity}
+    \begin{split}
+    BFS \sim \ & \mathcal{O} \big( (T + L) \log(T + L) \big) \\
+    DFS \sim \ & \mathcal{O} \big( T \log T + L \log L \big)
+    \end{split}
+\end{equation*}

### 2025-11-09 sections/3-datasets.tex
down to six pages
brought back Partition

@@
 
-\subsection{Datasets}
-\label{sec:datasets-and-benchmarks:ann-benchmark-datasets}
+% \subsection{Datasets}
+% \label{sec:datasets-and-benchmarks:ann-benchmark-datasets}
 
@@
 For both datasets, we hold out 1,000 random samples as queries and build the tree on the remaining samples.
-These were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM running Manjaro Linux 5.15.164-1-MANJARO.
+These benchmarks were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM running Manjaro Linux 5.15.164-1.
 
-\begin{table}
+\begin{table}[h]
+  \vskip -0.1in
   \caption{Datasets used in benchmarks.}
+  \vskip -0.25in
   \label{tab:datasets:summary}
@@
   \end{center}
-    \vskip -0.1in
+  \vskip -0.25in
 \end{table}
@@
 
-In order to study scaling behavior as cardinality increases, we synthetically augment the ANN-Benchmarks datasets by generating new points by applying small random perturbations to existing points in the original dataset.
+In order to study how search performance scales as cardinality increases, we synthetically augment the ANN-Benchmarks datasets by generating new points by applying small random perturbations to existing points in the original dataset.
 Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, \ref{tab:results:qps-and-recall-random} show the results as the cardinality increases, with ``Mult.'' indicating the multiplier by which the cardinality was augmented.
@@
 Since the time complexity of CAKES algorithms scales with the LFD of the dataset, we examine, in Figure~\ref{fig:results:lfd-plots}, the LFD of each dataset we used for benchmarks.
-
-\begin{figure}[t]
+\begin{figure}[h]
+    \vskip -0.2in
     \centering
@@
       \includegraphics[width=0.48\columnwidth]{images/lfd/glove-25.png}}
+    \vskip -0.2in
 
@@
       \includegraphics[width=0.48\columnwidth]{images/lfd/random.png}}
+    \vskip -0.2in
 
@@
 Since HNSW, ANNOY, and FAISS do not support the Levenshtein or DTW distance functions, we only benchmark CAKES on the SilVA-18S and RadioML datasets and report results in Figures~\ref{fig:results:silva-scaling} and~\ref{fig:results:radioml-scaling}.
-All benchmarks were run with $k=10$, throughput is averaged over all queries, and recall is computed using a na\"ive linear scan.
+All benchmarks were run with $k=10$, throughput is averaged over all queries, and recall is computed using a na\"ive linear scan as ground truth.
 

### 2025-11-09 sections/4-results.tex
brought back Partition

@@
 
-For each of the ANN-benchmark datasets and Random, Fig.~\ref{fig:results:indexing} reports the time taken for each algorithm to build its index and to tune any hyper-parameters to maximize recall.
+For each of the vector datasets, Fig.~\ref{fig:results:indexing} reports the time taken for each algorithm to build its index assuming that the hyper-parameters are set to maximize recall.
 
@@
   }
-  \subfloat[Glove-25]{=
+  \subfloat[Glove-25]{
     \includegraphics[width=0.24\textwidth]{images/indexing/glove-25-indexing.png}

### 2025-11-09 sections/1-introduction.tex
fixed some more notation

@@
 FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
-However, many of these algorithms do not support \emph{exact} search (as defined in Section \ref{sec:introduction}).
+However, many of these algorithms do not support \emph{exact} search.
 
@@
 CAKES is a set of three algorithms for exact $k$-NN search: Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
-In Section~\ref{sec:methods}, we describe the clustering and search algorithms, and provide a theoretical analysis.
+In the next section, we describe the clustering and search algorithms, and provide a complexity analysis.
 These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a random uniform distribution.

### 2025-11-09 sections/2-methods.tex
fixed some more notation

@@
 
-Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$ and a distance function $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$, we define a \emph{cluster} $C$ as a set of points with a \emph{center} and \emph{radius}.
-The center $c$ is the geometric median of the points in the cluster (i.e., the point minimizing the sum of distances to all other points in $C$) and the radius $r$ is the maximum distance from $c$ to any point in $C$.
+Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$ and a distance function $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$, we define a \emph{cluster} $\mathcal{C}$ as a set of points with a \emph{center} and \emph{radius}.
+The center $c$ is the geometric median of the points in the cluster (i.e., the point that minimizes the sum of distances to all other points in $\mathcal{C}$) and the radius $r$ is the maximum distance from $c$ to any point in $\mathcal{C}$.
 Using previous work in CHESS~\cite{ishaq2019clustered}, and CLAM~\cite{ishaq2021clustered}, we build a binary tree $\mathcal{T}$ of clusters by recursively partitioning the dataset $\textbf{X}$ (See Alg.~\ref{alg:methods:partition}).
-In practice, we approximate $c$ using a random sample of points from $C$.
+In practice, we approximate $c$ using a random sample of points from $\mathcal{C}$.
 \begin{algorithm}[h] % enter the algorithm environment
-    \caption{\textsc{Partition}($C$, $criteria$)} % give the algorithm a caption
+    \caption{\textsc{Partition}($\mathcal{C}$, $criteria$)} % give the algorithm a caption
     \label{alg:methods:partition} % and a label for \ref{} commands later in the document
@@
         \Require $f$, a distance function
-        \Require $C$, a cluster
+        \Require $\mathcal{C}$, a cluster
         \Require $criteria$, user-specified continuation criteria
 
-        \If{$|C| > 1$ \textbf{and} $C$ satisfies $criteria$}
-        \State $c \Leftarrow$ geometric median of $C$
+        \If{$|C| > 1$ \textbf{and} $\mathcal{C}$ satisfies $criteria$}
+        \State $c \Leftarrow$ geometric median of $\mathcal{C}$
         \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
@@
 
-$k$-NN search in CAKES guarantees perfect recall when $f$ is a metric (i.e.\ it satisfies the triangle inequality) with performance that scales with LFD rather than $|X|$.
+$k$-NN search in CAKES guarantees perfect recall when $f$ is a metric (i.e.\ it satisfies the triangle inequality) with performance that scales with LFD rather than $|\textbf{X}|$.
 
-In the following, we use $Q$ for the data structure containing clusters that are in contention for containing the $k$ nearest neighbors and $H$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
-We denote the distances from the query to the center, the theoretically farthest, and the theoretically closest points in a cluster $C$ using $\delta_C$, $\delta^{+}_C$, and $\delta^{-}_C$, respectively.
-We also assume that the query $q$ is drawn from the same distribution as $X$.
+In the following, we use $\mathcal{Q}$ for the data structure containing clusters that are in contention for containing the $k$ nearest neighbors and $\mathcal{H}$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
+We assume that the query $q$ is drawn from the same distribution as $X$, and we
+denote the distances from $q$ to the center, the theoretically farthest, and the theoretically closest points in a cluster $\mathcal{C}$ using $\delta_\mathcal{C}$, $\delta^{+}_\mathcal{C}$, and $\delta^{-}_\mathcal{C}$, respectively.
 
@@
 
-We begin by using \textsc{TreeSearch} on the root cluster $R$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|X|$.
-If the returned set $Q$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $Q$ is non-empty.
-Now, so long as $\sum_{C \in Q} |C| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{C \in Q} |C|}} \big)^\frac{1}{\mu}$
-where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
+We begin by using \textsc{TreeSearch} on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
+If the returned set $\mathcal{Q}$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $\mathcal{Q}$ is non-empty.
+Now, so long as $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
+In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}|}} \big)^\frac{1}{\mu}$
+where $\mu$ is the harmonic mean of the LFD of the clusters in $\mathcal{Q}$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
 Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
-Once $\sum_{C \in Q} |C| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
-
+Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
 \begin{algorithm}[h]
-    \caption{\textsc{TreeSearch}($C$, $q$, $r$)}
+    \caption{\textsc{TreeSearch}($\mathcal{C}$, $q$, $r$)}
     \label{alg:methods:rnn-search:tree-search}
     \begin{algorithmic}
-        \Require $C$, a cluster
+        \Require $\mathcal{C}$, a cluster
         \Require $q$, a query
@@
         \EndIf
-        \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
+        \State $[L, R]$ $\Leftarrow$ \textit{children} of $\mathcal{C}$
         \State \textbf{return} \Call{TreeSearch}{$L, q, r$} $\cup$ \Call{TreeSearch}{$R, q, r$}
@@
 \begin{algorithm}[h]
-    \caption{\textsc{LeafSearch}($Q$, $q$, $r$)}
+    \caption{\textsc{LeafSearch}($\mathcal{Q}$, $q$, $r$)}
     \label{alg:methods:rnn-search:leaf-search}
@@
         \Require $f$, a distance function
-        \Require $Q$, a set of clusters
+        \Require $\mathcal{Q}$, a set of clusters
         \Require $q$, a query
@@
             \For{$p \in C$ \textbf{if} $f(p, q) \leq r$}
-                \State $H$ $\Leftarrow$ $H \cup \{p\}$
+                \State $\mathcal{H}$ $\Leftarrow$ $H \cup \{p\}$
             \EndFor
         \EndFor
-        \State \textbf{return} $H$
+        \State \textbf{return} $\mathcal{H}$
     \end{algorithmic}
@@
         \underbrace{
-            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
+            \log~\overbrace{\mathcal{N}_{\hat{r}}(\textbf{X})}^{\textrm{metric entropy}}
         }_{\textrm{\textsc{TreeSearch}}}
@@
 
-To extend this to Repeated $\rho$-NN, we first need to estimate the number of iterations of \textsc{TreeSearch} needed to find a radius that guarantees at least one neighbor.
-Since we expect $q$ to be from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the nearest neighbor.
-Thus, the multiplicative factor described above suggests that in the expected case, we need only one iteration of \textsc{TreeSearch} to find the first cluster, and one more to find enough clusters to guarantee $k$ neighbors.
-Since this is a constant factor, Repeated $\rho$-NN has the same complexity for \textsc{TreeSearch} as $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(X)\big)$.
+To extend this to Repeated $\rho$-NN, we first need to estimate the number of repetitions of \textsc{TreeSearch} needed to find a radius that guarantees at least one neighbor.
+Since we expect $q$ to be from the same distribution as $\textbf{X}$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$.
+Thus, the multiplicative factor described above suggests that, in the expected case, we need only one iteration of \textsc{TreeSearch} to find the first cluster, and one more to find enough clusters to guarantee $k$ neighbors.
+Since this is a constant factor, Repeated $\rho$-NN has the same complexity for \textsc{TreeSearch} as $\rho$-NN search, i.e.,\,$\mathcal{O}\big(\log\mathcal{N}_{\hat{r}}(\textbf{X})\big)$.
 
-For \textsc{LeafSearch}, we must estimate $\sum_{C \in Q} |C|$.
+For \textsc{LeafSearch}, we must estimate $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}|$.
 Since we must examine every point in each such cluster, the time complexity of \textsc{LeafSearch} is linear in this quantity.
 Let $\rho_k$ be the distance from the query to the $k^{th}$ nearest neighbor.
-Then, we see that $Q$ is the set of clusters that overlap with a ball of radius $\rho_k$ around the query.
-We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $Q$.
-Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot ( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} )^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of LFDs in $Q$.
+Then, we see that $\mathcal{Q}$ is the set of clusters that overlap with a ball of radius $\rho_k$ around the query.
+We can estimate this region as a ball of radius $\rho_k + 2\hat{r}$, where $\hat{r}$ is the mean radius of the clusters in $\mathcal{Q}$.
+Thus, the \textsc{LeafSearch} component of Equation~\ref{eq:methods:rnn-search-complexity} becomes $k \cdot ( \frac{\rho_k + 2 \cdot \hat{r}}{\rho_k} )^{\hat{d}}$, where $\hat{d}$ is the harmonic mean of LFDs in $\mathcal{Q}$.
 It remains to estimate $\rho_k$.
 
-While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|C|}$, of clusters in $Q$ at a radius equal to the mean of their radii, $\hat{r}$.
-Assuming that $q$ is from the same distribution as $X$, by Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|C|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|C|}}{k} )^{\hat{d}^{-1}}$.
-This simplifies the \textsc{LeafSearch} component to $k ( 1 + 2 \cdot ( \frac{\hat{|C|}}{k} ) ^ {\hat{d}^{-1}} )^d$.
+While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at a radius equal to the mean of their radii, $\hat{r}$.
+Assuming that $q$ is from the same distribution as $\textbf{X}$, by Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|\mathcal{C}|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|\mathcal{C}|}}{k} )^{\hat{d}^{-1}}$.
+This simplifies the \textsc{LeafSearch} component to $k ( 1 + 2 \cdot ( \frac{\hat{|\mathcal{C}|}}{k} ) ^ {\hat{d}^{-1}} )^d$.
 
-Again using the assumption that $q$ is from the same distribution as $X$, we have that $\hat{d} \approx d$.
+Again using the assumption that $q$ is from the same distribution as $\textbf{X}$, we have that $\hat{d} \approx d$.
 Thus, the asymptotic complexity of Repeated $\rho$-NN is
@@
             k \cdot
-            \overbrace{\Big( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \Big)^d}^{\textrm{scaling factor}}
+            \overbrace{\Big( 1 + 2 \cdot \Big( \frac{\hat{|\mathcal{C}|}}{k} \Big) ^ {d^{-1}} \Big)^d}^{\textrm{scaling factor}}
         }_{\textrm{\textsc{LeafSearch}}}
@@
 
-We begin by letting $Q$ be a list of 3-tuples $(C, \delta^{+}, m)$, where $m$ is the multiplicity of $C$ in $Q$.
-$Q$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $H$ starts with the center of $\mathcal{R}$, where $\mathcal{R}$ is the root of $\mathcal{T}$.
-We then use QuickSelect, modified to account for multiplicities and to reorder $Q$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $H$).
-Since this step may require a binary search of $Q$ and reordering $Q$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|Q| \log |Q|\right)$ time complexity.
+We begin by letting $\mathcal{Q}$ be a list of 3-tuples $(\mathcal{C}, \delta^{+}_{\mathcal{C}}, m)$, where $m$ is the multiplicity of $\mathcal{C}$ in $\mathcal{Q}$.
+$\mathcal{Q}$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ starts with the center of $\mathcal{R}$, where $\mathcal{R}$ is the root of $\mathcal{T}$.
+We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $\mathcal{H}$).
+Since this step may require a binary search of $\mathcal{Q}$ and reordering $\mathcal{Q}$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|\mathcal{Q}| \log |\mathcal{Q}|\right)$ time complexity.
 
-Next, we check every cluster in $Q$, removing any for which $\delta^{-} > \tau$ because they cannot contain one of the $k$ nearest neighbors.
-For each of the remaining clusters, if it is a leaf, we remove it from $Q$ and add all its points to $H$.
-Otherwise (since it must be a parent cluster), we remove it from $Q$ and, for each of its child clusters, we compute $\delta^{+}$ and add the corresponding 3-tuple to $Q$ and the child's center to $H$, thus expanding the search to the next level in $T$.
-We continue until $Q$ is empty or the farthest point in $H$ is closer to $q$ than the $\delta^{-}$ of every cluster in $Q$.
-Thus, $H$ contains the $k$ nearest neighbors to the query.
+Next, we check every cluster in $\mathcal{Q}$, removing any for which $\delta^{-} > \tau$ because they cannot contain one of the $k$ nearest neighbors.
+For each of the remaining clusters, if it is a leaf, we remove it from $\mathcal{Q}$ and add all its points to $\mathcal{H}$.
+Otherwise (since it must be a parent cluster), we remove it from $\mathcal{Q}$ and, for each of its children, we compute $\delta^{+}$ and add the corresponding 3-tuple to $\mathcal{Q}$ and the child's center to $\mathcal{H}$, thus expanding the search to the next level in $\mathcal{T}$.
+We continue until $\mathcal{Q}$ is empty or the farthest point in $\mathcal{H}$ is closer to $q$ than the $\delta^{-}$ of every cluster in $\mathcal{Q}$.
+Thus, $\mathcal{H}$ contains the $k$ nearest neighbors to the query.
 
@@
 
-Let $Q$ be a min-queue of clusters prioritized by $\delta^{-}$ starting with only $\mathcal{R}$ while $H$ starts empty.
-So long as $H$ is not full or the top-priority items from $H$ and $Q$ have $\delta_H \geq \delta_Q$, we repeat following:
+Let $\mathcal{Q}$ be a min-queue of clusters prioritized by $\delta^{-}$ starting with only $\mathcal{R}$ while $\mathcal{H}$ starts empty.
+So long as $\mathcal{H}$ is not full or the top-priority items from $\mathcal{H}$ and $\mathcal{Q}$ have $\delta_{\mathcal{H}} \geq \delta_{\mathcal{Q}}$, we repeat following:
 \begin{itemize}
-    \item While the top-priority cluster is not a leaf, remove it from $Q$ and add its children to $Q$.
-    \item Remove the top-priority cluster (a leaf) from $Q$, add all its points to $H$, and trim $H$ to size $k$.
+    \item While the top-priority cluster is not a leaf, remove it from $\mathcal{Q}$ and add its children to $\mathcal{Q}$.
+    \item Remove the top-priority cluster (a leaf) from $\mathcal{Q}$, add all its points to $\mathcal{H}$, and trim $\mathcal{H}$ to size $k$.
 \end{itemize}
-This process terminates when $|H| = k$ and the top-priority items in $H$ and $Q$ are such that the next closest point to be visited in $Q$ is farther from the query than the $k^{th}$ nearest neighbor found so far.
-Thus, $H$ contains the $k$ nearest neighbors to the query.
+This process terminates when $|\mathcal{H}| = k$ and the top-priority items in $\mathcal{H}$ and $\mathcal{Q}$ are such that the next closest point to be visited in $\mathcal{Q}$ is farther from the query than the $k^{th}$ nearest neighbor found so far.
+Thus, $\mathcal{H}$ contains the $k$ nearest neighbors to the query.
 
@@
 Then the number of such clusters is bounded above by $2d$, where the bound is achieved by having a cluster overlap the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
-In the worst-case scenario for \textsc{TreeSearch}, these clusters would all come from different branches of the tree, and \textsc{TreeSearch} would look at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
-Thus, the asymptotic complexity of this component is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
-For \textsc{LeafSearch}, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the complexity is $L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|C|}}{k} \big) ^ {d^{-1}} \Big)^d \Big)$.
+In the worst-case scenario for \textsc{TreeSearch}, these clusters would all come from different branches of the tree, and \textsc{TreeSearch} would look at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(\textbf{X})$ clusters.
+Thus, the asymptotic complexity of this component is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(\textbf{X}) \big)$.
+For \textsc{LeafSearch}, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the complexity is $L \coloneqq \mathcal{O} \Big( k \cdot \Big( 1 + 2 \cdot \big( \frac{\hat{|\mathcal{C}|}}{k} \big) ^ {d^{-1}} \Big)^d \Big)$.
 
-The asymptotic complexity of BFS is dominated by our modified QuickSelect algorithm, which accounts for the clusters in $Q$ and the points in $H$, and is log-linear in their sum.
+The asymptotic complexity of BFS is dominated by our modified QuickSelect algorithm, which accounts for the clusters in $\mathcal{Q}$ and the points in $\mathcal{H}$, and is log-linear in their sum.
 Similarly, that of DFS is dominated by the priority queue operations for $T$ clusters and $L$ points.
 These complexities are:
-\begin{equation*}
-    % \label{eq:methods:sieve-complexity}
+\begin{equation}
+    \label{eq:methods:sieve-complexity}
     \begin{split}
@@
     \end{split}
-\end{equation*}
+\end{equation}

### 2025-11-09 sections/3-datasets.tex
fixed some more notation

@@
 Since HNSW, ANNOY, and FAISS do not support the Levenshtein or DTW distance functions, we only benchmark CAKES on the SilVA-18S and RadioML datasets and report results in Figures~\ref{fig:results:silva-scaling} and~\ref{fig:results:radioml-scaling}.
-All benchmarks were run with $k=10$, throughput is averaged over all queries, and recall is computed using a na\"ive linear scan as ground truth.
+All benchmarks were run with $k=10$, throughput was averaged over all queries, and recall was computed using a na\"ive linear scan as ground truth.
 
 HNSW, ANNOY, and FAISS have various hyper-parameters that can be tuned to trade off throughput, recall, and indexing time.
-We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the configuration with the highest throughput that still achieves perfect recall (or as close to it as possible) for each dataset and cardinality.
-To make the comparison fair, we also allow CAKES to select the fastest of its algorithms for each dataset and cardinality by testing each with a small number of the centers of clusters at low depth as representative queries.
+We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the configuration with the highest throughput that still achieves perfect recall (or as close to it as possible).
+To make the comparison fair, we also allow CAKES to select the fastest of its algorithms by testing each with a small number of the centers of clusters at low depth as representative queries.

### 2025-11-11 sections/2-methods.tex
made minor edits and fixed typos

@@
 
-        \If{$|C| > 1$ \textbf{and} $\mathcal{C}$ satisfies $criteria$}
+        \If{$|\mathcal{C}| > 1$ \textbf{and} $\mathcal{C}$ satisfies $criteria$}
         \State $c \Leftarrow$ geometric median of $\mathcal{C}$
-        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in C$
-        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in C$
-        \State $L \Leftarrow \{x \ | \ x \in C \land f(l, x) \le f(r, x)\}$
-        \State $R \Leftarrow \{x \ | \ x \in C \land f(r, x) < f(l, x)\}$
-        \State \Call{Partition}{L, criteria}
-        \State \Call{Partition}{R, criteria}
+        \State $l \Leftarrow \argmax f(c, x) \ \forall \ x \in \mathcal{C}$
+        \State $r \Leftarrow \argmax f(l, x) \ \forall \ x \in \mathcal{C}$
+        \State $\mathcal{L} \Leftarrow \{x \ | \ x \in \mathcal{C} \land f(l, x) \le f(r, x)\}$
+        \State $\mathcal{R} \Leftarrow \{x \ | \ x \in \mathcal{C} \land f(r, x) < f(l, x)\}$
+        \State \Call{Partition}{$\mathcal{L}$, criteria}
+        \State \Call{Partition}{$\mathcal{R}$, criteria}
         \EndIf
@@
 
-We begin by using \textsc{TreeSearch} on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
+We begin by using \textsc{TreeSearch} (Alg~\ref{alg:methods:rnn-search:tree-search}) on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
 If the returned set $\mathcal{Q}$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $\mathcal{Q}$ is non-empty.
@@
 Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
-Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} to find the $k$ nearest neighbors.
+Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} (Alg~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors.
 \begin{algorithm}[h]
@@
         \Require $r$, a search radius
-        \If{$\delta^+_C \leq r$}
-            \State \textbf{return} $\{C\}$
+        \If{$\delta^+_\mathcal{C} \leq r$}
+            \State \textbf{return} $\{\mathcal{C}\}$
         \EndIf
-        \State $[L, R]$ $\Leftarrow$ \textit{children} of $\mathcal{C}$
-        \State \textbf{return} \Call{TreeSearch}{$L, q, r$} $\cup$ \Call{TreeSearch}{$R, q, r$}
+        \State $[\mathcal{L}, \mathcal{R}]$ $\Leftarrow$ \textit{children} of $\mathcal{C}$
+        \State \textbf{return} \Call{TreeSearch}{$\mathcal{L}, q, r$} $\cup$ \Call{TreeSearch}{$\mathcal{R}, q, r$}
     \end{algorithmic}
@@
         \Require $r$, a search radius
-        \State $H \Leftarrow \emptyset$
-        \For{$C \in Q$}
-            \For{$p \in C$ \textbf{if} $f(p, q) \leq r$}
-                \State $\mathcal{H}$ $\Leftarrow$ $H \cup \{p\}$
+        \State $\mathcal{H} \Leftarrow \emptyset$
+        \For{$\mathcal{C} \in \mathcal{Q}$}
+            \For{$p \in \mathcal{C}$ \textbf{if} $f(p, q) \leq r$}
+                \State $\mathcal{H}$ $\Leftarrow$ $\mathcal{H} \cup \{p\}$
             \EndFor
@@
 We begin by letting $\mathcal{Q}$ be a list of 3-tuples $(\mathcal{C}, \delta^{+}_{\mathcal{C}}, m)$, where $m$ is the multiplicity of $\mathcal{C}$ in $\mathcal{Q}$.
-$\mathcal{Q}$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ starts with the center of $\mathcal{R}$, where $\mathcal{R}$ is the root of $\mathcal{T}$.
+$\mathcal{Q}$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ starts with the center of $\mathcal{R}$.
 We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $\mathcal{H}$).
-Since this step may require a binary search of $\mathcal{Q}$ and reordering $\mathcal{Q}$ takes linear time, this version of QuickSelect has $\mathcal{O}\left(|\mathcal{Q}| \log |\mathcal{Q}|\right)$ time complexity.
 

### 2025-11-11 sections/3-datasets.tex
made minor edits and fixed typos

@@
 We allow for a hyper-parameter search (in the ranges recommended by the documentation accompanying the respective libraries) to find the configuration with the highest throughput that still achieves perfect recall (or as close to it as possible).
-To make the comparison fair, we also allow CAKES to select the fastest of its algorithms by testing each with a small number of the centers of clusters at low depth as representative queries.
+To make the comparison fair, we also allow CAKES to select the fastest of its algorithms by testing each with a small number of the centers of clusters at low depth in the tree as representative queries.

### 2025-11-11 sections/4-results.tex
made minor edits and fixed typos

@@
 
-On Silva (Fig.~\ref{fig:results:silva-scaling}), all CAKES algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then level off at higher cardinalities.
+On Silva (Fig.~\ref{fig:results:silva-scaling}), all CAKES algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then levels off at higher cardinalities.
 DFS is consistently the fastest algorithm on this dataset.
@@
   \vskip 0.1in
-  \includegraphics[width=0.75\columnwidth]{plots/legend.png}
+  \includegraphics[width=0.75\columnwidth]{plots/legend_only_cakes.png}
 

### 2025-11-12 sections/1-introduction.tex
made minor edits and fixed typos
polishing

@@
 FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
-However, many of these algorithms do not support \emph{exact} search.
+Crucially, many of these algorithms do not support \emph{exact} search.
 
@@
 
-CAKES is a set of three algorithms for exact $k$-NN search: Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
+CAKES is a set of three such algorithms for exact $k$-NN search: Repeated $\rho$-NN, Breadth-First Sieve, and Depth-First Sieve.
 In the next section, we describe the clustering and search algorithms, and provide a complexity analysis.

### 2025-11-12 sections/2-methods.tex
polishing

@@
 
-Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$ and a distance function $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$, we define a \emph{cluster} $\mathcal{C}$ as a set of points with a \emph{center} and \emph{radius}.
+Given a dataset $\textbf{X} = \{x_1, \dots, x_n\}$ and a distance function $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^{\geq 0}$, we define a \emph{cluster} $\mathcal{C}$ as a set of points with a \emph{center} and \emph{radius}.
 The center $c$ is the geometric median of the points in the cluster (i.e., the point that minimizes the sum of distances to all other points in $\mathcal{C}$) and the radius $r$ is the maximum distance from $c$ to any point in $\mathcal{C}$.
@@
 \end{equation}
-where $p \in X$ is a point and $B(p, r)$ is the set of points within radius $r$ from $p$.
-When $d \ll D$ for the majority of points, we simply say that the dataset has low LFD.
+where $p$ is a dataset point and $B(p, r)$ is the set of points within radius $r$ of $p$.
+When $d \ll D$ for the majority of points in the dataset, we say the dataset has low LFD.
 
@@
 
-$k$-NN search in CAKES guarantees perfect recall when $f$ is a metric (i.e.\ it satisfies the triangle inequality) with performance that scales with LFD rather than $|\textbf{X}|$.
+CAKES guarantees perfect $k$-NN search recall when $f$ is a metric (in particular, it satisfies the triangle inequality) with performance that scales with LFD rather than cardinality $|\textbf{X}|$.
 
 In the following, we use $\mathcal{Q}$ for the data structure containing clusters that are in contention for containing the $k$ nearest neighbors and $\mathcal{H}$ for a max-priority queue (of size $k$) containing the current $k$ nearest neighbors found.
-We assume that the query $q$ is drawn from the same distribution as $X$, and we
-denote the distances from $q$ to the center, the theoretically farthest, and the theoretically closest points in a cluster $\mathcal{C}$ using $\delta_\mathcal{C}$, $\delta^{+}_\mathcal{C}$, and $\delta^{-}_\mathcal{C}$, respectively.
+We assume that the query $q$ is drawn from the same distribution as $\textbf{X}$. 
+For a cluster $\mathcal{C}$, we denote by $\delta_\mathcal{C}$, $\delta^{+}_\mathcal{C}$, and $\delta^{-}_\mathcal{C}$, the distances from $q$ to the cluster center, to the theoretically farthest point, and to the theoretically closest point, respectively.
 
@@
 Now, so long as $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| < k$, we repeat \textsc{TreeSearch}, but instead of doubling $r$, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
-In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}|}} \big)^\frac{1}{\mu}$
-where $\mu$ is the harmonic mean of the LFD of the clusters in $\mathcal{Q}$ to ensure that this factor is not dominated by outlier clusters with very high LFD.
-Intuitively, this factor will be large when we have found few points or the LFD is low because we need to increase the radius quickly, and small when we have found almost all $k$ points or LFD is large and we need to increase the radius slowly.
+In particular, we increase the radius by a factor of $\big( {\frac{k}{\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}|}} \big)^\frac{1}{\mu}$,
+where $\mu$ is the harmonic mean of the LFD of the clusters in $\mathcal{Q}$, to ensure that this factor is not dominated by outlier clusters with very high LFD.
+Intuitively, this factor will be large when we have found few points or the LFD is low, because in these cases, we can afford to increase the radius quickly.
+It will be small when we have found almost all $k$ points or LFD is large, and thus we want to increase the radius slowly.
 Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} (Alg~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors.
@@
 
-In CHESS~\cite{ishaq2019clustered}, we showed that the complexity of ranged search is
+In CHESS~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
 \begin{gather}
@@
 
-While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{around the query} of radius $\rho_k$ to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at a radius equal to the mean of their radii, $\hat{r}$.
+While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{centered at the query} of radius $\rho_k$ to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at a radius equal to the mean of their radii, $\hat{r}$.
 Assuming that $q$ is from the same distribution as $\textbf{X}$, by Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|\mathcal{C}|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|\mathcal{C}|}}{k} )^{\hat{d}^{-1}}$.
-This simplifies the \textsc{LeafSearch} component to $k ( 1 + 2 \cdot ( \frac{\hat{|\mathcal{C}|}}{k} ) ^ {\hat{d}^{-1}} )^d$.
+This simplifies the \textsc{LeafSearch} component to $k ( 1 + 2 \cdot ( \frac{\hat{|\mathcal{C}|}}{k} ) ^ {\hat{d}^{-1}} )^{\hat{d}}$.
 
@@
 We begin by letting $\mathcal{Q}$ be a list of 3-tuples $(\mathcal{C}, \delta^{+}_{\mathcal{C}}, m)$, where $m$ is the multiplicity of $\mathcal{C}$ in $\mathcal{Q}$.
-$\mathcal{Q}$ starts containing only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ starts with the center of $\mathcal{R}$.
-We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| \geq k$ (including the points in $\mathcal{H}$).
+$\mathcal{Q}$ initially contains only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ initially contains only the center of $\mathcal{R}$.
+We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| + |\mathcal{H}| \geq k$.
 
@@
     BFS \sim \ & \mathcal{O} \big( (T + L) \log(T + L) \big) \\
-    DFS \sim \ & \mathcal{O} \big( T \log T + L \log L \big)
+    DFS \sim \ & \mathcal{O} \big( T \log T + L \log L \big).
     \end{split}

### 2025-11-12 sections/3-datasets.tex
polishing

@@
 The Radio-ML dataset consists of 1,024-dimensional complex-valued time-series samples, spanning 24 modulation modes and 26 SNR levels from $-20$dB to $30$dB.
-We use the 10dB SNR subset with 97,304 samples and Dynamic Time Warping~\cite{muller2007dynamic} as the distance metric.
-For both datasets, we hold out 1,000 random samples as queries and build the tree on the remaining samples.
+We use the 10dB SNR subset with 97,304 samples and Dynamic Time Warping~\cite{muller2007dynamic} (DTW) as the distance metric.
+For both datasets, we withhold 1,000 random samples as queries and build the tree on the remaining samples.
 These benchmarks were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM running Manjaro Linux 5.15.164-1.
@@
     \caption{Local fractal dimension vs.
-cluster depth across six datasets; note the different y-axis.
+cluster depth across six datasets; note the different y-axes.
 The x-axis is tree depth; the y-axis is LFD at that depth.
 We plot the 5th, 25th, 50th, 75th, and 95th percentiles, and min/max.
-The Fashion-MNIST, Glove-25, Sift and Silva datasets are real-world datasets and exhibit low LFD indicating the presence of manifold structure.
+The Fashion-MNIST, Glove-25, Sift and Silva datasets are real-world datasets and exhibit low LFD, indicating a manifold structure.
 The Random dataset has high LFD, indicating lack of manifold structure.
@@
 
-Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, and \ref{tab:results:qps-and-recall-random} show the throughput (QPS) and recall for each algorithm on each of the vector datasets.
+Tables~\ref{tab:results:qps-and-recall-fmn}, \ref{tab:results:qps-and-recall-glove}, \ref{tab:results:qps-and-recall-sift}, and \ref{tab:results:qps-and-recall-random} show the throughput (queries per second) and recall for each algorithm on each of the vector datasets.
 Since HNSW, ANNOY, and FAISS do not support the Levenshtein or DTW distance functions, we only benchmark CAKES on the SilVA-18S and RadioML datasets and report results in Figures~\ref{fig:results:silva-scaling} and~\ref{fig:results:radioml-scaling}.

### 2025-11-14 sections/2-methods.tex
yet another pass

@@
         }_{\textrm{\textsc{LeafSearch}}}
-    \Bigg).
+    \Bigg)
     \label{eq:methods:rnn-search-complexity}
 \end{gather}
+where $\mathcal{N}_{\hat{r}}(\textbf{X})$ is the minimum number of clusters of radius $\hat{r}$ needed to cover $\textbf{X}$, and $\rho$ is the search radius.
 
@@
 
-While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball \textit{centered at the query} of radius $\rho_k$ to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at a radius equal to the mean of their radii, $\hat{r}$.
+While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball of radius $\rho_k$ \textit{centered at the query} to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at a radius equal to the mean of their radii, $\hat{r}$.
 Assuming that $q$ is from the same distribution as $\textbf{X}$, by Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|\mathcal{C}|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|\mathcal{C}|}}{k} )^{\hat{d}^{-1}}$.
@@
 We begin by letting $\mathcal{Q}$ be a list of 3-tuples $(\mathcal{C}, \delta^{+}_{\mathcal{C}}, m)$, where $m$ is the multiplicity of $\mathcal{C}$ in $\mathcal{Q}$.
-$\mathcal{Q}$ initially contains only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ while $\mathcal{H}$ initially contains only the center of $\mathcal{R}$.
+Initially, $\mathcal{Q}$ contains only $(\mathcal{R}, \delta^{+}_{\mathcal{R}}, |\mathcal{R}| - 1)$ and $\mathcal{H}$ contains only the center of $\mathcal{R}$.
 We then use QuickSelect, modified to account for multiplicities and to reorder $\mathcal{Q}$ in-place, to find $\tau$, the smallest $\delta^{+}$ in $\mathcal{Q}$ such that $\left| B(q, \tau) \right| + |\mathcal{H}| \geq k$.

### 2025-11-14 sections/3-datasets.tex
yet another pass

@@
 cluster depth across six datasets; note the different y-axes.
-The x-axis is tree depth; the y-axis is LFD at that depth.
+The x-axis is tree depth; the y-axis is LFD of clusters at that depth, computed using Equation~\ref{eq:methods:lfd-original} with $r_1 = r$ and $r_2 = \frac{r}{2}$ where $r$ is the radius of the cluster.
 We plot the 5th, 25th, 50th, 75th, and 95th percentiles, and min/max.
@@
 The RadioML dataset is also a randomly generated dataset, but with a much more elaborate generation process.
-It seems to have little manifold structure, if any, instead exhibiting behavior akin to the Random dataset across three separate radial scales, indicated by the three peaks in LFD at different depths in the tree.}
+It seems to have little manifold structure, if any, instead exhibiting behavior akin to the Random dataset across three separate radial scales, as indicated by the three peaks in LFD at different depths in the tree.}
     \label{fig:results:lfd-plots}

### 2025-11-14 sections/4-results.tex
yet another pass

@@
 \label{sec:results}
-
-\subsection{Scaling Behavior and Recall}
-\label{sec:results:scaling-behavior-and-recall}
 
@@
 
-On Silva (Fig.~\ref{fig:results:silva-scaling}), all CAKES algorithms initially exhibit throughput that decreases approximately linearly with cardinality, but then levels off at higher cardinalities.
+On Silva (Fig.~\ref{fig:results:silva-scaling}), the throughput of all CAKES algorithms initially decreases with cardinality, but then levels off at higher cardinalities.
 DFS is consistently the fastest algorithm on this dataset.
@@
 
-For each of the vector datasets, Fig.~\ref{fig:results:indexing} reports the time taken for each algorithm to build its index assuming that the hyper-parameters are set to maximize recall.
+For each of the vector datasets, Fig.~\ref{fig:results:indexing} reports the time taken for each algorithm to build its index.
 
@@
   \caption{Indexing and tuning time.
-  We report the time taken by various algorithms to build and tune its index on various datasets.
+  We report the time taken by each algorithm to build and tune its index on various datasets.
   For CAKES, this includes the time required to build the tree and find the fastest internal algorithm.
+  For HNSW, ANNOY, and FAISS-IVF, this excludes any hyper-parameter tuning time, including only the time taken to build an index assuming that the recall-maximizing hyper-parameters are already known.
   All axes are on logarithmic scales.

### 2025-11-14 sections/5-discussion.tex
brought back Partition
revived availability line
yet another pass

@@
 Examining the geometric and topological properties of datasets that favor one CAKES algorithm over another would also be valuable.
-% An examination of the geometric and topological properties of datasets that favor one CAKES algorithm over another would also be valuable.
+Development of approximate variants of CAKES algorithms would allow users to trade off recall for higher throughput.
+Finally, the cluster tree could be extended to support dynamic insertion and deletion of points, enabling CAKES to be used in streaming applications.
 
-CAKES is implemented in Rust and available under the MIT license at https://github.com/URI-ABD/clam.
+We provide a Rust implementation of CAKES under the MIT license at \url{https://github.com/URI-ABD/clam}.

### 2025-11-15 sections/1-introduction.tex
changes I mentioned in slack

@@
 
-Given a distance function for the data, $k$-nearest neighbors ($k$-NN) search finds the $k$ points in a dataset that are closest to a given query point.
-We use the term \textit{approximate} to refer to algorithms that have imperfect recall, and \textit{exact} to refer to algorithms that have perfect recall as measured against a brute-force search.
-Existing fast algorithms are often approximate~\cite{gao2023high}, which may suffice for some applications, but the need for \textit{fast and exact} search remains~\cite{ukey2023survey}.
+Given a distance function on a dataset, $k$-nearest neighbors ($k$-NN) search returns the $k$ closest points to a query.
+We call algorithms with imperfect recall \textit{approximate} and those matching brute-force recall \textit{exact}.
+Most fast methods are approximate~\cite{gao2023high}; while sufficient for some applications, there remains demand for \textit{fast exact} search~\cite{ukey2023survey}.
 
@@
 Modern k-NN search strategies include entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}, Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, and random projection and tree building (ANNOY)~\cite{annoy}.
-HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
-FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
+% HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
+% FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
 Crucially, many of these algorithms do not support \emph{exact} search.

### 2025-11-15 sections/2-methods.tex
changes I mentioned in slack

@@
 where $p$ is a dataset point and $B(p, r)$ is the set of points within radius $r$ of $p$.
-When $d \ll D$ for the majority of points in the dataset, we say the dataset has low LFD.
+When we refer to the LFD of a cluster $\mathcal{C}$, we compute it using the cluster center as $p$, the cluster radius as $r_1$, and half the cluster radius as $r_2$.
+If the median cluster LFD at most depths is $\leq 3$, we say the dataset has ``low LFD''.
 
@@
 We assume that the query $q$ is drawn from the same distribution as $\textbf{X}$. 
-For a cluster $\mathcal{C}$, we denote by $\delta_\mathcal{C}$, $\delta^{+}_\mathcal{C}$, and $\delta^{-}_\mathcal{C}$, the distances from $q$ to the cluster center, to the theoretically farthest point, and to the theoretically closest point, respectively.
+For a cluster $\mathcal{C}$ with radius $r$, we denote by $\delta_\mathcal{C}$ the distance from $q$ to the cluster center.
+We define $\delta^{+}_\mathcal{C} \coloneqq \delta_\mathcal{C} + r$ and $\delta^{-}_\mathcal{C} \coloneqq \min(0, \delta_\mathcal{C} - r)$ respectively.
 
@@
 \label{sec:methods:knn-search:repeated-rnn}
-
+This algorithm is based on repeating $\rho$-NN search, which finds all points within a radius $\rho$ around a query point under the chosen distance function.
 We begin by using \textsc{TreeSearch} (Alg~\ref{alg:methods:rnn-search:tree-search}) on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
@@
 where $\mu$ is the harmonic mean of the LFD of the clusters in $\mathcal{Q}$, to ensure that this factor is not dominated by outlier clusters with very high LFD.
-Intuitively, this factor will be large when we have found few points or the LFD is low, because in these cases, we can afford to increase the radius quickly.
+% Intuitively, this factor will be large when we have found few points or the LFD is low, because in these cases, we can afford to increase the radius quickly.
 It will be small when we have found almost all $k$ points or LFD is large, and thus we want to increase the radius slowly.
@@
 \end{gather}
-where $\mathcal{N}_{\hat{r}}(\textbf{X})$ is the minimum number of clusters of radius $\hat{r}$ needed to cover $\textbf{X}$, and $\rho$ is the search radius.
+where $q$ is the query, $\hat{r}$ is the mean cluster radius, $\mathcal{N}_{\hat{r}}(\textbf{X})$ is the minimum number of clusters of radius $\hat{r}$ needed to cover $\textbf{X}$, and $\rho$ is the search radius.
 
@@
 
-While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball of radius $\rho_k$ \textit{centered at the query} to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at a radius equal to the mean of their radii, $\hat{r}$.
+While ordinarily we would compute LFD by comparing cardinalities of two balls with two different radii centered at \textit{the same} point, in order to estimate $\rho_k$, we instead compare the cardinality of a ball of radius $\rho_k$ \textit{centered at the query} to the mean cardinality, $\hat{|\mathcal{C}|}$, of clusters in $\mathcal{Q}$ at radius $\hat{r}$.
 Assuming that $q$ is from the same distribution as $\textbf{X}$, by Equation~\ref{eq:methods:lfd-original}, we have $\hat{d} = \frac{\log{}\frac{\hat{|\mathcal{C}|}}{k}}{\log{}\frac{\hat{r}}{\rho_k}}$, and rearranging, $\frac{\hat{r}}{\rho_k} = ( \frac{\hat{|\mathcal{C}|}}{k} )^{\hat{d}^{-1}}$.

### 2025-11-15 sections/3-datasets.tex
changes I mentioned in slack

--- 2025-11-12 sections/3-datasets.tex
@@
 cluster depth across six datasets; note the different y-axes.
-The x-axis is tree depth; the y-axis is LFD at that depth.
+The x-axis is tree depth; the y-axis is the LFD of clusters at that depth.
 We plot the 5th, 25th, 50th, 75th, and 95th percentiles, and min/max.
@@
 The RadioML dataset is also a randomly generated dataset, but with a much more elaborate generation process.
-It seems to have little manifold structure, if any, instead exhibiting behavior akin to the Random dataset across three separate radial scales, indicated by the three peaks in LFD at different depths in the tree.}
+It exhibits behavior akin to the Random dataset across three separate radial scales, as indicated by the three peaks in LFD at different depths in the tree.}
     \label{fig:results:lfd-plots}

### 2025-11-15 sections/5-discussion.tex
changes I mentioned in slack

@@
 In Figure~\ref{fig:results:lfd-plots}, we show the extent to which each dataset exhibits a manifold structure, as quantified by LFD.
-As expected, CAKES scales sub-linearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with high LFD;
+As expected, CAKES scales sub-linearly with cardinality on real-world datasets with low LFD and linearly on synthetic datasets with higher LFD;
 Sift versus Random illustrates this contrast, as does Radio-ML.

### 2025-11-15 sections/1-introduction.tex
brought back some stuff

@@
 
-Given a distance function on a dataset, $k$-nearest neighbors ($k$-NN) search returns the $k$ closest points to a query.
-We call algorithms with imperfect recall \textit{approximate} and those matching brute-force recall \textit{exact}.
-Most fast methods are approximate~\cite{gao2023high}; while sufficient for some applications, there remains demand for \textit{fast exact} search~\cite{ukey2023survey}.
+Given a distance function on a dataset, $k$-nearest neighbors ($k$-NN) search finds the $k$ closest points to a query.
+We call algorithms with imperfect recall \textit{approximate} and those matching brute-force \textit{exact}.
+Most fast methods are approximate~\cite{gao2023high}, and while they are sufficient for some applications, there remains a need for \textit{fast exact} search~\cite{ukey2023survey}.
 
@@
 
-Modern k-NN search strategies include entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}, Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw}, InVerted File indexing (FAISS-IVF)~\cite{faissivf}, and random projection and tree building (ANNOY)~\cite{annoy}.
-% HNSW~\cite{malkov2016hnsw} relies on navigable small world (NSW) networks and skip lists.
-% FAISS-IVF~\cite{faissivf, sacks1987multikey, kent1990signature} partitions data into high-dimensional Voronoi cells, and searches only within cells near the query.
+Modern k-NN search strategies include entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered} which leverages manifold structure, Hierarchical Navigable Small World networks (HNSW)~\cite{malkov2016hnsw} which relies on NSW networks and skip lists, InVerted File indexing (FAISS-IVF)~\cite{faissivf} which creates high-dimensional Voronoi cells, and ANNOY~\cite{annoy}, which builds several trees under different random projections.
 Crucially, many of these algorithms do not support \emph{exact} search.
@@
 In the next section, we describe the clustering and search algorithms, and provide a complexity analysis.
-These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a random uniform distribution.
-Since CAKES is designed for datasets with a manifold structure, an analysis assuming a uniform distribution would be uninformative.
-Therefore, our analysis assumes datasets with manifold structure, reflecting the conditions for which CAKES is designed.
+Our analysis assumes datasets with manifold structure, reflecting the conditions for which CAKES is designed.
+We expect, and show in Section~\ref{sec:results}, that CAKES performs well on such datasets, and poorly when manifold structure is absent.

### 2025-11-15 sections/2-methods.tex
brought back some stuff

--- 2025-11-14 sections/2-methods.tex
@@
 where $p$ is a dataset point and $B(p, r)$ is the set of points within radius $r$ of $p$.
-When $d \ll D$ for the majority of points in the dataset, we say the dataset has low LFD.
+For a cluster $\mathcal{C}$ with center $c$ and radius $r$, we define its LFD as $\text{LFD}(c, r, r/2)$.
+If the median LFD of clusters at most (and particularly high) depths in the tree is $\leq 3$, we say the dataset has ``low LFD''.
 
@@
 We assume that the query $q$ is drawn from the same distribution as $\textbf{X}$. 
-For a cluster $\mathcal{C}$, we denote by $\delta_\mathcal{C}$, $\delta^{+}_\mathcal{C}$, and $\delta^{-}_\mathcal{C}$, the distances from $q$ to the cluster center, to the theoretically farthest point, and to the theoretically closest point, respectively.
+For a cluster $\mathcal{C}$ with center $c$ and radius $r$, we define $\delta_\mathcal{C} = f(q, c)$, $\delta^{+}_\mathcal{C} = \delta_\mathcal{C} + r$, and $\delta^{-}_\mathcal{C} = \min(0, \delta_\mathcal{C} - r)$.
 
@@
 \label{sec:methods:knn-search:repeated-rnn}
-
-We begin by using \textsc{TreeSearch} (Alg~\ref{alg:methods:rnn-search:tree-search}) on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
+This algorithm repeats parts of $\rho$-NN search~\cite{ishaq2019clustered}, which finds all points within a radius $\rho$ around a query.
+Repeated $\rho$-NN begins by using \textsc{TreeSearch} (Alg~\ref{alg:methods:rnn-search:tree-search}) on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
 If the returned set $\mathcal{Q}$ is empty, we double $r$ and repeat \textsc{TreeSearch} until $\mathcal{Q}$ is non-empty.
@@
 Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} (Alg~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors.
+\vspace{-0.1in}
 \begin{algorithm}[h]
@@
 \end{gather}
-where $\mathcal{N}_{\hat{r}}(\textbf{X})$ is the minimum number of clusters of radius $\hat{r}$ needed to cover $\textbf{X}$, and $\rho$ is the search radius.
+where $\hat{r}$ is the mean cluster radius, $\mathcal{N}_{\hat{r}}(\textbf{X})$ is the minimum number of clusters of radius $\hat{r}$ needed to cover $\textbf{X}$.
 

### 2025-11-15 sections/3-datasets.tex
brought back some stuff

@@
 
-    \caption{Local fractal dimension vs.
-cluster depth across six datasets; note the different y-axes.
-The x-axis is tree depth; the y-axis is the LFD of clusters at that depth.
+    \caption{Local fractal dimension vs. cluster depth across six datasets;
+    note the different y-axes.
+    The x-axis is tree depth;
+    the y-axis is the LFD of clusters at that depth.
     We plot the 5th, 25th, 50th, 75th, and 95th percentiles, and min/max.
@@
     The Random dataset has high LFD, indicating lack of manifold structure.
+    The measured LFD of ``only'' 20 is a consequence of Equation~\ref{eq:methods:lfd-original} and only having sampled $10^6$ points in a 128-dimensional space $(\log_2 10^6 \approx 20)$;
+    We would need to sample many more points ($>2^{128}$), for the LFD to approach the embedding dimension of 128.
     The RadioML dataset is also a randomly generated dataset, but with a much more elaborate generation process.
-It exhibits behavior akin to the Random dataset across three separate radial scales, as indicated by the three peaks in LFD at different depths in the tree.}
+    It exhibits behavior akin to the Random dataset across three separate radial scales, as indicated by the distinct peaks in LFD at different depths.}
     \label{fig:results:lfd-plots}

### 2025-11-15 sections/2-methods.tex
couple more typos

@@
 \label{sec:methods:knn-search:repeated-rnn}
-This algorithm repeats parts of $\rho$-NN search~\cite{ishaq2019clustered}, which finds all points within a radius $\rho$ around a query.
+This algorithm builds on $\rho$-NN search from CHESS~\cite{ishaq2019clustered}, which finds all points within a radius $\rho$ around a query.
 Repeated $\rho$-NN begins by using \textsc{TreeSearch} (Alg~\ref{alg:methods:rnn-search:tree-search}) on the root cluster $\mathcal{R}$ with the search radius $r$ equal to the radius of $\mathcal{R}$ divided by $|\textbf{X}|$.
@@
 Once $\sum_{\mathcal{C} \in \mathcal{Q}} |\mathcal{C}| \geq k$, we perform \textsc{LeafSearch} (Alg~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors.
-\vspace{-0.1in}
+\vspace{-0.075in}
 \begin{algorithm}[h]
@@
 \end{algorithm}
-\vspace{-0.2in}
+\vspace{-0.225in}
 

### 2025-11-15 sections/3-datasets.tex
couple more typos

@@
     The Random dataset has high LFD, indicating lack of manifold structure.
-    The measured LFD of ``only'' 20 is a consequence of Equation~\ref{eq:methods:lfd-original} and only having sampled $10^6$ points in a 128-dimensional space $(\log_2 10^6 \approx 20)$;
+    Its measured LFD of ``only'' 20 is a consequence of Equation~\ref{eq:methods:lfd-original} and having sampled only $10^6$ points in a 128-dimensional space $(\log_2 10^6 \approx 20)$.
     We would need to sample many more points ($>2^{128}$), for the LFD to approach the embedding dimension of 128.
